### 2021-05-06 FMCAD2021/paper.tex
paper draft

\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{times} 
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{
->, % makes the edges directed
>=stealth, % makes the arrow heads bold
node distance=2cm, % specifies the minimum distance between two nodes. Change if necessary.
every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
initial text=$ $, % sets the text that appears on the start arrow
}

\RequirePackage{todonotes}
%\newcommand{\znj}[1]{\todo[inline,color=orange!10]{\textbf{ZHAN:} #1}}
\newcommand{\znjSide}[1]{\todo[color=orange!10]{\textbf{ZHAN:} #1}}
\newcommand\znj[1]{\textcolor{red}{#1}}

\newcommand{\parseInt}{\textsf{parseInt}}
\newcommand{\paexp}{$\textsf{PA}_{\exp}$}
%\newcommand{\paexp}{$(\mathbb{N},+,2^x)$}
\newcommand{\Def}{\hat{=}}

\title{A Decision Procedure for String Constraints with {\parseInt} Function and Flat Regular Constraints}
\author{ }
\institute{ } 
\begin{document}
\maketitle{}

\begin{abstract}
    The problem of solving string constraints is currently one of the focuses of theoretical computer science, particular related to SMT-based program verification, as it is inevitably that most of 
    programs manipulate string data structures. 
    The general problem of string constraint solving 
    is undecidable, therefore people try to identify 
    some sub-problems with decidability, e.g., string constraints restricted to generic flat languages. 
    Certainly, it is challenging and 
    makes theoretical and practical significance to 
    identify such subset with as much expressive power as possible. 
    In this paper, we identify 
    the subset of flat string constraints with 
     the string-number conversion function $\mathbf{parseInt}$, and 
     provide a decision procedure to the subset by reduction to
   the decidability problem of Presburger Arithmetic with exponential functions, 
    which is decidable. To the best of our knowledge, such a subset is 
    the largest subset of string constraints with decidability so far.

\keywords{ String Constraints \and Decidability \and Exponential Functions \and Quantifier Elimination.}

\end{abstract}

\iffalse     
    A Decision Procedure for String Constraints with the {\parseInt} Function and Flat Regular Constraints

    {\parseInt} converts strings to integers and is an important function in string manipulating programs. 
    Solving string constraints with {\parseInt} is challenging and undecidable in general. 
    The state-of-the-art string solvers resort to heuristics to solve such constraints. 
    In this work, we propose a decision procedure for a class of string constraints which 
    are boolean combinations of word equations, flat regular constraints, and {\parseInt} function. 
    The decision procedure is obtained by reducing the satisfiability problem to that of
    an extension of Presburger arithmetic with exponential functions, which is decidable. 
    We implement the decision procedure and carry out experiments to evaluate the performance. 
    The experimental results show the improvement against the state-of-the-art string solvers 
    in both precision and performance. To the best of our knowledge, this work represents the 
    first decision procedure for a class of string constraints with {\parseInt} function.
\fi

\section{Introduction}

solve string constraint is hard

strategy: do unsat and sat separately

strategy: using different procedure to (dis)prove validaity

for disprove validatity, there are two approches, first is bound string length

cannot handle $x.y \neq z  \wedge |x| > 2000$

more recent approach is flattening

it is known that word equaltion + flat regular constraints + len constraints
is decidable

it was unknown that whether word equations + flat regular constraints + len constraints + parseInt is decidable

\section{Preliminaries} \label{sec:pre}

In this section, we introduce some basic concepts and theories that will be used later. 
\subsection{Basic Concepts}

\paragraph{Sets and Strings}
We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the set of natural numbers and integers, respectively. 
$\mathbb{N}^+$ stands for the set of non-zero natural numbers. 
Let $\Sigma$ be a finite alphabet,
a string $w$ over $\Sigma$ is a sequence $a_1....a_n$ of characters from $\Sigma$.
Empty string is denoted by $\epsilon$.
$\Sigma^*$ denotes the set of all finite strings over $\Sigma$,
and $\Sigma_{\epsilon}$ stands for $\Sigma \cup \{\epsilon\}$.
For any string $w_1, w_2\in \Sigma^*$, 
we use $|w_1|$ to denote the length of $w_1$,
and $w_1\cdot w_2$ to denote the concatenation of $w_1$ and $w_2$.
A language $L$ over $\Sigma$ is a subset of $\Sigma^*$.

There are two types of variables in string constraints,
i.e., $X$, a set of string variables ranged over $\Sigma^*$,  
and $Z$,  a set of integer variables ranged over $\mathbb{Z}$.
As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
essentially a pair of two mappings 
$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.

\paragraph{Finite State Automata}
A Finite State Automaton is a tuple 
$\mathcal{A}=\langle Q,\Sigma,\Delta,$ $q_{\textit{init}},$ $q_{\textit{acc}}\rangle$, 
where $Q$ is a finite set of states, 
$\Sigma$ is the given alphabet,
$\Delta\subseteq Q\times \Sigma_\epsilon\times Q$ 
defines the transition relations in $\mathcal{A}$.
$q_{\textit{init}},q_{\textit{acc}}\in Q$ is the initial state and accepting state. 
A sequence $q_0 \langle a_1 \rangle  q_1 ... \langle a_n\rangle q_n$ is called accepting if $q_0 = q_{\textit{init}}$, $q_n = q_{\textit{acc}}$ and $q_{i-1}\langle a_i \rangle q_i \in \Delta$ for $1\le i\le n$.

\paragraph{Presburger Arithmetic} \label{PA}
The Presburger Arithmetic (PA) is a first order theory over signature 
$\Sigma_\mathbb{N}\Def  \{0,1,+,=\}$, where $0,1$ are constants, 
$+$ is a binary function and $=$ is a binary predicate.

PA can be axiomatized by the following axioms \cite{PA}

\begin{itemize}
    \item $\forall x, \neg (x+1=0)$
    \item $\forall x \forall y. x+1=y+1 \to x=y$
    \item $F(0) \wedge (\forall x. F(x)\to F(x+1)) \to \forall x. F(x)$ 
    \item $\forall x. x+0=x$
    \item $\forall x \forall y. x+(y+1)=(x+y)+1$
\end{itemize}

Given the domain $\mathbb{N}$,
the standard interpretation of PA interprets 
$0,1$ to $0_\mathbb{N},1_\mathbb{N}\in \mathbb{N}$
and $+,=$ to addition and equality over $\mathbb{N}$.
We call a PA formula without quantifiers a quantifier-free PA formula.

PA is a decidable theory, 
and the complexity of decidability is related to 
the number and locations of quantifiers.
Generally, 
the upper bound (on deterministic time and space) 
for deciding a formula of length $n$ is $2^{2^{2^{p n log(n)}}}$,
where $p>1$ is a constant\cite{Oppen69}.

\paragraph{Parikh Image}
Given an alphabet $\Sigma$ and a string $w\in \Sigma^*$, 
we define the set of Parikh variables 
$\Sigma^\# \Def \{a^\# \mid a\in \Sigma\}$.
The Parikh image of $w$ is a function 
$\mathbb{P}(w): \Sigma^\# \mapsto \mathbb{N}$,
which maps each symbol $a^\#\in \Sigma^\#$ to the number of occurrences
of $a$ in $w$.
For example, let $w\Def aabba$,
then $\mathbb{P}(w)(a^\#)=3,\mathbb{P}(w)(b^\#)=2$.

For a language $L\subseteq \Sigma^*$, 
define the Parikh image of $L$ to be 
$\mathbb{P}(L)\Def \{\mathbb{P}(w) | w\in L\}$.
We say a language $L$ is \emph{Parikh-definable} 
if $\mathbb{P}(L)$ can be characterized by a quantifier-free PA formula over $\Sigma^\#$, 
where $a^\#$ in the formula encodes the number of occurrences of $a$.
It is well known that
any context-free language (therefore regular language) 
is Parikh definable \cite{Parikh66}.

\paragraph{String Terms}
Given a finite alphabet $\Sigma$ and 
a finite set $X$ of string variables over $\Sigma^*$,
we define the set of terms $\textit{Terms}(\Sigma,X)$ 
to be the smallest set satisfying
\begin{itemize}
    \item[1] $\Sigma\cup \{\epsilon\} \cup X \subseteq \textit{Terms}(\Sigma,X)$;
    \item[2] if $t_1,t_2\in \textit{Terms}(\Sigma,X)$, then $t_1 \cdot t_2 \in \textit{Terms}(\Sigma,X)$.
\end{itemize}

We extend $I_X$ to $\textit{Terms}(\Sigma,X)$ by letting $I_X(\epsilon)=\epsilon$, 
for $a\in \Sigma, I_X(a)=a$,
and $I_X(t_1\cdot t_2)= I_X(t_1)\cdot I_X(t_2)$.

\paragraph{String Constraints} \label{par: string constraints}
Given a constraint $\phi$ and an interpretation $I$,
$I\models \phi$ denotes that $I$ satisfies $\phi$,
and $I$ is called a \emph{model} of $\phi$.
We use $\lVert \phi\rVert$ to denote the set of all models of $\phi$.

We define the following three forms of atomic string constraints:
\begin{itemize}
\item An equality constraint $\phi_e$ is of the form 
$t_1 = t_2$, where $t_1,t_2\in \textit{Terms}(\Sigma,X)$.
We define $\lVert \phi_e \rVert = \{I\mid I(t_1)=I(t_2)\}$.
Inequality constraints can be defined analogously.

\item A regular constraint $\phi_r$ is of the form 
$x\in L(\mathcal{A})$,
where $x\in \mathbb{X}$ and $\mathcal{A}$ is a finite state automaton.
We define $\lVert \phi_r \rVert = \{I\mid I(x)\in L(\mathcal{A})\}$.

\item A length constraint $\phi_l$ is a linear constraint over 
$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
where $|\cdot |$ is the length function.
We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
\end{itemize}

In Section~\ref{sec:string-solving},  
we will introduce a new form of atomic string constraints, i.e., \emph{string-number conversion constraints}.
A string constraint is a Boolean combination of atomic string constraints possibly with quantifications over $X\cup Z$. Other notions are same as in the first-order logic.

Giving a string constraint $\Psi$,
the problem of string constraints solving is to decide 
whether $\lVert \Psi \rVert$ is empty,
if not, to compute an interpretation $I$ that satisfies $\Psi$.

\subsection{Flat Automata and Flat Languages}

Given a string constraint $\Psi$,
the general problem of deciding whether $\lVert \Psi \rVert$ is empty is undecidable.
However, 
the problem becomes decidable when certain restrictions are imposed.
One of the restriction is by flat automata and flat languages, 
defined below.

\paragraph{Flat Languages and Automata}
For a fixed alphabet $\Sigma$,
we say a language $L$ over $\Sigma$ to be \emph{$\langle p,q \rangle$-flat} if 
there exist strings $w_1,...,w_q \in \Sigma^*$ such that
$|w_i|\le p$ for all $i:1\le i \le q$ 
and $L = (w_1)^*(w_2)^*...(w_q)^*$. 
We use $\alpha$ to denote $\langle p,q \rangle$, 
and call it the \emph{abstraction parameter} of $L$.
Intuitively,
a flat language with abstraction parameter $\alpha = \langle p,q \rangle$
consists of $q$ loops and the length of each loop body is equal or less than $p$.
For example,
$L = (ab)^*(a)^*(bb)^*$ is a $\langle 2,3 \rangle$-flat language.

Flat automata are a special form of finite state automata that 
recognize flat languages.
Fix the abstraction parameter $\alpha=\langle p,q\rangle$,
a flat automaton consists of $q$ loops,
each loop is a circle of $p$ states.
Formally, 
an $\alpha$-flat automaton contains $p q$ states at most,
and we name the states from $1$ to $p q$,
$1$ is the initial state and $(p q - p + 1)$ is the accepting state.
We use $\cdot$ as a placeholder for some symbol in $\Sigma_\epsilon$,
the transition relations of state $i$ are defined as 
\begin{itemize}
    \item if $i\  \text{mod}\  p = 1$ and $i \neq pq-p+1$, then 
    $(i,\epsilon,i+p)\in \Delta$,
    $(i, \cdot ,i+1) \in \Delta$;
    \item if $i\  \text{mod}\  p = 0$, then 
    $(i,\cdot , i-p+1) \in \Delta$;
    \item otherwise, $(i,\cdot, i+1) \in \Delta$.
\end{itemize}

A $\langle 2,3 \rangle$-flat automaton 
that recognizes $L\Def  (ab)^*(a)^*(bb)^*$ is shown in 
figure (\ref{fig: FA}).

\begin{figure}[ht]
    \centering 
    \begin{tikzpicture}
        \node[state,           ] (4) {$4$};
        \node[state, left  of=4] (2) {$2$};
        \node[state, right of=4] (6) {$6$};
        \node[state, initial, below of=2] (1) {$1$};
        \node[state, below of=4] (3) {$3$};
        \node[state, accepting, right of=3] (5) {$5$};

        \draw 
        (1) edge[above] node{$\epsilon$} (3)
        (3) edge[above] node{$\epsilon$} (5)

        (1) edge[bend left,left] node{$a$} (2)
        (2) edge[bend left,right] node{$b$} (1)

        (3) edge[bend left,left] node{$a$} (4)
        (4) edge[bend left,right] node{$\epsilon$} (3)

        (5) edge[bend left,left] node{$b$} (6)
        (6) edge[bend left,right] node{$b$} (5);
    \end{tikzpicture}
    \caption{A $\langle 2,3 \rangle$-flat automaton 
that recognizes $L\Def  (ab)^*(a)^*(bb)^*$}
    \label{fig: FA}
\end{figure}

\paragraph{Generic Flat Languages and Automata}
Fix $\alpha = \langle p,q \rangle$,
we define the \emph{generic $\alpha$-flat language} is the union of all $\alpha$-flat languages, denoted by $\mathbb{F}(\alpha)$.
Now, we try to define an automaton that recognizes all $\alpha$-flat languages,
i.e., collects all behaviors of $\alpha$-flat automata.

Intuitively, 
the generic automaton is obtained by introducing a new alphabet (
a multi-set with $p q$ copies of the original alphabet) and 
adding more transitions (labels),
the states and the overall framework remain unchanged. 
In details, a generic $\alpha$-flat automaton is still a finite state automaton over
$\Sigma(\alpha)\Def \{a(i)| (a\in \Sigma_\epsilon) \wedge i\in \mathbb{N}:1\le i \le pq\}\cup \{\epsilon\}$.
The states are still named from $1$ to $pq$, 
the initial state is $1$ and the accepting state is $(pq-p+1)$.
The transition relations for state $i$ are defined as 
\begin{itemize}
    \item if $i\  \text{mod}\  p = 1$ and $i\neq pq-p+1$, then 
    $(i,\epsilon,i+p)\in \Delta$
    and $\forall s\in \Sigma_{\epsilon}. (i, s(i) ,i+1) \in \Delta$;
    \item if $i\  \text{mod}\  p = 0$, 
    $\forall s \in \Sigma_{\epsilon}. (i,s(i), i-p+1) \in \Delta$;
    \item otherwise, $\forall s \in \Sigma_{\epsilon}. (i,s(i), i+1) \in \Delta$.
\end{itemize}

For $\Sigma = \{a,b\}$, an example of generic $\langle 2,3 \rangle$-flat automaton is shown in figure (\ref{fig: GFA}).

\begin{figure}[ht]
    \centering 
    \begin{tikzpicture}
        \node[state,           ] (4) {$4$};
        \node[state, left  = 2cm of 4] (2) {$2$};
        \node[state, right = 2cm of 4] (6) {$6$};
        \node[state, initial, below of=2] (1) {$1$};
        \node[state, below of=4] (3) {$3$};
        \node[state, accepting, below of=6] (5) {$5$};

        \draw 
        (1) edge[above] node{$\epsilon$} (3)
        (3) edge[above] node{$\epsilon$} (5)

        (1) edge[bend left, pos =0.2 ,left] node{$a(1)$} (2)
        (1) edge[bend left, pos =0.5 ,left] node{$b(1)$} (2)
        (1) edge[bend left, pos =0.8 ,left] node{$\epsilon(1)$} (2)

        (2) edge[bend left, pos = 0.2 ,right] node{$\epsilon(2)$} (1)
        (2) edge[bend left, pos = 0.5 ,right] 
        node{$b(2)$} (1)
        (2) edge[bend left, pos = 0.8 ,right] 
        node{$a(2)$} (1)

        (3) edge[bend left, pos =0.2 ,left] node{$a(3)$} (4)
        (3) edge[bend left, pos =0.5 ,left] node{$b(3)$} (4)
        (3) edge[bend left, pos =0.8 ,left] node{$\epsilon(3)$} (4)

        (4) edge[bend left, pos = 0.2 ,right] node{$\epsilon(4)$} (3)
        (4) edge[bend left, pos = 0.5 ,right] 
        node{$b(4)$} (3)
        (4) edge[bend left, pos = 0.8 ,right] 
        node{$a(4)$} (3)

        (5) edge[bend left, pos =0.2 ,left] node{$a(5)$} (6)
        (5) edge[bend left, pos =0.5 ,left] node{$b(5)$} (6)
        (5) edge[bend left, pos =0.8 ,left] node{$\epsilon(5)$} (6)

        (6) edge[bend left, pos = 0.2 ,right] node{$\epsilon(6)$} (5)
        (6) edge[bend left, pos = 0.5 ,right] 
        node{$b(6)$} (5)
        (6) edge[bend left, pos = 0.8 ,right] 
        node{$a(6)$} (5);
    \end{tikzpicture}
    \caption{The generic $\langle 2,3 \rangle$-flat automaton}
    \label{fig: GFA}
\end{figure}

However,
the resulted automaton may accept languages that are not in $\mathbb{F}(\alpha)$,
because in different passes inside a loop, 
the automaton can choose different symbols between identical pairs. 
To avoid this problem,
we add a so-called purity condition on the accepting language of generic flat automata,
which is equivalent to intersecting the language of a generic flat automaton 
with a language that encodes the purity condition.

We say a string $w\in (\Sigma(\alpha))^*$ is pure if for all $i: 1\le i \le p q$,
and $a,b\in \Sigma$, 
$a\neq b \wedge \#w(a(i))>0$ implies $\#w(b(i))=0$.
Formally, the purity condition is defined by 
\begin{equation} \label{eq:purity}
 \bigwedge_{1\le i\le pq}\bigwedge_{a,b\in \Sigma, a\neq b} ({a(i)}^\#>0)\to ({b(i)}^\#=0)\, . 
\end{equation}

We denote the accepting language of the generic $\alpha$-flat automaton by $\mathbb{G}(\alpha)$.
Note that $\mathbb{G}(\alpha)$ is a language over $\Sigma_\alpha$,
but what we want is a language over $\Sigma$.
So we define a renaming function $R:\Sigma(\alpha)\mapsto \Sigma$ such that for all $a(i) \in \Sigma_\alpha, R(a(i))=a$,
and $R(\epsilon) = \epsilon$.
Define $\mathbb{G}'(\alpha) \, \Def \, \{R(w) \mid w\in \mathbb{G}(\alpha)\}$, 
for simplicity, we write
$\mathbb{G}'(\alpha)=R(\mathbb{G}(\alpha))$.

The important feature of generic flat autamata
is that every word $w\in \mathbb{G}(\alpha)$ is uniquely determined by its Parikh image $\mathbb{P}(w)$.

\subsection{Flattening}

The flattening technique was first introduced 
in \cite{Abdulla 2017}.
Fix an alphabet $\Sigma$ and an abstraction parameter $\alpha$, 
for a given atomic string constraint $\phi$, 
flattening $\phi$ with parameter $\alpha$ 
results in a new string constraint $\phi_\alpha$,
such that 
$R(\lVert \phi_\alpha \rVert) = \lVert \phi \rVert \cap \{I \mid \forall x\in X, I(x)\in \mathbb{G}'(\alpha)\}$,
where $R$ is the renaming function with its domain extended to interpretations in the normal manner.
Intuitively, it restricts $\phi$ to interpret 
string variables over $\mathbb{G}'(\alpha)$.

\cite{Abdulla 2017} discussed the flattening of basic 
string constraints including equality, integer, (regular) grammar and transducer constraints. 
For an atomic string constraint $\phi$,
the flattening $\phi_\alpha$ is still an atomic string constraint 
and is Parikh definable,
so its Parikh image can be expressed by a quantifier free PA formula.
Together with the purity condition,
we obtain an existential quantified PA formula $\rho$.
$\rho$ will be sent to a SMT solver,
if the solver returns an solution $\theta$,
then we can construct an interpretation for $\phi_\alpha$ from $\theta$,
otherwise it means $\phi$ is unsatistiable when 
string variables are interpreted  to $\alpha$-flat languages.

Take a regular constraint $\phi = x\in L(A)$ for example,
the flattening of $\phi$ resutls in a new finite state automaton $A'$ over $\Sigma(\alpha)$,
which encodes running $A$ ``in parallel" 
with the generic $\alpha$-flat automaton.
Let $\rho_1$ be the formula describing the Parikh image of $A'$,
which is a formula over variable sets $\Sigma(\alpha)^\#$.
Let $\rho_2$ be the purity condition \eqref{eq:purity}.
Then we obtain the PA formula $\exists (\Sigma(\alpha))^\#. \rho_1 \wedge \rho_2$.
In order to distinguish between different string variables,
we may replace $a(i)^\# \in \Sigma(\alpha)^\#$ by $(x,a(i))^\#$.

Since the structure of a flat automaton is decided 
by its abstraction parameter $\alpha$, 
a Counter-Example Guided Abstraction 
Refinement (CEGAR) framework is designed, which contains both an under- and an 
over-approximation module, to search the possible values of $\alpha$.
The termination for the overall algorithm is not guaranteed.

\section{Solving String Constraints with \textsf{parseInt} Function}\label{sec:string-solving}

The string-number conversion functions are commonly used functions
in most of programming languages,
for example,
\verb+parseInt()+ in Java and \verb+Int()+ in Python.
The functions usually take two parameters, 
a string over the agreed alphabet $\Sigma$
and an optional parameter denotes the radix.
They parse the string according to the rules indicated by the radix,
and return an integer denoted by the string.

From the view of string constraints,
string-number conversion functions give rise to a new form of string constraints
and are more expressive than length constraints.
So we consider extending string constraints with 
{\parseInt} function. 
As the general problem of string constraints is undecidable,
we still adopt the idea of flattening, 
i.e., variables are restricted to (generic) flat languages.
This problem has been investigated in \cite{POPL20},
which defined a special form of flat restriction (straight-line PFA) and 
proposed a heuristic search method.

In this section, 
we describe the problem of interest first, and then 
present an reduction from the problem of solving flat string constraints with {\parseInt} function
to the decidability problem of Presburger Arithmetic with exponential functions.
Hence, we identify a decidable subset of string constraints, which is the largest one with decidability so far to the best of our knowledge.

\subsection{String-Number Conversion Function}

Commonly, 
{\parseInt} function takes a string representation of a decimal
number and returns an integer.
Since we focus on the decidability,
we define a binary version of {\parseInt},
which takes a binary string and returns a decimal integer number.
For example,
${\parseInt}('111')=7$. 
Clearly,  our decision procedure given in this paper 
can be adapted to string constraints with other string to number conversion function without substantial change. 
Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.

\begin{definition} Let $\Sigma_{\textit{num}}=\{'0','1'\}$,
 ${\parseInt}:\Sigma_{\textit{num}}^*\mapsto \mathbb{N}$ is recursively defined by:
    for $w\in \Sigma_{\textit{num}}^*$
    \begin{itemize}
        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
        \item if $|w|=1, {\parseInt}('w')=w_\mathbb{N}$;
        \item for $|w|\ge 2$ and $w=w_1 \cdot w_2$, 
        where $|w_1|, |w_2| \ge 1$, 
        ${\parseInt}(w) = {\parseInt}(w_1)2^{|w_2|}+{\parseInt}(w_2)$.
    \end{itemize} 
\end{definition}

Now, we introduce a new form of atomic string constraints: 
a {\parseInt} constraint $\phi$ is of the form 
$n \sim {\parseInt}(t)$,
where $n$ is an integer term, $\sim \in \{\le,<,=,>,\ge\}$ and $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$ is a string term.
$\lVert \phi\rVert \Def  \{I \mid I(n)\sim {\parseInt}(I(t))\}$.

In what follows, we only  consider the problem in the case when $t=x$ and $x$ is restricted to (generic) flat languages. 
%If $t$ is not a single variable, 
For the general case $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$, 
it can be reduced to this special case by induction on the structure of $t$. 
%we can reduce the problem by separating $t$ (corresponding to the $|w|>2$ case in definition).

Given an $\alpha$-flat language $L$,
we assume $\alpha=\langle p,q \rangle$ and $L=(w_1)^*...(w_q)^*$,
where $p,q$ and $w_i(1\le i \le q) $ are known.
We further assume that $x = (w_1)^{\beta_1} ... (w_q)^{\beta_q}$,
then we have
\begin{align}
    &{\parseInt}(x) \notag \\
    =&{\parseInt}((w_1)^{\beta_1} ... (w_q)^{\beta_q}) \notag \\
    =&{\parseInt}((w_1)^{\beta_1} ... (w_{q-1})^{\beta_{q-1}})\cdot 2^{\beta_q |w_q|}
    + {\parseInt}((w_q)^{\beta_q}) \label{parse}
\end{align}
(\ref{parse}) is a recursive expression.
So we only need to deal with the basic case ${\parseInt}((w_q)^{\beta_q})$, 
where $w_q \neq \epsilon$
\begin{align}
    {\parseInt}((w_q)^{\beta_q}) &= \sum_{i=0}^{\beta_q-1} {\parseInt}(w_q)\cdot 2^{|w_q|\cdot i}  \notag \\
    &={\parseInt}(w_q)\frac{2^{|w_q|\cdot \beta_q}-1}{2^{|w_q|}-1}
    \label{parseInt-basic}
\end{align}
In (\ref{parseInt-basic}), since $w_q$ and $|w_q|$ are known, 
they can be regarded as constants.
The only unknown variable is $\beta_q$.

Combine (\ref{parse}) and (\ref{parseInt-basic}),
the constraint $n={\parseInt}(x)$ can be expressed by an arithmetic expression with 
$n$ and $(\beta_1,...,\beta_q)$,
inevitably with exponential components.

Take $n={\parseInt}((11)^a(10)^b)$ for example.
\begin{align}
    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
\end{align}

So we have 
\begin{equation}
    3\cdot n = 3\cdot 2^{2a+2b}-3\cdot 2^{2b}+2\cdot 2^{2b}-2. \notag
\end{equation}

Observe the form of the above equation,
$a,b,n$ are integer variables and either occur in 
an exponential term or a linear term.
This is always the case,
so the problem can be reduced to 
the decidability of PA with exponential function.

When $x$ of {\parseInt} constraints is restricted to the
generic $\alpha$-flat language ($\alpha$ is fixed),
the (\ref{parse}) and (\ref{parseInt-basic}) still hold 
but $w_i(1\le i\le q)$ is known.
However,
by definition,
the generic $\alpha$-flat language is the finite union of all $\alpha$-flat languages,
so we can enumerate all possible values for $w_i$.
In this way,
the problem can still be reduced to the decidability of PA with exponential function (\paexp), i.e.,
\begin{theorem} \label{thm:string-parInt}
If {\paexp} is decidable, then the satisfiability (validity) of string constraints with {\parseInt} in which all string variables 
are ranged over flat strings is decidable. 
\end{theorem}

\subsection{Decidability of \paexp} \label{sec:3.2}

For a first order theory $T$,
we say theory $T$ admits quantifier elimination (QE) if for any formula in $T$, 
there is a quantifier-free formula equivalent to it.
It is well-known that if a theory admits QE, 
then it is a decidable theory.

The formal definition of PA is given in section \ref{PA}.
Here we add the ordering predicate $\le$ into the signature,
which can be defined by $x \le y\, \Def \, \exists z. x+z=y$.
However, the theory PA so far does not admit QE, 
for example, consider the formula $\exists x.x = y + y$. 
We augment the theory with countable unary divisible predicates
$n|x$, where $n\in \mathbb{N}$, 
$n|x$ is true if and only if $x\ \text{mod}\ n=0$ holds.
This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.

We then introduce a theory {\paexp}, denoted by $(\mathbb{N},+,2^x)$, that we work on.
\begin{definition}
    Let $\mathcal{L}=\{0,1,+,\le,n \, |x(n\in \mathbb{N}),2^x,l_2(x)\}$, 
     $(\mathbb{N},+,2^x)$ be a $\mathcal{L}$-theory that has domain $\mathbb{N}$, where 
    \begin{itemize}
        \item  $2^x$ is interpreted to the exponential function of $2$ over $\mathbb{N}$; 
        \item interpretations of $0,1,+,\le,=$ are consistent with PA;
        \item for $n\ge 1$,$n|x$ holds iff $\exists y.x=ny$;
        \item $2^0=1$, for $n \ge 1, 2^n = 2^{n-1}+2^{n-1}$;
        We further assume that if $m,n\in \mathbb{N}, m\le n$, 
        then $2^{m-n} = 1$
        \item $l_2(0)=0$; for $n \ge 1,l_2(n) = y$ iff $2^y \le n < 2^{y+1}$;  $l_2(m-n) = 0$
        if $m,n\in \mathbb{N}$ and $m\le n$. 
    \end{itemize}
\end{definition}

$\lambda_2(x) = 2^{l_2(x)}$ can be defined by $l_2(x)$,
intuitively, $\lambda_2(x)$ means the maximal power of 2 that is not larger than $x$.
Then we have $\lambda_2(x) \le x \le 2\lambda_2(x)-1$,
which will be useful in our proof.

\begin{definition}
    For a strictly increasing function $f:\mathbb{N}\mapsto \mathbb{N}$, 
    $f$ is said to be compatible with addition 
    if for every $m\in \mathbb{N}^+$,  $f$ modulo $m$ is periodic, 
    and for any term $A(x)=\sum_{1\le i\le n} a_i f(x+b_i)$, 
    where $n \in \mathbb{N}^+, a_i ,b_i \in \mathbb{Z}$,
    one of the following holds:
    \begin{itemize}
        \item $A(x)$ is bounded; 
        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
        $\forall x. A(x+\Delta_A)\ge f(x)$;
        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
        $\forall x. -A(x+\Delta_A)\ge f(x)$.
    \end{itemize}
\end{definition}

%\znjSide{\textbf{Solved}. I think you need to explain the meanings of compatibility.}

Semenov proved that for any function $f$ that is compatible with addition, 
theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov}. 
Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
by \cite{semenov}.
Particularly, in \cite{Francoise} Francoise gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
Unfortunately, Francoise's algorithm is flawed, one problem 
 lies in the discussion to eliminate exponential terms. 
% making the algorithm fail to handle "$a_0<0$" case.
For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
the expected result should be $y\le 2^{10} =1024$,
but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
% 这里好像还有另一个与Francoise一起的作者

In the next section, within Francoise's framework, we give a revised algorithm
 to the decision problem of {\paexp} with 
both subtle improvements and crucial corrections, that 
provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to 
Theorem~\ref{thm:string-parInt}.

\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
Based on Francoise's work \cite{Francoise}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 
Since $\forall x. F = \neg \exists x. \neg F$, we further assume the quantifier $Q$ to be the existential quantifier, 
that is, $\exists x.\theta(x,\bar{y})$. 
For eliminating quantifiers in arbitrary formula, 
we apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.

We divide the whole QE algorithm into 4 steps.
The first step $\textbf{Normalization}$ can be viewed as a pre-processing step, which substitutes ``complex" terms like $2^{2^x}$, $l_2(3x+y)$ with ``simple" terms by introducing new variables $x_i$. Now we move forward to handle a ``simpler" formula, however, at the cost of more quantified variables.

The other three steps undertake the task to eliminate the introduced variables $x_i$ and the original variable $x$ (will be denoted by $x_0$) one by one. 
The main body of \textbf{QE-with-Order} step is a loop to enumerate all possible orders among quantified variables. 
In each iteration, according the given (decreasing) order (corresponding to a for loop), we invoke \textbf{QE-exp} or 
\textbf{QE-linear} to eliminate these variables one by one. 
%depending on  eliminates the maximal variable in the remaining $\bar{x}$. So it is necessary to specify an ordering for $\bar{x}$ at the beginning, as we will see, this is the main reason for the complexity of this algorithm.
During each iteration of the inner loop (the for loop), 
if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atomic formula, 
we will invoke \textbf{QE-exp} to produce a formula equivalent to the atom where $x_i$ occurs linearly;  
otherwise, if $x_i$ occurs linearly in the formula, \textbf{QE-linear} will be invoked to eliminate all occurrences of $x_i$, 
and this procedure is similar to the classic QE algorithm for PA.

A more detailed description is given below.

\subsection{Normalization}

In order to show $(\mathbb{N},+,2^x)$ admits QE , 
it is sufficient to show that any 1-existential formula $\exists x.\theta(x,\bar{y})$ indeed does, 
where $\theta(x,\bar{y})$ is a quantifier-free formula. 
However, the form of $\exists x.\theta(x,\bar{y})$ is unknown so it may contain terms difficult to handle such as $2^{3x+y+1}$ or $l_2(x)$.
The \textbf{Normalization} step simplifies these terms by introducing new variables, 
for example, $\exists x.2^{3x+y+1}>10$ is equivalent to
$$\exists x\exists x_1.2^{x_1}>10 \wedge x_1 =3x+y+1\,.$$

Logarithm functions are replaced by exponential functions, take $\exists x.l_2(x)>3$ for example, $l_2(x)$ is replaced by a new variable $x_1$, and we have  
$$\exists x\exists x_1. x_1 > 3 \wedge 2^{x_1}\le x \wedge x\le 2^{x_1+1}-1\,.$$

The \textbf{Normalization} step goes like this, first transform the given formula $\theta(x,\bar{y})$ into the Negation Normal Form (NNF).
Then $\theta(x,\bar{y})$ becomes a Boolean combination (with only $\wedge$ and $\vee$) of literals.

Make substitutions by introducing new variables according to the \textit{while} statements in the pseudo-code.For consistency, we rename the original $x$ to be $x_0$ and assume $n$ new quantified variables are introduced.

After introducing new variables and substitution, 
we obtain a formula where if an exponential term contains a quantified variable $x_i(0\le i\le n)$, the term should be $2^{x_i}$.
We then collect all terms with $x_i(0\le i\le n)$ together, 
it will be of the form $s(\bar{x})\Def \sum_{i=0}^{n} a_i 2^{x_i} + \sum_{i=0}^n b_i x_i$, where $a_i,b_i(0\le i \le n)$ are all constants.
Other terms including constants and terms of $\bar{y}$ are collected,  
denoted by $t(\bar{y})$. 
Since $\bar{y}$ are free variables, 
we will regard $t(\bar{y})$ as a constant.
Inequalities and equalities will all be transformed into $s(\bar{x})\le t(\bar{y})$, for example, $s(\bar{x}) = t(\bar{y})$ will be replaced by $s(\bar{x})\le t(\bar{y})\wedge t(\bar{y})\le s(\bar{x})$.

At the end, the resulted formula $\theta'$ only contains literals of the forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ and $\neg  (k|s(\bar{x})+t(\bar{y}))$.

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{1-existential formula $\exists x.\theta(x,\bar{y})$}
    \KwOut{n-existential formula $\exists \bar{x}.\theta'(\bar{x},\bar{y})$}

    $\theta' :=  \theta(x,\bar{y})$ \;
    Transform $\theta'$ into NNF\;
    \tcp{ i is used for counting the introduced variables }
    i = 1\; 
    \tcp{ $\bar{y}$ are regarded as constants} 
    \While{there is a term $l_2(t)$, $t$ is not a constant}
    {
        $\theta' :=  \theta'[x_i/l_2(t)] \wedge (2^{x_i}\le t) \wedge (t\le 2^{x_i+1}-1)$\;
        $i:= i+1$\;
    }
    \While{there is a term $2^t$, $t$ is not a variable $x_j(j<i)$ or a constant}
    {
        $\theta' :=  \theta'[2^{x_i}/2^t]\wedge (x_i\le t) \wedge (t \le x_i)$\;
        $i:= i+1$\;
    }
    $n:= i-1$\;
    \tcp{ Collect quantified variables $x_i$}
    Transform all atoms into forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ or $\neg  (k|s(\bar{x})+t(\bar{y}))$\;
    Return $\exists x_0,...,\exists x_n. \theta'$

    \caption{Normalization}
\end{algorithm}

\subsection{QE-with-order} 
During the \textbf{Normalization} step, 
we simplify the origin formula at the cost of more quantified variables. 
Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
and it has $n$ new variables $x_i$ ($x$ is denoted by $x_0$).
As a consequence, we need to eliminate all quantified variables one by one.

To the end, we denote the set of all orders among the $n+1$ variables by 
 $S_{n+1}$, and then enumerate these orders one by one in the outer \textit{for} loop. 
 For the considered order $\sigma$, we first add the ordering information to 
 the quantifier free formula, and then according to the 
 order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{QE-exp} first, and then invoking \textbf{QE-linear} (the inner \textit{for} loop). 
 In each iteration of the inner \textit{for} loop, 
if $x_{\sigma{(i)}}$ occurs in an exponential term in an atomic formula, 
\textbf{QE-exp} is invoked and a formula   equivalent to the atom is returned, in which $x_i$ occurs linearly;  
then $x_i$ occurs linearly in the formula, thus \textbf{QE-linear} is further invoked to eliminate all occurrences of $x_i$. 
The returned formula is the disjunction of all formulas resulted in all iterations of the outer \textit{for} loop.

\iffalse
put , We now specify an order of $\bar{x}$, and recursively eliminate the maximal element in the remaining $\bar{x}$. 
That is to say,
every execution of \textbf{QE-exp} and \textbf{QE-linear} will eliminate the largest element $x_i$ in $\bar{x}$.
However, as you may notice, if there is no clue about the order of $\bar{x}$, 
we will have to use the idea of permutation group and the number of sub-cases will blow up, in the worst case, $(n+1)!$ cases.

Here is how \textbf{QE-exp} and \textbf{QE-linear} are invoked in
\textbf{Ordering and QE}. 
\fi

%我们这里还是考虑用循环群来做，因为如果每一次只找到最大的$x_i$，那么递归的执行还是需要整个$\bar{x}$的顺序。但如果一开始就知道了，后面有些操作可以简化

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{a normalized $(n+1)$-existential formula $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
    where $\bar{x}=(x_0,...,x_n)$} 
    \KwOut{a equivalent quantifier free formula without $\bar{x}$}

    Let $S_{n+1}$ denote the group of permutations on $\{0,1,...,n\}$\;
    $\phi :=  \textit{False}$\;
    \For{each $\sigma \in S_{n+1}$ }
    {
        $\theta_{\sigma}:=  \theta(\bar{x},\bar{y})\wedge \bigwedge_{j=0}^{n-1}(x_{\sigma(j)}\le x_{\sigma(j+1)})$\;

        \tcp{recursively eliminate the maximal $x_i$ in $\bar{x}$}
        %\znj{How to effectively determine the maximal $x_i$?} 
        \For{i from n to 0}
        {
            \tcp{if $x_{\sigma(i)}$ occurs exponentially in $\theta_\sigma$}
            $\theta_{\sigma}:= \textbf{QE-exp}(x_{\sigma(i)},\theta_{\sigma})$\;
            \tcp{now $x_{\sigma(i)}$ occurs only linearly in $\theta_\sigma$}
            $\theta_{\sigma}:= \textbf{QE-linear}(x_{\sigma(i)},\theta_{\sigma})$\;
        }
        $\phi :=  \phi \vee \theta_{\sigma}$\;
    }
\iffalse
    \For{i from $0$ to $n$}
    {
        \tcp{ specify the maximal quantified variable}
        $\theta_i :=  \theta(\bar{x},\bar{y}) \wedge (\bigwedge_{0\le j \le n} x_i - x_j \ge 0)$\;
        \tcp{ eliminate $x_i$}
        \While{$x_i$ occurs in an exponential term in formula $\tau(x_i,\bar{x},\bar{y})$}{
            \tcp{ QE-exp outputs a formula equivalent with $\tau(x_i,\bar{x},\bar{y})$ where $x_i$ occurs linearly }
            $\theta_i :=  \theta_i[\text{QE-exp}(\tau(x_i,\bar{x},\bar{y}))/\tau(x_i,\bar{x},\bar{y})]$\;
        }   
        transform $\theta_i$ into disjunction normal form $\theta_i = \bigvee_s \theta_{i,s}$\;
        \For{ every $\theta_{i,s}$ }{
            $\theta'_{i,s} :=  \text{QE-linear}(x_i,\theta_{i,s})$
        }
        $\theta'_i =\bigvee_s \theta'_{i,s}$\;
    }
\fi
    output $\phi$
    \caption{QE-with-order}
\end{algorithm}
%\znj{I cannot understand why you need to do QE for all permutation $\sigma$? I think 
%one is enough.}

\subsection{QE-exp}

\textbf{QE-exp} and \textbf{QE-linear} are the most technical parts of our QE algorithm.
As mentioned, 
\textbf{QE-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as 
inputs, and  outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{QE-linear}.
For simplicity, we will abuse   $x_i$ for $x_{\sigma(i)}$, 
and  $\bar{x}$ for $(x_{\sigma(0)},...,x_{\sigma(i-1)})$ ( remind that $x_{\sigma(i+1)},...,x_n$ have been eliminated already).

After normalization, the formula $\theta(\bar{x},\bar{y})$ contains atoms of three forms corresponding to 
the predicates $\le$, $|$ and negation of $|$, respectively.
The problem will be discussed in 2 cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
corresponding to \textbf{QE-exp-ineq} or \textbf{QE-exp-div}.

We first give the whole procedure of \textbf{QE-exp},
and then provide the details of the two sub-routines.
\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{$x_i$ and $\theta$, $x_i$ is larger than $x_0$ to $x_{i-1}$
    and occurs exponentially in $\theta$} 
    \KwOut{a formula equivalent to $\theta$ where $x_i$ occurs linearly}

    \While{there is an atom $\tau(x_i,\bar{x},\bar{y})$ contains $2^{x_i}$}
    {
        $\Psi:= \textit{False}$\;
        \eIf{$\tau$ is of the form $a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$}
        { 
            \tcp{ \textbf{QE-exp-ineq} case}
            $A :=  \sum_{j=0}^{i-1}|a_j|$, $B:= b_i + \sum_{j=0}^{i-1}|b_j|$\;
            $B':= 2(l_2(B)+3)$, $\delta:=  l_2(A)+3$\;
            If $a_i>0$ then $\alpha := l_2(t(y))-l_2(a_i)$
            else $\alpha := l_2(-t(y))-l_2(-a_i)$\;

            $\rho_1 :=  (x_i \le \alpha -1 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$

            $\quad \vee ( x_i \le \alpha -1 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta} (x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))$\;

            $\rho_2 :=  (x_i = \alpha) \wedge \tau[\alpha/x_i]$\;
            $\rho_3 :=  (x_i = \alpha+1) \wedge \tau[\alpha+1/x_i]$\;
            $\rho_4 :=  (x_i \ge \alpha+2 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$

            $\quad \vee (x_i \ge \alpha +2 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta}(x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))
            $\;
            \eIf{$a_i>0$}
            {
                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
                \vee [x_i \le \alpha -1 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$
            }
            {
                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
                \vee [x_i \ge \alpha +2 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$    
            }
            }
        {
            \tcp{ \textbf{QE-exp-div} case, the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
            \tcp{ let $d=2^r d_0$ where $d_0$ is odd}
            $\rho_5 := \bigvee_{p=0}^{r-1} [\tau(p,\bar{x},\bar{y})\wedge x=p]$\;
            $\rho_6 := \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
            +  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r]
            $\;
            $\Psi :=  \rho_5 \vee \rho_6$
        }
        replace $\tau$ by $\Psi$ in $\theta$\;
    }
    \caption{QE-exp}
\end{algorithm}

\subsubsection{QE-exp-ineq}

This algorithm deals with the case where $\tau$ is an inequality of the form 
$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$

We now try to eliminate $2^{x_i}$ in $\tau$, the idea is to find a bound for $x_i$, 
either by constants or by other variables.
We will prove the following theorem.

\begin{theorem} \label{thm:exp-ineq}
Assume that $x_i$ is the maximal variable in $\bar{x}$ according to the given order. 
Given an inequality $\tau(x_i,\bar{x},\bar{y})$ of the form %, suppose $x_i$ occurs in an exponential term, that is 
$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
with $a_i \neq 0$, 
let $A\Def \sum_{j=0}^{i-1}|a_j|$, 
$B\Def |b_i| + \sum_{j=0}^{i-1}|b_j|$, 
$B'\Def 2(l_2(B)+3)$,
$\delta\Def  l_2(A)+3$, then 
\begin{itemize}
    \item if $a_i > 0$, let $\alpha \Def l_2(t(y))-l_2(a_i)$
    \begin{itemize}
        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
    \end{itemize}
    \item if $a_i < 0$, let $\alpha \Def l_2(-t(y))-l_2(-(a_i))$
    \begin{itemize}
        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
    \end{itemize}
\end{itemize}
\end{theorem}

Before proving \textbf{Theorem 1}, we need  the following lemma to estimate linear terms.

\begin{lemma} \label{lem:1}
For any $n,m\in \mathbb{N}$, if $n\ge m\ge 1$ and $x\ge 2(l_2(n)-l_2(m)+1)$, then 
$nx\le m2^x$ holds.
\end{lemma}

\begin{proof}
First we can prove that for any $N\in \mathbb{N}$, $x\ge 2N \implies 2^N x\le 2^x$. Let $N\Def l_2(n)-l_2(m)+1$, then we have $x\ge  2N \implies n x\le 2\lambda_2(n) x\le 2^N \lambda_2(m) x \le m 2^x $. \qed 
\end{proof}

Then we give the proof for Theorem~\ref{thm:exp-ineq}.
\begin{proof}
We only prove for the $a_i > 0$ case, the other case is analogous. 
The goal is to find a bound for $x_i$ such that 
the values of the atoms containing $x_i$ keep constant when $x_i$ is greater than the bound.

Note that $x_i$ is the largest among $\bar{x}$. 
suppose $x_i > x_{i-1} + \delta$, and let 
 $\delta \Def  l_2(A)+3$, then 
 \begin{equation} 
   2^{-\delta}A = \frac{A}{8\lambda_2(A)}\le \frac{1}{4}.   \label{eq:thm-ineq-1}
 \end{equation}
 When $x_i \ge B' = 2(l_2(4B)-l_2(1)+1)$, according to Lemma~\ref{lem:1}, 
\begin{equation} 
    4 B x_i \le 2^{x_i}. \label{eq:thm-ineq-2}
\end{equation}

When $x_i \ge \alpha+2$,
\begin{align}
    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
      (\mbox{ Def. of } \alpha) & \notag \\
     & \ge  t(\bar{y}) & \notag 
\end{align}
So we conclude that %when $x_i \ge \alpha+2,x_i\ge B',x_i \ge x_{i-1}+\delta$,
$\tau(x_i,\bar{x},\bar{y})$ keeps \textit{false} in this case.

%\znj{I think there are some mistakes in the above proof, pls check it!}
When $x_i \le \alpha-1$, similarly we have
\begin{align}
    \tau(x_i,\bar{x},\bar{y})\notag 
    & \le  a_i 2^{x_i} + 2^{x_i -\delta}A  + Bx_i &  \notag \\
    & \le  2^{x_i}(a_i +\frac{1}{4} +\frac{1}{4}) &(\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}}) \notag \\
   & \le  2\lambda_2(a_i)\frac{\lambda_2(t(\bar{y}))}{2\lambda_2(a_i)} &  (\mbox{ Def. of } \alpha) \notag \\
   &  \le  t(\bar{y}) & \notag 
\end{align}
which indicates that when  $x_i \le \alpha-1$, $\tau(x_i,\bar{x},\bar{y})$ keeps \textit{true}. \qed 
\end{proof}

\subsubsection{QE-exp-div}

If $x_i$ occurs exponentially in an atom $\tau(x_i,\bar{x},\bar{y})$ of the form
$$\tau(x_i,\bar{x},\bar{y}) \Def  d| a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k+t(\bar{y}) \quad (a_i \neq 0)\,,$$
the algorithm \textbf{QE-exp-div} outputs an equivalent formula without $2^{x_i}$ terms.
The idea is that $a_i 2^x$ modulo $d$ is a periodic function when $x$ is large enough and the period can be computed.

Let $d = 2^rd_0$, $d_0$ is an odd natural number. 
According to Euler's Theorem, 
$gcd(2,d_0)=1$ implies
$2^{\phi(d_0)} \text{mod}\ d_0 = 1$,
where $\phi$ is the Euler function.
Consider function $f(n)=(2^n\ \text{mod} \ d)$,
when $n\ge r$, 
$f(n)$ becomes a periodic function and its period is a divisor of $\phi(d_0)$ because
\begin{align}
    f(n+\phi(d_0)) 
    &=2^{n+\phi(d_0)}\ \text{mod}\ d &\notag\\
    &=2^n\cdot 2^{\phi(d_0)}\ \text{mod}\ d &\notag\\
    &=2^n\cdot (k d_0 + 1) \ \text{mod}\ d &
    (\mbox{assume } 2^{\phi(d_0)}=k d_0 +1) \notag \\
    &= 2^n \ \text{mod}\ d &
    (\mbox{when } n\ge r, 2^n d_0\ \text{mod}\ d = 0 )\notag\\
    &= f(n) \notag
\end{align}

When $x_i\le r-1$, we just enumerate all possible value of $x_i$, i.e., 
$$\rho_5 \Def \bigvee_{p=0}^{r-1} \tau(p,\bar{x},\bar{y})\,.$$
When $x_i\ge r$, $\tau(x_i,\bar{x},\bar{y})$ is equivalent to
$$
\rho_6 \Def \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
+  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r] \,.
$$
            Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.

%\znj{\textbf{Solved}.More explanation or a formal proof is needed. }

\subsection{QE-linear}

After \textbf{QE-exp}, 
we obtain a formula (still denoted by $\theta(x_i,\bar{x},\bar{y})$) without $2^{x_i}$ terms,
i.e., $x_i$ occurs only linearly.
Now we wish to construct a formula without $x_i$ equivalent to $\exists x_i.\theta(x_i,\bar{x},\bar{y})$.
This can be done by 
following Cooper's QE algorithm for PA,
treating all $x_j (j<i)$ as free variables (like $\bar{y}$).
The procedure contains 3 steps.

Step 1: Put $\theta(x_i,\bar{x},\bar{y})$ in NNF form, replace atomic formulae containing symbols other than $\le,|$ by equivalent formulae only with $\le$, that is, $x=a$ by $x\le a \wedge a\le x$, $x< a$ by $x\le a+1$, $x\neq a$ by $x>a \vee x<a$,  $x\geq a$ by $-x \le -a$, and $x>a$ by $-x \le -a-1$.
In inequality atoms, collect terms of $x_i$
to one side and guarantee the coefficients of $x_i$ are positive.

Step 2: Let $d$ be the least common multiple of all coefficients of $x_i$. 
For atoms with $x_i$, multiply all terms by a factor so that the coefficient of $x_i$ is $d$.
We introduce a fresh variable $x_i'$, replace all occurrence of $dx_i$ by $x_i'$,
and denote the resulted formula by $\theta'(x_i')$.
Note that $x_i'$ is a multiple of $d$, 
so we set $\theta'=\theta'\wedge d|x'$.

We classify all atoms in $\theta$ into the following three sets:  
$L$ denotes all atoms of the form $t_l(\bar{x},\bar{y})\le x_i'$,
$U$ denotes all atoms of the form $x_i'\le t_u(\bar{x},\bar{y})$,
$M$ denotes all atoms of the form $k_m|x_i'+t_m(\bar{x},\bar{y})$ and their negations,
where $t(\bar{x},\bar{y})$ with a subscript is any term of $\bar{x}$ and $\bar{y}$ and 
$k_m \in \mathbb{N}$ for $m\in M$. % are integers in $$. 
%Then all atoms with $x_i'$ are of one of the types $L,U,M$.

Step 3: Let $\delta$ be the least common multiple of $\{k_m \mid m\in M\}$.
Construct $\theta'_{+\infty}(x_i')$ from $\theta'(x_i')$ by  
 replacing $\textit{true}$ for all atoms in $L$,
and $\textit{false}$ for all atoms in  $U$.
%We use $\theta'[t]$ to denote replacing all $x_i'$ by a term $t$. 
Then $\exists x_i'.\theta'(x_i')$
is equivalent to

$$\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee 
\bigvee_{j=0}^{\delta-1} \bigvee_{u\in U} \theta'[t_u(\bar{x},\bar{y})-j/x_i']$$

The first disjunction $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i']$ corresponds to the case where $x_i'$ is large enough 
so that all atoms in $L$ are \textit{true} and all atoms in  $U$ are \textit{false}. Thus $\theta'_{+\infty}$  contains only divisibility literals.
If there is a number $n (0\le n \le \delta-1)$ such that $ \theta'_{+\infty}[n/x_i']$ is evaluated to be \textit{true},
then for every $\lambda\in \mathbb{N}$,
$\theta'_{+\infty}[n+\lambda\delta/x_i']$ is also evaluated to be \textit{true}.
Hence, there exists $x_i'$ large enough to satisfy $\theta'$.

The second disjunction corresponds to the case that for 
some $u\in U, x_i'\le t_u(\bar{x},\bar{y})$ holds.
In this case, 
select the minimal $t_u(\bar{x},\bar{y})$
from atoms in $U$ that holds,
then there exists a solution for $x_i'$ such that $x_i'$ is in the interval $[t_u(\bar{x},\bar{y})-\delta+1,t_u(\bar{x},\bar{y})]$.

Here we describe Cooper's algorithm in pseudo code.
We will use $\theta\{\tau'/\tau\}$ to denote substitute all atoms $\tau$ with $\tau'$,
distinguished from term substitution $\theta[t'/t]$.

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{$x_i,\theta(x_i,\bar{x},\bar{y})$,
    where $x_i$ occurs linearly} 
    \KwOut{A equivalent formula without $x_i$}

    Collect terms of $x_i$ in each atom\;
    Let $d$ to be the least common multiple of all coefficients of $x_i$\;
    Multiply each atom by a factor so that coefficient of $x_i $ is $d$\;
    Let $\theta'(x_i'):= \theta[x_i'/dx_i]$\;
    $\theta'(x_i') := \theta'(x_i') \wedge d| x_i'$\;
    \tcp{ Now the atoms have 3 different forms}
    \tcp{ $t_l(\bar{x},\bar{y})\le x_i'$, $x_i' \le t_u(\bar{x},\bar{y})$, $k_m|x_i'+t_m(\bar{x},\bar{y})$(or its negation)}
    \tcp{ where $l\in L,u\in U,m\in M$ are the index sets}
    $\delta :=  lcm\{k_m|m\in M\}$\;
    \For{all atom $\tau_l \in L$}
    {
        $\theta'_{+\infty}:=  \theta'\{\textit{true}/\tau_l\}$
    }
    \For{all atom $\tau_u \in U$}
    {
        $\theta'_{+\infty}:=  \theta'_{+\infty}\{\textit{false}/\tau_u\}$
    }
    output $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee \bigvee_{j=0}^{\delta-1} \bigvee_{u\in U}\theta'[t_u(\bar{x},\bar{y})-j/x_i']$
    \caption{QE-linear (Cooper's QE algorithm for PA)}
\end{algorithm}

%\znj{\textbf{Solved, missing $\bigvee$ for $\theta_{+\infty}$}.Pls check whether the output is correct or not, my intuition is not correct.} 
% 之前的算法

\subsection{Compared with Francoise's Origin Algorithm}

Francoise's algorithm \cite{Francoise} has four steps,
corresponding to the four steps of our algorithm respectively.
So we will use the names of our steps to refer to them. 
The divisible predicate $n|x$ is not included in his theory,
instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
The \textbf{Normalization} and \textbf{QE-with-order} steps in two algorithms are similar,
with some minor differences due to the setting of the theory.
The main differences between the two algorithms lies in the following aspects.

In \textbf{QE-exp},
we adopt the same idea to find sufficient conditions
and necessary conditions for the formula to holds.
But our strategy for choosing parameters are easier to understand and use.
A flaw lies in the $a_i<0$ case in Francoise's algorithm 
\footnote{
In the $a_i<0$ case of \textbf{QE-exp-ineq},
Francoise's algorithm assumes falsely $t(y)\ge 0$.
The reason for this mistake might be the confusion of 
two subtraction operators.
The example mentioned in \ref{sec:3.2} shows this case.
},
our \textbf{QE-exp-ineq} corrects this flaw by 
treating $a_i<0$ case similarly to the $a_i>0$ case.

\textbf{QE-linear} in Francoise's algorithm in some sense includes our
\textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
In this step,
Francoise uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
We instead invoke the QE algorithm of PA without referring to permutation groups, 
other improved QE algorithm can also be used to simplify this step.

Francoise's algorithm has more restrictions on the form
of the formula.
Before \textbf{QE-linear} and \textbf{QE-exp} step,
it needs to transform the formula into disjunction normal form (DNF)
and deals with the basic case when input formula $\theta$ is 
a conjunction of atoms. 
It is well known that transforming an arbitrary formula into 
DNF is costly and the length of the formula may increase exponentially.
So our algorithm remove these limitations and assume no special forms of
the formula other than NNF.

\section{Complexity Analysis and Case Study}
In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
\subsection{Complexity Analysis}

In this part, we estimate the complexity of our algorithm.
It can be divided into two part, 
the first part convert a string constraint $\phi$ to a PA formula with exponential functions,
and the second part is the QE procedure on the obtained formula.

Given a finite alphabet $\Sigma$ and $\alpha=\langle p,q \rangle$, the size of $\Sigma$ is denoted by
$|\Sigma|$. 
The generic $\alpha$-flat language contains $O(|\Sigma|^{pq})$ $\alpha$-flat languages

\paragraph{Complexity for QE}
In \textbf{QE-with-order},
it reduces the origin problem to $(n+1)!$ sub-cases,
and for each case,
we recursively invoke \textbf{QE-exp} and \textbf{QE-linear}.
So the key is to analysis the inner \textit{for}
loop in \textbf{QE-with-order}.
Note that if we omit the \textbf{QE-exp} step,
the sub-cases are equivalent to Cooper's algorithm,
so we adopt the idea of Oppen's analysis to analyze the complexity
\cite{Oppen}.

Apply Cooper's algorithm on the quantified PA formula 
$Q_m x_m Q_{m-1}$ 
$x_{m-1} ... Q_1 x_1. $
$F(x_1,...,x_m)$,
the algorithm repeats $m$ times to eliminate
$x_1,...,x_m$ one by one.
Let $F_k \Def Q_m x_m Q_{m-1} 
x_{m-1} ... Q_{k+1} x_{k+1}. F_k'(x_{k+1},...,x_m)$
be the formula produced after $k$th iterations of the algorithm. 
Let $a_k$ denote the number of atoms in $F_k$,
$c_k$ denote the number of distinct $\delta_i$ in atom $\delta_i | t$ (and its negation) plus the number of distinct coefficients of variables.
$s_k$ denote the largest value of integer constant.
Use $a_0,c_0,s_0$ to denote the initial value of $a_k,c_k,s_k$ respectively.
The following theorem holds.

\begin{theorem}
for all $k: 1\le k\le m$, the following relations holds
\begin{itemize}
    \item $c_k \le {c_{k-1}}^4$
    \item $s_k \le {s_{k-1}}^{4c_{k-1}}$
    \item $a_k \le {a_{k-1}}^4 \cdot {s_{k-1}}^{2c_{k-1}}$
\end{itemize}
and from the above relations we have 
\begin{itemize}
    \item $c_k \le {{c_0}^4}^k$
    \item $s_k \le {{{s_0}^{4c_0}}^4}^k$
    \item $a_k \le {{a_0}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
\end{itemize}
\end{theorem}

For each $k$,
the space required to store $F_k$ 
is $a_k \cdot (m+1) \cdot s_k \cdot q$,
where $(m+1)$ is the maximum number of constants per atom and $q>1$ is some constant. Assume $m\le n, c\le n, a\le n, s\le n$, we obtain the deterministic space complexity is ${{2^2}^2}^{p n log_n}$, which is also the bound for deterministic time.

In our algorithm,
we use the same denotations and have the following theorem. 
Note that each iteration corresponds to \textbf{QE-exp} and \textbf{QE-linear}.
\begin{theorem}
for all $k: 1\le k\le m$, the following relations holds
\begin{itemize}
    \item $c_k \le {c_{k-1}}^4$
    \item $s_k \le {(n s_{k-1})}^{4c_{k-1}}$
    \item $a_k \le {{(l_2(ns_{k-1})\cdot a_{k-1})}}^4 \cdot {(n s_{k-1})}^{2c_{k-1}}$
\end{itemize}
and from the above relations we have 
\begin{itemize}
    \item $c_k \le {{c_0}^4}^k$
    \item $s_k \le {{{n}^{4c_0}}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
    \item $a_k \le {{a_0}^4}^k \cdot 
     {{{n}^{4c_0}}^4}^k
     \cdot {{{s_0}^{4c_0}}^4}^k$
\end{itemize}
\end{theorem}

WAIT TO CHECK!

Again assume $m,c,a,s\le n$, which gives 
$$
\textit{space}\le q \cdot {n^4}^n \cdot {{n^{(4n)}}^4}^n \cdot
{{n^{(4n)}}^4}^n \cdot (n+1) \cdot 
\cdot {{n^{(4n)}}^4}^n \cdot {{n^{(4n)}}^4}^n
\le 2^{2^{2^{p n \textit{log}n}}}
$$

Then the upper bound for Cooper's algorithm still holds for the inner \textit{for} loop of \textbf{QE-with-order}.
When considering \textbf{Normalization} and permutations groups in \textbf{QE-with-order},
since the super exponential term dominates the number of permutations $n^{O(n)}$,
this upper bound still holds.

\subsection{Example}

We apply our algorithm to a simple example to 
illustrate its function.
Given $\Sigma = \{0,1\}$
Suppose $\mathcal{A}$ is a finite state automaton
that recognizes language $L = \{ w \mid \#w('1') \text{ is odd}\}$.
Consider the following string constraint
$$y\in L(\mathcal{A})\wedge 4\le  \parseInt(y) < 16\,.$$

We need an abstraction parameter $alpha$ to restrict interpretations
for $x$, for simplicity, set $\alpha\Def \langle 2,1 \rangle$.
The generic $\langle 2,1 \rangle$-flat language is the union of 
$\langle 2,1 \rangle$-flat languages $(11)^*,(10)^*,(01)^*,(00)^*,1^*,0^*$.
Here we divide the problem into 6 sub-cases by restricting $y$
to one of the $\langle 2,1 \rangle$-flat language.
Another approach is to directly restrict $y$ to 
the generic $\langle 2,1 \rangle$-flat language,
at the cost of more introduced variables.
Since the complexity is related to the number of variables,
we wish it to be as little as possible.

When $y$ is in language $1^*$, assume $y = 1^x$ where $x$
is a fresh integer variable.
Note that the Parikh image of $L(\mathcal{A})$ is $2|( x+1)$, 
so the constraint is converted to a PA formula $\phi$
$$\phi\Def \exists x. 2|(x+1) \wedge 5 \le 2^{x} < 17\,.$$

Observe that when $x=3$ the formula holds,
so we can easily deduce that $\phi$ is true.  
By applying our algorithm,
after \textbf{Normalization},
$\phi$ is transformed into the form

$$\exists x_0. 2|(x+1)\wedge  2^{x}\le 16 \wedge -2^{x}\le -5\,.$$

Since there is only one variable, 
\textbf{QE-with-order} can be omitted.
We send $2^x\le 16$ and $-2^x \le -5$ to \textbf{QE-exp} and obtain
$$2^x\le 16 \equiv x \le 4$$
$$-2^x\le -5 \equiv 3\le x\,.$$ 
Then we apply \textbf{QE-linear} on formula $\exists x_0. 2|(x+1) \wedge x\le 4 \wedge 3\le x$ and get
$$\bigvee_{j=0}^1(2| j+1 \wedge \textit{false} \wedge \textit{true})\vee \bigvee_{j=0}^1 (2|4-j+i\wedge 4-j\le 4 \wedge 3\le 4-j) \,.$$
which evaluates \textit{true} and we can deduce $j=1\implies x=3$, so $y$ is string $111$.

Other sub-cases are similar. However, when the abstraction parameter becomes larger,
say, $\alpha =\langle 2,2 \rangle$,
the complexity will grow rapidly,
because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.

\section{Conclusion and Future Works}

%test bib
\bibliographystyle{plain}
\bibliography{reference}

\end{document}

### 2021-05-06 FMCAD2021/paper.tex
rename

@@
+% !TEX program = xelatex
 \documentclass[runningheads]{llncs}
@@
 \RequirePackage{todonotes}
-%\newcommand{\znj}[1]{\todo[inline,color=orange!10]{\textbf{ZHAN:} #1}}
 \newcommand{\znjSide}[1]{\todo[color=orange!10]{\textbf{ZHAN:} #1}}
 \newcommand\znj[1]{\textcolor{red}{#1}}
-
 \newcommand{\parseInt}{\textsf{parseInt}}
 \newcommand{\paexp}{$\textsf{PA}_{\exp}$}
-%\newcommand{\paexp}{$(\mathbb{N},+,2^x)$}
 \newcommand{\Def}{\hat{=}}
@@
 \begin{abstract}
-    The problem of solving string constraints is currently one of the focuses of theoretical computer science, particular related to SMT-based program verification, as it is inevitably that most of 
-    programs manipulate string data structures. 
-    The general problem of string constraint solving 
-    is undecidable, therefore people try to identify 
-    some sub-problems with decidability, e.g., string constraints restricted to generic flat languages. 
-    Certainly, it is challenging and 
-    makes theoretical and practical significance to 
-    identify such subset with as much expressive power as possible. 
-    In this paper, we identify 
-    the subset of flat string constraints with 
-     the string-number conversion function $\mathbf{parseInt}$, and 
-     provide a decision procedure to the subset by reduction to
-   the decidability problem of Presburger Arithmetic with exponential functions, 
-    which is decidable. To the best of our knowledge, such a subset is 
-    the largest subset of string constraints with decidability so far.
-
+    String constraint solving is the core of various testing and verification approaches for scripting languages. 
+    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
+    In the PLDI 2020 paper of Abdulla et al., the authors extended the flattening approach to support the string-integer conversion function, which appears in almost all scripting languages.
+    However, their approach supports only a specific flattening pattern and leaves the support of the general class as an open problem.
+    In this paper, we bridge the gap and propose a complete flattening approach for the string-integer conversion, based upon a decision procedure for solving the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions).
+    Besides the theoretical results, we also develop various optimizations and implement the decision procedure.
+    Moreover, we evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as generated randomly.
+    The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer constraints as well as linear-exponential constraints, in both precision and efficiency.
 \keywords{ String Constraints \and Decidability \and Exponential Functions \and Quantifier Elimination.}

### 2021-05-07 FMCAD2021/paper.tex
abstract and references

@@
 
-\title{A Decision Procedure for String Constraints with {\parseInt} Function and Flat Regular Constraints}
+\title{A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints}
 \author{ }
@@
     Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
-    In the PLDI 2020 paper of Abdulla et al., the authors extended the flattening approach to support the string-integer conversion function, which appears in almost all scripting languages.
-    However, their approach supports only a specific flattening pattern and leaves the support of the general class as an open problem.
-    In this paper, we bridge the gap and propose a complete flattening approach for the string-integer conversion, based upon a decision procedure for solving the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions).
-    Besides the theoretical results, we also develop various optimizations and implement the decision procedure.
-    Moreover, we evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as generated randomly.
-    The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer constraints as well as linear-exponential constraints, in both precision and efficiency.
+    In the PLDI 2020 paper of Abdulla et al., the authors extended the flattening approach to support the string-integer conversion, which is an important function appearing in almost all scripting languages.
+    However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
+    In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion.
+The approach is built upon a decision procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions).
+    Besides the theoretical results, we also develop various optimizations and provide an implementation.
+    We evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as randomly.
+    The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer conversion constraints as well as linear-exponential constraints, in both precision and efficiency.
 \keywords{ String Constraints \and Decidability \and Exponential Functions \and Quantifier Elimination.}
@@
 Semenov proved that for any function $f$ that is compatible with addition, 
-theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov}. 
+theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
 Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
-by \cite{semenov}.
-Particularly, in \cite{Francoise} Francoise gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
-Unfortunately, Francoise's algorithm is flawed, one problem 
+by \cite{Semenov84}.
+Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
+Unfortunately, Point's algorithm is flawed, one problem 
  lies in the discussion to eliminate exponential terms. 
@@
 
-In the next section, within Francoise's framework, we give a revised algorithm
+In the next section, within Point's framework, we give a revised algorithm
  to the decision problem of {\paexp} with 
@@
 \section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
-Based on Francoise's work \cite{Francoise}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
+Based on Point's work \cite{Point86}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
 where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 

### 2021-05-07 FMCAD2021/paper.tex
abstract

@@
     In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion.
-The approach is built upon a decision procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions).
-    Besides the theoretical results, we also develop various optimizations and provide an implementation.
+The approach is built upon a decision procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
+While the decision procedure by Point relies on expensive quantifier elimination, we introduce various optimizations and provide an efficient implementation.
+%    Besides the theoretical results, we also develop various optimizations and provide an implementation.
     We evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as randomly.

### 2021-05-07 Benchmark/NIA-unbound/2-3-3-4/runcvc4.sh
stash

@@ -1,5 +1,5 @@
 #!/bin/bash
-echo "begin test using Z3 ..."
+echo "begin test using CVC4 ..."
 
 count=1
 while(( $count<= 100 ))

### 2021-05-07 Benchmark/NIA-unbound/3-4-3-4-eq/runcvc4.sh

@@ -1,17 +1,0 @@
-#!/bin/bash
-echo "begin test using CVC4 ..."
-
-count=1
-while(( $count<= 100 ))
-do
-    START=$(date +%s)
-    # do something
-    # start your script work here
-    echo $count
-    cvc4 test$count -L smt --tlimit=60000 --fmf-fun --produce-models
-    let "count++"
-    # your logic ends here
-    END=$(date +%s)
-    DIFF=$(( $END - $START ))
-    echo "It took $DIFF seconds"
-done

### 2021-05-07 Benchmark/produce_nia_test.py

@@ -1,15 +1,15 @@
 from z3 import *
 import random
 
-expnum = 3
-linearnum = 4
+expnum = 2
+linearnum = 3
 expfml = 3
 linearfml = 4
 totaltests = 100
 
 x=[0]*(expnum+linearnum+2)
 
-g = open("NIA-bound/3-4-3-4-eq/paras","w+")
+g = open("NIA-unbound/2-3-3-4-eq/paras","w+")
 g.write("%d\n" %totaltests)
 g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
 
@@ -24,21 +24,19 @@ for testno in range(1,totaltests+1):
         s.add(x[i] >= 0)
 
     # add an equality
-    g.write("{")
-    expr = 0
-    for j in range (1,expnum+1):
-        coef = random.randint(-10,10)
-        expr = expr + coef * pow(x[j])
-        g.write("%d," %coef)
-    for j in range (1,expnum+linearnum+1):
-        coef = random.randint(-10**2,10**2)
-        expr = expr + coef * x[j]
-        g.write("%d," %coef)
-
-    num = random.randint(-10**5,10**5)
-    s.add(expr == num)
-    g.write("%d}\n" % num)
-
+    #g.write("{")
+    #expr = 0
+    #for j in range (1,expnum+1):
+    #    coef = random.randint(-10,10)
+    #    expr = expr + coef * pow(x[j])
+    #    g.write("%d," %coef)
+    #for j in range (1,expnum+linearnum+1):
+    #    coef = random.randint(-10**2,10**2)
+    #    expr = expr + coef * x[j]
+    #    g.write("%d," %coef)
+    #num = random.randint(-10**5,10**5)
+    #s.add(expr == num)
+    #g.write("%d}\n" % num)
 
     # with eq 2
     for i in range (2,expfml+1):
@@ -46,21 +44,21 @@ for testno in range(1,totaltests+1):
         expr = 0
         for j in range (1,expnum+1):
             coef = random.randint(-10,10)
-            #if coef == 0:
-            #    coef =1
+            if coef == 0:
+                coef = 1
             expr = expr + coef * pow(x[j])
             g.write("%d," %coef)
         # bound cases
-        for j in range (1,expnum+linearnum+1):
+        #for j in range (1,expnum+linearnum+1):
         # unbound cases
-        #for j in range (1,expnum+1):
-            coef = random.randint(-10**2,10**2)
+        for j in range (1,expnum+1):
[…]

### 2021-05-07 Benchmark/runall.sh

@@ -1,12 +1,11 @@
 #!/bin/bash
 echo "begin test..."
 
-cd NIA-bound
-    cd 2-3-2-3-eq
-    ./runz3.sh > z3-result 
+cd NIA-unbound
+    cd 3-4-4-5
     ./runcvc4.sh > cvc4-result
     cd ..
-    cd 3-4-3-4-eq
+    cd 4-5-5-6
     ./runz3.sh > z3-result 
     ./runcvc4.sh > cvc4-result
     cd ..

### 2021-05-07 FMCAD2021/paper.tex
12

@@
     The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer conversion constraints as well as linear-exponential constraints, in both precision and efficiency.
-\keywords{ String Constraints \and Decidability \and Exponential Functions \and Quantifier Elimination.}
+\keywords{ String-integer conversion \and Flat regular constraints \and Presburger arithmetic \and Exponential function \and Quantifier elimination.}
 
 \end{abstract}
-
-\iffalse     
-    A Decision Procedure for String Constraints with the {\parseInt} Function and Flat Regular Constraints
-
-    {\parseInt} converts strings to integers and is an important function in string manipulating programs. 
-    Solving string constraints with {\parseInt} is challenging and undecidable in general. 
-    The state-of-the-art string solvers resort to heuristics to solve such constraints. 
-    In this work, we propose a decision procedure for a class of string constraints which 
-    are boolean combinations of word equations, flat regular constraints, and {\parseInt} function. 
-    The decision procedure is obtained by reducing the satisfiability problem to that of
-    an extension of Presburger arithmetic with exponential functions, which is decidable. 
-    We implement the decision procedure and carry out experiments to evaluate the performance. 
-    The experimental results show the improvement against the state-of-the-art string solvers 
-    in both precision and performance. To the best of our knowledge, this work represents the 
-    first decision procedure for a class of string constraints with {\parseInt} function.
-\fi
 

### 2021-05-07 FMCAD2021/paper.tex
change to IEEE Trans template

--- 2021-05-07 FMCAD2021/paper.tex
@@
 % !TEX program = xelatex
-\documentclass[runningheads]{llncs}
+%
+\documentclass[10pt,letterpaper]{IEEEtran}
+
+% *** CITATION PACKAGES ***
+%
+\ifCLASSOPTIONcompsoc
+  % The IEEE Computer Society needs nocompress option
+  % requires cite.sty v4.0 or later (November 2003)
+  \usepackage[nocompress]{cite}
+\else
+  % normal IEEE
+  \usepackage{cite}
+\fi
+% cite.sty was written by Donald Arseneau
+% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
+% \cite{} output to follow that of the IEEE. Loading the cite package will
+% result in citation numbers being automatically sorted and properly
+% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
+% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
+% \cite will automatically add leading space, if needed. Use cite.sty's
+% noadjust option (cite.sty V3.8 and later) if you want to turn this off
+% such as if a citation ever needs to be enclosed in parenthesis.
+% cite.sty is already installed on most LaTeX systems. Be sure and use
+% version 5.0 (2009-03-20) and later if using hyperref.sty.
+% The latest version can be obtained at:
+% http://www.ctan.org/pkg/cite
+% The documentation is contained in the cite.sty file itself.
+%
+% Note that some packages require special options to format as the Computer
+% Society requires. In particular, Computer Society  papers do not use
+% compressed citation ranges as is done in typical IEEE papers
+% (e.g., [1]-[4]). Instead, they list every citation separately in order
+% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
+% package with the nocompress option which is supported by cite.sty v4.0
+% and later.
+
+% *** GRAPHICS RELATED PACKAGES ***
+%
+\ifCLASSINFOpdf
+  % \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../pdf/}{../jpeg/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  % \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
+% graphicx was written by David Carlisle and Sebastian Rahtz. It is
+% required if you want graphics, photos, etc. graphicx.sty is already
+% installed on most LaTeX systems. The latest version and documentation
+% can be obtained at: 
+% http://www.ctan.org/pkg/graphicx
+% Another good source of documentation is "Using Imported Graphics in
+% LaTeX2e" by Keith Reckdahl which can be found at:
+% http://www.ctan.org/pkg/epslatex
+%
+% latex, and pdflatex in dvi mode, support graphics in encapsulated
+% postscript (.eps) format. pdflatex in pdf mode supports graphics
+% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
+% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
+% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
+% which can result in "jaggedy"/blurry rendering of lines and letters as
+% well as large increases in file sizes.
+%
+% You can find documentation about the pdfTeX application at:
+% http://www.tug.org/applications/pdftex
+
+% *** MATH PACKAGES ***
+%
+%\usepackage{amsmath}
+% A popular package from the American Mathematical Society that provides
+% many useful and powerful commands for dealing with mathematics.
+%
+% Note that the amsmath package sets \interdisplaylinepenalty to 10000
+% thus preventing page breaks from occurring within multiline equations. Use:
+%\interdisplaylinepenalty=2500
+% after loading amsmath to restore such page breaks as IEEEtran.cls normally
+% does. amsmath.sty is already installed on most LaTeX systems. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/pkg/amsmath
+
+% *** SPECIALIZED LIST PACKAGES ***
+%\usepackage{acronym}
+% acronym.sty was written by Tobias Oetiker. This package provides tools for
+% managing documents with large numbers of acronyms. (You don't *have* to
+% use this package - unless you have a lot of acronyms, you may feel that
+% such package management of them is bit of an overkill.)
+% Do note that the acronym environment (which lists acronyms) will have a
+% problem when used under IEEEtran.cls because acronym.sty relies on the
+% description list environment - which IEEEtran.cls has customized for
+% producing IEEE style lists. A workaround is to declared the longest
+% label width via the IEEEtran.cls \IEEEiedlistdecl global control:
+%
+% \renewcommand{\IEEEiedlistdecl}{\IEEEsetlabelwidth{SONET}}
+% \begin{acronym}
+%
+% \end{acronym}
+% \renewcommand{\IEEEiedlistdecl}{\relax}% remember to reset \IEEEiedlistdecl
+%
+% instead of using the acronym environment's optional argument.
+% The latest version and documentation can be obtained at:
+% http://www.ctan.org/pkg/acronym
+
+%\usepackage{algorithmic}
+% algorithmic.sty was written by Peter Williams and Rogerio Brito.
+% This package provides an algorithmic environment fo describing algorithms.
+% You can use the algorithmic environment in-text or within a figure
+% environment to provide for a floating algorithm. Do NOT use the algorithm
+% floating environment provided by algorithm.sty (by the same authors) or
+% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
+% algorithm float types and packages that provide these will not provide
+% correct IEEE style captions. The latest version and documentation of
+% algorithmic.sty can be obtained at:
+% http://www.ctan.org/pkg/algorithms
+% Also of interest may be the (relatively newer and more customizable)
+% algorithmicx.sty package by Szasz Janos:
+% http://www.ctan.org/pkg/algorithmicx
+
+% *** ALIGNMENT PACKAGES ***
+%
+%\usepackage{array}
+% Frank Mittelbach's and David Carlisle's array.sty patches and improves
+% the standard LaTeX2e array and tabular environments to provide better
+% appearance and additional user controls. As the default LaTeX2e table
+% generation code is lacking to the point of almost being broken with
+% respect to the quality of the end results, all users are strongly
+% advised to use an enhanced (at the very least that provided by array.sty)
+% set of table tools. array.sty is already installed on most systems. The
+% latest version and documentation can be obtained at:
+% http://www.ctan.org/pkg/array
+
+%\usepackage{mdwmath}
+%\usepackage{mdwtab}
+% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
+% especially mdwmath.sty and mdwtab.sty which are used to format equations
+% and tables, respectively. The MDWtools set is already installed on most
+% LaTeX systems. The lastest version and documentation is available at:
+% http://www.ctan.org/pkg/mdwtools
+
+% IEEEtran contains the IEEEeqnarray family of commands that can be used to
+% generate multiline equations as well as matrices, tables, etc., of high
+% quality.
+
+%\usepackage{eqparbox}
+% Also of notable interest is Scott Pakin's eqparbox package for creating
+% (automatically sized) equal width boxes - aka "natural width parboxes".
+% Available at:
+% http://www.ctan.org/pkg/eqparbox
+
+% *** SUBFIGURE PACKAGES ***
+%\ifCLASSOPTIONcompsoc
+%  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
+%\else
+%  \usepackage[caption=false,font=footnotesize]{subfig}
+%\fi
+% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
+% for subfigure.sty, the latter of which is no longer maintained and is
+% incompatible with some LaTeX packages including fixltx2e. However,
+% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
+% which will override IEEEtran.cls' handling of captions and this will result
+% in non-IEEE style figure/table captions. To prevent this problem, be sure
+% and invoke subfig.sty's "caption=false" package option (available since
+% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
+% handling of captions.
+% Note that the Computer Society format requires a sans serif font rather
+% than the serif font used in traditional IEEE formatting and thus the need
+% to invoke different subfig.sty package options depending on whether
+% compsoc mode has been enabled.
+%
+% The latest version and documentation of subfig.sty can be obtained at:
+% http://www.ctan.org/pkg/subfig
+
+% *** FLOAT PACKAGES ***
+%
+%\usepackage{fixltx2e}
+% fixltx2e, the successor to the earlier fix2col.sty, was written by
+% Frank Mittelbach and David Carlisle. This package corrects a few problems
+% in the LaTeX2e kernel, the most notable of which is that in current
+% LaTeX2e releases, the ordering of single and double column floats is not
+% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
+% single column figure to be placed prior to an earlier double column
+% figure.
+% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
+% corrections already built into the system in which case a warning will
+% be issued if an attempt is made to load fixltx2e.sty as it is no longer
+% needed.
+% The latest version and documentation can be found at:
+% http://www.ctan.org/pkg/fixltx2e
+
+%\usepackage{stfloats}
+% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
+% the ability to do double column floats at the bottom of the page as well
+% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
+% LaTeX2e). It also provides a command:
+%\fnbelowfloat
+% to enable the placement of footnotes below bottom floats (the standard
+% LaTeX2e kernel puts them above bottom floats). This is an invasive package
+% which rewrites many portions of the LaTeX2e float routines. It may not work
+% with other packages that modify the LaTeX2e float routines. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/pkg/stfloats
+% Do not use the stfloats baselinefloat ability as the IEEE does not allow
+% \baselineskip to stretch. Authors submitting work to the IEEE should note
+% that the IEEE rarely uses double column equations and that authors should try
+% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
+% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
+% such ways.
+% Do not attempt to use stfloats with fixltx2e as they are incompatible.
+% Instead, use Morten Hogholm'a dblfloatfix which combines the features
+% of both fixltx2e and stfloats:
+%
+% \usepackage{dblfloatfix}
+% The latest version can be found at:
+% http://www.ctan.org/pkg/dblfloatfix
+
+%\ifCLASSOPTIONcaptionsoff
+%  \usepackage[nomarkers]{endfloat}
+% \let\MYoriglatexcaption\caption
+% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
+%\fi
+% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
+% Axel Sommerfeldt. This package may be useful when used in conjunction with 
+% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
+% submissions have lists of figures/tables at the end of the paper and that
+% figures/tables without any captions are placed on a page by themselves at
+% the end of the document. If needed, the draftcls IEEEtran class option or
+% \CLASSINPUTbaselinestretch interface can be used to increase the line
+% spacing as well. Be sure and use the nomarkers option of endfloat to
+% prevent endfloat from "marking" where the figures would have been placed
+% in the text. The two hack lines of code above are a slight modification of
+% that suggested by in the endfloat docs (section 8.4.1) to ensure that
+% the full captions always appear in the list of figures/tables - even if
+% the user used the short optional argument of \caption[]{}.
+% IEEE papers do not typically make use of \caption[]'s optional argument,
+% so this should not be an issue. A similar trick can be used to disable
+% captions of packages such as subfig.sty that lack options to turn off
+% the subcaptions:
+% For subfig.sty:
+% \let\MYorigsubfloat\subfloat
+% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
+% However, the above trick will not work if both optional arguments of
+% the \subfloat command are used. Furthermore, there needs to be a
+% description of each subfigure *somewhere* and endfloat does not add
+% subfigure captions to its list of figures. Thus, the best approach is to
+% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
+% and instead reference/explain all the subfigures within the main caption.
+% The latest version of endfloat.sty and its documentation can obtained at:
+% http://www.ctan.org/pkg/endfloat
+%
+% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
+% later in the document, say, to conditionally put the References on a 
+% page by themselves.
+
+% *** PDF, URL AND HYPERLINK PACKAGES ***
+%
+%\usepackage{url}
+% url.sty was written by Donald Arseneau. It provides better support for
+% handling and breaking URLs. url.sty is already installed on most LaTeX
+% systems. The latest version and documentation can be obtained at:
+% http://www.ctan.org/pkg/url
+% Basically, \url{my_url_here}.
+
+% NOTE: PDF thumbnail features are not required in IEEE papers
+%       and their use requires extra complexity and work.
+%\ifCLASSINFOpdf
+%  \usepackage[pdftex]{thumbpdf}
+%\else
+%  \usepackage[dvips]{thumbpdf}
+%\fi
+% thumbpdf.sty and its companion Perl utility were written by Heiko Oberdiek.
+% It allows the user a way to produce PDF documents that contain fancy
+% thumbnail images of each of the pages (which tools like acrobat reader can
+% utilize). This is possible even when using dvi->ps->pdf workflow if the
+% correct thumbpdf driver options are used. thumbpdf.sty incorporates the
+% file containing the PDF thumbnail information (filename.tpm is used with
+% dvips, filename.tpt is used with pdftex, where filename is the base name of
+% your tex document) into the final ps or pdf output document. An external
+% utility, the thumbpdf *Perl script* is needed to make these .tpm or .tpt
+% thumbnail files from a .ps or .pdf version of the document (which obviously
+% does not yet contain pdf thumbnails). Thus, one does a:
+% 
+% thumbpdf filename.pdf 
+%
+% to make a filename.tpt, and:
+%
+% thumbpdf --mode dvips filename.ps
+%
+% to make a filename.tpm which will then be loaded into the document by
+% thumbpdf.sty the NEXT time the document is compiled (by pdflatex or
+% latex->dvips->ps2pdf). Users must be careful to regenerate the .tpt and/or
+% .tpm files if the main document changes and then to recompile the
+% document to incorporate the revised thumbnails to ensure that thumbnails
+% match the actual pages. It is easy to forget to do this!
+% 
+% Unix systems come with a Perl interpreter. However, MS Windows users
+% will usually have to install a Perl interpreter so that the thumbpdf
+% script can be run. The Ghostscript PS/PDF interpreter is also required.
+% See the thumbpdf docs for details. The latest version and documentation
+% can be obtained at.
+% http://www.ctan.org/pkg/thumbpdf
+
+% NOTE: PDF hyperlink and bookmark features are not required in IEEE
+%       papers and their use requires extra complexity and work.
+% *** IF USING HYPERREF BE SURE AND CHANGE THE EXAMPLE PDF ***
+% *** TITLE/SUBJECT/AUTHOR/KEYWORDS INFO BELOW!!           ***
+\newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
+pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
+colorlinks=true,linkcolor={black},citecolor={black},urlcolor={black},
+pdftitle={A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints},%<!CHANGE!
+pdfsubject={String constraint solving},%<!CHANGE!
+pdfauthor={Hao Wu, Yu-Fang Chen, Zhilin Wu, Naijun Zhan},%<!CHANGE!
+pdfkeywords={String-integer conversion, Flat regular constraints, Presburger arithmetic, Exponential function, Quantifier elimination.}}%<^!CHANGE!
+%\ifCLASSINFOpdf
+%\usepackage[\MYhyperrefoptions,pdftex]{hyperref}
+%\else
+%\usepackage[\MYhyperrefoptions,breaklinks=true,dvips]{hyperref}
+%\usepackage{breakurl}
+%\fi
+% One significant drawback of using hyperref under DVI output is that the
+% LaTeX compiler cannot break URLs across lines or pages as can be done
+% under pdfLaTeX's PDF output via the hyperref pdftex driver. This is
+% probably the single most important capability distinction between the
+% DVI and PDF output. Perhaps surprisingly, all the other PDF features
+% (PDF bookmarks, thumbnails, etc.) can be preserved in
+% .tex->.dvi->.ps->.pdf workflow if the respective packages/scripts are
+% loaded/invoked with the correct driver options (dvips, etc.). 
+% As most IEEE papers use URLs sparingly (mainly in the references), this
+% may not be as big an issue as with other publications.
+%
+% That said, Vilar Camara Neto created his breakurl.sty package which
+% permits hyperref to easily break URLs even in dvi mode.
+% Note that breakurl, unlike most other packages, must be loaded
+% AFTER hyperref. The latest version of breakurl and its documentation can
+% be obtained at:
+% http://www.ctan.org/pkg/breakurl
+% breakurl.sty is not for use under pdflatex pdf mode.
+%
+% The advanced features offer by hyperref.sty are not required for IEEE
+% submission, so users should weigh these features against the added
+% complexity of use.
+% The package options above demonstrate how to enable PDF bookmarks
+% (a type of table of contents viewable in Acrobat Reader) as well as
+% PDF document information (title, subject, author and keywords) that is
+% viewable in Acrobat reader's Document_Properties menu. PDF document
+% information is also used extensively to automate the cataloging of PDF
+% documents. The above set of options ensures that hyperlinks will not be
+% colored in the text and thus will not be visible in the printed page,
+% but will be active on "mouse over". USING COLORS OR OTHER HIGHLIGHTING
+% OF HYPERLINKS CAN RESULT IN DOCUMENT REJECTION BY THE IEEE, especially if
+% these appear on the "printed" page. IF IN DOUBT, ASK THE RELEVANT
+% SUBMISSION EDITOR. You may need to add the option hypertexnames=false if
+% you used duplicate equation numbers, etc., but this should not be needed
+% in normal IEEE work.
+% The latest version of hyperref and its documentation can be obtained at:
+% http://www.ctan.org/pkg/hyperref
+
+% *** Do not adjust lengths that control margins, column widths, etc. ***
+% *** Do not use packages that alter fonts (such as pslatex).         ***
+% There should be no need to do such things with IEEEtran.cls V1.6 and later.
+% (Unless specifically asked to do so by the journal or conference you plan
+% to submit to, of course. )
+
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
 
@@
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{corollary}[theorem]{Corollary}
+%\newtheorem{resume}{Resume}
+%\newtheorem{claim}{Claim}
+
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{remark}[theorem]{Remark}
+%\newtheorem{openq}{Open Question}
+% \newcommand{\sextract}{\mathcal{S}}
+% \newcommand{\ainput}{Input}
+
+\begin{document}
+%
+% paper title
+% Titles are generally capitalized except for words such as a, an, and, as,
+% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
+% not capitalized unless they are the first or last word of the title.
+% Linebreaks \\ can be used within to get better formatting as desired.
+% Do not put math or special symbols in the title.
 \title{A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints}
-\author{ }
-\institute{ } 
-\begin{document}
+
+\author{Hao Wu,
+        Yu-Fang Chen,
+        Zhilin Wu, 
+        Naijun Zhan% <-this % stops a space
+\IEEEcompsocitemizethanks{
+\IEEEcompsocthanksitem Academica Sinica.\protect\\
+% note need leading \protect in front of \\ to get a newline within \thanks as
+% \\ is fragile and will error, could use \hfil\break instead.
+E-mail: see http://www.michaelshell.org/contact.html
+\IEEEcompsocthanksitem xxx.}% <-this % stops a space
+\thanks{xxxx.}
+}
+
+\author{\IEEEauthorblockN{Hao Wu\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Yu-Fang Chen\IEEEauthorrefmark{2}, 
+Zhilin Wu\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Naijun Zhan\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}
+\\
+\IEEEauthorblockA{
+\IEEEauthorrefmark{1} State Key Laboratory of Computer Science,\\
+Institute of Software, Chinese Academy of Sciences, 100190, Beijing, China\\
+Email: \{wuhao, wuzl, znj\}@ios.ac.cn}\\
+\IEEEauthorblockA{\IEEEauthorrefmark{2}
+Institute of Information Science, Academia Sinica, Taiwan\\
+Email: yfc@iis.sinica.edu.tw}\\
+\IEEEauthorblockA{\IEEEauthorrefmark{3}University of Chinese Academy of Sciences,
+Beijing, China}
+}
+
+% The paper headers
+\markboth{FMCAD 2021}%
+{Wu \MakeLowercase{\textit{et al.}}: A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints}
+% The only time the second header will appear is for the odd numbered pages
+% after the title page when using the twoside option.
+% 
+% *** Note that you probably will NOT want to include the author's ***
+% *** name in the headers of peer review papers.                   ***
+% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
+% you desire.
+
+%\begin{document}
 \maketitle{}
@@
 The approach is built upon a decision procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-While the decision procedure by Point relies on expensive quantifier elimination, we introduce various optimizations and provide an efficient implementation.
+While the decision procedure by Point relies on quantifier eliminations and is expensive, we introduce various optimizations and provide an efficient implementation.
 %    Besides the theoretical results, we also develop various optimizations and provide an implementation.
@@
     The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer conversion constraints as well as linear-exponential constraints, in both precision and efficiency.
-\keywords{ String Constraints \and Decidability \and Exponential Functions \and Quantifier Elimination.}
+%\keywords{ String-integer conversion \and Flat regular constraints \and Presburger arithmetic \and Exponential function \and Quantifier elimination.}
 

### 2021-05-07 FMCAD2021/paper.tex

@@
 
-\iffalse     
-    A Decision Procedure for String Constraints with the {\parseInt} Function and Flat Regular Constraints
-
-    {\parseInt} converts strings to integers and is an important function in string manipulating programs. 
-    Solving string constraints with {\parseInt} is challenging and undecidable in general. 
-    The state-of-the-art string solvers resort to heuristics to solve such constraints. 
-    In this work, we propose a decision procedure for a class of string constraints which 
-    are boolean combinations of word equations, flat regular constraints, and {\parseInt} function. 
-    The decision procedure is obtained by reducing the satisfiability problem to that of
-    an extension of Presburger arithmetic with exponential functions, which is decidable. 
-    We implement the decision procedure and carry out experiments to evaluate the performance. 
-    The experimental results show the improvement against the state-of-the-art string solvers 
-    in both precision and performance. To the best of our knowledge, this work represents the 
-    first decision procedure for a class of string constraints with {\parseInt} function.
-\fi
-
 \section{Introduction}

### 2021-05-07 Benchmark/HashFunction/produce-head.py
generate string tests

@@ -1,95 +1,0 @@
-from z3 import *
-import random
-
-expnum = 2
-linearnum = 3
-expfml = 3
-linearfml = 4
-totaltests = 100
-
-x=[0]*(expnum+linearnum+2)
-
-g = open("NIA-unbound/2-3-3-4-eq/paras","w+")
-g.write("%d\n" %totaltests)
-g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
-
-for testno in range(1,totaltests+1):
-    for i in range(1,expnum+linearnum+1):
-        x[i] = Int('x'+str(i))
-    pow = Function('pow',IntSort(),IntSort())
-
-    s = Solver()
-
-    for i in range (1,expnum+linearnum+1):
-        s.add(x[i] >= 0)
-
-    # add an equality
-    #g.write("{")
-    #expr = 0
-    #for j in range (1,expnum+1):
-    #    coef = random.randint(-10,10)
-    #    expr = expr + coef * pow(x[j])
-    #    g.write("%d," %coef)
-    #for j in range (1,expnum+linearnum+1):
-    #    coef = random.randint(-10**2,10**2)
-    #    expr = expr + coef * x[j]
-    #    g.write("%d," %coef)
-    #num = random.randint(-10**5,10**5)
-    #s.add(expr == num)
-    #g.write("%d}\n" % num)
-
-    # with eq 2
-    for i in range (2,expfml+1):
-        g.write("{")
-        expr = 0
-        for j in range (1,expnum+1):
-            coef = random.randint(-10,10)
-            if coef == 0:
-                coef = 1
-            expr = expr + coef * pow(x[j])
-            g.write("%d," %coef)
-        # bound cases
-        #for j in range (1,expnum+linearnum+1):
-        # unbound cases
-        for j in range (1,expnum+1):
-            coef = random.randint(-10**5,10**5)
-            expr = expr + coef * x[j]
-            g.write("%d," %coef)
-        # unbound cases
-        for j in range (expnum+1,expnum+linearnum+1):
-            coef = 0
-            g.write("%d," %coef)
-
-        num = random.randint(-10**5,10**5)
-        s.add(expr <= num)
-        g.write("%d}\n" % num)
-    
-    for i in range (1,linearfml+1):
-        expr = 0
-        g.write("{")
-        for j in range (1,expnum+linearnum+1):
[…]

### 2021-05-07 Benchmark/HashFunction/runstring.sh

@@ -3,7 +3,7 @@ echo "--------------------------------"
 
 echo "begin string test Using Z3 ..."
 START=$(date +%s)
-z3 -smt2 string -T:120
+z3 -smt2 ./all/head/test1 -T:60
 END=$(date +%s)
 DIFF=$(( $END - $START ))
 echo "Z3 took $DIFF seconds"
@@ -12,7 +12,7 @@ echo "--------------------------------"
 
 echo "begin string test Using CVC4 ..."
 START=$(date +%s)
-cvc4 string -L smt --tlimit=120000 --strings-exp   
+cvc4 ./all/head/test1 -L smt --tlimit=60000 --strings-exp   
 END=$(date +%s)
 DIFF=$(( $END - $START ))
 echo "CVC4 took $DIFF seconds"

### 2021-05-07 Benchmark/NIA-test/produce-bound.py

@@ -1,0 +1,84 @@
+from z3 import *
+import random
+
+expnum = 3
+linearnum = 4
+expfml = 4
+linearfml = 5
+totaltests = 100
+
+x=[0]*(expnum+linearnum+2)
+
+g = open("NIA-bound/3-4-4-5-eq/paras","w+")
+g.write("%d\n" %totaltests)
+g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
+
+for testno in range(1,totaltests+1):
+    for i in range(1,expnum+linearnum+1):
+        x[i] = Int('x'+str(i))
+    pow = Function('pow',IntSort(),IntSort())
+
+    s = Solver()
+
+    for i in range (1,expnum+linearnum+1):
+        s.add(x[i] >= 0)
+
+    g.write("{")
+    expr = 0
+    for j in range (1,expnum+1):
+        coef = random.randint(-10,10)
+        expr = expr + coef * pow(x[j])
+        g.write("%d," %coef)
+    for j in range (1,expnum+linearnum+1):
+        coef = random.randint(-10**5,10**5)
+        expr = expr + coef * x[j]
+        g.write("%d," %coef)
+    num = random.randint(-10**5,10**5)
+    s.add(expr == num)
+    g.write("%d}\n" % num)
+
+
+    for i in range (2,expfml+1):
+        g.write("{")
+        expr = 0
+        for j in range (1,expnum+1):
+            coef = random.randint(-10,10)
+            expr = expr + coef * pow(x[j])
+            g.write("%d," %coef)
+        # bound cases
+        for j in range (1,expnum+linearnum+1):
+            coef = random.randint(-10**5,10**5)
+            expr = expr + coef * x[j]
+            g.write("%d," %coef)
+        
+        num = random.randint(-10**5,10**5)
+        s.add(expr <= num)
+        g.write("%d}\n" % num)
+    
+    for i in range (1,linearfml+1):
+        expr = 0
+        g.write("{")
+        for j in range (1,expnum+linearnum+1):
+            coef = random.randint(-10**2,10**2)
+            expr = expr + coef * x[j]
+            g.write("%d," %coef)
+        num = random.randint(-10**5,10**5)
+        s.add(expr <= num)
+        g.write("%d}\n" % num)
+            
[…]

### 2021-05-07 Benchmark/NIA-test/produce-unbound.py

@@ -1,0 +1,76 @@
+from z3 import *
+import random
+
+expnum = 3
+linearnum = 4
+expfml = 4
+linearfml = 5
+totaltests = 100
+
+x=[0]*(expnum+linearnum+2)
+
+g = open("NIA-unbound/3-4-4-5-eq/paras","w+")
+g.write("%d\n" %totaltests)
+g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
+
+for testno in range(1,totaltests+1):
+    for i in range(1,expnum+linearnum+1):
+        x[i] = Int('x'+str(i))
+    pow = Function('pow',IntSort(),IntSort())
+
+    s = Solver()
+
+    for i in range (1,expnum+linearnum+1):
+        s.add(x[i] >= 0)
+
+    for i in range (1,expfml+1):
+        g.write("{")
+        expr = 0
+        for j in range (1,expnum+1):
+            coef = random.randint(-10,10)
+            while coef == 0:
+                coef = random.randint(-10,10)
+            expr = expr + coef * pow(x[j])
+            g.write("%d," %coef)
+        
+        # unbound cases
+        for j in range (1,expnum+1):
+            coef = random.randint(-10**5,10**5)
+            expr = expr + coef * x[j]
+            g.write("%d," %coef)
+        # unbound cases
+        for j in range (expnum+1,expnum+linearnum+1):
+            coef = 0
+            g.write("%d," %coef)
+
+        num = random.randint(-10**5,10**5)
+        s.add(expr <= num)
+        g.write("%d}\n" % num)
+    
+    for i in range (1,linearfml+1):
+        expr = 0
+        g.write("{")
+        for j in range (1,expnum+linearnum+1):
+            coef = random.randint(-10**5,10**5)
+            expr = expr + coef * x[j]
+            g.write("%d," %coef)
+        num = random.randint(-10**5,10**5)
+        s.add(expr <= num)
+        g.write("%d}\n" % num)
+            
+    f = open('NIA-unbound/2-3-3-4/test'+str(testno),mode='w+')
+    f.write("(set-logic ALL)\n")
+    f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
+    f.write(s.sexpr())
+    f.write("(check-sat)\n")
+    f.write("(get-model)\n")
+    f.seek(0)
[…]

### 2021-05-07 Benchmark/produce-bound.py

@@ -1,69 +1,0 @@
-from z3 import *
-import random
-
-expnum = 2
-linearnum = 3
-expfml = 3
-linearfml = 4
-totaltests = 100
-
-x=[0]*(expnum+linearnum+2)
-
-g = open("NIA-bound/2-3-3-4/paras","w+")
-g.write("%d\n" %totaltests)
-g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
-
-for testno in range(1,totaltests+1):
-    for i in range(1,expnum+linearnum+1):
-        x[i] = Int('x'+str(i))
-    pow = Function('pow',IntSort(),IntSort())
-
-    s = Solver()
-
-    for i in range (1,expnum+linearnum+1):
-        s.add(x[i] >= 0)
-
-    for i in range (1,expfml+1):
-        g.write("{")
-        expr = 0
-        for j in range (1,expnum+1):
-            coef = random.randint(-10,10)
-            expr = expr + coef * pow(x[j])
-            g.write("%d," %coef)
-        # bound cases
-        for j in range (1,expnum+linearnum+1):
-            coef = random.randint(-10**5,10**5)
-            expr = expr + coef * x[j]
-            g.write("%d," %coef)
-        
-        num = random.randint(-10**5,10**5)
-        s.add(expr <= num)
-        g.write("%d}\n" % num)
-    
-    for i in range (1,linearfml+1):
-        expr = 0
-        g.write("{")
-        for j in range (1,expnum+linearnum+1):
-            coef = random.randint(-10**2,10**2)
-            expr = expr + coef * x[j]
-            g.write("%d," %coef)
-        num = random.randint(-10**5,10**5)
-        s.add(expr <= num)
-        g.write("%d}\n" % num)
-            
-    f = open('NIA-bound/2-3-3-4/test'+str(testno),mode='w+')
-    f.write("(set-logic ALL)\n")
-    f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
-    f.write(s.sexpr())
-    f.write("(check-sat)\n")
-    f.write("(get-model)\n")
-    f.seek(0)
-    d = f.readlines()
-    f.seek(0)
-    for i in d:
-        if i != "(declare-fun pow (Int) Int)\n":
-            f.write(i)
-    f.truncate()    
-    f.close()
-
-g.close()

### 2021-05-07 Benchmark/produce-unbound.py

@@ -1,76 +1,0 @@
-from z3 import *
-import random
-
-expnum = 2
-linearnum = 3
-expfml = 3
-linearfml = 4
-totaltests = 100
-
-x=[0]*(expnum+linearnum+2)
-
-g = open("NIA-unbound/2-3-3-4/paras","w+")
-g.write("%d\n" %totaltests)
-g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
-
-for testno in range(1,totaltests+1):
-    for i in range(1,expnum+linearnum+1):
-        x[i] = Int('x'+str(i))
-    pow = Function('pow',IntSort(),IntSort())
-
-    s = Solver()
-
-    for i in range (1,expnum+linearnum+1):
-        s.add(x[i] >= 0)
-
-    for i in range (1,expfml+1):
-        g.write("{")
-        expr = 0
-        for j in range (1,expnum+1):
-            coef = random.randint(-10,10)
-            while coef == 0:
-                coef = random.randint(-10,10)
-            expr = expr + coef * pow(x[j])
-            g.write("%d," %coef)
-        
-        # unbound cases
-        for j in range (1,expnum+1):
-            coef = random.randint(-10**5,10**5)
-            expr = expr + coef * x[j]
-            g.write("%d," %coef)
-        # unbound cases
-        for j in range (expnum+1,expnum+linearnum+1):
-            coef = 0
-            g.write("%d," %coef)
-
-        num = random.randint(-10**5,10**5)
-        s.add(expr <= num)
-        g.write("%d}\n" % num)
-    
-    for i in range (1,linearfml+1):
-        expr = 0
-        g.write("{")
-        for j in range (1,expnum+linearnum+1):
-            coef = random.randint(-10**5,10**5)
-            expr = expr + coef * x[j]
-            g.write("%d," %coef)
-        num = random.randint(-10**5,10**5)
-        s.add(expr <= num)
-        g.write("%d}\n" % num)
-            
-    f = open('NIA-unbound/2-3-3-4/test'+str(testno),mode='w+')
-    f.write("(set-logic ALL)\n")
-    f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
-    f.write(s.sexpr())
-    f.write("(check-sat)\n")
-    f.write("(get-model)\n")
-    f.seek(0)
-    d = f.readlines()
[…]

### 2021-05-07 FMCAD2021/paper.tex
affiliation
restructure the paper

@@
 \IEEEauthorblockA{\IEEEauthorrefmark{2}
-Institute of Information Science, Academia Sinica, Taiwan\\
+Institute of Information Science, Academia Sinica, Taiwan, Republic of China\\
 Email: yfc@iis.sinica.edu.tw}\\
@@
 
-solve string constraint is hard
-
-strategy: do unsat and sat separately
-
-strategy: using different procedure to (dis)prove validaity
-
-for disprove validatity, there are two approches, first is bound string length
-
-cannot handle $x.y \neq z  \wedge |x| > 2000$
-
-more recent approach is flattening
-
-it is known that word equaltion + flat regular constraints + len constraints
-is decidable
-
-it was unknown that whether word equations + flat regular constraints + len constraints + parseInt is decidable
+\input{1_intro.tex}
 
@@
 
-In this section, we introduce some basic concepts and theories that will be used later. 
-\subsection{Basic Concepts}
+\input{2_preliminary.tex}
 
-\paragraph{Sets and Strings}
-We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the set of natural numbers and integers, respectively. 
-$\mathbb{N}^+$ stands for the set of non-zero natural numbers. 
-Let $\Sigma$ be a finite alphabet,
-a string $w$ over $\Sigma$ is a sequence $a_1....a_n$ of characters from $\Sigma$.
-Empty string is denoted by $\epsilon$.
-$\Sigma^*$ denotes the set of all finite strings over $\Sigma$,
-and $\Sigma_{\epsilon}$ stands for $\Sigma \cup \{\epsilon\}$.
-For any string $w_1, w_2\in \Sigma^*$, 
-we use $|w_1|$ to denote the length of $w_1$,
-and $w_1\cdot w_2$ to denote the concatenation of $w_1$ and $w_2$.
-A language $L$ over $\Sigma$ is a subset of $\Sigma^*$.
+\section{Flattening string constraints with string-integer conversion}\label{sec:string-solving}
 
-There are two types of variables in string constraints,
-i.e., $X$, a set of string variables ranged over $\Sigma^*$,  
-and $Z$,  a set of integer variables ranged over $\mathbb{Z}$.
-As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
-essentially a pair of two mappings 
-$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
-where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.
+\input{3_flatten.tex}
 
-\paragraph{Finite State Automata}
-A Finite State Automaton is a tuple 
-$\mathcal{A}=\langle Q,\Sigma,\Delta,$ $q_{\textit{init}},$ $q_{\textit{acc}}\rangle$, 
-where $Q$ is a finite set of states, 
-$\Sigma$ is the given alphabet,
-$\Delta\subseteq Q\times \Sigma_\epsilon\times Q$ 
-defines the transition relations in $\mathcal{A}$.
-$q_{\textit{init}},q_{\textit{acc}}\in Q$ is the initial state and accepting state. 
-A sequence $q_0 \langle a_1 \rangle  q_1 ...
-\langle a_n\rangle q_n$ is called accepting if $q_0 = q_{\textit{init}}$, $q_n = q_{\textit{acc}}$ and $q_{i-1}\langle a_i \rangle q_i \in \Delta$ for $1\le i\le n$.
+\section{Decision procedure for $(\mathbb{N},+,2^x)$}
 
-\paragraph{Presburger Arithmetic} \label{PA}
-The Presburger Arithmetic (PA) is a first order theory over signature 
-$\Sigma_\mathbb{N}\Def  \{0,1,+,=\}$, where $0,1$ are constants, 
-$+$ is a binary function and $=$ is a binary predicate.
+\input{4_dec_proc.tex}
 
-PA can be axiomatized by the following axioms \cite{PA}
+\section{Optimizations}
 
-\begin{itemize}
-    \item $\forall x, \neg (x+1=0)$
-    \item $\forall x \forall y. x+1=y+1 \to x=y$
-    \item $F(0) \wedge (\forall x. F(x)\to F(x+1)) \to \forall x. F(x)$ 
-    \item $\forall x. x+0=x$
-    \item $\forall x \forall y. x+(y+1)=(x+y)+1$
-\end{itemize}
+\input{5_optimization.tex}
 
-Given the domain $\mathbb{N}$,
-the standard interpretation of PA interprets 
-$0,1$ to $0_\mathbb{N},1_\mathbb{N}\in \mathbb{N}$
-and $+,=$ to addition and equality over $\mathbb{N}$.
-We call a PA formula without quantifiers a quantifier-free PA formula.
+\section{Implementation and Experiments}
 
-PA is a decidable theory, 
-and the complexity of decidability is related to 
-the number and locations of quantifiers.
-Generally, 
-the upper bound (on deterministic time and space) 
-for deciding a formula of length $n$ is $2^{2^{2^{p n log(n)}}}$,
-where $p>1$ is a constant\cite{Oppen69}.
+\input{6_impl_exp.tex}
 
-\paragraph{Parikh Image}
-Given an alphabet $\Sigma$ and a string $w\in \Sigma^*$, 
-we define the set of Parikh variables 
-$\Sigma^\# \Def \{a^\# \mid a\in \Sigma\}$.
-The Parikh image of $w$ is a function 
-$\mathbb{P}(w): \Sigma^\# \mapsto \mathbb{N}$,
-which maps each symbol $a^\#\in \Sigma^\#$ to the number of occurrences
-of $a$ in $w$.
-For example, let $w\Def aabba$,
-then $\mathbb{P}(w)(a^\#)=3,\mathbb{P}(w)(b^\#)=2$.
+\section{Conclusion}
 
-For a language $L\subseteq \Sigma^*$, 
-define the Parikh image of $L$ to be 
-$\mathbb{P}(L)\Def \{\mathbb{P}(w) | w\in L\}$.
-We say a language $L$ is \emph{Parikh-definable} 
-if $\mathbb{P}(L)$ can be characterized by a quantifier-free PA formula over $\Sigma^\#$, 
-where $a^\#$ in the formula encodes the number of occurrences of $a$.
-It is well known that
-any context-free language (therefore regular language) 
-is Parikh definable \cite{Parikh66}.
-
-\paragraph{String Terms}
-Given a finite alphabet $\Sigma$ and 
-a finite set $X$ of string variables over $\Sigma^*$,
-we define the set of terms $\textit{Terms}(\Sigma,X)$ 
-to be the smallest set satisfying
-\begin{itemize}
-    \item[1] $\Sigma\cup \{\epsilon\} \cup X \subseteq \textit{Terms}(\Sigma,X)$;
-    \item[2] if $t_1,t_2\in \textit{Terms}(\Sigma,X)$, then $t_1 \cdot t_2 \in \textit{Terms}(\Sigma,X)$.
-\end{itemize}
-
-We extend $I_X$ to $\textit{Terms}(\Sigma,X)$ by letting $I_X(\epsilon)=\epsilon$, 
-for $a\in \Sigma, I_X(a)=a$,
-and $I_X(t_1\cdot t_2)= I_X(t_1)\cdot I_X(t_2)$.
-
-\paragraph{String Constraints} \label{par: string constraints}
-Given a constraint $\phi$ and an interpretation $I$,
-$I\models \phi$ denotes that $I$ satisfies $\phi$,
-and $I$ is called a \emph{model} of $\phi$.
-We use $\lVert \phi\rVert$ to denote the set of all models of $\phi$.
-
-We define the following three forms of atomic string constraints:
-\begin{itemize}
-\item An equality constraint $\phi_e$ is of the form 
-$t_1 = t_2$, where $t_1,t_2\in \textit{Terms}(\Sigma,X)$.
-We define $\lVert \phi_e \rVert = \{I\mid I(t_1)=I(t_2)\}$.
-Inequality constraints can be defined analogously.
-
-\item A regular constraint $\phi_r$ is of the form 
-$x\in L(\mathcal{A})$,
-where $x\in \mathbb{X}$ and $\mathcal{A}$ is a finite state automaton.
-We define $\lVert \phi_r \rVert = \{I\mid I(x)\in L(\mathcal{A})\}$.
-
-\item A length constraint $\phi_l$ is a linear constraint over 
-$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
-where $|\cdot |$ is the length function.
-We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
-\end{itemize}
-
-In Section~\ref{sec:string-solving},  
-we will introduce a new form of atomic string constraints, i.e., \emph{string-number conversion constraints}.
-A string constraint is a Boolean combination of atomic string constraints possibly with quantifications over $X\cup Z$. Other notions are same as in the first-order logic.
-
-Giving a string constraint $\Psi$,
-the problem of string constraints solving is to decide 
-whether $\lVert \Psi \rVert$ is empty,
-if not, to compute an interpretation $I$ that satisfies $\Psi$.
-
-\subsection{Flat Automata and Flat Languages}
-
-Given a string constraint $\Psi$,
-the general problem of deciding whether $\lVert \Psi \rVert$ is empty is undecidable.
-However, 
-the problem becomes decidable when certain restrictions are imposed.
-One of the restriction is by flat automata and flat languages, 
-defined below.
-
-\paragraph{Flat Languages and Automata}
-For a fixed alphabet $\Sigma$,
-we say a language $L$ over $\Sigma$ to be \emph{$\langle p,q \rangle$-flat} if 
-there exist strings $w_1,...,w_q \in \Sigma^*$ such that
-$|w_i|\le p$ for all $i:1\le i \le q$ 
-and $L = (w_1)^*(w_2)^*...(w_q)^*$. 
-We use $\alpha$ to denote $\langle p,q \rangle$, 
-and call it the \emph{abstraction parameter} of $L$.
-Intuitively,
-a flat language with abstraction parameter $\alpha = \langle p,q \rangle$
-consists of $q$ loops and the length of each loop body is equal or less than $p$.
-For example,
-$L = (ab)^*(a)^*(bb)^*$ is a $\langle 2,3 \rangle$-flat language.
-
-Flat automata are a special form of finite state automata that 
-recognize flat languages.
-Fix the abstraction parameter $\alpha=\langle p,q\rangle$,
-a flat automaton consists of $q$ loops,
-each loop is a circle of $p$ states.
-Formally, 
-an $\alpha$-flat automaton contains $p q$ states at most,
-and we name the states from $1$ to $p q$,
-$1$ is the initial state and $(p q - p + 1)$ is the accepting state.
-We use $\cdot$ as a placeholder for some symbol in $\Sigma_\epsilon$,
-the transition relations of state $i$ are defined as 
-\begin{itemize}
-    \item if $i\  \text{mod}\  p = 1$ and $i \neq pq-p+1$, then 
-    $(i,\epsilon,i+p)\in \Delta$,
-    $(i, \cdot ,i+1) \in \Delta$;
-    \item if $i\  \text{mod}\  p = 0$, then 
-    $(i,\cdot , i-p+1) \in \Delta$;
-    \item otherwise, $(i,\cdot, i+1) \in \Delta$.
-\end{itemize}
-
-A $\langle 2,3 \rangle$-flat automaton 
-that recognizes $L\Def  (ab)^*(a)^*(bb)^*$ is shown in 
-figure (\ref{fig: FA}).
-
-\begin{figure}[ht]
-    \centering 
-    \begin{tikzpicture}
-        \node[state,           ] (4) {$4$};
-        \node[state, left  of=4] (2) {$2$};
-        \node[state, right of=4] (6) {$6$};
-        \node[state, initial, below of=2] (1) {$1$};
-        \node[state, below of=4] (3) {$3$};
-        \node[state, accepting, right of=3] (5) {$5$};
-
-        \draw 
-        (1) edge[above] node{$\epsilon$} (3)
-        (3) edge[above] node{$\epsilon$} (5)
-
-        (1) edge[bend left,left] node{$a$} (2)
-        (2) edge[bend left,right] node{$b$} (1)
-
-        (3) edge[bend left,left] node{$a$} (4)
-        (4) edge[bend left,right] node{$\epsilon$} (3)
-
-        (5) edge[bend left,left] node{$b$} (6)
-        (6) edge[bend left,right] node{$b$} (5);
-    \end{tikzpicture}
-    \caption{A $\langle 2,3 \rangle$-flat automaton 
-that recognizes $L\Def  (ab)^*(a)^*(bb)^*$}
-    \label{fig: FA}
-\end{figure}
-
-\paragraph{Generic Flat Languages and Automata}
-Fix $\alpha = \langle p,q \rangle$,
-we define the \emph{generic $\alpha$-flat language} is the union of all $\alpha$-flat languages, denoted by $\mathbb{F}(\alpha)$.
-Now, we try to define an automaton that recognizes all $\alpha$-flat languages,
-i.e., collects all behaviors of $\alpha$-flat automata.
-
-Intuitively, 
-the generic automaton is obtained by introducing a new alphabet (
-a multi-set with $p q$ copies of the original alphabet) and 
-adding more transitions (labels),
-the states and the overall framework remain unchanged. 
-In details, a generic $\alpha$-flat automaton is still a finite state automaton over
-$\Sigma(\alpha)\Def \{a(i)| (a\in \Sigma_\epsilon) \wedge i\in \mathbb{N}:1\le i \le pq\}\cup \{\epsilon\}$.
-The states are still named from $1$ to $pq$, 
-the initial state is $1$ and the accepting state is $(pq-p+1)$.
-The transition relations for state $i$ are defined as 
-\begin{itemize}
-    \item if $i\  \text{mod}\  p = 1$ and $i\neq pq-p+1$, then 
-    $(i,\epsilon,i+p)\in \Delta$
-    and $\forall s\in \Sigma_{\epsilon}. (i, s(i) ,i+1) \in \Delta$;
-    \item if $i\  \text{mod}\  p = 0$, 
-    $\forall s \in \Sigma_{\epsilon}. (i,s(i), i-p+1) \in \Delta$;
-    \item otherwise, $\forall s \in \Sigma_{\epsilon}. (i,s(i), i+1) \in \Delta$.
-\end{itemize}
-
-For $\Sigma = \{a,b\}$, an example of generic $\langle 2,3 \rangle$-flat automaton is shown in figure (\ref{fig: GFA}).
-
-\begin{figure}[ht]
-    \centering 
-    \begin{tikzpicture}
-        \node[state,           ] (4) {$4$};
-        \node[state, left  = 2cm of 4] (2) {$2$};
-        \node[state, right = 2cm of 4] (6) {$6$};
-        \node[state, initial, below of=2] (1) {$1$};
-        \node[state, below of=4] (3) {$3$};
-        \node[state, accepting, below of=6] (5) {$5$};
-
-        \draw 
-        (1) edge[above] node{$\epsilon$} (3)
-        (3) edge[above] node{$\epsilon$} (5)
-
-        (1) edge[bend left, pos =0.2 ,left] node{$a(1)$} (2)
-        (1) edge[bend left, pos =0.5 ,left] node{$b(1)$} (2)
-        (1) edge[bend left, pos =0.8 ,left] node{$\epsilon(1)$} (2)
-
-        (2) edge[bend left, pos = 0.2 ,right] node{$\epsilon(2)$} (1)
-        (2) edge[bend left, pos = 0.5 ,right] 
-        node{$b(2)$} (1)
-        (2) edge[bend left, pos = 0.8 ,right] 
-        node{$a(2)$} (1)
-
-        (3) edge[bend left, pos =0.2 ,left] node{$a(3)$} (4)
-        (3) edge[bend left, pos =0.5 ,left] node{$b(3)$} (4)
-        (3) edge[bend left, pos =0.8 ,left] node{$\epsilon(3)$} (4)
-
-        (4) edge[bend left, pos = 0.2 ,right] node{$\epsilon(4)$} (3)
-        (4) edge[bend left, pos = 0.5 ,right] 
-        node{$b(4)$} (3)
-        (4) edge[bend left, pos = 0.8 ,right] 
-        node{$a(4)$} (3)
-
-        (5) edge[bend left, pos =0.2 ,left] node{$a(5)$} (6)
-        (5) edge[bend left, pos =0.5 ,left] node{$b(5)$} (6)
-        (5) edge[bend left, pos =0.8 ,left] node{$\epsilon(5)$} (6)
-
-        (6) edge[bend left, pos = 0.2 ,right] node{$\epsilon(6)$} (5)
-        (6) edge[bend left, pos = 0.5 ,right] 
-        node{$b(6)$} (5)
-        (6) edge[bend left, pos = 0.8 ,right] 
-        node{$a(6)$} (5);
-    \end{tikzpicture}
-    \caption{The generic $\langle 2,3 \rangle$-flat automaton}
-    \label{fig: GFA}
-\end{figure}
-
-However,
-the resulted automaton may accept languages that are not in $\mathbb{F}(\alpha)$,
-because in different passes inside a loop, 
-the automaton can choose different symbols between identical pairs. 
-To avoid this problem,
-we add a so-called purity condition on the accepting language of generic flat automata,
-which is equivalent to intersecting the language of a generic flat automaton 
-with a language that encodes the purity condition.
-
-We say a string $w\in (\Sigma(\alpha))^*$ is pure if for all $i: 1\le i \le p q$,
-and $a,b\in \Sigma$, 
-$a\neq b \wedge \#w(a(i))>0$ implies $\#w(b(i))=0$.
-Formally, the purity condition is defined by 
-\begin{equation} \label{eq:purity}
- \bigwedge_{1\le i\le pq}\bigwedge_{a,b\in \Sigma, a\neq b} ({a(i)}^\#>0)\to ({b(i)}^\#=0)\, . 
-\end{equation}
-
-We denote the accepting language of the generic $\alpha$-flat automaton by $\mathbb{G}(\alpha)$.
-Note that $\mathbb{G}(\alpha)$ is a language over $\Sigma_\alpha$,
-but what we want is a language over $\Sigma$.
-So we define a renaming function $R:\Sigma(\alpha)\mapsto \Sigma$ such that for all $a(i) \in \Sigma_\alpha, R(a(i))=a$,
-and $R(\epsilon) = \epsilon$.
-Define $\mathbb{G}'(\alpha) \, \Def \, \{R(w) \mid w\in \mathbb{G}(\alpha)\}$, 
-for simplicity, we write
-$\mathbb{G}'(\alpha)=R(\mathbb{G}(\alpha))$.
-
-The important feature of generic flat autamata
-is that every word $w\in \mathbb{G}(\alpha)$ is uniquely determined by its Parikh image $\mathbb{P}(w)$.
-
-\subsection{Flattening}
-
-The flattening technique was first introduced 
-in \cite{Abdulla 2017}.
-Fix an alphabet $\Sigma$ and an abstraction parameter $\alpha$, 
-for a given atomic string constraint $\phi$, 
-flattening $\phi$ with parameter $\alpha$ 
-results in a new string constraint $\phi_\alpha$,
-such that 
-$R(\lVert \phi_\alpha \rVert) = \lVert \phi \rVert \cap \{I \mid \forall x\in X, I(x)\in \mathbb{G}'(\alpha)\}$,
-where $R$ is the renaming function with its domain extended to interpretations in the normal manner.
-Intuitively, it restricts $\phi$ to interpret 
-string variables over $\mathbb{G}'(\alpha)$.
-
-\cite{Abdulla 2017} discussed the flattening of basic 
-string constraints including equality, integer, (regular) grammar and transducer constraints. 
-For an atomic string constraint $\phi$,
-the flattening $\phi_\alpha$ is still an atomic string constraint 
-and is Parikh definable,
-so its Parikh image can be expressed by a quantifier free PA formula.
-Together with the purity condition,
-we obtain an existential quantified PA formula $\rho$.
-$\rho$ will be sent to a SMT solver,
-if the solver returns an solution $\theta$,
-then we can construct an interpretation for $\phi_\alpha$ from $\theta$,
-otherwise it means $\phi$ is unsatistiable when 
-string variables are interpreted  to $\alpha$-flat languages.
-
-Take a regular constraint $\phi = x\in L(A)$ for example,
-the flattening of $\phi$ resutls in a new finite state automaton $A'$ over $\Sigma(\alpha)$,
-which encodes running $A$ ``in parallel" 
-with the generic $\alpha$-flat automaton.
-Let $\rho_1$ be the formula describing the Parikh image of $A'$,
-which is a formula over variable sets $\Sigma(\alpha)^\#$.
-Let $\rho_2$ be the purity condition \eqref{eq:purity}.
-Then we obtain the PA formula $\exists (\Sigma(\alpha))^\#. \rho_1 \wedge \rho_2$.
-In order to distinguish between different string variables,
-we may replace $a(i)^\# \in \Sigma(\alpha)^\#$ by $(x,a(i))^\#$.
-
-Since the structure of a flat automaton is decided 
-by its abstraction parameter $\alpha$, 
-a Counter-Example Guided Abstraction 
-Refinement (CEGAR) framework is designed, which contains both an under- and an 
-over-approximation module, to search the possible values of $\alpha$.
-The termination for the overall algorithm is not guaranteed.
-
-\section{Solving String Constraints with \textsf{parseInt} Function}\label{sec:string-solving}
-
-The string-number conversion functions are commonly used functions
-in most of programming languages,
-for example,
-\verb+parseInt()+ in Java and \verb+Int()+ in Python.
-The functions usually take two parameters, 
-a string over the agreed alphabet $\Sigma$
-and an optional parameter denotes the radix.
-They parse the string according to the rules indicated by the radix,
-and return an integer denoted by the string.
-
-From the view of string constraints,
-string-number conversion functions give rise to a new form of string constraints
-and are more expressive than length constraints.
-So we consider extending string constraints with 
-{\parseInt} function. 
-As the general problem of string constraints is undecidable,
-we still adopt the idea of flattening, 
-i.e., variables are restricted to (generic) flat languages.
-This problem has been investigated in \cite{POPL20},
-which defined a special form of flat restriction (straight-line PFA) and 
-proposed a heuristic search method.
-
-In this section, 
-we describe the problem of interest first, and then 
-present an reduction from the problem of solving flat string constraints with {\parseInt} function
-to the decidability problem of Presburger Arithmetic with exponential functions.
-Hence, we identify a decidable subset of string constraints, which is the largest one with decidability so far to the best of our knowledge.
-
-\subsection{String-Number Conversion Function}
-
-Commonly, 
-{\parseInt} function takes a string representation of a decimal
-number and returns an integer.
-Since we focus on the decidability,
-we define a binary version of {\parseInt},
-which takes a binary string and returns a decimal integer number.
-For example,
-${\parseInt}('111')=7$. 
-Clearly,  our decision procedure given in this paper 
-can be adapted to string constraints with other string to number conversion function without substantial change. 
-Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
-from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.
-
-\begin{definition} Let $\Sigma_{\textit{num}}=\{'0','1'\}$,
- ${\parseInt}:\Sigma_{\textit{num}}^*\mapsto \mathbb{N}$ is recursively defined by:
-    for $w\in \Sigma_{\textit{num}}^*$
-    \begin{itemize}
-        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
-        \item if $|w|=1, {\parseInt}('w')=w_\mathbb{N}$;
-        \item for $|w|\ge 2$ and $w=w_1 \cdot w_2$, 
-        where $|w_1|, |w_2| \ge 1$, 
-        ${\parseInt}(w) = {\parseInt}(w_1)2^{|w_2|}+{\parseInt}(w_2)$.
-    \end{itemize} 
-\end{definition}
-
-Now, we introduce a new form of atomic string constraints: 
-a {\parseInt} constraint $\phi$ is of the form 
-$n \sim {\parseInt}(t)$,
-where $n$ is an integer term, $\sim \in \{\le,<,=,>,\ge\}$ and $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$ is a string term.
-$\lVert \phi\rVert \Def  \{I \mid I(n)\sim {\parseInt}(I(t))\}$.
-
-In what follows, we only  consider the problem in the case when $t=x$ and $x$ is restricted to (generic) flat languages. 
-%If $t$ is not a single variable, 
-For the general case $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$, 
-it can be reduced to this special case by induction on the structure of $t$. 
-%we can reduce the problem by separating $t$ (corresponding to the $|w|>2$ case in definition).
-
-Given an $\alpha$-flat language $L$,
-we assume $\alpha=\langle p,q \rangle$ and $L=(w_1)^*...(w_q)^*$,
-where $p,q$ and $w_i(1\le i \le q) $ are known.
-We further assume that $x = (w_1)^{\beta_1} ... (w_q)^{\beta_q}$,
-then we have
-\begin{align}
-    &{\parseInt}(x) \notag \\
-    =&{\parseInt}((w_1)^{\beta_1} ... (w_q)^{\beta_q}) \notag \\
-    =&{\parseInt}((w_1)^{\beta_1} ... (w_{q-1})^{\beta_{q-1}})\cdot 2^{\beta_q |w_q|}
-    + {\parseInt}((w_q)^{\beta_q}) \label{parse}
-\end{align}
-(\ref{parse}) is a recursive expression.
-So we only need to deal with the basic case ${\parseInt}((w_q)^{\beta_q})$, 
-where $w_q \neq \epsilon$
-\begin{align}
-    {\parseInt}((w_q)^{\beta_q}) &= \sum_{i=0}^{\beta_q-1} {\parseInt}(w_q)\cdot 2^{|w_q|\cdot i}  \notag \\
-    &={\parseInt}(w_q)\frac{2^{|w_q|\cdot \beta_q}-1}{2^{|w_q|}-1}
-    \label{parseInt-basic}
-\end{align}
-In (\ref{parseInt-basic}), since $w_q$ and $|w_q|$ are known, 
-they can be regarded as constants.
-The only unknown variable is $\beta_q$.
-
-Combine (\ref{parse}) and (\ref{parseInt-basic}),
-the constraint $n={\parseInt}(x)$ can be expressed by an arithmetic expression with 
-$n$ and $(\beta_1,...,\beta_q)$,
-inevitably with exponential components.
-
-Take $n={\parseInt}((11)^a(10)^b)$ for example.
-\begin{align}
-    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
-    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
-    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
-\end{align}
-
-So we have 
-\begin{equation}
-    3\cdot n = 3\cdot 2^{2a+2b}-3\cdot 2^{2b}+2\cdot 2^{2b}-2. \notag
-\end{equation}
-
-Observe the form of the above equation,
-$a,b,n$ are integer variables and either occur in 
-an exponential term or a linear term.
-This is always the case,
-so the problem can be reduced to 
-the decidability of PA with exponential function.
-
-When $x$ of {\parseInt} constraints is restricted to the
-generic $\alpha$-flat language ($\alpha$ is fixed),
-the (\ref{parse}) and (\ref{parseInt-basic}) still hold 
-but $w_i(1\le i\le q)$ is known.
-However,
-by definition,
-the generic $\alpha$-flat language is the finite union of all $\alpha$-flat languages,
-so we can enumerate all possible values for $w_i$.
-In this way,
-the problem can still be reduced to the decidability of PA with exponential function (\paexp), i.e.,
-\begin{theorem} \label{thm:string-parInt}
-If {\paexp} is decidable, then the satisfiability (validity) of string constraints with {\parseInt} in which all string variables 
-are ranged over flat strings is decidable. 
-\end{theorem}
-
-\subsection{Decidability of \paexp} \label{sec:3.2}
-
-For a first order theory $T$,
-we say theory $T$ admits quantifier elimination (QE) if for any formula in $T$, 
-there is a quantifier-free formula equivalent to it.
-It is well-known that if a theory admits QE, 
-then it is a decidable theory.
-
-The formal definition of PA is given in section \ref{PA}.
-Here we add the ordering predicate $\le$ into the signature,
-which can be defined by $x \le y\, \Def \, \exists z. x+z=y$.
-However, the theory PA so far does not admit QE, 
-for example, consider the formula $\exists x.x = y + y$. 
-We augment the theory with countable unary divisible predicates
-$n|x$, where $n\in \mathbb{N}$, 
-$n|x$ is true if and only if $x\ \text{mod}\ n=0$ holds.
-This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.
-
-We then introduce a theory {\paexp}, denoted by $(\mathbb{N},+,2^x)$, that we work on.
-\begin{definition}
-    Let $\mathcal{L}=\{0,1,+,\le,n \, |x(n\in \mathbb{N}),2^x,l_2(x)\}$, 
-     $(\mathbb{N},+,2^x)$ be a $\mathcal{L}$-theory that has domain $\mathbb{N}$, where 
-    \begin{itemize}
-        \item  $2^x$ is interpreted to the exponential function of $2$ over $\mathbb{N}$; 
-        \item interpretations of $0,1,+,\le,=$ are consistent with PA;
-        \item for $n\ge 1$,$n|x$ holds iff $\exists y.x=ny$;
-        \item $2^0=1$, for $n \ge 1, 2^n = 2^{n-1}+2^{n-1}$;
-        We further assume that if $m,n\in \mathbb{N}, m\le n$, 
-        then $2^{m-n} = 1$
-        \item $l_2(0)=0$; for $n \ge 1,l_2(n) = y$ iff $2^y \le n < 2^{y+1}$;  $l_2(m-n) = 0$
-        if $m,n\in \mathbb{N}$ and $m\le n$. 
-    \end{itemize}
-\end{definition}
-
-$\lambda_2(x) = 2^{l_2(x)}$ can be defined by $l_2(x)$,
-intuitively, $\lambda_2(x)$ means the maximal power of 2 that is not larger than $x$.
-Then we have $\lambda_2(x) \le x \le 2\lambda_2(x)-1$,
-which will be useful in our proof.
-
-\begin{definition}
-    For a strictly increasing function $f:\mathbb{N}\mapsto \mathbb{N}$, 
-    $f$ is said to be compatible with addition 
-    if for every $m\in \mathbb{N}^+$,  $f$ modulo $m$ is periodic, 
-    and for any term $A(x)=\sum_{1\le i\le n} a_i f(x+b_i)$, 
-    where $n \in \mathbb{N}^+, a_i ,b_i \in \mathbb{Z}$,
-    one of the following holds:
-    \begin{itemize}
-        \item $A(x)$ is bounded; 
-        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
-        $\forall x. A(x+\Delta_A)\ge f(x)$;
-        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
-        $\forall x. -A(x+\Delta_A)\ge f(x)$.
-    \end{itemize}
-\end{definition}
-
-%\znjSide{\textbf{Solved}. I think you need to explain the meanings of compatibility.}
-
-Semenov proved that for any function $f$ that is compatible with addition, 
-theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
-Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
-by \cite{Semenov84}.
-Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
-Unfortunately, Point's algorithm is flawed, one problem 
- lies in the discussion to eliminate exponential terms. 
-% making the algorithm fail to handle "$a_0<0$" case.
-For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
-the expected result should be $y\le 2^{10} =1024$,
-but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
-% 这里好像还有另一个与Francoise一起的作者
-
-In the next section, within Point's framework, we give a revised algorithm
- to the decision problem of {\paexp} with 
-both subtle improvements and crucial corrections, that 
-provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to 
-Theorem~\ref{thm:string-parInt}.
-
-\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
-Based on Point's work \cite{Point86}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
-where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 
-Since $\forall x. F = \neg \exists x. \neg F$, we further assume the quantifier $Q$ to be the existential quantifier, 
-that is, $\exists x.\theta(x,\bar{y})$. 
-For eliminating quantifiers in arbitrary formula, 
-we apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.
-
-We divide the whole QE algorithm into 4 steps.
-The first step $\textbf{Normalization}$ can be viewed as a pre-processing step, which substitutes ``complex" terms like $2^{2^x}$, $l_2(3x+y)$ with ``simple" terms by introducing new variables $x_i$.
-Now we move forward to handle a ``simpler" formula, however, at the cost of more quantified variables.
-
-The other three steps undertake the task to eliminate the introduced variables $x_i$ and the original variable $x$ (will be denoted by $x_0$) one by one. 
-The main body of \textbf{QE-with-Order} step is a loop to enumerate all possible orders among quantified variables. 
-In each iteration, according the given (decreasing) order (corresponding to a for loop), we invoke \textbf{QE-exp} or 
-\textbf{QE-linear} to eliminate these variables one by one. 
-%depending on  eliminates the maximal variable in the remaining $\bar{x}$.
-So it is necessary to specify an ordering for $\bar{x}$ at the beginning, as we will see, this is the main reason for the complexity of this algorithm.
-During each iteration of the inner loop (the for loop), 
-if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atomic formula, 
-we will invoke \textbf{QE-exp} to produce a formula equivalent to the atom where $x_i$ occurs linearly;  
-otherwise, if $x_i$ occurs linearly in the formula, \textbf{QE-linear} will be invoked to eliminate all occurrences of $x_i$, 
-and this procedure is similar to the classic QE algorithm for PA.
-
-A more detailed description is given below.
-
-\subsection{Normalization}
-
-In order to show $(\mathbb{N},+,2^x)$ admits QE , 
-it is sufficient to show that any 1-existential formula $\exists x.\theta(x,\bar{y})$ indeed does, 
-where $\theta(x,\bar{y})$ is a quantifier-free formula. 
-However, the form of $\exists x.\theta(x,\bar{y})$ is unknown so it may contain terms difficult to handle such as $2^{3x+y+1}$ or $l_2(x)$.
-The \textbf{Normalization} step simplifies these terms by introducing new variables, 
-for example, $\exists x.2^{3x+y+1}>10$ is equivalent to
-$$\exists x\exists x_1.2^{x_1}>10 \wedge x_1 =3x+y+1\,.$$
-
-Logarithm functions are replaced by exponential functions, take $\exists x.l_2(x)>3$ for example, $l_2(x)$ is replaced by a new variable $x_1$, and we have  
-$$\exists x\exists x_1. x_1 > 3 \wedge 2^{x_1}\le x \wedge x\le 2^{x_1+1}-1\,.$$
-
-The \textbf{Normalization} step goes like this, first transform the given formula $\theta(x,\bar{y})$ into the Negation Normal Form (NNF).
-Then $\theta(x,\bar{y})$ becomes a Boolean combination (with only $\wedge$ and $\vee$) of literals.
-
-Make substitutions by introducing new variables according to the \textit{while} statements in the pseudo-code.For consistency, we rename the original $x$ to be $x_0$ and assume $n$ new quantified variables are introduced.
-
-After introducing new variables and substitution, 
-we obtain a formula where if an exponential term contains a quantified variable $x_i(0\le i\le n)$, the term should be $2^{x_i}$.
-We then collect all terms with $x_i(0\le i\le n)$ together, 
-it will be of the form $s(\bar{x})\Def \sum_{i=0}^{n} a_i 2^{x_i} + \sum_{i=0}^n b_i x_i$, where $a_i,b_i(0\le i \le n)$ are all constants.
-Other terms including constants and terms of $\bar{y}$ are collected,  
-denoted by $t(\bar{y})$. 
-Since $\bar{y}$ are free variables, 
-we will regard $t(\bar{y})$ as a constant.
-Inequalities and equalities will all be transformed into $s(\bar{x})\le t(\bar{y})$, for example, $s(\bar{x}) = t(\bar{y})$ will be replaced by $s(\bar{x})\le t(\bar{y})\wedge t(\bar{y})\le s(\bar{x})$.
-
-At the end, the resulted formula $\theta'$ only contains literals of the forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ and $\neg  (k|s(\bar{x})+t(\bar{y}))$.
-
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{1-existential formula $\exists x.\theta(x,\bar{y})$}
-    \KwOut{n-existential formula $\exists \bar{x}.\theta'(\bar{x},\bar{y})$}
-
-    $\theta' :=  \theta(x,\bar{y})$ \;
-    Transform $\theta'$ into NNF\;
-    \tcp{ i is used for counting the introduced variables }
-    i = 1\; 
-    \tcp{ $\bar{y}$ are regarded as constants} 
-    \While{there is a term $l_2(t)$, $t$ is not a constant}
-    {
-        $\theta' :=  \theta'[x_i/l_2(t)] \wedge (2^{x_i}\le t) \wedge (t\le 2^{x_i+1}-1)$\;
-        $i:= i+1$\;
-    }
-    \While{there is a term $2^t$, $t$ is not a variable $x_j(j<i)$ or a constant}
-    {
-        $\theta' :=  \theta'[2^{x_i}/2^t]\wedge (x_i\le t) \wedge (t \le x_i)$\;
-        $i:= i+1$\;
-    }
-    $n:= i-1$\;
-    \tcp{ Collect quantified variables $x_i$}
-    Transform all atoms into forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ or $\neg  (k|s(\bar{x})+t(\bar{y}))$\;
-    Return $\exists x_0,...,\exists x_n. \theta'$
-
-    \caption{Normalization}
-\end{algorithm}
-
-\subsection{QE-with-order} 
-During the \textbf{Normalization} step, 
-we simplify the origin formula at the cost of more quantified variables. 
-Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
-and it has $n$ new variables $x_i$ ($x$ is denoted by $x_0$).
-As a consequence, we need to eliminate all quantified variables one by one.
-
-To the end, we denote the set of all orders among the $n+1$ variables by 
- $S_{n+1}$, and then enumerate these orders one by one in the outer \textit{for} loop. 
- For the considered order $\sigma$, we first add the ordering information to 
- the quantifier free formula, and then according to the 
- order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{QE-exp} first, and then invoking \textbf{QE-linear} (the inner \textit{for} loop). 
- In each iteration of the inner \textit{for} loop, 
-if $x_{\sigma{(i)}}$ occurs in an exponential term in an atomic formula, 
-\textbf{QE-exp} is invoked and a formula   equivalent to the atom is returned, in which $x_i$ occurs linearly;  
-then $x_i$ occurs linearly in the formula, thus \textbf{QE-linear} is further invoked to eliminate all occurrences of $x_i$. 
-The returned formula is the disjunction of all formulas resulted in all iterations of the outer \textit{for} loop.
-
-\iffalse
-put , We now specify an order of $\bar{x}$, and recursively eliminate the maximal element in the remaining $\bar{x}$. 
-That is to say,
-every execution of \textbf{QE-exp} and \textbf{QE-linear} will eliminate the largest element $x_i$ in $\bar{x}$.
-However, as you may notice, if there is no clue about the order of $\bar{x}$, 
-we will have to use the idea of permutation group and the number of sub-cases will blow up, in the worst case, $(n+1)!$ cases.
-
-Here is how \textbf{QE-exp} and \textbf{QE-linear} are invoked in
-\textbf{Ordering and QE}. 
-\fi
-
-%我们这里还是考虑用循环群来做，因为如果每一次只找到最大的$x_i$，那么递归的执行还是需要整个$\bar{x}$的顺序。但如果一开始就知道了，后面有些操作可以简化
-
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{a normalized $(n+1)$-existential formula $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
-    where $\bar{x}=(x_0,...,x_n)$} 
-    \KwOut{a equivalent quantifier free formula without $\bar{x}$}
-
-    Let $S_{n+1}$ denote the group of permutations on $\{0,1,...,n\}$\;
-    $\phi :=  \textit{False}$\;
-    \For{each $\sigma \in S_{n+1}$ }
-    {
-        $\theta_{\sigma}:=  \theta(\bar{x},\bar{y})\wedge \bigwedge_{j=0}^{n-1}(x_{\sigma(j)}\le x_{\sigma(j+1)})$\;
-
-        \tcp{recursively eliminate the maximal $x_i$ in $\bar{x}$}
-        %\znj{How to effectively determine the maximal $x_i$?} 
-        \For{i from n to 0}
-        {
-            \tcp{if $x_{\sigma(i)}$ occurs exponentially in $\theta_\sigma$}
-            $\theta_{\sigma}:= \textbf{QE-exp}(x_{\sigma(i)},\theta_{\sigma})$\;
-            \tcp{now $x_{\sigma(i)}$ occurs only linearly in $\theta_\sigma$}
-            $\theta_{\sigma}:= \textbf{QE-linear}(x_{\sigma(i)},\theta_{\sigma})$\;
-        }
-        $\phi :=  \phi \vee \theta_{\sigma}$\;
-    }
-\iffalse
-    \For{i from $0$ to $n$}
-    {
-        \tcp{ specify the maximal quantified variable}
-        $\theta_i :=  \theta(\bar{x},\bar{y}) \wedge (\bigwedge_{0\le j \le n} x_i - x_j \ge 0)$\;
-        \tcp{ eliminate $x_i$}
-        \While{$x_i$ occurs in an exponential term in formula $\tau(x_i,\bar{x},\bar{y})$}{
-            \tcp{ QE-exp outputs a formula equivalent with $\tau(x_i,\bar{x},\bar{y})$ where $x_i$ occurs linearly }
-            $\theta_i :=  \theta_i[\text{QE-exp}(\tau(x_i,\bar{x},\bar{y}))/\tau(x_i,\bar{x},\bar{y})]$\;
-        }   
-        transform $\theta_i$ into disjunction normal form $\theta_i = \bigvee_s \theta_{i,s}$\;
-        \For{ every $\theta_{i,s}$ }{
-            $\theta'_{i,s} :=  \text{QE-linear}(x_i,\theta_{i,s})$
-        }
-        $\theta'_i =\bigvee_s \theta'_{i,s}$\;
-    }
-\fi
-    output $\phi$
-    \caption{QE-with-order}
-\end{algorithm}
-%\znj{I cannot understand why you need to do QE for all permutation $\sigma$? I think 
-%one is enough.}
-
-\subsection{QE-exp}
-
-\textbf{QE-exp} and \textbf{QE-linear} are the most technical parts of our QE algorithm.
-As mentioned, 
-\textbf{QE-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as 
-inputs, and  outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
-The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{QE-linear}.
-For simplicity, we will abuse   $x_i$ for $x_{\sigma(i)}$, 
-and  $\bar{x}$ for $(x_{\sigma(0)},...,x_{\sigma(i-1)})$ ( remind that $x_{\sigma(i+1)},...,x_n$ have been eliminated already).
-
-After normalization, the formula $\theta(\bar{x},\bar{y})$ contains atoms of three forms corresponding to 
-the predicates $\le$, $|$ and negation of $|$, respectively.
-The problem will be discussed in 2 cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
-corresponding to \textbf{QE-exp-ineq} or \textbf{QE-exp-div}.
-
-We first give the whole procedure of \textbf{QE-exp},
-and then provide the details of the two sub-routines.
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{$x_i$ and $\theta$, $x_i$ is larger than $x_0$ to $x_{i-1}$
-    and occurs exponentially in $\theta$} 
-    \KwOut{a formula equivalent to $\theta$ where $x_i$ occurs linearly}
-
-    \While{there is an atom $\tau(x_i,\bar{x},\bar{y})$ contains $2^{x_i}$}
-    {
-        $\Psi:= \textit{False}$\;
-        \eIf{$\tau$ is of the form $a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$}
-        { 
-            \tcp{ \textbf{QE-exp-ineq} case}
-            $A :=  \sum_{j=0}^{i-1}|a_j|$, $B:= b_i + \sum_{j=0}^{i-1}|b_j|$\;
-            $B':= 2(l_2(B)+3)$, $\delta:=  l_2(A)+3$\;
-            If $a_i>0$ then $\alpha := l_2(t(y))-l_2(a_i)$
-            else $\alpha := l_2(-t(y))-l_2(-a_i)$\;
-
-            $\rho_1 :=  (x_i \le \alpha -1 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$
-
-            $\quad \vee ( x_i \le \alpha -1 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta} (x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))$\;
-
-            $\rho_2 :=  (x_i = \alpha) \wedge \tau[\alpha/x_i]$\;
-            $\rho_3 :=  (x_i = \alpha+1) \wedge \tau[\alpha+1/x_i]$\;
-            $\rho_4 :=  (x_i \ge \alpha+2 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$
-
-            $\quad \vee (x_i \ge \alpha +2 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta}(x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))
-            $\;
-            \eIf{$a_i>0$}
-            {
-                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
-                \vee [x_i \le \alpha -1 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$
-            }
-            {
-                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
-                \vee [x_i \ge \alpha +2 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$    
-            }
-            }
-        {
-            \tcp{ \textbf{QE-exp-div} case, the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
-            \tcp{ let $d=2^r d_0$ where $d_0$ is odd}
-            $\rho_5 := \bigvee_{p=0}^{r-1} [\tau(p,\bar{x},\bar{y})\wedge x=p]$\;
-            $\rho_6 := \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
-            +  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r]
-            $\;
-            $\Psi :=  \rho_5 \vee \rho_6$
-        }
-        replace $\tau$ by $\Psi$ in $\theta$\;
-    }
-    \caption{QE-exp}
-\end{algorithm}
-
-\subsubsection{QE-exp-ineq}
-
-This algorithm deals with the case where $\tau$ is an inequality of the form 
-$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
-
-We now try to eliminate $2^{x_i}$ in $\tau$, the idea is to find a bound for $x_i$, 
-either by constants or by other variables.
-We will prove the following theorem.
-
-\begin{theorem} \label{thm:exp-ineq}
-Assume that $x_i$ is the maximal variable in $\bar{x}$ according to the given order. 
-Given an inequality $\tau(x_i,\bar{x},\bar{y})$ of the form %, suppose $x_i$ occurs in an exponential term, that is 
-$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
-with $a_i \neq 0$, 
-let $A\Def \sum_{j=0}^{i-1}|a_j|$, 
-$B\Def |b_i| + \sum_{j=0}^{i-1}|b_j|$, 
-$B'\Def 2(l_2(B)+3)$,
-$\delta\Def  l_2(A)+3$, then 
-\begin{itemize}
-    \item if $a_i > 0$, let $\alpha \Def l_2(t(y))-l_2(a_i)$
-    \begin{itemize}
-        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
-        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
-    \end{itemize}
-    \item if $a_i < 0$, let $\alpha \Def l_2(-t(y))-l_2(-(a_i))$
-    \begin{itemize}
-        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
-        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
-    \end{itemize}
-\end{itemize}
-\end{theorem}
-
-Before proving \textbf{Theorem 1}, we need  the following lemma to estimate linear terms.
-
-\begin{lemma} \label{lem:1}
-For any $n,m\in \mathbb{N}$, if $n\ge m\ge 1$ and $x\ge 2(l_2(n)-l_2(m)+1)$, then 
-$nx\le m2^x$ holds.
-\end{lemma}
-
-\begin{proof}
-First we can prove that for any $N\in \mathbb{N}$, $x\ge 2N \implies 2^N x\le 2^x$.
-Let $N\Def l_2(n)-l_2(m)+1$, then we have $x\ge  2N \implies n x\le 2\lambda_2(n) x\le 2^N \lambda_2(m) x \le m 2^x $.
-\qed
-\end{proof}
-
-Then we give the proof for Theorem~\ref{thm:exp-ineq}.
-\begin{proof}
-We only prove for the $a_i > 0$ case, the other case is analogous. 
-The goal is to find a bound for $x_i$ such that 
-the values of the atoms containing $x_i$ keep constant when $x_i$ is greater than the bound.
-
-Note that $x_i$ is the largest among $\bar{x}$. 
-suppose $x_i > x_{i-1} + \delta$, and let 
- $\delta \Def  l_2(A)+3$, then 
- \begin{equation} 
-   2^{-\delta}A = \frac{A}{8\lambda_2(A)}\le \frac{1}{4}.   \label{eq:thm-ineq-1}
- \end{equation}
- When $x_i \ge B' = 2(l_2(4B)-l_2(1)+1)$, according to Lemma~\ref{lem:1}, 
-\begin{equation} 
-    4 B x_i \le 2^{x_i}. \label{eq:thm-ineq-2}
-\end{equation}
-
-When $x_i \ge \alpha+2$,
-\begin{align}
-    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
-    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
-     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
-    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
-    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
-    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
-      (\mbox{ Def. of } \alpha) & \notag \\
-     & \ge  t(\bar{y}) & \notag 
-\end{align}
-So we conclude that %when $x_i \ge \alpha+2,x_i\ge B',x_i \ge x_{i-1}+\delta$,
-$\tau(x_i,\bar{x},\bar{y})$ keeps \textit{false} in this case.
-
-%\znj{I think there are some mistakes in the above proof, pls check it!}
-When $x_i \le \alpha-1$, similarly we have
-\begin{align}
-    \tau(x_i,\bar{x},\bar{y})\notag 
-    & \le  a_i 2^{x_i} + 2^{x_i -\delta}A  + Bx_i &  \notag \\
-    & \le  2^{x_i}(a_i +\frac{1}{4} +\frac{1}{4}) &(\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}}) \notag \\
-   & \le  2\lambda_2(a_i)\frac{\lambda_2(t(\bar{y}))}{2\lambda_2(a_i)} &  (\mbox{ Def. of } \alpha) \notag \\
-   &  \le  t(\bar{y}) & \notag 
-\end{align}
-which indicates that when  $x_i \le \alpha-1$, $\tau(x_i,\bar{x},\bar{y})$ keeps \textit{true}. \qed 
-\end{proof}
-
-\subsubsection{QE-exp-div}
-
-If $x_i$ occurs exponentially in an atom $\tau(x_i,\bar{x},\bar{y})$ of the form
-$$\tau(x_i,\bar{x},\bar{y}) \Def  d| a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k+t(\bar{y}) \quad (a_i \neq 0)\,,$$
-the algorithm \textbf{QE-exp-div} outputs an equivalent formula without $2^{x_i}$ terms.
-The idea is that $a_i 2^x$ modulo $d$ is a periodic function when $x$ is large enough and the period can be computed.
-
-Let $d = 2^rd_0$, $d_0$ is an odd natural number. 
-According to Euler's Theorem, 
-$gcd(2,d_0)=1$ implies
-$2^{\phi(d_0)} \text{mod}\ d_0 = 1$,
-where $\phi$ is the Euler function.
-Consider function $f(n)=(2^n\ \text{mod} \ d)$,
-when $n\ge r$, 
-$f(n)$ becomes a periodic function and its period is a divisor of $\phi(d_0)$ because
-\begin{align}
-    f(n+\phi(d_0)) 
-    &=2^{n+\phi(d_0)}\ \text{mod}\ d &\notag\\
-    &=2^n\cdot 2^{\phi(d_0)}\ \text{mod}\ d &\notag\\
-    &=2^n\cdot (k d_0 + 1) \ \text{mod}\ d &
-    (\mbox{assume } 2^{\phi(d_0)}=k d_0 +1) \notag \\
-    &= 2^n \ \text{mod}\ d &
-    (\mbox{when } n\ge r, 2^n d_0\ \text{mod}\ d = 0 )\notag\\
-    &= f(n) \notag
-\end{align}
-
-When $x_i\le r-1$, we just enumerate all possible value of $x_i$, i.e., 
-$$\rho_5 \Def \bigvee_{p=0}^{r-1} \tau(p,\bar{x},\bar{y})\,.$$
-When $x_i\ge r$, $\tau(x_i,\bar{x},\bar{y})$ is equivalent to
-$$
-\rho_6 \Def \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
-+  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r] \,.
-$$
-            Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.
-
-%\znj{\textbf{Solved}.More explanation or a formal proof is needed. }
-
-\subsection{QE-linear}
-
-After \textbf{QE-exp}, 
-we obtain a formula (still denoted by $\theta(x_i,\bar{x},\bar{y})$) without $2^{x_i}$ terms,
-i.e., $x_i$ occurs only linearly.
-Now we wish to construct a formula without $x_i$ equivalent to $\exists x_i.\theta(x_i,\bar{x},\bar{y})$.
-This can be done by 
-following Cooper's QE algorithm for PA,
-treating all $x_j (j<i)$ as free variables (like $\bar{y}$).
-The procedure contains 3 steps.
-
-Step 1: Put $\theta(x_i,\bar{x},\bar{y})$ in NNF form, replace atomic formulae containing symbols other than $\le,|$ by equivalent formulae only with $\le$, that is, $x=a$ by $x\le a \wedge a\le x$, $x< a$ by $x\le a+1$, $x\neq a$ by $x>a \vee x<a$,  $x\geq a$ by $-x \le -a$, and $x>a$ by $-x \le -a-1$.
-In inequality atoms, collect terms of $x_i$
-to one side and guarantee the coefficients of $x_i$ are positive.
-
-Step 2: Let $d$ be the least common multiple of all coefficients of $x_i$. 
-For atoms with $x_i$, multiply all terms by a factor so that the coefficient of $x_i$ is $d$.
-We introduce a fresh variable $x_i'$, replace all occurrence of $dx_i$ by $x_i'$,
-and denote the resulted formula by $\theta'(x_i')$.
-Note that $x_i'$ is a multiple of $d$, 
-so we set $\theta'=\theta'\wedge d|x'$.
-
-We classify all atoms in $\theta$ into the following three sets:  
-$L$ denotes all atoms of the form $t_l(\bar{x},\bar{y})\le x_i'$,
-$U$ denotes all atoms of the form $x_i'\le t_u(\bar{x},\bar{y})$,
-$M$ denotes all atoms of the form $k_m|x_i'+t_m(\bar{x},\bar{y})$ and their negations,
-where $t(\bar{x},\bar{y})$ with a subscript is any term of $\bar{x}$ and $\bar{y}$ and 
-$k_m \in \mathbb{N}$ for $m\in M$. % are integers in $$. 
-%Then all atoms with $x_i'$ are of one of the types $L,U,M$.
-
-Step 3: Let $\delta$ be the least common multiple of $\{k_m \mid m\in M\}$.
-Construct $\theta'_{+\infty}(x_i')$ from $\theta'(x_i')$ by  
- replacing $\textit{true}$ for all atoms in $L$,
-and $\textit{false}$ for all atoms in  $U$.
-%We use $\theta'[t]$ to denote replacing all $x_i'$ by a term $t$. 
-Then $\exists x_i'.\theta'(x_i')$
-is equivalent to
-
-$$\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee 
-\bigvee_{j=0}^{\delta-1} \bigvee_{u\in U} \theta'[t_u(\bar{x},\bar{y})-j/x_i']$$
-
-The first disjunction $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i']$ corresponds to the case where $x_i'$ is large enough 
-so that all atoms in $L$ are \textit{true} and all atoms in  $U$ are \textit{false}. Thus $\theta'_{+\infty}$  contains only divisibility literals.
-If there is a number $n (0\le n \le \delta-1)$ such that $ \theta'_{+\infty}[n/x_i']$ is evaluated to be \textit{true},
-then for every $\lambda\in \mathbb{N}$,
-$\theta'_{+\infty}[n+\lambda\delta/x_i']$ is also evaluated to be \textit{true}.
-Hence, there exists $x_i'$ large enough to satisfy $\theta'$.
-
-The second disjunction corresponds to the case that for 
-some $u\in U, x_i'\le t_u(\bar{x},\bar{y})$ holds.
-In this case, 
-select the minimal $t_u(\bar{x},\bar{y})$
-from atoms in $U$ that holds,
-then there exists a solution for $x_i'$ such that $x_i'$ is in the interval $[t_u(\bar{x},\bar{y})-\delta+1,t_u(\bar{x},\bar{y})]$.
-
-Here we describe Cooper's algorithm in pseudo code.
-We will use $\theta\{\tau'/\tau\}$ to denote substitute all atoms $\tau$ with $\tau'$,
-distinguished from term substitution $\theta[t'/t]$.
-
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{$x_i,\theta(x_i,\bar{x},\bar{y})$,
-    where $x_i$ occurs linearly} 
-    \KwOut{A equivalent formula without $x_i$}
-
-    Collect terms of $x_i$ in each atom\;
-    Let $d$ to be the least common multiple of all coefficients of $x_i$\;
-    Multiply each atom by a factor so that coefficient of $x_i $ is $d$\;
-    Let $\theta'(x_i'):= \theta[x_i'/dx_i]$\;
-    $\theta'(x_i') := \theta'(x_i') \wedge d| x_i'$\;
-    \tcp{ Now the atoms have 3 different forms}
-    \tcp{ $t_l(\bar{x},\bar{y})\le x_i'$, $x_i' \le t_u(\bar{x},\bar{y})$, $k_m|x_i'+t_m(\bar{x},\bar{y})$(or its negation)}
-    \tcp{ where $l\in L,u\in U,m\in M$ are the index sets}
-    $\delta :=  lcm\{k_m|m\in M\}$\;
-    \For{all atom $\tau_l \in L$}
-    {
-        $\theta'_{+\infty}:=  \theta'\{\textit{true}/\tau_l\}$
-    }
-    \For{all atom $\tau_u \in U$}
-    {
-        $\theta'_{+\infty}:=  \theta'_{+\infty}\{\textit{false}/\tau_u\}$
-    }
-    output $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee \bigvee_{j=0}^{\delta-1} \bigvee_{u\in U}\theta'[t_u(\bar{x},\bar{y})-j/x_i']$
-    \caption{QE-linear (Cooper's QE algorithm for PA)}
-\end{algorithm}
-
-%\znj{\textbf{Solved, missing $\bigvee$ for $\theta_{+\infty}$}.Pls check whether the output is correct or not, my intuition is not correct.} 
-% 之前的算法
-
-\subsection{Compared with Francoise's Origin Algorithm}
-
-Francoise's algorithm \cite{Francoise} has four steps,
-corresponding to the four steps of our algorithm respectively.
-So we will use the names of our steps to refer to them. 
-The divisible predicate $n|x$ is not included in his theory,
-instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
-The \textbf{Normalization} and \textbf{QE-with-order} steps in two algorithms are similar,
-with some minor differences due to the setting of the theory.
-The main differences between the two algorithms lies in the following aspects.
-
-In \textbf{QE-exp},
-we adopt the same idea to find sufficient conditions
-and necessary conditions for the formula to holds.
-But our strategy for choosing parameters are easier to understand and use.
-A flaw lies in the $a_i<0$ case in Francoise's algorithm 
-\footnote{
-In the $a_i<0$ case of \textbf{QE-exp-ineq},
-Francoise's algorithm assumes falsely $t(y)\ge 0$.
-The reason for this mistake might be the confusion of 
-two subtraction operators.
-The example mentioned in \ref{sec:3.2} shows this case.
-},
-our \textbf{QE-exp-ineq} corrects this flaw by 
-treating $a_i<0$ case similarly to the $a_i>0$ case.
-
-\textbf{QE-linear} in Francoise's algorithm in some sense includes our
-\textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
-In this step,
-Francoise uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
-We instead invoke the QE algorithm of PA without referring to permutation groups, 
-other improved QE algorithm can also be used to simplify this step.
-
-Francoise's algorithm has more restrictions on the form
-of the formula.
-Before \textbf{QE-linear} and \textbf{QE-exp} step,
-it needs to transform the formula into disjunction normal form (DNF)
-and deals with the basic case when input formula $\theta$ is 
-a conjunction of atoms. 
-It is well known that transforming an arbitrary formula into 
-DNF is costly and the length of the formula may increase exponentially.
-So our algorithm remove these limitations and assume no special forms of
-the formula other than NNF.
-
-\section{Complexity Analysis and Case Study}
-In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
-\subsection{Complexity Analysis}
-
-In this part, we estimate the complexity of our algorithm.
-It can be divided into two part, 
-the first part convert a string constraint $\phi$ to a PA formula with exponential functions,
-and the second part is the QE procedure on the obtained formula.
-
-Given a finite alphabet $\Sigma$ and $\alpha=\langle p,q \rangle$, the size of $\Sigma$ is denoted by
-$|\Sigma|$. 
-The generic $\alpha$-flat language contains $O(|\Sigma|^{pq})$ $\alpha$-flat languages
-
-\paragraph{Complexity for QE}
-In \textbf{QE-with-order},
-it reduces the origin problem to $(n+1)!$ sub-cases,
-and for each case,
-we recursively invoke \textbf{QE-exp} and \textbf{QE-linear}.
-So the key is to analysis the inner \textit{for}
-loop in \textbf{QE-with-order}.
-Note that if we omit the \textbf{QE-exp} step,
-the sub-cases are equivalent to Cooper's algorithm,
-so we adopt the idea of Oppen's analysis to analyze the complexity
-\cite{Oppen}.
-
-Apply Cooper's algorithm on the quantified PA formula 
-$Q_m x_m Q_{m-1}$ 
-$x_{m-1} ... Q_1 x_1. $
-$F(x_1,...,x_m)$,
-the algorithm repeats $m$ times to eliminate
-$x_1,...,x_m$ one by one.
-Let $F_k \Def Q_m x_m Q_{m-1} 
-x_{m-1} ... Q_{k+1} x_{k+1}. F_k'(x_{k+1},...,x_m)$
-be the formula produced after $k$th iterations of the algorithm. 
-Let $a_k$ denote the number of atoms in $F_k$,
-$c_k$ denote the number of distinct $\delta_i$ in atom $\delta_i | t$ (and its negation) plus the number of distinct coefficients of variables.
-$s_k$ denote the largest value of integer constant.
-Use $a_0,c_0,s_0$ to denote the initial value of $a_k,c_k,s_k$ respectively.
-The following theorem holds.
-
-\begin{theorem}
-for all $k: 1\le k\le m$, the following relations holds
-\begin{itemize}
-    \item $c_k \le {c_{k-1}}^4$
-    \item $s_k \le {s_{k-1}}^{4c_{k-1}}$
-    \item $a_k \le {a_{k-1}}^4 \cdot {s_{k-1}}^{2c_{k-1}}$
-\end{itemize}
-and from the above relations we have 
-\begin{itemize}
-    \item $c_k \le {{c_0}^4}^k$
-    \item $s_k \le {{{s_0}^{4c_0}}^4}^k$
-    \item $a_k \le {{a_0}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
-\end{itemize}
-\end{theorem}
-
-For each $k$,
-the space required to store $F_k$ 
-is $a_k \cdot (m+1) \cdot s_k \cdot q$,
-where $(m+1)$ is the maximum number of constants per atom and $q>1$ is some constant.
-Assume $m\le n, c\le n, a\le n, s\le n$, we obtain the deterministic space complexity is ${{2^2}^2}^{p n log_n}$, which is also the bound for deterministic time.
-
-In our algorithm,
-we use the same denotations and have the following theorem. 
-Note that each iteration corresponds to \textbf{QE-exp} and \textbf{QE-linear}.
-\begin{theorem}
-for all $k: 1\le k\le m$, the following relations holds
-\begin{itemize}
-    \item $c_k \le {c_{k-1}}^4$
-    \item $s_k \le {(n s_{k-1})}^{4c_{k-1}}$
-    \item $a_k \le {{(l_2(ns_{k-1})\cdot a_{k-1})}}^4 \cdot {(n s_{k-1})}^{2c_{k-1}}$
-\end{itemize}
-and from the above relations we have 
-\begin{itemize}
-    \item $c_k \le {{c_0}^4}^k$
-    \item $s_k \le {{{n}^{4c_0}}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
-    \item $a_k \le {{a_0}^4}^k \cdot 
-     {{{n}^{4c_0}}^4}^k
-     \cdot {{{s_0}^{4c_0}}^4}^k$
-\end{itemize}
-\end{theorem}
-
-WAIT TO CHECK!
-
-Again assume $m,c,a,s\le n$, which gives 
-$$
-\textit{space}\le q \cdot {n^4}^n \cdot {{n^{(4n)}}^4}^n \cdot
-{{n^{(4n)}}^4}^n \cdot (n+1) \cdot 
-\cdot {{n^{(4n)}}^4}^n \cdot {{n^{(4n)}}^4}^n
-\le 2^{2^{2^{p n \textit{log}n}}}
-$$
-
-Then the upper bound for Cooper's algorithm still holds for the inner \textit{for} loop of \textbf{QE-with-order}.
-When considering \textbf{Normalization} and permutations groups in \textbf{QE-with-order},
-since the super exponential term dominates the number of permutations $n^{O(n)}$,
-this upper bound still holds.
-
-\subsection{Example}
-
-We apply our algorithm to a simple example to 
-illustrate its function.
-Given $\Sigma = \{0,1\}$
-Suppose $\mathcal{A}$ is a finite state automaton
-that recognizes language $L = \{ w \mid \#w('1') \text{ is odd}\}$.
-Consider the following string constraint
-$$y\in L(\mathcal{A})\wedge 4\le  \parseInt(y) < 16\,.$$
-
-We need an abstraction parameter $alpha$ to restrict interpretations
-for $x$, for simplicity, set $\alpha\Def \langle 2,1 \rangle$.
-The generic $\langle 2,1 \rangle$-flat language is the union of 
-$\langle 2,1 \rangle$-flat languages $(11)^*,(10)^*,(01)^*,(00)^*,1^*,0^*$.
-Here we divide the problem into 6 sub-cases by restricting $y$
-to one of the $\langle 2,1 \rangle$-flat language.
-Another approach is to directly restrict $y$ to 
-the generic $\langle 2,1 \rangle$-flat language,
-at the cost of more introduced variables.
-Since the complexity is related to the number of variables,
-we wish it to be as little as possible.
-
-When $y$ is in language $1^*$, assume $y = 1^x$ where $x$
-is a fresh integer variable.
-Note that the Parikh image of $L(\mathcal{A})$ is $2|( x+1)$, 
-so the constraint is converted to a PA formula $\phi$
-$$\phi\Def \exists x. 2|(x+1) \wedge 5 \le 2^{x} < 17\,.$$
-
-Observe that when $x=3$ the formula holds,
-so we can easily deduce that $\phi$ is true.  
-By applying our algorithm,
-after \textbf{Normalization},
-$\phi$ is transformed into the form
-
-$$\exists x_0. 2|(x+1)\wedge  2^{x}\le 16 \wedge -2^{x}\le -5\,.$$
-
-Since there is only one variable, 
-\textbf{QE-with-order} can be omitted.
-We send $2^x\le 16$ and $-2^x \le -5$ to \textbf{QE-exp} and obtain
-$$2^x\le 16 \equiv x \le 4$$
-$$-2^x\le -5 \equiv 3\le x\,.$$ 
-Then we apply \textbf{QE-linear} on formula $\exists x_0. 2|(x+1) \wedge x\le 4 \wedge 3\le x$ and get
-$$\bigvee_{j=0}^1(2| j+1 \wedge \textit{false} \wedge \textit{true})\vee \bigvee_{j=0}^1 (2|4-j+i\wedge 4-j\le 4 \wedge 3\le 4-j) \,.$$
-which evaluates \textit{true} and we can deduce $j=1\implies x=3$, so $y$ is string $111$.
-
-Other sub-cases are similar. However, when the abstraction parameter becomes larger,
-say, $\alpha =\langle 2,2 \rangle$,
-the complexity will grow rapidly,
-because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.
-
-\section{Conclusion and Future Works}
+\input{7_conclusion.tex}
 

### 2021-05-07 FMCAD2021/1_intro.tex
restructure the paper

solve string constraint is hard

strategy: do unsat and sat separately

strategy: using different procedure to (dis)prove validaity

for disprove validatity, there are two approches, first is bound string length

cannot handle $x.y \neq z  \wedge |x| > 2000$

more recent approach is flattening

it is known that word equaltion + flat regular constraints + len constraints
is decidable

it was unknown that whether word equations + flat regular constraints + len constraints + parseInt is decidable

### 2021-05-07 FMCAD2021/2_preliminary.tex
restructure the paper

In this section, we introduce some basic concepts and theories that will be used later. 
\subsection{Basic Concepts}

\paragraph{Sets and Strings}
We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the set of natural numbers and integers, respectively. 
$\mathbb{N}^+$ stands for the set of non-zero natural numbers. 
Let $\Sigma$ be a finite alphabet,
a string $w$ over $\Sigma$ is a sequence $a_1....a_n$ of characters from $\Sigma$.
Empty string is denoted by $\epsilon$.
$\Sigma^*$ denotes the set of all finite strings over $\Sigma$,
and $\Sigma_{\epsilon}$ stands for $\Sigma \cup \{\epsilon\}$.
For any string $w_1, w_2\in \Sigma^*$, 
we use $|w_1|$ to denote the length of $w_1$,
and $w_1\cdot w_2$ to denote the concatenation of $w_1$ and $w_2$.
A language $L$ over $\Sigma$ is a subset of $\Sigma^*$.

There are two types of variables in string constraints,
i.e., $X$, a set of string variables ranged over $\Sigma^*$,  
and $Z$,  a set of integer variables ranged over $\mathbb{Z}$.
As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
essentially a pair of two mappings 
$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.

\paragraph{Finite State Automata}
A Finite State Automaton is a tuple 
$\mathcal{A}=\langle Q,\Sigma,\Delta,$ $q_{\textit{init}},$ $q_{\textit{acc}}\rangle$, 
where $Q$ is a finite set of states, 
$\Sigma$ is the given alphabet,
$\Delta\subseteq Q\times \Sigma_\epsilon\times Q$ 
defines the transition relations in $\mathcal{A}$.
$q_{\textit{init}},q_{\textit{acc}}\in Q$ is the initial state and accepting state. 
A sequence $q_0 \langle a_1 \rangle  q_1 ... \langle a_n\rangle q_n$ is called accepting if $q_0 = q_{\textit{init}}$, $q_n = q_{\textit{acc}}$ and $q_{i-1}\langle a_i \rangle q_i \in \Delta$ for $1\le i\le n$.

\paragraph{Presburger Arithmetic} \label{PA}
The Presburger Arithmetic (PA) is a first order theory over signature 
$\Sigma_\mathbb{N}\Def  \{0,1,+,=\}$, where $0,1$ are constants, 
$+$ is a binary function and $=$ is a binary predicate.

PA can be axiomatized by the following axioms \cite{PA}

\begin{itemize}
    \item $\forall x, \neg (x+1=0)$
    \item $\forall x \forall y. x+1=y+1 \to x=y$
    \item $F(0) \wedge (\forall x. F(x)\to F(x+1)) \to \forall x. F(x)$ 
    \item $\forall x. x+0=x$
    \item $\forall x \forall y. x+(y+1)=(x+y)+1$
\end{itemize}

Given the domain $\mathbb{N}$,
the standard interpretation of PA interprets 
$0,1$ to $0_\mathbb{N},1_\mathbb{N}\in \mathbb{N}$
and $+,=$ to addition and equality over $\mathbb{N}$.
We call a PA formula without quantifiers a quantifier-free PA formula.

PA is a decidable theory, 
and the complexity of decidability is related to 
the number and locations of quantifiers.
Generally, 
the upper bound (on deterministic time and space) 
for deciding a formula of length $n$ is $2^{2^{2^{p n log(n)}}}$,
where $p>1$ is a constant\cite{Oppen69}.

\paragraph{Parikh Image}
Given an alphabet $\Sigma$ and a string $w\in \Sigma^*$, 
we define the set of Parikh variables 
$\Sigma^\# \Def \{a^\# \mid a\in \Sigma\}$.
The Parikh image of $w$ is a function 
$\mathbb{P}(w): \Sigma^\# \mapsto \mathbb{N}$,
which maps each symbol $a^\#\in \Sigma^\#$ to the number of occurrences
of $a$ in $w$.
For example, let $w\Def aabba$,
then $\mathbb{P}(w)(a^\#)=3,\mathbb{P}(w)(b^\#)=2$.

For a language $L\subseteq \Sigma^*$, 
define the Parikh image of $L$ to be 
$\mathbb{P}(L)\Def \{\mathbb{P}(w) | w\in L\}$.
We say a language $L$ is \emph{Parikh-definable} 
if $\mathbb{P}(L)$ can be characterized by a quantifier-free PA formula over $\Sigma^\#$, 
where $a^\#$ in the formula encodes the number of occurrences of $a$.
It is well known that
any context-free language (therefore regular language) 
is Parikh definable \cite{Parikh66}.

\paragraph{String Terms}
Given a finite alphabet $\Sigma$ and 
a finite set $X$ of string variables over $\Sigma^*$,
we define the set of terms $\textit{Terms}(\Sigma,X)$ 
to be the smallest set satisfying
\begin{itemize}
    \item[1] $\Sigma\cup \{\epsilon\} \cup X \subseteq \textit{Terms}(\Sigma,X)$;
    \item[2] if $t_1,t_2\in \textit{Terms}(\Sigma,X)$, then $t_1 \cdot t_2 \in \textit{Terms}(\Sigma,X)$.
\end{itemize}

We extend $I_X$ to $\textit{Terms}(\Sigma,X)$ by letting $I_X(\epsilon)=\epsilon$, 
for $a\in \Sigma, I_X(a)=a$,
and $I_X(t_1\cdot t_2)= I_X(t_1)\cdot I_X(t_2)$.

\paragraph{String Constraints} \label{par: string constraints}
Given a constraint $\phi$ and an interpretation $I$,
$I\models \phi$ denotes that $I$ satisfies $\phi$,
and $I$ is called a \emph{model} of $\phi$.
We use $\lVert \phi\rVert$ to denote the set of all models of $\phi$.

We define the following three forms of atomic string constraints:
\begin{itemize}
\item An equality constraint $\phi_e$ is of the form 
$t_1 = t_2$, where $t_1,t_2\in \textit{Terms}(\Sigma,X)$.
We define $\lVert \phi_e \rVert = \{I\mid I(t_1)=I(t_2)\}$.
Inequality constraints can be defined analogously.

\item A regular constraint $\phi_r$ is of the form 
$x\in L(\mathcal{A})$,
where $x\in \mathbb{X}$ and $\mathcal{A}$ is a finite state automaton.
We define $\lVert \phi_r \rVert = \{I\mid I(x)\in L(\mathcal{A})\}$.

\item A length constraint $\phi_l$ is a linear constraint over 
$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
where $|\cdot |$ is the length function.
We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
\end{itemize}

In Section~\ref{sec:string-solving},  
we will introduce a new form of atomic string constraints, i.e., \emph{string-number conversion constraints}.
A string constraint is a Boolean combination of atomic string constraints possibly with quantifications over $X\cup Z$. Other notions are same as in the first-order logic.

Giving a string constraint $\Psi$,
the problem of string constraints solving is to decide 
whether $\lVert \Psi \rVert$ is empty,
if not, to compute an interpretation $I$ that satisfies $\Psi$.

\subsection{Flat Automata and Flat Languages}

Given a string constraint $\Psi$,
the general problem of deciding whether $\lVert \Psi \rVert$ is empty is undecidable.
However, 
the problem becomes decidable when certain restrictions are imposed.
One of the restriction is by flat automata and flat languages, 
defined below.

\paragraph{Flat Languages and Automata}
For a fixed alphabet $\Sigma$,
we say a language $L$ over $\Sigma$ to be \emph{$\langle p,q \rangle$-flat} if 
there exist strings $w_1,...,w_q \in \Sigma^*$ such that
$|w_i|\le p$ for all $i:1\le i \le q$ 
and $L = (w_1)^*(w_2)^*...(w_q)^*$. 
We use $\alpha$ to denote $\langle p,q \rangle$, 
and call it the \emph{abstraction parameter} of $L$.
Intuitively,
a flat language with abstraction parameter $\alpha = \langle p,q \rangle$
consists of $q$ loops and the length of each loop body is equal or less than $p$.
For example,
$L = (ab)^*(a)^*(bb)^*$ is a $\langle 2,3 \rangle$-flat language.

Flat automata are a special form of finite state automata that 
recognize flat languages.
Fix the abstraction parameter $\alpha=\langle p,q\rangle$,
a flat automaton consists of $q$ loops,
each loop is a circle of $p$ states.
Formally, 
an $\alpha$-flat automaton contains $p q$ states at most,
and we name the states from $1$ to $p q$,
$1$ is the initial state and $(p q - p + 1)$ is the accepting state.
We use $\cdot$ as a placeholder for some symbol in $\Sigma_\epsilon$,
the transition relations of state $i$ are defined as 
\begin{itemize}
    \item if $i\  \text{mod}\  p = 1$ and $i \neq pq-p+1$, then 
    $(i,\epsilon,i+p)\in \Delta$,
    $(i, \cdot ,i+1) \in \Delta$;
    \item if $i\  \text{mod}\  p = 0$, then 
    $(i,\cdot , i-p+1) \in \Delta$;
    \item otherwise, $(i,\cdot, i+1) \in \Delta$.
\end{itemize}

A $\langle 2,3 \rangle$-flat automaton 
that recognizes $L\Def  (ab)^*(a)^*(bb)^*$ is shown in 
figure (\ref{fig: FA}).

\begin{figure}[ht]
    \centering 
    \begin{tikzpicture}
        \node[state,           ] (4) {$4$};
        \node[state, left  of=4] (2) {$2$};
        \node[state, right of=4] (6) {$6$};
        \node[state, initial, below of=2] (1) {$1$};
        \node[state, below of=4] (3) {$3$};
        \node[state, accepting, right of=3] (5) {$5$};

        \draw 
        (1) edge[above] node{$\epsilon$} (3)
        (3) edge[above] node{$\epsilon$} (5)

        (1) edge[bend left,left] node{$a$} (2)
        (2) edge[bend left,right] node{$b$} (1)

        (3) edge[bend left,left] node{$a$} (4)
        (4) edge[bend left,right] node{$\epsilon$} (3)

        (5) edge[bend left,left] node{$b$} (6)
        (6) edge[bend left,right] node{$b$} (5);
    \end{tikzpicture}
    \caption{A $\langle 2,3 \rangle$-flat automaton 
that recognizes $L\Def  (ab)^*(a)^*(bb)^*$}
    \label{fig: FA}
\end{figure}

\paragraph{Generic Flat Languages and Automata}
Fix $\alpha = \langle p,q \rangle$,
we define the \emph{generic $\alpha$-flat language} is the union of all $\alpha$-flat languages, denoted by $\mathbb{F}(\alpha)$.
Now, we try to define an automaton that recognizes all $\alpha$-flat languages,
i.e., collects all behaviors of $\alpha$-flat automata.

Intuitively, 
the generic automaton is obtained by introducing a new alphabet (
a multi-set with $p q$ copies of the original alphabet) and 
adding more transitions (labels),
the states and the overall framework remain unchanged. 
In details, a generic $\alpha$-flat automaton is still a finite state automaton over
$\Sigma(\alpha)\Def \{a(i)| (a\in \Sigma_\epsilon) \wedge i\in \mathbb{N}:1\le i \le pq\}\cup \{\epsilon\}$.
The states are still named from $1$ to $pq$, 
the initial state is $1$ and the accepting state is $(pq-p+1)$.
The transition relations for state $i$ are defined as 
\begin{itemize}
    \item if $i\  \text{mod}\  p = 1$ and $i\neq pq-p+1$, then 
    $(i,\epsilon,i+p)\in \Delta$
    and $\forall s\in \Sigma_{\epsilon}. (i, s(i) ,i+1) \in \Delta$;
    \item if $i\  \text{mod}\  p = 0$, 
    $\forall s \in \Sigma_{\epsilon}. (i,s(i), i-p+1) \in \Delta$;
    \item otherwise, $\forall s \in \Sigma_{\epsilon}. (i,s(i), i+1) \in \Delta$.
\end{itemize}

For $\Sigma = \{a,b\}$, an example of generic $\langle 2,3 \rangle$-flat automaton is shown in figure (\ref{fig: GFA}).

\begin{figure}[ht]
    \centering 
    \begin{tikzpicture}
        \node[state,           ] (4) {$4$};
        \node[state, left  = 2cm of 4] (2) {$2$};
        \node[state, right = 2cm of 4] (6) {$6$};
        \node[state, initial, below of=2] (1) {$1$};
        \node[state, below of=4] (3) {$3$};
        \node[state, accepting, below of=6] (5) {$5$};

        \draw 
        (1) edge[above] node{$\epsilon$} (3)
        (3) edge[above] node{$\epsilon$} (5)

        (1) edge[bend left, pos =0.2 ,left] node{$a(1)$} (2)
        (1) edge[bend left, pos =0.5 ,left] node{$b(1)$} (2)
        (1) edge[bend left, pos =0.8 ,left] node{$\epsilon(1)$} (2)

        (2) edge[bend left, pos = 0.2 ,right] node{$\epsilon(2)$} (1)
        (2) edge[bend left, pos = 0.5 ,right] 
        node{$b(2)$} (1)
        (2) edge[bend left, pos = 0.8 ,right] 
        node{$a(2)$} (1)

        (3) edge[bend left, pos =0.2 ,left] node{$a(3)$} (4)
        (3) edge[bend left, pos =0.5 ,left] node{$b(3)$} (4)
        (3) edge[bend left, pos =0.8 ,left] node{$\epsilon(3)$} (4)

        (4) edge[bend left, pos = 0.2 ,right] node{$\epsilon(4)$} (3)
        (4) edge[bend left, pos = 0.5 ,right] 
        node{$b(4)$} (3)
        (4) edge[bend left, pos = 0.8 ,right] 
        node{$a(4)$} (3)

        (5) edge[bend left, pos =0.2 ,left] node{$a(5)$} (6)
        (5) edge[bend left, pos =0.5 ,left] node{$b(5)$} (6)
        (5) edge[bend left, pos =0.8 ,left] node{$\epsilon(5)$} (6)

        (6) edge[bend left, pos = 0.2 ,right] node{$\epsilon(6)$} (5)
        (6) edge[bend left, pos = 0.5 ,right] 
        node{$b(6)$} (5)
        (6) edge[bend left, pos = 0.8 ,right] 
        node{$a(6)$} (5);
    \end{tikzpicture}
    \caption{The generic $\langle 2,3 \rangle$-flat automaton}
    \label{fig: GFA}
\end{figure}

However,
the resulted automaton may accept languages that are not in $\mathbb{F}(\alpha)$,
because in different passes inside a loop, 
the automaton can choose different symbols between identical pairs. 
To avoid this problem,
we add a so-called purity condition on the accepting language of generic flat automata,
which is equivalent to intersecting the language of a generic flat automaton 
with a language that encodes the purity condition.

We say a string $w\in (\Sigma(\alpha))^*$ is pure if for all $i: 1\le i \le p q$,
and $a,b\in \Sigma$, 
$a\neq b \wedge \#w(a(i))>0$ implies $\#w(b(i))=0$.
Formally, the purity condition is defined by 
\begin{equation} \label{eq:purity}
 \bigwedge_{1\le i\le pq}\bigwedge_{a,b\in \Sigma, a\neq b} ({a(i)}^\#>0)\to ({b(i)}^\#=0)\, . 
\end{equation}

We denote the accepting language of the generic $\alpha$-flat automaton by $\mathbb{G}(\alpha)$.
Note that $\mathbb{G}(\alpha)$ is a language over $\Sigma_\alpha$,
but what we want is a language over $\Sigma$.
So we define a renaming function $R:\Sigma(\alpha)\mapsto \Sigma$ such that for all $a(i) \in \Sigma_\alpha, R(a(i))=a$,
and $R(\epsilon) = \epsilon$.
Define $\mathbb{G}'(\alpha) \, \Def \, \{R(w) \mid w\in \mathbb{G}(\alpha)\}$, 
for simplicity, we write
$\mathbb{G}'(\alpha)=R(\mathbb{G}(\alpha))$.

The important feature of generic flat autamata
is that every word $w\in \mathbb{G}(\alpha)$ is uniquely determined by its Parikh image $\mathbb{P}(w)$.

\subsection{Flattening}

The flattening technique was first introduced 
in \cite{Abdulla 2017}.
Fix an alphabet $\Sigma$ and an abstraction parameter $\alpha$, 
for a given atomic string constraint $\phi$, 
flattening $\phi$ with parameter $\alpha$ 
results in a new string constraint $\phi_\alpha$,
such that 
$R(\lVert \phi_\alpha \rVert) = \lVert \phi \rVert \cap \{I \mid \forall x\in X, I(x)\in \mathbb{G}'(\alpha)\}$,
where $R$ is the renaming function with its domain extended to interpretations in the normal manner.
Intuitively, it restricts $\phi$ to interpret 
string variables over $\mathbb{G}'(\alpha)$.

\cite{Abdulla 2017} discussed the flattening of basic 
string constraints including equality, integer, (regular) grammar and transducer constraints. 
For an atomic string constraint $\phi$,
the flattening $\phi_\alpha$ is still an atomic string constraint 
and is Parikh definable,
so its Parikh image can be expressed by a quantifier free PA formula.
Together with the purity condition,
we obtain an existential quantified PA formula $\rho$.
$\rho$ will be sent to a SMT solver,
if the solver returns an solution $\theta$,
then we can construct an interpretation for $\phi_\alpha$ from $\theta$,
otherwise it means $\phi$ is unsatistiable when 
string variables are interpreted  to $\alpha$-flat languages.

Take a regular constraint $\phi = x\in L(A)$ for example,
the flattening of $\phi$ resutls in a new finite state automaton $A'$ over $\Sigma(\alpha)$,
which encodes running $A$ ``in parallel" 
with the generic $\alpha$-flat automaton.
Let $\rho_1$ be the formula describing the Parikh image of $A'$,
which is a formula over variable sets $\Sigma(\alpha)^\#$.
Let $\rho_2$ be the purity condition \eqref{eq:purity}.
Then we obtain the PA formula $\exists (\Sigma(\alpha))^\#. \rho_1 \wedge \rho_2$.
In order to distinguish between different string variables,
we may replace $a(i)^\# \in \Sigma(\alpha)^\#$ by $(x,a(i))^\#$.

Since the structure of a flat automaton is decided 
by its abstraction parameter $\alpha$, 
a Counter-Example Guided Abstraction 
Refinement (CEGAR) framework is designed, which contains both an under- and an 
over-approximation module, to search the possible values of $\alpha$.
The termination for the overall algorithm is not guaranteed.

### 2021-05-07 FMCAD2021/3_flatten.tex
restructure the paper

The string-number conversion functions are commonly used functions
in most of programming languages,
for example,
\verb+parseInt()+ in Java and \verb+Int()+ in Python.
The functions usually take two parameters, 
a string over the agreed alphabet $\Sigma$
and an optional parameter denotes the radix.
They parse the string according to the rules indicated by the radix,
and return an integer denoted by the string.

From the view of string constraints,
string-number conversion functions give rise to a new form of string constraints
and are more expressive than length constraints.
So we consider extending string constraints with 
{\parseInt} function. 
As the general problem of string constraints is undecidable,
we still adopt the idea of flattening, 
i.e., variables are restricted to (generic) flat languages.
This problem has been investigated in \cite{POPL20},
which defined a special form of flat restriction (straight-line PFA) and 
proposed a heuristic search method.

In this section, 
we describe the problem of interest first, and then 
present an reduction from the problem of solving flat string constraints with {\parseInt} function
to the decidability problem of Presburger Arithmetic with exponential functions.
Hence, we identify a decidable subset of string constraints, which is the largest one with decidability so far to the best of our knowledge.

\subsection{String-Number Conversion Function}

Commonly, 
{\parseInt} function takes a string representation of a decimal
number and returns an integer.
Since we focus on the decidability,
we define a binary version of {\parseInt},
which takes a binary string and returns a decimal integer number.
For example,
${\parseInt}('111')=7$. 
Clearly,  our decision procedure given in this paper 
can be adapted to string constraints with other string to number conversion function without substantial change. 
Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.

\begin{definition} Let $\Sigma_{\textit{num}}=\{'0','1'\}$,
 ${\parseInt}:\Sigma_{\textit{num}}^*\mapsto \mathbb{N}$ is recursively defined by:
    for $w\in \Sigma_{\textit{num}}^*$
    \begin{itemize}
        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
        \item if $|w|=1, {\parseInt}('w')=w_\mathbb{N}$;
        \item for $|w|\ge 2$ and $w=w_1 \cdot w_2$, 
        where $|w_1|, |w_2| \ge 1$, 
        ${\parseInt}(w) = {\parseInt}(w_1)2^{|w_2|}+{\parseInt}(w_2)$.
    \end{itemize} 
\end{definition}

Now, we introduce a new form of atomic string constraints: 
a {\parseInt} constraint $\phi$ is of the form 
$n \sim {\parseInt}(t)$,
where $n$ is an integer term, $\sim \in \{\le,<,=,>,\ge\}$ and $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$ is a string term.
$\lVert \phi\rVert \Def  \{I \mid I(n)\sim {\parseInt}(I(t))\}$.

In what follows, we only  consider the problem in the case when $t=x$ and $x$ is restricted to (generic) flat languages. 
%If $t$ is not a single variable, 
For the general case $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$, 
it can be reduced to this special case by induction on the structure of $t$. 
%we can reduce the problem by separating $t$ (corresponding to the $|w|>2$ case in definition).

Given an $\alpha$-flat language $L$,
we assume $\alpha=\langle p,q \rangle$ and $L=(w_1)^*...(w_q)^*$,
where $p,q$ and $w_i(1\le i \le q) $ are known.
We further assume that $x = (w_1)^{\beta_1} ... (w_q)^{\beta_q}$,
then we have
\begin{align}
    &{\parseInt}(x) \notag \\
    =&{\parseInt}((w_1)^{\beta_1} ... (w_q)^{\beta_q}) \notag \\
    =&{\parseInt}((w_1)^{\beta_1} ... (w_{q-1})^{\beta_{q-1}})\cdot 2^{\beta_q |w_q|}
    + {\parseInt}((w_q)^{\beta_q}) \label{parse}
\end{align}
(\ref{parse}) is a recursive expression.
So we only need to deal with the basic case ${\parseInt}((w_q)^{\beta_q})$, 
where $w_q \neq \epsilon$
\begin{align}
    {\parseInt}((w_q)^{\beta_q}) &= \sum_{i=0}^{\beta_q-1} {\parseInt}(w_q)\cdot 2^{|w_q|\cdot i}  \notag \\
    &={\parseInt}(w_q)\frac{2^{|w_q|\cdot \beta_q}-1}{2^{|w_q|}-1}
    \label{parseInt-basic}
\end{align}
In (\ref{parseInt-basic}), since $w_q$ and $|w_q|$ are known, 
they can be regarded as constants.
The only unknown variable is $\beta_q$.

Combine (\ref{parse}) and (\ref{parseInt-basic}),
the constraint $n={\parseInt}(x)$ can be expressed by an arithmetic expression with 
$n$ and $(\beta_1,...,\beta_q)$,
inevitably with exponential components.

Take $n={\parseInt}((11)^a(10)^b)$ for example.
\begin{align}
    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
\end{align}

So we have 
\begin{equation}
    3\cdot n = 3\cdot 2^{2a+2b}-3\cdot 2^{2b}+2\cdot 2^{2b}-2. \notag
\end{equation}

Observe the form of the above equation,
$a,b,n$ are integer variables and either occur in 
an exponential term or a linear term.
This is always the case,
so the problem can be reduced to 
the decidability of PA with exponential function.

When $x$ of {\parseInt} constraints is restricted to the
generic $\alpha$-flat language ($\alpha$ is fixed),
the (\ref{parse}) and (\ref{parseInt-basic}) still hold 
but $w_i(1\le i\le q)$ is known.
However,
by definition,
the generic $\alpha$-flat language is the finite union of all $\alpha$-flat languages,
so we can enumerate all possible values for $w_i$.
In this way,
the problem can still be reduced to the decidability of PA with exponential function (\paexp), i.e.,
\begin{theorem} \label{thm:string-parInt}
If {\paexp} is decidable, then the satisfiability (validity) of string constraints with {\parseInt} in which all string variables 
are ranged over flat strings is decidable. 
\end{theorem}

### 2021-05-07 FMCAD2021/4_dec_proc.tex
restructure the paper

\subsection{Decidability of \paexp} \label{sec:3.2}

For a first order theory $T$,
we say theory $T$ admits quantifier elimination (QE) if for any formula in $T$, 
there is a quantifier-free formula equivalent to it.
It is well-known that if a theory admits QE, 
then it is a decidable theory.

The formal definition of PA is given in section \ref{PA}.
Here we add the ordering predicate $\le$ into the signature,
which can be defined by $x \le y\, \Def \, \exists z. x+z=y$.
However, the theory PA so far does not admit QE, 
for example, consider the formula $\exists x.x = y + y$. 
We augment the theory with countable unary divisible predicates
$n|x$, where $n\in \mathbb{N}$, 
$n|x$ is true if and only if $x\ \text{mod}\ n=0$ holds.
This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.

We then introduce a theory {\paexp}, denoted by $(\mathbb{N},+,2^x)$, that we work on.
\begin{definition}
    Let $\mathcal{L}=\{0,1,+,\le,n \, |x(n\in \mathbb{N}),2^x,l_2(x)\}$, 
     $(\mathbb{N},+,2^x)$ be a $\mathcal{L}$-theory that has domain $\mathbb{N}$, where 
    \begin{itemize}
        \item  $2^x$ is interpreted to the exponential function of $2$ over $\mathbb{N}$; 
        \item interpretations of $0,1,+,\le,=$ are consistent with PA;
        \item for $n\ge 1$,$n|x$ holds iff $\exists y.x=ny$;
        \item $2^0=1$, for $n \ge 1, 2^n = 2^{n-1}+2^{n-1}$;
        We further assume that if $m,n\in \mathbb{N}, m\le n$, 
        then $2^{m-n} = 1$
        \item $l_2(0)=0$; for $n \ge 1,l_2(n) = y$ iff $2^y \le n < 2^{y+1}$;  $l_2(m-n) = 0$
        if $m,n\in \mathbb{N}$ and $m\le n$. 
    \end{itemize}
\end{definition}

$\lambda_2(x) = 2^{l_2(x)}$ can be defined by $l_2(x)$,
intuitively, $\lambda_2(x)$ means the maximal power of 2 that is not larger than $x$.
Then we have $\lambda_2(x) \le x \le 2\lambda_2(x)-1$,
which will be useful in our proof.

\begin{definition}
    For a strictly increasing function $f:\mathbb{N}\mapsto \mathbb{N}$, 
    $f$ is said to be compatible with addition 
    if for every $m\in \mathbb{N}^+$,  $f$ modulo $m$ is periodic, 
    and for any term $A(x)=\sum_{1\le i\le n} a_i f(x+b_i)$, 
    where $n \in \mathbb{N}^+, a_i ,b_i \in \mathbb{Z}$,
    one of the following holds:
    \begin{itemize}
        \item $A(x)$ is bounded; 
        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
        $\forall x. A(x+\Delta_A)\ge f(x)$;
        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
        $\forall x. -A(x+\Delta_A)\ge f(x)$.
    \end{itemize}
\end{definition}

%\znjSide{\textbf{Solved}. I think you need to explain the meanings of compatibility.}

Semenov proved that for any function $f$ that is compatible with addition, 
theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
by \cite{Semenov84}.
Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
Unfortunately, Point's algorithm is flawed, one problem 
 lies in the discussion to eliminate exponential terms. 
% making the algorithm fail to handle "$a_0<0$" case.
For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
the expected result should be $y\le 2^{10} =1024$,
but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
% 这里好像还有另一个与Francoise一起的作者

In the next section, within Point's framework, we give a revised algorithm
 to the decision problem of {\paexp} with 
both subtle improvements and crucial corrections, that 
provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to 
Theorem~\ref{thm:string-parInt}.

%\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
Based on Point's work \cite{Point86}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 
Since $\forall x. F = \neg \exists x. \neg F$, we further assume the quantifier $Q$ to be the existential quantifier, 
that is, $\exists x.\theta(x,\bar{y})$. 
For eliminating quantifiers in arbitrary formula, 
we apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.

We divide the whole QE algorithm into 4 steps.
The first step $\textbf{Normalization}$ can be viewed as a pre-processing step, which substitutes ``complex" terms like $2^{2^x}$, $l_2(3x+y)$ with ``simple" terms by introducing new variables $x_i$. Now we move forward to handle a ``simpler" formula, however, at the cost of more quantified variables.

The other three steps undertake the task to eliminate the introduced variables $x_i$ and the original variable $x$ (will be denoted by $x_0$) one by one. 
The main body of \textbf{QE-with-Order} step is a loop to enumerate all possible orders among quantified variables. 
In each iteration, according the given (decreasing) order (corresponding to a for loop), we invoke \textbf{QE-exp} or 
\textbf{QE-linear} to eliminate these variables one by one. 
%depending on  eliminates the maximal variable in the remaining $\bar{x}$. So it is necessary to specify an ordering for $\bar{x}$ at the beginning, as we will see, this is the main reason for the complexity of this algorithm.
During each iteration of the inner loop (the for loop), 
if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atomic formula, 
we will invoke \textbf{QE-exp} to produce a formula equivalent to the atom where $x_i$ occurs linearly;  
otherwise, if $x_i$ occurs linearly in the formula, \textbf{QE-linear} will be invoked to eliminate all occurrences of $x_i$, 
and this procedure is similar to the classic QE algorithm for PA.

A more detailed description is given below.

\subsection{Normalization}

In order to show $(\mathbb{N},+,2^x)$ admits QE , 
it is sufficient to show that any 1-existential formula $\exists x.\theta(x,\bar{y})$ indeed does, 
where $\theta(x,\bar{y})$ is a quantifier-free formula. 
However, the form of $\exists x.\theta(x,\bar{y})$ is unknown so it may contain terms difficult to handle such as $2^{3x+y+1}$ or $l_2(x)$.
The \textbf{Normalization} step simplifies these terms by introducing new variables, 
for example, $\exists x.2^{3x+y+1}>10$ is equivalent to
$$\exists x\exists x_1.2^{x_1}>10 \wedge x_1 =3x+y+1\,.$$

Logarithm functions are replaced by exponential functions, take $\exists x.l_2(x)>3$ for example, $l_2(x)$ is replaced by a new variable $x_1$, and we have  
$$\exists x\exists x_1. x_1 > 3 \wedge 2^{x_1}\le x \wedge x\le 2^{x_1+1}-1\,.$$

The \textbf{Normalization} step goes like this, first transform the given formula $\theta(x,\bar{y})$ into the Negation Normal Form (NNF).
Then $\theta(x,\bar{y})$ becomes a Boolean combination (with only $\wedge$ and $\vee$) of literals.

Make substitutions by introducing new variables according to the \textit{while} statements in the pseudo-code.For consistency, we rename the original $x$ to be $x_0$ and assume $n$ new quantified variables are introduced.

After introducing new variables and substitution, 
we obtain a formula where if an exponential term contains a quantified variable $x_i(0\le i\le n)$, the term should be $2^{x_i}$.
We then collect all terms with $x_i(0\le i\le n)$ together, 
it will be of the form $s(\bar{x})\Def \sum_{i=0}^{n} a_i 2^{x_i} + \sum_{i=0}^n b_i x_i$, where $a_i,b_i(0\le i \le n)$ are all constants.
Other terms including constants and terms of $\bar{y}$ are collected,  
denoted by $t(\bar{y})$. 
Since $\bar{y}$ are free variables, 
we will regard $t(\bar{y})$ as a constant.
Inequalities and equalities will all be transformed into $s(\bar{x})\le t(\bar{y})$, for example, $s(\bar{x}) = t(\bar{y})$ will be replaced by $s(\bar{x})\le t(\bar{y})\wedge t(\bar{y})\le s(\bar{x})$.

At the end, the resulted formula $\theta'$ only contains literals of the forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ and $\neg  (k|s(\bar{x})+t(\bar{y}))$.

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{1-existential formula $\exists x.\theta(x,\bar{y})$}
    \KwOut{n-existential formula $\exists \bar{x}.\theta'(\bar{x},\bar{y})$}

    $\theta' :=  \theta(x,\bar{y})$ \;
    Transform $\theta'$ into NNF\;
    \tcp{ i is used for counting the introduced variables }
    i = 1\; 
    \tcp{ $\bar{y}$ are regarded as constants} 
    \While{there is a term $l_2(t)$, $t$ is not a constant}
    {
        $\theta' :=  \theta'[x_i/l_2(t)] \wedge (2^{x_i}\le t) \wedge (t\le 2^{x_i+1}-1)$\;
        $i:= i+1$\;
    }
    \While{there is a term $2^t$, $t$ is not a variable $x_j(j<i)$ or a constant}
    {
        $\theta' :=  \theta'[2^{x_i}/2^t]\wedge (x_i\le t) \wedge (t \le x_i)$\;
        $i:= i+1$\;
    }
    $n:= i-1$\;
    \tcp{ Collect quantified variables $x_i$}
    Transform all atoms into forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ or $\neg  (k|s(\bar{x})+t(\bar{y}))$\;
    Return $\exists x_0,...,\exists x_n. \theta'$

    \caption{Normalization}
\end{algorithm}

\subsection{QE-with-order} 
During the \textbf{Normalization} step, 
we simplify the origin formula at the cost of more quantified variables. 
Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
and it has $n$ new variables $x_i$ ($x$ is denoted by $x_0$).
As a consequence, we need to eliminate all quantified variables one by one.

To the end, we denote the set of all orders among the $n+1$ variables by 
 $S_{n+1}$, and then enumerate these orders one by one in the outer \textit{for} loop. 
 For the considered order $\sigma$, we first add the ordering information to 
 the quantifier free formula, and then according to the 
 order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{QE-exp} first, and then invoking \textbf{QE-linear} (the inner \textit{for} loop). 
 In each iteration of the inner \textit{for} loop, 
if $x_{\sigma{(i)}}$ occurs in an exponential term in an atomic formula, 
\textbf{QE-exp} is invoked and a formula   equivalent to the atom is returned, in which $x_i$ occurs linearly;  
then $x_i$ occurs linearly in the formula, thus \textbf{QE-linear} is further invoked to eliminate all occurrences of $x_i$. 
The returned formula is the disjunction of all formulas resulted in all iterations of the outer \textit{for} loop.

\iffalse
put , We now specify an order of $\bar{x}$, and recursively eliminate the maximal element in the remaining $\bar{x}$. 
That is to say,
every execution of \textbf{QE-exp} and \textbf{QE-linear} will eliminate the largest element $x_i$ in $\bar{x}$.
However, as you may notice, if there is no clue about the order of $\bar{x}$, 
we will have to use the idea of permutation group and the number of sub-cases will blow up, in the worst case, $(n+1)!$ cases.

Here is how \textbf{QE-exp} and \textbf{QE-linear} are invoked in
\textbf{Ordering and QE}. 
\fi

%我们这里还是考虑用循环群来做，因为如果每一次只找到最大的$x_i$，那么递归的执行还是需要整个$\bar{x}$的顺序。但如果一开始就知道了，后面有些操作可以简化

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{a normalized $(n+1)$-existential formula $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
    where $\bar{x}=(x_0,...,x_n)$} 
    \KwOut{a equivalent quantifier free formula without $\bar{x}$}

    Let $S_{n+1}$ denote the group of permutations on $\{0,1,...,n\}$\;
    $\phi :=  \textit{False}$\;
    \For{each $\sigma \in S_{n+1}$ }
    {
        $\theta_{\sigma}:=  \theta(\bar{x},\bar{y})\wedge \bigwedge_{j=0}^{n-1}(x_{\sigma(j)}\le x_{\sigma(j+1)})$\;

        \tcp{recursively eliminate the maximal $x_i$ in $\bar{x}$}
        %\znj{How to effectively determine the maximal $x_i$?} 
        \For{i from n to 0}
        {
            \tcp{if $x_{\sigma(i)}$ occurs exponentially in $\theta_\sigma$}
            $\theta_{\sigma}:= \textbf{QE-exp}(x_{\sigma(i)},\theta_{\sigma})$\;
            \tcp{now $x_{\sigma(i)}$ occurs only linearly in $\theta_\sigma$}
            $\theta_{\sigma}:= \textbf{QE-linear}(x_{\sigma(i)},\theta_{\sigma})$\;
        }
        $\phi :=  \phi \vee \theta_{\sigma}$\;
    }
\iffalse
    \For{i from $0$ to $n$}
    {
        \tcp{ specify the maximal quantified variable}
        $\theta_i :=  \theta(\bar{x},\bar{y}) \wedge (\bigwedge_{0\le j \le n} x_i - x_j \ge 0)$\;
        \tcp{ eliminate $x_i$}
        \While{$x_i$ occurs in an exponential term in formula $\tau(x_i,\bar{x},\bar{y})$}{
            \tcp{ QE-exp outputs a formula equivalent with $\tau(x_i,\bar{x},\bar{y})$ where $x_i$ occurs linearly }
            $\theta_i :=  \theta_i[\text{QE-exp}(\tau(x_i,\bar{x},\bar{y}))/\tau(x_i,\bar{x},\bar{y})]$\;
        }   
        transform $\theta_i$ into disjunction normal form $\theta_i = \bigvee_s \theta_{i,s}$\;
        \For{ every $\theta_{i,s}$ }{
            $\theta'_{i,s} :=  \text{QE-linear}(x_i,\theta_{i,s})$
        }
        $\theta'_i =\bigvee_s \theta'_{i,s}$\;
    }
\fi
    output $\phi$
    \caption{QE-with-order}
\end{algorithm}
%\znj{I cannot understand why you need to do QE for all permutation $\sigma$? I think 
%one is enough.}

\subsection{QE-exp}

\textbf{QE-exp} and \textbf{QE-linear} are the most technical parts of our QE algorithm.
As mentioned, 
\textbf{QE-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as 
inputs, and  outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{QE-linear}.
For simplicity, we will abuse   $x_i$ for $x_{\sigma(i)}$, 
and  $\bar{x}$ for $(x_{\sigma(0)},...,x_{\sigma(i-1)})$ ( remind that $x_{\sigma(i+1)},...,x_n$ have been eliminated already).

After normalization, the formula $\theta(\bar{x},\bar{y})$ contains atoms of three forms corresponding to 
the predicates $\le$, $|$ and negation of $|$, respectively.
The problem will be discussed in 2 cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
corresponding to \textbf{QE-exp-ineq} or \textbf{QE-exp-div}.

We first give the whole procedure of \textbf{QE-exp},
and then provide the details of the two sub-routines.
\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{$x_i$ and $\theta$, $x_i$ is larger than $x_0$ to $x_{i-1}$
    and occurs exponentially in $\theta$} 
    \KwOut{a formula equivalent to $\theta$ where $x_i$ occurs linearly}

    \While{there is an atom $\tau(x_i,\bar{x},\bar{y})$ contains $2^{x_i}$}
    {
        $\Psi:= \textit{False}$\;
        \eIf{$\tau$ is of the form $a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$}
        { 
            \tcp{ \textbf{QE-exp-ineq} case}
            $A :=  \sum_{j=0}^{i-1}|a_j|$, $B:= b_i + \sum_{j=0}^{i-1}|b_j|$\;
            $B':= 2(l_2(B)+3)$, $\delta:=  l_2(A)+3$\;
            If $a_i>0$ then $\alpha := l_2(t(y))-l_2(a_i)$
            else $\alpha := l_2(-t(y))-l_2(-a_i)$\;

            $\rho_1 :=  (x_i \le \alpha -1 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$

            $\quad \vee ( x_i \le \alpha -1 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta} (x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))$\;

            $\rho_2 :=  (x_i = \alpha) \wedge \tau[\alpha/x_i]$\;
            $\rho_3 :=  (x_i = \alpha+1) \wedge \tau[\alpha+1/x_i]$\;
            $\rho_4 :=  (x_i \ge \alpha+2 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$

            $\quad \vee (x_i \ge \alpha +2 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta}(x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))
            $\;
            \eIf{$a_i>0$}
            {
                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
                \vee [x_i \le \alpha -1 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$
            }
            {
                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
                \vee [x_i \ge \alpha +2 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$    
            }
            }
        {
            \tcp{ \textbf{QE-exp-div} case, the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
            \tcp{ let $d=2^r d_0$ where $d_0$ is odd}
            $\rho_5 := \bigvee_{p=0}^{r-1} [\tau(p,\bar{x},\bar{y})\wedge x=p]$\;
            $\rho_6 := \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
            +  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r]
            $\;
            $\Psi :=  \rho_5 \vee \rho_6$
        }
        replace $\tau$ by $\Psi$ in $\theta$\;
    }
    \caption{QE-exp}
\end{algorithm}

\subsubsection{QE-exp-ineq}

This algorithm deals with the case where $\tau$ is an inequality of the form 
$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$

We now try to eliminate $2^{x_i}$ in $\tau$, the idea is to find a bound for $x_i$, 
either by constants or by other variables.
We will prove the following theorem.

\begin{theorem} \label{thm:exp-ineq}
Assume that $x_i$ is the maximal variable in $\bar{x}$ according to the given order. 
Given an inequality $\tau(x_i,\bar{x},\bar{y})$ of the form %, suppose $x_i$ occurs in an exponential term, that is 
$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
with $a_i \neq 0$, 
let $A\Def \sum_{j=0}^{i-1}|a_j|$, 
$B\Def |b_i| + \sum_{j=0}^{i-1}|b_j|$, 
$B'\Def 2(l_2(B)+3)$,
$\delta\Def  l_2(A)+3$, then 
\begin{itemize}
    \item if $a_i > 0$, let $\alpha \Def l_2(t(y))-l_2(a_i)$
    \begin{itemize}
        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
    \end{itemize}
    \item if $a_i < 0$, let $\alpha \Def l_2(-t(y))-l_2(-(a_i))$
    \begin{itemize}
        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
    \end{itemize}
\end{itemize}
\end{theorem}

Before proving \textbf{Theorem 1}, we need  the following lemma to estimate linear terms.

\begin{lemma} \label{lem:1}
For any $n,m\in \mathbb{N}$, if $n\ge m\ge 1$ and $x\ge 2(l_2(n)-l_2(m)+1)$, then 
$nx\le m2^x$ holds.
\end{lemma}

\begin{proof}
First we can prove that for any $N\in \mathbb{N}$, $x\ge 2N \implies 2^N x\le 2^x$. Let $N\Def l_2(n)-l_2(m)+1$, then we have $x\ge  2N \implies n x\le 2\lambda_2(n) x\le 2^N \lambda_2(m) x \le m 2^x $. \qed 
\end{proof}

Then we give the proof for Theorem~\ref{thm:exp-ineq}.
\begin{proof}
We only prove for the $a_i > 0$ case, the other case is analogous. 
The goal is to find a bound for $x_i$ such that 
the values of the atoms containing $x_i$ keep constant when $x_i$ is greater than the bound.

Note that $x_i$ is the largest among $\bar{x}$. 
suppose $x_i > x_{i-1} + \delta$, and let 
 $\delta \Def  l_2(A)+3$, then 
 \begin{equation} 
   2^{-\delta}A = \frac{A}{8\lambda_2(A)}\le \frac{1}{4}.   \label{eq:thm-ineq-1}
 \end{equation}
 When $x_i \ge B' = 2(l_2(4B)-l_2(1)+1)$, according to Lemma~\ref{lem:1}, 
\begin{equation} 
    4 B x_i \le 2^{x_i}. \label{eq:thm-ineq-2}
\end{equation}

When $x_i \ge \alpha+2$,
\begin{align}
    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
      (\mbox{ Def. of } \alpha) & \notag \\
     & \ge  t(\bar{y}) & \notag 
\end{align}
So we conclude that %when $x_i \ge \alpha+2,x_i\ge B',x_i \ge x_{i-1}+\delta$,
$\tau(x_i,\bar{x},\bar{y})$ keeps \textit{false} in this case.

%\znj{I think there are some mistakes in the above proof, pls check it!}
When $x_i \le \alpha-1$, similarly we have
\begin{align}
    \tau(x_i,\bar{x},\bar{y})\notag 
    & \le  a_i 2^{x_i} + 2^{x_i -\delta}A  + Bx_i &  \notag \\
    & \le  2^{x_i}(a_i +\frac{1}{4} +\frac{1}{4}) &(\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}}) \notag \\
   & \le  2\lambda_2(a_i)\frac{\lambda_2(t(\bar{y}))}{2\lambda_2(a_i)} &  (\mbox{ Def. of } \alpha) \notag \\
   &  \le  t(\bar{y}) & \notag 
\end{align}
which indicates that when  $x_i \le \alpha-1$, $\tau(x_i,\bar{x},\bar{y})$ keeps \textit{true}. \qed 
\end{proof}

\subsubsection{QE-exp-div}

If $x_i$ occurs exponentially in an atom $\tau(x_i,\bar{x},\bar{y})$ of the form
$$\tau(x_i,\bar{x},\bar{y}) \Def  d| a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k+t(\bar{y}) \quad (a_i \neq 0)\,,$$
the algorithm \textbf{QE-exp-div} outputs an equivalent formula without $2^{x_i}$ terms.
The idea is that $a_i 2^x$ modulo $d$ is a periodic function when $x$ is large enough and the period can be computed.

Let $d = 2^rd_0$, $d_0$ is an odd natural number. 
According to Euler's Theorem, 
$gcd(2,d_0)=1$ implies
$2^{\phi(d_0)} \text{mod}\ d_0 = 1$,
where $\phi$ is the Euler function.
Consider function $f(n)=(2^n\ \text{mod} \ d)$,
when $n\ge r$, 
$f(n)$ becomes a periodic function and its period is a divisor of $\phi(d_0)$ because
\begin{align}
    f(n+\phi(d_0)) 
    &=2^{n+\phi(d_0)}\ \text{mod}\ d &\notag\\
    &=2^n\cdot 2^{\phi(d_0)}\ \text{mod}\ d &\notag\\
    &=2^n\cdot (k d_0 + 1) \ \text{mod}\ d &
    (\mbox{assume } 2^{\phi(d_0)}=k d_0 +1) \notag \\
    &= 2^n \ \text{mod}\ d &
    (\mbox{when } n\ge r, 2^n d_0\ \text{mod}\ d = 0 )\notag\\
    &= f(n) \notag
\end{align}

When $x_i\le r-1$, we just enumerate all possible value of $x_i$, i.e., 
$$\rho_5 \Def \bigvee_{p=0}^{r-1} \tau(p,\bar{x},\bar{y})\,.$$
When $x_i\ge r$, $\tau(x_i,\bar{x},\bar{y})$ is equivalent to
$$
\rho_6 \Def \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
+  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r] \,.
$$
            Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.

%\znj{\textbf{Solved}.More explanation or a formal proof is needed. }

\subsection{QE-linear}

After \textbf{QE-exp}, 
we obtain a formula (still denoted by $\theta(x_i,\bar{x},\bar{y})$) without $2^{x_i}$ terms,
i.e., $x_i$ occurs only linearly.
Now we wish to construct a formula without $x_i$ equivalent to $\exists x_i.\theta(x_i,\bar{x},\bar{y})$.
This can be done by 
following Cooper's QE algorithm for PA,
treating all $x_j (j<i)$ as free variables (like $\bar{y}$).
The procedure contains 3 steps.

Step 1: Put $\theta(x_i,\bar{x},\bar{y})$ in NNF form, replace atomic formulae containing symbols other than $\le,|$ by equivalent formulae only with $\le$, that is, $x=a$ by $x\le a \wedge a\le x$, $x< a$ by $x\le a+1$, $x\neq a$ by $x>a \vee x<a$,  $x\geq a$ by $-x \le -a$, and $x>a$ by $-x \le -a-1$.
In inequality atoms, collect terms of $x_i$
to one side and guarantee the coefficients of $x_i$ are positive.

Step 2: Let $d$ be the least common multiple of all coefficients of $x_i$. 
For atoms with $x_i$, multiply all terms by a factor so that the coefficient of $x_i$ is $d$.
We introduce a fresh variable $x_i'$, replace all occurrence of $dx_i$ by $x_i'$,
and denote the resulted formula by $\theta'(x_i')$.
Note that $x_i'$ is a multiple of $d$, 
so we set $\theta'=\theta'\wedge d|x'$.

We classify all atoms in $\theta$ into the following three sets:  
$L$ denotes all atoms of the form $t_l(\bar{x},\bar{y})\le x_i'$,
$U$ denotes all atoms of the form $x_i'\le t_u(\bar{x},\bar{y})$,
$M$ denotes all atoms of the form $k_m|x_i'+t_m(\bar{x},\bar{y})$ and their negations,
where $t(\bar{x},\bar{y})$ with a subscript is any term of $\bar{x}$ and $\bar{y}$ and 
$k_m \in \mathbb{N}$ for $m\in M$. % are integers in $$. 
%Then all atoms with $x_i'$ are of one of the types $L,U,M$.

Step 3: Let $\delta$ be the least common multiple of $\{k_m \mid m\in M\}$.
Construct $\theta'_{+\infty}(x_i')$ from $\theta'(x_i')$ by  
 replacing $\textit{true}$ for all atoms in $L$,
and $\textit{false}$ for all atoms in  $U$.
%We use $\theta'[t]$ to denote replacing all $x_i'$ by a term $t$. 
Then $\exists x_i'.\theta'(x_i')$
is equivalent to

$$\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee 
\bigvee_{j=0}^{\delta-1} \bigvee_{u\in U} \theta'[t_u(\bar{x},\bar{y})-j/x_i']$$

The first disjunction $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i']$ corresponds to the case where $x_i'$ is large enough 
so that all atoms in $L$ are \textit{true} and all atoms in  $U$ are \textit{false}. Thus $\theta'_{+\infty}$  contains only divisibility literals.
If there is a number $n (0\le n \le \delta-1)$ such that $ \theta'_{+\infty}[n/x_i']$ is evaluated to be \textit{true},
then for every $\lambda\in \mathbb{N}$,
$\theta'_{+\infty}[n+\lambda\delta/x_i']$ is also evaluated to be \textit{true}.
Hence, there exists $x_i'$ large enough to satisfy $\theta'$.

The second disjunction corresponds to the case that for 
some $u\in U, x_i'\le t_u(\bar{x},\bar{y})$ holds.
In this case, 
select the minimal $t_u(\bar{x},\bar{y})$
from atoms in $U$ that holds,
then there exists a solution for $x_i'$ such that $x_i'$ is in the interval $[t_u(\bar{x},\bar{y})-\delta+1,t_u(\bar{x},\bar{y})]$.

Here we describe Cooper's algorithm in pseudo code.
We will use $\theta\{\tau'/\tau\}$ to denote substitute all atoms $\tau$ with $\tau'$,
distinguished from term substitution $\theta[t'/t]$.

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{$x_i,\theta(x_i,\bar{x},\bar{y})$,
    where $x_i$ occurs linearly} 
    \KwOut{A equivalent formula without $x_i$}

    Collect terms of $x_i$ in each atom\;
    Let $d$ to be the least common multiple of all coefficients of $x_i$\;
    Multiply each atom by a factor so that coefficient of $x_i $ is $d$\;
    Let $\theta'(x_i'):= \theta[x_i'/dx_i]$\;
    $\theta'(x_i') := \theta'(x_i') \wedge d| x_i'$\;
    \tcp{ Now the atoms have 3 different forms}
    \tcp{ $t_l(\bar{x},\bar{y})\le x_i'$, $x_i' \le t_u(\bar{x},\bar{y})$, $k_m|x_i'+t_m(\bar{x},\bar{y})$(or its negation)}
    \tcp{ where $l\in L,u\in U,m\in M$ are the index sets}
    $\delta :=  lcm\{k_m|m\in M\}$\;
    \For{all atom $\tau_l \in L$}
    {
        $\theta'_{+\infty}:=  \theta'\{\textit{true}/\tau_l\}$
    }
    \For{all atom $\tau_u \in U$}
    {
        $\theta'_{+\infty}:=  \theta'_{+\infty}\{\textit{false}/\tau_u\}$
    }
    output $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee \bigvee_{j=0}^{\delta-1} \bigvee_{u\in U}\theta'[t_u(\bar{x},\bar{y})-j/x_i']$
    \caption{QE-linear (Cooper's QE algorithm for PA)}
\end{algorithm}

%\znj{\textbf{Solved, missing $\bigvee$ for $\theta_{+\infty}$}.Pls check whether the output is correct or not, my intuition is not correct.} 
% 之前的算法

\subsection{Compared with Francoise's Origin Algorithm}

Francoise's algorithm \cite{Francoise} has four steps,
corresponding to the four steps of our algorithm respectively.
So we will use the names of our steps to refer to them. 
The divisible predicate $n|x$ is not included in his theory,
instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
The \textbf{Normalization} and \textbf{QE-with-order} steps in two algorithms are similar,
with some minor differences due to the setting of the theory.
The main differences between the two algorithms lies in the following aspects.

In \textbf{QE-exp},
we adopt the same idea to find sufficient conditions
and necessary conditions for the formula to holds.
But our strategy for choosing parameters are easier to understand and use.
A flaw lies in the $a_i<0$ case in Francoise's algorithm 
\footnote{
In the $a_i<0$ case of \textbf{QE-exp-ineq},
Francoise's algorithm assumes falsely $t(y)\ge 0$.
The reason for this mistake might be the confusion of 
two subtraction operators.
The example mentioned in \ref{sec:3.2} shows this case.
},
our \textbf{QE-exp-ineq} corrects this flaw by 
treating $a_i<0$ case similarly to the $a_i>0$ case.

\textbf{QE-linear} in Francoise's algorithm in some sense includes our
\textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
In this step,
Francoise uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
We instead invoke the QE algorithm of PA without referring to permutation groups, 
other improved QE algorithm can also be used to simplify this step.

Francoise's algorithm has more restrictions on the form
of the formula.
Before \textbf{QE-linear} and \textbf{QE-exp} step,
it needs to transform the formula into disjunction normal form (DNF)
and deals with the basic case when input formula $\theta$ is 
a conjunction of atoms. 
It is well known that transforming an arbitrary formula into 
DNF is costly and the length of the formula may increase exponentially.
So our algorithm remove these limitations and assume no special forms of
the formula other than NNF.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Complexity Analysis and Case Study}
In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
%\subsection{Complexity Analysis}

In this part, we estimate the complexity of our algorithm.
It can be divided into two part, 
the first part convert a string constraint $\phi$ to a PA formula with exponential functions,
and the second part is the QE procedure on the obtained formula.

Given a finite alphabet $\Sigma$ and $\alpha=\langle p,q \rangle$, the size of $\Sigma$ is denoted by
$|\Sigma|$. 
The generic $\alpha$-flat language contains $O(|\Sigma|^{pq})$ $\alpha$-flat languages

\paragraph{Complexity for QE}
In \textbf{QE-with-order},
it reduces the origin problem to $(n+1)!$ sub-cases,
and for each case,
we recursively invoke \textbf{QE-exp} and \textbf{QE-linear}.
So the key is to analysis the inner \textit{for}
loop in \textbf{QE-with-order}.
Note that if we omit the \textbf{QE-exp} step,
the sub-cases are equivalent to Cooper's algorithm,
so we adopt the idea of Oppen's analysis to analyze the complexity
\cite{Oppen}.

Apply Cooper's algorithm on the quantified PA formula 
$Q_m x_m Q_{m-1}$ 
$x_{m-1} ... Q_1 x_1. $
$F(x_1,...,x_m)$,
the algorithm repeats $m$ times to eliminate
$x_1,...,x_m$ one by one.
Let $F_k \Def Q_m x_m Q_{m-1} 
x_{m-1} ... Q_{k+1} x_{k+1}. F_k'(x_{k+1},...,x_m)$
be the formula produced after $k$th iterations of the algorithm. 
Let $a_k$ denote the number of atoms in $F_k$,
$c_k$ denote the number of distinct $\delta_i$ in atom $\delta_i | t$ (and its negation) plus the number of distinct coefficients of variables.
$s_k$ denote the largest value of integer constant.
Use $a_0,c_0,s_0$ to denote the initial value of $a_k,c_k,s_k$ respectively.
The following theorem holds.

\begin{theorem}
for all $k: 1\le k\le m$, the following relations holds
\begin{itemize}
    \item $c_k \le {c_{k-1}}^4$
    \item $s_k \le {s_{k-1}}^{4c_{k-1}}$
    \item $a_k \le {a_{k-1}}^4 \cdot {s_{k-1}}^{2c_{k-1}}$
\end{itemize}
and from the above relations we have 
\begin{itemize}
    \item $c_k \le {{c_0}^4}^k$
    \item $s_k \le {{{s_0}^{4c_0}}^4}^k$
    \item $a_k \le {{a_0}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
\end{itemize}
\end{theorem}

For each $k$,
the space required to store $F_k$ 
is $a_k \cdot (m+1) \cdot s_k \cdot q$,
where $(m+1)$ is the maximum number of constants per atom and $q>1$ is some constant. Assume $m\le n, c\le n, a\le n, s\le n$, we obtain the deterministic space complexity is ${{2^2}^2}^{p n log_n}$, which is also the bound for deterministic time.

In our algorithm,
we use the same denotations and have the following theorem. 
Note that each iteration corresponds to \textbf{QE-exp} and \textbf{QE-linear}.
\begin{theorem}
for all $k: 1\le k\le m$, the following relations holds
\begin{itemize}
    \item $c_k \le {c_{k-1}}^4$
    \item $s_k \le {(n s_{k-1})}^{4c_{k-1}}$
    \item $a_k \le {{(l_2(ns_{k-1})\cdot a_{k-1})}}^4 \cdot {(n s_{k-1})}^{2c_{k-1}}$
\end{itemize}
and from the above relations we have 
\begin{itemize}
    \item $c_k \le {{c_0}^4}^k$
    \item $s_k \le {{{n}^{4c_0}}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
    \item $a_k \le {{a_0}^4}^k \cdot 
     {{{n}^{4c_0}}^4}^k
     \cdot {{{s_0}^{4c_0}}^4}^k$
\end{itemize}
\end{theorem}

WAIT TO CHECK!

Again assume $m,c,a,s\le n$, which gives 
$$
\textit{space}\le q \cdot {n^4}^n \cdot {{n^{(4n)}}^4}^n \cdot
{{n^{(4n)}}^4}^n \cdot (n+1) \cdot 
\cdot {{n^{(4n)}}^4}^n \cdot {{n^{(4n)}}^4}^n
\le 2^{2^{2^{p n \textit{log}n}}}
$$

Then the upper bound for Cooper's algorithm still holds for the inner \textit{for} loop of \textbf{QE-with-order}.
When considering \textbf{Normalization} and permutations groups in \textbf{QE-with-order},
since the super exponential term dominates the number of permutations $n^{O(n)}$,
this upper bound still holds.

%\subsection{Example}

We apply our algorithm to a simple example to 
illustrate its function.
Given $\Sigma = \{0,1\}$
Suppose $\mathcal{A}$ is a finite state automaton
that recognizes language $L = \{ w \mid \#w('1') \text{ is odd}\}$.
Consider the following string constraint
$$y\in L(\mathcal{A})\wedge 4\le  \parseInt(y) < 16\,.$$

We need an abstraction parameter $alpha$ to restrict interpretations
for $x$, for simplicity, set $\alpha\Def \langle 2,1 \rangle$.
The generic $\langle 2,1 \rangle$-flat language is the union of 
$\langle 2,1 \rangle$-flat languages $(11)^*,(10)^*,(01)^*,(00)^*,1^*,0^*$.
Here we divide the problem into 6 sub-cases by restricting $y$
to one of the $\langle 2,1 \rangle$-flat language.
Another approach is to directly restrict $y$ to 
the generic $\langle 2,1 \rangle$-flat language,
at the cost of more introduced variables.
Since the complexity is related to the number of variables,
we wish it to be as little as possible.

When $y$ is in language $1^*$, assume $y = 1^x$ where $x$
is a fresh integer variable.
Note that the Parikh image of $L(\mathcal{A})$ is $2|( x+1)$, 
so the constraint is converted to a PA formula $\phi$
$$\phi\Def \exists x. 2|(x+1) \wedge 5 \le 2^{x} < 17\,.$$

Observe that when $x=3$ the formula holds,
so we can easily deduce that $\phi$ is true.  
By applying our algorithm,
after \textbf{Normalization},
$\phi$ is transformed into the form

$$\exists x_0. 2|(x+1)\wedge  2^{x}\le 16 \wedge -2^{x}\le -5\,.$$

Since there is only one variable, 
\textbf{QE-with-order} can be omitted.
We send $2^x\le 16$ and $-2^x \le -5$ to \textbf{QE-exp} and obtain
$$2^x\le 16 \equiv x \le 4$$
$$-2^x\le -5 \equiv 3\le x\,.$$ 
Then we apply \textbf{QE-linear} on formula $\exists x_0. 2|(x+1) \wedge x\le 4 \wedge 3\le x$ and get
$$\bigvee_{j=0}^1(2| j+1 \wedge \textit{false} \wedge \textit{true})\vee \bigvee_{j=0}^1 (2|4-j+i\wedge 4-j\le 4 \wedge 3\le 4-j) \,.$$
which evaluates \textit{true} and we can deduce $j=1\implies x=3$, so $y$ is string $111$.

Other sub-cases are similar. However, when the abstraction parameter becomes larger,
say, $\alpha =\langle 2,2 \rangle$,
the complexity will grow rapidly,
because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.

### 2021-05-07 FMCAD2021/5_optimization.tex
restructure the paper

Describe the optimizations for improving the scalability

### 2021-05-07 FMCAD2021/3_flatten.tex
section 3

@@
+%!TEX root = paper.tex
+
+In this section, we first define the class of string constraints with string-integer conversion, denoted by $\strint$.
+Then we define the extension of Presburger arithmetic with exponential functions, denoted by $\paexp$.
+Finally, we show how to flatten the string constraints in $\strint$ into the arithmetic constraints in $\paexp$.
+
+\subsection{String constraints with string-integer conversion ($\strint$)}
+
+Commonly, 
+{\parseInt} function takes a string representation of a decimal
+number and returns an integer.
+Since we focus on the decidability,
+we define a binary version of {\parseInt},
+which takes a binary string and returns a decimal integer number.
+For example,
+${\parseInt}('111')=7$. 
+Clearly,  our decision procedure given in this paper 
+can be adapted to string constraints with other string to number conversion function without substantial change. 
+Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
+from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.
+
+\begin{definition} Let $\Sigma_{\textit{num}}=\{'0','1'\}$,
+ ${\parseInt}:\Sigma_{\textit{num}}^*\mapsto \mathbb{N}$ is recursively defined by:
+    for $w\in \Sigma_{\textit{num}}^*$
+    \begin{itemize}
+        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
+        \item if $|w|=1, {\parseInt}('w')=w_\mathbb{N}$;
+        \item for $|w|\ge 2$ and $w=w_1 \cdot w_2$, 
+        where $|w_1|, |w_2| \ge 1$, 
+        ${\parseInt}(w) = {\parseInt}(w_1)2^{|w_2|}+{\parseInt}(w_2)$.
+    \end{itemize} 
+\end{definition}
+
+\paragraph{String Terms}
+Given a finite alphabet $\Sigma$ and 
+a finite set $X$ of string variables over $\Sigma^*$,
+we define the set of terms $\textit{Terms}(\Sigma,X)$ 
+to be the smallest set satisfying
+\begin{itemize}
+    \item[1] $\Sigma\cup \{\epsilon\} \cup X \subseteq \textit{Terms}(\Sigma,X)$;
+    \item[2] if $t_1,t_2\in \textit{Terms}(\Sigma,X)$, then $t_1 \cdot t_2 \in \textit{Terms}(\Sigma,X)$.
+\end{itemize}
+
+We extend $I_X$ to $\textit{Terms}(\Sigma,X)$ by letting $I_X(\epsilon)=\epsilon$, 
+for $a\in \Sigma, I_X(a)=a$,
+and $I_X(t_1\cdot t_2)= I_X(t_1)\cdot I_X(t_2)$.
+
+\paragraph{String Constraints} \label{par: string constraints}
+Given a constraint $\phi$ and an interpretation $I$,
+$I\models \phi$ denotes that $I$ satisfies $\phi$,
+and $I$ is called a \emph{model} of $\phi$.
+We use $\lVert \phi\rVert$ to denote the set of all models of $\phi$.
+
+We define the following three forms of atomic string constraints:
+\begin{itemize}
+\item An equality constraint $\phi_e$ is of the form 
+$t_1 = t_2$, where $t_1,t_2\in \textit{Terms}(\Sigma,X)$.
+We define $\lVert \phi_e \rVert = \{I\mid I(t_1)=I(t_2)\}$.
+Inequality constraints can be defined analogously.
+
+\item A regular constraint $\phi_r$ is of the form 
+$x\in L(\mathcal{A})$,
+where $x\in \mathbb{X}$ and $\mathcal{A}$ is a finite state automaton.
+We define $\lVert \phi_r \rVert = \{I\mid I(x)\in L(\mathcal{A})\}$.
+
+\item A length constraint $\phi_l$ is a linear constraint over 
+$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
+where $|\cdot |$ is the length function.
+We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
+\end{itemize}
+
+In Section~\ref{sec:string-solving},  
+we will introduce a new form of atomic string constraints, i.e., \emph{string-number conversion constraints}.
+A string constraint is a Boolean combination of atomic string constraints possibly with quantifications over $X\cup Z$. Other notions are same as in the first-order logic.
+
+Giving a string constraint $\Psi$,
+the problem of string constraints solving is to decide 
+whether $\lVert \Psi \rVert$ is empty,
+if not, to compute an interpretation $I$ that satisfies $\Psi$.
+
+\subsection{An Extension of Presburger Arithmetic with Exponential Functions ($\paexp$)}
+
+For a first order theory $T$,
+we say theory $T$ admits quantifier elimination (QE) if for any formula in $T$, 
+there is a quantifier-free formula equivalent to it.
+It is well-known that if a theory admits QE, 
+then it is a decidable theory.
+
+The formal definition of PA is given in section \ref{PA}.
+Here we add the ordering predicate $\le$ into the signature,
+which can be defined by $x \le y\, \Def \, \exists z. x+z=y$.
+However, the theory PA so far does not admit QE, 
+for example, consider the formula $\exists x.x = y + y$. 
+We augment the theory with countable unary divisible predicates
+$n|x$, where $n\in \mathbb{N}$, 
+$n|x$ is true if and only if $x\ \text{mod}\ n=0$ holds.
+This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.
+
+We then introduce a theory {$\paexp$}, denoted by $(\mathbb{N},+,2^x)$, that we work on.
+\begin{definition}
+    Let $\mathcal{L}=\{0,1,+,\le,n \, |x(n\in \mathbb{N}),2^x,l_2(x)\}$, 
+     $(\mathbb{N},+,2^x)$ be a $\mathcal{L}$-theory that has domain $\mathbb{N}$, where 
+    \begin{itemize}
+        \item  $2^x$ is interpreted to the exponential function of $2$ over $\mathbb{N}$; 
+        \item interpretations of $0,1,+,\le,=$ are consistent with PA;
+        \item for $n\ge 1$,$n|x$ holds iff $\exists y.x=ny$;
+        \item $2^0=1$, for $n \ge 1, 2^n = 2^{n-1}+2^{n-1}$;
+        We further assume that if $m,n\in \mathbb{N}, m\le n$, 
+        then $2^{m-n} = 1$
+        \item $l_2(0)=0$; for $n \ge 1,l_2(n) = y$ iff $2^y \le n < 2^{y+1}$;  $l_2(m-n) = 0$
+        if $m,n\in \mathbb{N}$ and $m\le n$. 
+    \end{itemize}
+\end{definition}
+
+$\lambda_2(x) = 2^{l_2(x)}$ can be defined by $l_2(x)$,
+intuitively, $\lambda_2(x)$ means the maximal power of 2 that is not larger than $x$.
+Then we have $\lambda_2(x) \le x \le 2\lambda_2(x)-1$,
+which will be useful in our proof.
+
+\subsection{Flattening $\strint$ into $\paexp$}
+
+The flattening technique was first introduced 
+in \cite{Abdulla 2017}.
+Fix an alphabet $\Sigma$ and an abstraction parameter $\alpha$, 
+for a given atomic string constraint $\phi$, 
+flattening $\phi$ with parameter $\alpha$ 
+results in a new string constraint $\phi_\alpha$,
+such that 
+$R(\lVert \phi_\alpha \rVert) = \lVert \phi \rVert \cap \{I \mid \forall x\in X, I(x)\in \mathbb{G}'(\alpha)\}$,
+where $R$ is the renaming function with its domain extended to interpretations in the normal manner.
+Intuitively, it restricts $\phi$ to interpret 
+string variables over $\mathbb{G}'(\alpha)$.
+
+\cite{Abdulla 2017} discussed the flattening of basic 
+string constraints including equality, integer, (regular) grammar and transducer constraints. 
+For an atomic string constraint $\phi$,
+the flattening $\phi_\alpha$ is still an atomic string constraint 
+and is Parikh definable,
+so its Parikh image can be expressed by a quantifier free PA formula.
+Together with the purity condition,
+we obtain an existential quantified PA formula $\rho$.
+$\rho$ will be sent to a SMT solver,
+if the solver returns an solution $\theta$,
+then we can construct an interpretation for $\phi_\alpha$ from $\theta$,
+otherwise it means $\phi$ is unsatistiable when 
+string variables are interpreted  to $\alpha$-flat languages.
+
+Take a regular constraint $\phi = x\in L(A)$ for example,
+the flattening of $\phi$ resutls in a new finite state automaton $A'$ over $\Sigma(\alpha)$,
+which encodes running $A$ ``in parallel" 
+with the generic $\alpha$-flat automaton.
+Let $\rho_1$ be the formula describing the Parikh image of $A'$,
+which is a formula over variable sets $\Sigma(\alpha)^\#$.
+Let $\rho_2$ be the purity condition \eqref{eq:purity}.
+Then we obtain the PA formula $\exists (\Sigma(\alpha))^\#. \rho_1 \wedge \rho_2$.
+In order to distinguish between different string variables,
+we may replace $a(i)^\# \in \Sigma(\alpha)^\#$ by $(x,a(i))^\#$.
+
+Since the structure of a flat automaton is decided 
+by its abstraction parameter $\alpha$, 
+a Counter-Example Guided Abstraction 
+Refinement (CEGAR) framework is designed, which contains both an under- and an 
+over-approximation module, to search the possible values of $\alpha$.
+The termination for the overall algorithm is not guaranteed.
+
 The string-number conversion functions are commonly used functions
@@
 Hence, we identify a decidable subset of string constraints, which is the largest one with decidability so far to the best of our knowledge.
-
-\subsection{String-Number Conversion Function}
-
-Commonly, 
-{\parseInt} function takes a string representation of a decimal
-number and returns an integer.
-Since we focus on the decidability,
-we define a binary version of {\parseInt},
-which takes a binary string and returns a decimal integer number.
-For example,
-${\parseInt}('111')=7$. 
-Clearly,  our decision procedure given in this paper 
-can be adapted to string constraints with other string to number conversion function without substantial change. 
-Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
-from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.
-
-\begin{definition} Let $\Sigma_{\textit{num}}=\{'0','1'\}$,
- ${\parseInt}:\Sigma_{\textit{num}}^*\mapsto \mathbb{N}$ is recursively defined by:
-    for $w\in \Sigma_{\textit{num}}^*$
-    \begin{itemize}
-        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
-        \item if $|w|=1, {\parseInt}('w')=w_\mathbb{N}$;
-        \item for $|w|\ge 2$ and $w=w_1 \cdot w_2$, 
-        where $|w_1|, |w_2| \ge 1$, 
-        ${\parseInt}(w) = {\parseInt}(w_1)2^{|w_2|}+{\parseInt}(w_2)$.
-    \end{itemize} 
-\end{definition}
 
@@
 In this way,
-the problem can still be reduced to the decidability of PA with exponential function (\paexp), i.e.,
+the problem can still be reduced to the decidability of PA with exponential function ($\paexp$), i.e.,
 \begin{theorem} \label{thm:string-parInt}
-If {\paexp} is decidable, then the satisfiability (validity) of string constraints with {\parseInt} in which all string variables 
+If {$\paexp$} is decidable, then the satisfiability (validity) of string constraints with {\parseInt} in which all string variables 
 are ranged over flat strings is decidable. 

### 2021-05-07 FMCAD2021/7_conclusion.tex
section 3

%!TEX root = paper.tex

Conclude this paper and discuss future work.

### 2021-05-07 Benchmark/HashFunction/runstring.sh
arithmetic tests done

@@ -1,19 +1,32 @@
 #!/bin/bash
-echo "--------------------------------"
+echo "begin test..."
 
-echo "begin string test Using Z3 ..."
-START=$(date +%s)
-z3 -smt2 ./all/head/test1 -T:60
-END=$(date +%s)
-DIFF=$(( $END - $START ))
-echo "Z3 took $DIFF seconds"
+cd all
+    cd head
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+    cd head-tail
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+    cd tail
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+cd ..
 
-echo "--------------------------------"
-
-echo "begin string test Using CVC4 ..."
-START=$(date +%s)
-cvc4 ./all/head/test1 -L smt --tlimit=60000 --strings-exp   
-END=$(date +%s)
-DIFF=$(( $END - $START ))
-echo "CVC4 took $DIFF seconds"
-echo "--------------------------------"
+cd flatten
+    cd head
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+    cd head-tail
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+    cd tail
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+cd ..

### 2021-05-07 Benchmark/NIA-test/produce-unbound.py

@@ -23,7 +23,24 @@ for testno in range(1,totaltests+1):
     for i in range (1,expnum+linearnum+1):
         s.add(x[i] >= 0)
 
-    for i in range (1,expfml+1):
+    g.write("{")
+    expr = 0
+    for j in range (1,expnum+1):
+        coef = random.randint(-10,10)
+        while coef == 0:
+            coef = random.randint(-10,10)
+        expr = expr + coef * pow(x[j])
+        g.write("%d," %coef)
+    for j in range (1,expnum+linearnum+1):
+        coef = random.randint(-10**5,10**5)
+        expr = expr + coef * x[j]
+        g.write("%d," %coef)
+    num = random.randint(-10**5,10**5)
+    s.add(expr == num)
+    g.write("%d}\n" % num)
+
+
+    for i in range (2,expfml+1):
         g.write("{")
         expr = 0
         for j in range (1,expnum+1):
@@ -58,7 +75,7 @@ for testno in range(1,totaltests+1):
         s.add(expr <= num)
         g.write("%d}\n" % num)
             
-    f = open('NIA-unbound/2-3-3-4/test'+str(testno),mode='w+')
+    f = open('NIA-unbound/3-4-4-5-eq/test'+str(testno),mode='w+')
     f.write("(set-logic ALL)\n")
     f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
     f.write(s.sexpr())

### 2021-05-07 Benchmark/NIA-test/runall.sh

@@ -2,10 +2,14 @@
 echo "begin test..."
 
 cd NIA-unbound
-    cd 3-4-4-5
+    cd 3-4-4-5-eq
+    ./runz3.sh > z3-result 
     ./runcvc4.sh > cvc4-result
     cd ..
-    cd 4-5-5-6
+cd ..
+
+cd NIA-bound
+    cd 3-4-4-5-eq
     ./runz3.sh > z3-result 
     ./runcvc4.sh > cvc4-result
     cd ..

### 2021-05-07 FMCAD2021/2_preliminary.tex

@@
+%!TEX root = paper.tex
+
 In this section, we introduce some basic concepts and theories that will be used later. 
@@
 is Parikh definable \cite{Parikh66}.
-
-\paragraph{String Terms}
-Given a finite alphabet $\Sigma$ and 
-a finite set $X$ of string variables over $\Sigma^*$,
-we define the set of terms $\textit{Terms}(\Sigma,X)$ 
-to be the smallest set satisfying
-\begin{itemize}
-    \item[1] $\Sigma\cup \{\epsilon\} \cup X \subseteq \textit{Terms}(\Sigma,X)$;
-    \item[2] if $t_1,t_2\in \textit{Terms}(\Sigma,X)$, then $t_1 \cdot t_2 \in \textit{Terms}(\Sigma,X)$.
-\end{itemize}
-
-We extend $I_X$ to $\textit{Terms}(\Sigma,X)$ by letting $I_X(\epsilon)=\epsilon$, 
-for $a\in \Sigma, I_X(a)=a$,
-and $I_X(t_1\cdot t_2)= I_X(t_1)\cdot I_X(t_2)$.
-
-\paragraph{String Constraints} \label{par: string constraints}
-Given a constraint $\phi$ and an interpretation $I$,
-$I\models \phi$ denotes that $I$ satisfies $\phi$,
-and $I$ is called a \emph{model} of $\phi$.
-We use $\lVert \phi\rVert$ to denote the set of all models of $\phi$.
-
-We define the following three forms of atomic string constraints:
-\begin{itemize}
-\item An equality constraint $\phi_e$ is of the form 
-$t_1 = t_2$, where $t_1,t_2\in \textit{Terms}(\Sigma,X)$.
-We define $\lVert \phi_e \rVert = \{I\mid I(t_1)=I(t_2)\}$.
-Inequality constraints can be defined analogously.
-
-\item A regular constraint $\phi_r$ is of the form 
-$x\in L(\mathcal{A})$,
-where $x\in \mathbb{X}$ and $\mathcal{A}$ is a finite state automaton.
-We define $\lVert \phi_r \rVert = \{I\mid I(x)\in L(\mathcal{A})\}$.
-
-\item A length constraint $\phi_l$ is a linear constraint over 
-$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
-where $|\cdot |$ is the length function.
-We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
-\end{itemize}
-
-In Section~\ref{sec:string-solving},  
-we will introduce a new form of atomic string constraints, i.e., \emph{string-number conversion constraints}.
-A string constraint is a Boolean combination of atomic string constraints possibly with quantifications over $X\cup Z$. Other notions are same as in the first-order logic.
-
-Giving a string constraint $\Psi$,
-the problem of string constraints solving is to decide 
-whether $\lVert \Psi \rVert$ is empty,
-if not, to compute an interpretation $I$ that satisfies $\Psi$.
 
@@
 is that every word $w\in \mathbb{G}(\alpha)$ is uniquely determined by its Parikh image $\mathbb{P}(w)$.
-
-\subsection{Flattening}
-
-The flattening technique was first introduced 
-in \cite{Abdulla 2017}.
-Fix an alphabet $\Sigma$ and an abstraction parameter $\alpha$, 
-for a given atomic string constraint $\phi$, 
-flattening $\phi$ with parameter $\alpha$ 
-results in a new string constraint $\phi_\alpha$,
-such that 
-$R(\lVert \phi_\alpha \rVert) = \lVert \phi \rVert \cap \{I \mid \forall x\in X, I(x)\in \mathbb{G}'(\alpha)\}$,
-where $R$ is the renaming function with its domain extended to interpretations in the normal manner.
-Intuitively, it restricts $\phi$ to interpret 
-string variables over $\mathbb{G}'(\alpha)$.
-
-\cite{Abdulla 2017} discussed the flattening of basic 
-string constraints including equality, integer, (regular) grammar and transducer constraints. 
-For an atomic string constraint $\phi$,
-the flattening $\phi_\alpha$ is still an atomic string constraint 
-and is Parikh definable,
-so its Parikh image can be expressed by a quantifier free PA formula.
-Together with the purity condition,
-we obtain an existential quantified PA formula $\rho$.
-$\rho$ will be sent to a SMT solver,
-if the solver returns an solution $\theta$,
-then we can construct an interpretation for $\phi_\alpha$ from $\theta$,
-otherwise it means $\phi$ is unsatistiable when 
-string variables are interpreted  to $\alpha$-flat languages.
-
-Take a regular constraint $\phi = x\in L(A)$ for example,
-the flattening of $\phi$ resutls in a new finite state automaton $A'$ over $\Sigma(\alpha)$,
-which encodes running $A$ ``in parallel" 
-with the generic $\alpha$-flat automaton.
-Let $\rho_1$ be the formula describing the Parikh image of $A'$,
-which is a formula over variable sets $\Sigma(\alpha)^\#$.
-Let $\rho_2$ be the purity condition \eqref{eq:purity}.
-Then we obtain the PA formula $\exists (\Sigma(\alpha))^\#. \rho_1 \wedge \rho_2$.
-In order to distinguish between different string variables,
-we may replace $a(i)^\# \in \Sigma(\alpha)^\#$ by $(x,a(i))^\#$.
-
-Since the structure of a flat automaton is decided 
-by its abstraction parameter $\alpha$, 
-a Counter-Example Guided Abstraction 
-Refinement (CEGAR) framework is designed, which contains both an under- and an 
-over-approximation module, to search the possible values of $\alpha$.
-The termination for the overall algorithm is not guaranteed.

### 2021-05-07 FMCAD2021/4_dec_proc.tex

@@
-\subsection{Decidability of \paexp} \label{sec:3.2}
+%!TEX root = paper.tex
 
-For a first order theory $T$,
-we say theory $T$ admits quantifier elimination (QE) if for any formula in $T$, 
-there is a quantifier-free formula equivalent to it.
-It is well-known that if a theory admits QE, 
-then it is a decidable theory.
-
-The formal definition of PA is given in section \ref{PA}.
-Here we add the ordering predicate $\le$ into the signature,
-which can be defined by $x \le y\, \Def \, \exists z. x+z=y$.
-However, the theory PA so far does not admit QE, 
-for example, consider the formula $\exists x.x = y + y$. 
-We augment the theory with countable unary divisible predicates
-$n|x$, where $n\in \mathbb{N}$, 
-$n|x$ is true if and only if $x\ \text{mod}\ n=0$ holds.
-This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.
-
-We then introduce a theory {\paexp}, denoted by $(\mathbb{N},+,2^x)$, that we work on.
-\begin{definition}
-    Let $\mathcal{L}=\{0,1,+,\le,n \, |x(n\in \mathbb{N}),2^x,l_2(x)\}$, 
-     $(\mathbb{N},+,2^x)$ be a $\mathcal{L}$-theory that has domain $\mathbb{N}$, where 
-    \begin{itemize}
-        \item  $2^x$ is interpreted to the exponential function of $2$ over $\mathbb{N}$; 
-        \item interpretations of $0,1,+,\le,=$ are consistent with PA;
-        \item for $n\ge 1$,$n|x$ holds iff $\exists y.x=ny$;
-        \item $2^0=1$, for $n \ge 1, 2^n = 2^{n-1}+2^{n-1}$;
-        We further assume that if $m,n\in \mathbb{N}, m\le n$, 
-        then $2^{m-n} = 1$
-        \item $l_2(0)=0$; for $n \ge 1,l_2(n) = y$ iff $2^y \le n < 2^{y+1}$;  $l_2(m-n) = 0$
-        if $m,n\in \mathbb{N}$ and $m\le n$. 
-    \end{itemize}
-\end{definition}
-
-$\lambda_2(x) = 2^{l_2(x)}$ can be defined by $l_2(x)$,
-intuitively, $\lambda_2(x)$ means the maximal power of 2 that is not larger than $x$.
-Then we have $\lambda_2(x) \le x \le 2\lambda_2(x)-1$,
-which will be useful in our proof.
+%\subsection{Decidability of $\paexp$} \label{sec:3.2}
 
@@
 theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
-Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
+Exponential functions are compatible with addition functions, and therefore {$\paexp$} is decidable 
 by \cite{Semenov84}.
@@
 In the next section, within Point's framework, we give a revised algorithm
- to the decision problem of {\paexp} with 
+ to the decision problem of {$\paexp$} with 
 both subtle improvements and crucial corrections, that 

### 2021-05-09 Benchmark/NIA-test/produce-bound.py
pruning opt

@@ -1,15 +1,15 @@
 from z3 import *
 import random
 
-expnum = 3
-linearnum = 4
-expfml = 4
-linearfml = 5
+expnum = 1
+linearnum = 2
+expfml = 2
+linearfml = 3
 totaltests = 100
 
 x=[0]*(expnum+linearnum+2)
 
-g = open("NIA-bound/3-4-4-5-eq/paras","w+")
+g = open("NIA-bound/1-2-2-3/paras","w+")
 g.write("%d\n" %totaltests)
 g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
 
@@ -23,22 +23,7 @@ for testno in range(1,totaltests+1):
     for i in range (1,expnum+linearnum+1):
         s.add(x[i] >= 0)
 
-    g.write("{")
-    expr = 0
-    for j in range (1,expnum+1):
-        coef = random.randint(-10,10)
-        expr = expr + coef * pow(x[j])
-        g.write("%d," %coef)
-    for j in range (1,expnum+linearnum+1):
-        coef = random.randint(-10**5,10**5)
-        expr = expr + coef * x[j]
-        g.write("%d," %coef)
-    num = random.randint(-10**5,10**5)
-    s.add(expr == num)
-    g.write("%d}\n" % num)
-
-
-    for i in range (2,expfml+1):
+    for i in range (1,expfml+1):
         g.write("{")
         expr = 0
         for j in range (1,expnum+1):
@@ -66,7 +51,7 @@ for testno in range(1,totaltests+1):
         s.add(expr <= num)
         g.write("%d}\n" % num)
             
-    f = open('NIA-bound/3-4-4-5-eq/test'+str(testno),mode='w+')
+    f = open('NIA-bound/1-2-2-3/test'+str(testno),mode='w+')
     f.write("(set-logic ALL)\n")
     f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
     f.write(s.sexpr())

### 2021-05-09 Benchmark/NIA-test/runall.sh

@@ -1,15 +1,8 @@
 #!/bin/bash
 echo "begin test..."
 
-cd NIA-unbound
-    cd 3-4-4-5-eq
-    ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
-    cd ..
-cd ..
-
 cd NIA-bound
-    cd 3-4-4-5-eq
+    cd 1-2-2-3
     ./runz3.sh > z3-result 
     ./runcvc4.sh > cvc4-result
     cd ..

### 2021-05-09 FMCAD2021/4_dec_proc.tex
opt section

@@
 
-%\znjSide{\textbf{Solved}. I think you need to explain the meanings of compatibility.}
-
 Semenov proved that for any function $f$ that is compatible with addition, 
@@
 the expected result should be $y\le 2^{10} =1024$,
-but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
+but Point's algorithm will return a formula equivalent to $y \le 2^6=64$.
 % 这里好像还有另一个与Francoise一起的作者
@@
 
-%\znj{\textbf{Solved, missing $\bigvee$ for $\theta_{+\infty}$}.Pls check whether the output is correct or not, my intuition is not correct.} 
-% 之前的算法
+\subsection{Compared with Point's Origin Algorithm}
 
-\subsection{Compared with Francoise's Origin Algorithm}
-
-Francoise's algorithm \cite{Francoise} has four steps,
+Point's algorithm \cite{Point86} has four steps,
 corresponding to the four steps of our algorithm respectively.
@@
 But our strategy for choosing parameters are easier to understand and use.
-A flaw lies in the $a_i<0$ case in Francoise's algorithm 
+A flaw lies in the $a_i<0$ case in Point's algorithm 
 \footnote{
 In the $a_i<0$ case of \textbf{QE-exp-ineq},
-Francoise's algorithm assumes falsely $t(y)\ge 0$.
+Point's algorithm assumes falsely $t(y)\ge 0$.
 The reason for this mistake might be the confusion of 
@@
 
-\textbf{QE-linear} in Francoise's algorithm in some sense includes our
+\textbf{QE-linear} in Point's algorithm in some sense includes our
 \textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
 In this step,
-Francoise uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
+Point uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
 We instead invoke the QE algorithm of PA without referring to permutation groups, 
@@
 
-Francoise's algorithm has more restrictions on the form
+Point's algorithm has more restrictions on the form
 of the formula.
@@
 
+\iffalse
 \subsection{Complexity Analysis and Case Study}
@@
 because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.
+\fi

### 2021-05-09 FMCAD2021/5_optimization.tex
pruning opt
update the result of arithmetic
opt section

@@
+%!TEX root = paper.tex
+
+In the last section, 
+we introduced quantifier elimination as a decision procedure for {$\paexp$}. According to theorem \ref{thm:string-parInt}, the satisfiability of {$\strint$} fragment is decidable.
+However, the quantifier elimination procedure can not be directly applied on solving string constraints over {$\strint$} due to the following two reasons.
+First, quantifier elimination does not return a model that satisfies the constraints.
+In constrast, it only leaves a formula with constants and free variables.
+And to obtain satisfiability, all free variables need to be treated as existential quantified variables and be eliminated one by one,
+Second, the quantifier elimination procedure is highly complicated.
+Note that we alternatively invoke QE-exp and Cooper's algorithm to eliminate exponential terms and linear terms of a quantified variable, the length of the formula grows super-exponentially in the procedure.
+
+In this section, we introduced some optimizations on the QE procedure when focus on the quantifier-free fragment in {$paexp$}.
+Given a quantifier-free formula, we wish to decide its satisfiability and find an assignment for variables in the formula if it exists.
+Since the Cooper's algorithm is the main contribution to the complexity, we try to only eliminate exponential terms using an variation of QE-exp and send the obtained linear formula to SMT-solver.
+The cost is that the algorithm is no longer complete in some situations, in which case some practical restrictions are needed.
+
+A quantifier-free formula is a boolean combination of constraints.
+Here we allow only equalities and inequalities, divisible predicates are replaced by equalities, for example, $2|x$ is replace to $x=2y$, where $y$ is a fresh variable.
+Since we only eliminate exponential terms, the newly introduced $y$ will not bother.
+For convenience, we call variables that occurs exponentially in the formula \emph{exponential variables}, variables that have only linear occurrences \emph{linear variables}.
+Similarly,
+constraints that have exponential terms are called exponential constraints, others are called linear constraints.
+
+\paragraph{QE-exp for all atoms}
+
+\paragraph{over-approximation}
+
+\paragraph{DFS and pruning}
+
 Describe the optimizations for improving the scalability
+
+existential PAexp formula, framework, eliminate exponential terms leaving linear terms, send the (equivalent) linear formula to SMT solver
+
+over-approximation for UNSAT and find possible orders
+
+eliminate exponential terms for the formula
+
+dfs search

### 2021-05-10 FMCAD2021/4_dec_proc.tex
update sec 4

@@
 %!TEX root = paper.tex
-
-%\subsection{Decidability of $\paexp$} \label{sec:3.2}
 
@@
     For a strictly increasing function $f:\mathbb{N}\mapsto \mathbb{N}$, 
-    $f$ is said to be compatible with addition 
+    $f$ is said to be \emph{compatible with addition}
     if for every $m\in \mathbb{N}^+$,  $f$ modulo $m$ is periodic, 
@@
 
-Semenov proved that for any function $f$ that is compatible with addition, 
-theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
-Exponential functions are compatible with addition functions, and therefore {$\paexp$} is decidable 
-by \cite{Semenov84}.
-Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
-Unfortunately, Point's algorithm is flawed, one problem 
- lies in the discussion to eliminate exponential terms. 
-% making the algorithm fail to handle "$a_0<0$" case.
-For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
-the expected result should be $y\le 2^{10} =1024$,
-but Point's algorithm will return a formula equivalent to $y \le 2^6=64$.
-% 这里好像还有另一个与Francoise一起的作者
+Semenov proved that for any function $f$ that is compatible with addition, the first order theory $(\mathbb{N},+,f)$ admits QE  and thus is complete and decidable \cite{Semenov84}.
+Exponential functions are a typical class of compatible with addition functions, and therefore {$\paexp$} is decidable.
+Particularly, in \cite{Point86}
+Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$.
+%下面这些感觉可以不用说，或者放在后面
+%Unfortunately, Point's algorithm is flawed, one problem lies in the discussion to eliminate exponential terms.
+For example, after  eliminating variable $x$ in formula $\exists x.
+y\le 2^x \wedge x\le 10$, the expected result should be $y\le 2^{10} =1024$, but Point's algorithm will return a formula equivalent to $y \le 2^6=64$.
 
-In the next section, within Point's framework, we give a revised algorithm
- to the decision problem of {$\paexp$} with 
-both subtle improvements and crucial corrections, that 
-provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to 
-Theorem~\ref{thm:string-parInt}.
+In this section, within Point's framework, we give a revised algorithm to the decision problem of {$\paexp$} with both improvements and corrections, that provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to Theorem~\ref{thm:string-parInt}.
 
 %\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
-Based on Point's work \cite{Point86}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
-where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 
-Since $\forall x. F = \neg \exists x. \neg F$, we further assume the quantifier $Q$ to be the existential quantifier, 
-that is, $\exists x.\theta(x,\bar{y})$. 
-For eliminating quantifiers in arbitrary formula, 
-we apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.
+% Based on Point's work \cite{Point86}, 
+We will present a revised QE algorithm for {$\paexp$} formulae of the form $Qx.\theta(x,\bar{y})$, where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula.
+Since $\forall x.
+F = \neg \exists x.
+\neg F$, we further assume the quantifier is an existential quantifier, that is, $\exists x.\theta(x,\bar{y})$.
+For eliminating quantifiers in arbitrary {$\paexp$} formula, we need to apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.
 
-We divide the whole QE algorithm into 4 steps.
-The first step $\textbf{Normalization}$ can be viewed as a pre-processing step, which substitutes ``complex" terms like $2^{2^x}$, $l_2(3x+y)$ with ``simple" terms by introducing new variables $x_i$.
-Now we move forward to handle a ``simpler" formula, however, at the cost of more quantified variables.
+The whole QE algorithm can be divided into 4 steps.
+The first step \textbf{Normalization} can be viewed as a pre-processing step, which substitutes ``complex" terms like $2^{2^x}$, $l_2(3x+y)$ with ``simple" terms by introducing new variables.
+After \textbf{Normalization}, we move forward to handle a ``simpler" formula, however, at the cost of more quantified variables.
 
-The other three steps undertake the task to eliminate the introduced variables $x_i$ and the original variable $x$ (will be denoted by $x_0$) one by one. 
-The main body of \textbf{QE-with-Order} step is a loop to enumerate all possible orders among quantified variables. 
-In each iteration, according the given (decreasing) order (corresponding to a for loop), we invoke \textbf{QE-exp} or 
-\textbf{QE-linear} to eliminate these variables one by one. 
-%depending on  eliminates the maximal variable in the remaining $\bar{x}$.
-So it is necessary to specify an ordering for $\bar{x}$ at the beginning, as we will see, this is the main reason for the complexity of this algorithm.
-During each iteration of the inner loop (the for loop), 
-if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atomic formula, 
-we will invoke \textbf{QE-exp} to produce a formula equivalent to the atom where $x_i$ occurs linearly;  
-otherwise, if $x_i$ occurs linearly in the formula, \textbf{QE-linear} will be invoked to eliminate all occurrences of $x_i$, 
-and this procedure is similar to the classic QE algorithm for PA.
+The other three steps undertake the task to eliminate the newly introduced variables in \textbf{Normalization} step, denoted by $\{x_i:1\le i\le n\}$, and the original variable $x$ (will be denoted by $x_0$) one by one.
+The main body of \textbf{Enumerate-Orders} step is a loop to enumerate all possible orders among quantified variables $\{x_i : 0\le i \le n\}$. 
+In each iteration, according the given (decreasing) order of quantified variables (corresponding to an outer \emph{for} loop), we invoke \textbf{Elim-exp} and \textbf{Elim-linear} to eliminate these variables one by one.
+
+During each iteration of the inner \emph{for} loop, if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atomic formula $\tau$, we will invoke \textbf{Elim-exp} to produce a new formula where $x_i$ occurs linearly equivalent to $\tau$.
+Then \textbf{Elim-linear} will eliminate all linear occurrences of $x_i$, which is similar to the classic Cooper's QE algorithm for PA \cite{Cooper72}.
 
@@
 
-Make substitutions by introducing new variables according to the \textit{while} statements in the pseudo-code.For consistency, we rename the original $x$ to be $x_0$ and assume $n$ new quantified variables are introduced.
+Make substitutions by introducing new variables according to the \textit{while} statements in the pseudo-code.
+For consistency, we rename the original $x$ to be $x_0$ and assume $n$ new quantified variables are introduced.
 
-After introducing new variables and substitution, 
-we obtain a formula where if an exponential term contains a quantified variable $x_i(0\le i\le n)$, the term should be $2^{x_i}$.
-We then collect all terms with $x_i(0\le i\le n)$ together, 
-it will be of the form $s(\bar{x})\Def \sum_{i=0}^{n} a_i 2^{x_i} + \sum_{i=0}^n b_i x_i$, where $a_i,b_i(0\le i \le n)$ are all constants.
-Other terms including constants and terms of $\bar{y}$ are collected,  
-denoted by $t(\bar{y})$. 
-Since $\bar{y}$ are free variables, 
-we will regard $t(\bar{y})$ as a constant.
+%After introducing new variables and substitution, 
+After substitutions, we obtain a formula where if an exponential term contains a quantified variable $x_i$, the term must be $2^{x_i}$.
+We then collect all terms with $x_i (0\le i\le n)$ together, it will be of the form $s(\bar{x}) \Def \sum_{i=0}^{n} a_i 2^{x_i} + \sum_{i=0}^n b_i x_i$, where $a_i,b_i (0\le i \le n)$ are all constants.
+Other terms including constants and terms of free variables $\bar{y}$ are collected, denoted by $t(\bar{y})$.
+Since $\bar{y}$ are free variables, we will regard $t(\bar{y})$ as a constant.
 Inequalities and equalities will all be transformed into $s(\bar{x})\le t(\bar{y})$, for example, $s(\bar{x}) = t(\bar{y})$ will be replaced by $s(\bar{x})\le t(\bar{y})\wedge t(\bar{y})\le s(\bar{x})$.
@@
     \tcp{ i is used for counting the introduced variables }
-    i = 1\; 
+    $i := 1$\; 
     \tcp{ $\bar{y}$ are regarded as constants} 
@@
     $n:= i-1$\;
-    \tcp{ Collect quantified variables $x_i$}
-    Transform all atoms into forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ or $\neg  (k|s(\bar{x})+t(\bar{y}))$\;
+    \tcp{ Collect quantified variables}
+    Transform all atoms of $\theta'$ into forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ or $\neg  (k|s(\bar{x})+t(\bar{y}))$\;
     Return $\exists x_0,...,\exists x_n. \theta'$
@@
 
-\subsection{QE-with-order} 
-During the \textbf{Normalization} step, 
-we simplify the origin formula at the cost of more quantified variables. 
+\subsection{Enumerate-Orders} 
+In the \textbf{Normalization} step, we simplify the origin formula at the cost of more quantified variables. 
 Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
@@
  the quantifier free formula, and then according to the 
- order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{QE-exp} first, and then invoking \textbf{QE-linear} (the inner \textit{for} loop). 
+ order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{Elim-exp} first, and then invoking \textbf{Elim-linear} (the inner \textit{for} loop). 
  In each iteration of the inner \textit{for} loop, 
 if $x_{\sigma{(i)}}$ occurs in an exponential term in an atomic formula, 
-\textbf{QE-exp} is invoked and a formula   equivalent to the atom is returned, in which $x_i$ occurs linearly;  
-then $x_i$ occurs linearly in the formula, thus \textbf{QE-linear} is further invoked to eliminate all occurrences of $x_i$. 
+\textbf{Elim-exp} is invoked and a formula   equivalent to the atom is returned, in which $x_i$ occurs linearly;  
+then $x_i$ occurs linearly in the formula, thus \textbf{Elim-linear} is further invoked to eliminate all occurrences of $x_i$. 
 The returned formula is the disjunction of all formulas resulted in all iterations of the outer \textit{for} loop.
@@
 That is to say,
-every execution of \textbf{QE-exp} and \textbf{QE-linear} will eliminate the largest element $x_i$ in $\bar{x}$.
+every execution of \textbf{Elim-exp} and \textbf{Elim-linear} will eliminate the largest element $x_i$ in $\bar{x}$.
 However, as you may notice, if there is no clue about the order of $\bar{x}$, 
@@
 
-Here is how \textbf{QE-exp} and \textbf{QE-linear} are invoked in
+Here is how \textbf{Elim-exp} and \textbf{Elim-linear} are invoked in
 \textbf{Ordering and QE}. 
@@
             \tcp{if $x_{\sigma(i)}$ occurs exponentially in $\theta_\sigma$}
-            $\theta_{\sigma}:= \textbf{QE-exp}(x_{\sigma(i)},\theta_{\sigma})$\;
+            $\theta_{\sigma}:= \textbf{Elim-exp}(x_{\sigma(i)},\theta_{\sigma})$\;
             \tcp{now $x_{\sigma(i)}$ occurs only linearly in $\theta_\sigma$}
-            $\theta_{\sigma}:= \textbf{QE-linear}(x_{\sigma(i)},\theta_{\sigma})$\;
+            $\theta_{\sigma}:= \textbf{Elim-linear}(x_{\sigma(i)},\theta_{\sigma})$\;
         }
@@
         \While{$x_i$ occurs in an exponential term in formula $\tau(x_i,\bar{x},\bar{y})$}{
-            \tcp{ QE-exp outputs a formula equivalent with $\tau(x_i,\bar{x},\bar{y})$ where $x_i$ occurs linearly }
-            $\theta_i :=  \theta_i[\text{QE-exp}(\tau(x_i,\bar{x},\bar{y}))/\tau(x_i,\bar{x},\bar{y})]$\;
+            \tcp{ Elim-exp outputs a formula equivalent with $\tau(x_i,\bar{x},\bar{y})$ where $x_i$ occurs linearly }
+            $\theta_i :=  \theta_i[\text{Elim-exp}(\tau(x_i,\bar{x},\bar{y}))/\tau(x_i,\bar{x},\bar{y})]$\;
         }   
@@
         \For{ every $\theta_{i,s}$ }{
-            $\theta'_{i,s} :=  \text{QE-linear}(x_i,\theta_{i,s})$
+            $\theta'_{i,s} :=  \text{Elim-linear}(x_i,\theta_{i,s})$
         }
@@
     output $\phi$
-    \caption{QE-with-order}
+    \caption{Enumerate-Orders}
 \end{algorithm}
@@
 
-\subsection{QE-exp}
+\subsection{Elim-exp}
 
-\textbf{QE-exp} and \textbf{QE-linear} are the most technical parts of our QE algorithm.
+\textbf{Elim-exp} and \textbf{Elim-linear} are the most technical parts of our QE algorithm.
 As mentioned, 
-\textbf{QE-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as 
+\textbf{Elim-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as 
 inputs, and  outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
-The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{QE-linear}.
+The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{Elim-linear}.
 For simplicity, we will abuse   $x_i$ for $x_{\sigma(i)}$, 
@@
 The problem will be discussed in 2 cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
-corresponding to \textbf{QE-exp-ineq} or \textbf{QE-exp-div}.
+corresponding to \textbf{Elim-exp-ineq} or \textbf{Elim-exp-div}.
 
-We first give the whole procedure of \textbf{QE-exp},
+We first give the whole procedure of \textbf{Elim-exp},
 and then provide the details of the two sub-routines.
@@
         { 
-            \tcp{ \textbf{QE-exp-ineq} case}
+            \tcp{ \textbf{Elim-exp-ineq} case}
             $A :=  \sum_{j=0}^{i-1}|a_j|$, $B:= b_i + \sum_{j=0}^{i-1}|b_j|$\;
@@
         {
-            \tcp{ \textbf{QE-exp-div} case, the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
+            \tcp{ \textbf{Elim-exp-div} case, the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
             \tcp{ let $d=2^r d_0$ where $d_0$ is odd}
@@
     }
-    \caption{QE-exp}
+    \caption{Elim-exp}
 \end{algorithm}
 
-\subsubsection{QE-exp-ineq}
+\subsubsection{Elim-exp-ineq}
 
@@
 
-\subsubsection{QE-exp-div}
+\subsubsection{Elim-exp-div}
 
@@
 $$\tau(x_i,\bar{x},\bar{y}) \Def  d| a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k+t(\bar{y}) \quad (a_i \neq 0)\,,$$
-the algorithm \textbf{QE-exp-div} outputs an equivalent formula without $2^{x_i}$ terms.
+the algorithm \textbf{Elim-exp-div} outputs an equivalent formula without $2^{x_i}$ terms.
 The idea is that $a_i 2^x$ modulo $d$ is a periodic function when $x$ is large enough and the period can be computed.
@@
 
-\subsection{QE-linear}
+\subsection{Elim-linear}
 
-After \textbf{QE-exp}, 
+After \textbf{Elim-exp}, 
 we obtain a formula (still denoted by $\theta(x_i,\bar{x},\bar{y})$) without $2^{x_i}$ terms,
@@
     output $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee \bigvee_{j=0}^{\delta-1} \bigvee_{u\in U}\theta'[t_u(\bar{x},\bar{y})-j/x_i']$
-    \caption{QE-linear (Cooper's QE algorithm for PA)}
+    \caption{Elim-linear (Cooper's QE algorithm for PA)}
 \end{algorithm}
@@
 instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
-The \textbf{Normalization} and \textbf{QE-with-order} steps in two algorithms are similar,
+The \textbf{Normalization} and \textbf{Enumerate-Orders} steps in two algorithms are similar,
 with some minor differences due to the setting of the theory.
@@
 
-In \textbf{QE-exp},
+In \textbf{Elim-exp},
 we adopt the same idea to find sufficient conditions
@@
 \footnote{
-In the $a_i<0$ case of \textbf{QE-exp-ineq},
+In the $a_i<0$ case of \textbf{Elim-exp-ineq},
 Point's algorithm assumes falsely $t(y)\ge 0$.
@@
 },
-our \textbf{QE-exp-ineq} corrects this flaw by 
+our \textbf{Elim-exp-ineq} corrects this flaw by 
 treating $a_i<0$ case similarly to the $a_i>0$ case.
 
-\textbf{QE-linear} in Point's algorithm in some sense includes our
-\textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
+\textbf{Elim-linear} in Point's algorithm in some sense includes our
+\textbf{Elim-exp-div}, because divisible predicates is not introduced in his setting.
 In this step,
-Point uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
+Point uses the permutation groups (just like $S_{(n+1)}$ in \textbf{Enumerate-Orders}) and needs recursively renaming of variables $x_i$.
 We instead invoke the QE algorithm of PA without referring to permutation groups, 
@@
 of the formula.
-Before \textbf{QE-linear} and \textbf{QE-exp} step,
+Before \textbf{Elim-linear} and \textbf{Elim-exp} step,
 it needs to transform the formula into disjunction normal form (DNF)
@@
 \paragraph{Complexity for QE}
-In \textbf{QE-with-order},
+In \textbf{Enumerate-Orders},
 it reduces the origin problem to $(n+1)!$ sub-cases,
 and for each case,
-we recursively invoke \textbf{QE-exp} and \textbf{QE-linear}.
+we recursively invoke \textbf{Elim-exp} and \textbf{Elim-linear}.
 So the key is to analysis the inner \textit{for}
-loop in \textbf{QE-with-order}.
-Note that if we omit the \textbf{QE-exp} step,
+loop in \textbf{Enumerate-Orders}.
+Note that if we omit the \textbf{Elim-exp} step,
 the sub-cases are equivalent to Cooper's algorithm,
@@
 we use the same denotations and have the following theorem. 
-Note that each iteration corresponds to \textbf{QE-exp} and \textbf{QE-linear}.
+Note that each iteration corresponds to \textbf{Elim-exp} and \textbf{Elim-linear}.
 \begin{theorem}
@@
 
-Then the upper bound for Cooper's algorithm still holds for the inner \textit{for} loop of \textbf{QE-with-order}.
-When considering \textbf{Normalization} and permutations groups in \textbf{QE-with-order},
+Then the upper bound for Cooper's algorithm still holds for the inner \textit{for} loop of \textbf{Enumerate-Orders}.
+When considering \textbf{Normalization} and permutations groups in \textbf{Enumerate-Orders},
 since the super exponential term dominates the number of permutations $n^{O(n)}$,
@@
 Since there is only one variable, 
-\textbf{QE-with-order} can be omitted.
-We send $2^x\le 16$ and $-2^x \le -5$ to \textbf{QE-exp} and obtain
+\textbf{Enumerate-Orders} can be omitted.
+We send $2^x\le 16$ and $-2^x \le -5$ to \textbf{Elim-exp} and obtain
 $$2^x\le 16 \equiv x \le 4$$
 $$-2^x\le -5 \equiv 3\le x\,.$$ 
-Then we apply \textbf{QE-linear} on formula $\exists x_0. 2|(x+1) \wedge x\le 4 \wedge 3\le x$ and get
+Then we apply \textbf{Elim-linear} on formula $\exists x_0. 2|(x+1) \wedge x\le 4 \wedge 3\le x$ and get
 $$\bigvee_{j=0}^1(2| j+1 \wedge \textit{false} \wedge \textit{true})\vee \bigvee_{j=0}^1 (2|4-j+i\wedge 4-j\le 4 \wedge 3\le 4-j) \,.$$
@@
 the complexity will grow rapidly,
-because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.
+because \textbf{Enumerate-Orders} can not always be omitted and formulae in \textbf{Elim-exp} can not be simplified easily.
 \fi

### 2021-05-10 FMCAD2021/4_dec_proc.tex
update sec 4

@@
 % Based on Point's work \cite{Point86}, 
-We will present a revised QE algorithm for {$\paexp$} formulae of the form $Qx.\theta(x,\bar{y})$, where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula.
+We will present a revised QE algorithm for {$\paexp$} formulas of the form $Qx.\theta(x,\bar{y})$, where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula.
 Since $\forall x.
@@
 
-During each iteration of the inner \emph{for} loop, if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atomic formula $\tau$, we will invoke \textbf{Elim-exp} to produce a new formula where $x_i$ occurs linearly equivalent to $\tau$.
+During each iteration of the inner \emph{for} loop, if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atom, we will invoke \textbf{Elim-exp} to produce an equivalent formula where $x_i$ occurs linearly.
 Then \textbf{Elim-linear} will eliminate all linear occurrences of $x_i$, which is similar to the classic Cooper's QE algorithm for PA \cite{Cooper72}.
@@
 Inequalities and equalities will all be transformed into $s(\bar{x})\le t(\bar{y})$, for example, $s(\bar{x}) = t(\bar{y})$ will be replaced by $s(\bar{x})\le t(\bar{y})\wedge t(\bar{y})\le s(\bar{x})$.
-
-At the end, the resulted formula $\theta'$ only contains literals of the forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ and $\neg  (k|s(\bar{x})+t(\bar{y}))$.
+ end, the resulted formula $\theta'$ only contains literals of the forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ and $\neg  (k|s(\bar{x})+t(\bar{y}))$.
 
@@
 In the \textbf{Normalization} step, we simplify the origin formula at the cost of more quantified variables.
-Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
-and it has $n$ new variables $x_i$ ($x$ is denoted by $x_0$).
-As a consequence, we need to eliminate all quantified variables one by one.
+Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$ and it has $n$ new variables $x_i$.
+As a consequence, we need to eliminate all $(n+1)$ quantified variables one by one.
 
-To the end, we denote the set of all orders among the $n+1$ variables by 
- $S_{n+1}$, and then enumerate these orders one by one in the outer \textit{for} loop. 
- For the considered order $\sigma$, we first add the ordering information to 
- the quantifier free formula, and then according to the 
- order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{Elim-exp} first, and then invoking \textbf{Elim-linear} (the inner \textit{for} loop). 
- In each iteration of the inner \textit{for} loop, 
-if $x_{\sigma{(i)}}$ occurs in an exponential term in an atomic formula, 
-\textbf{Elim-exp} is invoked and a formula   equivalent to the atom is returned, in which $x_i$ occurs linearly;  
-then $x_i$ occurs linearly in the formula, thus \textbf{Elim-linear} is further invoked to eliminate all occurrences of $x_i$. 
+%To the end, 
+We denote the set of all orders among the $(n+1)$ variables by $S_{n+1}$, and then enumerate these orders one by one in the outer \textit{for} loop.
+For the considered order $\sigma$, we first add the ordering information to $\theta$, and then according to the order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{Elim-exp} first, and then invoking \textbf{Elim-linear} (the inner \textit{for} loop).
+In each iteration of the inner \textit{for} loop, if $x_{\sigma{(i)}}$ occurs in an exponential term in an atom $\tau$, \textbf{Elim-exp} is invoked to produce a formula, in which $x_i$ occurs linearly, equivalent to $\theta$;
+then $x_i$ occurs only linearly, thus \textbf{Elim-linear} is further invoked to eliminate all (linear) occurrences of $x_i$.
 The returned formula is the disjunction of all formulas resulted in all iterations of the outer \textit{for} loop.
@@
 Here is how \textbf{Elim-exp} and \textbf{Elim-linear} are invoked in
-\textbf{Ordering and QE}. 
+\textb
+At thef{Ordering and QE}. 
 \fi
-
-%我们这里还是考虑用循环群来做，因为如果每一次只找到最大的$x_i$，那么递归的执行还是需要整个$\bar{x}$的顺序。但如果一开始就知道了，后面有些操作可以简化
 
@@
     \SetAlgoLined
-    \KwIn{a normalized $(n+1)$-existential formula $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
-    where $\bar{x}=(x_0,...,x_n)$} 
-    \KwOut{a equivalent quantifier free formula without $\bar{x}$}
-
-    Let $S_{n+1}$ denote the group of permutations on $\{0,1,...,n\}$\;
+    \KwIn{a normalized $(n+1)$-existential formula $\exists \bar{x}.\theta(\bar{x},\bar{y})$, where $\bar{x}=(x_0,...,x_n)$} 
+    \KwOut{an equivalent quantifier-free formula without $\bar{x}$}
+    \tcp{Let $S_{n+1}$ denote the group of permutations on $\{0,1,...,n\}$}
     $\phi := \textit{False}$\;
@@
     }
-\iffalse
-    \For{i from $0$ to $n$}
-    {
-        \tcp{ specify the maximal quantified variable}
-        $\theta_i :=  \theta(\bar{x},\bar{y}) \wedge (\bigwedge_{0\le j \le n} x_i - x_j \ge 0)$\;
-        \tcp{ eliminate $x_i$}
-        \While{$x_i$ occurs in an exponential term in formula $\tau(x_i,\bar{x},\bar{y})$}{
-            \tcp{ Elim-exp outputs a formula equivalent with $\tau(x_i,\bar{x},\bar{y})$ where $x_i$ occurs linearly }
-            $\theta_i :=  \theta_i[\text{Elim-exp}(\tau(x_i,\bar{x},\bar{y}))/\tau(x_i,\bar{x},\bar{y})]$\;
-        }   
-        transform $\theta_i$ into disjunction normal form $\theta_i = \bigvee_s \theta_{i,s}$\;
-        \For{ every $\theta_{i,s}$ }{
-            $\theta'_{i,s} :=  \text{Elim-linear}(x_i,\theta_{i,s})$
-        }
-        $\theta'_i =\bigvee_s \theta'_{i,s}$\;
-    }
-\fi
     output $\phi$
@@
 \end{algorithm}
-%\znj{I cannot understand why you need to do QE for all permutation $\sigma$? I think 
-%one is enough.}
 
@@
 
-\textbf{Elim-exp} and \textbf{Elim-linear} are the most technical parts of our QE algorithm.
-As mentioned, 
-\textbf{Elim-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as 
-inputs, and  outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
+\textbf{Elim-exp} is the most technical part of our QE algorithm.
+As mentioned, \textbf{Elim-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as inputs, and outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
 The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{Elim-linear}.
-For simplicity, we will abuse   $x_i$ for $x_{\sigma(i)}$, 
-and  $\bar{x}$ for $(x_{\sigma(0)},...,x_{\sigma(i-1)})$ ( remind that $x_{\sigma(i+1)},...,x_n$ have been eliminated already).
+For simplicity, we will abuse notations $x_i$ for $x_{\sigma(i)}$, and $\bar{x}$ for $(x_{\sigma(0)},...,x_{\sigma(i-1)})$ ( remind that $x_{\sigma(i+1)},...,x_n$ have been eliminated already).
 
@@
 the predicates $\le$, $|$ and negation of $|$, respectively.
-The problem will be discussed in 2 cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
+The problem will be discussed in two cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
 corresponding to \textbf{Elim-exp-ineq} or \textbf{Elim-exp-div}.
 
-We first give the whole procedure of \textbf{Elim-exp},
-and then provide the details of the two sub-routines.
+%We first give the whole procedure of \textbf{Elim-exp},and then provide the details of the two sub-routines.
+
 \begin{algorithm}[t]
@@
         {
-            \tcp{ \textbf{Elim-exp-div} case, the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
+            \tcp{the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
+            \tcp{ \textbf{Elim-exp-div} case}
             \tcp{ let $d=2^r d_0$ where $d_0$ is odd}
@@
 
-This algorithm deals with the case where $\tau$ is an inequality of the form 
+This part deals with the case when $\tau$ is an inequality of the form 
 $$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
@@
 \begin{theorem} \label{thm:exp-ineq}
-Assume that $x_i$ is the maximal variable in $\bar{x}$ according to the given order. 
-Given an inequality $\tau(x_i,\bar{x},\bar{y})$ of the form %, suppose $x_i$ occurs in an exponential term, that is 
+Assume that $x_i$ is the maximal variable in $\{x_0,...,x_{i}\}$ according to the given order, $\bar{x}$ denotes $\{x_0,...,x_{i-1}\}$.
+Given an inequality $\tau(x_i,\bar{x},\bar{y})$ of the form  
 $$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
-with $a_i \neq 0$, 
-let $A\Def \sum_{j=0}^{i-1}|a_j|$, 
+with $a_i \neq 0$. Define $A\Def \sum_{j=0}^{i-1}|a_j|$, 
 $B\Def |b_i| + \sum_{j=0}^{i-1}|b_j|$, 
@@
 
-Step 1: Put $\theta(x_i,\bar{x},\bar{y})$ in NNF form, replace atomic formulae containing symbols other than $\le,|$ by equivalent formulae only with $\le$, that is, $x=a$ by $x\le a \wedge a\le x$, $x< a$ by $x\le a+1$, $x\neq a$ by $x>a \vee x<a$,  $x\geq a$ by $-x \le -a$, and $x>a$ by $-x \le -a-1$.
+Step 1: Put $\theta(x_i,\bar{x},\bar{y})$ in NNF form, replace atomic formulas containing symbols other than $\le,|$ by equivalent formulas only with $\le$, that is, $x=a$ by $x\le a \wedge a\le x$, $x< a$ by $x\le a+1$, $x\neq a$ by $x>a \vee x<a$,  $x\geq a$ by $-x \le -a$, and $x>a$ by $-x \le -a-1$.
 In inequality atoms, collect terms of $x_i$
@@
 the complexity will grow rapidly,
-because \textbf{Enumerate-Orders} can not always be omitted and formulae in \textbf{Elim-exp} can not be simplified easily.
+because \textbf{Enumerate-Orders} can not always be omitted and formulas in \textbf{Elim-exp} can not be simplified easily.
 \fi

### 2021-05-10 FMCAD2021/3_flatten.tex
preliminaries

@@
 \subsection{String constraints with string-integer conversion ($\strint$)}
+
+There are two types of variables in string constraints,
+i.e., $X$, a set of string variables ranged over $\Sigma^*$,  
+and $Z$,  a set of integer variables ranged over $\mathbb{Z}$.
+As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
+essentially a pair of two mappings 
+$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
+where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.
 

### 2021-05-10 FMCAD2021/3_flatten.tex
flatten

--- 2021-05-07 FMCAD2021/3_flatten.tex
@@
 
-Commonly, 
-{\parseInt} function takes a string representation of a decimal
-number and returns an integer.
-Since we focus on the decidability,
-we define a binary version of {\parseInt},
-which takes a binary string and returns a decimal integer number.
+In the sequel, we shall define $\strint$, the class of string constraints with the string-integer conversion function {\parseInt}.
+
+The function  {\parseInt} takes a decimal string as the input and returns the integer represented by the string\footnote{The {\parseInt} function in scripting languages e.g.
+Javascript is more general in the sense that the base can be a number between 2 and 36.
+Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for readibility.}.
+%Since we focus on the decidability,
+%we define a binary version of {\parseInt},
+%which takes a binary string and returns a decimal integer number.
 For example,
-${\parseInt}('111')=7$. 
-Clearly,  our decision procedure given in this paper 
-can be adapted to string constraints with other string to number conversion function without substantial change. 
-Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
-from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.
+${\parseInt}('0123') = {\parseInt}('123')=10^2+10*2+3 = 123$. 
+Note here we use the quotation marks to delimit the strings, 
+%Clearly,  our decision procedure given in this paper 
+%can be adapted to string constraints with other string to number conversion function without substantial change. 
+%Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
+%from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.
 
-\begin{definition} Let $\Sigma_{\textit{num}}=\{'0','1'\}$,
- ${\parseInt}:\Sigma_{\textit{num}}^*\mapsto \mathbb{N}$ is recursively defined by:
-    for $w\in \Sigma_{\textit{num}}^*$
+Formally, the semantics of the $\parseInt$ function is defined as follows. 
+%\begin{definition} 
+Let $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$. Then ${\parseInt}: \Sigma_{\textit{num}}^+ \mapsto \Nat$ is recursively defined by
+    for every $w\in \Sigma_{\textit{num}}^+$,
     \begin{itemize}
-        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
-        \item if $|w|=1, {\parseInt}('w')=w_\mathbb{N}$;
-        \item for $|w|\ge 2$ and $w=w_1 \cdot w_2$, 
-        where $|w_1|, |w_2| \ge 1$, 
-        ${\parseInt}(w) = {\parseInt}(w_1)2^{|w_2|}+{\parseInt}(w_2)$.
+%        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
+        \item if $w={'i'}$ for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}('i')=i$;
+        \item for $w = w_1 {'i'}$ for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
+        ${\parseInt}(w) = 10*{\parseInt}(w_1)+{\parseInt}({'i'})$.
     \end{itemize} 
-\end{definition}
+%\end{definition}
+Note that $\parseInt$ is undefined with $\varepsilon$ as the input.
 
+In $\strint$, there are two types of variables, namely, the string variables $\svarx,\svary,\ldots \in \svars$ and the integer variables $\ivarx,\ivary,\ldots \in \ivars$.
+%
+A $\strint$ formula $\varphi$ is defined by the following rules, where $\len(\sterm)$ denotes the length of a string $\sterm$.
+\[
+\begin{array}{l c l}
+\sterm & ::= & a \mid \svarx \mid \sterm \concat \sterm, \\
+\iterm & ::= & n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,\\
+\varphi & ::= & \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi,
+\end{array}
+\]
+where $a \in \Sigma_\varepsilon$, $n \in \Nat$, $\Aut$ is an FA, and $\op \in \{=, < , >, \le, \ge\}$.
+Let us call $\sterm$ as string terms, $\iterm$ as integer terms, $\sterm = \sterm$ as string equality constraints, $x \in \Aut$ as regular constraints, $\iterm \op\ \iterm$ as arithmetic constraints.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 \paragraph{String Terms}
@@
 \end{itemize}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-In Section~\ref{sec:string-solving},  
-we will introduce a new form of atomic string constraints, i.e., \emph{string-number conversion constraints}.
-A string constraint is a Boolean combination of atomic string constraints possibly with quantifications over $X\cup Z$. Other notions are same as in the first-order logic.
+The $\strint$ formulas are interpreted on $I=(I_s, I_i)$ where $I_s$ is a partial function from $\svars$ (the set of string variables) to $\Sigma^*$ and $I_i$ is a partial function from $\ivars$ (the set of integer variables) to $\Nat$.
+Moreover, it is required that the domains of $I_s, I_i$ are finite.
+Given $I = (I_s, I_i)$, the interpretations of string terms, integer terms, as well as the formulas $\varphi$ under $I$ are easy to comprehend, thus omitted to avoid tediousness.
+For $\varphi \in \strint$ and $I = (I_s, I_i)$, $I$ satisfies $\varphi$ if the interpretation of $\varphi$ under $I$ is $\ltrue$.
+Let us use $\lVert \varphi \rVert$ to denote the set of $I = (I_s, I_i)$ satisfying $\varphi$.
 
-Giving a string constraint $\Psi$,
-the problem of string constraints solving is to decide 
-whether $\lVert \Psi \rVert$ is empty,
-if not, to compute an interpretation $I$ that satisfies $\Psi$.
+%As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
+%essentially a pair of two mappings 
+%$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
+%where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.
+
+The satisfiability problem of $\strint$ is to decide for a given constraint $\varphi \in \strint$,
+whether $\lVert  \varphi \rVert \neq \emptyset$.
+%if not, to compute an interpretation $I$ that satisfies $\Psi$.
 
 \subsection{An Extension of Presburger Arithmetic with Exponential Functions ($\paexp$)}
+
+\zhilin{stopped here}
 

### 2021-05-11 FMCAD2021/2_preliminary.tex

@@
 
-In this section, we introduce some basic concepts and theories that will be used later. 
-\subsection{Basic Concepts}
+%In this section, we introduce some basic concepts and theories that will be used later. 
+%\subsection{Basic Concepts}
 
-\paragraph{Sets and Strings}
-We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the set of natural numbers and integers, respectively. 
-$\mathbb{N}^+$ stands for the set of non-zero natural numbers. 
-Let $\Sigma$ be a finite alphabet,
-a string $w$ over $\Sigma$ is a sequence $a_1....a_n$ of characters from $\Sigma$.
-Empty string is denoted by $\epsilon$.
-$\Sigma^*$ denotes the set of all finite strings over $\Sigma$,
-and $\Sigma_{\epsilon}$ stands for $\Sigma \cup \{\epsilon\}$.
-For any string $w_1, w_2\in \Sigma^*$, 
-we use $|w_1|$ to denote the length of $w_1$,
-and $w_1\cdot w_2$ to denote the concatenation of $w_1$ and $w_2$.
+In this section, we fix the notations and introduce some basic concepts, including Presburger arithmetic, finite-state automata, Parikh images, flat automata and languages.
+
+\paragraph{Integers, strings, and languges}
+Let $\Nat$ denote the set of natural numbers, $\Int$ denote the set of integers, and $\Int^+$ denote the set of positive integers. For $n \in \Int^+$, let $[n]$ denote $\{1,\dots,n\}$.
+
+An \emph{alphabet} $\Sigma$ is a finite set. Elements of $\Sigma$ are called \emph{letters}.
+%Let $\Sigma$ denote a finite alphabet,
+A \emph{string} $w$ over $\Sigma$ is a (possibly empty) finite sequence $a_1\ldots a_n$ with $a_i \in \Sigma$ for every $i \in [n]$.
+Let $\varepsilon$ denote the empty string, namely, the empty sequence.
+For a string $w = a_1 \ldots a_n \in \Sigma^*$, let $|w|$ denote the \emph{length} of $|w|$, i.e.
+$n$.
+In particular, $|\varepsilon| = 0$.
+%
+For $w_1 = a_1 \ldots a_m, w_2 = b_1 \ldots b_n\in \Sigma^*$, 
+let $w_1\cdot w_2$ denote the \emph{concatenation} of $w_1$ and $w_2$, that is, $a_1 \ldots a_m b_1 \ldots b_n$.
+%
+Let $\Sigma^*$ denote the set of all strings over $\Sigma$ and $\Sigma^+$ denote the set of nonempty strings over $\Sigma$. 
+For convenience, we also use $\Sigma_{\epsilon}$ to denote $\Sigma \cup \{\epsilon\}$.
+%
 A language $L$ over $\Sigma$ is a subset of $\Sigma^*$.
 
-There are two types of variables in string constraints,
-i.e., $X$, a set of string variables ranged over $\Sigma^*$,  
-and $Z$,  a set of integer variables ranged over $\mathbb{Z}$.
-As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
-essentially a pair of two mappings 
-$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
-where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.
+\paragraph{Presburger Arithmetic} \label{PA}
+A Presburger Arithmetic (PA) formula 
+% is a first order theory over the signature 
+%$\Sigma_\mathbb{N}\Def  \{0,1,+, <, |\}$. 
+ is defined by the rules 
+ $$
+ \begin{array}{ l c l}
+ \phi &::= & t \ \op\ 0 \mid c | t \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,\\
+ t &::=& c \mid \ivarx \mid t+t \mid t - t,
+ \end{array}
+ $$
+where $\op \in \{=, < , >, \le, \ge\}$ and $\ivarx, c$ are integer variables and constants, respectively.
+A \emph{quantifier-free} PA (QFPA) formula is a PA formula containing no quantifiers.
+A PA formula is in \emph{negation normal form} (NNF) if all occurrences of the negation symbol $\neg$ are before the atomic formulas.
+A PA formula is called \emph{existential} if it is in NNF and contains no occurrences of universal quantifiers.
+%An \emph{existential} LIA (ELIA) formula is an LIA formula where there are no universal quantifiers and all existential quantifiers are in the scope of an even number of the negation symbols $\neg$. 
+%
+The set of free variables of $\phi$, denoted by $\free(\phi)$, is defined in a standard manner. 
+We usually write $\phi(\ivarx_1,\cdots, \ivarx_k)$ to denote an PA formula $\phi$ such that $\free(\phi) \subseteq \{\ivarx_1,\cdots, \ivarx_k\}$.
+Given an PA formula $\phi$, and an integer interpretation of $\free(\phi)$, i.e.
+a function $I: \free(\phi) \rightarrow \Int$, we denote by $I \models \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner, with $+$, $-$, and $|$ interpreted as the integer addition, subtraction, and divisibility relation respectively), and call $I$ a \emph{model} of $\phi$.
+We use $\llbracket \phi \rrbracket$ to denote the set of models of $\phi$.
 
-\paragraph{Finite State Automata}
-A Finite State Automaton is a tuple 
-$\mathcal{A}=\langle Q,\Sigma,\Delta,$ $q_{\textit{init}},$ $q_{\textit{acc}}\rangle$, 
-where $Q$ is a finite set of states, 
-$\Sigma$ is the given alphabet,
-$\Delta\subseteq Q\times \Sigma_\epsilon\times Q$ 
-defines the transition relations in $\mathcal{A}$.
-$q_{\textit{init}},q_{\textit{acc}}\in Q$ is the initial state and accepting state. 
-A sequence $q_0 \langle a_1 \rangle  q_1 ...
-\langle a_n\rangle q_n$ is called accepting if $q_0 = q_{\textit{init}}$, $q_n = q_{\textit{acc}}$ and $q_{i-1}\langle a_i \rangle q_i \in \Delta$ for $1\le i\le n$.
-
-\paragraph{Presburger Arithmetic} \label{PA}
-The Presburger Arithmetic (PA) is a first order theory over signature 
-$\Sigma_\mathbb{N}\Def  \{0,1,+,=\}$, where $0,1$ are constants, 
-$+$ is a binary function and $=$ is a binary predicate.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
+Presburger Arithmetic (PA) is a first order theory over the signature 
+$\Sigma_\mathbb{N}\Def  \{0,1,+, <, |\}$, where $0,1$ are constants, 
+$+$ is a binary function symbol, $<, |$ are  and $=$ is a binary predicate.
 
@@
 \end{itemize}
-
 Given the domain $\mathbb{N}$,
@@
 where $p>1$ is a constant\cite{Oppen69}. 
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Finite state automata}
+A finite state automaton (FA) is a tuple 
+$\Aut=\langle Q,\Sigma,\Delta, q_{\textit{init}}, F\rangle$, 
+where $Q$ is a finite set of states, 
+$\Sigma$ is a finite alphabet,
+$\Delta\subseteq Q\times \Sigma_\epsilon\times Q$ is the transition relation, 
+$q_{\textit{init}}$ is the initial state, $F \subseteq Q$ is the set of accepting states. 
+A \emph{run} of $\Aut$ on a string $w = a_1 \ldots a_n$ is a sequence $q_0 \xrightarrow{b_1} q_1  \xrightarrow{b_2}  \ldots   \xrightarrow{b_{m-1}}  q_{m-1} \xrightarrow{b_m} q_m$ such that $q_0 = q_{\textit{init}}$, $(q_{i-1}, b_i, q_i) \in \Delta$ for every $i \in [m]$, and $a_1 \ldots a_n = b_1 \ldots b_m$.
+A \emph{run} $q_0 \xrightarrow{b_1} q_1  \xrightarrow{b_2}  \ldots   \xrightarrow{b_{m-1}}  q_{m-1} \xrightarrow{b_m} q_m$ is \emph{accepting} if $q_m  \in F$.
+A string $w$ is \emph{accepted} by $\Aut$ if there is an accepting run of $\Aut$ on $w$.
+Let $\Lang(\Aut)$ denote the set of strings accepted by $\Aut$.
+A language $L \subseteq \Sigma^*$ is \emph{regular} if it can be defined by some FA $\Aut$, namely, $L = \Lang(\Aut)$.
 
 \paragraph{Parikh Image}
-Given an alphabet $\Sigma$ and a string $w\in \Sigma^*$, 
-we define the set of Parikh variables 
-$\Sigma^\# \Def \{a^\# \mid a\in \Sigma\}$.
-The Parikh image of $w$ is a function 
-$\mathbb{P}(w): \Sigma^\# \mapsto \mathbb{N}$,
-which maps each symbol $a^\#\in \Sigma^\#$ to the number of occurrences
-of $a$ in $w$.
-For example, let $w\Def aabba$,
-then $\mathbb{P}(w)(a^\#)=3,\mathbb{P}(w)(b^\#)=2$.
+%Given an alphabet $\Sigma$ and a string $w\in \Sigma^*$, 
+%we define the set of Parikh variables 
+%$\Sigma^\# \Def \{a^\# \mid a\in \Sigma\}$.
+%The Parikh image of $w$ is a function 
+%$\mathbb{P}(w): \Sigma^\# \mapsto \mathbb{N}$,
+%which maps each symbol $a^\#\in \Sigma^\#$ to the number of occurrences
+%of $a$ in $w$.
+%
+The \emph{Parikh image} of a word $w \in \Sigma^*$, denoted by $\parikh(w)$, maps each Parikh (integer) variable $\#a$, where $a \in \Sigma$, to the number of occurrences of $a$ in $w$. 
+For instance, let $\Sigma = \{a,b\}$ and $w= aabba$,
+then $\parikh(w)(\#a)=3$ and $\parikh(w)(\#b)=2$.
+Given an alphabet $\Sigma$, let $\#\Sigma$ denote the set of Parikh variables $\{\#a | a \in \Sigma\}$. 
+%The Parikh image of $w$ is a function $\parikh(w): \#\Sigma \rightarrow \mathbb{N}$ such that $\parikh(w)(\#a) = |w|_a$, for each $a \in \Sigma$. 
+The Parikh image of a language $L$ is defined as $\parikh(L) = \{\parikh(w) \mid w \in L\}$.
+It is well known that the Parikh image of a regular language can be characterized by an existential PA formula~\cite{SSMH04}.
 
-For a language $L\subseteq \Sigma^*$, 
-define the Parikh image of $L$ to be 
-$\mathbb{P}(L)\Def \{\mathbb{P}(w) | w\in L\}$.
-We say a language $L$ is \emph{Parikh-definable} 
-if $\mathbb{P}(L)$ can be characterized by a quantifier-free PA formula over $\Sigma^\#$, 
-where $a^\#$ in the formula encodes the number of occurrences of $a$.
-It is well known that
-any context-free language (therefore regular language) 
-is Parikh definable \cite{Parikh66}.
+%For a language $L\subseteq \Sigma^*$, 
+%define the Parikh image of $L$ to be 
+%$\mathbb{P}(L)\Def \{\mathbb{P}(w) | w\in L\}$.
+%We say a language $L$ is \emph{Parikh-definable} 
+%if $\mathbb{P}(L)$ can be characterized by a quantifier-free PA formula over $\Sigma^\#$, 
+%where $a^\#$ in the formula encodes the number of occurrences of $a$.
+%It is well known that
+%any context-free language (therefore regular language) 
+%is Parikh definable \cite{Parikh66}.
 
-\subsection{Flat Automata and Flat Languages}
+%\paragraph{Flat automata and languages}
 
-Given a string constraint $\Psi$,
-the general problem of deciding whether $\lVert \Psi \rVert$ is empty is undecidable.
-However, 
-the problem becomes decidable when certain restrictions are imposed.
-One of the restriction is by flat automata and flat languages, 
-defined below.
+%Given a string constraint $\Psi$,
+%the general problem of deciding whether $\lVert \Psi \rVert$ is empty is undecidable.
+%However, 
+%the problem becomes decidable when certain restrictions are imposed.
+%One of the restriction is by flat automata and flat languages, 
+%defined below.
 
-\paragraph{Flat Languages and Automata}
+\paragraph{Flat languages}
+
+We will present flat languages. This is a high level, language-theoretical view at the flat automata from \cite{Parosh:20:PLDI}.
+
+%For integers $k$ and $\ell$ and a string variable $x$, 
+%we define the family of indexed \emph{character variables} $\charvars = \{x_j^i \mid 1 \le i \le k, 1 \le j \le \ell\}$. 
+A \emph{flat language} (FL)  over $\Sigma$
+%with the \emph{period} at most $\ell$ and the \emph{cycle count} $k$ 
+is the set of strings 
+that conform to a regular expression of the form
+$$
+(a^1_1\ldots a^1_{\ell_1})^* \cdot \ldots \cdot (a^k_1\ldots a^k_{\ell_k})^*, 
+$$
+where $a^i_j \in \Sigma$ for each $i \in [k]$ and $j \in [\ell_i]$.
+Intuitively, an FL is a set of strings consisting of $k$ consecutive parts, each created by iterating a \emph{cycle}, a string $a^i_1\ldots a^i_{\ell_i}$ of $\ell_i$ letters. 
+%A $\langle \ell, k \rangle$-FL is an FL such that its \emph{period} is at most $\ell$ and the \emph{cycle count} is at most $k$.
+For instance, the language defined by $(ab)^*(a)^*(bb)^*$ is an FL.
+
+%A central property of FL is that the strings accepted by an FL are fully characterized by their Parikh images.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 For a fixed alphabet $\Sigma$,
@@
 
-A $\langle 2,3 \rangle$-flat automaton 
-that recognizes $L\Def  (ab)^*(a)^*(bb)^*$ is shown in 
-figure (\ref{fig: FA}).
-
 \begin{figure}[ht]
@@
 \end{figure}
-
+} 
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 \paragraph{Generic Flat Languages and Automata}
@@
 is that every word $w\in \mathbb{G}(\alpha)$ is uniquely determined by its Parikh image $\mathbb{P}(w)$.
+}

### 2021-05-11 FMCAD2021/1_intro.tex

@@
-solve string constraint is hard
+The emerging of scripting languages boosted the needs of better approaches and tools to ensure program quality.
+Comparing with traditional programming languages, string data type plays a more critical role in its analysis.
+String constraint solvers are the engine of modern scripting program analysis techniques. 
+Due to the high demand, in recent years, there is a boosting amount of publications on this subject.
+
+However, research progress of string constraint solving has been hampered by many major obstacles in both theory and tool implementation aspects (including long-standing open problems). 
+Logical theories over strings have to allow string concatenation, which is arguably the most fundamental operation of strings. 
+The most celebrated result concerning theories of strings is Makanin’s result on deciding the satisfiability problem for word equations~\cite{}.
+A simple example of a word equation is $xaby = ybax$, where $x, y$ are variables, and $a, b \in \Sigma$ are constant letters. 
+A word equation is satisfiable if it has a solution, i.e., an assignment that maps variables to strings over the alphabet $\Sigma$ which equate the left-hand side with the right-hand side of the equation.
+The correctness proof of Makanin’s algorithm is arguably one of the most complex termination proofs in computer science. 
+Makanin’s result can be extended to include regular constraints (a.k.a. regular expression matching, e.g., $x \in (ba)^*)$, and arbitrary Boolean connectives.
+This extension is called word equations with regular constraints. 
+The complexity of the satisfiability of word equations with regular constraints was proven to be PSPACE-complete by Plandowski~\cite{}, after decades of improvement of the original algorithm by Makanin.
+
+Satisfiability of word equations is a special instance of Hilbert’s 10th problem. 
+In the past, the original motivation of studying word equations was for finding an undecidability proof of Hilbert’s 10th problem. 
+However, the motivation is no longer valid since Makanin finds a decision procedure.
+Interestingly, in the last decade, driven by the need for program analysis, people start to revisit the problem and its extensions to describe the complete string library APIs in conventional programming languages.
+
+The satisfiability of string/number conversion constraints has been proven undecidable in~\cite{}. 
+It is undecidable even without combining with other types of string constraints.
+However, this kind of constraint is pervasive in scripting language programs. 
+For example, it is common that programs read string input from text files and conversion a part of the string input to numbers.
+Even more crucially, in many programming languages, the string-number conversion is a part of the definition of their core semantics. 
+JavaScript, which powers most interactive content on the Web and increasingly server- side code with Node.js, is one of such languages.
 

### 2021-05-11 FMCAD2021/2_preliminary.tex

@@
 \paragraph{Presburger Arithmetic} \label{PA}
-A Presburger Arithmetic (PA) formula 
+A Presburger Arithmetic ({\pa}) formula 
 % is a first order theory over the signature 
@@
  \begin{array}{ l c l}
- \phi &::= & t \ \op\ 0 \mid c | t \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,\\
- t &::=& c \mid \ivarx \mid t+t \mid t - t,
+ \iterm &::= & c \mid \ivarx \mid \iterm+\iterm \mid \iterm - \iterm, \\
+ \phi & ::= & \iterm \ \op\ 0 \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,
+
  \end{array}
@@
 where $\op \in \{=, < , >, \le, \ge\}$ and $\ivarx, c$ are integer variables and constants, respectively.
-A \emph{quantifier-free} PA (QFPA) formula is a PA formula containing no quantifiers.
-A PA formula is in \emph{negation normal form} (NNF) if all occurrences of the negation symbol $\neg$ are before the atomic formulas.
-A PA formula is called \emph{existential} if it is in NNF and contains no occurrences of universal quantifiers.
+A \emph{quantifier-free} PA (\qfpa) formula is a {\pa} formula containing no quantifiers.
+A {\pa} formula is in \emph{negation normal form} (NNF) if all occurrences of the negation symbol $\neg$ are before the atomic formulas.
+A {\pa} formula is called \emph{existential} if it is in NNF and contains no occurrences of universal quantifiers.
 %An \emph{existential} LIA (ELIA) formula is an LIA formula where there are no universal quantifiers and all existential quantifiers are in the scope of an even number of the negation symbols $\neg$. 

### 2021-05-11 FMCAD2021/3_flatten.tex

@@
 \]
-where $a \in \Sigma_\varepsilon$, $n \in \Nat$, $\Aut$ is an FA, and $\op \in \{=, < , >, \le, \ge\}$.
+where $a \in (\Sigma_{\textit{num}})_\varepsilon$, $n \in \Nat$, $\Aut$ is an FA, and $\op \in \{=, < , >, \le, \ge\}$.
 Let us call $\sterm$ as string terms, $\iterm$ as integer terms, $\sterm = \sterm$ as string equality constraints, $x \in \Aut$ as regular constraints, $\iterm \op\ \iterm$ as arithmetic constraints.
@@
 
+\begin{example}
+some example for $\strint$ here
+\end{example}
+
 %As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
@@
 
-\zhilin{stopped here}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 For a first order theory $T$,
@@
 This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We then introduce a theory {$\paexp$}, denoted by $(\mathbb{N},+,2^x)$, that we work on.
+{\paexp} extends Presburger arithmetic with the exponential function $10^x$ as well as the (partial) functions $\ell_{10}(\ivarx)$ and $\lambda_{10}(\ivarx)$ (cf.
+\cite{Point86}).
+The functions $\ell_{10}(\ivarx)$ and $\lambda_{10}(\ivarx)$ are inductively defined as follows.
+\begin{itemize}
+\item The (partial) function $\ell_{10}(\ivarx)$ is inductively defined as follows: For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$. (Note that $\ell_{10}(0)$ is undefined.) 
+%
+\item The (partial) function $\lambda_{10}(\ivarx)$ can be defined by $\ell_{10}(\ivarx)$: For $n \ge 1$, $\lambda_{10}(n) = 10^{\ell_{10}(n)}$.
+(Intuitively, $\lambda_{10}(\ivarx)$ is the maximum power of 10 that is no greater than $\ivarx$.)
+\end{itemize}
+
+The syntax of {\paexp} is obtained from that of {\pa} by adding $10^\iterm$, $\ell_{10}(\iterm)$ and $\lambda_{10}(\iterm)$ to the definition of integer terms.
+Specifically, {\paexp} formulas are defined by the rules,
+%
+ $$
+ \begin{array}{ l c l}
+ \iterm &::=& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \mid \lambda_{10}(\iterm), \\
+ \phi &::= & \iterm \ \op\ 0 \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi.
+ \end{array}
+ $$
+ The semantics of {\paexp} are defined similarly as {\pa}.
+
+\begin{example}
+Some example of  {\paexp} here.
+\end{example}
+
+%Moreover, we introduce an abbreviation $\lambda_{10}(\ivarx) \equiv 2^{\ell_{10}(\ivarx)}$.
+From the definition of $\lambda_{10}(\ivarx)$, we know that for all $n \ge 1$, $\lambda_{10}(n) \le n \le 10\lambda_{10}(n)-1$ holds.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 \begin{definition}
@@
 which will be useful in our proof. 
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-05-11 FMCAD2021/3_flatten.tex
flattening section

@@
 Let us call $\sterm$ as string terms, $\iterm$ as integer terms, $\sterm = \sterm$ as string equality constraints, $x \in \Aut$ as regular constraints, $\iterm \op\ \iterm$ as arithmetic constraints.
+Let  $\svar(\varphi)$ and $\ivar(\varphi)$ denote the set of string variables and integer variables respectively occurring in $\varphi$.
 
@@
 
+We first recall the flattening approach for string constraints in \cite{Parosh:20:PLDI}, then show how to extend it to deal with the {\parseInt} function.
+
+A \emph{flat domain restriction} for a string constraint $\varphi$ is a function $\flatfun_\varphi$ that maps each string variable $\svarx \in \svar(\varphi)$ to a flat language $(w_{\svarx,1})^* \cdots (w_{\svarx, k_\svarx})^*$, where $w_{\svarx, i} \in \Sigma_{\textit{num}}^+$ for every $i \in [k_\svarx]$.
+%
+The flattened semantics of $\phi \in \strint$ is defined as $\llbracket \varphi \rrbracket_{\flatfun_\varphi} = \{(I_s, I_i) \in \llbracket \varphi  \rrbracket \mid \forall \svarx \in \svar(\varphi).\ I_s(x) \in  \flatfun_\varphi(\svarx)\}$.
+
+The flattening of $\varphi \in \strint$ under a flat domain restriction $\flatfun_\varphi$ is a {\paexp} formula, denoted by $\flatten_{\flatfun_\varphi}(\varphi)$, that encodes its flattened semantics.
+%
+More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and Parikh variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$, called \emph{flattening variables}, plus some other auxiliary variables, such that
+%
+$$\llbracket \phi \rrbracket_\kell =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)})$$
+%
+The decoding function above decodes an interpretation of integer and flattening variables $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \Nat$ as a set $\decode_{\flatfun_\varphi}(I_e)$ of interpretations of the $\varphi$'s integer and string variables $(I_s, I_i)$ with $I_s: \svar(\varphi) \rightarrow \Sigma^*$ and $I_i: \ivar(\phi) \rightarrow \Nat$ such that
+\begin{itemize}
+\item  
+for every $ \svarx \in \svar(\varphi)$, $I_s(\svarx) = w_{\svarx,1}^{I_e(\#_{\svarx,1})} \ldots  w_{\svarx,k_\svarx}^{I_e(\#_{\svarx,k_\svarx})}$, 
+%
+\item for every $ \ivarx \in \ivar(\varphi)$, $I_i(\ivarx) = I_e(\ivarx)$.
+\end{itemize}
+
+The formula $\flatten_{\flatfun_\varphi}(\varphi)$ is constructed inductively by following the structure of $\varphi$: $\flatten_{\flatfun_\varphi}(\varphi_1\ \mathfrak{o}\ \varphi_2) = \flatten_{\flatfun_\varphi}(\varphi_1) \ \mathfrak{o}\  \flatten_{\flatfun_\varphi}(\varphi_2)$, where $\mathfrak{o} \in \{\wedge, \vee\}$, and $\flatten_{\flatfun_\varphi}(\neg \varphi_1) = \neg \flatten_{\flatfun_\varphi}(\varphi_1)$.
+Therefore, it is sufficient to show how to construct $\flatten_{\flatfun_\varphi}(\varphi)$ for atomic constraints $\varphi$.
+In the sequel, we will show how to construct $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ where $\parseInt(\sterm)$ may occur in $\iterm_1$ or $\iterm_2$.
+The construction of $\flatten_{\flatfun_\varphi}(\varphi)$ for the other atomic constraints is essentially the same as that in \cite{Parosh:20:PLDI} and thus omitted.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 The flattening technique was first introduced 
@@
 Hence, we identify a decidable subset of string constraints, which is the largest one with decidability so far to the best of our knowledge.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+Before presenting the construction of $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$, for every $\svarx \in \svar(\varphi)$, we define $\flatten_{\flatfun_\varphi}(\parseInt(\svarx)) = \iterm_{\svarx,1}$  such that $(\iterm_{\svarx, i})_{i \in [k_\svarx]}$ and $(\ell_{\svarx, i})_{i \in [k_\svarx]}$ are inductively defined as follows:
+\begin{itemize}
+\item for $i = k_\svarx$, 
+$$\iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{|w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx} } -1 )}{(10^{|w_{\svarx, k_\svarx}|} -1 )}$$ 
+and $\ell_{\svarx, i} = |w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx}$,
+%
+\item for $i \in [k_\svarx-1]$, 
+%
+$$\iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{|w_{\svarx, i} | \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{|w_{\svarx, i}|} -1 )} + \iterm_{\svarx, i+1}$$
+%
+\and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
+\end{itemize}
+
+Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by the following procedure, 
+\begin{itemize}
+\item Replace each integer term $\len(t)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$ by 
+ $\sum \limits_{i \in [m]} \theta_i$, where for every $i \in [m]$, $\theta_i = 1$ if $\alpha_i \in \Sigma_{\textit{num}}$ and $ \theta_i = \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j} | \#_{\alpha_i, j}$ otherwise.
+%
+\item  Replace each integer term $\parseInt(\sterm)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$ by $\iterm_{\parseInt(\sterm)}$, where $\iterm_{\parseInt(\sterm)} = \iterm_{t,1}$ such that $(\iterm_{t,i})_{i \in [m]}$ and $(\ell_{t, i})_{i \in [m]}$ are inductively defined as follows:
+\begin{itemize}
+\item if $\alpha_m \in \Sigma_{\textit{num}}$, then $\iterm_{t, m} = \alpha_m$ and $\ell_{t, m} = 1$, otherwise, $\iterm_{t, m} = \flatten_{\flatfun_\varphi}(\parseInt(\alpha_m))$ and $\ell_{t, m} = \sum \limits_{j \in [k_{\alpha_m}]} |w_{\alpha_m, j}| \#_{\alpha_m, j}$,
+%
+\item for $i \in [m-1]$, if $\alpha_i \in \Sigma_{\textit{num}}$, then $\iterm_{t, i} = \alpha_i 10^{\ell_{t, i+1}} + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1}+1$, otherwise, $\iterm_{t, i} = 
+\flatten_{\flatfun_\varphi}(\parseInt(\alpha_i))10^{\ell_{t, i+1}}  + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1} + \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j}| \#_{\alpha_i, j}$.
+\end{itemize}
+\end{itemize}
+
+\begin{example}
+\zhilin{this example to be cleaned}
+Take $n={\parseInt}((11)^a(10)^b)$ for example.
+\begin{align}
+    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
+    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
+    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
+\end{align}
+
+So we have 
+\begin{equation}
+    3\cdot n = 3\cdot 2^{2a+2b}-3\cdot 2^{2b}+2\cdot 2^{2b}-2. \notag
+\end{equation}
+\end{example}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 Now, we introduce a new form of atomic string constraints: 
@@
 
-Take $n={\parseInt}((11)^a(10)^b)$ for example.
-\begin{align}
-    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
-    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
-    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
-\end{align}
-
-So we have 
-\begin{equation}
-    3\cdot n = 3\cdot 2^{2a+2b}-3\cdot 2^{2b}+2\cdot 2^{2b}-2. \notag
-\end{equation}
-
 Observe the form of the above equation,
@@
 \end{theorem}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-11 FMCAD2021/5_optimization.tex

@@
 we introduced quantifier elimination as a decision procedure for {$\paexp$}. According to theorem \ref{thm:string-parInt}, the satisfiability of {$\strint$} fragment is decidable.
-However, the quantifier elimination procedure can not be directly applied on solving string constraints over {$\strint$} due to the following two reasons.
+However, it is not practical to directly apply quantifier elimination when solving constraints over {$\paexp$} due to two main reasons.
+
 First, quantifier elimination does not return a model that satisfies the constraints.
 In constrast, it only leaves a formula with constants and free variables.
-And to obtain satisfiability, all free variables need to be treated as existential quantified variables and be eliminated one by one,
-Second, the quantifier elimination procedure is highly complicated.
+To obtain satisfiability, all free variables need to be treated as existential quantified variables and be eliminated one by one until the formula is equal to \textit{true} or \textit{false}.
+Second, the quantifier elimination procedure has a high complexity, both time and space.
 Note that we alternatively invoke QE-exp and Cooper's algorithm to eliminate exponential terms and linear terms of a quantified variable, the length of the formula grows super-exponentially in the procedure.
 
-In this section, we introduced some optimizations on the QE procedure when focus on the quantifier-free fragment in {$paexp$}.
+In this section, we introduced some optimizations on the QE procedure restricted to quantifier-free fragment in {$\paexp$}.
 Given a quantifier-free formula, we wish to decide its satisfiability and find an assignment for variables in the formula if it exists.
-Since the Cooper's algorithm is the main contribution to the complexity, we try to only eliminate exponential terms using an variation of QE-exp and send the obtained linear formula to SMT-solver.
+Since the Cooper's algorithm is the main contribution to the complexity, we try to only eliminate exponential terms using an variation of \textbf{Elim-exp} and send the obtained linear formula to SMT-solver.
 The cost is that the algorithm is no longer complete in some situations, in which case some practical restrictions are needed.
@@
 Similarly,
-constraints that have exponential terms are called exponential constraints, others are called linear constraints.
-
-\paragraph{QE-exp for all atoms}
+constraints that have exponential terms are called \textit{exponential constraints}, others are called \textit{linear constraints.}
 
 \paragraph{over-approximation}
+
+Given any formula in $\paexp$, we first invoke \textbf{Normalization} to translate it into a simpler form.
+Then we  try to detect if there are obvious conflicts, for example, if the linear constraints combined are unsatisfiable, the whole formula is of course unsatisfiable.
+
+Suppose the normalized formula $\theta$ contains variables $\{x_i:i\in[n]\}$, so the exponential terms could be $e^{x_i}$.
+For each $e^{x_i}$, we subtitute $e^{x_i}$ by a fresh variable $y_i$ in $\theta$ and add the constraints $(e-1)x_i+1\le y_i$ to get $\theta'$.
+Since $(e-1)x_i+1\le e^{x_i}$ holds for all $e,x\in \Nat$, $\theta'$ is an over-approximation of $\theta$ and if $\theta'$ is unsatisfiable over positive reals, then $\theta$ is unsatisfiable over $\Nat$.
+
+This technique also helps us to rule out impossible orders of exponential variables. In this case, constraints $\bigwedge_{1\le i\le n-1} x_{\sigma(i)}<x_{\sigma(i+1)}$ is added in $\theta'$.
+
+\paragraph{modified \textbf{Elim-exp}}
+
+\textbf{Elim-exp} is performed recursively on atoms according to orders of all variables.
+The idea is that if the upper bound of linear variables is given, it is enough to consider orders of exponential variables and this step can be performed on all atoms in the formula simultaneously.
+
+Assume we have $m$ exponential variables with an order $x_1\le ...\le x_m$ and $n-m$ linear variables $x_{m+1},...,x_n$.
+We further assume that there is an upper bound for all variables, say $x_{m+1},...,x_n< e^\gamma$, the bound may be inferred from constraints or specified manually.
+When we try to eliminate exponential terms of $x_i(1\le i\le m)$, an atom with $e^{x_i}$ is of the form
+
+\begin{equation}\label{eq:modify-Elim-exp}
+    \sum_{j=1}^i a_j e^{x_j}+\sum_{k=1}^n b_j x_j\le t 
+\end{equation}
+where $a_m\neq 0$ and $t$ is a constant.
+Note that the form of (\ref{eq:modify-Elim-exp}) is different from formula in theorem \ref{thm:exp-ineq}, exponential variables $x_{i+1},...,x_{m}$ may appear linearly in formula (\ref{eq:modify-Elim-exp}) because we do not invoke \textbf{Elim-linear} to eliminate linear terms.
+Following the method of theorem \ref{thm:exp-ineq}, we seperate terms in the left hand side into the following form.
+\begin{equation}
+    \sum_{j=1}^i (a_j e^{x_j} + b_j x_j) \le t -\sum_{k=i+1}^n b_k x_k \notag
+\end{equation}
+
+The first part $\sum_{j=1}^i (a_j e^{x_j} + b_j x_j)$ can be estimated using the same technique in theorem \ref{thm:exp-ineq}.
+
+The problem now reduce to estimate terms of linear variables $\sum_{j=m+1}^{n} b_j x_j$ by $c e^{x_m}$
 

### 2021-05-12 FMCAD2021/3_flatten.tex
section 3

@@
 %
-\and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
+and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
 \end{itemize}
 
-Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by the following procedure, 
-\begin{itemize}
-\item Replace each integer term $\len(t)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$ by 
+Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by the following procedure.
+\begin{enumerate}
+\item Replace each integer term $\len(t)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by 
  $\sum \limits_{i \in [m]} \theta_i$, where for every $i \in [m]$, $\theta_i = 1$ if $\alpha_i \in \Sigma_{\textit{num}}$ and $ \theta_i = \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j} | \#_{\alpha_i, j}$ otherwise.
 %
-\item  Replace each integer term $\parseInt(\sterm)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$ by $\iterm_{\parseInt(\sterm)}$, where $\iterm_{\parseInt(\sterm)} = \iterm_{t,1}$ such that $(\iterm_{t,i})_{i \in [m]}$ and $(\ell_{t, i})_{i \in [m]}$ are inductively defined as follows:
+\item  Replace each integer term $\parseInt(\sterm)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by $\iterm_{\parseInt(\sterm)}$, where $\iterm_{\parseInt(\sterm)} = \iterm_{t,1}$ such that $(\iterm_{t,i})_{i \in [m]}$ and $(\ell_{t, i})_{i \in [m]}$ are inductively defined as follows:
 \begin{itemize}
@@
 \end{itemize}
-\end{itemize}
+%
+\item Let $\iterm'_1 \op \iterm'_2$ be the formula resulting from the aforementioned replacements.
+(Note that strictly speaking, $\iterm'_1 \op \iterm'_2$ is not a $\paexp$ formula since it contains divisions.) Let $N$ be the least common multiplier of the denominators of $\iterm'_1$ and $\iterm'_2$.
+Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained by multiplying the both sides of $\iterm'_1 \op \iterm'_2$ with $N$,  so that the division operator disappears.
+\end{enumerate}
 
 \begin{example}
-\zhilin{this example to be cleaned}
-Take $n={\parseInt}((11)^a(10)^b)$ for example.
-\begin{align}
-    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
-    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
-    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
-\end{align}
-
-So we have 
-\begin{equation}
-    3\cdot n = 3\cdot 2^{2a+2b}-3\cdot 2^{2b}+2\cdot 2^{2b}-2. \notag
-\end{equation}
+Suppose $\parseInt(\svarx) = 2\ivarx$ is an atomic constraint and $\flatfun_\varphi(\svarx) = (12)^*(34)^*$. Then 
+\[
+\small
+\begin{array}{l l}
+& \flatten_{\flatfun_\varphi}(\parseInt(\svarx)  =  2\ivarx)  \\
+\equiv & \parseInt(12) \frac{10^{2\#_{\svarx,1}}-1}{10^2-1}10^{2\#_{\svarx,2}}  + \parseInt(34) \frac{10^{2\#_{\svarx,2}}-1}{10^2-1} = 2\ivarx   \\
+\equiv & 12 (10^{2\#_{\svarx,1}+2\#_{\svarx,2}} - 10^{2\#_{\svarx,2}})  + 34 (10^{2\#_{\svarx,2}}-1) = 198\ivarx.
+\end{array}
+\]
+%Take $n={\parseInt}((11)^a(10)^b)$ for example.
+%\begin{align}
+%    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
+%    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
+%    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
+%\end{align}
 \end{example}

### 2021-05-12 FMCAD2021/5_optimization.tex
sec5: draft

@@
 \textbf{Elim-exp} is performed recursively on atoms according to orders of all variables.
-The idea is that if the upper bound of linear variables is given, it is enough to consider orders of exponential variables and this step can be performed on all atoms in the formula simultaneously.
+The idea is that if the upper bound of variables is given, it is enough to consider orders of exponential variables and this step can be performed on all atoms in the formula simultaneously.
 
@@
 
-The first part $\sum_{j=1}^i (a_j e^{x_j} + b_j x_j)$ can be estimated using the same technique in theorem \ref{thm:exp-ineq}.
+The left hand side $\sum_{j=1}^i (a_j e^{x_j} + b_j x_j)$ can be estimated using the same technique in theorem \ref{thm:exp-ineq}.
+For the right hand side, define  $ t_1 \Def t-\gamma (\sum_{k=i+1}^n b_k)$, $t_2\Def t+\gamma (\sum_{k=i+1}^n b_k)$, we have
+\begin{equation}
+    t_1  \le t - \sum_{k=i+1}^n b_k x_k \le  t_2 \notag
+\end{equation}
 
-The problem now reduce to estimate terms of linear variables $\sum_{j=m+1}^{n} b_j x_j$ by $c e^{x_m}$
+Similar to \textbf{Elim-exp}, we will discuss 3 subcases,
+assuming w.l.o.g $a_i>0$, define $\alpha_1 \Def l(t_1)-l(a_i)$, $\alpha_2 \Def l(t_2)-l(a_i)$.
 
-\paragraph{DFS and pruning}
+\begin{itemize}
+    \item $x_i\le \alpha_1 -1$, corresponds to $\rho_1$ (in \textbf{Elim-exp}), the atom is evaluated \textit{true}
+    \item $\alpha_1 -1\le x_i \le \alpha_2 +2$, corresponds to $\rho_2$ and $\rho_3$
+    \item $\alpha_2 + 2\le x_i$, corresponds to $\rho_4$, the atom is evaluated \textit{false}
+\end{itemize}
 
-Describe the optimizations for improving the scalability
+For each atom with $e^{x_i}$, we compute $\alpha_1$ and $\alpha_2$ according to coefficients in the atom.
+Select the minimal $\alpha_1$ and the maximal $\alpha_2$, in this way we can perform \textbf{Elim-exp} simultaneously on all atoms in the formula.
 
-existential PAexp formula, framework, eliminate exponential terms leaving linear terms, send the (equivalent) linear formula to SMT solver
+Note that the manually specified bound is for variables that only appear linearly in exponential constraints, if there are no variables of this form, the modified algorithm is still complete.
 
-over-approximation for UNSAT and find possible orders
+\paragraph{DFS pruning}
 
-eliminate exponential terms for the formula
+Given an order for exponential variables $x_1\le x_2 \le ...
+\le x_m$, the process of \textbf{Elim-exp} from $x_m$ to $x_1$ can be seen as a search tree with depth $m+1$.
+A depth $i(1\le i\le m)$ non-leaf node is a disjunction, each of its child corresponds to a subcase in \textbf{Elim-exp} of $x_i$, a leaf is labeled by a linear formula in which all exponential terms have been eliminated.
 
-dfs search
+In our implement, we adopt a DFS strategy to search the tree to avoid the explosion of length of the formula.
+When eliminating $x_{i-1}$, the upper bound for $x_i$ in some subcases are passed down for pruning.
+
+\paragraph{pre-search}
+To answer \textit{unsat}, the algorithm must search for all possible orders of exponential variables.
+Intuitively, in most satisfiable cases exponential variables are small, so we first give a small bound (roughly the biggest constants in the formula) for exponential variables to perform a pre-search.

### 2021-05-12 FMCAD2021/1_intro.tex
update intro (onging)

@@
 Logical theories over strings have to allow string concatenation, which is arguably the most fundamental operation of strings. 
-The most celebrated result concerning theories of strings is Makanin’s result on deciding the satisfiability problem for word equations~\cite{}.
+The most celebrated result concerning theories of strings is Makanin’s result on deciding the satisfiability problem for \emph{word equations}~\cite{}.
 A simple example of a word equation is $xaby = ybax$, where $x, y$ are variables, and $a, b \in \Sigma$ are constant letters. 
@@
 The correctness proof of Makanin’s algorithm is arguably one of the most complex termination proofs in computer science. 
-Makanin’s result can be extended to include regular constraints (a.k.a. regular expression matching, e.g., $x \in (ba)^*)$, and arbitrary Boolean connectives.
+Makanin’s result can be extended to include \emph{regular constraints} (a.k.a. regular expression matching, e.g., $x \in (ba)^*)$, and arbitrary Boolean connectives.
 This extension is called word equations with regular constraints. 
+However, the satisfiability problem of word equation together with length constraints (e.g., $|x|=|y|+1 \wedge wx=yx$) is still open.
 The complexity of the satisfiability of word equations with regular constraints was proven to be PSPACE-complete by Plandowski~\cite{}, after decades of improvement of the original algorithm by Makanin.
@@
 
-The satisfiability of string/number conversion constraints has been proven undecidable in~\cite{}. 
-It is undecidable even without combining with other types of string constraints.
+The satisfiability of \emph{string/number conversion constraints} \yfc{give example} has been proven undecidable in~\cite{}. 
+It is undecidable even without combining with other types of string constraints. \yfc{double check}
 However, this kind of constraint is pervasive in scripting language programs. 
@@
 
-strategy: do unsat and sat separately
+Due to the difficulties in solving string constraints and, in practice, satisfiable string constraints are more critical for automatic testing, one idea is to have separate specialized procedures for solving satisfiable instances.
+Currently, there are two main specialized approaches for proving satisfiability.
+The first is to consider only strings of bounded length.
+This approach is taken in earlier generation solvers such as Hampi and Kuluza~\cite{}.
+Although they are useful in handling many practical cases, they fail short when the all string solutions exceed the selected bound.
+For example, a constraint of the form $x.y \neq z  \wedge |x| > 2000$ would be quite challenging to handle using those solvers.
 
-strategy: using different procedure to (dis)prove validaity
+One more recent approach is flattening~\cite{}.
+The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat automata} (see Section~\ref{PA} for more details). 
+The major benefit of considering this class is two folds.
+First, even under the restriction, often the potential solution space is still infinite, which gives us a higher potential of finding solutions.
+For instance, we can find a solution for $x.y \neq z  \wedge |x| > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is a character in our alphabet.
+Second, and more importantly, because we can convert the membership problem of a flat automaton to a satisfiability problem of Presburgh formula, the class of word equation + regular constraints + length constraints is decidable.
 
-for disprove validatity, there are two approches, first is bound string length
+The paper of Abdulla et al. has considered adding string/number conversion constraints to the above class.
+The paper proposes a decision procedure with a restricted form of flat automata.
+For string/number conversion, their approach projects the solution to string/number conversion to, in terms of numerical solutions, a finite solution space.
+In this paper, we propose a complete solution to this problem.
+We propose a decision procedure for the class of word equation + regular constraints + length constraints + string/number conversion under arbitrary flat automata.
 
-cannot handle $x.y \neq z  \wedge |x| > 2000$
+How we did it? Also mention the quantifier elimination procedure. Optimizations
 
-more recent approach is flattening
+Say the problem is challenging for state-of-the-art solvers, mention the hash function example.
 
-it is known that word equaltion + flat regular constraints + len constraints
-is decidable
-
-it was unknown that whether word equations + flat regular constraints + len constraints + parseInt is decidable
+Experimental results.

### 2021-05-12 FMCAD2021/4_dec_proc.tex
section 4

@@
 
+Sem\"{e}nov first proved that  {\paexp} admits quantifier elimination in \cite{Semenov84}, thus its satisfiability problem is decidable.
+However, Sem\"{e}nov did not give a concrete quantifier elimination algorithm.
+Remedying this, Point proposed a quantifier elimination algorithm for {\paexp} in \cite{Point86}.
+In the sequel, we are going to illustrate how Point's quantifier elimination algorithm works.
+Points' algorithm is non-elementary in the worst case and a faithful implementation of the algorithm would not scale\footnote{We did implemented Point's algorithm and discovered that the implementation could only solve formulas of very small size.}.
+In the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
+
+Specifically, in the sequel, we are going to show that for every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
+% such that $\free(\varphi') = \free(\varphi) \setminus \{\ivarx\}$. 
+%From this, we can easily derive that every {\paexp} formula $\varphi$ can be transformed effectively into an equivalent quantifier-free Presburger arithmetic formula $\varphi'$.
+
+Before a formal description of the quantifier elimination procedure, let us use a simple example to illustrate the main idea and give an overview of the procedure.
+
+%\begin{example}
+\subsection{Overview of the quantifier elimination procedure}
+Consider $\varphi \Def \exists \ivarx_2.\ 10^{\ivarx_1 + \ivarx_2} - 10^{\ivarx_2} \le \ivary + 1001$.
+
+At first, we \emph{normalize} $\varphi$ by introducing a fresh variable $\ivarx_3$ for $\ivarx_1 + \ivarx_2$ and get the formula 
+$$\varphi_1 \Def \exists \ivarx_3 \exists \ivarx_2.\ 10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001 \wedge \ivarx_3 = \ivarx_1 + \ivarx_2.$$
+
+Then, we enumerate different \emph{orders} of the quantified variables, i.e.
+$\ivarx_2$ and $\ivarx_3$.
+Since $\ivarx_3 = \ivarx_1 + \ivarx_2$, there is only one possible order, that is, $\ivarx_3 \ge \ivarx_2$.
+
+Next, we illustrate how to eliminate the quantifier $\exists \ivarx_3$, assuming $\ivarx_3 \ge \ivarx_2$. The elimination of $\exists \ivarx_2$ is similar and simpler, thus omitted.
+
+The elimination of $\exists \ivarx_3$ consists of two steps, namely, eliminating the exponential occurrences of $\ivarx_3$ first, and the linear occurrences next.
+
+The main idea of the elimination of the exponential occurrences of $\ivarx_3$ is to observe that if $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$, then $10^{\ivarx_3} - 10^{\ivarx_2}$ is dominated by $10^{\ivarx_3}$, that is, $10^{\ivarx_3} - 10^{\ivarx_2} \ge 10^{\ivarx_3} - 10^{\ivarx_3 - 3} = (1-10^{-3}) 10^{\ivarx_3} \ge 10^{\ell_{10}(\ivary+1001)+1} = 10\lambda_{10}(\ivary+1001) > \ivary+1001$.
+Therefore, a necessary condition for $10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001$ is that either $\ivarx_3 \le \ell_{10}(\ivary+1001)+1$ or $\ivarx_3 \le \ivarx_2 + 2$ holds.
+\begin{itemize}
+\item If $\ivarx_3 \le \ell_{10}(\ivary+1001)+1$, then we distinguish between whether $\ivarx_3 \le \ell_{10}(\ivary+1001)$ or  $\ivarx_3 = \ell_{10}(\ivary+1001)+1$. 
+\begin{itemize}
+\item If $\ivarx_3 \le \ell_{10}(\ivary+1001)$, then $10^{\ivarx_3} - 10^{\ivarx_2} \le 10^{\ell_{10}(\ivary+1001)} = \lambda_{10}(\ivary+1001) \le \ivary + 1001$.
+In this case, $10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001 \wedge \ivarx_3 = \ivarx_1 + \ivarx_2$ can simplified into $\ltrue$.
+%
+\item If $\ivarx_3 = \ell_{10}(\ivary+1001)+1$, then $10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001$ can be turned into $10^{\ell_{10}(\ivary+1001)+1} - 10^{\ivarx_2} \le \ivary + 1001 \equiv 10 \lambda_{10}(\ivary+1001) - 10^{\ivarx_2} \le \ivary + 1001 $.
+\end{itemize} 
+%
+\item If $\ivarx_3 \le \ivarx_2 + 2$, then $\ivarx_3 = \ivarx_2 + j$ for $j \in \{0,1,2\}$.
+Thus $10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001$ can be transformed to $\bigvee \limits_{j \in \{0,1,2\}} 10^{\ivarx_2 + j} - 10^{\ivarx_2} \le \ivary + 1001$.
+\end{itemize}
+
+To summarize, $\varphi_1$ is transformed into 
+\[
+\small
+\begin{array}{l}
+\varphi_2 \Def \exists \ivarx_3 \exists \ivarx_2. \\
+\begin{array}{l}
+\vspace{2mm}
+\left(
+\begin{array}{l}
+\ivarx_3 \le \ell_{10}(\ivary+1001)\ \vee \\
+\left(
+\begin{array}{l}
+\ivarx_3 = \ell_{10}(\ivary+1001)+1\ \wedge \\
+10 \lambda_{10}(\ivary+1001) - 10^{\ivarx_2} \le \ivary + 1001 
+\end{array}
+\right) \vee \\
+%
+ \bigvee \limits_{j \in \{0,1,2\}}  \left(\ivarx_3 = \ivarx_2 +j \wedge 10^{\ivarx_2 + j} - 10^{\ivarx_2} \le \ivary + 1001\right)
+\end{array}
+\right) 
+\wedge \\
+ \ivarx_3 = \ivarx_1 + \ivarx_2.
+ \end{array}
+\end{array}
+\]
+Note that $\varphi_2$ contains \emph{only linear} occurrences of $\ivarx_3$.
+
+Finally, we can eliminate the linear occurrences of $\ivarx_3$, thus the quantifier $\exists \ivarx_3$, by applying the quantifier elimination algorithm of {\pa}, e.g.
+Cooper's algorithm in \cite{Cooper72}.
+The elimination of $\ivarx_3$ in $\varphi_2$ is simple, with $\ivarx_3$ replaced by $\ivarx_1 + \ivarx_2$, resulting into
+\[
+\small
+\begin{array}{l}
+\varphi_3 \Def \exists \ivarx_2.\ 
+\ivarx_1 + \ivarx_2 \le \ell_{10}(\ivary+1001)\ \vee \\
+\left(
+\begin{array}{l}
+\ivarx_1 + \ivarx_2 = \ell_{10}(\ivary+1001)+1\ \wedge \\
+10 \lambda_{10}(\ivary+1001) - 10^{\ivarx_2} \le \ivary + 1001 
+\end{array}
+\right) \vee \\
+%
+ \bigvee \limits_{j \in \{0,1,2\}}  \left(\ivarx_1 + \ivarx_2 = \ivarx_2 +j \wedge 10^{\ivarx_2 + j} - 10^{\ivarx_2} \le \ivary + 1001\right).
+\end{array}
+\]
+
+In the remaining part of this section, we are going to describe the steps of the decision procedures, namely, the normalization, the enumeration of the orders of variables, 
+the elimination of the exponential occurrences of variables, and the elimination of the  linear occurrences.
+
+\zhilin{stopped here}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 \begin{definition}
@@
 A more detailed description is given below.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-05-12 FMCAD2021/4_dec_proc.tex
normalization

@@
 
-In the remaining part of this section, we are going to describe the steps of the decision procedures, namely, the normalization, the enumeration of the orders of variables, 
-the elimination of the exponential occurrences of variables, and the elimination of the  linear occurrences.
+\smallskip
 
-\zhilin{stopped here}
+In the remainder of this section, we are going to describe the aforementioned steps of the decision procedures: Normalization, the enumeration of the variable orders, 
+the elimination of the exponential occurrences of variables. The elimination of the  linear occurrences of variables is essentially the quantifier elimination of the {\pa} and omitted.
+
+Let us assume that $\varphi \Def \exists \ivarx.\ \varphi'(\ivarx, \vec{\ivary})$, where $\varphi'$ is a quantifier-free formula.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 In order to show $(\mathbb{N},+,2^x)$ admits QE , 
@@
 $$\exists x\exists x_1. x_1 > 3 \wedge 2^{x_1}\le x \wedge x\le 2^{x_1+1}-1\,.$$
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The \textbf{Normalization} step goes like this, first transform the given formula $\theta(x,\bar{y})$ into the Negation Normal Form (NNF).
+The normalization step comprises the following sub-steps.
+\begin{enumerate}
+\item At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form).
+Moreover, we remove the occurrences of $\neg$ by replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, $\neg \iterm_1 = \iterm_2$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, $\neg \iterm_1 < \iterm_2$ with $\iterm_2 \le \iterm_1$, $\neg \iterm_1 \le \iterm_2$ with $\iterm_2 < \iterm_1$, and so on.
+%
+\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ where $\ivarx$ occurs: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
+Let the resulting formula be $\varphi''$.
+%
+\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ where $\ivarx$ occurs, we have $\iterm = \ivarx$: For each occurrence of the $\ell_{10}$ in $\varphi''$, say $\ell_{10}(\iterm)$, such that it contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$,
+%and let $\varphi''':= \exists \ivarz.\ \varphi''[\ivarz/\ell_{10}(\iterm)] \wedge 10^\ivarz \le \iterm < 10^{\ivarz + 1}$. 
+and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct. Let $\varphi'''$ denote the resulting formula.  
+%
+\item Do the following replacements to $\varphi'''$ so that all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
+$\iterm_1 > \iterm_2$) with $\iterm_1 \le \iterm_2 - 1$ (resp.
+$\iterm_2 \le \iterm_1-1$).
+Replace ever occurrence of $\iterm_1 = \iterm_2$ wtih $\iterm_2 \le \iterm_1 \wedge \iterm_1 \le \iterm_2$.
+Let $\varphi^\dag$ the resulting formula.
+%
+\item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_k$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
+$\exists \vec{\ivarz}\exists \ivarx.\ \varphi^\dag$.
+\end{enumerate}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 Then $\theta(x,\bar{y})$ becomes a Boolean combination (with only $\wedge$ and $\vee$) of literals.
@@
 \end{algorithm}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection{Enumerate-Orders} 
+\subsection{Enumeration of the variable orders}
+
+\zhilin{stopped here}
+
 In the \textbf{Normalization} step, we simplify the origin formula at the cost of more quantified variables.

### 2021-05-13 FMCAD2021/4_dec_proc.tex
order enumeration

@@
 %
-\item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_k$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
+\item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_n$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
 $\exists \vec{\ivarz}\exists \ivarx.\ \varphi^\dag$.
@@
 
-\zhilin{stopped here}
+Let the output of the normalization procedure be $\exists \vec{\ivarx}.\ \varphi'$ with $\vec{\ivarx} = (\ivarx_1,\ldots, \ivarx_n)$. 
+We then enumerate all the linear orders of $\{\ivarx_1,\ldots, \ivarx_n\}$.
+Each linear order can represented by a permutation $\sigma \in \mathcal{S}_n$ (where $\mathcal{S}_n$ is the permutation group on $[n]$), with the intention that $\ivarx_{\sigma(n)} \ge \cdots \ge \ivarx_{\sigma(1)}$.
 
+Assuming a linear order $\sigma \in \mathcal{S}_n$ of $\{\ivarx_1,\ldots, \ivarx_n\}$, we then consider $\varphi'_\sigma  = \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ and eliminate the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(1)}$,  one by one and from left to right.
+Let $\varphi''_\sigma$ denote the resulting formula.
+
+Finally, $\exists \vec{\ivarx}.\ \varphi'$ is transformed into the quantifier-free formula $\bigvee \limits_{\sigma \in \mathcal{S}_n} \varphi''_{\sigma}$.
+
+In the sequel, assuming a linear order $\sigma \in \mathcal{S}_n$, for $i \in [n]$, let $\exists \ivarx_{\sigma(i)} \ldots \exists \ivarx_{\sigma(1)}.\ \varphi''_{\sigma,i}$ be the formula obtained from $\varphi'_\sigma$ by eliminating the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(i+1)}$, we show how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\exists \ivarx_{\sigma(i)} \ldots \exists \ivarx_{\sigma(1)}.\ \varphi''_{\sigma,i}$.
+We would like to remark that the linear occurrences of $\ivarx_{\sigma(i)}$ should be eliminated further so that the quantifier $\exists \ivarx_{\sigma(i)}$ can be eliminated.
+The elimination of linear occurrences of $\ivarx_{\sigma(i)}$ is essentially the quantifier elimination algorithm of {\pa}.
+%$\varphi'_\sigma =  \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 In the \textbf{Normalization} step, we simplify the origin formula at the cost of more quantified variables.
@@
 \end{algorithm}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection{Elim-exp}
+\subsection{Elimination of  exponential occurrences of variables}\label{sec-elim-exp}
 

### 2021-05-13 FMCAD2021/6_impl_exp.tex
pruning opt
opt draft
implement ongoing

--- 2021-05-07 FMCAD2021/7_conclusion.tex
@@
 
-Conclude this paper and discuss future work.
+We have implemented a prototype of our algorithm in Wolfram Mathematica, called  $\paexp$-Solver.
+Any postive Integer is allowed to be the base of exponential functions in our algorithm, but $\parseInt$ in Z3 and CVC4 only takes a numeric string as input.
+For this reason, we fix the exponential function to be $10^x$ in our experiments.
+
+Our solver takes an existential quantified {$\paexp$} formula together with an upper bound for variables.
+As we have mentioned in \textbf{modified Elim-exp}, if there is no variables only occurs linearly in exponential constraints, the upper bound can be omitted, otherwise it needs to be specified manually.
+So for arithmetic constraints, we have generated benchmarks of two forms, named \textbf{BOUND} and \textbf{UNBOUND}, depends on whether the bound is needed.
+
+The outputs of our solver can be SAT, UNSAT, B-UNSAT (short for bounded unsat) or TIMEOUT.
+For satisfiable cases, if a satisfying assignment is found, our solver returns ``SAT" together with the model.
+For unsatisfiable cases, our solver will decide whether the manually specified bound for variables is needed.
+The output ``UNSAT" means the formula is unsatisfiable over $\Nat$, while ``B-UNSAT" means no solution within the range.
+``TIMEOUT" will be returned if the computation excesses the time limit.
+
+We are not aware of any SMT solver or constraints solver that directly support arithmetic formulas with exponential functions over Integer domain.
+In Z3 and CVC4, we need to define exponential functions as recursive functions.
+Whereas string constraints with $\parseInt$ constraints are supported in Z3, CVC4 and other string constraints solvers like Trau, HAMPI etc.
+Therefore, we designed two benchmarks to evaluate our $\paexp$-Solver.
+The first benchmark is pure arithmetic formulas with exponential functions generated randomly, we compared our solver with Z3 and CVC4.
+The second benchmark is string constraints about hash functions, we tranlated them to arithmetic constraints for our solver and compared the result with Z3, CVC4 and Trau, who treated this benchmark as just string constraints.
+The details of the two benchmarks will be illustrate following:
+
+\paragraph{arithmetic}
+
+We have 8 groups of experiments for arithmetic constraints in all, including 4 groups for \textbf{BOUND} case and 4 for \textbf{UNBOUND} case.
+
+Each group consists of 100 trails paratrized by 4 natuaral number parameters, which in turn represent the number of exponential variables, linear variables, exponential inequalities and linear inequalities.For \textbf{BOUND} case, assume $m$ exponential variables and $(n-m)$ linear variables, an exponential inequalities is of the form
+$$\sum_{i=1}^m a_i 10^{x_i} + \sum_{j=1}^n b_j x_j \le c$$
+
+For \textbf{UNBOUND} case, we further restrict in the exponential inequalities, $a_i\neq 0$ for $i:1\le i\le m$ and $b_j=0$ for $j:m+1\le j \le n$.
+
+For both cases, we always require the number of constraints is more than the number variables.
+It is reasonable because the coefficients are randomly generated and some constraints may become inactive.
+We also have 2 groups include equalities by changing the first exponential inequality into equality.
+
+Coefficients $a_i$, $b_j$ and $c$ are generated randomly for each constarint within the same range.
+We set $a_i\in [-10,10]$, and $b_j,c\in[-10^5,10^5]$.
+$a_i$ is set smaller so that exponential terms will not always dominate the left hand side of an inequality.
+
+\paragraph{string hash function}
+ flat pattern and arbitrarys pattern
+
+All experiments were run on an Intel Core i5 1.4GHz with 8GB RAM.
+
+In our benchmark, the coefficients in inequalities are generated randomly within a range.
+As a consequence, some inequalities may not actually work, for example, $-10^{x_1} - x_1 \le 1$.
+So we always set the number of exponential  inequalities is one more than number of exponential variables, and the same for the linear comterpart.
+Since exponential terms with large coefficients are more likely to dominate the summation, coefficients of exponential terms are in the range $[-10,10]$ while other coefficients and constarints are generated randomly in $[-10^5,10^5]$.
+
+For each

### 2021-05-13 FMCAD2021/4_dec_proc.tex
section 4

@@
 
+Let $i \in [n]$ and $\exists \ivarx_{\sigma(i)} \ldots \exists \ivarx_{\sigma(1)}.\ \varphi''_{\sigma,i}(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ be the formula obtained from $\varphi'_\sigma$ by eliminating the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(i+1)}$.
+We show how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\varphi''_{\sigma,i}$.
+The elimination is \emph{local} in the sense that it is applied to the atomic formulas independently.
+
+Recall that after normalization, the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+Therefore, we can assume that the atomic formulas in $\varphi''_{\sigma,i}$ are  of the form
+%
+$$a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary})$$
+or  
+$$c \big | \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big).$$
+%
+For convenience, let us call these formulas as inequality respectively divisibility atomic formulas.
+In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these two types of atomic formulas.
+
+\medskip
+
+\paragraph{Inequality atomic formulas} Let us consider 
+$$
+\begin{array}{l}
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
+\hspace{4mm} 
+a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
+\end{array}
+$$
+
+The elimination of the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ relies on the following lemma.
+Intuitively, the lemma states the fact that if $\ivarx_{\sigma(i)}$ is sufficiently greater than $\ivarx_{\sigma(i-1)}$, then the left-hand-side of $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is \emph{dominated} by $a_i10^{\ivarx_{\sigma(i)}}$, moreover, if $a_i > 0$ and the value of $\ivarx_{\sigma(i)}$ is sufficiently small (resp.
+big), then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds (resp.
+does not hold), similarly for $a_i < 0$.
+
+\begin{lemma} \label{lem:exp-ineq}
+Let  
+%
+$$
+\begin{array}{l}
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
+\hspace{4mm} 
+a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
+\end{array}
+$$
+%
+with $a_i \neq 0$, $A\Def \sum_{j=1}^{i-1}|a_j|$, 
+%$B\Def  \sum_{j=1}^{i}|b_j|$, 
+$B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
+and $\delta\Def  \ell_{10}(A)+3$. 
+\begin{itemize}
+    \item If $a_i > 0$, let $\alpha(\vec{\ivary}) \Def \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$, then 
+    \begin{itemize}
+        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds,
+        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)}  \ge \ivarx_{\sigma(i-1)} +\delta$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold.
+    \end{itemize}
+    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-(a_i))$, then 
+    \begin{itemize}
+        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold,
+        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
+    \end{itemize}
+\end{itemize}
+\end{lemma}
+
+If $a_i > 0$, then from Lemma~\ref{lem:exp-ineq}, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
+\[
+\small
+\begin{array}{l}
+\bigvee \limits_{p=0}^{B-1} a_i 10^{p}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_i p + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}) \\
+\bigvee \big(\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1  \wedge \ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta \big)\\
+\bigvee 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} = \alpha(\vec{\ivary})\ \wedge \\
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\alpha(\vec{\ivary})/\ivarx_{\sigma(i)}]
+\end{array}
+\right)  \\
+\bigvee 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} = \alpha(\vec{\ivary})+1\ \wedge \\
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\alpha(\vec{\ivary})+1/\ivarx_{\sigma(i)}]
+\end{array}
+\right)  \\
+\bigvee \bigvee \limits_{p=0}^{\delta-1} 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2\ \wedge \\
+ \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\ivarx_{\sigma(i-1)}+ p /\ivarx_{\sigma(i)}]
+\end{array}
+\right),
+\end{array}
+\]
+where exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+
+The elimination of the exponential occurrences of $\ivarx_{\sigma(i)}$ for the case $a_i < 0$ is similar.
+
+\medskip
+
+\paragraph{Divisibility atomic formulas} Consider
+%
+$$
+\begin{array}{l}
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
+c\ \big |\ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
++ \iterm(\vec{\ivary}) \big)
+\end{array}
+$$
+with $a_i \neq 0$.
+%
+
+Let $c = 2^{r_1} 5^{r_2}  c_0$ such that $c_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $c | (10^rc_0)$.
+
+If $c_0 = 1$, then $10^r$ is divisible by $c = 2^{r_1}5^{r_2}$.
+Thus for every $n \ge r$, $c \ |\ 10^n$.
+Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to
+\[
+\small
+\begin{array}{l}
+\bigvee \limits_{p = 0}^{r - 1} c\ \big | \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
++ \iterm(\vec{\ivary}) \big)\\
+%
+\vee \big(\ivarx_{\sigma(i)} \ge r \wedge c\ \big | \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
++ \iterm(\vec{\ivary}) \big)\big),
+\end{array}
+\]
+where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+
+Next, let us assume $c_0 > 1$.
+Since $10$ and $c_0$ are relatively prime, according to Euler's theorem (cf.
+\cite{HW80}), $10^{\phi(c_0)} \equiv 1 \bmod c_0$, where $\phi$ is the Euler function.
+Suppose $10^{\phi(c_0)} = kc_0 + 1$ for some $k \in \Nat$.
+Then for every $n \in \Nat$ with $n \ge r$, 
+$$
+\begin{array}{l}
+10^{n + \phi(c_0)} \bmod c =10^{n-r} 10^r (kc_0 + 1) \bmod c = \\
+10^{n-r} (k 10^rc_0 + 10^r) \bmod c = \\
+10^{n-r} (0+10^r) \bmod c = 10^n \bmod c.
+\end{array}
+$$
+
+Then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
+\[
+\begin{array}{l}
+\bigvee \limits_{p=0}^{r-1} \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[p/\ivarx_{\sigma(i)}]\ \vee \\
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i)} \ge r\ \wedge \\
+\bigvee \limits_{q = 0}^{\phi(c_0)-1} 
+\left(
+\begin{array}{l}
+\phi(c_0)\ \big |\ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
+c\ \big | 
+\left(
+\begin{array}{l}
+a_i 10^{r+q} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \\
+\sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})
+\end{array}
+\right) 
+\end{array}
+\right)
+\end{array}
+\right),
+\end{array}
+\]
+where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
+Let $d = 2^rd_0$, $d_0$ is an odd natural number. 
+According to Euler's Theorem, 
+$gcd(2,d_0)=1$ implies
+$2^{\phi(d_0)} \text{mod}\ d_0 = 1$,
+where $\phi$ is the Euler function.
+Consider function $f(n)=(2^n\ \text{mod} \ d)$,
+when $n\ge r$, 
+$f(n)$ becomes a periodic function and its period is a divisor of $\phi(d_0)$ because
+\begin{align}
+    f(n+\phi(d_0)) 
+    &=2^{n+\phi(d_0)}\ \text{mod}\ d &\notag\\
+    &=2^n\cdot 2^{\phi(d_0)}\ \text{mod}\ d &\notag\\
+    &=2^n\cdot (k d_0 + 1) \ \text{mod}\ d &
+    (\mbox{assume } 2^{\phi(d_0)}=k d_0 +1) \notag \\
+    &= 2^n \ \text{mod}\ d &
+    (\mbox{when } n\ge r, 2^n d_0\ \text{mod}\ d = 0 )\notag\\
+    &= f(n) \notag
+\end{align}
+
+When $x_i\le r-1$, we just enumerate all possible value of $x_i$, i.e., 
+$$\rho_5 \Def \bigvee_{p=0}^{r-1} \tau(p,\bar{x},\bar{y})\,.$$
+When $x_i\ge r$, $\tau(x_i,\bar{x},\bar{y})$ is equivalent to
+$$
+\rho_6 \Def \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
++  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r] \,.
+$$
+            Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.
+
 \textbf{Elim-exp} is the most technical part of our QE algorithm.
@@
             Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 \subsection{Elim-linear}
@@
 the formula other than NNF.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-05-13 FMCAD2021/4_dec_proc.tex
section 4
some change on sec2

@@
 We then enumerate all the linear orders of $\{\ivarx_1,\ldots, \ivarx_n\}$.
-Each linear order can represented by a permutation $\sigma \in \mathcal{S}_n$ (where $\mathcal{S}_n$ is the permutation group on $[n]$), with the intention that $\ivarx_{\sigma(n)} \ge \cdots \ge \ivarx_{\sigma(1)}$.
+Each linear order can be represented by a permutation $\sigma \in \mathcal{S}_n$ (where $\mathcal{S}_n$ is the permutation group on $[n]$), with the intention that $\ivarx_{\sigma(n)} \ge \cdots \ge \ivarx_{\sigma(1)}$.
 
@@
 \]
-where exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
 
@@
 
-%\znj{I think there are some mistakes in the above proof, pls check it!}
 When $x_i \le \alpha-1$, similarly we have

### 2021-05-14 FMCAD2021/4_dec_proc.tex
section 4

@@
 %$B\Def  \sum_{j=1}^{i}|b_j|$, 
-$B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
+$B \Def 10(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
 and $\delta\Def  \ell_{10}(A)+3$. 
@@
 
-If $a_i > 0$, then from Lemma~\ref{lem:exp-ineq}, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
+If $a_i > 0$, then the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ can be eliminated by utilizing  Lemma~\ref{lem:exp-ineq} and enumerating the constraints on $\ivarx_{\sigma(i)}$ and $\ivarx_{\sigma(i-1)}$.
+Specifically,
+$\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
 \[
@@
 \bigvee \big(\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1  \wedge \ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta \big)\\
+%
+\bigvee \bigvee \limits_{p=0}^{\delta-1} 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1 \ \wedge \\
+ \ivarx_{\sigma(i)} = \ivarx_{\sigma(i-1)} +p\ \wedge\\
+ \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\ivarx_{\sigma(i-1)} + p /\ivarx_{\sigma(i)}] 
+\end{array}
+\right)\\
 \bigvee 
@@
 \begin{array}{l}
-\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2\ \wedge \\
+\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2\ \wedge\\
+ \ivarx_{\sigma(i)} = \ivarx_{\sigma(i-1)} +p\ \wedge \\
  \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\ivarx_{\sigma(i-1)}+ p /\ivarx_{\sigma(i)}]

### 2021-05-14 FMCAD2021/4_dec_proc.tex
section 4

@@
 
-Specifically, in the sequel, we are going to show that for every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
+Specifically, in the sequel, we are going to show that every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
 % such that $\free(\varphi') = \free(\varphi) \setminus \{\ivarx\}$. 
@@
 %\begin{example}
-\subsection{Overview of the quantifier elimination procedure}
+\subsection{An overview of the quantifier elimination procedure}
 Consider $\varphi \Def \exists \ivarx_2.\ 10^{\ivarx_1 + \ivarx_2} - 10^{\ivarx_2} \le \ivary + 1001$.

### 2021-05-14 FMCAD2021/8_appendix.tex
appendix

%!TEX root = paper.tex

\noindent {\bf Lemma~\ref{lem:exp-ineq}}.
\emph{Let  
%
$$
\begin{array}{l}
\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
\hspace{4mm} 
a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
\end{array}
$$
%
with $a_i \neq 0$, $A\Def \sum_{j=1}^{i-1}|a_j|$, 
%$B\Def  \sum_{j=1}^{i}|b_j|$, 
$B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
and $\delta\Def  \ell_{10}(A)+3$. 
\begin{itemize}
    \item If $a_i > 0$, let $\alpha(\vec{\ivary}) \Def \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$, then 
    \begin{itemize}
        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds,
        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)}  \ge \ivarx_{\sigma(i-1)} +\delta$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold.
    \end{itemize}
    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-(a_i))$, then 
    \begin{itemize}
        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold,
        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
    \end{itemize}
\end{itemize}
}

We need the following proposition for the proof of Lemma~\ref{lem:exp-ineq}.

\begin{proposition} \label{prop:1}
If $n\ge m\ge 1$ and $p \ge 10(\ell_{10}(n)- \ell_{10}(m)+1)$, then 
$n p \le m10^p$ holds.
\end{proposition}

\begin{proof}
By an induction on $n'$, we first show that for any $n' \in \Nat$, if $p \ge 10n'$, then $10^{n'} p \le 10^p$.

If $ p \ge 10 (\ell_{10}(n) - \ell_{10}(m)+1)$, then  $n p \le 10\lambda_{10}(n) p = 10 * 10^{\ell_{10}(n)} p = 10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p$.

Because $p \ge 10(\ell_{10}(n)-\ell_{10}(m) + 1)$, we deduce that $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  p \le 10^p$.  Therefore, $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p \le 10^{\ell_{10}(m)} 10^p \le m 10^p$.
We conclude that $np \le m 10^p$.
\end{proof}

%Then we give the proof for Theorem~\ref{thm:exp-ineq}.
\begin{proof}[Proof of Lemma~\ref{lem:exp-ineq}]
We only prove for the case $a_i > 0$, the other case is symmetric.

Let $A\Def \sum_{j=1}^{i-1}|a_j|$, $B \Def 10(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$, and $\delta\Def  \ell_{10}(A)+3$.

Suppose $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$. Moreover, let $\alpha(\vec{\ivary}) \Def  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$.

Note that
 \begin{equation} 
   A10^{-\delta} =A 10^{-\ell_{10}(A)-3} = \frac{A}{1000\lambda_{10}(A)} \le \frac{1}{100}.   \label{eq:thm-ineq-1}
 \end{equation}

 From $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$ and $\ivarx_{\sigma(i-1)} \ge \ldots \ge \ivarx_{\sigma(1)}$, 
we know   
$$-A 10^{\ivarx_{\sigma(i)} - \delta} \le  \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} \le A 10^{\ivarx_{\sigma(i)} - \delta}$$
and
$$-(\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)} \le \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} \le (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)}.$$

Moreover, let $n = 100\sum_{j=1}^{i}|b_j|$, $m = 1$, and $p = \ivarx_{\sigma(i)}$, then 
$n \ge m \ge 1$. From $10(\ell_{10}(n)- \ell_{10}(m)+1) = 10 (\ell_{10}(100 \sum_{j=1}^{i}|b_j|) + 1) = 10 (\ell_{10}(\sum_{j=1}^{i}|b_j|) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 10(\ell_{10}(n)- \ell_{10}(m)+1)$.
Then according to Proposition~\ref{prop:1}, 
$$100(\sum_{j=1}^{i}|b_j|)\ivarx_{\sigma(i)} = np \le m 10^p = 10^{\ivarx_{\sigma(i)}}.$$
Thus 
$(\sum_{j=1}^{i}|b_j|)\ivarx_{\sigma(i)}  \le \frac{1}{100} 10^{\ivarx_{\sigma(i)}}.$

If $\ivarx_{\sigma(i)} \ge   \alpha(\vec{\ivary}) + 2$, then

$
\begin{array}{l}
a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \ge \\
a_i 10^{\ivarx_{\sigma(i)}} - A 10^{\ivarx_{\sigma(i)} - \delta} - (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)}  \ge \\
a_i 10^{\ivarx_{\sigma(i)}} - \frac{1}{100}10^{\ivarx_{\sigma(i)}} - \frac{1}{100} 10^{\ivarx_{\sigma(i)}} = \\
(a_i - \frac{1}{50}) 10^{\ivarx_{\sigma(i)}} \ge (a_i - \frac{1}{50}) 10^{ \alpha(\vec{\ivary}) + 2} = \\
(a_i - \frac{1}{50}) 10^{  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i) + 2} = \\
\frac{10(a_i - \frac{1}{50})}{10^{\ell_{10}(a_i)} } 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \frac{10(a_i - \frac{1}{50})} {a_i} 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \\
(10 - \frac{1}{5a_i}) 10^{ \ell_{10}(\iterm(\ivary))+1} > 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \iterm(\ivary).
\end{array}
$

Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \emph{does not} hold.

If $\ivarx_{\sigma(i)} \le   \alpha(\vec{\ivary}) - 1$, then

$
\begin{array}{l}
a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \\
a_i 10^{\ivarx_{\sigma(i)}} + A 10^{\ivarx_{\sigma(i)} - \delta} + (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)} \le \\
a_i 10^{\ivarx_{\sigma(i)}} + \frac{A} {10^\delta}10^{\ivarx_{\sigma(i)}} + \frac{1}{100} 10^{\ivarx_{\sigma(i)}} \le \\
(a_i + \frac{1}{100} + \frac{1}{100})10^{\ivarx_{\sigma(i)}} = (a_i + \frac{1}{50}) 10^{\ivarx_{\sigma(i)}} \le \\
(a_i + \frac{1}{50}) 10^{\alpha(\vec{\ivary}) - 1} = (a_i + \frac{1}{50}) 10^{\ell_{10}(\iterm(\ivary))- \ell_{10}(a_i) - 1} = \\
\frac{a_i + \frac{1}{50}} {10^{\ell_{10}(a_i) + 1}} 10^{\ell_{10}(\iterm(\ivary))} = \frac{a_i + \frac{1}{50}} {10 \lambda_{10}(a_i)} 10^{\ell_{10}(\iterm(\ivary))} \le \\
\frac{a_i + \frac{1}{50}} {a_i+1} 10^{\ell_{10}(\iterm(\ivary))} \le 10^{\ell_{10}(\iterm(\ivary))} \le \iterm(\ivary).
\end{array}
$

Therefore,  in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
\begin{align}
    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
      (\mbox{ Def. of } \alpha) & \notag \\
     & \ge  t(\bar{y}) & \notag 
\end{align}

If $\ivarx_{\sigma(i)} \ge  \alpha(\vec{\ivary}) + 2$, then

The goal is to find a bound for $x_i$ such that 
the values of the atoms containing $x_i$ keep constant when $x_i$ is greater than the bound.

When $x_i \ge \alpha+2$,
\begin{align}
    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
      (\mbox{ Def. of } \alpha) & \notag \\
     & \ge  t(\bar{y}) & \notag 
\end{align}
So we conclude that %when $x_i \ge \alpha+2,x_i\ge B',x_i \ge x_{i-1}+\delta$,
$\tau(x_i,\bar{x},\bar{y})$ keeps \textit{false} in this case.

%\znj{I think there are some mistakes in the above proof, pls check it!}
When $x_i \le \alpha-1$, similarly we have
\begin{align}
    \tau(x_i,\bar{x},\bar{y})\notag 
    & \le  a_i 2^{x_i} + 2^{x_i -\delta}A  + Bx_i &  \notag \\
    & \le  2^{x_i}(a_i +\frac{1}{4} +\frac{1}{4}) &(\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}}) \notag \\
   & \le  2\lambda_2(a_i)\frac{\lambda_2(t(\bar{y}))}{2\lambda_2(a_i)} &  (\mbox{ Def. of } \alpha) \notag \\
   &  \le  t(\bar{y}) & \notag 
\end{align}
which indicates that when  $x_i \le \alpha-1$, $\tau(x_i,\bar{x},\bar{y})$ keeps \textit{true}. \qed 
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{proof}

### 2021-05-14 FMCAD2021/1_intro.tex
sec6 half done

@@
 
-One more recent approach is flattening~\cite{}.
+One more recent approach is flattening~\cite{A}.
 The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat automata} (see Section~\ref{PA} for more details). 
@@
 For instance, we can find a solution for $x.y \neq z  \wedge |x| > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is a character in our alphabet.
-Second, and more importantly, because we can convert the membership problem of a flat automaton to a satisfiability problem of Presburgh formula, the class of word equation + regular constraints + length constraints is decidable.
+Second, and more importantly, because we can convert the membership problem of a flat automaton to a satisfiability problem of Presburgher formula, the class of word equation + regular constraints + length constraints is decidable.
 
@@
 For string/number conversion, their approach projects the solution to string/number conversion to, in terms of numerical solutions, a finite solution space.
+
 In this paper, we propose a complete solution to this problem.
-We propose a decision procedure for the class of word equation + regular constraints + length constraints + string/number conversion under arbitrary flat automata.
+We propose a decision procedure for the class of word equation + regular constraints + length constraints + string/number conversion under arbitrary flat automata (denoted by $\strint$).
+The basic idea of our approach can be sketched as follows:  similar to \cite{A}, we first reduce the satisfiability problem of $\strint$ to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$); then, according to the decidability of $\paexp$ \cite{Semenov84}, we therefore obtain the decidability of the satisfiability of $\strint$.
+
+In \cite{Semenov84}, Semen\"{o}v just proved the decidability and did not provide an explicit decision procedure.
+So, in  \cite{Point86}, Point presented a first decision procedure for the satisfiability of $\paexp$ in a mathematical way, which is difficult to understand and therefore no implementation available based on which up to now to the best of our knowledge.
+So, another contribution of this paper is that we present a more understandable algorithm for $\paexp$'s satisfiability in a computational way and its prototypical implementation.
+Although the general framework of our algorithm is quite similar to Point's,  it contains the following improvements and corrections:
+
+\wuhao{to do}
 

### 2021-05-14 FMCAD2021/4_dec_proc.tex
sec6 half done

@@
 
-\iffalse
-\subsection{Complexity Analysis and Case Study}
+\subsection{Complexity Analysis}
+\wuhao{To be checked}
 In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
@@
 this upper bound still holds.
-
-%\subsection{Example}
-
-We apply our algorithm to a simple example to 
-illustrate its function.
-Given $\Sigma = \{0,1\}$
-Suppose $\mathcal{A}$ is a finite state automaton
-that recognizes language $L = \{ w \mid \#w('1') \text{ is odd}\}$.
-Consider the following string constraint
-$$y\in L(\mathcal{A})\wedge 4\le  \parseInt(y) < 16\,.$$
-
-We need an abstraction parameter $alpha$ to restrict interpretations
-for $x$, for simplicity, set $\alpha\Def \langle 2,1 \rangle$.
-The generic $\langle 2,1 \rangle$-flat language is the union of 
-$\langle 2,1 \rangle$-flat languages $(11)^*,(10)^*,(01)^*,(00)^*,1^*,0^*$.
-Here we divide the problem into 6 sub-cases by restricting $y$
-to one of the $\langle 2,1 \rangle$-flat language.
-Another approach is to directly restrict $y$ to 
-the generic $\langle 2,1 \rangle$-flat language,
-at the cost of more introduced variables.
-Since the complexity is related to the number of variables,
-we wish it to be as little as possible.
-
-When $y$ is in language $1^*$, assume $y = 1^x$ where $x$
-is a fresh integer variable.
-Note that the Parikh image of $L(\mathcal{A})$ is $2|( x+1)$, 
-so the constraint is converted to a PA formula $\phi$
-$$\phi\Def \exists x. 2|(x+1) \wedge 5 \le 2^{x} < 17\,.$$
-
-Observe that when $x=3$ the formula holds,
-so we can easily deduce that $\phi$ is true.  
-By applying our algorithm,
-after \textbf{Normalization},
-$\phi$ is transformed into the form
-
-$$\exists x_0. 2|(x+1)\wedge  2^{x}\le 16 \wedge -2^{x}\le -5\,.$$
-
-Since there is only one variable, 
-\textbf{Enumerate-Orders} can be omitted.
-We send $2^x\le 16$ and $-2^x \le -5$ to \textbf{Elim-exp} and obtain
-$$2^x\le 16 \equiv x \le 4$$
-$$-2^x\le -5 \equiv 3\le x\,.$$ 
-Then we apply \textbf{Elim-linear} on formula $\exists x_0. 2|(x+1) \wedge x\le 4 \wedge 3\le x$ and get
-$$\bigvee_{j=0}^1(2| j+1 \wedge \textit{false} \wedge \textit{true})\vee \bigvee_{j=0}^1 (2|4-j+i\wedge 4-j\le 4 \wedge 3\le 4-j) \,.$$
-which evaluates \textit{true} and we can deduce $j=1\implies x=3$, so $y$ is string $111$.
-
-Other sub-cases are similar. However, when the abstraction parameter becomes larger,
-say, $\alpha =\langle 2,2 \rangle$,
-the complexity will grow rapidly,
-because \textbf{Enumerate-Orders} can not always be omitted and formulas in \textbf{Elim-exp} can not be simplified easily.
-\fi

### 2021-05-14 FMCAD2021/6_impl_exp.tex
implement ongoing
sec6 half done

@@
 
+\subsection{Implementation}
 We have implemented a prototype of our algorithm in Wolfram Mathematica, called  $\paexp$-Solver.
@@
 \paragraph{string hash function}
- flat pattern and arbitrarys pattern
 
-All experiments were run on an Intel Core i5 1.4GHz with 8GB RAM.
+In the sequel, we consider string hash functions of the form  
+\begin{equation}
+    \text{hash}(x)= \parseInt(x) \mod m_1  \notag
+\end{equation}
+where $m_1$ is usually a large (prime) number.
+We would like to find if there exists strings that match desired patterns and are hashed into certain integer values.
+In our benchmark, we focus string constraints of the form
+\begin{equation}
+    \left( \parseInt(x) \mod m_1\right) \mod m_2 = 0   \notag
+\end{equation}
+where $x$ either has a prefix ``12345'' or a suffix ``6789'' or both, for example $x=``12345"y``6789"$. We choose 2 patterns for $y$ to match:
 
-In our benchmark, the coefficients in inequalities are generated randomly within a range.
-As a consequence, some inequalities may not actually work, for example, $-10^{x_1} - x_1 \le 1$.
-So we always set the number of exponential  inequalities is one more than number of exponential variables, and the same for the linear comterpart.
-Since exponential terms with large coefficients are more likely to dominate the summation, coefficients of exponential terms are in the range $[-10,10]$ while other coefficients and constarints are generated randomly in $[-10^5,10^5]$.
+The first pattern is a FL with 2 cycles, $y \in (a)^+(b)^+$.
+Here we assume each cycle repeat at least once so that it will not degrade into FL with 1 cycle.
+The constraints is encoded to a $\paexp$ formula where two exponential variables denote the number of cycle $a$ and $b$ respectively.\wuhao{Need example?}
 
-For each
+The Second pattern is arbitrary string, that is $y\in \Sigma_{num}^*$.
+We use languages of the form $(a^1_1)^* (a^2_1...a^2_l)$ to under-approximate arbitrary strings, the first part is a single digit and the second part is a string length $l$, which can be seen as a FL with 2 cycles and force the second cycle do not repeat.
+We first set $l=1$ and enumerate $a^1_1$ from $1$ to $9$, for each case we obtain an arithmetic formula with $1$ exponential variables.
+If no satisfying model is found, we increase $l$ by 1 and repeat this procedure until it finds a model or timeout.
+This strategy is similar to \cite{Abdulla2020} with the difference that they force $a^1_1=0$ so that no exponential terms will occur.
+
+\subsection{Experiment Results}
+
+All experiments are run on a Intel i5 1.4GHz with 8GB RAM with TIMEOUT set to be 60 seconds.
+
+\wuhao{where to place tables?}
+
+From Table \ref{table:arithmetic}, we can see that for all cases except bounded 4-5-5-6 case, $\paexp$-Solver answers more \textit{sat} and \textit{unsat} than Z3 and CVC4.
+Our solver significantly outperforms when equalities are included.
+The average time needed to assert \textit{sat} or \textit{unsat} increase when we add more exponential variables.\wuhao{more details needed}
+
+For string hash function benchmark, we did compared our solver with Z3, CVC4 and Trau, but the experiments show that Trau would incorrectly answer \textit{unsat} in most cases even though a model can be found.
+So in \ref{table:string}, we still show the results of the other 3 solvers.
+As we can see, for flat patterns $(a)^*(b)^*$, Z3 and CVC4 would run out of time on most cases while $\paexp$-solver can give an answer for almost all trails.
+For arbitrary strings $(\Sigma_{num})^*$, CVC4 and Z3 seem to be biased against trails with surfix or prefix.
+By applying the under-approximation technique, our tool is able to find a match on all trails.
+\wuhao{more details needed}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table*}[]
+    \label{table:arithmetic}
+    \caption{Results of Z3, CVC4 and $\paexp$-Solver on arithmetic benchmark}
+    \begin{tabular}{cc|cc|cc|cc}
+    \multicolumn{1}{l}{} &  & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c}{MySolver} \\
+    \multicolumn{1}{l}{} &  & number & avg & number & avg & number & avg \\ \hline
+    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 2-3-3-4\end{tabular}} & SAT & 31 & \textless{}0.1 & 26 & 1.2 & 35 & \textless{}0.1 \\
+     & UNSAT & 37 & \textless{}0.1 & 38 & \textless{}0.1 & 65 & 0.3 \\
+     & TIMEOUT & 32 &  & 36 &  & 0 &  \\ \hline
+    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 3-4-4-5\end{tabular}} & SAT & 27 & \textless{}0.1 & 16 & \textless{}0.1 & 34 & 0.4 \\
+     & UNSAT & 22 & \textless{}0.1 & 24 & \textless{}0.1 & 66 & 2.8 \\
+     & TIMEOUT & 51 &  & 60 &  & 0 &  \\ \hline
+    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 3-4-4-5 \\ with equality\end{tabular}} & SAT & 4 & 19 & 0 & \textless{}0.1 & 16 & 3.7 \\
+     & UNSAT & 30 & \textless{}0.1 & 31 & 0.2 & 44 & \textless{}0.1 \\
+     & TIMEOUT & 66 &  & 69 &  & 28 & 19.6 \\ \hline
+    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 4-5-5-6\end{tabular}} & SAT & 20 & 0.2 & 15 & 8.2 & 33 & 5.2 \\
+     & UNSAT & 22 & \textless{}0.1 & 23 & 0.2 & 50 & 4.8 \\
+     & TIMEOUT & 58 &  & 62 &  & 17 &  \\ \hline\hline
+    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 2-3-3-4\end{tabular}} & SAT & 31 & \textless{}0.1 & 14 & \textless{}0.1 & 37 & 0,6 \\
+     & UNSAT & 31 & \textless{}0.1 & 30 & \textless{}0.1 & 46 & \textless{}0.1 \\
+     & B-UNSAT &  &  &  &  & 17 & 9.3 \\
+     & TIMEOUT & 38 &  & 56 &  & 0 &  \\ \hline
+    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 3-4-4-5\end{tabular}} & SAT & 31 & 0.8 & 13 & \textless{}0.1 & 36 & 3.2 \\
+     & UNSAT & 20 & \textless{}0.1 & 20 & \textless{}0.1 & 42 & \textless{}0.1 \\
+     & B-UNSAT &  &  &  &  & 3 & 54 \\
+     & TIMEOUT & 49 &  & 67 &  & 19 &  \\ \hline
+    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 3-4-4-5 \\ with equality\end{tabular}} & SAT & 3 & 23 & 1 & 39 & 14 & 27.5 \\
+     & UNSAT & 25 & \textless{}0.1 & 25 & \textless{}0.1 & 47 & \textless{}0.1 \\
+     & B-UNSAT &  &  &  &  & 2 & 57.9 \\
+     & TIMEOUT & 72 &  & 74 &  & 37 &  \\ \hline
+    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 4-5-5-6\end{tabular}} & SAT & 29 & 2 & 11 & \textless{}0.1 & 28 & 10 \\
+     & UNSAT & 20 & \textless{}0.1 & 20 & \textless{}0.1 & 38 & \textless{}0.1 \\
+     & B-UNSAT &  &  &  &  & 0 &  \\
+     & TIMEOUT & 51 &  & 69 &  & 34 & 
+    \end{tabular}
+\end{table*}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table*}[]
+    \label{table:string}
+    \caption{Results of Z3, CVC4 and $\paexp$-Solver on string hash function benchmark}
+    \begin{tabular}{c|c|ccc|ccc|ccc}
+                                                       &     & \multicolumn{3}{c|}{Z3} & \multicolumn{3}{c|}{CVC4} & \multicolumn{3}{c}{MySolver}       \\
+    PATTERN                                            & $\ell_{10}(m_1)$   & SAT  & UNSAT  & TIMEOUT & SAT   & UNSAT  & TIMEOUT  & SAT & UNSAT & TIMEOUT              \\ \hline
+    \multirow{5}{*}{12345(a)*(b)*}                     & 2   & 5    &        & 20      & 23    &        & 2        & 24  & 1     & 0                    \\
+                                                       & 3   & 0    &        & 25      & 4     &        & 21       & 10  & 14    & 1                    \\
+                                                       & 4   & 0    &        & 25      & 1     &        & 24       & 2   & 22    & 1                    \\
+                                                       & 5   & 0    &        & 25      & 1     &        & 24       & 1   & 23    & 1                    \\
+                                                       & total & 5    &        & 95      & 29    &        & 71       & 37  & 60    & 3                    \\ \hline
+    \multirow{5}{*}{12345(a)*(b)*6789}                 & 2   & 10   &        & 15      & 18    &        & 7        & 24  & 1     & 0                    \\
+                                                       & 3   & 1    &        & 24      & 8     &        & 17       & 10  & 15    & 0                    \\
+                                                       & 4   & 0    &        & 25      & 3     &        & 22       & 3   & 22    & 0                    \\
+                                                       & 5   & 0    &        & 25      & 0     &        & 25       & 0   & 25    & 0                    \\
+                                                       & total & 11   &        & 89      & 29    &        & 71       & 37  & 63    & 0                    \\ \hline
+    \multirow{5}{*}{(a)*(b)*6789}                      & 2   & 14   &        & 11      & 23    &        & 2        & 22  & 3     & 0                    \\
+                                                       & 3   & 4    &        & 21      & 6     & 1      & 18       & 15  & 10    & 0                    \\
+                                                       & 4   & 0    &        & 25      & 1     &        & 24       & 3   & 22    & 0                    \\
+                                                       & 5   & 0    &        & 25      & 0     &        & 25       & 1   & 24    & 0                    \\
+                                                       & total & 18   &        & 82      & 30    & 1      & 69       & 41  & 59    & 0                    \\ \hline\hline
+    \multirow{5}{*}{12345$(\Sigma_{num})^*$}     & 2   & 25   &        & 0       & 25    &        & 0        & 25  &       &                      \\
+                                                       & 3   & 25   &        & 0       & 25    &        & 0        & 25  &       &                      \\
+                                                       & 4   & 23   &        & 2       & 25    &        & 0        & 25  &       &                      \\
+                                                       & 5   & 9    &        & 16      & 25    &        & 0        & 25  &       &                      \\
+                                                       & total & 82   &        & 18      & 100   &        & 0        & 100 &       &                      \\ \hline
+    \multirow{5}{*}{12345$(\Sigma_{num})^*$6789} & 2   & 25   &        & 0       & 15    &        & 10       & 25  &       & \multicolumn{1}{l}{} \\
+                                                       & 3   & 23   &        & 2       & 1     &        & 24       & 25  &       & \multicolumn{1}{l}{} \\
+                                                       & 4   & 9    &        & 16      & 1     &        & 24       & 25  &       & \multicolumn{1}{l}{} \\
+                                                       & 5   & 3    &        & 22      & 0     &        & 25       & 25  &       &                      \\
+                                                       & total & 60   &        & 40      & 17    &        & 83       & 100 &       &                      \\ \hline
+    \multirow{5}{*}{$(\Sigma_{num})^*$6789}      & 2   & 25   &        & 0       & 19    &        & 6        & 25  &       & \multicolumn{1}{l}{} \\
+                                                       & 3   & 25   &        & 0       & 4     &        & 21       & 25  &       & \multicolumn{1}{l}{} \\
+                                                       & 4   & 16   &        & 9       & 4     &        & 21       & 25  &       & \multicolumn{1}{l}{} \\
+                                                       & 5   & 2    &        & 23      & 0     &        & 25       & 25  &       & \multicolumn{1}{l}{} \\
+                                                       & total & 68   &        & 32      & 27    &        & 73       & 100 &       &                     
+    \end{tabular}
+\end{table*}

### 2021-05-14 FMCAD2021/1_intro.tex
intro: improvements & corrections

@@
 
-\wuhao{to do}
+The syntax of Presburger arithmetic with exponential functions (denoted by $\paexp$) in our setting is more natural and easy to understand.
+In the algorihm, Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables.
+In our algorithm, we assume no special form of the formula other than NNF (negation normal form).
+A sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it in our lemma \ref{lem:exp-ineq} $a_i<0$ case.
 

### 2021-05-15 FMCAD2021/3_flatten.tex
section 4
some change on sec2
sec6 half done
optimization

@@
 \begin{array}{l c l}
-\sterm & ::= & a \mid \svarx \mid \sterm \concat \sterm, \\
-\iterm & ::= & n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,\\
-\varphi & ::= & \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi,
+\sterm & \Def & a \mid \svarx \mid \sterm \concat \sterm, \\
+\iterm & \Def & n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,\\
+\varphi & \Def & \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi,
 \end{array}
@@
 
-\subsection{An Extension of Presburger Arithmetic with Exponential Functions ($\paexp$)}
+\subsection{An Extension of Presburger Arithmetic with Exponential Functions (\paexp)}
 
@@
  \begin{array}{ l c l}
- \iterm &::=& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \mid \lambda_{10}(\iterm), \\
- \phi &::= & \iterm \ \op\ 0 \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi.
+ \iterm &\Def& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \mid \lambda_{10}(\iterm), \\
+ \phi &\Def & \iterm \ \op\ \iterm \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi.
  \end{array}
  $$
- The semantics of {\paexp} are defined similarly as {\pa}.
+ The semantics of {\paexp} are defined similarly to {\pa}. The quantifier-free as well as existential {\paexp} formulas are defined similarly to {\pa} as well.
 
@@
 \begin{example}
-Suppose $\parseInt(\svarx) = 2\ivarx$ is an atomic constraint and $\flatfun_\varphi(\svarx) = (12)^*(34)^*$. Then 
+Suppose $\parseInt(\svarx) = 2\ivarx$ is an atomic constraint and $\flatfun_\varphi(\svarx) = 1^*2^*$. Then 
 \[
@@
 & \flatten_{\flatfun_\varphi}(\parseInt(\svarx)  =  2\ivarx)  \\
-\equiv & \parseInt(12) \frac{10^{2\#_{\svarx,1}}-1}{10^2-1}10^{2\#_{\svarx,2}}  + \parseInt(34) \frac{10^{2\#_{\svarx,2}}-1}{10^2-1} = 2\ivarx   \\
-\equiv & 12 (10^{2\#_{\svarx,1}+2\#_{\svarx,2}} - 10^{2\#_{\svarx,2}})  + 34 (10^{2\#_{\svarx,2}}-1) = 198\ivarx.
+\Def & 1 \frac{10^{\#_{\svarx,1}}-1}{10-1}10^{\#_{\svarx,2}}  + 2 \frac{10^{\#_{\svarx,2}}-1}{10-1} = 2\ivarx   \\
+\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} - 10^{\#_{\svarx,2}}  + 2 (10^{2\#_{\svarx,2}}-1) = 18\ivarx \\
+\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} +  10^{\#_{\svarx,2}} = 18\ivarx + 2.
 \end{array}

### 2021-05-15 FMCAD2021/5_optimization.tex
optimization

@@
 %!TEX root = paper.tex
+
+In the last section, we illustrated the main idea of the decision procedure for  {\paexp}.
+The decision procedure has a high complexity in the sense that the elimination of the exponential occurrences   of each variable incurs an exponential blow-up, similarly for linear occurrences.
+Note that this high complexity holds even for quantifier-free formulas containing exponential terms: For a quantifier-free formula $\varphi$ containing exponential terms, we solve its satisfiability problem by adding the existential quantifiers for all the variables occurring in $\varphi$, then eliminate the quantifiers one by one, resulting into $\ltrue$ or $\lfalse$ in the end.
+The original formula $\varphi$ is satisfiable if $\ltrue$ is obtained in the end.
+
+In this section, we focus on quantifier-free {\paexp} formulas (or existential {$\paexp$} formulas since they are satisfiability-equivalent), and present various optimizations of the decision procedure for {\paexp}, aiming at an efficient implementation.
+The focus on quantifier-free {\paexp} formulas is motivated by the following two facts: 1) the flattening of {$\strint$} string constraints results into such formulas, 2) these formulas are already challenging for the state-of-the-art SMT solvers e.g.
+CVC4 and Z3 (with exponential functions defined as recursive functions).
+
+Let $\varphi$ be a quantifier-free {\paexp} formula in the remainder of this section.
+Moreover, we assume that $\varphi$ is normalized since the optimizations presented in the sequel are for normalized formulas.
+
+\paragraph{Reducing the number of enumerated variable orders by over approximation}
+
+Recall that in the decision procedure for {\paexp}, after the normalization, the orders of variables are enumerated and for each order, the exponential as well as linear occurrences of variables are eliminated.
+Since the quantifier elimination is expensive and applied to each possible order of variables, if we could reduce the candidate variable orders in the very beginning, it would facilitate considerable speed-up for the decision procedure.
+
+Our main idea is to consider an over approximation of $\varphi$, which is a {\pa} formula $\varphi'$, and use $\varphi'$ to remove the infeasible candidate variable orders.
+
+Note that all the exponential terms in $\varphi$ is of the form $10^{\ivarz}$ for some integer variable $\ivarz$. 
+%
+The over approximation is based on the observation that $10^n \ge 9 n + 1$ for every $n \in \Nat$.
+Then we obtain the over approximation $\varphi'$ from $\varphi$ by replacing each exponential term $10^{\ivarz}$ with a fresh variable $\ivarz'$ and add $\ivarz' \ge 9 \ivarz +1$ as a conjunct.
+
+Then during the enumeration of the linear orders for the variables $\ivarx_1, \ldots, \ivarx_n$, we can quickly remove those infeasible candidates $\sigma$ such that $\varphi' \wedge \bigwedge \limits_{i \in [n-1]}\ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ is unsatisfiable.
+A special case is that if $\varphi'$ is unsatisfiable, then we can directly conclude that original formula $\varphi$ is unsatisfiable.
+
+%Given any formula in $\paexp$, we first invoke \textbf{Normalization} to translate it into a simpler form.
+Then we  try to detect if there are obvious conflicts, for example, if the linear constraints combined are unsatisfiable, the whole formula is of course unsatisfiable.
+
+%Suppose the normalized formula $\theta$ contains variables $\{x_i:i\in[n]\}$, so the exponential terms could be $e^{x_i}$.
+For each $e^{x_i}$, we subtitute $e^{x_i}$ by a fresh variable $y_i$ in $\theta$ and add the constraints $(e-1)x_i+1\le y_i$ to get $\theta'$.
+Since $(e-1)x_i+1\le e^{x_i}$ holds for all $e,x\in \Nat$, $\theta'$ is an over-approximation of $\theta$ and if $\theta'$ is unsatisfiable over positive reals, then $\theta$ is unsatisfiable over $\Nat$.
+
+\zhilin{stopped here}
+
+\paragraph{Avoid the elimination of linear occurrences of variables by under approximation}
+
+xxx
+
+\paragraph{Avoid the formula-size blow-up by depth-first search}
+
+xxx
+
+\paragraph{pruning and preprocessing}
+
+xxx
 
@@
 
-\paragraph{modified \textbf{Elim-exp}}
+\paragraph{modified \textbf{Elim-exp}}\label{para: modified Elim-exp}
 

### 2021-05-15 FMCAD2021/1_intro.tex
Update 1_intro.tex

--- 2021-05-14 FMCAD2021/1_intro.tex
@@
 Although the general framework of our algorithm is quite similar to Point's,  it contains the following improvements and corrections:
-
-\wuhao{to do}
+   The syntax of Presburger arithmetic with exponential functions (denoted by $\paexp$) in our setting is more natural and easy to understand; Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables.
+In contrast, our algorithm just assumes that formulas are NNF (negation normal form).
+Therefore, our algorithm is much easy to follow and more efficient; Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} ( see later).
 

### 2021-05-15 FMCAD2021/5_optimization.tex
optimization

@@
 Moreover, we assume that $\varphi$ is normalized since the optimizations presented in the sequel are for normalized formulas.
+Furthermore, for technical convenience, we assume that all the inequality atomic formulas are of the form
+%
+\[\sum_{j=1}^{n} a_j 10^{\ivarx_{j}} + \sum_{k=1}^{n}b_k \ivarx_{k} \le c\]
+%
+where $c$ is an integer constant. (Implicitly, we assume that there are no free variables and all the variables are existentially quantified.)
 
-\paragraph{Reducing the number of enumerated variable orders by over approximation}
+\paragraph{Reduce the number of enumerated variable orders by over approximation}
 
@@
 
-\zhilin{stopped here}
-
 \paragraph{Avoid the elimination of linear occurrences of variables by under approximation}
 
-xxx
+In Section~\ref{sec-dec}, the decision procedure of {\paexp} requires the elimination of both exponential and linear occurrences of variables.
+Considering the fact that {\pa} formulas can be solved efficiently by the state-of-the-art solvers, e.g.
+CVC4 and Z3, one natural idea is to try to eliminate only the exponential occurrences, but not the linear occurrences, of variables, and obtain the {\pa} formulas in the end, which can then be solved by the state-of-the-art solvers.
+
+Recall that Lemma~\ref{lem:exp-ineq} enables us to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ from 
+$$
+\begin{array}{l}
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
+\hspace{4mm} 
+a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c.
+\end{array}
+$$
+Actually, Lemma~\ref{lem:exp-ineq} does more in the sense that all occurrences of $\ivarx_{\sigma(i)}$, including the linear ones, are eliminated from the atomic formulas resulted from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, e.g.
+$\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-1)} + p /\ivarx_{\sigma(i)}]$.
+Then we can continue eliminating the exponential occurrences of $\ivarx_{\sigma(i-1)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-1)} + p /\ivarx_{\sigma(i)}]$, provided that the coefficient of $\ivarx_{\sigma(i-1)}$ therein is nonzero.
+Iterating this process would produce a {\pa} formula eventually.
+
+Nevertheless, the side condition of Lemma~\ref{lem:exp-ineq}, namely $a_i \neq 0$, undermines the aforementioned natural idea. If $a_i = 0$, but $b_i \neq 0$, namely, 
+$$
+\begin{array}{l}
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
+\hspace{4mm} 
+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c,
+\end{array}
+$$
+then we are unable to utilize Lemma~\ref{lem:exp-ineq} to eliminate the linear occurrences of $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$.
+In this case, the quantifier elimination algorithm of {\pa} has to be applied to eliminate $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, so that later on, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-1)}$, which requires that $\ivarx_{\sigma(i-1)}$ is the maximum variable in the left-hand side of the inequality.
+
+To avoid applying the quantifier elimination algorithm of {\pa}, we consider the following under approximation of $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, namely, we additionally assume that $\ivarx_{\sigma(i)} \le 10^{u}$ for some constant bound $u \in \Nat$ with $u \ge 1$.
+Then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$ can be rewritten as
+%\begin{equation}\label{eqn-org}
+\[\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c - b_i  \ivarx_{\sigma(i)}.\]
+%\end{equation}
+Let us assume $a_{i-1} > 0$.
+It is easy to observe that $c - |b_i|10^u \le c - b_i \ivarx_{\sigma(i)} \le c + |b_i|10^u$.
+Then we can apply Lemma~\ref{lem:exp-ineq} to the following two inequalities to eliminate $10^{\ivarx_{\sigma(i-1)}}$,
+\begin{equation}\label{eqn-lb}
+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c - |b_i|10^u
+\end{equation}
+and
+\begin{equation}\label{eqn-ub}
+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c + |b_i|10^u.
+\end{equation}
+Let $\alpha_1 = \ell_{10}(c - |b_i|10^u) -  \ell_{10}(a_{i-1})$ and $\alpha_2 = \ell_{10}(c + |b_i|10^u) -  \ell_{10}(a_{i-1})$. Then from Lemma~\ref{lem:exp-ineq}, 
+\begin{itemize}
+\item if $\ivarx_{\sigma(i-1)} \ge B$, $\ivarx_{\sigma(i-1)} \le \alpha_1 - 1$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-lb}), thus also $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, is evaluated to $\ltrue$,
+\item if $\ivarx_{\sigma(i-1)} \ge B$, $\ivarx_{\sigma(i-1)} \ge \alpha_2 + 2$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-ub}), thus also $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, is evaluated to $\lfalse$.
+\end{itemize}
+Therefore, $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, thus also $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, is equivalent to 
+\[
+\small
+\begin{array}{l}
+\bigvee \limits_{p=0}^{B-1} ( \ivarx_{\sigma(i-1)} = p \wedge \tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) [p/\ivarx_{\sigma(i-1)}])   \\
+\bigvee \big(\ivarx_{\sigma(i-1)} \ge B \wedge \ivarx_{\sigma(i-1)} \le \alpha_1  - 1  \wedge \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} +\delta \big)\\
+%
+\bigvee \bigvee \limits_{p=0}^{\delta-1} 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i-1)} \ge B \wedge \ivarx_{\sigma(i-1)} \le \alpha_1  -1 \ \wedge \\
+ \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
+ \tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
+\end{array}
+\right)\\
+\bigvee \bigvee \limits_{p=0}^{\delta-1} 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i-1)} \ge B \wedge \ivarx_{\sigma(i-1)} \ge \alpha_2 + 2 \ \wedge \\
+ \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
+ \tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
+\end{array}
+\right)\\
+\bigvee \bigvee \limits_{p = \alpha_1}^{\alpha_2 + 1}
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i-1)} \ge B \wedge \ivarx_{\sigma(i-1)} = p\ \wedge \\
+ \tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}] 
+\end{array}
+\right),
+\end{array}
+\]
+where all exponential occurrences of $\ivarx_{\sigma(i-1)}$ are eliminated.
+
+Similarly, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-2)}$ from $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}]$ and  $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}]$.
+Eventually, we obtain a {\pa} formula.
 
@@
 
-xxx
+The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size.
+If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory.
+Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
+If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' result is reported.
 
-\paragraph{pruning and preprocessing}
+\paragraph{Preprocessing with small upper bound}
 
-xxx
+We believe that if a quantifier-free {\paexp} formula is satisfiable, then most probably it is satisfiable with small values assigned to variables.
+Consequently, as a preprocessing step, we put a small upper bound, e.g.
+the biggest constant occurring in the formula, on the values of variables, and perform a depth-first search, so that a model can be quickly found, if there is any.
+If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.
 
+%\zhilin{stopped here}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 In the last section, 
@@
 Intuitively, in most satisfiable cases exponential variables are small, so we first give a small bound (roughly the biggest constants in the formula) for exponential variables to perform a pre-search.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-15 FMCAD2021/1_intro.tex

@@
 In contrast, our algorithm just assumes that formulas are NNF (negation normal form).
-Therefore, our algorithm is much easy to follow and more efficient; Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} ( see later).
+Therefore, our algorithm is much easy to follow and more efficient; Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} (see later).
 

### 2021-05-16 FMCAD2021/6_impl_exp.tex
unbound result changed, not finished

@@
 Our solver significantly outperforms when equalities are included.
-The average time needed to assert \textit{sat} or \textit{unsat} increase when we add more exponential variables.\wuhao{more details needed}
+The average time needed to assert \textit{sat} or \textit{unsat} increase when we add more exponential variables.\wuhao{results of unsat cases are not correct!}
 

### 2021-05-16 Benchmark/NIA-test/produce-unbound.py

@@ -31,9 +31,12 @@ for testno in range(1,totaltests+1):
             coef = random.randint(-10,10)
         expr = expr + coef * pow(x[j])
         g.write("%d," %coef)
-    for j in range (1,expnum+linearnum+1):
+    for j in range (1,expnum+1):
         coef = random.randint(-10**5,10**5)
         expr = expr + coef * x[j]
+        g.write("%d," %coef)
+    for j in range (1,linearnum+1):
+        coef = 0
         g.write("%d," %coef)
     num = random.randint(-10**5,10**5)
     s.add(expr == num)

### 2021-05-16 FMCAD2021/6_impl_exp.tex
implementation+experiments

@@
 \subsection{Implementation}
-We have implemented a prototype of our algorithm in Wolfram Mathematica, called  $\paexp$-Solver.
+We implemented the decision procedure in Wolfram Mathematica, and obtain a solver, called the {\paexp}-solver, which is able to solve the satisfiability of quantifier-free {\paexp} formulas.
+
+The {\paexp}-solver takes a quantifier-free {\paexp} formula as the input.
+Moreover, it allows specifying a constant upper bound $10^u$ for the values of variables.
+If a constant upper bound $10^u$ is specified, then the problem is to decide whether there is an assignment of values no more than $10^u$ to variables satisfying the given {\paexp} formula.
+
+The outputs of the {\paexp}-solver are either ``SAT'', ``UNSAT'', ``B-UNSAT'', or ``TIMEOUT'', corresponding to the facts that the given formula is satisfiable, unsatisfiable, unsatisfiable up to $10^u$, or the search goes beyond the predetermined time limit.
+If the output is ``SAT'', then a model (namely, an assignment of values to variables) is also returned.
+
+%a prototype of our algorithm in Wolfram Mathematica, called  $\paexp$-Solver.
 Any postive Integer is allowed to be the base of exponential functions in our algorithm, but $\parseInt$ in Z3 and CVC4 only takes a numeric string as input.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 Our solver takes an existential quantified {$\paexp$} formula together with an upper bound for variables.
@@
 ``TIMEOUT" will be returned if the computation excesses the time limit.
+}
+
+\subsection{Benchmarks}
+
+To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC, and STRINGHASH.
+\begin{itemize}
+\item The ARITHMETIC benchmark suite comprises the {\paexp} formulas.
+%
+\item The STRINGHASH benchmark suite comprises 
+\end{itemize}
+
+\zhilin{stopped here}
 

### 2021-05-16 Benchmark/NIA-test/runall.sh
unbound benchmark corrected

@@ -1,8 +1,8 @@
 #!/bin/bash
 echo "begin test..."
 
-cd NIA-bound
-    cd 1-2-2-3
+cd NIA-unbound
+    cd 3-4-4-5-eq
     ./runz3.sh > z3-result 
     ./runcvc4.sh > cvc4-result
     cd ..

### 2021-05-16 FMCAD2021/6_impl_exp.tex
benchmarks

@@
 }
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC, and STRINGHASH.
-\begin{itemize}
-\item The ARITHMETIC benchmark suite comprises the {\paexp} formulas.
+%\begin{itemize}
+%\item
+
+\paragraph{The ARITHMETIC benchmark suite} 
+This suite comprises four groups of randomly generated {\paexp} formulas.
+Each group is characterized by four parameters ($EV$, $LV$, $EI$, $LI$), where $EV, LV$ represent the number of variables with exponential occurrences and with only linear occurrences respectively, and $EI, LI$ represent the number of inequalities with exponential terms and with only linear terms respectively.
+We consider the following parameter classes, 
+The ARITHMETIC benchmark suite consists of 200 randomly generated problem instances.
+The coefficients of exponential terms are randomly selected from the interval $[-10, 10]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
+The two intervals are chosen with intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.
+
 %
-\item The STRINGHASH benchmark suite comprises 
-\end{itemize}
+%\item
+
+\paragraph{The STRINGHASH benchmark suite} 
+This suite comprises 
+%\end{itemize}
 

### 2021-05-16 FMCAD2021/6_impl_exp.tex
benchmark

@@
 \paragraph{The ARITHMETIC benchmark suite} 
-This suite comprises four groups of randomly generated {\paexp} formulas.
+This suite comprises three groups of randomly generated {\paexp} formulas.
 Each group is characterized by four parameters ($EV$, $LV$, $EI$, $LI$), where $EV, LV$ represent the number of variables with exponential occurrences and with only linear occurrences respectively, and $EI, LI$ represent the number of inequalities with exponential terms and with only linear terms respectively.
-We consider the following parameter classes, 
-The ARITHMETIC benchmark suite consists of 200 randomly generated problem instances.
+We consider three parameter classes, $(2, 3, 3, 4)$, $(3, 4, 4, 5)$, and $(4, 5, 5, 6)$. 
+Each group of the ARITHMETIC benchmark suite consists of 200\zhilin{check later} randomly generated problem instances.
 The coefficients of exponential terms are randomly selected from the interval $[-10, 10]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
-The two intervals are chosen with intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.
+The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.
+Moreover, we randomly replace some occurrences of $\le$ by $=$ in order to include some equalities in the problem instances.
+We also create SMTLib2 files for each problem instance, in order to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
+Because neither CVC4 nor Z3 supports the exponential functions natively,  in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 10$ and $f(n+1) = 10*f(n)$.
 
@@
 \paragraph{The STRINGHASH benchmark suite} 
-This suite comprises 
+This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ such that for each $w= a_1 \ldots a_n$, 
+${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m$, where $p, m \in \Int^+$ with $p, m \ge 2$.
+It is easy to see that ${\sf hash}(w)$ can be seen as a generalization of $\parseInt$ followed by a modulo operation.
+In particular, if $\Sigma = \Sigma_{\sf num}$ and $p = 10$, then ${\sf hash}(w) = \parseInt(w) \bmod m$.
+Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, e.g.
+CVC4, Z3, and Trau, we assume $\Sigma =\Sigma_{\sf num}$ and $p = 10$, so that the string hash function can be encoded by $\parseInt$.
+
+The string constraints in the STRINGHASH benchmark suite are of the form 
+$$\svarx \in \Aut \wedge \left( \parseInt(\svarx) \bmod m \right) \bmod m^\prime = 0   \notag \wedge \len(\svarx) < 100,$$ 
+where $\Aut$ is an FA, $m, m^\prime \ge 2.$ 
+The six groups of string constraints are characterized by different classes of regular constraints, namely, $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, $w^+_1 w^+_2 6789$, $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, and $\Sigma_{\sf num}^* 6789$, where $w_1,w_2 \in \Sigma^+_{\sf num}$.
+The first three are flat languages, while the latter three are not.
+We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
+For the {\paexp}-solver, we do the following.
+\begin{itemize}
+\item For flat regular constraints, we generate the {\paexp} formulas corresponding to the string constraints, as the input to the {\paexp}-solver.
+%
+\item For non-flat regular constraints, we use flat languages $(a^1_1)^* (a^2_1...a^2_k)$ to under approximate 
+$\Sigma_{\sf num}^*$,  where $a^i_j \in \Sigma_{\sf num}$. 
+%
+%the first part is a single digit and the second part is a string length $l$, which can be seen as a FL with 2 cycles and force the second cycle do not repeat. 
+%
+We iterate the following procedure to find a model or reach the time limit: first set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a^1_1$.
+For each assignment, we encode the resulting string constraints into the {\paexp} formulas with only one exponential variables.
+If the resulting {\paexp} formula is unsatisfiable, then we increase $k$ by 1 and repeat this process.
+\end{itemize}
+We would like to remark that the flattening strategy for non-flat regular constraints is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(a^2_1...a^2_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
+On the other hand, we consider patterns of the form e.g.
+$12345 (a^1_1)^* (a^2_1...a^2_k) 6789$, which requires {\paexp} formulas to encode.
+
 %\end{itemize}
 
-\zhilin{stopped here}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 We are not aware of any SMT solver or constraints solver that directly support arithmetic formulas with exponential functions over Integer domain.
@@
 
-\paragraph{string hash function}
+%\paragraph{string hash function}
 
@@
 This strategy is similar to \cite{Abdulla2020} with the difference that they force $a^1_1=0$ so that no exponential terms will occur.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\zhilin{stopped here}
 

### 2021-05-16 FMCAD2021/6_impl_exp.tex
benmchmarks

@@
 %
-\item For non-flat regular constraints, we use flat languages $(a^1_1)^* (a^2_1...a^2_k)$ to under approximate 
-$\Sigma_{\sf num}^*$,  where $a^i_j \in \Sigma_{\sf num}$. 
+\item For non-flat regular constraints, we use flat languages $(a)^* (b_{1} \ldots b_{k})$ to under approximate 
+$\Sigma_{\sf num}^*$,  where $a, b_1,\ldots, b_k \in \Sigma_{\sf num}$. 
 %
@@
 %
-We iterate the following procedure to find a model or reach the time limit: first set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a^1_1$.
+We iterate the following procedure to find a model or reach the time limit: first set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a$.
 For each assignment, we encode the resulting string constraints into the {\paexp} formulas with only one exponential variables.
@@
 \end{itemize}
-We would like to remark that the flattening strategy for non-flat regular constraints is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(a^2_1...a^2_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
+We would like to remark that the flattening strategy for non-flat regular constraints is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(b_1...b_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
 On the other hand, we consider patterns of the form e.g.
-$12345 (a^1_1)^* (a^2_1...a^2_k) 6789$, which requires {\paexp} formulas to encode.
+$12345 (a)^* (b_1 \ldots b_k) 6789$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode.
 

### 2021-05-16 FMCAD2021/6_impl_exp.tex
experiments

@@
 \begin{itemize}
-\item For flat regular constraints, we generate the {\paexp} formulas corresponding to the string constraints, as the input to the {\paexp}-solver.
+\item For flat regular constraints, we generate the {\paexp} formulas corresponding to the string constraints, as the inputs to the {\paexp}-solver.
 %
@@
 
-\zhilin{stopped here}
-
 \subsection{Experiment Results}
 
-All experiments are run on a Intel i5 1.4GHz with 8GB RAM with TIMEOUT set to be 60 seconds.
+We compare the {\paexp}-solver against the state-of-the-art SMT solvers on the generated benchmarks. Specifically, 
+\begin{itemize}
+\item over the ARITHMETIC benchmark suite, we compare the {\paexp}-solver against CVC4 and Z3,
+\item over the STRINGHASH benchmark suite, we compare the {\paexp}-solver against CVC4, Z3, and Trau. 
+\end{itemize}
+All the experiments are run on a lap-top with the Intel i5 1.4GHz CPU and 8GB memory. We set the timeout threshold as 60 seconds per problem instance.
+
+The experiment results are summarized in Table ~\ref{xxx}.
+
+\zhilin{stopped here}
 

### 2021-05-16 FMCAD2021/6_impl_exp.tex
2 results

--- 2021-05-16 FMCAD2021/6_impl_exp.tex
@@
 }
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC, and STRINGHASH.
-%\begin{itemize}
-%\item
+\begin{itemize}
+\item The ARITHMETIC benchmark suite comprises the {\paexp} formulas.
+%
+\item The STRINGHASH benchmark suite comprises 
+\end{itemize}
 
-\paragraph{The ARITHMETIC benchmark suite} 
-This suite comprises three groups of randomly generated {\paexp} formulas.
-Each group is characterized by four parameters ($EV$, $LV$, $EI$, $LI$), where $EV, LV$ represent the number of variables with exponential occurrences and with only linear occurrences respectively, and $EI, LI$ represent the number of inequalities with exponential terms and with only linear terms respectively.
-We consider three parameter classes, $(2, 3, 3, 4)$, $(3, 4, 4, 5)$, and $(4, 5, 5, 6)$. 
-Each group of the ARITHMETIC benchmark suite consists of 200\zhilin{check later} randomly generated problem instances.
-The coefficients of exponential terms are randomly selected from the interval $[-10, 10]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
-The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.
-Moreover, we randomly replace some occurrences of $\le$ by $=$ in order to include some equalities in the problem instances.
-We also create SMTLib2 files for each problem instance, in order to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
-Because neither CVC4 nor Z3 supports the exponential functions natively,  in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 10$ and $f(n+1) = 10*f(n)$.
+\zhilin{stopped here}
 
-%
-%\item
-
-\paragraph{The STRINGHASH benchmark suite} 
-This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ such that for each $w= a_1 \ldots a_n$, 
-${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m$, where $p, m \in \Int^+$ with $p, m \ge 2$.
-It is easy to see that ${\sf hash}(w)$ can be seen as a generalization of $\parseInt$ followed by a modulo operation.
-In particular, if $\Sigma = \Sigma_{\sf num}$ and $p = 10$, then ${\sf hash}(w) = \parseInt(w) \bmod m$.
-Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, e.g.
-CVC4, Z3, and Trau, we assume $\Sigma =\Sigma_{\sf num}$ and $p = 10$, so that the string hash function can be encoded by $\parseInt$.
-
-The string constraints in the STRINGHASH benchmark suite are of the form 
-$$\svarx \in \Aut \wedge \left( \parseInt(\svarx) \bmod m \right) \bmod m^\prime = 0   \notag \wedge \len(\svarx) < 100,$$ 
-where $\Aut$ is an FA, $m, m^\prime \ge 2.$ 
-The six groups of string constraints are characterized by different classes of regular constraints, namely, $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, $w^+_1 w^+_2 6789$, $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, and $\Sigma_{\sf num}^* 6789$, where $w_1,w_2 \in \Sigma^+_{\sf num}$.
-The first three are flat languages, while the latter three are not.
-We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
-For the {\paexp}-solver, we do the following.
-\begin{itemize}
-\item For flat regular constraints, we generate the {\paexp} formulas corresponding to the string constraints, as the input to the {\paexp}-solver.
-%
-\item For non-flat regular constraints, we use flat languages $(a^1_1)^* (a^2_1...a^2_k)$ to under approximate 
-$\Sigma_{\sf num}^*$,  where $a^i_j \in \Sigma_{\sf num}$. 
-%
-%the first part is a single digit and the second part is a string length $l$, which can be seen as a FL with 2 cycles and force the second cycle do not repeat. 
-%
-We iterate the following procedure to find a model or reach the time limit: first set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a^1_1$.
-For each assignment, we encode the resulting string constraints into the {\paexp} formulas with only one exponential variables.
-If the resulting {\paexp} formula is unsatisfiable, then we increase $k$ by 1 and repeat this process.
-\end{itemize}
-We would like to remark that the flattening strategy for non-flat regular constraints is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(a^2_1...a^2_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
-On the other hand, we consider patterns of the form e.g.
-$12345 (a^1_1)^* (a^2_1...a^2_k) 6789$, which requires {\paexp} formulas to encode.
-
-%\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
 We are not aware of any SMT solver or constraints solver that directly support arithmetic formulas with exponential functions over Integer domain.
@@
 
-%\paragraph{string hash function}
+\paragraph{string hash function}
 
@@
 This strategy is similar to \cite{Abdulla2020} with the difference that they force $a^1_1=0$ so that no exponential terms will occur.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\zhilin{stopped here}
 
@@
 % \usepackage{multirow}
+% Merge bound and unbound 
+\begin{table*}[]
+    \label{table:arithmetic}
+    \caption{Results of Z3, CVC4 and $\paexp$-Solver on arithmetic benchmark}
+    \begin{tabular}{cc|cc|cc|cc}
+    \multicolumn{1}{l}{} &  & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c}{MySolver} \\
+    \multicolumn{1}{l}{} &  & number & avg & number & avg & number & avg \\ \hline
+    \multirow{4}{*}{2-3-3-4} & SAT & 62 & \textless{}0.1 & 40 & 0.8 & 72 & 0.1 \\
+     & UNSAT & 68 & \textless{}0.1 & 68 & \textless{}0.1 & 111 & 0.3 \\
+     & \multicolumn{1}{l|}{B-UNSAT} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 17 & 9.3 \\
+     & TIMEOUT & 70 &  & 92 &  & 0 &  \\ \hline
+    \multirow{4}{*}{3-4-4-5} & SAT & 58 & 0.5 & 29 & 0.3 & 70 & 2.1 \\
+     & UNSAT & 42 & \textless{}0.1 & 44 & \textless{}0.1 & 97 & 1.8 \\
+     & \multicolumn{1}{l|}{B-UNSAT} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 14 & 18.0 \\
+     & TIMEOUT & 100 &  & 127 &  & 19 &  \\ \hline
+    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}3-4-4-5 \\ with equality\end{tabular}} & SAT & 3 & 23 & 1 & 39.0 & 14 & 7.5 \\
+     & UNSAT & 63 & \textless{}0.1 & 66 & 0.2 & 124 & 3.4 \\
+     & \multicolumn{1}{l|}{B-UNSAT} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 25 & 13.4 \\
+     & TIMEOUT & 134 &  & 133 &  & 37 &  \\ \hline
+    \multirow{4}{*}{4-5-5-6} & SAT & 49 & 1.2 & 26 & 4.8 & 58 & 13.4 \\
+     & UNSAT & 42 & \textless{}0.1 & 43 & 0.1 & 82 & \textless{}0.1 \\
+     & \multicolumn{1}{l|}{B-UNSAT} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 1 & 39 \\
+     & TIMEOUT & 109 &  & 131 &  & 59 & 
+    \end{tabular}
+\end{table*}
+
+\hide{
 \begin{table*}[]
@@
 \end{table*}
+}
 

### 2021-05-16 Benchmark/NIA-test/produce.py

@@ -1,65 +1,66 @@
 from z3 import *
 import random
 
-expnum = 2
-linearnum = 3
-expfml = 3
-linearfml = 4
+expnum = 4
+linearnum = 5
+expfml = 5
+linearfml = 6
 totaltests = 100
 
 x=[0]*(expnum+linearnum+2)
 
-g = open("NIA-unbound/2-3-3-4-eq/paras","w+")
+g = open("new-test/4-5-5-6/paras","w+")
 g.write("%d\n" %totaltests)
 g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
 
 for testno in range(1,totaltests+1):
-    for i in range(1,expnum+linearnum+1):
+    
+    for i in range(1, expnum+linearnum+1):
         x[i] = Int('x'+str(i))
     pow = Function('pow',IntSort(),IntSort())
 
     s = Solver()
-
     for i in range (1,expnum+linearnum+1):
         s.add(x[i] >= 0)
 
-    # add an equality
-    #g.write("{")
-    #expr = 0
-    #for j in range (1,expnum+1):
-    #    coef = random.randint(-10,10)
-    #    expr = expr + coef * pow(x[j])
-    #    g.write("%d," %coef)
-    #for j in range (1,expnum+linearnum+1):
-    #    coef = random.randint(-10**2,10**2)
-    #    expr = expr + coef * x[j]
-    #    g.write("%d," %coef)
-    #num = random.randint(-10**5,10**5)
-    #s.add(expr == num)
-    #g.write("%d}\n" % num)
+    # first exp formula, 1/2 equality 1/2 inequality
+    g.write("{")
+    expr = 0
+    for j in range (1,expnum+1):
+        coef = random.randint(-100,100)
+        expr = expr + coef * pow(x[j])
+        g.write("%d," %coef)
 
-    # with eq 2
+    for j in range (1,expnum+linearnum+1):
+        coef = random.randint(-10**5,10**5)
+        expr = expr + coef * x[j]
+        g.write("%d," %coef)
+        
+    num = random.randint(-10**5,10**5)
+    
+    if random.random() <= 0.5:
+        s.add(expr <= num)
+        g.write("0,")
+    else:
+        s.add(expr == num)
+        g.write("1,")
+
+    g.write("%d}\n" % num)
+
+
     for i in range (2,expfml+1):
         g.write("{")
         expr = 0
         for j in range (1,expnum+1):
             coef = random.randint(-10,10)
-            if coef == 0:
[…]

### 2021-05-16 Benchmark/NIA-test/runall.sh

@@ -1,8 +1,16 @@
 #!/bin/bash
 echo "begin test..."
 
-cd NIA-unbound
-    cd 3-4-4-5-eq
+cd new-test
+    cd 2-3-3-4
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+    cd 3-4-4-5
+    ./runz3.sh > z3-result 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+    cd 4-5-5-6
     ./runz3.sh > z3-result 
     ./runcvc4.sh > cvc4-result
     cd ..

### 2021-05-17 FMCAD2021/5_optimization.tex
optimization
Z3&mysolver result on Arith

@@
 Let us assume $a_{i-1} > 0$.
-It is easy to observe that $c - |b_i|10^u \le c - b_i \ivarx_{\sigma(i)} \le c + |b_i|10^u$.
+Define $c_1, c_2$ as follows: If $b_i > 0$, then $c_1 = c- b_i 10^u$ and $c_2 = c$, otherwise, $c_1 = c$ and $c_2 = c - b_i 10^u$.
+It is easy to observe that $c_1 \le c - b_i \ivarx_{\sigma(i)} \le c_2$.
 Then we can apply Lemma~\ref{lem:exp-ineq} to the following two inequalities to eliminate $10^{\ivarx_{\sigma(i-1)}}$,
 \begin{equation}\label{eqn-lb}
-\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c - |b_i|10^u
+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c_1
 \end{equation}
@@
 \begin{equation}\label{eqn-ub}
-\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c + |b_i|10^u.
+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c_2.
 \end{equation}
-Let $\alpha_1 = \ell_{10}(c - |b_i|10^u) -  \ell_{10}(a_{i-1})$ and $\alpha_2 = \ell_{10}(c + |b_i|10^u) -  \ell_{10}(a_{i-1})$. Then from Lemma~\ref{lem:exp-ineq}, 
+Let $\alpha_1 = \ell_{10}(c_1) -  \ell_{10}(a_{i-1})$ and $\alpha_2 = \ell_{10}(c_2) -  \ell_{10}(a_{i-1})$. Then from Lemma~\ref{lem:exp-ineq}, 
 \begin{itemize}
@@
 Eventually, we obtain a {\pa} formula.
+
+\paragraph{Perform the elimination of exponential occurrences of variables in parallel on atomic formulas}
+
+Note that when Lemma~\ref{lem:exp-ineq} is applied to an atomic formula $\tau$ to eliminate exponential occurrences of $\ivarx$, it computes 4 parameters, namingly $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$, and obtains an equivalent {\pa} formula according to these parameters.
+Since the parameters are used as boundary conditions in enumeration, we can choose the strictest boundary conditions and apply to all atomic formulas.
+
+More specifically, we first compute $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$ 
+for all atomic formulas that contain $10^\ivarx$.
+Let $\alpha^{\min}_1$ denote the minimal element in $\{\alpha^\tau_1\}_\tau$, let $\alpha^{\max}_2,B^{\max},\delta^{\max}$ denote the maximal element in $\{\alpha^\tau_2\}_\tau$, $\{B^\tau\}_\tau$ and $\{\delta^\tau\}_\tau$ respectively.
+Then, we apply $(\alpha^{\min}_1,\alpha^{\max}_2,B^{\max},\delta^{\max})$ to all atomic formulas $\tau$ in parallel and obtain $\varphi'$ equivalent to $\varphi$.
+
+For instance, assume coefficients of $10^\ivarx$ in atomic formulas of $\varphi$ are all non-negative.
+Let $\varphi^+$ denote the formula obtained by replace atomic formulas where coefficients of $10^\ivarx$ are positive by $\ltrue$.
+The resulted formula $\varphi'$ is
+\[
+\small
+\begin{array}{l}
+\bigvee \limits_{p=0}^{B^{\max}-1} ( \ivarx_{\sigma(i-1)} = p \wedge \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) [p/\ivarx_{\sigma(i-1)}])   \\
+\bigvee \left(
+    \begin{array}{l}
+        \ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  - 1  \wedge \\
+        \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} +\delta^{\max} \wedge \varphi^+  
+    \end{array}
+    \right)\\
+%
+\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  -1 \ \wedge \\
+ \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
+ \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
+\end{array}
+\right)\\
+\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \ge \alpha^{\max}_2 + 2 \ \wedge \\
+ \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
+ \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
+\end{array}
+\right)\\
+\bigvee \bigvee \limits_{p = \alpha^{\min}_1}^{\alpha^{\max}_2 + 1}
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} = p\ \wedge \\
+ \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}] 
+\end{array}
+\right),
+\end{array}
+\]
 

### 2021-05-17 FMCAD2021/6_impl_exp.tex
Z3&mysolver result on Arith

--- 2021-05-16 FMCAD2021/6_impl_exp.tex
@@
 % \usepackage{multirow}
+% Merge bound and unbound 
 \begin{table*}[]
@@
     \caption{Results of Z3, CVC4 and $\paexp$-Solver on arithmetic benchmark}
+    \centering
     \begin{tabular}{cc|cc|cc|cc}
@@
         \multicolumn{1}{l}{} &  & number & avg & number & avg & number & avg \\ \hline
-    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 2-3-3-4\end{tabular}} & SAT & 31 & \textless{}0.1 & 26 & 1.2 & 35 & \textless{}0.1 \\
-     & UNSAT & 37 & \textless{}0.1 & 38 & \textless{}0.1 & 65 & 0.3 \\
-     & TIMEOUT & 32 &  & 36 &  & 0 &  \\ \hline
-    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 3-4-4-5\end{tabular}} & SAT & 27 & \textless{}0.1 & 16 & \textless{}0.1 & 34 & 0.4 \\
-     & UNSAT & 22 & \textless{}0.1 & 24 & \textless{}0.1 & 66 & 2.8 \\
-     & TIMEOUT & 51 &  & 60 &  & 0 &  \\ \hline
-    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 3-4-4-5 \\ with equality\end{tabular}} & SAT & 4 & 19 & 0 & \textless{}0.1 & 16 & 3.7 \\
-     & UNSAT & 30 & \textless{}0.1 & 31 & 0.2 & 44 & \textless{}0.1 \\
-     & TIMEOUT & 66 &  & 69 &  & 28 & 19.6 \\ \hline
-    \multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}unbound\\ 4-5-5-6\end{tabular}} & SAT & 20 & 0.2 & 15 & 8.2 & 33 & 5.2 \\
-     & UNSAT & 22 & \textless{}0.1 & 23 & 0.2 & 50 & 4.8 \\
-     & TIMEOUT & 58 &  & 62 &  & 17 &  \\ \hline\hline
-    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 2-3-3-4\end{tabular}} & SAT & 31 & \textless{}0.1 & 14 & \textless{}0.1 & 37 & 0,6 \\
-     & UNSAT & 31 & \textless{}0.1 & 30 & \textless{}0.1 & 46 & \textless{}0.1 \\
-     & B-UNSAT &  &  &  &  & 17 & 9.3 \\
-     & TIMEOUT & 38 &  & 56 &  & 0 &  \\ \hline
-    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 3-4-4-5\end{tabular}} & SAT & 31 & 0.8 & 13 & \textless{}0.1 & 36 & 3.2 \\
-     & UNSAT & 20 & \textless{}0.1 & 20 & \textless{}0.1 & 42 & \textless{}0.1 \\
-     & B-UNSAT &  &  &  &  & 3 & 54 \\
-     & TIMEOUT & 49 &  & 67 &  & 19 &  \\ \hline
-    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 3-4-4-5 \\ with equality\end{tabular}} & SAT & 3 & 23 & 1 & 39 & 14 & 27.5 \\
-     & UNSAT & 25 & \textless{}0.1 & 25 & \textless{}0.1 & 47 & \textless{}0.1 \\
-     & B-UNSAT &  &  &  &  & 2 & 57.9 \\
-     & TIMEOUT & 72 &  & 74 &  & 37 &  \\ \hline
-    \multirow{4}{*}{\begin{tabular}[c]{@{}c@{}}bound\\ 4-5-5-6\end{tabular}} & SAT & 29 & 2 & 11 & \textless{}0.1 & 28 & 10 \\
-     & UNSAT & 20 & \textless{}0.1 & 20 & \textless{}0.1 & 38 & \textless{}0.1 \\
-     & B-UNSAT &  &  &  &  & 0 &  \\
-     & TIMEOUT & 51 &  & 69 &  & 34 & 
+        \multirow{4}{*}{2-3-3-4} & SAT & 27 & 0.6 &  &  & 31 & 0.4 \\
+         & UNSAT & 37 & \textless{}0.1 &  &  & 49 & \textless{}0.1 \\
+         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 20 & 10.4 \\
+         & TIMEOUT & 36 &  &  &  & 0 &  \\ \hline
+        \multirow{4}{*}{3-4-4-5} & SAT & 20 & 0.1 &  &  & 29 & 3.5 \\
+         & UNSAT & 26 & \textless{}0.1 &  &  & 42 & \textless{}0.1 \\
+         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 1 & 53.8 \\
+         & TIMEOUT & 54 &  &  &  & 28 &  \\ \hline
+        \multirow{4}{*}{4-5-5-6} & SAT & 22 & 1.8 &  &  & 27 & 20.0 \\
+         & UNSAT & 21 & 0.4 &  &  & 36 & 0.1 \\
+         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 0 &  \\
+         & TIMEOUT & 57 &  &  &  & 37 & 
     \end{tabular}

### 2021-05-17 Benchmark/NIA-test/produce.py

@@ -1,15 +1,15 @@
 from z3 import *
 import random
 
-expnum = 4
-linearnum = 5
-expfml = 5
-linearfml = 6
+expnum = 1
+linearnum = 2
+expfml = 2
+linearfml = 3
 totaltests = 100
 
 x=[0]*(expnum+linearnum+2)
 
-g = open("new-test/4-5-5-6/paras","w+")
+g = open("new-test/1-2-2-3/paras","w+")
 g.write("%d\n" %totaltests)
 g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
 
@@ -52,7 +52,7 @@ for testno in range(1,totaltests+1):
         g.write("{")
         expr = 0
         for j in range (1,expnum+1):
-            coef = random.randint(-10,10)
+            coef = random.randint(-100,100)
             expr = expr + coef * pow(x[j])
             g.write("%d," %coef)
         # bound cases
@@ -69,14 +69,14 @@ for testno in range(1,totaltests+1):
         expr = 0
         g.write("{")
         for j in range (1,expnum+linearnum+1):
-            coef = random.randint(-10**2,10**2)
+            coef = random.randint(-10**5,10**5)
             expr = expr + coef * x[j]
             g.write("%d," %coef)
         num = random.randint(-10**5,10**5)
         s.add(expr <= num)
         g.write("%d}\n" % num)
             
-    f = open('new-test/4-5-5-6/test'+str(testno),mode='w+')
+    f = open('new-test/1-2-2-3/test'+str(testno),mode='w+')
     f.write("(set-logic ALL)\n")
     f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
     f.write(s.sexpr())

### 2021-05-17 Benchmark/NIA-test/runall.sh

@@ -2,16 +2,19 @@
 echo "begin test..."
 
 cd new-test
-    cd 2-3-3-4
+    cd 1-2-2-3
     ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
     cd ..
-    cd 3-4-4-5
-    ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
-    cd ..
-    cd 4-5-5-6
-    ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
-    cd ..
+    #cd 2-3-3-4
+    #./runz3.sh > z3-result 
+    #./runcvc4.sh > cvc4-result
+    #cd ..
+    #cd 3-4-4-5
+    #./runz3.sh > z3-result 
+    #./runcvc4.sh > cvc4-result
+    #cd ..
+    #cd 4-5-5-6
+    #./runz3.sh > z3-result 
+    #./runcvc4.sh > cvc4-result
+    #cd ..
 cd ..

### 2021-05-17 Benchmark/NIA-test/new-test/3-4-4-4/runcvc4.sh
sec:experiment

@@ -1,0 +1,17 @@
+#!/bin/bash
+echo "begin test using CVC4 ..."
+
+count=1
+while(( $count<= 100 ))
+do
+    START=$(date +%s)
+    # do something
+    # start your script work here
+    echo $count
+    cvc4 test$count -L smt --tlimit=60000 --fmf-fun --produce-models
+    let "count++"
+    # your logic ends here
+    END=$(date +%s)
+    DIFF=$(( $END - $START ))
+    echo "It took $DIFF seconds"
+done

### 2021-05-17 Benchmark/NIA-test/produce.py

@@ -1,15 +1,15 @@
 from z3 import *
 import random
 
-expnum = 1
-linearnum = 2
-expfml = 2
-linearfml = 3
+expnum = 3
+linearnum = 4
+expfml = 4
+linearfml = 4
 totaltests = 100
 
 x=[0]*(expnum+linearnum+2)
 
-g = open("new-test/1-2-2-3/paras","w+")
+g = open("new-test/3-4-4-4/paras","w+")
 g.write("%d\n" %totaltests)
 g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
 
@@ -76,7 +76,7 @@ for testno in range(1,totaltests+1):
         s.add(expr <= num)
         g.write("%d}\n" % num)
             
-    f = open('new-test/1-2-2-3/test'+str(testno),mode='w+')
+    f = open('new-test/3-4-4-4/test'+str(testno),mode='w+')
     f.write("(set-logic ALL)\n")
     f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
     f.write(s.sexpr())

### 2021-05-17 Benchmark/NIA-test/runall.sh

@@ -2,7 +2,7 @@
 echo "begin test..."
 
 cd new-test
-    cd 1-2-2-3
+    cd 3-4-4-4
     ./runz3.sh > z3-result 
     cd ..
     #cd 2-3-3-4

### 2021-05-18 FMCAD2021/6_impl_exp.tex
new arith result

@@
 Each group of the ARITHMETIC benchmark suite consists of 200\zhilin{check later} randomly generated problem instances.
-The coefficients of exponential terms are randomly selected from the interval $[-10, 10]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
-The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.
-Moreover, we randomly replace some occurrences of $\le$ by $=$ in order to include some equalities in the problem instances.
-We also create SMTLib2 files for each problem instance, in order to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
+The coefficients of exponential terms are randomly selected from the interval $[-100, 100]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
+The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.\wuhao{Moreover, in order to include some equalities in the problem instances, we randomly select some instances and replace the occurrence of $\le$ by $=$ in the first inequality with exponential terms.} We also create SMTLib2 files for each problem instance, in order to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
 Because neither CVC4 nor Z3 supports the exponential functions natively, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 10$ and $f(n+1) = 10*f(n)$.
@@
         \multicolumn{1}{l}{} &  & number & avg & number & avg & number & avg \\ \hline
-        \multirow{4}{*}{2-3-3-4} & SAT & 27 & 0.6 &  &  & 31 & 0.4 \\
-         & UNSAT & 37 & \textless{}0.1 &  &  & 49 & \textless{}0.1 \\
-         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 20 & 10.4 \\
-         & TIMEOUT & 36 &  &  &  & 0 &  \\ \hline
-        \multirow{4}{*}{3-4-4-5} & SAT & 20 & 0.1 &  &  & 29 & 3.5 \\
-         & UNSAT & 26 & \textless{}0.1 &  &  & 42 & \textless{}0.1 \\
-         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 1 & 53.8 \\
-         & TIMEOUT & 54 &  &  &  & 28 &  \\ \hline
-        \multirow{4}{*}{4-5-5-6} & SAT & 22 & 1.8 &  &  & 27 & 20.0 \\
-         & UNSAT & 21 & 0.4 &  &  & 36 & 0.1 \\
+        \multirow{4}{*}{2-3-3-4} & SAT & 56 & 0.4 & 42 & 2.3 & 64 & 0.4 \\
+         & UNSAT & 69 & \textless{}0.1 & 72 & \textless{}0.1 & 89 & \textless{}0.1 \\
+         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 47 & 9.5 \\
+         & TIMEOUT & 75 &  & 86 &  & 0 &  \\ \hline
+        \multirow{4}{*}{3-4-4-5} & SAT & 33 & 1.4 & 25 & 2.9 & 52 & 3.3 \\
+         & UNSAT & 59 & \textless{}0.1 & 60 & \textless{}0.1 & 88 & \textless{}0.1 \\
+         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 1 & 54.0 \\
+         & TIMEOUT & 108 &  & 115 &  & 59 &  \\ \hline
+        \multirow{4}{*}{4-5-5-6} & SAT & 35 & 1.8 & 19 & 6.6 & 47 & 22.4 \\
+         & UNSAT & 36 & 0.3 & 39 & 0.4 & 72 & 0.1 \\
          & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 0 &  \\
-         & TIMEOUT & 57 &  &  &  & 37 & 
+         & TIMEOUT & 129 &  & 142 &  & 81 & 
         \end{tabular}

### 2021-05-18 Benchmark/NIA-test/new-test/2-3-3-4/runcvc4.sh

@@ -2,7 +2,7 @@
 echo "begin test using CVC4 ..."
 
 count=1
-while(( $count<= 100 ))
+while(( $count<= 200 ))
 do
     START=$(date +%s)
     # do something

### 2021-05-18 Benchmark/NIA-test/new-test/4-5-5-6/runz3.sh

@@ -1,8 +1,8 @@
 #!/bin/bash
 echo "begin test using Z3 ..."
 
-count=1
-while(( $count<= 100 ))
+count=101
+while(( $count<= 200 ))
 do
     START=$(date +%s)
     # do something

### 2021-05-18 Benchmark/NIA-test/produce.py

@@ -1,19 +1,19 @@
 from z3 import *
 import random
 
-expnum = 3
-linearnum = 4
-expfml = 4
-linearfml = 4
+expnum = 4
+linearnum = 5
+expfml = 5
+linearfml = 6
 totaltests = 100
 
 x=[0]*(expnum+linearnum+2)
 
-g = open("new-test/3-4-4-4/paras","w+")
+g = open("new-test/4-5-5-6/paras2","w+")
 g.write("%d\n" %totaltests)
 g.write("{%d,%d,%d,%d}\n" %(expnum,linearnum,expfml,linearfml))
 
-for testno in range(1,totaltests+1):
+for testno in range(100+1,100+totaltests+1):
     
     for i in range(1, expnum+linearnum+1):
         x[i] = Int('x'+str(i))
@@ -76,7 +76,7 @@ for testno in range(1,totaltests+1):
         s.add(expr <= num)
         g.write("%d}\n" % num)
             
-    f = open('new-test/3-4-4-4/test'+str(testno),mode='w+')
+    f = open('new-test/4-5-5-6/test'+str(testno),mode='w+')
     f.write("(set-logic ALL)\n")
     f.write("(define-fun-rec pow ((x Int)) Int (ite (= x 0) 1 (* 10 (pow (- x 1)) ) ) )\n")
     f.write(s.sexpr())

### 2021-05-18 Benchmark/NIA-test/runall.sh

@@ -2,19 +2,16 @@
 echo "begin test..."
 
 cd new-test
-    cd 3-4-4-4
-    ./runz3.sh > z3-result 
+    cd 2-3-3-4
+    #./runz3.sh > z3-result2 
+    ./runcvc4.sh > cvc4-result
     cd ..
-    #cd 2-3-3-4
-    #./runz3.sh > z3-result 
-    #./runcvc4.sh > cvc4-result
-    #cd ..
-    #cd 3-4-4-5
-    #./runz3.sh > z3-result 
-    #./runcvc4.sh > cvc4-result
-    #cd ..
-    #cd 4-5-5-6
-    #./runz3.sh > z3-result 
-    #./runcvc4.sh > cvc4-result
-    #cd ..
+    cd 3-4-4-5
+    #./runz3.sh > z3-result2 
+    ./runcvc4.sh > cvc4-result
+    cd ..
+    cd 4-5-5-6
+    #./runz3.sh > z3-result2
+    ./runcvc4.sh > cvc4-result
+    cd ..
 cd ..

### 2021-05-18 FMCAD2021/6_impl_exp.tex
add trau result

@@
     \begin{tabular}{cc|cc|cc|cc}
-        \multicolumn{1}{l}{} &  & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c}{MySolver} \\
+        \multicolumn{1}{l}{} &  & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c}{$\paexp$-solver} \\
         \multicolumn{1}{l}{} &  & number & avg & number & avg & number & avg \\ \hline
@@
     \caption{Results of Z3, CVC4 and $\paexp$-Solver on string hash function benchmark}
-    \begin{tabular}{c|c|ccc|ccc|ccc}
-                                                       &     & \multicolumn{3}{c|}{Z3} & \multicolumn{3}{c|}{CVC4} & \multicolumn{3}{c}{MySolver}       \\
-    PATTERN                                            & $\ell_{10}(m_1)$   & SAT  & UNSAT  & TIMEOUT & SAT   & UNSAT  & TIMEOUT  & SAT & UNSAT & TIMEOUT              \\ \hline
-    \multirow{5}{*}{12345(a)*(b)*}                     & 2   & 5    &        & 20      & 23    &        & 2        & 24  & 1     & 0                    \\
-                                                       & 3   & 0    &        & 25      & 4     &        & 21       & 10  & 14    & 1                    \\
-                                                       & 4   & 0    &        & 25      & 1     &        & 24       & 2   & 22    & 1                    \\
-                                                       & 5   & 0    &        & 25      & 1     &        & 24       & 1   & 23    & 1                    \\
-                                                       & total & 5    &        & 95      & 29    &        & 71       & 37  & 60    & 3                    \\ \hline
-    \multirow{5}{*}{12345(a)*(b)*6789}                 & 2   & 10   &        & 15      & 18    &        & 7        & 24  & 1     & 0                    \\
-                                                       & 3   & 1    &        & 24      & 8     &        & 17       & 10  & 15    & 0                    \\
-                                                       & 4   & 0    &        & 25      & 3     &        & 22       & 3   & 22    & 0                    \\
-                                                       & 5   & 0    &        & 25      & 0     &        & 25       & 0   & 25    & 0                    \\
-                                                       & total & 11   &        & 89      & 29    &        & 71       & 37  & 63    & 0                    \\ \hline
-    \multirow{5}{*}{(a)*(b)*6789}                      & 2   & 14   &        & 11      & 23    &        & 2        & 22  & 3     & 0                    \\
-                                                       & 3   & 4    &        & 21      & 6     & 1      & 18       & 15  & 10    & 0                    \\
-                                                       & 4   & 0    &        & 25      & 1     &        & 24       & 3   & 22    & 0                    \\
-                                                       & 5   & 0    &        & 25      & 0     &        & 25       & 1   & 24    & 0                    \\
-                                                       & total & 18   &        & 82      & 30    & 1      & 69       & 41  & 59    & 0                    \\ \hline\hline
-    \multirow{5}{*}{12345$(\Sigma_{num})^*$}     & 2   & 25   &        & 0       & 25    &        & 0        & 25  &       &                      \\
-                                                       & 3   & 25   &        & 0       & 25    &        & 0        & 25  &       &                      \\
-                                                       & 4   & 23   &        & 2       & 25    &        & 0        & 25  &       &                      \\
-                                                       & 5   & 9    &        & 16      & 25    &        & 0        & 25  &       &                      \\
-                                                       & total & 82   &        & 18      & 100   &        & 0        & 100 &       &                      \\ \hline
-    \multirow{5}{*}{12345$(\Sigma_{num})^*$6789} & 2   & 25   &        & 0       & 15    &        & 10       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 3   & 23   &        & 2       & 1     &        & 24       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 4   & 9    &        & 16      & 1     &        & 24       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 5   & 3    &        & 22      & 0     &        & 25       & 25  &       &                      \\
-                                                       & total & 60   &        & 40      & 17    &        & 83       & 100 &       &                      \\ \hline
-    \multirow{5}{*}{$(\Sigma_{num})^*$6789}      & 2   & 25   &        & 0       & 19    &        & 6        & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 3   & 25   &        & 0       & 4     &        & 21       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 4   & 16   &        & 9       & 4     &        & 21       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 5   & 2    &        & 23      & 0     &        & 25       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & total & 68   &        & 32      & 27    &        & 73       & 100 &       &                     
+    \begin{tabular}{cc|ccc|ccc|ccc|ccc}
+        &  & \multicolumn{3}{c|}{Z3} & \multicolumn{3}{c|}{CVC4} & \multicolumn{3}{c|}{Trau} & \multicolumn{3}{c}{$\paexp$-solver} \\
+       PATTERN & k & SAT & UNSAT & FAIL & SAT & UNSAT & FAIL & SAT & UNSAT & FAIL & SAT & UNSAT & FAIL \\ \hline
+       \multirow{5}{*}{$12345(a)^+(b)^+$} & 2 & 5 &  & 20 & 23 &  & 2 & 3 & 1 & 21 & 24 & 1 & 0 \\
+        & 3 & 0 &  & 25 & 4 &  & 21 & 0 & 14 & 11 & 10 & 14 & 1 \\
+        & 4 & 0 &  & 25 & 1 &  & 24 & 0 & 22 & 3 & 2 & 22 & 1 \\
+        & 5 & 0 &  & 25 & 1 &  & 24 & 0 & 23 & 2 & 1 & 23 & 1 \\
+        & all & 5 &  & 95 & 29 &  & 71 & 5 & 60 & 32 & 37 & 60 & 3 \\ \hline
+       \multirow{5}{*}{$12345(a)^+(b)^+6789$} & 2 & 10 &  & 15 & 18 &  & 7 & 0 & 1 & 24 & 24 & 1 &  \\
+        & 3 & 1 &  & 24 & 8 &  & 17 & 0 & 15 & 10 & 10 & 15 &  \\
+        & 4 & 0 &  & 25 & 3 &  & 22 & 0 & 22 & 3 & 3 & 22 &  \\
+        & 5 & 0 &  & 25 & 0 &  & 25 & 0 & 25 & 0 & 0 & 25 &  \\
+        & all & 11 &  & 89 & 29 &  & 71 & 10 & 63 & 37 & 37 & 63 &  \\ \hline
+       \multirow{5}{*}{$(a)^+(b)^+6789$} & 2 & 14 &  & 11 & 23 & 0 & 2 & 2 & 3 & 20 & 22 & 3 &  \\
+        & 3 & 4 &  & 21 & 6 & 1 & 18 & 0 & 10 & 15 & 15 & 10 &  \\
+        & 4 & 0 &  & 25 & 1 & 0 & 24 & 0 & 22 & 3 & 3 & 22 &  \\
+        & 5 & 0 &  & 25 & 0 & 0 & 25 & 0 & 24 & 1 & 1 & 24 &  \\
+        & all & 18 &  & 82 & 30 & 1 & 69 & 2 & 59 & 39 & 41 & 59 &  \\ \hline
+       \multirow{5}{*}{$12345\Sigma^*$} & 2 & 25 &  & 0 & 25 &  & 0 & 25 &  & 0 & 25 &  &  \\
+        & 3 & 25 &  & 0 & 25 &  & 0 & 2 &  & 23 & 25 &  &  \\
+        & 4 & 23 &  & 2 & 25 &  & 0 & 1 &  & 24 & 25 &  &  \\
+        & 5 & 9 &  & 16 & 25 &  & 0 & 0 &  & 25 & 25 &  &  \\
+        & all & 82 &  & 18 & 100 &  & 0 & 28 &  & 72 & 100 &  &  \\ \hline
+       \multirow{5}{*}{$12345\Sigma^*6789$} & 2 & 25 &  & 0 & 15 &  & 10 & 3 &  & 22 & 25 &  & \multicolumn{1}{l}{} \\
+        & 3 & 23 &  & 2 & 1 &  & 24 & 0 &  & 25 & 25 &  & \multicolumn{1}{l}{} \\
+        & 4 & 9 &  & 16 & 1 &  & 24 & 0 &  & 25 & 25 &  & \multicolumn{1}{l}{} \\
+        & 5 & 3 &  & 22 & 0 &  & 25 & 0 &  & 25 & 25 &  &  \\
+        & all & 60 &  & 40 & 17 &  & 83 & 3 &  & 97 & 100 &  &  \\ \hline
+       \multirow{5}{*}{$\Sigma^*6789$} & 2 & 25 &  & 0 & 19 &  & 6 & 22 &  & 3 & 25 &  & \multicolumn{1}{l}{} \\
+        & 3 & 25 &  & 0 & 4 &  & 21 & 2 &  & 24 & 25 &  & \multicolumn{1}{l}{} \\
+        & 4 & 16 &  & 9 & 4 &  & 21 & 0 &  & 25 & 25 &  & \multicolumn{1}{l}{} \\
+        & 5 & 2 &  & 23 & 0 &  & 25 & 0 &  & 25 & 25 &  & \multicolumn{1}{l}{} \\
+        & all & 68 &  & 32 & 27 &  & 73 & 24 &  & 76 & 100 &  & 
        \end{tabular}

### 2021-05-18 FMCAD2021/4_dec_proc.tex
Z3&mysolver result on Arith
section 3

@@
 Remedying this, Point proposed a quantifier elimination algorithm for {\paexp} in \cite{Point86}.
-In the sequel, we are going to illustrate how Point's quantifier elimination algorithm works.
-Points' algorithm is non-elementary in the worst case and a faithful implementation of the algorithm would not scale\footnote{We did implemented Point's algorithm and discovered that the implementation could only solve formulas of very small size.}.
-In the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
+The goal of this section is to illustrate how Point's quantifier elimination algorithm works. Moreover, we give a complexity analysis to Point's algorithm, which was missing in \cite{Point86}.
+%
+\begin{theorem}\label{thm-paexp}
+{\paexp} admits quantifier elimination of doubly-exponential complexity.
+\end{theorem}
+%
+As Point's algorithm is quite expensive and a faithful implementation would not scale\footnote{We did implemented Point's algorithm and discovered that the implementation could only solve formulas of very small size.}, in the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
 
-Specifically, in the sequel, we are going to show that every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
+In the sequel, we are going to show that every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
 % such that $\free(\varphi') = \free(\varphi) \setminus \{\ivarx\}$. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 
@@
 this upper bound still holds.
+
+}

### 2021-05-18 FMCAD2021/3_flatten.tex
section 3
polish

@@
 Javascript is more general in the sense that the base can be a number between 2 and 36.
-Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for readibility.}.
+Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for readability.}.
 %Since we focus on the decidability,
@@
 %\begin{definition} 
-Let $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$. Then ${\parseInt}: \Sigma_{\textit{num}}^+ \mapsto \Nat$ is recursively defined by
+In order to simplify presentation, we assume all string variables ranging over numerical symbols $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$.
+Note that one can easily extend our approach to allow arbitrary finite alphabet.
+Then ${\parseInt}: \Sigma_{\textit{num}}^+ \mapsto \Nat$ is recursively defined by
     for every $w\in \Sigma_{\textit{num}}^+$,
@@
 %        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
-        \item if $w={'i'}$ for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}('i')=i$;
+        \item \yfc{better use `$i$' or $``i"$ instead of $'i'$} if $w={'i'}$ for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}('i')=i$;
         \item for $w = w_1 {'i'}$ for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
@@
 %
-$$\llbracket \phi \rrbracket_\kell =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)})$$
+$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)})$$
 %
@@
 \end{itemize}
+Notice that here $|w_{\svarx, \textunderscore } |$ and $\ell_{\svarx, \textunderscore}$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%The correctness of the flattening approach is stated below.
+%\begin{proposition}
+%$\llbracket \varphi \rrbracket_{\flatfun_\varphi}$ is nonempty iff $\flatten_{\flatfun_\varphi}(\varphi)$ is satisfiable.
+%\end{proposition}

### 2021-05-18 FMCAD2021/6_impl_exp.tex
modify stringhash table

--- 2021-05-18 FMCAD2021/6_impl_exp.tex
@@
     \caption{Results of Z3, CVC4 and $\paexp$-Solver on string hash function benchmark}
-    \begin{tabular}{c|c|ccc|ccc|ccc}
-                                                       &     & \multicolumn{3}{c|}{Z3} & \multicolumn{3}{c|}{CVC4} & \multicolumn{3}{c}{MySolver}       \\
-    PATTERN                                            & $\ell_{10}(m_1)$   & SAT  & UNSAT  & TIMEOUT & SAT   & UNSAT  & TIMEOUT  & SAT & UNSAT & TIMEOUT              \\ \hline
-    \multirow{5}{*}{12345(a)*(b)*}                     & 2   & 5    &        & 20      & 23    &        & 2        & 24  & 1     & 0                    \\
-                                                       & 3   & 0    &        & 25      & 4     &        & 21       & 10  & 14    & 1                    \\
-                                                       & 4   & 0    &        & 25      & 1     &        & 24       & 2   & 22    & 1                    \\
-                                                       & 5   & 0    &        & 25      & 1     &        & 24       & 1   & 23    & 1                    \\
-                                                       & total & 5    &        & 95      & 29    &        & 71       & 37  & 60    & 3                    \\ \hline
-    \multirow{5}{*}{12345(a)*(b)*6789}                 & 2   & 10   &        & 15      & 18    &        & 7        & 24  & 1     & 0                    \\
-                                                       & 3   & 1    &        & 24      & 8     &        & 17       & 10  & 15    & 0                    \\
-                                                       & 4   & 0    &        & 25      & 3     &        & 22       & 3   & 22    & 0                    \\
-                                                       & 5   & 0    &        & 25      & 0     &        & 25       & 0   & 25    & 0                    \\
-                                                       & total & 11   &        & 89      & 29    &        & 71       & 37  & 63    & 0                    \\ \hline
-    \multirow{5}{*}{(a)*(b)*6789}                      & 2   & 14   &        & 11      & 23    &        & 2        & 22  & 3     & 0                    \\
-                                                       & 3   & 4    &        & 21      & 6     & 1      & 18       & 15  & 10    & 0                    \\
-                                                       & 4   & 0    &        & 25      & 1     &        & 24       & 3   & 22    & 0                    \\
-                                                       & 5   & 0    &        & 25      & 0     &        & 25       & 1   & 24    & 0                    \\
-                                                       & total & 18   &        & 82      & 30    & 1      & 69       & 41  & 59    & 0                    \\ \hline\hline
-    \multirow{5}{*}{12345$(\Sigma_{num})^*$}     & 2   & 25   &        & 0       & 25    &        & 0        & 25  &       &                      \\
-                                                       & 3   & 25   &        & 0       & 25    &        & 0        & 25  &       &                      \\
-                                                       & 4   & 23   &        & 2       & 25    &        & 0        & 25  &       &                      \\
-                                                       & 5   & 9    &        & 16      & 25    &        & 0        & 25  &       &                      \\
-                                                       & total & 82   &        & 18      & 100   &        & 0        & 100 &       &                      \\ \hline
-    \multirow{5}{*}{12345$(\Sigma_{num})^*$6789} & 2   & 25   &        & 0       & 15    &        & 10       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 3   & 23   &        & 2       & 1     &        & 24       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 4   & 9    &        & 16      & 1     &        & 24       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 5   & 3    &        & 22      & 0     &        & 25       & 25  &       &                      \\
-                                                       & total & 60   &        & 40      & 17    &        & 83       & 100 &       &                      \\ \hline
-    \multirow{5}{*}{$(\Sigma_{num})^*$6789}      & 2   & 25   &        & 0       & 19    &        & 6        & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 3   & 25   &        & 0       & 4     &        & 21       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 4   & 16   &        & 9       & 4     &        & 21       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & 5   & 2    &        & 23      & 0     &        & 25       & 25  &       & \multicolumn{1}{l}{} \\
-                                                       & total & 68   &        & 32      & 27    &        & 73       & 100 &       &                     
+    \centering
+    \begin{tabular}{cc|cc|cc|cc|cc}
+        &  & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c}{mysovler} \\
+       PATTERN &  & number & time & number & time & number & time & number & time \\ \hline
+       \multirow{3}{*}{12345(a)+(b)+} & SAT & 5 & 14.0 & 29 & 8.5 & 3 & \textless{}0.1 & 37 & 9.9 \\
+        & UNSAT & 0 &  & 0 &  & 60 & 1.3 & 60 & 47.2 \\
+        & FAIL & 95 &  & 71 &  & 32 &  & 3 &  \\ \hline
+       \multirow{3}{*}{12345(a)+(b)+6789} & SAT & 11 & 13.0 & 29 & 12.0 & 0 &  & 37 & 10.6 \\
+        & UNSAT & 0 &  & 0 &  & 63 & 1.2 & 63 & 50.0 \\
+        & FAIL & 89 &  & 71 &  & 37 &  & 0 &  \\ \hline
+       \multirow{3}{*}{(a)+(b)+6789} & SAT & 18 & 24.0 & 30 & 9.3 & 2 & \textless{}0.1 & 41 & 16.1 \\
+        & UNSAT & 0 &  & 1 & 4.0 & 59 & 2.5 & 59 & 45.8 \\
+        & FAIL & 82 &  & 69 &  & 39 &  & 0 &  \\ \hline
+       \multirow{3}{*}{12345(\textbackslash{}Sigma)*} & SAT & 82 & 8.7 & 100 & 2.2 & 28 & 5.9 & 100 & 18.5 \\
+        & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
+        & FAIL & 18 &  & 0 &  & 72 &  & 0 &  \\ \hline
+       \multirow{3}{*}{12345(\textbackslash{}Sigma)*6789} & SAT & 60 & 9.3 & 17 & 7.8 & 3 & 0.3 & 100 & 16.0 \\
+        & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
+        & FAIL & 40 &  & 83 &  & 97 &  & 0 &  \\ \hline
+       \multirow{3}{*}{(\textbackslash{}Sigma)*6789} & SAT & 68 & 5.5 & 27 & 13.0 & 24 & 9.0 & 100 & 15.7 \\
+        & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
+        & FAIL & 32 &  & 73 &  & 76 &  & 0 & 
        \end{tabular}

### 2021-05-18 Benchmark/HashFunction/flatten/head/runz3.sh

@@ -8,7 +8,8 @@ do
     # do something
     # start your script work here
     echo $count
-    z3 -smt2 test$count -T:60
+    #z3 -smt2 test$count -T:60
+    z3 smt.string_solver=z3str3 -smt2 test$count -T:60
     let "count++"
     # your logic ends here
     END=$(date +%s)

### 2021-05-18 FMCAD2021/4_dec_proc.tex

@@
 Therefore, a necessary condition for $10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001$ is that either $\ivarx_3 \le \ell_{10}(\ivary+1001)+1$ or $\ivarx_3 \le \ivarx_2 + 2$ holds.
+
+\yfc{why $+2$ and $+3$ in $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$?}
 \begin{itemize}
@@
 \item At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form).
-Moreover, we remove the occurrences of $\neg$ by replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, $\neg \iterm_1 = \iterm_2$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, $\neg \iterm_1 < \iterm_2$ with $\iterm_2 \le \iterm_1$, $\neg \iterm_1 \le \iterm_2$ with $\iterm_2 < \iterm_1$, and so on.
+Moreover, we remove the occurrences of $\neg$ by (a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, (b)  $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, (c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, (d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
 %
-\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ where $\ivarx$ occurs: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
+\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
 Let the resulting formula be $\varphi''$.
 %
-\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ where $\ivarx$ occurs, we have $\iterm = \ivarx$: For each occurrence of the $\ell_{10}$ in $\varphi''$, say $\ell_{10}(\iterm)$, such that it contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$,
+\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $\ell_{10}$ in $\varphi''$, say $\ell_{10}(\iterm)$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$,
 %and let $\varphi''':= \exists \ivarz.\ \varphi''[\ivarz/\ell_{10}(\iterm)] \wedge 10^\ivarz \le \iterm < 10^{\ivarz + 1}$. 
@@
 
+\yfc{Maybe describe the shape of normalized formula here. After normalization, all formulae are in the form ...}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\yfc{Explain why only the largest variable can  be eliminated. Which part of the procedure we need this condition? Or if we do not have the order, what can be wrong?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-18 FMCAD2021/1_intro.tex
Update 1_intro.tex

@@
 
-In \cite{Semenov84}, Semen\"{o}v just proved the decidability and did not provide an explicit decision procedure.
+In \cite{Semenov84}, Semen\"{o}v just proved that $\paexp$ admits quantifier elimination property and 
+therefore obtained that its satisfiability is decidable, but  did not provide an explicit decision procedure.
 So, in  \cite{Point86}, Point presented a first decision procedure for the satisfiability of $\paexp$ in a mathematical way, which is difficult to understand and therefore no implementation available based on which up to now to the best of our knowledge.
 So, another contribution of this paper is that we present a more understandable algorithm for $\paexp$'s satisfiability in a computational way and its prototypical implementation.
-Although the general framework of our algorithm is quite similar to Point's,  it contains the following improvements and corrections:
-   The syntax of Presburger arithmetic with exponential functions (denoted by $\paexp$) in our setting is more natural and easy to understand; Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables.
+Although the general framework of our algorithm is quite akin to Point's,  it contains the following improvements and corrections:
+\begin{itemize}
+	\item[i)] The syntax of Presburger arithmetic with exponential functions in our setting is more natural and easy to understand; Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables, that could increase the length of the considered $\paexp$ exponentially.
 In contrast, our algorithm just assumes that formulas are NNF (negation normal form).
-Therefore, our algorithm is much easy to follow and more efficient; Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} (see later).
+Therefore, our algorithm is much easy to follow and more efficient.
+\item[ii)] Unlike in Point's algorithm to eliminate the exponential part and then the linear part of the largest variable according to the given order one by one, in contrast, our algorithm eliminates the exponential parts of
+	all variables according to the order and then the resulted linear part, that improve the efficiency very much.
+\item[iii)] Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} (see later).
+\end{itemize}
 

### 2021-05-18 FMCAD2021/1_intro.tex
Update 1_intro.tex
QE to general case

@@
 \begin{itemize}
-	\item[i)] The syntax of Presburger arithmetic with exponential functions in our setting is more natural and easy to understand; Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables, that could increase the length of the considered $\paexp$ exponentially.
+	\item[i)] The syntax of Presburger arithmetic with exponential functions in our setting is more natural and easy to understand; Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables, that could increase the length of the considered $\paexp$ formula exponentially.
 In contrast, our algorithm just assumes that formulas are NNF (negation normal form).

### 2021-05-18 FMCAD2021/6_impl_exp.tex
experimental results

@@
     \label{table:arithmetic}
-    \caption{Results of Z3, CVC4 and $\paexp$-Solver on arithmetic benchmark}
+        \caption{Experimental Results on ARITHMETIC, \\
+    $\#_{\rm prb}$: number of problems, $T_{\rm avg}$: average time in seconds}
     \centering
-    \begin{tabular}{cc|cc|cc|cc}
-        \multicolumn{1}{l}{} &  & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c}{$\paexp$-solver} \\
-        \multicolumn{1}{l}{} &  & number & avg & number & avg & number & avg \\ \hline
-        \multirow{4}{*}{2-3-3-4} & SAT & 56 & 0.4 & 42 & 2.3 & 64 & 0.4 \\
-         & UNSAT & 69 & \textless{}0.1 & 72 & \textless{}0.1 & 89 & \textless{}0.1 \\
-         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 47 & 9.5 \\
-         & TIMEOUT & 75 &  & 86 &  & 0 &  \\ \hline
-        \multirow{4}{*}{3-4-4-5} & SAT & 33 & 1.4 & 25 & 2.9 & 52 & 3.3 \\
-         & UNSAT & 59 & \textless{}0.1 & 60 & \textless{}0.1 & 88 & \textless{}0.1 \\
-         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 1 & 54.0 \\
-         & TIMEOUT & 108 &  & 115 &  & 59 &  \\ \hline
-        \multirow{4}{*}{4-5-5-6} & SAT & 35 & 1.8 & 19 & 6.6 & 47 & 22.4 \\
-         & UNSAT & 36 & 0.3 & 39 & 0.4 & 72 & 0.1 \\
-         & B-UNSAT & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l|}{} & 0 &  \\
-         & TIMEOUT & 129 &  & 142 &  & 81 & 
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|}
+    \hline
+        \multirow{2}{*}{Param. }  & \multirow{2}{*}{Output} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c|}{$\paexp$} \\
+        \cline{3-8}
+        &  & $\#_{\rm prb}$ & $T_{\rm avg}$  & $\#_{\rm prb}$ & $T_{\rm avg}$  & $\#_{\rm prb}$ & $T_{\rm avg}$  \\ \hline
+        \multirow{4}{*}{(2,3,3,4)} & SAT & 56 & {\bf 0.4} & 42 & 2.3 & {\bf 64} & {\bf 0.4} \\
+        \cline{2-8}
+         & UNSAT & 69 & {\bf \textless{}0.1} & 72 &  {\bf \textless{}0.1} & {\bf 89} & {\bf \textless{}0.1} \\
+         \cline{2-8}
+         & B-UNSAT & - & - & - & - & 47 & 9.5 \\
+         \cline{2-8}
+         & TIMEOUT & 75 & - & 86 & - & {\bf 0} & - \\ \hline
+         \cline{1-8}
+        \multirow{4}{*}{(3,4,4,5)} & SAT & 33 & {\bf 1.4} & 25 & 2.9 & {\bf 52} & 3.3 \\
+        \cline{2-8}
+         & UNSAT & 59 & {\bf \textless{}0.1} & 60 & {\bf \textless{}0.1} & {\bf 88} & {\bf \textless{}0.1} \\
+         \cline{2-8}
+         & B-UNSAT & -  & - &  - & -  & 1 & 54.0 \\
+         \cline{2-8}
+         & TIMEOUT & 108 & - & 115 & - & {\bf 59} & - \\ \hline
+         \cline{1-8}
+        \multirow{4}{*}{(4,5,5,6)} & SAT & 35 & {\bf 1.8} & 19 & 6.6 & {\bf 47} & 22.4 \\
+        \cline{2-8}
+         & UNSAT & 36 & 0.3 & 39 & 0.4 & {\bf 72} & {\bf 0.1} \\
+         \cline{2-8}
+         & B-UNSAT & -  & -  & - & - & 0 & - \\
+         \cline{2-8}
+         & TIMEOUT & 129 & - & 142 & - & {\bf 81} & -\\
+         \hline
         \end{tabular}
@@
     \label{table:string}
-    \caption{Results of Z3, CVC4 and $\paexp$-Solver on string hash function benchmark}
+    \caption{Experimental results on STRINGHASH, \\ $\#_{\rm prb}$: number of problems, $T_{\rm avg}$: average time in seconds}
     \centering
-    \begin{tabular}{cc|cc|cc|cc|cc}
-        &  & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c}{mysovler} \\
-       PATTERN &  & number & time & number & time & number & time & number & time \\ \hline
-       \multirow{3}{*}{12345(a)+(b)+} & SAT & 5 & 14.0 & 29 & 8.5 & 3 & \textless{}0.1 & 37 & 9.9 \\
-        & UNSAT & 0 &  & 0 &  & 60 & 1.3 & 60 & 47.2 \\
-        & FAIL & 95 &  & 71 &  & 32 &  & 3 &  \\ \hline
-       \multirow{3}{*}{12345(a)+(b)+6789} & SAT & 11 & 13.0 & 29 & 12.0 & 0 &  & 37 & 10.6 \\
-        & UNSAT & 0 &  & 0 &  & 63 & 1.2 & 63 & 50.0 \\
-        & FAIL & 89 &  & 71 &  & 37 &  & 0 &  \\ \hline
-       \multirow{3}{*}{(a)+(b)+6789} & SAT & 18 & 24.0 & 30 & 9.3 & 2 & \textless{}0.1 & 41 & 16.1 \\
-        & UNSAT & 0 &  & 1 & 4.0 & 59 & 2.5 & 59 & 45.8 \\
-        & FAIL & 82 &  & 69 &  & 39 &  & 0 &  \\ \hline
-       \multirow{3}{*}{12345(\textbackslash{}Sigma)*} & SAT & 82 & 8.7 & 100 & 2.2 & 28 & 5.9 & 100 & 18.5 \\
-        & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
-        & FAIL & 18 &  & 0 &  & 72 &  & 0 &  \\ \hline
-       \multirow{3}{*}{12345(\textbackslash{}Sigma)*6789} & SAT & 60 & 9.3 & 17 & 7.8 & 3 & 0.3 & 100 & 16.0 \\
-        & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
-        & FAIL & 40 &  & 83 &  & 97 &  & 0 &  \\ \hline
-       \multirow{3}{*}{(\textbackslash{}Sigma)*6789} & SAT & 68 & 5.5 & 27 & 13.0 & 24 & 9.0 & 100 & 15.7 \\
-        & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
-        & FAIL & 32 &  & 73 &  & 76 &  & 0 & 
+    \renewcommand{\arraystretch}{1.2}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+    \hline
+        \multirow{2}{*} {Pattern}  &  \multirow{2}{*} {Output}  &  \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
+        \cline{3-10}
+       &  & $\#_{\rm prb}$ & $T_{\rm avg}$ &  $\#_{\rm prb}$ & $T_{\rm avg}$ & $\#_{\rm prb}$ & $T_{\rm avg}$ &  $\#_{\rm prb}$ & $T_{\rm avg}$ \\ 
+       \hline
+       \cline{1-10}
+       \multirow{3}{*}{$12345(w_1)^+(w_2)^+$} & SAT & 5 & 14.0 & 29 & 8.5 & 3 & {\bf \textless{}0.1} & {\bf 37} & 9.9 \\
+         \cline{2-10}
+        & UNSAT & 0 & - & 0 & - & {\bf 60} & {\bf 1.3} & {\bf 60} & 47.2 \\
+         \cline{2-10}
+        & FAIL & 95 & - & 71 & - & 32 & - & {\bf 3} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{$12345(w_1)^+(w_2)^+6789$} & SAT & 11 & 13.0 & 29 & 12.0 & 0 & - & {\bf 37} & {\bf 10.6} \\
+        \cline{2-10}
+        & UNSAT & 0 & - & 0 & - & {\bf 63} & {\bf 1.2} & {\bf 63} & 50.0 \\
+         \cline{2-10}
+        & FAIL & 89 & - & 71 & - & 37 & - & {\bf 0} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{$(w_1)^+(w_2)^+6789$} & SAT & 18 & 24.0 & 30 & 9.3 & 2 & {\bf \textless{}0.1} & {\bf 41} & 16.1 \\
+      	 \cline{2-10}
+        & UNSAT & 0 & - & 1 & 4.0 & {\bf 59} & {\bf 2.5} & {\bf 59} & 45.8 \\
+         \cline{2-10}
+        & FAIL & 82 &  & 69 &  & 39 &  & {\bf 0} &  \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{$12345\Sigma^*_{\sf num}$} & SAT & 82 & 8.7 & {\bf 100} & {\bf 2.2} & 28 & 5.9 & {\bf 100} & 18.5 \\
+        \cline{2-10}
+        & UNSAT & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & FAIL & 18 & - & {\bf 0} & - & 72 & - & {\bf 0} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{$12345\Sigma^*_{\sf num}6789$} & SAT & 60 & 9.3 & 17 & 7.8 & 3 & {\bf 0.3} & {\bf 100} & 16.0 \\
+        \cline{2-10}
+        & UNSAT & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & FAIL & 40 & - & 83 & - & 97 & - & {\bf 0} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{$\Sigma^*_{\sf num}6789$} & SAT & 68 & 5.5 & 27 & 13.0 & 24 & {\bf 9.0} & {\bf 100} & 15.7 \\
+        \cline{2-10}
+        & UNSAT & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & FAIL & 32 & - & 73 & - & 76 & - & {\bf 0} &- \\
+        \hline
        \end{tabular}

### 2021-05-18 FMCAD2021/6_impl_exp.tex
exp results

--- 2021-05-18 FMCAD2021/6_impl_exp.tex
@@
 We consider three parameter classes, $(2, 3, 3, 4)$, $(3, 4, 4, 5)$, and $(4, 5, 5, 6)$. 
-Each group of the ARITHMETIC benchmark suite consists of 200\zhilin{check later} randomly generated problem instances.
-The coefficients of exponential terms are randomly selected from the interval $[-100, 100]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
-The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.\wuhao{Moreover, in order to include some equalities in the problem instances, we randomly select some instances and replace the occurrence of $\le$ by $=$ in the first inequality with exponential terms.} We also create SMTLib2 files for each problem instance, in order to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
+Each group of the ARITHMETIC benchmark suite consists of 200 randomly generated problem instances.
+The coefficients of exponential terms are randomly selected from the interval $[-10^2, 10^2]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
+The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.\wuhao{Moreover, in order to include some equalities in the problem instances, we randomly select some instances and replace the occurrence of $\le$ by $=$ in the first inequality with exponential terms.
+The upper bound for variables is set to be $10^{20}$ for $\paexp$-solver because the largest number represented by 64 bits is less than $10^{20}$} We also create SMTLib2 files for each problem instance, in order to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
 Because neither CVC4 nor Z3 supports the exponential functions natively, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 10$ and $f(n+1) = 10*f(n)$.
@@
 The first three are flat languages, while the latter three are not.
-We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
-For the {\paexp}-solver, we do the following.
+Each group has 100 randomly generated problem instances, where $m$ is a prime number in the interval $[10^2, 10^5]$ and $m' \le m-1$.
+\wuhao{selection of $m,m'$}We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
+For the {\paexp}-solver, we do the following:
 \begin{itemize}
@@
 
-The experiment results are summarized in Table ~\ref{xxx}.
+The experiment results are summarized in Table \ref{table:arithmetic} and Table \ref{table:string}.
 
@@
 
-\wuhao{where to place tables?}
+On ARITHMETIC benchmark (Table \ref{table:arithmetic}), $\paexp$-Solver answers roughly 50\% more SAT and (B-)UNSAT than Z3 and CVC4, which means it can tackle some difficult instances.
+However, there are some instances are answered SAT or UNSAT by other solvers but timeout in $\paexp$-solver, e.g.
+12 instances in 3-4-4-5 group.
+The reason is that $\paexp$-solver must seach for all possible assignments before answering B-UNSAT, and the number of variables with exponential occurrences has a dramatic impact on the time complexity due to the non-elementary complexity upper bound.
 
-From Table \ref{table:arithmetic}, we can see that for all cases except bounded 4-5-5-6 case, $\paexp$-Solver answers more \textit{sat} and \textit{unsat} than Z3 and CVC4.
-Our solver significantly outperforms when equalities are included.
-The average time needed to assert \textit{sat} or \textit{unsat} increase when we add more exponential variables.\wuhao{results of unsat cases are not correct!}
-
-For string hash function benchmark, we did compared our solver with Z3, CVC4 and Trau, but the experiments show that Trau would incorrectly answer \textit{unsat} in most cases even though a model can be found.
-So in \ref{table:string}, we still show the results of the other 3 solvers.
-As we can see, for flat patterns $(a)^*(b)^*$, Z3 and CVC4 would run out of time on most cases while $\paexp$-solver can give an answer for almost all trails.
-For arbitrary strings $(\Sigma_{num})^*$, CVC4 and Z3 seem to be biased against trails with surfix or prefix.
-By applying the under-approximation technique, our tool is able to find a match on all trails.
-\wuhao{more details needed}
+On STRINGHASH benchmark (Table \ref{table:string}), $\paexp$-solver significantly outperforms than all other three solvers on both two types of constraints.
+For flat regular constraints, Z3 and CVC fail on most of the instances while Trau may answer UNSAT incorrectly to some satisfiable instances.
+$\paexp$-solver work out all but 3 instances in 60 seconds and an extra experiment shows it will the remaining 3 instances can be solved in 70 seconds.
+For non-flat regular constraints, the results show a different bias for differenct pattern (with or without preffix and suffix) both in Z3 and CVC4, whereas $\paexp$-solver work out all problem instances in similar time.
 
@@
        PATTERN &  & number & time & number & time & number & time & number & time \\ \hline
-       \multirow{3}{*}{12345(a)+(b)+} & SAT & 5 & 14.0 & 29 & 8.5 & 3 & \textless{}0.1 & 37 & 9.9 \\
+       \multirow{3}{*}{$12345(w_1)^+(w_2)^+$} & SAT & 5 & 14.0 & 29 & 8.5 & 3 & \textless{}0.1 & 37 & 9.9 \\
         & UNSAT & 0 &  & 0 &  & 60 & 1.3 & 60 & 47.2 \\
         & FAIL & 95 &  & 71 &  & 32 &  & 3 &  \\ \hline
-       \multirow{3}{*}{12345(a)+(b)+6789} & SAT & 11 & 13.0 & 29 & 12.0 & 0 &  & 37 & 10.6 \\
+       \multirow{3}{*}{$12345(w_1)^+(w_2)^+6789$} & SAT & 11 & 13.0 & 29 & 12.0 & 0 &  & 37 & 10.6 \\
         & UNSAT & 0 &  & 0 &  & 63 & 1.2 & 63 & 50.0 \\
         & FAIL & 89 &  & 71 &  & 37 &  & 0 &  \\ \hline
-       \multirow{3}{*}{(a)+(b)+6789} & SAT & 18 & 24.0 & 30 & 9.3 & 2 & \textless{}0.1 & 41 & 16.1 \\
+       \multirow{3}{*}{$(w_1)^+(w_2)^+6789$} & SAT & 18 & 24.0 & 30 & 9.3 & 2 & \textless{}0.1 & 41 & 16.1 \\
         & UNSAT & 0 &  & 1 & 4.0 & 59 & 2.5 & 59 & 45.8 \\
         & FAIL & 82 &  & 69 &  & 39 &  & 0 &  \\ \hline
-       \multirow{3}{*}{12345(\textbackslash{}Sigma)*} & SAT & 82 & 8.7 & 100 & 2.2 & 28 & 5.9 & 100 & 18.5 \\
+       \multirow{3}{*}{$12345\Sigma_{\sf num}^*$} & SAT & 82 & 8.7 & 100 & 2.2 & 28 & 5.9 & 100 & 18.5 \\
         & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
         & FAIL & 18 &  & 0 &  & 72 &  & 0 &  \\ \hline
-       \multirow{3}{*}{12345(\textbackslash{}Sigma)*6789} & SAT & 60 & 9.3 & 17 & 7.8 & 3 & 0.3 & 100 & 16.0 \\
+       \multirow{3}{*}{$12345\Sigma_{\sf num}^*6789$} & SAT & 60 & 9.3 & 17 & 7.8 & 3 & 0.3 & 100 & 16.0 \\
         & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\
         & FAIL & 40 &  & 83 &  & 97 &  & 0 &  \\ \hline
-       \multirow{3}{*}{(\textbackslash{}Sigma)*6789} & SAT & 68 & 5.5 & 27 & 13.0 & 24 & 9.0 & 100 & 15.7 \\
+       \multirow{3}{*}{$\Sigma_{\sf num}^*6789$} & SAT & 68 & 5.5 & 27 & 13.0 & 24 & 9.0 & 100 & 15.7 \\
         & UNSAT & 0 &  & 0 &  & 0 &  & 0 &  \\

### 2021-05-18 Benchmark/HashFunction/runstring.sh

@@ -3,30 +3,23 @@ echo "begin test..."
 
 cd all
     cd head
-    ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
+    ./runz3str.sh > z3str-result 
     cd ..
     cd head-tail
-    ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
+    ./runz3str.sh > z3str-result
     cd ..
     cd tail
-    ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
+    ./runz3str.sh > z3str-result 
     cd ..
 cd ..
 
 cd flatten
-    cd head
-    ./runz3.sh > z3-result 
-    ./runcvc4.sh > cvc4-result
-    cd ..
     cd head-tail
-    ./runz3.sh > z3-result 
+    ./runz3str.sh > z3str-result 
     ./runcvc4.sh > cvc4-result
     cd ..
     cd tail
-    ./runz3.sh > z3-result 
+    ./runz3str.sh > z3str-result 
     ./runcvc4.sh > cvc4-result
     cd ..
 cd ..

### 2021-05-18 FMCAD2021/6_impl_exp.tex

--- 2021-05-18 FMCAD2021/6_impl_exp.tex
@@
     \centering
+
     \renewcommand{\arraystretch}{1.2}
@@
         \hline
+>>>>>>> 95d362b87ce0ea0996f188ec3ea54edfff835d3d
        \end{tabular}

### 2021-05-18 FMCAD2021/6_impl_exp.tex
delete a git message
experiments

@@
 
-The experiment results are summarized in Table \ref{table:arithmetic} and Table \ref{table:string}.
+The experiment results are summarized in Table~\ref{table:arithmetic}-\ref{table:string}, where ``Fail'' means either timeout, unknown, or wrong answers.
 
-\zhilin{stopped here}
+On the ARITHMETIC benchmark suite (Table~\ref{table:arithmetic}), the $\paexp$-solver solves around $20\%$-$60\%$ more instances than Z3, and $30\%$-$100\%$ more instances than CVC4.
+Moreover, the gap becomes bigger as the the sizes of the formulas increase, which demonstrates that the $\paexp$-solver is more scalable in solving formulas of greater sizes.
+The average time of the $\paexp$-solver is comparable with Z3 and CVC4.
+The $\paexp$-solver reports ``B-UNSAT'' for 47 instances of the $(2, 3, 3, 4)$-group, while it does not report ``B-UNSAT'' (except one) for the other two groups.
+If more time is allowed, the $\paexp$-solver can report ``B-UNSAT'' for the ``TIMEOUT'' instances in the ARITHMETIC benchmark.
 
-On ARITHMETIC benchmark (Table \ref{table:arithmetic}), $\paexp$-Solver answers roughly 50\% more SAT and (B-)UNSAT than Z3 and CVC4, which means it can tackle some difficult instances.
+%which means it can tackle some difficult instances.
 However, there are some instances are answered SAT or UNSAT by other solvers but timeout in $\paexp$-solver, e.g.
@@
 
-On STRINGHASH benchmark (Table \ref{table:string}), $\paexp$-solver significantly outperforms than all other three solvers on both two types of constraints.
+On the STRINGHASH benchmark suite (Table~\ref{table:string}), in overall, the $\paexp$-solver solves significantly more problem instances, especially those satisfiable instances, than Z3, CVC4, and Trau.
+For instance, for flat regular constraints, the $\paexp$-solver solves almost all 300 problem instances, except 3 of them\footnote{These three instances can actually be solved in 70 seconds.}, while  Z3, CVC4, Trau solve only 34, 89, 187 instances respectively.
+Trau gets wrong answers for some problem instances, e.g.
+it reports ``UNSAT'' for some satisfiable instances.
+While the $\paexp$-solver is more precise than the other solvers, it is slower than the other solvers (although still in a reasonable amount of time), especially for those unsatisfiable problem instances.
+
+% for the pattern $12345\Sigma^*_{\sf num}6789$,  the $\paexp$-solver solves all the problem instances, while Z3, CVC4, and Trau solve only 60, 17, 3 of them.
+%
+% than all other three solvers on both two types of constraints.
 For flat regular constraints, Z3 and CVC fail on most of the instances while Trau may answer UNSAT incorrectly to some satisfiable instances.
@@
 % Merge bound and unbound 
-\begin{table*}[]
-    \label{table:arithmetic}
+\begin{table*}[ht]
         \caption{Experimental Results on ARITHMETIC, \\
@@
         \end{tabular}
+            \label{table:arithmetic}
 \end{table*}
@@
 % \usepackage{multirow}
-\begin{table*}[]
-    \label{table:string}
+\begin{table*}[ht]
     \caption{Experimental results on STRINGHASH, \\ $\#_{\rm prb}$: number of problems, $T_{\rm avg}$: average time in seconds}
@@
         \hline
->>>>>>> 95d362b87ce0ea0996f188ec3ea54edfff835d3d
        \end{tabular}
+           \label{table:string}
 \end{table*}

### 2021-05-18 FMCAD2021/5_optimization.tex
optimization

@@
 
-Similarly, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-2)}$ from $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}]$ and  $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}]$.
+Similarly, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-2)}$ from $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}]$ as well as  $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}]$, and so on.
 Eventually, we obtain a {\pa} formula.
 
-\paragraph{Perform the elimination of exponential occurrences of variables in parallel on atomic formulas}
+\paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}
 
-Note that when Lemma~\ref{lem:exp-ineq} is applied to an atomic formula $\tau$ to eliminate exponential occurrences of $\ivarx$, it computes 4 parameters, namingly $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$, and obtains an equivalent {\pa} formula according to these parameters.
+Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be synchronized.
+That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, B^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$.
+Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
+Note that when Lemma~\ref{lem:exp-ineq} is applied to an atomic formula $\tau$ to eliminate exponential occurrences of $\ivarx$, it computes 4 parameters, namely $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$, and obtains an equivalent {\pa} formula according to these parameters.
 Since the parameters are used as boundary conditions in enumeration, we can choose the strictest boundary conditions and apply to all atomic formulas.
@@
 \]
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
-If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' result is reported.
+If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.
 
-\paragraph{Preprocessing with small upper bound}
+\paragraph{Preprocess with small upper bound}
 

### 2021-05-19 FMCAD2021/1_intro.tex
exp results
intro add optimization part

@@
 Logical theories over strings have to allow string concatenation, which is arguably the most fundamental operation of strings. 
-The most celebrated result concerning theories of strings is Makanin’s result on deciding the satisfiability problem for \emph{word equations}~\cite{}.
+The most celebrated result concerning theories of strings is Makanin’s result on deciding the satisfiability problem for \emph{word equations}~\cite{makanin77}.
 A simple example of a word equation is $xaby = ybax$, where $x, y$ are variables, and $a, b \in \Sigma$ are constant letters. 
@@
 However, the satisfiability problem of word equation together with length constraints (e.g., $|x|=|y|+1 \wedge wx=yx$) is still open.
-The complexity of the satisfiability of word equations with regular constraints was proven to be PSPACE-complete by Plandowski~\cite{}, after decades of improvement of the original algorithm by Makanin.
+The complexity of the satisfiability of word equations with regular constraints was proven to be PSPACE-complete by Plandowski~\cite{plandowski99}, after decades of improvement of the original algorithm by Makanin.
 
@@
 
+We propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implement.
+Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and perform the elimination of exponential occurrences of variables parallelly for all atomic formulas.
+We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.
+
 How we did it? Also mention the quantifier elimination procedure. Optimizations

### 2021-05-19 FMCAD2021/5_optimization.tex
intro add optimization part

--- 2021-05-17 FMCAD2021/5_optimization.tex
@@
 
-\paragraph{Avoid the elimination of linear occurrences of variables by under approximation}
+\paragraph{Avoid the elimination of linear occurrences of variables by under approximation}\label{para: opt under appro}
 
@@
 
-\paragraph{Perform the elimination of exponential occurrences of variables in parallel on atomic formulas}
+\paragraph{Perform the elimination of exponential occurrences of variables in parallel on atomic formulas} \label{para: opt parallel}
 

### 2021-05-19 Benchmark/HashFunction/runstring.sh

@@ -2,12 +2,6 @@
 echo "begin test..."
 
 cd all
-    cd head
-    ./runz3str.sh > z3str-result 
-    cd ..
-    cd head-tail
-    ./runz3str.sh > z3str-result
-    cd ..
     cd tail
     ./runz3str.sh > z3str-result 
     cd ..
@@ -16,10 +10,8 @@ cd ..
 cd flatten
     cd head-tail
     ./runz3str.sh > z3str-result 
-    ./runcvc4.sh > cvc4-result
     cd ..
     cd tail
     ./runz3str.sh > z3str-result 
-    ./runcvc4.sh > cvc4-result
     cd ..
 cd ..

### 2021-05-19 FMCAD2021/1_intro.tex

@@
 We propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implement.
-Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and perform the elimination of exponential occurrences of variables parallelly for all atomic formulas.
+Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
 We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.

### 2021-05-19 FMCAD2021/6_impl_exp.tex

@@
 The coefficients of exponential terms are randomly selected from the interval $[-10^2, 10^2]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
-The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.\wuhao{Moreover, in order to include some equalities in the problem instances, we randomly select some instances and replace the occurrence of $\le$ by $=$ in the first inequality with exponential terms.
-The upper bound for variables is set to be $10^{20}$ for $\paexp$-solver because the largest number represented by 64 bits is less than $10^{20}$} We also create SMTLib2 files for each problem instance, in order to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
-Because neither CVC4 nor Z3 supports the exponential functions natively, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 10$ and $f(n+1) = 10*f(n)$.
+The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.
+Moreover, aiming at a better coverage of the syntactical ingredients of {\paexp}, we randomly choose some problem instances and replace the $\le$ symbol of their first inequalities by $=$.
+The constant upper bound for the values of variables is set to be $10^{20}$, motivated by the fact that the largest 64-bit integer is less than $10^{20}$.
+We also create an SMTLib2 file for each problem instance, to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
+Because neither CVC4 nor Z3 supports the exponential functions natively, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 1$ and $f(n+1) = 10*f(n)$.
 
@@
 The first three are flat languages, while the latter three are not.
-Each group has 100 randomly generated problem instances, where $m$ is a prime number in the interval $[10^2, 10^5]$ and $m' \le m-1$.
-\wuhao{selection of $m,m'$}We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
+Each group has 100 randomly generated problem instances, where $m$ is a randomly chosen prime number in the interval $[10^2, 10^5]$ and $1 \le m' < m$ ($m'$ is not necessarily a prime number).
+We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
 For the {\paexp}-solver, we do the following:
@@
 %
-We iterate the following procedure to find a model or reach the time limit: first set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a$.
-For each assignment, we encode the resulting string constraints into the {\paexp} formulas with only one exponential variables.
+We iterate the following procedure until a model is found or the time limit is reached: Initially, set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a$.
+For each assignment, we encode the resulting string constraint into an {\paexp} formula with only one exponential variable.
 If the resulting {\paexp} formula is unsatisfiable, then we increase $k$ by 1 and repeat this process.
@@
 On the other hand, we consider patterns of the form e.g.
-$12345 (a)^* (b_1 \ldots b_k) 6789$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode.
+$(a)^* (b_1 \ldots b_k)$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode.
 
@@
 
-\subsection{Experiment Results}
+\subsection{Experiments}
 

### 2021-05-19 FMCAD2021/6_impl_exp.tex
Update 6_impl_exp.tex

@@
 % Merge bound and unbound
+
 \begin{table*}[ht]
-        \caption{Experimental Results on ARITHMETIC, \\
-    $\#_{\rm prb}$: number of problems, $T_{\rm avg}$: average time in seconds}
+        \caption{Experimental Results on ARITHMETIC, 
+    $\#_{\mathrm{prb}}$: number of problems, $T_{\mathrm{avg}}$: average time in seconds}
     \centering
@@
         \cline{3-8}
-        &  & $\#_{\rm prb}$ & $T_{\rm avg}$  & $\#_{\rm prb}$ & $T_{\rm avg}$  & $\#_{\rm prb}$ & $T_{\rm avg}$  \\ \hline
+        &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  \\ \hline
         \multirow{4}{*}{(2,3,3,4)} & SAT & 56 & {\bf 0.4} & 42 & 2.3 & {\bf 64} & {\bf 0.4} \\
@@
 \begin{table*}[ht]
-    \caption{Experimental results on STRINGHASH, \\ $\#_{\rm prb}$: number of problems, $T_{\rm avg}$: average time in seconds}
+    \caption{Experimental results on STRINGHASH,  $\#_{\mathrm{prb}}$: number of problems, $T_{\mathrm{avg}}$: average time in seconds}
     \centering
@@
         \cline{3-10}
-       &  & $\#_{\rm prb}$ & $T_{\rm avg}$ &  $\#_{\rm prb}$ & $T_{\rm avg}$ & $\#_{\rm prb}$ & $T_{\rm avg}$ &  $\#_{\rm prb}$ & $T_{\rm avg}$ \\ 
+       &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ &  $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ &  $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ \\ 
        \hline

### 2021-05-19 FMCAD2021/4_dec_proc.tex
move some paragraphs to the appendix

@@
 
-In the sequel, we are going to show that every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
+As $\forall \ivarx.
+\ \varphi$ is equivalent to $\neg \exists \ivarx.
+\ \neg \varphi$,  thus in the sequel, we only need to  show that every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
+
+% when the quantifier elimination problem of $\paexp$  can be easily reduced to the above special case.
+
 % such that $\free(\varphi') = \free(\varphi) \setminus \{\ivarx\}$. 
@@
 
+The elimination of the exponential occurrences of the divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.
+
+\hide{
 \medskip
@@
 where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+}
 

### 2021-05-19 FMCAD2021/5_optimization.tex
move some paragraphs to the appendix

--- 2021-05-18 FMCAD2021/5_optimization.tex
@@
 
+We also propose some additional optimizations, which are omitted due to the page limit and can be found in Appendix~\ref{app-opt}.
+
+\hide{
 \paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}
@@
 Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.
+}
 
@@
 
+\hide{
 \paragraph{Avoid the formula-size blow-up by depth-first search}
@@
 If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.
+}
 

### 2021-05-19 FMCAD2021/8_appendix.tex
move some paragraphs to the appendix

@@
 %!TEX root = paper.tex
+
+\subsection{Proof of Lemma~\ref{lem:exp-ineq}}
 
@@
 \end{proof}
+
+%\subsection{Complexity analysis of Point's algorithm}
+
+\subsection{Elimination of exponential occurrences of variables for divisibility atomic formulas}\label{app-div}
+
+Consider
+%
+$$
+\begin{array}{l}
+\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
+c\ \big |\ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
++ \iterm(\vec{\ivary}) \big)
+\end{array}
+$$
+with $a_i \neq 0$.
+%
+
+Let $c = 2^{r_1} 5^{r_2}  c_0$ such that $c_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $c | (10^rc_0)$.
+
+If $c_0 = 1$, then $10^r$ is divisible by $c = 2^{r_1}5^{r_2}$.
+Thus for every $n \ge r$, $c \ |\ 10^n$.
+Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to
+\[
+\small
+\begin{array}{l}
+\bigvee \limits_{p = 0}^{r - 1} c\ \big | \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
++ \iterm(\vec{\ivary}) \big)\\
+%
+\vee \big(\ivarx_{\sigma(i)} \ge r \wedge c\ \big | \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
++ \iterm(\vec{\ivary}) \big)\big),
+\end{array}
+\]
+where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+
+Next, let us assume $c_0 > 1$.
+Since $10$ and $c_0$ are relatively prime, according to Euler's theorem (cf.
+\cite{HW80}), $10^{\phi(c_0)} \equiv 1 \bmod c_0$, where $\phi$ is the Euler function.
+Suppose $10^{\phi(c_0)} = kc_0 + 1$ for some $k \in \Nat$.
+Then for every $n \in \Nat$ with $n \ge r$, 
+$$
+\begin{array}{l}
+10^{n + \phi(c_0)} \bmod c =10^{n-r} 10^r (kc_0 + 1) \bmod c = \\
+10^{n-r} (k 10^rc_0 + 10^r) \bmod c = \\
+10^{n-r} (0+10^r) \bmod c = 10^n \bmod c.
+\end{array}
+$$
+
+Then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
+\[
+\begin{array}{l}
+\bigvee \limits_{p=0}^{r-1} \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[p/\ivarx_{\sigma(i)}]\ \vee \\
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i)} \ge r\ \wedge \\
+\bigvee \limits_{q = 0}^{\phi(c_0)-1} 
+\left(
+\begin{array}{l}
+\phi(c_0)\ \big |\ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
+c\ \big | 
+\left(
+\begin{array}{l}
+a_i 10^{r+q} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \\
+\sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})
+\end{array}
+\right) 
+\end{array}
+\right)
+\end{array}
+\right),
+\end{array}
+\]
+where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+
+\subsection{Some additional optimizations of the decision procedure for {\paexp}}\label{app-opt}
+
+\paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}
+
+Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be synchronized.
+That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, B^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$.
+Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.
+
+\paragraph{Avoid the formula-size blow-up by depth-first search}
+
+The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size.
+If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory.
+Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
+If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.
+
+\paragraph{Preprocess with small upper bound}
+
+We believe that if a quantifier-free {\paexp} formula is satisfiable, then most probably it is satisfiable with small values assigned to variables.
+Consequently, as a preprocessing step, we put a small upper bound, e.g.
+the biggest constant occurring in the formula, on the values of variables, and perform a depth-first search, so that a model can be quickly found, if there is any.
+If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.

### 2021-05-19 FMCAD2021/3_flatten.tex
examples and comments

@@
 %        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
-        \item \yfc{better use `$i$' or $``i"$ instead of $'i'$} if $w={'i'}$ for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}('i')=i$;
-        \item for $w = w_1 {'i'}$ for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
-        ${\parseInt}(w) = 10*{\parseInt}(w_1)+{\parseInt}({'i'})$.
+        \item  if $w=$`$i$' for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}$$($`$i$'$)=i$;
+        \item for $w = w_1$`$i$' for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
+        ${\parseInt}(w) = 10*{\parseInt}(w_1)+{\parseInt}($`$i$'$)$.
     \end{itemize} 
@@
 \begin{example}
-some example for $\strint$ here
+$\svarx \in \Aut \wedge \parseInt(\svarx) = 109\ivarx \wedge \len(\svarx) < 100$, where $\Aut$ is an FA, is an $\strint$ formula.
 \end{example}
@@
 \begin{example}
-Some example of  {\paexp} here.
+$10^{\ivarx+ \ivary} + 3 * 10^{\ivary}  = 2 \ivarz +1 $ is an {\paexp} formula.
 \end{example}
@@
 %Moreover, we introduce an abbreviation $\lambda_{10}(\ivarx) \equiv 2^{\ell_{10}(\ivarx)}$.
-From the definition of $\lambda_{10}(\ivarx)$, we know that for all $n \ge 1$, $\lambda_{10}(n) \le n \le 10\lambda_{10}(n)-1$ holds.
+From the definition of $\lambda_{10}(\ivarx)$, we know that for all $n \ge 1$, $\lambda_{10}(n) \le n < 10\lambda_{10}(n)$ holds.
 

### 2021-05-19 FMCAD2021/4_dec_proc.tex
examples and comments

@@
 The main idea of the elimination of the exponential occurrences of $\ivarx_3$ is to observe that if $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$, then $10^{\ivarx_3} - 10^{\ivarx_2}$ is dominated by $10^{\ivarx_3}$, that is, $10^{\ivarx_3} - 10^{\ivarx_2} \ge 10^{\ivarx_3} - 10^{\ivarx_3 - 3} = (1-10^{-3}) 10^{\ivarx_3} \ge 10^{\ell_{10}(\ivary+1001)+1} = 10\lambda_{10}(\ivary+1001) > \ivary+1001$.
+(See Lemma~\ref{lem:exp-ineq} for the choice of the constants $2$ and $3$ in $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$.)
 Therefore, a necessary condition for $10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001$ is that either $\ivarx_3 \le \ell_{10}(\ivary+1001)+1$ or $\ivarx_3 \le \ivarx_2 + 2$ holds.
 
-\yfc{why $+2$ and $+3$ in $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$?}
+%\yfc{why $+2$ and $+3$ in $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$?}
 \begin{itemize}
@@
 Cooper's algorithm in \cite{Cooper72}.
-The elimination of $\ivarx_3$ in $\varphi_2$ is simple, with $\ivarx_3$ replaced by $\ivarx_1 + \ivarx_2$, resulting into
+The elimination of $\ivarx_3$ in $\varphi_2$ here is simple, with $\ivarx_3$ replaced by $\ivarx_1 + \ivarx_2$, resulting into
 \[
@@
 %
-\item Do the following replacements to $\varphi'''$ so that all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+\item Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
 Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
@@
 
-\yfc{Maybe describe the shape of normalized formula here. After normalization, all formulae are in the form ...}
+Note that after the normalization, all the exponential terms are of the form $10^{\ivarx}$ for some variable $\ivarx$.
+
+%\yfc{Maybe describe the shape of normalized formula here. After normalization, all formulae are in the form ...}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Assuming a linear order $\sigma \in \mathcal{S}_n$ of $\{\ivarx_1,\ldots, \ivarx_n\}$, we then consider $\varphi'_\sigma  = \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ and eliminate the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(1)}$,  one by one and from left to right.
+Assuming a linear order $\sigma \in \mathcal{S}_n$ of $\{\ivarx_1,\ldots, \ivarx_n\}$, we then consider $\varphi'_\sigma  = \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ and eliminate the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(1)}$,  one by one and from the innermost to the outmost.
 Let $\varphi''_\sigma$ denote the resulting formula.
@@
 
-In the sequel, assuming a linear order $\sigma \in \mathcal{S}_n$, for $i \in [n]$, let $\exists \ivarx_{\sigma(i)} \ldots \exists \ivarx_{\sigma(1)}.\ \varphi''_{\sigma,i}$ be the formula obtained from $\varphi'_\sigma$ by eliminating the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(i+1)}$, we show how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\exists \ivarx_{\sigma(i)} \ldots \exists \ivarx_{\sigma(1)}.\ \varphi''_{\sigma,i}$.
+In the sequel, assuming a linear order $\sigma \in \mathcal{S}_n$, for $i \in [n]$, let $\exists \ivarx_{\sigma(1)} \ldots \exists \ivarx_{\sigma(i)}.\ \varphi''_{\sigma,i}$ be the formula obtained from $\varphi'_\sigma$ by eliminating the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(i+1)}$, we show how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\exists \ivarx_{\sigma(1)} \ldots \exists \ivarx_{\sigma(i)}.\ \varphi''_{\sigma,i}$.
 We would like to remark that the linear occurrences of $\ivarx_{\sigma(i)}$ should be eliminated further so that the quantifier $\exists \ivarx_{\sigma(i)}$ can be eliminated.
@@
 
-\yfc{Explain why only the largest variable can  be eliminated. Which part of the procedure we need this condition? Or if we do not have the order, what can be wrong?}
+Note that the order $\ivarx_{\sigma(i)} \ge \ldots \ge \ivarx_{\sigma(1)}$ guarantees the maximum of $\ivarx_{\sigma(i)}$ among $\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}$, which is essential for the elimination of $10^{\ivarx_{\sigma(i)}}$ from $\varphi''_{\sigma,i}$ (see Lemma~\ref{lem:exp-ineq}).
+
+%\yfc{Explain why only the largest variable can  be eliminated. Which part of the procedure we need this condition? Or if we do not have the order, what can be wrong?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-20 FMCAD2021/2_preliminary.tex
section 4
normalization
move some paragraphs to the appendix
section 3 and 4

@@
 
-In this section, we fix the notations and introduce some basic concepts, including Presburger arithmetic, finite-state automata, Parikh images, flat automata and languages.
+In this section, we fix the notations and introduce some basic concepts, including Presburger arithmetic, finite-state automata, and flat languages.
 
@@
 Let $\varepsilon$ denote the empty string, namely, the empty sequence.
-For a string $w = a_1 \ldots a_n \in \Sigma^*$, let $|w|$ denote the \emph{length} of $|w|$, i.e.
+For a string $w = a_1 \ldots a_n \in \Sigma^*$, let $|w|$ denote the \emph{length} of $w$, i.e.
 $n$.
@@
  \begin{array}{ l c l}
- \iterm &::= & c \mid \ivarx \mid \iterm+\iterm \mid \iterm - \iterm, \\
- \phi & ::= & \iterm \ \op\ 0 \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,
+ \iterm &\Def & c \mid \ivarx \mid \iterm+\iterm \mid \iterm - \iterm, \\
+ \phi & \Def & \iterm \ \op\ \iterm \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,
 
@@
 where $\op \in \{=, < , >, \le, \ge\}$ and $\ivarx, c$ are integer variables and constants, respectively.
-A \emph{quantifier-free} PA (\qfpa) formula is a {\pa} formula containing no quantifiers.
+A \emph{quantifier-free} {\pa} (\qfpa) formula is a {\pa} formula containing no quantifiers.
 A {\pa} formula is in \emph{negation normal form} (NNF) if all occurrences of the negation symbol $\neg$ are before the atomic formulas.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 \paragraph{Parikh Image}
@@
 It is well known that the Parikh image of a regular language can be characterized by an existential PA formula~\cite{SSMH04}.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 We will present flat languages. This is a high level, language-theoretical view at the flat automata from \cite{Parosh:20:PLDI}. 
-
+%
 %For integers $k$ and $\ell$ and a string variable $x$, 
@@
 that conform to a regular expression of the form
-$$
+$
 (a^1_1\ldots a^1_{\ell_1})^* \cdot \ldots \cdot (a^k_1\ldots a^k_{\ell_k})^*, 
-$$
+$
 where $a^i_j \in \Sigma$ for each $i \in [k]$ and $j \in [\ell_i]$.

### 2021-05-20 FMCAD2021/4_dec_proc.tex
section 3 and 4

@@
 Sem\"{e}nov first proved that  {\paexp} admits quantifier elimination in \cite{Semenov84}, thus its satisfiability problem is decidable.
-However, Sem\"{e}nov did not give a concrete quantifier elimination algorithm.
-Remedying this, Point proposed a quantifier elimination algorithm for {\paexp} in \cite{Point86}.
-The goal of this section is to illustrate how Point's quantifier elimination algorithm works. Moreover, we give a complexity analysis to Point's algorithm, which was missing in \cite{Point86}.
+However, Sem\"{e}nov did not give a concrete quantifier elimination procedure.
+Remedying this, Point proposed a quantifier elimination procedure for {\paexp} in \cite{Point86}.
+
+In this section, we describe how Point's quantifier elimination procedure works. 
 %
-\begin{theorem}\label{thm-paexp}
-{\paexp} admits quantifier elimination of doubly-exponential complexity.
+\begin{theorem}[\cite{Point86}]
+\label{thm-paexp}
+{\paexp} admits quantifier elimination.
 \end{theorem}
 %
-As Point's algorithm is quite expensive and a faithful implementation would not scale\footnote{We did implemented Point's algorithm and discovered that the implementation could only solve formulas of very small size.}, in the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
+Compared to \cite{Point86}, the presentation here is more accessible to computer science researchers.
+Moreover, the procedure presented here slightly improves Point's procedure, in the following two aspects: 1) DNF (disjunctive normal form) was required in Point's procedure, which is not required here, 2) in Point's procedure, the divisibility constraints produced by the elimination of linear occurrences of a variable should be converted to equality constraints (by introducing fresh variables) before the elimination of exponential occurrences of some other variable, which is unnecessary here, since the divisibility constraints are directly dealt with in the elimination of exponential occurrences of variables.
+%Moreover, we give a complexity analysis to Point's algorithm, which was missing in \cite{Point86}.
+%
+%
+
+Furthermore, as Point's procedure is quite expensive and a faithful implementation would not scale\footnote{We did implemented Point's algorithm and discovered that the implementation could only solve formulas of very small size.},
+in the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
+
+\smallskip
 
@@
 Before a formal description of the quantifier elimination procedure, let us use a simple example to illustrate the main idea and give an overview of the procedure.
+
+\vspace{-4mm}
 
@@
 
-\smallskip
-
 In the remainder of this section, we are going to describe the aforementioned steps of the decision procedures: Normalization, the enumeration of the variable orders, 
@@
 
-The normalization step comprises the following sub-steps.
-\begin{enumerate}
-\item At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form).
-Moreover, we remove the occurrences of $\neg$ by (a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, (b)  $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, (c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, (d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
-%
-\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
-Let the resulting formula be $\varphi''$.
-%
-\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $\ell_{10}$ in $\varphi''$, say $\ell_{10}(\iterm)$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$,
-%and let $\varphi''':= \exists \ivarz.\ \varphi''[\ivarz/\ell_{10}(\iterm)] \wedge 10^\ivarz \le \iterm < 10^{\ivarz + 1}$. 
-and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct. Let $\varphi'''$ denote the resulting formula.  
-%
-\item Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
-Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
-$\iterm_1 > \iterm_2$) with $\iterm_1 \le \iterm_2 - 1$ (resp.
-$\iterm_2 \le \iterm_1-1$).
-Replace ever occurrence of $\iterm_1 = \iterm_2$ wtih $\iterm_2 \le \iterm_1 \wedge \iterm_1 \le \iterm_2$.
-Let $\varphi^\dag$ the resulting formula.
-%
-\item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_n$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
-$\exists \vec{\ivarz}\exists \ivarx.\ \varphi^\dag$.
-\end{enumerate}
-
-Note that after the normalization, all the exponential terms are of the form $10^{\ivarx}$ for some variable $\ivarx$.
+The normalization step first removes the occurrences of $\lambda_{10}(\iterm)$ and $\ell_{10}(\iterm)$  such that $\ivarx$ occurs  in $\iterm$, by encoding them with the exponential function.
+Moreover, for each occurrence of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, it introduces a fresh variable $\ivarz$, replaces $10^\iterm$ with $10^\ivarz$, and adds the equality $\ivarz = \iterm$.
+It also applies some additional transformations.
+After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\lambda_{10}(\iterm)$ as well as $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) moreover, all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+More details can be found in Appendix~\ref{app-norm}.
 
@@
 
+\vspace{-2mm}
 \subsection{Enumeration of the variable orders}
@@
 
-Note that the order $\ivarx_{\sigma(i)} \ge \ldots \ge \ivarx_{\sigma(1)}$ guarantees the maximum of $\ivarx_{\sigma(i)}$ among $\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}$, which is essential for the elimination of $10^{\ivarx_{\sigma(i)}}$ from $\varphi''_{\sigma,i}$ (see Lemma~\ref{lem:exp-ineq}).
+Note that the order $\ivarx_{\sigma(i)} \ge \ldots \ge \ivarx_{\sigma(1)}$ guarantees the maximality of $\ivarx_{\sigma(i)}$ among $\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}$, which is essential for the elimination of $10^{\ivarx_{\sigma(i)}}$ from $\varphi''_{\sigma,i}$ (see Lemma~\ref{lem:exp-ineq}).
 
@@
 
-Let $i \in [n]$ and $\exists \ivarx_{\sigma(i)} \ldots \exists \ivarx_{\sigma(1)}.\ \varphi''_{\sigma,i}(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ be the formula obtained from $\varphi'_\sigma$ by eliminating the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(i+1)}$.
+Let $i \in [n]$ and $\exists \ivarx_{\sigma(1)} \ldots \exists \ivarx_{\sigma(i)}.\ \varphi''_{\sigma,i}(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ be the formula obtained from $\varphi'_\sigma$ by eliminating the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(i+1)}$.
 We show how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\varphi''_{\sigma,i}$.
@@
 %
-$$a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary})$$
+$a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary})$
 or  
-$$c \big | \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big).$$
+$c \ \big  | \ \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big).$
 %
 For convenience, let us call these formulas as inequality respectively divisibility atomic formulas.
-In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these two types of atomic formulas.
+In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these the inequality atomic formulas.
+The elimination of the exponential occurrences of the divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.
 
-\medskip
-
-\paragraph{Inequality atomic formulas} Let us consider 
+Let us consider 
 $$
@@
 $$
-
+%
 The elimination of the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ relies on the following lemma.
@@
 where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.  
-
+%
 The elimination of the exponential occurrences of $\ivarx_{\sigma(i)}$ for the case $a_i < 0$ is similar.
-
-The elimination of the exponential occurrences of the divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.
 

### 2021-05-20 FMCAD2021/8_appendix.tex
section 3 and 4

@@
 %!TEX root = paper.tex
+
+\subsection{Details of the normalization step}\label{app-norm}
+%
+The normalization step comprises the following sub-steps.
+\begin{enumerate}
+\item At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form).
+Moreover, we remove the occurrences of $\neg$ by (a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, (b)  $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, (c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, (d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
+%
+\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
+Let the resulting formula be $\varphi''$.
+%
+\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $\ell_{10}$ in $\varphi''$, say $\ell_{10}(\iterm)$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$,
+%and let $\varphi''':= \exists \ivarz.\ \varphi''[\ivarz/\ell_{10}(\iterm)] \wedge 10^\ivarz \le \iterm < 10^{\ivarz + 1}$. 
+and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct. Let $\varphi'''$ denote the resulting formula.  
+%
+\item Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
+$\iterm_1 > \iterm_2$) with $\iterm_1 \le \iterm_2 - 1$ (resp.
+$\iterm_2 \le \iterm_1-1$).
+Replace ever occurrence of $\iterm_1 = \iterm_2$ wtih $\iterm_2 \le \iterm_1 \wedge \iterm_1 \le \iterm_2$.
+Let $\varphi^\dag$ the resulting formula.
+%
+\item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_n$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
+$\exists \vec{\ivarz}\exists \ivarx.\ \varphi^\dag$.
+\end{enumerate}
 

### 2021-05-20 FMCAD2021/6_impl_exp.tex
section VI

@@
 \paragraph{The STRINGHASH benchmark suite} 
-This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ such that for each $w= a_1 \ldots a_n$, 
-${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m$, where $p, m \in \Int^+$ with $p, m \ge 2$.
-It is easy to see that ${\sf hash}(w)$ can be seen as a generalization of $\parseInt$ followed by a modulo operation.
-In particular, if $\Sigma = \Sigma_{\sf num}$ and $p = 10$, then ${\sf hash}(w) = \parseInt(w) \bmod m$.
-Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, e.g.
-CVC4, Z3, and Trau, we assume $\Sigma =\Sigma_{\sf num}$ and $p = 10$, so that the string hash function can be encoded by $\parseInt$.
+This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
+%Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, we assume that $\Sigma =\Sigma_{\sf num}$ and $p = 10$ and the string hash function is encoded by $\parseInt$.
+%, so that the string hash function can be encoded by $\parseInt$.
 
@@
 Each group has 100 randomly generated problem instances, where $m$ is a randomly chosen prime number in the interval $[10^2, 10^5]$ and $1 \le m' < m$ ($m'$ is not necessarily a prime number).
-We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
-For the {\paexp}-solver, we do the following:
+We generate the SMTLib2 files for these string constraints, as inputs to the string constraint solvers.
+On the other hand, for the {\paexp}-solver, we do the following:
 \begin{itemize}
@@
 \end{itemize}
-We would like to remark that the flattening strategy for non-flat regular constraints is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(b_1...b_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
+We would like to remark that the flattening strategy for non-flat regular constraints here is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(b_1...b_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
 On the other hand, we consider patterns of the form e.g.
-$(a)^* (b_1 \ldots b_k)$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode.
+$(a)^* (b_1 \ldots b_k)$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode in general.
 

### 2021-05-20 FMCAD2021/6_impl_exp.tex
section VI

@@
 We consider three parameter classes, $(2, 3, 3, 4)$, $(3, 4, 4, 5)$, and $(4, 5, 5, 6)$. 
-Each group of the ARITHMETIC benchmark suite consists of 200 randomly generated problem instances.
-The coefficients of exponential terms are randomly selected from the interval $[-10^2, 10^2]$ and the other coefficients/constants are randomly selected from the interval $[-10^5, 10^5]$.
+Each group of the benchmark suite consists of 200 randomly generated problem instances.
+The coefficients of exponential terms are randomly selected from the interval $[-10^2, 10^2]$ and the other coefficients/constants are randomly selected from $[-10^5, 10^5]$.
 The two intervals are chosen with the intention that the coefficients of exponential terms are smaller so that they do not always dominate the left-hand side of the inequalities.
@@
 \end{itemize}
-All the experiments are run on a lap-top with the Intel i5 1.4GHz CPU and 8GB memory. We set the timeout threshold as 60 seconds per problem instance.
+All the experiments are run on a lap-top with the Intel i5 1.4GHz CPU and 8GB memory. We set the time limit as 60 seconds per problem instance.
 
@@
 The $\paexp$-solver reports ``B-UNSAT'' for 47 instances of the $(2, 3, 3, 4)$-group, while it does not report ``B-UNSAT'' (except one) for the other two groups.
-If more time is allowed, the $\paexp$-solver can report ``B-UNSAT'' for the ``TIMEOUT'' instances in the ARITHMETIC benchmark.
+If more time is allowed, the $\paexp$-solver can report ``B-UNSAT'' for the ``TIMEOUT'' instances.
 
@@
 it reports ``UNSAT'' for some satisfiable instances.
-While the $\paexp$-solver is more precise than the other solvers, it is slower than the other solvers (although still in a reasonable amount of time), especially for those unsatisfiable problem instances.
+While the $\paexp$-solver solves more instances than the other solvers, it is slower (although still in a reasonable amount of time), especially for those unsatisfiable instances.
 

### 2021-05-20 FMCAD2021/6_impl_exp.tex
experiments
expriment

@@
 
-\begin{table*}[ht]
-        \caption{Experimental Results on ARITHMETIC, 
+\begin{table}[ht]
+        \caption{Experimental Results, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
     $\#_{\mathrm{prb}}$: number of problems, $T_{\mathrm{avg}}$: average time in seconds}
@@
     \renewcommand{\arraystretch}{1.2}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
     \hline
-        \multirow{2}{*}{Param. }  & \multirow{2}{*}{Output} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c|}{$\paexp$} \\
-        \cline{3-8}
-        &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  \\ \hline
-        \multirow{4}{*}{(2,3,3,4)} & SAT & 56 & {\bf 0.4} & 42 & 2.3 & {\bf 64} & {\bf 0.4} \\
-        \cline{2-8}
-         & UNSAT & 69 & {\bf \textless{}0.1} & 72 &  {\bf \textless{}0.1} & {\bf 89} & {\bf \textless{}0.1} \\
-         \cline{2-8}
-         & B-UNSAT & - & - & - & - & 47 & 9.5 \\
-         \cline{2-8}
-         & TIMEOUT & 75 & - & 86 & - & {\bf 0} & - \\ \hline
-         \cline{1-8}
-        \multirow{4}{*}{(3,4,4,5)} & SAT & 33 & {\bf 1.4} & 25 & 2.9 & {\bf 52} & 3.3 \\
-        \cline{2-8}
-         & UNSAT & 59 & {\bf \textless{}0.1} & 60 & {\bf \textless{}0.1} & {\bf 88} & {\bf \textless{}0.1} \\
-         \cline{2-8}
-         & B-UNSAT & -  & - &  - & -  & 1 & 54.0 \\
-         \cline{2-8}
-         & TIMEOUT & 108 & - & 115 & - & {\bf 59} & - \\ \hline
-         \cline{1-8}
-        \multirow{4}{*}{(4,5,5,6)} & SAT & 35 & {\bf 1.8} & 19 & 6.6 & {\bf 47} & 22.4 \\
-        \cline{2-8}
-         & UNSAT & 36 & 0.3 & 39 & 0.4 & {\bf 72} & {\bf 0.1} \\
-         \cline{2-8}
-         & B-UNSAT & -  & -  & - & - & 0 & - \\
-         \cline{2-8}
-         & TIMEOUT & 129 & - & 142 & - & {\bf 81} & -\\
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{Out} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
+        \cline{3-10}
+        &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & - & - & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  \\ \hline
+        \multirow{4}{*}{
+        \newline (2,3,3,4)} & S & 56 & {\bf 0.4} & 42 & 2.3 & - & - &  {\bf 64} & {\bf 0.4} \\
+        \cline{2-10}
+         & U & 69 & {\bf \textless{}0.1} & 72 &  {\bf \textless{}0.1} & - & - &  {\bf 89} & {\bf \textless{}0.1} \\
+         \cline{2-10}
+         & B & - & - & - & - & - & - &  47 & 9.5 \\
+         \cline{2-10}
+         & F & 75 & - & 86 & - & - & - &  {\bf 0} & - \\ \hline
+         \cline{1-10}
+        \multirow{4}{*}{(3,4,4,5)} & S & 33 & {\bf 1.4} & 25 & 2.9 & - & - &  {\bf 52} & 3.3 \\
+        \cline{2-10}
+         & U & 59 & {\bf \textless{}0.1} & 60 & {\bf \textless{}0.1} & - & - &  {\bf 88} & {\bf \textless{}0.1} \\
+         \cline{2-10}
+         & B & -  & - &  - & -  & - & - &  1 & 54.0 \\
+         \cline{2-10}
+         & F & 108 & - & 115 & - & - & - &  {\bf 59} & - \\ \hline
+         \cline{1-10}
+        \multirow{4}{*}{(4,5,5,6)} & S & 35 & {\bf 1.8} & 19 & 6.6 & - & - &  {\bf 47} & 22.4 \\
+        \cline{2-10}
+         & U & 36 & 0.3 & 39 & 0.4 & - & - &  {\bf 72} & {\bf 0.1} \\
+         \cline{2-10}
+         & B & -  & -  & - & - & - & - &  0 & - \\
+         \cline{2-10}
+         & F & 129 & - & 142 & - & - & - &  {\bf 81} & -\\
          \hline
+         \cline{1-10}
+        \multirow{3}{*}{Flat} & S &  &  &  &  & &&   & \\
+        \cline{2-10}
+         & U &  &  &  &  & &&  &  \\
+         \cline{2-10}
+         & F &  &  &  & & &&  &  \\ \hline
+         \cline{1-10}
+        \multirow{3}{*}{Non-flat} & S &  &  &  &  & &&  & \\
+        \cline{2-10}
+         & U &  &  &  &  & && &  \\
+         \cline{2-10}
+         & F &  &  &  & & &&  &  \\ \hline
+         \cline{1-10}
         \end{tabular}
             \label{table:arithmetic}
-\end{table*}
-
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-\begin{table*}[ht]
-    \caption{Experimental results on STRINGHASH,  $\#_{\mathrm{prb}}$: number of problems, $T_{\mathrm{avg}}$: average time in seconds}
-    \centering
-
-    \renewcommand{\arraystretch}{1.2}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-    \hline
-        \multirow{2}{*} {Pattern}  &  \multirow{2}{*} {Output}  &  \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} & \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
-        \cline{3-10}
-       &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ &  $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ &  $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ \\ 
-       \hline
-       \cline{1-10}
-       \multirow{3}{*}{$12345(w_1)^+(w_2)^+$} & SAT & 5 & 14.0 & 29 & 8.5 & 3 & {\bf \textless{}0.1} & {\bf 37} & 9.9 \\
-         \cline{2-10}
-        & UNSAT & 0 & - & 0 & - & {\bf 60} & {\bf 1.3} & {\bf 60} & 47.2 \\
-         \cline{2-10}
-        & FAIL & 95 & - & 71 & - & 32 & - & {\bf 3} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{$12345(w_1)^+(w_2)^+6789$} & SAT & 11 & 13.0 & 29 & 12.0 & 0 & - & {\bf 37} & {\bf 10.6} \\
-        \cline{2-10}
-        & UNSAT & 0 & - & 0 & - & {\bf 63} & {\bf 1.2} & {\bf 63} & 50.0 \\
-         \cline{2-10}
-        & FAIL & 89 & - & 71 & - & 37 & - & {\bf 0} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{$(w_1)^+(w_2)^+6789$} & SAT & 18 & 24.0 & 30 & 9.3 & 2 & {\bf \textless{}0.1} & {\bf 41} & 16.1 \\
-      	 \cline{2-10}
-        & UNSAT & 0 & - & 1 & 4.0 & {\bf 59} & {\bf 2.5} & {\bf 59} & 45.8 \\
-         \cline{2-10}
-        & FAIL & 82 &  & 69 &  & 39 &  & {\bf 0} &  \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{$12345\Sigma^*_{\sf num}$} & SAT & 82 & 8.7 & {\bf 100} & {\bf 2.2} & 28 & 5.9 & {\bf 100} & 18.5 \\
-        \cline{2-10}
-        & UNSAT & 0 & - & 0 & - & 0 & - & 0 & - \\
-         \cline{2-10}
-        & FAIL & 18 & - & {\bf 0} & - & 72 & - & {\bf 0} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{$12345\Sigma^*_{\sf num}6789$} & SAT & 60 & 9.3 & 17 & 7.8 & 3 & {\bf 0.3} & {\bf 100} & 16.0 \\
-        \cline{2-10}
-        & UNSAT & 0 & - & 0 & - & 0 & - & 0 & - \\
-         \cline{2-10}
-        & FAIL & 40 & - & 83 & - & 97 & - & {\bf 0} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{$\Sigma^*_{\sf num}6789$} & SAT & 68 & 5.5 & 27 & 13.0 & 24 & {\bf 9.0} & {\bf 100} & 15.7 \\
-        \cline{2-10}
-        & UNSAT & 0 & - & 0 & - & 0 & - & 0 & - \\
-         \cline{2-10}
-        & FAIL & 32 & - & 73 & - & 76 & - & {\bf 0} &- \\
-        \hline
-       \end{tabular}
-           \label{table:string}
-\end{table*}
+\end{table}

### 2021-05-20 FMCAD2021/6_impl_exp.tex
m

--- 2021-05-20 FMCAD2021/6_impl_exp.tex
@@
 \paragraph{The STRINGHASH benchmark suite} 
-This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
-%Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, we assume that $\Sigma =\Sigma_{\sf num}$ and $p = 10$ and the string hash function is encoded by $\parseInt$.
-%, so that the string hash function can be encoded by $\parseInt$.
+This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ such that for each $w= a_1 \ldots a_n$, 
+${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m$, where $p, m \in \Int^+$ with $p, m \ge 2$.
+It is easy to see that ${\sf hash}(w)$ can be seen as a generalization of $\parseInt$ followed by a modulo operation.
+In particular, if $\Sigma = \Sigma_{\sf num}$ and $p = 10$, then ${\sf hash}(w) = \parseInt(w) \bmod m$.
+Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, e.g.
+CVC4, Z3, and Trau, we assume $\Sigma =\Sigma_{\sf num}$ and $p = 10$, so that the string hash function can be encoded by $\parseInt$.
 
@@
 Each group has 100 randomly generated problem instances, where $m$ is a randomly chosen prime number in the interval $[10^2, 10^5]$ and $1 \le m' < m$ ($m'$ is not necessarily a prime number).
-We generate the SMTLib2 files for these string constraints, as inputs to the string constraint solvers.
-On the other hand, for the {\paexp}-solver, we do the following:
+We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers.
+For the {\paexp}-solver, we do the following:
 \begin{itemize}
@@
 \end{itemize}
-We would like to remark that the flattening strategy for non-flat regular constraints here is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(b_1...b_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
+We would like to remark that the flattening strategy for non-flat regular constraints is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(b_1...b_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns.
 On the other hand, we consider patterns of the form e.g.
-$(a)^* (b_1 \ldots b_k)$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode in general.
+$(a)^* (b_1 \ldots b_k)$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode.
 
@@
 % Merge bound and unbound 
-
 \begin{table*}[ht]
-        \caption{Experimental Results on ARITHMETIC, 
-    $\#_{\mathrm{prb}}$: number of problems, $T_{\mathrm{avg}}$: average time in seconds}
+        \caption{Experimental Results on ARITHMETIC, \\
+    $\#_{\rm prb}$: number of problems, $T_{\mathrm avg}$: average time in seconds}
     \centering
@@
         \cline{3-8}
-        &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  \\ \hline
+        &  & $\#_{\rm prb}$ & $T_{\rm avg}$  & $\#_{\rm prb}$ & $T_{\rm avg}$  & $\#_{\rm prb}$ & $T_{\rm avg}$  \\ \hline
         \multirow{4}{*}{(2,3,3,4)} & SAT & 56 & {\bf 0.4} & 42 & 2.3 & {\bf 64} & {\bf 0.4} \\
@@
 \begin{table*}[ht]
-    \caption{Experimental results on STRINGHASH,  $\#_{\mathrm{prb}}$: number of problems, $T_{\mathrm{avg}}$: average time in seconds}
+    \caption{Experimental results on STRINGHASH, \\ $\#_{\rm prb}$: number of problems, $T_{\rm avg}$: average time in seconds}
     \centering
@@
         \cline{3-10}
-       &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ &  $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ &  $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$ \\ 
+       &  & $\#_{\rm prb}$ & $T_{\rm avg}$ &  $\#_{\rm prb}$ & $T_{\rm avg}$ & $\#_{\rm prb}$ & $T_{\rm avg}$ &  $\#_{\rm prb}$ & $T_{\rm avg}$ \\ 
        \hline

### 2021-05-20 FMCAD2021/3_flatten.tex

@@
 For example,
-${\parseInt}('0123') = {\parseInt}('123')=10^2+10*2+3 = 123$. 
-Note here we use the quotation marks to delimit the strings, 
+${\parseInt}($`$0123$'$) = {\parseInt}($`$123$'$)=10^2+10*2+3 = 123$. 
+Note here we use the quotation marks to delimit the strings.
 %Clearly,  our decision procedure given in this paper 
@@
 %\begin{definition} 
-In order to simplify presentation, we assume all string variables ranging over numerical symbols $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$.
+In order to simplify the presentation, we assume all string variables ranging over numerical symbols $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$.
 Note that one can easily extend our approach to allow arbitrary finite alphabet.
@@
 %
-The flattened semantics of $\phi \in \strint$ is defined as $\llbracket \varphi \rrbracket_{\flatfun_\varphi} = \{(I_s, I_i) \in \llbracket \varphi  \rrbracket \mid \forall \svarx \in \svar(\varphi).\ I_s(x) \in  \flatfun_\varphi(\svarx)\}$.
+The flattened semantics of $\varphi \in \strint$ is defined as $\llbracket \varphi \rrbracket_{\flatfun_\varphi} = \{(I_s, I_i) \in \llbracket \varphi  \rrbracket \mid \forall \svarx \in \svar(\varphi).\ I_s(x) \in  \flatfun_\varphi(\svarx)\}$.
 
@@
 %
-More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and Parikh variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$, called \emph{flattening variables}, plus some other auxiliary variables, such that
+More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and flattening variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$, plus some other auxiliary variables, such that
 %
@@
 
-Before presenting the construction of $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$, for every $\svarx \in \svar(\varphi)$, we define $\flatten_{\flatfun_\varphi}(\parseInt(\svarx)) = \iterm_{\svarx,1}$  such that $(\iterm_{\svarx, i})_{i \in [k_\svarx]}$ and $(\ell_{\svarx, i})_{i \in [k_\svarx]}$ are inductively defined as follows:
+For simplicity, we assume that each occurrence of $\parseInt$ (resp.
+$\len(t)$) in $\iterm_1 \op \iterm_2$ is of the form $\parseInt(\svarx)$ (resp.
+$\len(\svarx)$) for a string variable $\svarx$.
+(Otherwise, we can always introduce a fresh string variable $\svarx'$ for $\parseInt(\sterm)$ or $\len(\sterm)$ and add the constraint $\svarx' = \sterm$.)
+Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by replacing $\parseInt(\svarx)$ with $\flatten_{\flatfun_\varphi}(\parseInt(\svarx))$ and $\len(\svarx)$ with $\flatten_{\flatfun_\varphi}(\len(\svarx))$, where
+\begin{itemize}
+\item 
+$\flatten_{\flatfun_\varphi}(\parseInt(\svarx)) \Def \iterm_{\svarx,1}$  such that $(\iterm_{\svarx, i})_{i \in [k_\svarx]}$ and $(\ell_{\svarx, i})_{i \in [k_\svarx]}$ are inductively defined as follows:
 \begin{itemize}
 \item for $i = k_\svarx$, 
-$$\iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{|w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx} } -1 )}{(10^{|w_{\svarx, k_\svarx}|} -1 )}$$ 
+$$ \hspace{-4mm} \iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{|w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx} } -1 )}{(10^{|w_{\svarx, k_\svarx}|} -1 )}$$ 
 and $\ell_{\svarx, i} = |w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx}$,
@@
 %
-$$\iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{|w_{\svarx, i} | \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{|w_{\svarx, i}|} -1 )} + \iterm_{\svarx, i+1}$$
+$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{|w_{\svarx, i} | \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{|w_{\svarx, i}|} -1 )} + \iterm_{\svarx, i+1}$$
 %
 and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
+%
 \end{itemize}
 Notice that here $|w_{\svarx, \textunderscore } |$ and $\ell_{\svarx, \textunderscore}$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables.
+%
+\item $\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} |w_{\svarx, i} | \#_{\svarx, i}$. 
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
+Before presenting the construction of $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$, for every $\svarx \in \svar(\varphi)$,
 
@@
 \end{enumerate}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \Def & 1 \frac{10^{\#_{\svarx,1}}-1}{10-1}10^{\#_{\svarx,2}}  + 2 \frac{10^{\#_{\svarx,2}}-1}{10-1} = 2\ivarx   \\
-\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} - 10^{\#_{\svarx,2}}  + 2 (10^{2\#_{\svarx,2}}-1) = 18\ivarx \\
+\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} - 10^{\#_{\svarx,2}}  + 2*10^{2\#_{\svarx,2}}- 2 = 18\ivarx \\
 \equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} +  10^{\#_{\svarx,2}} = 18\ivarx + 2.

### 2021-05-20 FMCAD2021/5_optimization.tex

@@
 %
-\[\sum_{j=1}^{n} a_j 10^{\ivarx_{j}} + \sum_{k=1}^{n}b_k \ivarx_{k} \le c\]
+$\sum_{j=1}^{n} a_j 10^{\ivarx_{j}} + \sum_{k=1}^{n}b_k \ivarx_{k} \le c$, 
 %
 where $c$ is an integer constant. (Implicitly, we assume that there are no free variables and all the variables are existentially quantified.)
+
+In the sequel, we present two major optimizations and omit some additional optimizations due to the page limit, which can be found in Appendix~\ref{app-opt}.
 
@@
 
-In Section~\ref{sec-dec}, the decision procedure of {\paexp} requires the elimination of both exponential and linear occurrences of variables.
+The decision procedure of {\paexp} in Section~\ref{sec-dec} requires the elimination of both exponential and linear occurrences of variables.
 Considering the fact that {\pa} formulas can be solved efficiently by the state-of-the-art solvers, e.g.
-CVC4 and Z3, one natural idea is to try to eliminate only the exponential occurrences, but not the linear occurrences, of variables, and obtain the {\pa} formulas in the end, which can then be solved by the state-of-the-art solvers.
+CVC4 and Z3, one natural idea is to try to only eliminate the exponential occurrences, but not the linear occurrences, of variables, and obtain the {\pa} formulas in the end, which can then be solved by the state-of-the-art solvers.
 
@@
 
-Nevertheless, the side condition of Lemma~\ref{lem:exp-ineq}, namely $a_i \neq 0$, undermines the aforementioned natural idea. If $a_i = 0$, but $b_i \neq 0$, namely, 
-$$
-\begin{array}{l}
-\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
-\hspace{4mm} 
-\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c,
-\end{array}
-$$
+Nevertheless, the side condition of Lemma~\ref{lem:exp-ineq}, namely $a_i \neq 0$, undermines the aforementioned natural idea. If $a_i = 0$, but $b_i \neq 0$, 
+%namely, 
+%$$
+%\begin{array}{l}
+%\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
+%\hspace{4mm} 
+%\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c,
+%\end{array}
+%$$
 then we are unable to utilize Lemma~\ref{lem:exp-ineq} to eliminate the linear occurrences of $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$.
@@
 
-We also propose some additional optimizations, which are omitted due to the page limit and can be found in Appendix~\ref{app-opt}.
+%We also propose some additional optimizations, which are omitted due to the page limit and can be found in Appendix~\ref{app-opt}.
 

### 2021-05-20 FMCAD2021/1_intro.tex
section VI

@@
 
-The satisfiability of \emph{string/number conversion constraints} \yfc{give example} has been proven undecidable in~\cite{}. 
-It is undecidable even without combining with other types of string constraints. \yfc{double check}
+The satisfiability of \emph{string/number conversion constraints}, e.g., $wx=yx \wedge |x| > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
 However, this kind of constraint is pervasive in scripting language programs. 
@@
 The first is to consider only strings of bounded length.
-This approach is taken in earlier generation solvers such as Hampi and Kuluza~\cite{}.
+This approach is taken in earlier generation solvers such as Hampi~\cite{KiezunGAGHE12} and Kuluza~\cite{SaxenaAHMMS10}.
 Although they are useful in handling many practical cases, they fail short when the all string solutions exceed the selected bound.
@@
 
-One more recent approach is flattening~\cite{A}.
-The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat automata} (see Section~\ref{PA} for more details). 
+One more recent approach is flattening~\cite{Parosh:20:PLDI,Parosh:20:PLDI,AbdullaACDHRR18,AbdullaACDHRR17}.
+The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat automata} (see Section~\ref{sec:string-solving} for more details). 
 The major benefit of considering this class is two folds.
@@
 We propose a decision procedure for the class of word equation + regular constraints + length constraints + string/number conversion under arbitrary flat automata (denoted by $\strint$).
-The basic idea of our approach can be sketched as follows:  similar to \cite{A}, we first reduce the satisfiability problem of $\strint$ to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$); then, according to the decidability of $\paexp$ \cite{Semenov84}, we therefore obtain the decidability of the satisfiability of $\strint$.
+The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem of $\strint$ to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$); then, according to the decidability of $\paexp$ \cite{Semenov84}, we therefore obtain the decidability of the satisfiability of $\strint$.
 
@@
 
-How we did it? Also mention the quantifier elimination procedure. Optimizations
+In fact, other than the theoretical difficulties, in practice, the string-number constraints are quite challenging for state-of-the-art solvers. 
+Here we illustrate a toy example that mimic the ``mining'' step of block-chain construction.
+Essentially, giving a hash function ${\sf hash}(w)$, the goal of the mining step is to find an input $w$ such that ${\sf hash}(w)$ satisfies certain pattern, e.g., the last $k$ digits are zeros.
+Below we consider a simple hash function:
+$${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m,$$ 
+where $p, m \in \Int^+$ with $p, m \ge 2$.
+It is easy to see that ${\sf hash}(w)$ can be seen as a generalization of $\parseInt$ followed by a modulo operation.
+In particular, if $\Sigma = \Sigma_{\sf num}$ and $p = 10$, then ${\sf hash}(w) = \parseInt(w) \bmod m$.
+Thus the problem of finding a suitable input $w$ such that the last $k$ digits of ${\sf hash}(w)$ are zeros can be modeled as a $\strint$ formula.
+Although the example is seemingly simple, most state-of-the-art solvers have difficulty to handle it, c.f., Section~\ref{sec-exp} for our experimental results.
 
-Say the problem is challenging for state-of-the-art solvers, mention the hash function example.
+With several optimizations introduces in Section~\ref{sec-opt}, we manage to solve several variants of the hash function examples as well as some more randomly generated cases better than most state-of-the-art solvers.
 
-Experimental results.
+\yfc{todo: a short description of optimizations}

### 2021-05-20 FMCAD2021/6_impl_exp.tex
experiments

--- 2021-05-20 FMCAD2021/6_impl_exp.tex
@@
 \paragraph{The STRINGHASH benchmark suite} 
-This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
+This suite comprises two groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
 %Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, we assume that $\Sigma =\Sigma_{\sf num}$ and $p = 10$ and the string hash function is encoded by $\parseInt$.
@@
 where $\Aut$ is an FA, $m, m^\prime \ge 2.$ 
-The six groups of string constraints are characterized by different classes of regular constraints, namely, $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, $w^+_1 w^+_2 6789$, $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, and $\Sigma_{\sf num}^* 6789$, where $w_1,w_2 \in \Sigma^+_{\sf num}$.
-The first three are flat languages, while the latter three are not.
-Each group has 100 randomly generated problem instances, where $m$ is a randomly chosen prime number in the interval $[10^2, 10^5]$ and $1 \le m' < m$ ($m'$ is not necessarily a prime number).
+The two groups of string constraints are characterized by flat and non-flat regular constraints respectively.
+The flat group comprises 300 problem instances, where the flat languages are of the form $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, or $w^+_1 w^+_2 6789$, with $w_1,w_2 \in \Sigma^+_{\sf num}$.
+The non-flat group comprises 300 problem instances, where the non-flat languages are of the form $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, or $\Sigma_{\sf num}^* 6789$.
+Moreover, the number $m$ is a randomly chosen prime number in the interval $[10^2, 10^5]$ and $1 \le m' < m$ ($m'$ is not necessarily a prime number).
 We generate the SMTLib2 files for these string constraints, as inputs to the string constraint solvers.
@@
 \begin{itemize}
-\item For flat regular constraints, we generate the {\paexp} formulas corresponding to the string constraints, as the inputs to the {\paexp}-solver.
+\item For flat instances, we generate the {\paexp} formulas corresponding to the string constraints, as the inputs to the {\paexp}-solver.
 %
-\item For non-flat regular constraints, we use flat languages $(a)^* (b_{1} \ldots b_{k})$ to under approximate 
+\item For non-flat instances, we use flat languages $(a)^* (b_{1} \ldots b_{k})$ to under approximate 
 $\Sigma_{\sf num}^*$,  where $a, b_1,\ldots, b_k \in \Sigma_{\sf num}$. 
@@
 % Merge bound and unbound 
-
 \begin{table}[ht]
-        \caption{Experimental Results, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
-    $\#_{\mathrm{prb}}$: number of problems, $T_{\mathrm{avg}}$: average time in seconds}
+        \caption{Experimental Results, O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
+    $\#$: number of problems, $T$: average time in seconds}
     \centering
-    \renewcommand{\arraystretch}{1.2}
+    \renewcommand{\arraystretch}{1.1}
     \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
     \hline
-        \multirow{2}{*}{Group }  & \multirow{2}{*}{Out} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
         \cline{3-10}
-        &  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  & - & - & $\#_{\mathrm{prb}}$ & $T_{\mathrm{avg}}$  \\ \hline
+        &  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$} & \scriptsize{$\#$} & \scriptsize{$T$}  \\ \hline
         \multirow{4}{*}{
-        \newline (2,3,3,4)} & S & 56 & {\bf 0.4} & 42 & 2.3 & - & - &  {\bf 64} & {\bf 0.4} \\
+        \newline \scriptsize{(2,3,3,4)}} & S & 56 & {\bf 0.4} & 42 & 2.3 & - & - &  {\bf 64} & {\bf 0.4} \\
         \cline{2-10}
-         & U & 69 & {\bf \textless{}0.1} & 72 &  {\bf \textless{}0.1} & - & - &  {\bf 89} & {\bf \textless{}0.1} \\
+         & U & 69 & {\bf 0.1} & 72 &  {\bf 0.1} & - & - &  {\bf 89} & {\bf 0.1} \\
          \cline{2-10}
@@
          \cline{1-10}
-        \multirow{4}{*}{(3,4,4,5)} & S & 33 & {\bf 1.4} & 25 & 2.9 & - & - &  {\bf 52} & 3.3 \\
+        \multirow{4}{*}{\scriptsize{(3,4,4,5)}} & S & 33 & {\bf 1.4} & 25 & 2.9 & - & - &  {\bf 52} & 3.3 \\
         \cline{2-10}
-         & U & 59 & {\bf \textless{}0.1} & 60 & {\bf \textless{}0.1} & - & - &  {\bf 88} & {\bf \textless{}0.1} \\
+         & U & 59 & {\bf 0.1} & 60 & {\bf 0.1} & - & - &  {\bf 88} & {\bf 0.1} \\
          \cline{2-10}
@@
          \cline{1-10}
-        \multirow{4}{*}{(4,5,5,6)} & S & 35 & {\bf 1.8} & 19 & 6.6 & - & - &  {\bf 47} & 22.4 \\
+        \multirow{4}{*}{\scriptsize{(4,5,5,6)}} & S & 35 & {\bf 1.8} & 19 & 6.6 & - & - &  {\bf 47} & 22.4 \\
         \cline{2-10}
@@
          \cline{1-10}
-        \multirow{3}{*}{Flat} & S &  &  &  &  & &&   & \\
+        \multirow{3}{*}{\scriptsize{Flat}} 
+         & S & 34 & 19.0 & 88 & 12.7 & 5 & {\bf 0.1}& {\bf 115}  & 12.3\\
         \cline{2-10}
-         & U &  &  &  &  & &&  &  \\
+         & U & 0 & - & 1 & 4.0 & 182 & {\bf 2.5} & {\bf 182} & 47.7 \\
          \cline{2-10}
-         & F &  &  &  & & &&  &  \\ \hline
+         & F & 266 & - & 211 & - & 113 & - & {\bf 3} & - \\ \hline
          \cline{1-10}
-        \multirow{3}{*}{Non-flat} & S &  &  &  &  & &&  & \\
+        \multirow{3}{*}{\scriptsize{Non-flat}}
+         & S & 210 & 7.8 & 144 & 4.9 & 55 & 5.9 &{\bf 300} & 16.7 \\
         \cline{2-10}
-         & U &  &  &  &  & && &  \\
+         & U & 0 & - & 0 & - & 0 & - & 0 & - \\
          \cline{2-10}
-         & F &  &  &  & & &&  &  \\ \hline
+         & F & 90 & - & 156 & - & 245 & - & {\bf 0} & - \\ \hline
          \cline{1-10}

### 2021-05-20 FMCAD2021/6_impl_exp.tex
experiments

@@
 
-The experiment results are summarized in Table~\ref{table:arithmetic}-\ref{table:string}, where ``Fail'' means either timeout, unknown, or wrong answers.
+The experiment results are summarized in Table~\ref{table:arithmetic}, where ``Fail'' means either timeout, unknown, or wrong answers.
 
-On the ARITHMETIC benchmark suite (Table~\ref{table:arithmetic}), the $\paexp$-solver solves around $20\%$-$60\%$ more instances than Z3, and $30\%$-$100\%$ more instances than CVC4.
+On the ARITHMETIC benchmark suite, the $\paexp$-solver solves around $20\%$-$60\%$ more instances than Z3, and $30\%$-$100\%$ more instances than CVC4.
 Moreover, the gap becomes bigger as the the sizes of the formulas increase, which demonstrates that the $\paexp$-solver is more scalable in solving formulas of greater sizes.
@@
 
-On the STRINGHASH benchmark suite (Table~\ref{table:string}), in overall, the $\paexp$-solver solves significantly more problem instances, especially those satisfiable instances, than Z3, CVC4, and Trau.
+On the STRINGHASH benchmark suite, in overall, the $\paexp$-solver solves significantly more problem instances, especially those satisfiable instances, than Z3, CVC4, and Trau.
 For instance, for flat regular constraints, the $\paexp$-solver solves almost all 300 problem instances, except 3 of them\footnote{These three instances can actually be solved in 70 seconds.}, while  Z3, CVC4, Trau solve only 34, 89, 187 instances respectively.
@@
 it reports ``UNSAT'' for some satisfiable instances.
-While the $\paexp$-solver solves more instances than the other solvers, it is slower (although still in a reasonable amount of time), especially for those unsatisfiable instances.
+%While the $\paexp$-solver solves more instances than the other solvers,  
+From the results, we can see that the $\paexp$-solver achieves a good tradeoff between precision and efficiency, although it is slower than the other solvers.
+(More detailed experiment results on STRINGHASH can be found in Appendix~\ref{app-exp}.)
 

### 2021-05-20 FMCAD2021/1_intro.tex
some grammar mistakes

@@
 A simple example of a word equation is $xaby = ybax$, where $x, y$ are variables, and $a, b \in \Sigma$ are constant letters. 
-A word equation is satisfiable if it has a solution, i.e., an assignment that maps variables to strings over the alphabet $\Sigma$ which equate the left-hand side with the right-hand side of the equation.
+A word equation is satisfiable if it has a solution, i.e., an assignment that maps variables to strings over the alphabet $\Sigma$ which equates the left-hand side with the right-hand side of the equation.
 The correctness proof of Makanin’s algorithm is arguably one of the most complex termination proofs in computer science. 
@@
 This extension is called word equations with regular constraints. 
-However, the satisfiability problem of word equation together with length constraints (e.g., $|x|=|y|+1 \wedge wx=yx$) is still open.
+However, the satisfiability problem of word equations together with length constraints (e.g., $|x|=|y|+1 \wedge wx=yx$) is still open.
 The complexity of the satisfiability of word equations with regular constraints was proven to be PSPACE-complete by Plandowski~\cite{plandowski99}, after decades of improvement of the original algorithm by Makanin.
@@
 However, this kind of constraint is pervasive in scripting language programs. 
-For example, it is common that programs read string input from text files and conversion a part of the string input to numbers.
+For example, it is common that programs read string input from text files and converts a part of the string input to numbers.
 Even more crucially, in many programming languages, the string-number conversion is a part of the definition of their core semantics. 
-JavaScript, which powers most interactive content on the Web and increasingly server- side code with Node.js, is one of such languages.
+JavaScript, which powers most interactive content on the Web and increasingly server-side code with Node.js, is one of such languages.
 
@@
 
-One more recent approach is flattening~\cite{Parosh:20:PLDI,Parosh:20:PLDI,AbdullaACDHRR18,AbdullaACDHRR17}.
+One more recent approach is flattening~\cite{Parosh:20:PLDI,AbdullaACDHRR18,AbdullaACDHRR17}.
 The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat automata} (see Section~\ref{sec:string-solving} for more details). 
@@
 
-The paper of Abdulla et al. has considered adding string/number conversion constraints to the above class.
+The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string/number conversion constraints to the above class.
 The paper proposes a decision procedure with a restricted form of flat automata.
@@
 
-We propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implement.
+We propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implementation.
 Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
@@
 Although the example is seemingly simple, most state-of-the-art solvers have difficulty to handle it, c.f., Section~\ref{sec-exp} for our experimental results.
-
 With several optimizations introduces in Section~\ref{sec-opt}, we manage to solve several variants of the hash function examples as well as some more randomly generated cases better than most state-of-the-art solvers.
-
-\yfc{todo: a short description of optimizations}

### 2021-05-20 FMCAD2021/paper.tex
opt section
update sec 4
preliminaries
flattening section
section 3
normalization
section 4
sec6 half done
intro add optimization part
section 3 and 4
abstract

@@
-% !TEX program = xelatex
+% !TEX program = pdflatex
 %
@@
 \usepackage{tikz}
+\usepackage{multirow}
+
 \usetikzlibrary{automata, positioning, arrows}
@@
 \RequirePackage{todonotes}
-\newcommand{\znjSide}[1]{\todo[color=orange!10]{\textbf{ZHAN:} #1}}
-\newcommand\znj[1]{\textcolor{red}{#1}}
 \newcommand{\parseInt}{\textsf{parseInt}}
-\newcommand{\paexp}{$\textsf{PA}_{\exp}$}
+\newcommand{\pa}{{\sf PA}}
+\newcommand{\qfpa}{{\sf QFPA}}
+\newcommand{\paexp}{{\sf ExpPA}}
 \newcommand{\Def}{\hat{=}}
+\newcommand{\strint}{{\sf STR}_\parseInt}
 
+\newcommand{\Nat}{\mathbb{N}}
+\newcommand{\Int}{\mathbb{Z}}
+\newcommand\Aut{\mathcal{A}}
+\newcommand\Lang{\mathcal{L}}
+\newcommand\free{\mathrm{Free}}
+\newcommand\op{\odot}
+\newcommand\parikh{\mathbb{P}}
+\newcommand{\hide}[1]{}
+\newcommand\encode{\mathit{encode}}
+\newcommand\decode{\mathit{decode}}
+\newcommand\len{\textsf{len}}
+\newcommand\svars{\mathcal{X}}
+\newcommand\ivars{\mathbbm{X}}
+\newcommand\pos{\mathit{pos}}
+\newcommand\ltrue{\mathtt{true}}
+\newcommand\lfalse{\mathtt{false}}
+\newcommand\flatdom{\mathcal{R}}
+\newcommand\flatten{\mathit{flatten}}
+\newcommand\svarx{x}
+\newcommand\svary{y}
+\newcommand\ivarx{\mathbbm{x}}
+\newcommand\ivary{\mathbbm{y}}
+\newcommand\ivarz{\mathbbm{z}}
+%\newcommand\notsubstr{\tt{NotSubstr}}
+\newcommand\concat{\cdot}
+\newcommand\sterm{t}
+\newcommand\iterm{\mathbbm{t}}
+\newcommand{\kell}{{(k,\ell)}}
+
+\newcommand{\svar}{\textsf{SVar}}
+\newcommand{\ivar}{\textsf{IVar}}
+\newcommand{\pvar}{\textsf{PVar}}
+\newcommand{\flatfun}{\mathcal{F}}
+
+\usepackage{bbm}
 \usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
+\usepackage{stmaryrd}
+
+\newtheorem{theorem}{Theorem}
 \newtheorem{lemma}[theorem]{Lemma}
@@
 % \newcommand{\ainput}{Input}
+
+\newcommand{\znj}[1]{\color{red} {NJ: #1 :JN} \color{black}}
+\newcommand{\znjSide}[1]{\todo[color=orange!10]{\textbf{ZHAN:} #1}}
+\newcommand{\zhilin}[1]{\color{brown} {ZL: #1 :LZ} \color{black}}
+\newcommand{\wuhao}[1]{\color{cyan} {WH: #1 :HW} \color{black}}
+\newcommand{\yfc}[1]{\color{violet} {YF: #1 :FY} \color{black}}
 
@@
     Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
-    In the PLDI 2020 paper of Abdulla et al., the authors extended the flattening approach to support the string-integer conversion, which is an important function appearing in almost all scripting languages.
+    As string-integer conversion is an important function appearing in almost all scripting languages, Abdulla et al. extended the flattening approach to this function recently.
     However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
     In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion.
-The approach is built upon a decision procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-While the decision procedure by Point relies on quantifier eliminations and is expensive, we introduce various optimizations and provide an efficient implementation.
+The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
+While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide an efficient implementation.
 %    Besides the theoretical results, we also develop various optimizations and provide an implementation.
-    We evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as randomly.
-    The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer conversion constraints as well as linear-exponential constraints, in both precision and efficiency.
+    We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.
+    The experimental results show that our implementation outperforms the state-of-the-art solvers. 
+    %for string-integer conversion as well as linear-exponential constraints.
+    %, in both precision and efficiency.     
 %\keywords{ String-integer conversion \and Flat regular constraints \and Presburger arithmetic \and Exponential function \and Quantifier elimination.}
@@
 
-\section{Decision procedure for $(\mathbb{N},+,2^x)$}
+\section{Decision procedure for {\paexp}}\label{sec-dec}
 
@@
 
-\section{Optimizations}
+\section{Optimizations}\label{sec-opt}
 
@@
 
-\section{Implementation and Experiments}
+\section{Implementation and Experiments}\label{sec-exp}
 
@@
 
+\newpage
+
 %test bib
 \bibliographystyle{plain}
+
 \bibliography{reference}
 
+\newpage
+
+\begin{appendix}
+
+\input{8_appendix.tex}
+
+\end{appendix}
+
 \end{document}

### 2021-05-20 FMCAD2021/1_intro.tex

@@
 
-Due to the difficulties in solving string constraints and, in practice, satisfiable string constraints are more critical for automatic testing, one idea is to have separate specialized procedures for solving satisfiable instances.
+Due to the difficulties in solving string constraints and, in practice, satisfiable string constraints are more critical for automatic testing, one idea is to have separate specialized procedures for solving satisfiable sub-problems.
 Currently, there are two main specialized approaches for proving satisfiability.
@@
 	all variables according to the order and then the resulted linear part, that improve the efficiency very much.
+
+Moreover, we propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implementation.
+Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
+We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.
+
 \item[iii)] Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} (see later).
 \end{itemize}
-
-We propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implementation.
-Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
-We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.
 

### 2021-05-20 FMCAD2021/2_preliminary.tex

@@
 where $a^i_j \in \Sigma$ for each $i \in [k]$ and $j \in [\ell_i]$.
-Intuitively, an FL is a set of strings consisting of $k$ consecutive parts, each created by iterating a \emph{cycle}, a string $a^i_1\ldots a^i_{\ell_i}$ of $\ell_i$ letters. 
+Intuitively, an FL is a  sequence of loops,  and the body of each loop is a string of the form  $a^i_1\ldots a^i_{\ell_i}$ with $\ell_i$ letters. 
 %A $\langle \ell, k \rangle$-FL is an FL such that its \emph{period} is at most $\ell$ and the \emph{cycle count} is at most $k$.

### 2021-05-20 FMCAD2021/1_intro.tex
abstract+intro

@@
+%!TEX root = paper.tex
+
 The emerging of scripting languages boosted the needs of better approaches and tools to ensure program quality.
@@
 However, the motivation is no longer valid since Makanin finds a decision procedure.
-Interestingly, in the last decade, driven by the need for program analysis, people start to revisit the problem and its extensions to describe the complete string library APIs in conventional programming languages.
+Interestingly, in the last decade, driven by the need for program analysis, people started to revisit the problem and its extensions to describe the complete string library APIs in conventional programming languages.
 
-The satisfiability of \emph{string/number conversion constraints}, e.g., $wx=yx \wedge |x| > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
-However, this kind of constraint is pervasive in scripting language programs. 
-For example, it is common that programs read string input from text files and converts a part of the string input to numbers.
-Even more crucially, in many programming languages, the string-number conversion is a part of the definition of their core semantics. 
+The satisfiability of \emph{string-number conversion constraints}, e.g., $wx=yx \wedge |x| > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
+However, this kind of constraints is pervasive in scripting language programs. 
+For example, it is common that programs read string inputs from text files and converts a part of the string input to numbers.
+Even more crucially, in many programming languages, the string-number conversion is a part of the definition of their core semantics \zhilin{reference here ?}. 
 JavaScript, which powers most interactive content on the Web and increasingly server-side code with Node.js, is one of such languages.
@@
 The first is to consider only strings of bounded length.
-This approach is taken in earlier generation solvers such as Hampi~\cite{KiezunGAGHE12} and Kuluza~\cite{SaxenaAHMMS10}.
+This approach is taken in the first-generation solvers such as Hampi~\cite{KiezunGAGHE12} and Kuluza~\cite{SaxenaAHMMS10}.
 Although they are useful in handling many practical cases, they fail short when the all string solutions exceed the selected bound.
@@
 One more recent approach is flattening~\cite{Parosh:20:PLDI,AbdullaACDHRR18,AbdullaACDHRR17}.
-The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat automata} (see Section~\ref{sec:string-solving} for more details). 
+The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat languages} (see Section~\ref{sec:string-solving} for more details). 
 The major benefit of considering this class is two folds.
@@
 For instance, we can find a solution for $x.y \neq z  \wedge |x| > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is a character in our alphabet.
-Second, and more importantly, because we can convert the membership problem of a flat automaton to a satisfiability problem of Presburgher formula, the class of word equation + regular constraints + length constraints is decidable.
+Second, and more importantly, because we can convert the membership problem of a flat language to the satisfiability problem of a Presburgher formula, the class of word equation + flat languages + length constraints is decidable.
 
-The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string/number conversion constraints to the above class.
-The paper proposes a decision procedure with a restricted form of flat automata.
-For string/number conversion, their approach projects the solution to string/number conversion to, in terms of numerical solutions, a finite solution space.
+The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string-number conversion constraints to the above class.
+The paper proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
+For string-number conversion, their approach projects the solution to string-number conversion to
+%, in terms of numerical solutions, 
+a finite solution space.
 
 In this paper, we propose a complete solution to this problem.
-We propose a decision procedure for the class of word equation + regular constraints + length constraints + string/number conversion under arbitrary flat automata (denoted by $\strint$).
-The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem of $\strint$ to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$); then, according to the decidability of $\paexp$ \cite{Semenov84}, we therefore obtain the decidability of the satisfiability of $\strint$.
+We propose a decision procedure for the class of word equation + flat languages + length constraints + string-number conversion.
+The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$); then, according to the decidability of $\paexp$, we obtain the decidability of the original satisfiability problem.
 
-In \cite{Semenov84}, Semen\"{o}v just proved that $\paexp$ admits quantifier elimination property and 
-therefore obtained that its satisfiability is decidable, but  did not provide an explicit decision procedure.
-So, in  \cite{Point86}, Point presented a first decision procedure for the satisfiability of $\paexp$ in a mathematical way, which is difficult to understand and therefore no implementation available based on which up to now to the best of our knowledge.
-So, another contribution of this paper is that we present a more understandable algorithm for $\paexp$'s satisfiability in a computational way and its prototypical implementation.
+The decidability of $\paexp$ was first shown by Semen\"{o}v in \cite{Semenov84}.
+Nevertheless, Semen\"{o}v did not provide an explicit decision procedure.
+To remedy this, in \cite{Point86}, Point presented the first quantifier elimination procedure for the satisfiability of $\paexp$.
+Partially attributed to the fact that Semen\"{o}v decidability proof in \cite{Semenov84} was quite abstract and Point's procedure in \cite{Point86} was presented  in a mathematical and dense way, this line of work has eluded the attentions of computer science researchers, to the best of our knowledge.
+Therefore, no implementation based on Point's procedure has been available up to now.
+
+Aiming at introducing this line of work to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}).
+This, by our humble opinion, can be seen as another contribution of this paper.
+%
+Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).
+
+%
+%The optimizations include 
+%over approximation to reduce the number of subcases
+% the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
+We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 Although the general framework of our algorithm is quite akin to Point's,  it contains the following improvements and corrections: 
@@
 	all variables according to the order and then the resulted linear part, that improve the efficiency very much.
-
-Moreover, we propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implementation.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
+%Moreover, we propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implementation. 
 Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
@@
 \end{itemize}
+ }
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-In fact, other than the theoretical difficulties, in practice, the string-number constraints are quite challenging for state-of-the-art solvers. 
+In fact, other than the theoretical difficulties, in practice, the string-number conversion is quite challenging for state-of-the-art solvers. 
 Here we illustrate a toy example that mimic the ``mining'' step of block-chain construction. 
-Essentially, giving a hash function ${\sf hash}(w)$, the goal of the mining step is to find an input $w$ such that ${\sf hash}(w)$ satisfies certain pattern, e.g., the last $k$ digits are zeros. 
+Essentially, given a string hash function ${\sf hash}(w)$, the goal of the mining step is to find an input $w$ such that ${\sf hash}(w)$ satisfies a certain pattern, e.g., the last $k$ digits are zeros.
 Below we consider a simple hash function:
@@
 In particular, if $\Sigma = \Sigma_{\sf num}$ and $p = 10$, then ${\sf hash}(w) = \parseInt(w) \bmod m$.
-Thus the problem of finding a suitable input $w$ such that the last $k$ digits of ${\sf hash}(w)$ are zeros can be modeled as a $\strint$ formula.
-Although the example is seemingly simple, most state-of-the-art solvers have difficulty to handle it, c.f., Section~\ref{sec-exp} for our experimental results.
-With several optimizations introduces in Section~\ref{sec-opt}, we manage to solve several variants of the hash function examples as well as some more randomly generated cases better than most state-of-the-art solvers.
+Thus, the problem of finding a suitable input $w$ such that the last $k$ digits of ${\sf hash}(w)$ are zeros can be modeled as a string constraint with $\parseInt$.
+Although the example is seemingly simple, it is already challenging for most state-of-the-art solvers, as shown by our experiment results in Section~\ref{sec-exp}.
+With the optimizations introduced in Section~\ref{sec-opt}, 
+our implementation manages to solve several variants of the string-hash examples as well as some randomly generated problem instances better than the state-of-the-art solvers.

### 2021-05-20 FMCAD2021/paper.tex
abstract+intro

@@
 The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide an efficient implementation.
+While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
 %    Besides the theoretical results, we also develop various optimizations and provide an implementation.

### 2021-05-21 FMCAD2021/1_intro.tex
intro

@@
 To remedy this, in \cite{Point86}, Point presented the first quantifier elimination procedure for the satisfiability of $\paexp$.
-Partially attributed to the fact that Semen\"{o}v decidability proof in \cite{Semenov84} was quite abstract and Point's procedure in \cite{Point86} was presented  in a mathematical and dense way, this line of work has eluded the attentions of computer science researchers, to the best of our knowledge.
-Therefore, no implementation based on Point's procedure has been available up to now.
+Partially attributed to the fact that Point's procedure in \cite{Point86} was presented  in a mathematical and dense way, this quantifier elimination procedure has mostly eluded the attentions of computer science researchers\footnote{There are only two citations in Google Scholar.}.
+To the best of our knowledge, no implementation based on Point's procedure has been available up to now.
 
-Aiming at introducing this line of work to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}).
-This, by our humble opinion, can be seen as another contribution of this paper.
+Aiming at introducing Point's procedure to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}).
+This can be seen as another contribution of this paper.
 %

### 2021-05-21 FMCAD2021/1_intro.tex
conclusion

@@
 With the optimizations introduced in Section~\ref{sec-opt}, 
-our implementation manages to solve several variants of the string-hash examples as well as some randomly generated problem instances better than the state-of-the-art solvers.
+our implementation manages to solve several variants of the string-hash examples as well as some randomly generated problem instances better than the state-of-the-art solvers (Section~\ref{sec-exp}).

### 2021-05-21 FMCAD2021/7_conclusion.tex
conclusion

@@
 
-Conclude this paper and discuss future work.
+In this paper, we proposed a complete flattening approach for string constraints with string-integer conversion and flat regular constraints, based on a quantifier elimination procedure by Point in 1986, for the extension of Presburger arithmetic with exponential functions.
+We gave a more accessible reformulation of Point's procedure and proposed various optimizations.
+Moreover, we achieved the first prototypical implementation of Point's procedure.
+We also did extensive experiments to evaluate the performance of the implementation.
+The experiment results show the efficacy of our implementation, compared with the state-of-the-art solvers.
+An interesting question left open in this paper is a more refined complexity analysis of the quantifier elimination procedure, considering the fact that quantifier elimination procedure for Presburger arithmetic, e.g.
+Cooper's algorithm, is triply exponential.

### 2021-05-21 FMCAD2021/7_conclusion.tex
conclusion

@@
 The experiment results show the efficacy of our implementation, compared with the state-of-the-art solvers.
-An interesting question left open in this paper is a more refined complexity analysis of the quantifier elimination procedure, considering the fact that quantifier elimination procedure for Presburger arithmetic, e.g.
-Cooper's algorithm, is triply exponential.
+An interesting question left open in this paper is a more refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.

### 2021-05-21 FMCAD2021/1_intro.tex
space adjustment

@@
 One more recent approach is flattening~\cite{Parosh:20:PLDI,AbdullaACDHRR18,AbdullaACDHRR17}.
-The idea is to restrict the solution space of string variables to some pattern defined by \emph{flat languages} (see Section~\ref{sec:string-solving} for more details). 
-The major benefit of considering this class is two folds.
-First, even under the restriction, often the potential solution space is still infinite, which gives us a higher potential of finding solutions.
-For instance, we can find a solution for $x.y \neq z  \wedge |x| > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is a character in our alphabet.
-Second, and more importantly, because we can convert the membership problem of a flat language to the satisfiability problem of a Presburgher formula, the class of word equation + flat languages + length constraints is decidable.
+The idea is to restrict the solution space of string variables to 
+%some pattern defined by 
+\emph{flat languages} (see Section~\ref{sec:string-solving}). 
+The major benefit of considering this class is two-fold.
+First, under the restriction, the potential solution space is still infinite, which gives us a higher potential of finding solutions.
+For instance, we can find a solution for $x.y \neq z  \wedge |x| > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is an element of the alphabet.
+Second, and more importantly, because we can convert the membership problem of a flat language to the satisfiability problem of a Presburger arithmetic formula, the class of word equation + flat languages + length constraints is decidable.
 
@@
 Below we consider a simple hash function:
-$${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m,$$ 
+${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m,$
 where $p, m \in \Int^+$ with $p, m \ge 2$. 

### 2021-05-21 FMCAD2021/3_flatten.tex
some grammar mistakes
space adjustment

@@
 
+\vspace*{-3mm}
 \subsection{String constraints with string-integer conversion ($\strint$)}
@@
 
-In $\strint$, there are two types of variables, namely, the string variables $\svarx,\svary,\ldots \in \svars$ and the integer variables $\ivarx,\ivary,\ldots \in \ivars$.
+In $\strint$, there are two types of variables, i.e. the string variables $\svarx,\svary,\ldots \in \svars$ and the integer variables $\ivarx,\ivary,\ldots \in \ivars$.
 %
@@
 
+\vspace*{-2mm}
 \subsection{An Extension of Presburger Arithmetic with Exponential Functions (\paexp)}
@@
 \begin{itemize}
-\item The (partial) function $\ell_{10}(\ivarx)$ is inductively defined as follows: For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$. (Note that $\ell_{10}(0)$ is undefined.) 
+\item The (partial) function $\ell_{10}(\ivarx)$ is\hide{inductively} defined as follows: For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$. (Note that $\ell_{10}(0)$ is undefined.) 
 %
@@
 
+\vspace*{-2mm}
 \subsection{Flattening $\strint$ into $\paexp$}
 
-We first recall the flattening approach for string constraints in \cite{Parosh:20:PLDI}, then show how to extend it to deal with the {\parseInt} function.
+We first recall the flattening approach for string constraints in \cite{Parosh:20:PLDI}, then show how to extend it to deal with {\parseInt}.
 
@@
 $\len(\svarx)$) for a string variable $\svarx$.
-(Otherwise, we can always introduce a fresh string variable $\svarx'$ for $\parseInt(\sterm)$ or $\len(\sterm)$ and add the constraint $\svarx' = \sterm$.)
+(Otherwise, we can introduce a fresh variable $\svarx'$ for $\sterm$ in $\parseInt(\sterm)$ or $\len(\sterm)$ and add the constraint $\svarx' = \sterm$.)
 Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by replacing $\parseInt(\svarx)$ with $\flatten_{\flatfun_\varphi}(\parseInt(\svarx))$ and $\len(\svarx)$ with $\flatten_{\flatfun_\varphi}(\len(\svarx))$, where

### 2021-05-21 FMCAD2021/5_optimization.tex
space adjustment

@@
 
-\paragraph{Reduce the number of enumerated variable orders by over approximation}
+\vspace*{-3mm}
+\subsection{Reduce the number of enumerated variable orders by over approximation}
 
@@
 
-\paragraph{Avoid the elimination of linear occurrences of variables by under approximation}\label{para: opt under appro}
+\vspace*{-3mm}
+\subsection{Avoid the elimination of linear occurrences of variables by under approximation}\label{para: opt under appro}
 

### 2021-05-21 FMCAD2021/6_impl_exp.tex
experiments
space adjustment

@@
 
+\vspace*{-1mm}
 \subsection{Implementation}
@@
 
+\vspace*{-3mm}
 \subsection{Benchmarks}
@@
 %
-\item For non-flat instances, we use flat languages $(a)^* (b_{1} \ldots b_{k})$ to under approximate 
+\item For non-flat instances, we use flat languages $a^* (b_{1} \ldots b_{k})$ to under approximate 
 $\Sigma_{\sf num}^*$,  where $a, b_1,\ldots, b_k \in \Sigma_{\sf num}$. 
@@
 
+\vspace*{-2mm}
 \subsection{Experiments}
@@
 The $\paexp$-solver reports ``B-UNSAT'' for 47 instances of the $(2, 3, 3, 4)$-group, while it does not report ``B-UNSAT'' (except one) for the other two groups.
-If more time is allowed, the $\paexp$-solver can report ``B-UNSAT'' for the ``TIMEOUT'' instances.
+If more time is allowed, the $\paexp$-solver is able to report ``B-UNSAT'' for the ``TIMEOUT'' instances.
 
@@
 % Merge bound and unbound 
+\vspace*{-3mm}
 \begin{table}[ht]

### 2021-05-21 FMCAD2021/paper.tex
space adjustment

@@
 \IEEEauthorblockA{
-\IEEEauthorrefmark{1} State Key Laboratory of Computer Science,\\
-Institute of Software, Chinese Academy of Sciences, 100190, Beijing, China\\
+\IEEEauthorrefmark{1} State Key Laboratory of Computer Science,
+Institute of Software, Chinese Academy of Sciences, Beijing, China\\
 Email: \{wuhao, wuzl, znj\}@ios.ac.cn}\\
@@
 %\keywords{ String-integer conversion \and Flat regular constraints \and Presburger arithmetic \and Exponential function \and Quantifier elimination.}
+%
+\end{abstract}
 
-\end{abstract}
+\vspace*{-2mm}
 
@@
 
+\vspace*{-2mm}
 \section{Preliminaries} \label{sec:pre}
@@
 
+\vspace*{-2mm}
 \section{Flattening string constraints with string-integer conversion}\label{sec:string-solving}
@@
 
+\vspace*{-4mm}
 \section{Decision procedure for {\paexp}}\label{sec-dec}
@@
 
+\vspace*{-2mm}
 \section{Optimizations}\label{sec-opt}
@@
 
+\vspace*{-3mm}
 \section{Implementation and Experiments}\label{sec-exp}
@@
 
+\vspace*{-3mm}
 \section{Conclusion}

### 2021-05-21 FMCAD2021/1_intro.tex
polish intro

--- 2021-05-21 FMCAD2021/1_intro.tex
@@
 
-However, research progress of string constraint solving has been hampered by many major obstacles in both theory and tool implementation aspects (including long-standing open problems). 
+However, research progress of string constraint solving has been hampered by many major challenges in both theory and tool implementation aspects (including long-standing open problems). 
 Logical theories over strings have to allow string concatenation, which is arguably the most fundamental operation of strings. 
@@
 
-The satisfiability of \emph{string-number conversion constraints}, e.g., $wx=yx \wedge |x| > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
+The satisfiability of \emph{string-integer conversion constraints}, e.g., $wx=yx \wedge |x| > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
 However, this kind of constraints is pervasive in scripting language programs. 
-For example, it is common that programs read string inputs from text files and converts a part of the string input to numbers.
-Even more crucially, in many programming languages, the string-number conversion is a part of the definition of their core semantics \zhilin{reference here ?}. 
+For example, it is common that programs read string inputs from text files and converts a part of the string input to integers.
+Even more crucially, in many programming languages, the string-integer conversion is a part of the definition of their core semantics ~\cite{ecmascript2019ecmascript}. 
 JavaScript, which powers most interactive content on the Web and increasingly server-side code with Node.js, is one of such languages.
@@
 
-The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string-number conversion constraints to the above class.
+The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string-integer conversion constraints to the above class.
 The paper proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
-For string-number conversion, their approach projects the solution to string-number conversion to
+For string-integer conversion, their approach projects the solution to string-integer conversion to
 %, in terms of numerical solutions, 
@@
 In this paper, we propose a complete solution to this problem.
-We propose a decision procedure for the class of word equation + flat languages + length constraints + string-number conversion.
+We propose a decision procedure for the class of word equation + flat languages + length constraints + string-integer conversion.
 The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$); then, according to the decidability of $\paexp$, we obtain the decidability of the original satisfiability problem.
@@
 
-In fact, other than the theoretical difficulties, in practice, the string-number conversion is quite challenging for state-of-the-art solvers. 
+In fact, other than the theoretical difficulties, in practice, the string-integer conversion is quite challenging for state-of-the-art solvers. 
 Here we illustrate a toy example that mimic the ``mining'' step of block-chain construction. 
@@
 our implementation manages to solve several variants of the string-hash examples as well as some randomly generated problem instances better than the state-of-the-art solvers (Section~\ref{sec-exp}).
+
+\paragraph*{Structure}
+After an introduction to background knowledge in Section~\ref{sec:pre}, we describe how to flatten a string constraint with string-integer conversion to an $\paexp$ formula in Section~\ref{sec:string-solving}.
+A decision procedure of $\paexp$ formulae is presented in Section~\ref{sec-dec}.
+We present optimizations in Section~\ref{sec-opt} and show the empirical evaluation results in Section~\ref{sec-exp}.

### 2021-05-21 FMCAD2021/4_dec_proc.tex
space adjustment
Update 4_dec_proc.tex

@@
 
-Furthermore, as Point's procedure is quite expensive and a faithful implementation would not scale\footnote{We did implemented Point's algorithm and discovered that the implementation could only solve formulas of very small size.},
+Furthermore, as Point's procedure is quite expensive and a faithful implementation would not scale\footnote{We did implement Point's algorithm and discovered that the implementation could only solve formulas of very small size.},
 in the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.

### 2021-05-21 FMCAD2021/1_intro.tex
update intro

@@
 However, the motivation is no longer valid since Makanin finds a decision procedure.
-Interestingly, in the last decade, driven by the need for program analysis, people started to revisit the problem and its extensions to describe the complete string library APIs in conventional programming languages.
+Recently, driven by the need for program analysis, people started to revisit the problem and its extensions to describe the complete string library APIs in conventional programming languages.
+Many highly efficient solvers for string constraints, to name a few, CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, Z3-Str3~\cite{zheng2013z3}, S3~\cite{trinh2014s3}, Norn~\cite{abdulla2015norn}, Ostrich~\cite{chen2017decidable}, Sloth~\cite{sloth}, ABC~\cite{aydin2018parameterized}, Stranger~\cite{yu2010stranger}, and Trau~\cite{abdulla2018trau}, are developed in the last decade.
 
@@
 For example, it is common that programs read string inputs from text files and converts a part of the string input to integers.
-Even more crucially, in many programming languages, the string-integer conversion is a part of the definition of their core semantics ~\cite{ecmascript2019ecmascript}. 
+Even more crucially, in many programming languages, the string-integer conversion is a part of the definition of their core semantics~\cite{ecmascript2019ecmascript}. 
 JavaScript, which powers most interactive content on the Web and increasingly server-side code with Node.js, is one of such languages.
@@
 %, in terms of numerical solutions, 
-a finite solution space.
+a finite solution space (in a way similar to the PASS~\cite{goudong2013pass} approach).
 

### 2021-05-21 FMCAD2021/1_intro.tex
explain better the hash problem

@@
 Here we illustrate a toy example that mimic the ``mining'' step of block-chain construction. 
-Essentially, given a string hash function ${\sf hash}(w)$, the goal of the mining step is to find an input $w$ such that ${\sf hash}(w)$ satisfies a certain pattern, e.g., the last $k$ digits are zeros.
+Essentially, given a string hash function ${\sf hash}(w)$, the goal of the mining step is to find a nonce $n$ such that when inserting $n$ to the text to be protected, say $w_1,w_2$, ${\sf hash}(w_1.n.w_2)$ satisfies a certain pattern, e.g., the last $k$ digits are zeros.
+If $w_1$ or $w_2$ are modified, one needs to compute another $n$ satisfies the desired pattern.
 Below we consider a simple hash function:

### 2021-05-21 FMCAD2021/6_impl_exp.tex
explain better the hash problem

--- 2021-05-20 FMCAD2021/6_impl_exp.tex
@@
 This suite comprises two groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
+For one of them, we restrict the nonce string conforming some flat pattern, while for the other one, we allow any word from $\Sigma^+_{\sf num}$ to be used as the nonce.
 %Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, we assume that $\Sigma =\Sigma_{\sf num}$ and $p = 10$ and the string hash function is encoded by $\parseInt$.
@@
 The two groups of string constraints are characterized by flat and non-flat regular constraints respectively.
-The flat group comprises 300 problem instances, where the flat languages are of the form $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, or $w^+_1 w^+_2 6789$, with $w_1,w_2 \in \Sigma^+_{\sf num}$.
+The flat group comprises 300 problem instances, where the flat languages are of the form $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, or $w^+_1 w^+_2 6789$, with $w_1,w_2 \in \Sigma^+_{\sf num}$, where $1234$ and $6789$ are the text to be protected, and $w^+_1 w^+_2$ is the pattern for nonce string.
 The non-flat group comprises 300 problem instances, where the non-flat languages are of the form $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, or $\Sigma_{\sf num}^* 6789$.

### 2021-05-21 FMCAD2021/3_flatten.tex
Update 3_flatten.tex

--- 2021-05-20 FMCAD2021/3_flatten.tex
@@
 
+By the above reduction, and the decidability of $\paexp$'s satisfiability (see Theorem~\ref{thm-paexp}), we have 
+\begin{theorem} \label{thm:string-parInt}
+	The satisfiability of  $\strint$ is decidable. 
+\end{theorem}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-21 FMCAD2021/4_dec_proc.tex
Update 4_dec_proc.tex

@@
 
-The main idea of the elimination of the exponential occurrences of $\ivarx_3$ is to observe that if $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$, then $10^{\ivarx_3} - 10^{\ivarx_2}$ is dominated by $10^{\ivarx_3}$, that is, $10^{\ivarx_3} - 10^{\ivarx_2} \ge 10^{\ivarx_3} - 10^{\ivarx_3 - 3} = (1-10^{-3}) 10^{\ivarx_3} \ge 10^{\ell_{10}(\ivary+1001)+1} = 10\lambda_{10}(\ivary+1001) > \ivary+1001$.
-(See Lemma~\ref{lem:exp-ineq} for the choice of the constants $2$ and $3$ in $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$.)
+The main idea of the elimination of the exponential occurrences of $\ivarx_3$ is to observe that if $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$, then $10^{\ivarx_3} - 10^{\ivarx_2}$ is dominated by $10^{\ivarx_3}$, that is, $10^{\ivarx_3} - 10^{\ivarx_2} \ge 10^{\ivarx_3} - 10^{\ivarx_3 - 3} = (1-10^{-3}) 10^{\ivarx_3} \ge 10^{\ell_{10}(\ivary+1001)+1} = 10\lambda_{10}(\ivary+1001) > \ivary+1001$ (see Lemma~\ref{lem:exp-ineq} for the choice of the constants $2$ and $3$ in $\ivarx_3 \ge \ell_{10}(\ivary+1001)+2$ and $\ivarx_3 \ge \ivarx_2 + 3$.).
 Therefore, a necessary condition for $10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001$ is that either $\ivarx_3 \le \ell_{10}(\ivary+1001)+1$ or $\ivarx_3 \le \ivarx_2 + 2$ holds.
@@
 Cooper's algorithm in \cite{Cooper72}.
-The elimination of $\ivarx_3$ in $\varphi_2$ here is simple, with $\ivarx_3$ replaced by $\ivarx_1 + \ivarx_2$, resulting into
+The elimination of $\ivarx_3$ in $\varphi_2$ here is simple, with $\ivarx_3$ replaced by $\ivarx_1 + \ivarx_2$.
+\hide{, resulting into  
 \[
@@
 \end{array}
-\]
+\]}
 

### 2021-05-21 FMCAD2021/1_intro.tex
intro

@@
 \paragraph*{Structure}
-After an introduction to background knowledge in Section~\ref{sec:pre}, we describe how to flatten a string constraint with string-integer conversion to an $\paexp$ formula in Section~\ref{sec:string-solving}.
-A decision procedure of $\paexp$ formulae is presented in Section~\ref{sec-dec}.
-We present optimizations in Section~\ref{sec-opt} and show the empirical evaluation results in Section~\ref{sec-exp}.
+After the preliminaries in Section~\ref{sec:pre}, we describe how to flatten a string constraint with string-integer conversion to an $\paexp$ formula in Section~\ref{sec:string-solving}.
+We present the quantifier elimination procedure for $\paexp$ in Section~\ref{sec-dec} and its optimizations in Section~\ref{sec-opt}.
+Finally, we describe the implementation and experiment results in Section~\ref{sec-exp}.

### 2021-05-21 FMCAD2021/4_dec_proc.tex
space adjustment

@@
 
+\vspace*{-3mm}
 \subsection{Normalization}
@@
 It also applies some additional transformations.
-After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\lambda_{10}(\iterm)$ as well as $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) moreover, all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\lambda_{10}(\iterm)$ as well as $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
 More details can be found in Appendix~\ref{app-norm}.
@@
 
+\vspace*{-3mm}
 \subsection{Elimination of  exponential occurrences of variables}\label{sec-elim-exp}
@@
 For convenience, let us call these formulas as inequality respectively divisibility atomic formulas.
-In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these the inequality atomic formulas.
+In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these inequality atomic formulas.
 The elimination of the exponential occurrences of the divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.

### 2021-05-21 FMCAD2021/5_optimization.tex
space adjustment

@@
 
-In this section, we focus on quantifier-free {\paexp} formulas (or existential {$\paexp$} formulas since they are satisfiability-equivalent), and present various optimizations of the decision procedure for {\paexp}, aiming at an efficient implementation.
-The focus on quantifier-free {\paexp} formulas is motivated by the following two facts: 1) the flattening of {$\strint$} string constraints results into such formulas, 2) these formulas are already challenging for the state-of-the-art SMT solvers e.g.
-CVC4 and Z3 (with exponential functions defined as recursive functions).
+In this section, we focus on quantifier-free {\paexp} formulas (or existential {\paexp} formulas since they are satisfiability-equivalent), and present various optimizations of the quantifier elimination procedure for {\paexp}, aiming at an efficient implementation.
+The focus on quantifier-free {\paexp} formulas is motivated by the following two facts: 1) the flattening of {$\strint$} constraints results into such formulas, 2) these formulas are already challenging for state-of-the-art SMT solvers (with exponential functions defined as recursive functions).
 
@@
 
-Recall that in the decision procedure for {\paexp}, after the normalization, the orders of variables are enumerated and for each order, the exponential as well as linear occurrences of variables are eliminated.
+Recall that in the decision procedure for {\paexp}, after the normalization, the variable orders are enumerated and for each order, the exponential and linear occurrences of variables are eliminated.
 Since the quantifier elimination is expensive and applied to each possible order of variables, if we could reduce the candidate variable orders in the very beginning, it would facilitate considerable speed-up for the decision procedure.
@@
 Recall that Lemma~\ref{lem:exp-ineq} enables us to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ from 
+\vspace{-1mm}
 $$
@@
 \hspace{4mm} 
+\vspace{-1mm}
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c.
@@
 %\begin{equation}\label{eqn-org}
+\vspace{-2mm}
 \[\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c - b_i  \ivarx_{\sigma(i)}.\]
+\vspace{-1mm}
 %\end{equation}
@@
 Then we can apply Lemma~\ref{lem:exp-ineq} to the following two inequalities to eliminate $10^{\ivarx_{\sigma(i-1)}}$,
+\vspace{-2mm}
 \begin{equation}\label{eqn-lb}
@@
 and
+\vspace{-2mm}
 \begin{equation}\label{eqn-ub}

### 2021-05-21 FMCAD2021/6_impl_exp.tex
space adjustment

@@
 If a constant upper bound $10^u$ is specified, then the problem is to decide whether there is an assignment of values no more than $10^u$ to variables satisfying the given {\paexp} formula.
-
+%
 The outputs of the {\paexp}-solver are either ``SAT'', ``UNSAT'', ``B-UNSAT'', or ``TIMEOUT'', corresponding to the facts that the given formula is satisfiable, unsatisfiable, unsatisfiable up to $10^u$, or the search goes beyond the predetermined time limit.
@@
 
-\vspace*{-3mm}
+\vspace*{-4mm}
 \subsection{Benchmarks}
 
-To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC, and STRINGHASH.
+To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC and STRINGHASH.
 %\begin{itemize}
@@
 The string constraints in the STRINGHASH benchmark suite are of the form 
-$$\svarx \in \Aut \wedge \left( \parseInt(\svarx) \bmod m \right) \bmod m^\prime = 0   \notag \wedge \len(\svarx) < 100,$$ 
+%
+$\svarx \in \Aut \wedge \left( \parseInt(\svarx) \bmod m \right) \bmod m^\prime = 0   \notag \wedge \len(\svarx) < 100,$ 
+%
 where $\Aut$ is an FA, $m, m^\prime \ge 2.$ 
@@
 
-\vspace*{-2mm}
+\vspace*{-4mm}
 \subsection{Experiments}

### 2021-05-22 FMCAD2021/4_dec_proc.tex
space adjustment
changes on lemma 6

@@
 
+\vspace{-1mm}
 \begin{lemma} \label{lem:exp-ineq}
@@
 $$
+\small
 \begin{array}{l}
@@
 \hspace{4mm} 
+\vspace{-1mm}
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
@@
 %$B\Def  \sum_{j=1}^{i}|b_j|$, 
-$B \Def 10(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
+$B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
 and $\delta\Def  \ell_{10}(A)+3$. 
 \begin{itemize}
-    \item If $a_i > 0$, let $\alpha(\vec{\ivary}) \Def \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$, then 
+    \item If $a_i > 0$, let $\alpha(\vec{\ivary}) \Def \ell_{10}(\max \{1 ,\iterm(\ivary)\})- \ell_{10}(a_i)$, then 
     \begin{itemize}
@@
     \end{itemize}
-    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-(a_i))$, then 
+    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(\max \{1 ,-\iterm(\ivary)\})- \ell_{10}(-(a_i))$, then 
     \begin{itemize}
@@
 
-If $a_i > 0$, then the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ can be eliminated by utilizing  Lemma~\ref{lem:exp-ineq} and enumerating the constraints on $\ivarx_{\sigma(i)}$ and $\ivarx_{\sigma(i-1)}$.
+If $a_i > 0$ and $\iterm(\ivary)>0$, then the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ can be eliminated by utilizing  Lemma~\ref{lem:exp-ineq} and enumerating the constraints on $\ivarx_{\sigma(i)}$ and $\ivarx_{\sigma(i-1)}$.
 Specifically,
 $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
+\vspace{-2mm}
 \[
@@
  \ivarx_{\sigma(i)} = \ivarx_{\sigma(i-1)} +p\ \wedge \\
+ \vspace{-1mm}
  \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\ivarx_{\sigma(i-1)}+ p /\ivarx_{\sigma(i)}]

### 2021-05-22 FMCAD2021/8_appendix.tex
expriment
experiments
a mistake
changes on lemma 6

@@
 %
-\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
+\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ or $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
+Then replace each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct.
 Let the resulting formula be $\varphi''$.
 %
-\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $\ell_{10}$ in $\varphi''$, say $\ell_{10}(\iterm)$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$,
-%and let $\varphi''':= \exists \ivarz.\ \varphi''[\ivarz/\ell_{10}(\iterm)] \wedge 10^\ivarz \le \iterm < 10^{\ivarz + 1}$. 
-and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct. Let $\varphi'''$ denote the resulting formula.  
-%
+
+\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $10^{\iterm}$ in $\varphi''$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $10^{\iterm}$ by $10^\ivarz$, moreover, add the constraint $\ivarz = \iterm$ as a conjunct.
+Let $\varphi'''$ denote the resulting formula.
+
 \item Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
@@
 \subsection{Proof of Lemma~\ref{lem:exp-ineq}}
+
+Note that $\ell_{10}(n)$ is undefined for $n\le 0$ in $\paexp$. For convenience, we define $\ell_{10}(n)=0$ for $n\le 0$ in the following proof.
 
@@
     \end{itemize}
-    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-(a_i))$, then 
+    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-a_i)$, then 
     \begin{itemize}
@@
 \begin{proposition} \label{prop:1}
-If $n\ge m\ge 1$ and $p \ge 10(\ell_{10}(n)- \ell_{10}(m)+1)$, then 
+If $n\ge m\ge 1$ and $p \ge 2(\ell_{10}(n)- \ell_{10}(m)+1)$, then 
 $n p \le m10^p$ holds.
@@
 \begin{proof}
-By an induction on $n'$, we first show that for any $n' \in \Nat$, if $p \ge 10n'$, then $10^{n'} p \le 10^p$.
+By an induction on $n'$, we first show that for any $n' \in \Nat$, if $p \ge 2n'$, then $10^{n'} p \le 10^p$.
 
-If $ p \ge 10 (\ell_{10}(n) - \ell_{10}(m)+1)$, then  $n p \le 10\lambda_{10}(n) p = 10 * 10^{\ell_{10}(n)} p = 10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p$.
+If $ p \ge 2 (\ell_{10}(n) - \ell_{10}(m)+1)$, then  $n p \le 10\lambda_{10}(n) p = 10 * 10^{\ell_{10}(n)} p = 10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p$.
 
-Because $p \ge 10(\ell_{10}(n)-\ell_{10}(m) + 1)$, we deduce that $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  p \le 10^p$.
+Because $p \ge 2(\ell_{10}(n)-\ell_{10}(m) + 1)$, we deduce that $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  p \le 10^p$.
 Therefore, $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p \le 10^{\ell_{10}(m)} 10^p \le m 10^p$.
@@
 
-Let $A\Def \sum_{j=1}^{i-1}|a_j|$, $B \Def 10(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$, and $\delta\Def  \ell_{10}(A)+3$.
+Let $A\Def \sum_{j=1}^{i-1}|a_j|$, $B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$, and $\delta\Def  \ell_{10}(A)+3$.
 
@@
 $n \ge m \ge 1$.
-From $10(\ell_{10}(n)- \ell_{10}(m)+1) = 10 (\ell_{10}(100 \sum_{j=1}^{i}|b_j|) + 1) = 10 (\ell_{10}(\sum_{j=1}^{i}|b_j|) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 10(\ell_{10}(n)- \ell_{10}(m)+1)$.
+From $2(\ell_{10}(n)- \ell_{10}(m)+1) = 2 (\ell_{10}(100 \sum_{j=1}^{i}|b_j|) + 1) = 2 (\ell_{10}(\sum_{j=1}^{i}|b_j|) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 2(\ell_{10}(n)- \ell_{10}(m)+1)$.
 Then according to Proposition~\ref{prop:1}, 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-\begin{align}
-    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
-    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
-     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
-    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
-    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
-    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
-      (\mbox{ Def. of } \alpha) & \notag \\
-     & \ge  t(\bar{y}) & \notag 
-\end{align}
+\end{proof}
 
-If $\ivarx_{\sigma(i)} \ge  \alpha(\vec{\ivary}) + 2$, then
-
-The goal is to find a bound for $x_i$ such that 
-the values of the atoms containing $x_i$ keep constant when $x_i$ is greater than the bound.
-
-When $x_i \ge \alpha+2$,
-\begin{align}
-    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
-    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
-     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
-    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
-    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
-    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
-      (\mbox{ Def. of } \alpha) & \notag \\
-     & \ge  t(\bar{y}) & \notag 
-\end{align}
-So we conclude that %when $x_i \ge \alpha+2,x_i\ge B',x_i \ge x_{i-1}+\delta$,
-$\tau(x_i,\bar{x},\bar{y})$ keeps \textit{false} in this case.
-
-%\znj{I think there are some mistakes in the above proof, pls check it!}
-When $x_i \le \alpha-1$, similarly we have
-\begin{align}
-    \tau(x_i,\bar{x},\bar{y})\notag 
-    & \le  a_i 2^{x_i} + 2^{x_i -\delta}A  + Bx_i &  \notag \\
-    & \le  2^{x_i}(a_i +\frac{1}{4} +\frac{1}{4}) &(\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}}) \notag \\
-   & \le  2\lambda_2(a_i)\frac{\lambda_2(t(\bar{y}))}{2\lambda_2(a_i)} &  (\mbox{ Def. of } \alpha) \notag \\
-   &  \le  t(\bar{y}) & \notag 
-\end{align}
-which indicates that when  $x_i \le \alpha-1$, $\tau(x_i,\bar{x},\bar{y})$ keeps \textit{true}. \qed 
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{proof}
+We would like to remark that Lemma~\ref{lem:exp-ineq} still holds when the base of exponential function is changed to any natural number $n\ge 2$.
 
@@
 If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.
+
+\subsection{Detailed experiment results on the STRINGHASH benchmark suite}\label{app-exp}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table}[ht]
+    \caption{Experimental results on STRINGHASH,  O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
+    $\#$: number of problems, $T$: average time in seconds}
+    \centering
+
+    \renewcommand{\arraystretch}{1.1}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+    \hline
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
+        \cline{3-10}
+       &  & $\#$ & $T$ &  $\#$ & $T$ & $\#$ & $T$ &  $\#$ & $T$ \\ 
+       \hline
+       \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345(w_1)^+(w_2)^+$}} & S & 5 & 14.0 & 29 & 8.5 & 3 & {\bf  0.1} & {\bf 37} & 9.9 \\
+         \cline{2-10}
+        & U & 0 & - & 0 & - & {\bf 60} & {\bf 1.3} & {\bf 60} & 47.2 \\
+         \cline{2-10}
+        & F & 95 & - & 71 & - & 37 & - & {\bf 3} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$\begin{array}{l}12345(w_1)^+ \\
+       \hspace{2mm}(w_2)^+6789\end{array}$}} & S & 11 & 13.0 & 29 & 12.0 & 0 & - & {\bf 37} & {\bf 10.6} \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & {\bf 63} & {\bf 1.2} & {\bf 63} & 50.0 \\
+         \cline{2-10}
+        & F & 89 & - & 71 & - & 37 & - & {\bf 0} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$(w_1)^+(w_2)^+6789$}} & S & 18 & 24.0 & 30 & 9.3 & 2 & {\bf  0.1} & {\bf 41} & 16.1 \\
+      	 \cline{2-10}
+        & U & 0 & - & 1 & 4.0 & {\bf 59} & {\bf 2.5} & {\bf 59} & 45.8 \\
+         \cline{2-10}
+        & F & 82 &  & 69 &  & 39 &  & {\bf 0} &  \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}$}} & S & 82 & 8.7 & {\bf 100} & {\bf 2.2} & 28 & 5.9 & {\bf 100} & 18.5 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 18 & - & {\bf 0} & - & 72 & - & {\bf 0} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}6789$}} & S & 60 & 9.3 & 17 & 7.8 & 3 & {\bf 0.3} & {\bf 100} & 16.0 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 40 & - & 83 & - & 97 & - & {\bf 0} & - \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$\Sigma^*_{\sf num}6789$}} & S & 68 & 5.5 & 27 & 13.0 & 24 & {\bf 9.0} & {\bf 100} & 15.7 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 32 & - & 73 & - & 76 & - & {\bf 0} &- \\
+        \hline
+       \end{tabular}
+           \label{table:string}
+\end{table}

### 2021-05-22 FMCAD2021/3_flatten.tex
intro
space adjustment
vairables in paexp are non-neg Interger

--- 2021-05-21 FMCAD2021/3_flatten.tex
@@
 A $\strint$ formula $\varphi$ is defined by the following rules, where $\len(\sterm)$ denotes the length of a string $\sterm$.
+\vspace{-2mm}
 \[
@@
 \iterm & \Def & n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,\\
+\vspace{-2mm}
 \varphi & \Def & \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi,
@@
 
-\vspace*{-2mm}
+\vspace*{-5mm}
 \subsection{An Extension of Presburger Arithmetic with Exponential Functions (\paexp)}
@@
 %
+\vspace{-1mm}
  $$
@@
  \iterm &\Def& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \mid \lambda_{10}(\iterm), \\
+\vspace{-1mm}
  \phi &\Def & \iterm \ \op\ \iterm \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi.
@@
  $$
- The semantics of {\paexp} are defined similarly to {\pa}. The quantifier-free as well as existential {\paexp} formulas are defined similarly to {\pa} as well.
+ The semantics of {\paexp} are defined similarly to {\pa} with the difference that variables are interpreted over $\Nat$, that is, an interpretation is a function $I: \free(\phi) \rightarrow \Nat$.
+The quantifier-free as well as existential {\paexp} formulas are defined similarly to {\pa} as well.
 
@@
 %
-$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)})$$
+$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)}).$$
 %
@@
 
+By the above reduction, and the decidability of $\paexp$'s satisfiability (see Theorem~\ref{thm-paexp}), we have 
+\begin{theorem} \label{thm:string-parInt}
+	The satisfiability of  $\strint$ under flat domain restrictions  is decidable. 
+\end{theorem}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-22 FMCAD2021/6_impl_exp.tex
small changes

@@
 This suite comprises two groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
-For one of them, we restrict the nonce string conforming some flat pattern, while for the other one, we allow any word from $\Sigma^+_{\sf num}$ to be used as the nonce.
+For one of them, we restrict the nonce string conforming some flat pattern, while for the other one, we allow any word from $\Sigma^*_{\sf num}$ to be used as the nonce.
 %Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, we assume that $\Sigma =\Sigma_{\sf num}$ and $p = 10$ and the string hash function is encoded by $\parseInt$.
@@
 The two groups of string constraints are characterized by flat and non-flat regular constraints respectively.
-The flat group comprises 300 problem instances, where the flat languages are of the form $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, or $w^+_1 w^+_2 6789$, with $w_1,w_2 \in \Sigma^+_{\sf num}$, where $1234$ and $6789$ are the text to be protected, and $w^+_1 w^+_2$ is the pattern for nonce string.
+The flat group comprises 300 problem instances, where the flat languages are of the form $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, or $w^+_1 w^+_2 6789$, with $w_1,w_2 \in \Sigma^+_{\sf num}$, where $12345$ and $6789$ are the text to be protected, and $w^+_1 w^+_2$ is the pattern for nonce string.
 The non-flat group comprises 300 problem instances, where the non-flat languages are of the form $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, or $\Sigma_{\sf num}^* 6789$.

### 2021-05-22 FMCAD2021/8_appendix.tex
small changes
prop7 proof

@@
 
-Note that $\ell_{10}(n)$ is undefined for $n\le 0$ in $\paexp$. For convenience, we define $\ell_{10}(n)=0$ for $n\le 0$ in the following proof.
+Note that $\ell_{10}(n)$ is undefined for $n\le 0$ in $\paexp$. For convenience, we define for all $n\le 0$ that $\ell_{10}(n)=\ell_{10}(\max\{1,n\})=0$ in the following proof.
 
@@
 \begin{proof}
-By an induction on $n'$, we first show that for any $n' \in \Nat$, if $p \ge 2n'$, then $10^{n'} p \le 10^p$.
+
+First we show that for any $n' \in \Nat$, if $p \ge 2n'$, then $10^p \ge 10^{n'}10^{(p-n')}\ge 10^{n'}10(p-n')\ge 10^{n'}(5p+5(p-2n'))\ge  10^{n'}p$.
 

### 2021-05-22 FMCAD2021/7_conclusion.tex
add link for benchmarks

@@
 The experiment results show the efficacy of our implementation, compared with the state-of-the-art solvers.
-An interesting question left open in this paper is a more refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.
+An interesting open question is a refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.

### 2021-05-23 FMCAD2021/7_conclusion.tex
complexity

@@
 The experiment results show the efficacy of our implementation, compared with the state-of-the-art solvers.
-An interesting open question is a refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.
+\wuhao{An interesting question left open in this paper is a more refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.
+}

### 2021-05-23 FMCAD2021/8_appendix.tex
complexity

@@
 \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-c\ \big |\ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+d\ \big |\ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
 + \iterm(\vec{\ivary}) \big)
@@
 
-Let $c = 2^{r_1} 5^{r_2}  c_0$ such that $c_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $c | (10^rc_0)$.
+Let $d = 2^{r_1} 5^{r_2}  d_0$ such that $d_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $d | (10^rd_0)$.
 
-If $c_0 = 1$, then $10^r$ is divisible by $c = 2^{r_1}5^{r_2}$.
-Thus for every $n \ge r$, $c \ |\ 10^n$.
+If $d_0 = 1$, then $10^r$ is divisible by $d = 2^{r_1}5^{r_2}$.
+Thus for every $n \ge r$, $d \ |\ 10^n$.
 Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to
@@
 \begin{array}{l}
-\bigvee \limits_{p = 0}^{r - 1} c\ \big | \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
+\bigvee \limits_{p = 0}^{r - 1} d\ \big | \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
 + \iterm(\vec{\ivary}) \big)\\
 %
-\vee \big(\ivarx_{\sigma(i)} \ge r \wedge c\ \big | \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+\vee \big(\ivarx_{\sigma(i)} \ge r \wedge d\ \big | \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
 + \iterm(\vec{\ivary}) \big)\big),
@@
 
-Next, let us assume $c_0 > 1$.
-Since $10$ and $c_0$ are relatively prime, according to Euler's theorem (cf.
-\cite{HW80}), $10^{\phi(c_0)} \equiv 1 \bmod c_0$, where $\phi$ is the Euler function.
-Suppose $10^{\phi(c_0)} = kc_0 + 1$ for some $k \in \Nat$.
+Next, let us assume $d_0 > 1$.
+Since $10$ and $d_0$ are relatively prime, according to Euler's theorem (cf.
+\cite{HW80}), $10^{\phi(d_0)} \equiv 1 \bmod d_0$, where $\phi$ is the Euler function.
+Suppose $10^{\phi(d_0)} = kd_0 + 1$ for some $k \in \Nat$.
 Then for every $n \in \Nat$ with $n \ge r$, 
@@
 \begin{array}{l}
-10^{n + \phi(c_0)} \bmod c =10^{n-r} 10^r (kc_0 + 1) \bmod c = \\
-10^{n-r} (k 10^rc_0 + 10^r) \bmod c = \\
-10^{n-r} (0+10^r) \bmod c = 10^n \bmod c.
+10^{n + \phi(d_0)} \bmod d =10^{n-r} 10^r (kd_0 + 1) \bmod d = \\
+10^{n-r} (k 10^rd_0 + 10^r) \bmod d = \\
+10^{n-r} (0+10^r) \bmod d = 10^n \bmod d.
 \end{array}
@@
 \ivarx_{\sigma(i)} \ge r\ \wedge \\
-\bigvee \limits_{q = 0}^{\phi(c_0)-1} 
+\bigvee \limits_{q = 0}^{\phi(d_0)-1} 
 \left(
 \begin{array}{l}
-\phi(c_0)\ \big |\ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
-c\ \big | 
+\phi(d_0)\ \big |\ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
+d\ \big | 
 \left(
@@
 where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+
+\subsection{Complexity analysis}
+
+In this part, we will analysis the complexity of the decision procedure described in Section~\ref{sec-dec}.
+For convenience, we will slightly change the algorithm: the normalization procedure is invoked for all quantified variables in the beginning, which means step 2 and step 3 in the normalization procedure are performed for all quantified variables instead of a specified $\ivarx$.
+
+Let the formula after the normalization step be $\varphi = \exists \vec{\ivarx}. \theta$ with $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$. Let $n$ denote the length of formula $\varphi$.
+
+Then we eliminate the exponential occurrences and linear occurrences of variables one by one according to a specified linear order of $\vec{\ivarx}$.
+Suppose the linear order is $\ivarx_1\ge \dots \ge \ivarx_m$.
+
+We adopt the method in the analysis of the complexity of \pa \ (cf.
+\cite{Oppen73}).
+Let $\varphi'_k = \exists \ivarx_m \dots  \exists \ivarx_{k}\theta'_k$ denote the formula after $k-1$ variables and exponential occurrences of $\ivarx_k$ are eliminated from $\varphi$.
+Let $\varphi_k$ denote the formula obtained by eliminating linear occurrences of $\ivarx_k$ from $\varphi'_k$ and $\varphi_0=\varphi$.
+
+Let $c_k$ be the number of distinct $d$ in atoms of the form $d \ \big | \ \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of variables in $\varphi_k$.
+Let $s_k$ be the largest constant (including coefficients) and $a_k$ be the number of atomic formulas in $\varphi_k$.
+Similarly we define $c'_k$, $s'_k$ and $a'_k$ for $\varphi'_k$.
+And let $c,s,a$ be $c_0,s_0,a_0$ respectively.
+We will prove the following lemma.
+\begin{lemma}\label{lem:cpx exp}
+    \begin{align}
+        c'_1&\le c^2 \notag \\
+        s'_1&\le ms^2 \notag\\
+        a'_1&\le sa\notag 
+    \end{align} 
+\end{lemma}
+
+\begin{proof}
+We analyze the complexity in the elimination of exponential occurrences of variable $\ivarx_1$. It will be discussed into 2 cases.
+
+If all atoms are inequalities atomic formulas of the form in Lemma~\ref{lem:exp-ineq}.
+We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_1$ is replaced by a new formula.
+Only the coefficients of linear occurrences of $\ivarx_1$ and $\ivarx_2$ will be changed: constant coefficient $1$ is introduced, and if we substitute $x_1$ by $x_2+p$ for some $p$, coefficient of linear occurrence of $x_2$ will become $b_1+b_2$ ($b_1,b_2$ are coefficients for $x_1$ and $x_2$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
+Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le c^2$.
+Note that $\delta$ and $B$ are at most $\ell_{10}(ms)$, when we substitute $\ivarx_1$ by $\ivarx_2+\delta-1$ or by $B-1$, the largest constant in the formula is at most $s\cdot  10^{\ell_{10}(ms)}\le ms^2$.
+And an inequality is replaced by at most $4\ell_{10}(ms)$ atomic formulas, so $a'_1\le 4\ell_{10}(ms)a$.
+
+If all atoms are divisiblilty atomic formulas of the form $d \ \big | \ \iterm$.
+It is easy to see $c'_1<2c$ because a divisibility atomic formula  will produce at most two form of atomic formulas $d \ \big | \ \iterm$ and $\phi(d) \ \big | \ \iterm$.
+Note that any constant in $\iterm$ , say $l$, can be replaced by ($l \bmod d)$, so $s'_1\le s$.
+When $d$ is a large prime number, $\phi(d)=d-1$, a divisibility atomic formula is replaced by roughly $d$ atomic formulas, so $a'_1\le sa$
+
+Choose larger upper bound for $c'_1$, $s'_1$ and $a'_1$ respectively then the lemma is proved.
+
+\end{proof}
+
+Since linear occurrences of $\ivarx_1$ are eliminated using Cooper's algorithm, we combine Oppen's analysis and obtain:
+\begin{align}
+    c_1&\le c^8 \notag \\
+    s_1&\le (ms^2)^{4c^2} \notag\\
+    a_1&\le (sa)^4(ms^2)^{2c^2}\notag 
+\end{align}
+
+Assume $m\le n$, by induction on $k$ we can prove the following lemma.
+
+\begin{lemma}
+    \begin{align}
+        c_k&\le c^{8^k} \notag \\
+        s_k&\le n^{(4c)^{ 8^k}} s^{(8c)^{ 8^k}} \notag\\
+        a_k&\le a^{4^k}n^{(4c)^{ 8^k}} s^{(8c)^{ 8^k}} \notag
+    \end{align} 
+\end{lemma}
+
+Similarly to \cite{Oppen73}, we can assume $c\le n$, $a\le n$ and $s\le n$.So the space required to store $\varphi_k$ is the product of number of atoms $a_k$, the maximum number of constants $2m+2$ per atom, the maximum amount of space $s_k$ to store each constant and some constant $q$:
+
+$${\sf space}\le q \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
+for some large constant $p$. It is also a bound for deterministic time.
+Since for arbitrary $\paexp$ formula length $n'$, the normalization for all variables results into a formula at most $n'^3$, so the bound is still 3-EXPTIME for $n'$.
 

### 2021-05-23 FMCAD2021/7_conclusion.tex

--- 2021-05-22 FMCAD2021/7_conclusion.tex
@@
 The experiment results show the efficacy of our implementation, compared with the state-of-the-art solvers.
-An interesting open question is a refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.
+\wuhao{An interesting open question is a refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.
+}

### 2021-05-24 FMCAD2021/1_intro.tex
changes about complexity

@@
 %
+\wuhao{We also analyze the deterministic upper bound of the  complexity is triply exponential for the first time (Appendix~\ref{app:cpx}).
+Since the best known upper bound for Cooper's elimination algorithm is also triply exponential, it may not be easy to improve the deterministic upper bound.
+}
 Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

### 2021-05-24 FMCAD2021/paper.tex
space adjustment
changes about complexity

@@
 The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
-%    Besides the theoretical results, we also develop various optimizations and provide an implementation.
+\wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
     We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.
     The experimental results show that our implementation outperforms the state-of-the-art solvers.
-    %for string-integer conversion as well as linear-exponential constraints.
-    %, in both precision and efficiency.     
+
 %\keywords{ String-integer conversion \and Flat regular constraints \and Presburger arithmetic \and Exponential function \and Quantifier elimination.}
@@
 
-\vspace*{-2mm}
+\vspace*{-4mm}
 
@@
 
-\vspace*{-2mm}
+\vspace*{-3mm}
 \section{Optimizations}\label{sec-opt}
@@
 
-\vspace*{-3mm}
+\vspace*{-4mm}
 \section{Conclusion}

### 2021-05-24 FMCAD2021/1_intro.tex
sec1,4: complexity

@@
 %
-\wuhao{We also analyze the deterministic upper bound of the  complexity is triply exponential for the first time (Appendix~\ref{app:cpx}).
-Since the best known upper bound for Cooper's elimination algorithm is also triply exponential, it may not be easy to improve the deterministic upper bound.
+We also analyze the deterministic upper bound of the  complexity is triply exponential for the first time (Appendix~\ref{app:cpx}).
+\wuhao{Since the best known upper bound for Cooper's elimination algorithm is also triply exponential, it may not be easy to improve the bound.
 }

### 2021-05-24 FMCAD2021/paper.tex
sec1,4: complexity

@@
 The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-\wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
+\wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in deterministic time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
     We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.

### 2021-08-18 FMCAD2021/4_dec_proc.tex
small changes
complexity
changes about complexity
sec1,4: complexity
1

--- 2021-05-21 FMCAD2021/4_dec_proc.tex
@@
 \label{thm-paexp}
-{\paexp} admits quantifier elimination.
+{\paexp} admits quantifier elimination\wuhao{ in triply exponential deterministic time and space}. 
 \end{theorem}
@@
 Moreover, the procedure presented here slightly improves Point's procedure, in the following two aspects: 1) DNF (disjunctive normal form) was required in Point's procedure, which is not required here, 2) in Point's procedure, the divisibility constraints produced by the elimination of linear occurrences of a variable should be converted to equality constraints (by introducing fresh variables) before the elimination of exponential occurrences of some other variable, which is unnecessary here, since the divisibility constraints are directly dealt with in the elimination of exponential occurrences of variables.
-%Moreover, we give a complexity analysis to Point's algorithm, which was missing in \cite{Point86}.
+Moreover, we give a complexity analysis in Appendix~\ref{app:cpx}.% which was missing in \cite{Point86}.
 %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-In order to show $(\mathbb{N},+,2^x)$ admits QE , 
-it is sufficient to show that any 1-existential formula $\exists x.\theta(x,\bar{y})$ indeed does, 
-where $\theta(x,\bar{y})$ is a quantifier-free formula. 
-However, the form of $\exists x.\theta(x,\bar{y})$ is unknown so it may contain terms difficult to handle such as $2^{3x+y+1}$ or $l_2(x)$.
-The \textbf{Normalization} step simplifies these terms by introducing new variables, 
-for example, $\exists x.2^{3x+y+1}>10$ is equivalent to
-$$\exists x\exists x_1.2^{x_1}>10 \wedge x_1 =3x+y+1\,.$$
-
-Logarithm functions are replaced by exponential functions, take $\exists x.l_2(x)>3$ for example, $l_2(x)$ is replaced by a new variable $x_1$, and we have  
-$$\exists x\exists x_1. x_1 > 3 \wedge 2^{x_1}\le x \wedge x\le 2^{x_1+1}-1\,.$$
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 The normalization step first removes the occurrences of $\lambda_{10}(\iterm)$ and $\ell_{10}(\iterm)$  such that $\ivarx$ occurs  in $\iterm$, by encoding them with the exponential function.
@@
 After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\lambda_{10}(\iterm)$ as well as $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+A formula is \textit{normalized} if it satisfies the above properties.
 More details can be found in Appendix~\ref{app-norm}.
@@
 %$B\Def  \sum_{j=1}^{i}|b_j|$, 
-$B \Def 10(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
+$B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
 and $\delta\Def  \ell_{10}(A)+3$. 
@@
 $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
+\vspace{-2mm}
 \[

### 2021-08-18 FMCAD2021/8_appendix.tex
changes about complexity
1

@@
 
-\subsection{Complexity analysis}
+\subsection{Complexity analysis}\label{app:cpx}
 
-In this part, we will analysis the complexity of the decision procedure described in Section~\ref{sec-dec}.
-For convenience, we will slightly change the algorithm: the normalization procedure is invoked for all quantified variables in the beginning, which means step 2 and step 3 in the normalization procedure are performed for all quantified variables instead of a specified $\ivarx$.
+In this part, we analyse the complexity of the decision procedure for $\paexp$ described in Section~\ref{sec-dec}. For simplicity, we assume that the given formula is normalized
 
-Let the formula after the normalization step be $\varphi = \exists \vec{\ivarx}. \theta$ with $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$. Let $n$ denote the length of formula $\varphi$.
+We first analyse the preprocessing step, namely the normalization procedure. For the rest of the decision procedure, we adopt the style in the complexity analysis of \pa \ (cf. \cite{Oppen73}).
+
+Recall that in the normalization step, the formula is first transformed into the NNF and then fresh variables are introduced for terms of different forms.
+The transformation would double the length of the formula in the worst case when all atomic formulas are of the form $\iterm_1 = \iterm_2$ or its negation.
+\wuhao{How to compute the number of fresh variables in normalization?}
+
+After the normalization step, denote the obtained formula by $\varphi = \exists \vec{\ivarx}.
+\theta(\vec{\ivarx},\vec{\ivary})$ with quantified variables $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$ and free variables $\vec{\ivary}$.
+$\theta$ contains no quantifiers.
+Let $n$ denote the length of formula $\varphi$.
 

### 2021-09-13 FMCAD2021/8_appendix.tex
complexity: 40%

@@
 
-In this part, we analyse the complexity of the decision procedure for $\paexp$ described in Section~\ref{sec-dec}. For simplicity, we assume that the given formula is normalized
+In this part, we analyse the complexity of the quantifier elimination procedure applied on a formula with one existential quantifer, that is, a formula of the form $\exists \ivarx.
+\varphi(\ivarx, \vec{\ivary})$.
+We will prove that the complexity of eliminating one quantifer has a 3-EXPTIME upper bound.
+Since the quantifier elimination procedure for $\pa$ works as a subprocedure and also has a 3-EXPTIME complexity, the bound may not be improved easily\wuhao{ref Oppen73}.
 
-We first analyse the preprocessing step, namely the normalization procedure. For the rest of the decision procedure, we adopt the style in the complexity analysis of \pa \ (cf. \cite{Oppen73}).
+We first analyse the changes of formula length in the preprocessing step (normalization step).
+For the rest of the decision procedure, we adopt the style in the complexity analysis of \pa \ (cf.
+\cite{Oppen73}).
+The idea is that the formula length is bounded (up to a constant) by the product of the number of atoms, the number of coefficients and the length of the maximum constant.
+So we track the changes of such quantities throughout the procedure.
 
-Recall that in the normalization step, the formula is first transformed into the NNF and then fresh variables are introduced for terms of different forms.
-The transformation would double the length of the formula in the worst case when all atomic formulas are of the form $\iterm_1 = \iterm_2$ or its negation.
-\wuhao{How to compute the number of fresh variables in normalization?}
+Given a formula $\exists \ivarx.
+\varphi(\ivarx, \vec{\ivary})$ with length $n$.
+After the normalization step, suppose we obtain a formula with $m$ quantified variables $\exists \ivarx_m \dots\exists \ivarx_1.
+\varphi'(\ivarx_1,\dots, \ivarx_m, \vec{\ivary})$.
+We show that the length of the new formula is at most $6n$ and the number of quantified variables $m\le n$.
+In each sub-step of normalization, we analyse the worst situation: (1) \textit{Eliminate negations} suppose all atoms are of the form $\iterm_1=\iterm_2$, then taking negation will double the number of atoms, the length increases to $2n$ at most.
+(2) \textit{eliminate $\ell_10(\iterm)$ terms} the original formula has at most $n$ terms of the form $\ell_10(\iterm)$, for each of them, we introduce a fresh variable and two conjuncts.
+The formula length increases to $4n$.
+(3) \textit{flatten $10^\iterm$ terms} similar to (2), the formula has at most $n$ such terms and for each term, a fresh variable and a conjunct are added.
+The formula length increases to $5n$.
+(4) \textit{$\le$ atoms} similar to (1), here we assume all atoms are of the form $\iterm_1  = \iterm_2$, so the length of the formula with increase to $10n$ at most.
+
+The analysis for the normalization step is coarse, for example, the worst case in (1) and (4) or (2) and (3) can not happen at the same time.
+However, what we need is that the increased length of the formula is bounded by a constant factor.
+In addition, we can also conclude that the number of fresh variables are less than $n$ since there are at most $n$ different forms.
+
+\wuhao{stop here}
 

### 2021-09-16 FMCAD2021/8_appendix.tex
complexity done

@@
 
-%\subsection{Complexity analysis of Point's algorithm}
-
 \subsection{Elimination of exponential occurrences of variables for divisibility atomic formulas}\label{app-div}
@@
 \varphi'(\ivarx_1,\dots, \ivarx_m, \vec{\ivary})$.
-We show that the length of the new formula is at most $6n$ and the number of quantified variables $m\le n$.
+We show that the length of the new formula is at most $10n$ and the number of quantified variables $m\le n$.
 In each sub-step of normalization, we analyse the worst situation: (1) \textit{Eliminate negations} suppose all atoms are of the form $\iterm_1=\iterm_2$, then taking negation will double the number of atoms, the length increases to $2n$ at most.
@@
 
-\wuhao{stop here}
+After the normalization step, denote the obtained formula by $\exists \vec{\ivarx}.
+\varphi(\vec{\ivarx},\vec{\ivary})$ with quantified variables $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$ and free variables $\vec{\ivary}$.
+$\varphi$ contains no quantifiers.
+Let $n'$ denote the length of the formula.
 
-After the normalization step, denote the obtained formula by $\varphi = \exists \vec{\ivarx}.
-\theta(\vec{\ivarx},\vec{\ivary})$ with quantified variables $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$ and free variables $\vec{\ivary}$.
-$\theta$ contains no quantifiers.
-Let $n$ denote the length of formula $\varphi$.
+According to our quantifier elimination procedure,we eliminate the quantified variables one by one: each time select the largest variable among $\ivarx_i,1\le i\le m$, first eliminate the exponential occurrences, then linear occurrences.
+The linear order of quantified variables are given by a permutation.
+For $m$ quantified variables, there are $m!$ possible linear orders.
 
-Then we eliminate the exponential occurrences and linear occurrences of variables one by one according to a specified linear order of $\vec{\ivarx}$.
-Suppose the linear order is $\ivarx_1\ge \dots \ge \ivarx_m$.
+Suppose the specified linear order is $\ivarx_1\ge \ivarx_2 \ge \dots \ge \ivarx_m$, otherwise we rename the variables. While eliminating $\ivarx_k$
+The elimination of exponential
+Let $\exists \ivarx_m \dots \exists \ivarx_{k}\varphi'_k$ denote the formula after $k-1$ variables and exponential occurrences of $\ivarx_k$ are eliminated.
+Let $\varphi_k$ denote the formula obtained by eliminating linear occurrences of $\ivarx_k$ from $\varphi'_k$.
+For convenience, set $\varphi_0=\varphi$.
 
-We adopt the method in the analysis of the complexity of \pa \ (cf.
-\cite{Oppen73}).
-Let $\varphi'_k = \exists \ivarx_m \dots  \exists \ivarx_{k}\theta'_k$ denote the formula after $k-1$ variables and exponential occurrences of $\ivarx_k$ are eliminated from $\varphi$.
-Let $\varphi_k$ denote the formula obtained by eliminating linear occurrences of $\ivarx_k$ from $\varphi'_k$ and $\varphi_0=\varphi$.
-
-Let $c_k$ be the number of distinct $d$ in atoms of the form $d \ \big | \ \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of variables in $\varphi_k$.
+Let $c_k$ be the number of distinct $d$ in atoms of the form $d \ \big | \ \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of quantified variables in $\varphi_k$.
 Let $s_k$ be the largest constant (including coefficients) and $a_k$ be the number of atomic formulas in $\varphi_k$.
 Similarly we define $c'_k$, $s'_k$ and $a'_k$ for $\varphi'_k$.
-And let $c,s,a$ be $c_0,s_0,a_0$ respectively.
-We will prove the following lemma.
+Let $c,s,a$ be $c_0,s_0,a_0$ respectively.
+
+First we analyse the sub-procedure to eliminate exponential occurrences of $\ivarx_1$. We prove the following lemma.
+
 \begin{lemma}\label{lem:cpx exp}
@@
 \begin{proof}
-We analyze the complexity in the elimination of exponential occurrences of variable $\ivarx_1$. It will be discussed into 2 cases.
+
+The analysis is separated into 2 cases where all atoms are of the same form (inequalities atoms or divisiblilty atoms).
 
@@
 
-If all atoms are divisiblilty atomic formulas of the form $d \ \big | \ \iterm$.
-It is easy to see $c'_1<2c$ because a divisibility atomic formula  will produce at most two form of atomic formulas $d \ \big | \ \iterm$ and $\phi(d) \ \big | \ \iterm$.
+If all atoms are divisiblilty atomic formulas of the form $d \ \big | \ \iterm$ or $d \ \nmid \ \iterm$.
+We have $c'_1<2c$ since a divisibility atomic formula  will produce at most two forms of atomic formulas $d \ \big | \ \iterm$ and $\phi(d) \ \big | \ \iterm$.
 Note that any constant in $\iterm$ , say $l$, can be replaced by ($l \bmod d)$, so $s'_1\le s$.
-When $d$ is a large prime number, $\phi(d)=d-1$, a divisibility atomic formula is replaced by roughly $d$ atomic formulas, so $a'_1\le sa$
+When $d$ is a large prime number, $\phi(d)=d-1$, a divisibility atomic formula is replaced by roughly $d$ atomic formulas, so $a'_1\le sa$.
 
-Choose larger upper bound for $c'_1$, $s'_1$ and $a'_1$ respectively then the lemma is proved.
+Choose larger upper bound for $c'_1$, $s'_1$ and $a'_1$ respectively, then the lemma is proved.
 
@@
 
-Since linear occurrences of $\ivarx_1$ are eliminated using Cooper's algorithm, we combine Oppen's analysis and obtain:
+Since linear occurrences of $\ivarx_1$ are eliminated using Cooper's algorithm, we combine Oppen's analysis for this sub-procedure:
+
+\begin{lemma}\label{lem:cpx oppen}
+    \begin{align}
+        c_1&\le c'^4 \notag \\
+        s_1&\le s'^{4c'} \notag\\
+        a_1&\le {a'}^4 s'^{2c'}\notag 
+    \end{align} 
+\end{lemma}
+
+From Lemma~\ref{lem:cpx exp} and Lemma~\ref{lem:cpx oppen}, we have
 \begin{align}
@@
 
-Assume $m\le n$, by induction on $k$ we can prove the following lemma.
+Assuming $m\le n$, by induction on $k$ we get
 
@@
 
-Similarly to \cite{Oppen73}, we can assume $c\le n$, $a\le n$ and $s\le n$.So the space required to store $\varphi_k$ is the product of number of atoms $a_k$, the maximum number of constants $2m+2$ per atom, the maximum amount of space $s_k$ to store each constant and some constant $q$:
+If we assume $c\le n$, $a\le n$ and $s\le n$, the space required to store the quantifier free formula, $\varphi_k$, is bounded by the product of the number of linear orders $m!$, the number of atoms $a_k$, the maximum number of constants $2m+2$ per atom, the maximum amount of space $s_k$ to store each constant and some constant $q$:
 
-$${\sf space}\le q \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
-for some large constant $p$. It is also a bound for deterministic time.
-Since for arbitrary $\paexp$ formula length $n'$, the normalization for all variables results into a formula at most $n'^3$, so the bound is still 3-EXPTIME for $n'$.
+$${\sf space}\le q \cdot m! \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
+for some large constant $p$. It is also a bound for deterministic time\wuhao{why}.
 

### 2021-09-17 FMCAD2021/1_intro.tex
.

--- 2021-05-21 FMCAD2021/1_intro.tex
@@
 This can be seen as another contribution of this paper.
-%
+\wuhao{We also analyze the deterministic upper bound of the complexity for the first time, the time complexity of eliminating one quantifier is shown to be triply exponential (Appendix~\ref{app:cpx}).}
 Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).
-
-%
-%The optimizations include 
-%over approximation to reduce the number of subcases
-% the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
-We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-Although the general framework of our algorithm is quite akin to Point's,  it contains the following improvements and corrections: 
-\begin{itemize}
-	\item[i)] The syntax of Presburger arithmetic with exponential functions in our setting is more natural and easy to understand; Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables, that could increase the length of the considered $\paexp$ formula exponentially.
-In contrast, our algorithm just assumes that formulas are NNF (negation normal form).
-Therefore, our algorithm is much easy to follow and more efficient.
-\item[ii)] Unlike in Point's algorithm to eliminate the exponential part and then the linear part of the largest variable according to the given order one by one, in contrast, our algorithm eliminates the exponential parts of
-	all variables according to the order and then the resulted linear part, that improve the efficiency very much.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
-%Moreover, we propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implementation. 
-Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas.
-We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.
-
-\item[iii)] Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} (see later).
-\end{itemize}
- }
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-09-17 FMCAD2021/paper.tex
complexity done
.

--- 2021-05-24 FMCAD2021/paper.tex
@@
 \usepackage{multirow}
-
+\usepackage{amssymb}
 \usetikzlibrary{automata, positioning, arrows}
@@
 The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-\wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
+\wuhao{The complexity of the algorithm is analyzed for the first time and we show that the elimination of one existential quantifier is 3-EXPSPACE.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
     We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.

### 2021-09-18 FMCAD2021/3_flatten.tex
cpx analysis done

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-For a first order theory $T$,
-we say theory $T$ admits quantifier elimination (QE) if for any formula in $T$, 
-there is a quantifier-free formula equivalent to it.
-It is well-known that if a theory admits QE, 
-then it is a decidable theory.
-
-The formal definition of PA is given in section \ref{PA}.
-Here we add the ordering predicate $\le$ into the signature,
-which can be defined by $x \le y\, \Def \, \exists z. x+z=y$.
-However, the theory PA so far does not admit QE, 
-for example, consider the formula $\exists x.x = y + y$. 
-We augment the theory with countable unary divisible predicates
-$n|x$, where $n\in \mathbb{N}$, 
-$n|x$ is true if and only if $x\ \text{mod}\ n=0$ holds.
-This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 {\paexp} extends Presburger arithmetic with the exponential function $10^x$ as well as the (partial) functions $\ell_{10}(\ivarx)$ and $\lambda_{10}(\ivarx)$ (cf.
@@
 
-The syntax of {\paexp} is obtained from that of {\pa} by adding $10^\iterm$, $\ell_{10}(\iterm)$ and $\lambda_{10}(\iterm)$ to the definition of integer terms.
-Specifically, {\paexp} formulas are defined by the rules,
+The syntax of {\paexp} is obtained from that of {\pa} by adding $10^\iterm$ and $\ell_{10}(\iterm)$ to the definition of integer terms. Specifically, {\paexp} formulas are defined by the rules,
 %
@@
  \begin{array}{ l c l}
- \iterm &\Def& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \mid \lambda_{10}(\iterm), \\
+ \iterm &\Def& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \\
 \vspace{-1mm}
@@
 %Moreover, we introduce an abbreviation $\lambda_{10}(\ivarx) \equiv 2^{\ell_{10}(\ivarx)}$.
-From the definition of $\lambda_{10}(\ivarx)$, we know that for all $n \ge 1$, $\lambda_{10}(n) \le n < 10\lambda_{10}(n)$ holds.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-\begin{definition}
-    Let $\mathcal{L}=\{0,1,+,\le,n \, |x(n\in \mathbb{N}),2^x,l_2(x)\}$, 
-     $(\mathbb{N},+,2^x)$ be a $\mathcal{L}$-theory that has domain $\mathbb{N}$, where 
-    \begin{itemize}
-        \item  $2^x$ is interpreted to the exponential function of $2$ over $\mathbb{N}$; 
-        \item interpretations of $0,1,+,\le,=$ are consistent with PA;
-        \item for $n\ge 1$,$n|x$ holds iff $\exists y.x=ny$;
-        \item $2^0=1$, for $n \ge 1, 2^n = 2^{n-1}+2^{n-1}$;
-        We further assume that if $m,n\in \mathbb{N}, m\le n$, 
-        then $2^{m-n} = 1$
-        \item $l_2(0)=0$; for $n \ge 1,l_2(n) = y$ iff $2^y \le n < 2^{y+1}$;  $l_2(m-n) = 0$
-        if $m,n\in \mathbb{N}$ and $m\le n$. 
-    \end{itemize}
-\end{definition}
-
-$\lambda_2(x) = 2^{l_2(x)}$ can be defined by $l_2(x)$,
-intuitively, $\lambda_2(x)$ means the maximal power of 2 that is not larger than $x$.
-Then we have $\lambda_2(x) \le x \le 2\lambda_2(x)-1$,
-which will be useful in our proof. 
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We will use the notation $\lambda_{10}(\iterm)$ to denote $10^{\ell_{10}(\iterm)}$.
+From the definition of $\lambda_{10}(\iterm)$, we know that for all $n \ge 1$, $\lambda_{10}(n) \le n < 10\lambda_{10}(n)$ holds.
 
@@
 The construction of $\flatten_{\flatfun_\varphi}(\varphi)$ for the other atomic constraints is essentially the same as that in \cite{Parosh:20:PLDI} and thus omitted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-The flattening technique was first introduced 
-in \cite{Abdulla 2017}.
-Fix an alphabet $\Sigma$ and an abstraction parameter $\alpha$, 
-for a given atomic string constraint $\phi$, 
-flattening $\phi$ with parameter $\alpha$ 
-results in a new string constraint $\phi_\alpha$,
-such that 
-$R(\lVert \phi_\alpha \rVert) = \lVert \phi \rVert \cap \{I \mid \forall x\in X, I(x)\in \mathbb{G}'(\alpha)\}$,
-where $R$ is the renaming function with its domain extended to interpretations in the normal manner.
-Intuitively, it restricts $\phi$ to interpret 
-string variables over $\mathbb{G}'(\alpha)$.
-
-\cite{Abdulla 2017} discussed the flattening of basic 
-string constraints including equality, integer, (regular) grammar and transducer constraints. 
-For an atomic string constraint $\phi$,
-the flattening $\phi_\alpha$ is still an atomic string constraint 
-and is Parikh definable,
-so its Parikh image can be expressed by a quantifier free PA formula.
-Together with the purity condition,
-we obtain an existential quantified PA formula $\rho$.
-$\rho$ will be sent to a SMT solver,
-if the solver returns an solution $\theta$,
-then we can construct an interpretation for $\phi_\alpha$ from $\theta$,
-otherwise it means $\phi$ is unsatistiable when 
-string variables are interpreted  to $\alpha$-flat languages.
-
-Take a regular constraint $\phi = x\in L(A)$ for example,
-the flattening of $\phi$ resutls in a new finite state automaton $A'$ over $\Sigma(\alpha)$,
-which encodes running $A$ ``in parallel" 
-with the generic $\alpha$-flat automaton.
-Let $\rho_1$ be the formula describing the Parikh image of $A'$,
-which is a formula over variable sets $\Sigma(\alpha)^\#$.
-Let $\rho_2$ be the purity condition \eqref{eq:purity}.
-Then we obtain the PA formula $\exists (\Sigma(\alpha))^\#. \rho_1 \wedge \rho_2$.
-In order to distinguish between different string variables,
-we may replace $a(i)^\# \in \Sigma(\alpha)^\#$ by $(x,a(i))^\#$.
-
-Since the structure of a flat automaton is decided 
-by its abstraction parameter $\alpha$, 
-a Counter-Example Guided Abstraction 
-Refinement (CEGAR) framework is designed, which contains both an under- and an 
-over-approximation module, to search the possible values of $\alpha$.
-The termination for the overall algorithm is not guaranteed.
-
-The string-number conversion functions are commonly used functions
-in most of programming languages,
-for example,
-\verb+parseInt()+ in Java and \verb+Int()+ in Python.
-The functions usually take two parameters, 
-a string over the agreed alphabet $\Sigma$
-and an optional parameter denotes the radix.
-They parse the string according to the rules indicated by the radix,
-and return an integer denoted by the string.
-
-From the view of string constraints,
-string-number conversion functions give rise to a new form of string constraints
-and are more expressive than length constraints.
-So we consider extending string constraints with 
-{\parseInt} function. 
-As the general problem of string constraints is undecidable,
-we still adopt the idea of flattening, 
-i.e., variables are restricted to (generic) flat languages.
-This problem has been investigated in \cite{POPL20},
-which defined a special form of flat restriction (straight-line PFA) and 
-proposed a heuristic search method.
-
-In this section, 
-we describe the problem of interest first, and then 
-present an reduction from the problem of solving flat string constraints with {\parseInt} function
-to the decidability problem of Presburger Arithmetic with exponential functions.
-Hence, we identify a decidable subset of string constraints, which is the largest one with decidability so far to the best of our knowledge.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-09-18 FMCAD2021/4_dec_proc.tex
cpx analysis done

@@
 
-Sem\"{e}nov first proved that  {\paexp} admits quantifier elimination in \cite{Semenov84}, thus its satisfiability problem is decidable.
-However, Sem\"{e}nov did not give a concrete quantifier elimination procedure.
+Semënov first proved that  {\paexp} admits quantifier elimination in \cite{Semenov84}, thus its satisfiability problem is decidable.
+However, Semënov did not give a concrete quantifier elimination procedure.
 Remedying this, Point proposed a quantifier elimination procedure for {\paexp} in \cite{Point86}.
@@
 \label{thm-paexp}
-{\paexp} admits quantifier elimination\wuhao{ in triply exponential deterministic time and space}. 
+{\paexp} admits quantifier elimination. 
 \end{theorem}
@@
 Moreover, the procedure presented here slightly improves Point's procedure, in the following two aspects: 1) DNF (disjunctive normal form) was required in Point's procedure, which is not required here, 2) in Point's procedure, the divisibility constraints produced by the elimination of linear occurrences of a variable should be converted to equality constraints (by introducing fresh variables) before the elimination of exponential occurrences of some other variable, which is unnecessary here, since the divisibility constraints are directly dealt with in the elimination of exponential occurrences of variables.
-Moreover, we give a complexity analysis in Appendix~\ref{app:cpx}.% which was missing in \cite{Point86}.
-%
+Furthermore, we give a complexity analysis in Appendix~\ref{app:cpx}, which shows that the elimination of one quantifier has a space complexity of 3-EXPSPACE. \wuhao{relation with PA}
 %
 
-Furthermore, as Point's procedure is quite expensive and a faithful implementation would not scale\footnote{We did implement Point's algorithm and discovered that the implementation could only solve formulas of very small size.},
-in the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
+However, Point's procedure is quite expensive and a faithful implementation would not scale\footnote{We did implement Point's algorithm and discovered that the implementation could only solve formulas of very small size.}.
+In the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
 
@@
 \ \varphi$ is equivalent to $\neg \exists \ivarx.
-\ \neg \varphi$,  thus in the sequel, we only need to  show that every {\paexp} formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
+\ \neg \varphi$,  thus in the sequel, we only need to  show that every $\paexp$ formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
 
@@
 The elimination of $\ivarx_3$ in $\varphi_2$ here is simple, with $\ivarx_3$ replaced by $\ivarx_1 + \ivarx_2$.
-\hide{, resulting into  
-\[
-\small
-\begin{array}{l}
-\varphi_3 \Def \exists \ivarx_2.\ 
-\ivarx_1 + \ivarx_2 \le \ell_{10}(\ivary+1001)\ \vee \\
-\left(
-\begin{array}{l}
-\ivarx_1 + \ivarx_2 = \ell_{10}(\ivary+1001)+1\ \wedge \\
-10 \lambda_{10}(\ivary+1001) - 10^{\ivarx_2} \le \ivary + 1001 
-\end{array}
-\right) \vee \\
-%
- \bigvee \limits_{j \in \{0,1,2\}}  \left(\ivarx_1 + \ivarx_2 = \ivarx_2 +j \wedge 10^{\ivarx_2 + j} - 10^{\ivarx_2} \le \ivary + 1001\right).
-\end{array}
-\]}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-\begin{definition}
-    For a strictly increasing function $f:\mathbb{N}\mapsto \mathbb{N}$, 
-    $f$ is said to be \emph{compatible with addition}
-    if for every $m\in \mathbb{N}^+$,  $f$ modulo $m$ is periodic, 
-    and for any term $A(x)=\sum_{1\le i\le n} a_i f(x+b_i)$, 
-    where $n \in \mathbb{N}^+, a_i ,b_i \in \mathbb{Z}$,
-    one of the following holds:
-    \begin{itemize}
-        \item $A(x)$ is bounded; 
-        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
-        $\forall x. A(x+\Delta_A)\ge f(x)$;
-        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
-        $\forall x. -A(x+\Delta_A)\ge f(x)$.
-    \end{itemize}
-\end{definition}
-
-Semenov proved that for any function $f$ that is compatible with addition, the first order theory $(\mathbb{N},+,f)$ admits QE  and thus is complete and decidable \cite{Semenov84}.
-Exponential functions are a typical class of compatible with addition functions, and therefore {$\paexp$} is decidable.
-Particularly, in \cite{Point86}
-Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$.
-%下面这些感觉可以不用说，或者放在后面
-%Unfortunately, Point's algorithm is flawed, one problem lies in the discussion to eliminate exponential terms.
-For example, after  eliminating variable $x$ in formula $\exists x.
-y\le 2^x \wedge x\le 10$, the expected result should be $y\le 2^{10} =1024$, but Point's algorithm will return a formula equivalent to $y \le 2^6=64$.
-
-In this section, within Point's framework, we give a revised algorithm to the decision problem of {$\paexp$} with both improvements and corrections, that provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to Theorem~\ref{thm:string-parInt}.
-
-%\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
-% Based on Point's work \cite{Point86}, 
-We will present a revised QE algorithm for {$\paexp$} formulas of the form $Qx.\theta(x,\bar{y})$, where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula.
-Since $\forall x.
-F = \neg \exists x.
-\neg F$, we further assume the quantifier is an existential quantifier, that is, $\exists x.\theta(x,\bar{y})$.
-For eliminating quantifiers in arbitrary {$\paexp$} formula, we need to apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.
-
-The whole QE algorithm can be divided into 4 steps.
-The first step \textbf{Normalization} can be viewed as a pre-processing step, which substitutes ``complex" terms like $2^{2^x}$, $l_2(3x+y)$ with ``simple" terms by introducing new variables.
-After \textbf{Normalization}, we move forward to handle a ``simpler" formula, however, at the cost of more quantified variables.
-
-The other three steps undertake the task to eliminate the newly introduced variables in \textbf{Normalization} step, denoted by $\{x_i:1\le i\le n\}$, and the original variable $x$ (will be denoted by $x_0$) one by one.
-The main body of \textbf{Enumerate-Orders} step is a loop to enumerate all possible orders among quantified variables $\{x_i : 0\le i \le n\}$. 
-In each iteration, according the given (decreasing) order of quantified variables (corresponding to an outer \emph{for} loop), we invoke \textbf{Elim-exp} and \textbf{Elim-linear} to eliminate these variables one by one.
-
-During each iteration of the inner \emph{for} loop, if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atom, we will invoke \textbf{Elim-exp} to produce an equivalent formula where $x_i$ occurs linearly.
-Then \textbf{Elim-linear} will eliminate all linear occurrences of $x_i$, which is similar to the classic Cooper's QE algorithm for PA \cite{Cooper72}.
-
-A more detailed description is given below.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \vspace*{-3mm}
@@
 
-The normalization step first removes the occurrences of $\lambda_{10}(\iterm)$ and $\ell_{10}(\iterm)$  such that $\ivarx$ occurs  in $\iterm$, by encoding them with the exponential function.
+The normalization step first removes the occurrences of $\ell_{10}(\iterm)$ where $\ivarx$ occurs in $\iterm$, by encoding them with the exponential function.
 Moreover, for each occurrence of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, it introduces a fresh variable $\ivarz$, replaces $10^\iterm$ with $10^\ivarz$, and adds the equality $\ivarz = \iterm$.
 It also applies some additional transformations.
-After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\lambda_{10}(\iterm)$ as well as $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+Denote the negation of $c | \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
 A formula is \textit{normalized} if it satisfies the above properties.
 More details can be found in Appendix~\ref{app-norm}.
-
-%\yfc{Maybe describe the shape of normalized formula here. After normalization, all formulae are in the form ...}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-Then $\theta(x,\bar{y})$ becomes a Boolean combination (with only $\wedge$ and $\vee$) of literals.
-
-Make substitutions by introducing new variables according to the \textit{while} statements in the pseudo-code.
-For consistency, we rename the original $x$ to be $x_0$ and assume $n$ new quantified variables are introduced.
-
-%After introducing new variables and substitution, 
-After substitutions, we obtain a formula where if an exponential term contains a quantified variable $x_i$, the term must be $2^{x_i}$.
-We then collect all terms with $x_i (0\le i\le n)$ together, it will be of the form $s(\bar{x}) \Def \sum_{i=0}^{n} a_i 2^{x_i} + \sum_{i=0}^n b_i x_i$, where $a_i,b_i (0\le i \le n)$ are all constants.
-Other terms including constants and terms of free variables $\bar{y}$ are collected, denoted by $t(\bar{y})$.
-Since $\bar{y}$ are free variables, we will regard $t(\bar{y})$ as a constant.
-Inequalities and equalities will all be transformed into $s(\bar{x})\le t(\bar{y})$, for example, $s(\bar{x}) = t(\bar{y})$ will be replaced by $s(\bar{x})\le t(\bar{y})\wedge t(\bar{y})\le s(\bar{x})$.
- end, the resulted formula $\theta'$ only contains literals of the forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ and $\neg  (k|s(\bar{x})+t(\bar{y}))$.
-
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{1-existential formula $\exists x.\theta(x,\bar{y})$}
-    \KwOut{n-existential formula $\exists \bar{x}.\theta'(\bar{x},\bar{y})$}
-
-    $\theta' :=  \theta(x,\bar{y})$ \;
-    Transform $\theta'$ into NNF\;
-    \tcp{ i is used for counting the introduced variables }
-    $i := 1$\; 
-    \tcp{ $\bar{y}$ are regarded as constants} 
-    \While{there is a term $l_2(t)$, $t$ is not a constant}
-    {
-        $\theta' :=  \theta'[x_i/l_2(t)] \wedge (2^{x_i}\le t) \wedge (t\le 2^{x_i+1}-1)$\;
-        $i:= i+1$\;
-    }
-    \While{there is a term $2^t$, $t$ is not a variable $x_j(j<i)$ or a constant}
-    {
-        $\theta' :=  \theta'[2^{x_i}/2^t]\wedge (x_i\le t) \wedge (t \le x_i)$\;
-        $i:= i+1$\;
-    }
-    $n:= i-1$\;
-    \tcp{ Collect quantified variables}
-    Transform all atoms of $\theta'$ into forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ or $\neg  (k|s(\bar{x})+t(\bar{y}))$\;
-    Return $\exists x_0,...,\exists x_n. \theta'$
-
-    \caption{Normalization}
-\end{algorithm}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+Suppose $n-1$ fresh variables are introduced in the normalization procedure, rename the original variable $\ivarx$ and the $n-1$ introduced variables $\ivarx_i,1\le i\le n$.
 Let the output of the normalization procedure be $\exists \vec{\ivarx}.\ \varphi'$ with $\vec{\ivarx} = (\ivarx_1,\ldots, \ivarx_n)$. 
@@
 
-Assuming a linear order $\sigma \in \mathcal{S}_n$ of $\{\ivarx_1,\ldots, \ivarx_n\}$, we then consider $\varphi'_\sigma  = \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ and eliminate the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(1)}$,  one by one and from the innermost to the outmost.
+Assuming a linear order $\sigma \in \mathcal{S}_n$ of $\{\ivarx_1,\ldots, \ivarx_n\}$, we then consider $\varphi'_\sigma  = \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ and eliminate the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(1)}$,  one by one and from $\ivarx_n$ to $\ivarx_1$.
 Let $\varphi''_\sigma$ denote the resulting formula.
@@
 %\yfc{Explain why only the largest variable can  be eliminated. Which part of the procedure we need this condition? Or if we do not have the order, what can be wrong?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-In the \textbf{Normalization} step, we simplify the origin formula at the cost of more quantified variables.
-Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$ and it has $n$ new variables $x_i$.
-As a consequence, we need to eliminate all $(n+1)$ quantified variables one by one.
-
-%To the end, 
-We denote the set of all orders among the $(n+1)$ variables by $S_{n+1}$, and then enumerate these orders one by one in the outer \textit{for} loop.
-For the considered order $\sigma$, we first add the ordering information to $\theta$, and then according to the order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{Elim-exp} first, and then invoking \textbf{Elim-linear} (the inner \textit{for} loop).
-In each iteration of the inner \textit{for} loop, if $x_{\sigma{(i)}}$ occurs in an exponential term in an atom $\tau$, \textbf{Elim-exp} is invoked to produce a formula, in which $x_i$ occurs linearly, equivalent to $\theta$;
-then $x_i$ occurs only linearly, thus \textbf{Elim-linear} is further invoked to eliminate all (linear) occurrences of $x_i$.
-The returned formula is the disjunction of all formulas resulted in all iterations of the outer \textit{for} loop.
-
-\iffalse
-put , We now specify an order of $\bar{x}$, and recursively eliminate the maximal element in the remaining $\bar{x}$. 
-That is to say,
-every execution of \textbf{Elim-exp} and \textbf{Elim-linear} will eliminate the largest element $x_i$ in $\bar{x}$.
-However, as you may notice, if there is no clue about the order of $\bar{x}$, 
-we will have to use the idea of permutation group and the number of sub-cases will blow up, in the worst case, $(n+1)!$ cases.
-
-Here is how \textbf{Elim-exp} and \textbf{Elim-linear} are invoked in
-\textb
-At thef{Ordering and QE}. 
-\fi
-
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{a normalized $(n+1)$-existential formula $\exists \bar{x}.\theta(\bar{x},\bar{y})$, where $\bar{x}=(x_0,...,x_n)$} 
-    \KwOut{an equivalent quantifier-free formula without $\bar{x}$}
-    \tcp{Let $S_{n+1}$ denote the group of permutations on $\{0,1,...,n\}$}
-    $\phi := \textit{False}$\;
-    \For{each $\sigma \in S_{n+1}$ }
-    {
-        $\theta_{\sigma}:=  \theta(\bar{x},\bar{y})\wedge \bigwedge_{j=0}^{n-1}(x_{\sigma(j)}\le x_{\sigma(j+1)})$\;
-
-        \tcp{recursively eliminate the maximal $x_i$ in $\bar{x}$}
-        %\znj{How to effectively determine the maximal $x_i$?} 
-        \For{i from n to 0}
-        {
-            \tcp{if $x_{\sigma(i)}$ occurs exponentially in $\theta_\sigma$}
-            $\theta_{\sigma}:= \textbf{Elim-exp}(x_{\sigma(i)},\theta_{\sigma})$\;
-            \tcp{now $x_{\sigma(i)}$ occurs only linearly in $\theta_\sigma$}
-            $\theta_{\sigma}:= \textbf{Elim-linear}(x_{\sigma(i)},\theta_{\sigma})$\;
-        }
-        $\phi :=  \phi \vee \theta_{\sigma}$\;
-    }
-    output $\phi$
-    \caption{Enumerate-Orders}
-\end{algorithm}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Recall that after normalization, the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+Recall that after normalization, the atomic formulas are of the form $\iterm_1 \le \iterm_2$, $c | \iterm$ or $c \nmid \iterm$.
 Therefore, we can assume that the atomic formulas in $\varphi''_{\sigma,i}$ are  of the form
@@
 or  
-$c \ \big  | \ \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big).$
-%
-For convenience, let us call these formulas as inequality respectively divisibility atomic formulas.
+$c \ \big  | \ \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big)$
+(or $\nmid$).
+
+For convenience, let us call these formulas as inequality respectively (in)divisibility atomic formulas.
 In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these inequality atomic formulas.
-The elimination of the exponential occurrences of the divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.
+The elimination of the exponential occurrences of the (in)divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.
 

### 2021-09-18 FMCAD2021/8_appendix.tex
.
cpx analysis done

@@
 \begin{enumerate}
-\item At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form).
-Moreover, we remove the occurrences of $\neg$ by (a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, (b)  $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, (c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, (d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
+\item \textit{NNF transformation} At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form). Moreover, we remove the occurrences of $\neg$ by replacing 
+\hide{(a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, }
+(a) $\neg c | \iterm$ with $ c \nmid \iterm$, 
+(b) $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, 
+(c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, 
+(d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
 %
-\item Repeat the following procedure, until there are no $\lambda_{10}(\iterm)$ or $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: Replace each occurrence of $\lambda_{10}$ in $\varphi'$, say $\lambda_{10}(\iterm)$, such that $\ivarx$ occurs in $\iterm$, by $10^{\ell_{10}(\iterm)}$.
-Then replace each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct.
+\item \textit{replace $\ell_{10}(\iterm)$ terms} Repeat the following procedure, until there are no $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: for each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct.
+Note that if $\iterm$ contains no variables, then $\ell_{10}(\iterm)$ is a constant.
+In this case, we can also assume that $\iterm$ contains $\ivarx$ and perform the same replacements, which helps in the analysis of complexity in \ref{app:cpx}.
 Let the resulting formula be $\varphi''$.
@@
 
-\item Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $10^{\iterm}$ in $\varphi''$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $10^{\iterm}$ by $10^\ivarz$, moreover, add the constraint $\ivarz = \iterm$ as a conjunct.
+\item \textit{flatten $10^\iterm$ terms} Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $10^{\iterm}$ in $\varphi''$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $10^{\iterm}$ by $10^\ivarz$, moreover, add the constraint $\ivarz = \iterm$ as a conjunct.
 Let $\varphi'''$ denote the resulting formula.
 
-\item Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c | \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
 Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
@@
 Let $\varphi^\dag$ the resulting formula.
-%
+
 \item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_n$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
@@
 
-\subsection{Elimination of exponential occurrences of variables for divisibility atomic formulas}\label{app-div}
+\subsection{Elimination of exponential occurrences of variables for (in)divisibility atomic formulas}\label{app-div}
 
-Consider
+Consider a divisiblilty atomic formula 
 %
@@
 $$
-with $a_i \neq 0$.
+with $a_i \neq 0$. The indivisibility case can be treated analogously by taking a negation.
 %
@@
 \varphi(\ivarx, \vec{\ivary})$.
-We will prove that the complexity of eliminating one quantifer has a 3-EXPTIME upper bound.
-Since the quantifier elimination procedure for $\pa$ works as a subprocedure and also has a 3-EXPTIME complexity, the bound may not be improved easily\wuhao{ref Oppen73}.
+We will prove that the complexity of eliminating one quantifer has a 3-EXPSPACE upper bound.
+Since the quantifier elimination procedure for $\pa$ works as a subprocedure (in eliminating linear occurrences of variables) and also has a 3-EXPSPACE complexity, the bound may not be easily improved.
 
-We first analyse the changes of formula length in the preprocessing step (normalization step).
+We first analyse the changes of formula length in the normalization step.
 For the rest of the decision procedure, we adopt the style in the complexity analysis of \pa \ (cf.
 \cite{Oppen73}).
-The idea is that the formula length is bounded (up to a constant) by the product of the number of atoms, the number of coefficients and the length of the maximum constant.
+The idea is that the upper bound of the formula length can be expressed using the product of the number of atoms, the number of coefficients and the length of the maximum constant.
 So we track the changes of such quantities throughout the procedure.
@@
 \varphi(\ivarx, \vec{\ivary})$ with length $n$.
-After the normalization step, suppose we obtain a formula with $m$ quantified variables $\exists \ivarx_m \dots\exists \ivarx_1.
+After the normalization step, suppose we obtain a formula with $m$ quantified variables $\exists \ivarx_1 \dots\exists \ivarx_m.
 \varphi'(\ivarx_1,\dots, \ivarx_m, \vec{\ivary})$.
 We show that the length of the new formula is at most $10n$ and the number of quantified variables $m\le n$.
-In each sub-step of normalization, we analyse the worst situation: (1) \textit{Eliminate negations} suppose all atoms are of the form $\iterm_1=\iterm_2$, then taking negation will double the number of atoms, the length increases to $2n$ at most.
-(2) \textit{eliminate $\ell_10(\iterm)$ terms} the original formula has at most $n$ terms of the form $\ell_10(\iterm)$, for each of them, we introduce a fresh variable and two conjuncts.
+In each sub-step of normalization, we analyse the worst situation:
+(1) \textit{NNF transformation} suppose all atoms are of the form $\iterm_1=\iterm_2$, then taking negation will double the number of atoms, the length increases to $2n$ at most. 
+(2) \textit{replace $\ell_{10}(\iterm)$ terms} the original formula has at most $n$ terms of the form $\ell_{10}(\iterm)$, for each of them, we introduce a fresh variable and two conjuncts.
 The formula length increases to $4n$.
-(3) \textit{flatten $10^\iterm$ terms} similar to (2), the formula has at most $n$ such terms and for each term, a fresh variable and a conjunct are added.
-The formula length increases to $5n$.
-(4) \textit{$\le$ atoms} similar to (1), here we assume all atoms are of the form $\iterm_1  = \iterm_2$, so the length of the formula with increase to $10n$ at most.
+(3) \textit{flatten $10^\iterm$ terms} similar to (2), the formula has at most $n$ such terms and for each term, a fresh variable and a conjunct are added. The formula length increases to $5n$. 
+(4) \textit{$\le$ transformation} similar to (1), here we assume all atoms are of the form $\iterm_1  = \iterm_2$, so the length of the formula with increase to $10n$ at most.
 
-The analysis for the normalization step is coarse, for example, the worst case in (1) and (4) or (2) and (3) can not happen at the same time.
-However, what we need is that the increased length of the formula is bounded by a constant factor.
-In addition, we can also conclude that the number of fresh variables are less than $n$ since there are at most $n$ different forms.
+The analysis for the normalization step is coarse, for example, the worst case in (1)\&(4) or (2)\&(3) can not happen at the same time.
+However, what we need is that the increased length of the formula is bounded by a constant factor,i.e.
+10.
+In addition, we can also conclude that the number of fresh variables, $m$, is less than $n$ since there are at most $n$ different forms of terms.
 
@@
 
-Suppose the specified linear order is $\ivarx_1\ge \ivarx_2 \ge \dots \ge \ivarx_m$, otherwise we rename the variables. While eliminating $\ivarx_k$
-The elimination of exponential
-Let $\exists \ivarx_m \dots \exists \ivarx_{k}\varphi'_k$ denote the formula after $k-1$ variables and exponential occurrences of $\ivarx_k$ are eliminated.
-Let $\varphi_k$ denote the formula obtained by eliminating linear occurrences of $\ivarx_k$ from $\varphi'_k$.
-For convenience, set $\varphi_0=\varphi$.
+Suppose the specified linear order is $\ivarx_m\ge \dots \ge \ivarx_1$.
+Let $\exists \ivarx_1\dots \exists\ivarx_{m-k}.\varphi_k$ denote the formula obtained from $\exists \vec{\ivarx}.
+\varphi(\vec{\ivarx},\vec{\ivary})$ by eliminating quantifiers $\exists \ivarx_m,\dots,\exists\ivarx_{m-k+1}$.
+Let $\exists \ivarx_1 \dots \exists \ivarx_{m-k+1}.\varphi'_k$ denote the formula obtained from $\exists \ivarx_1\dots \exists\ivarx_{m-k+1}.\varphi_{k-1}$ by eliminating linear occurrences of $\ivarx_{m-k+1}$.
+Let $\varphi_0$ denote $\varphi$.
 
@@
 
-First we analyse the sub-procedure to eliminate exponential occurrences of $\ivarx_1$. We prove the following lemma.
+First we analyse the sub-procedure to eliminate exponential occurrences of $\ivarx_m$. We prove the following lemma.
 
@@
 If all atoms are inequalities atomic formulas of the form in Lemma~\ref{lem:exp-ineq}.
-We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_1$ is replaced by a new formula.
-Only the coefficients of linear occurrences of $\ivarx_1$ and $\ivarx_2$ will be changed: constant coefficient $1$ is introduced, and if we substitute $x_1$ by $x_2+p$ for some $p$, coefficient of linear occurrence of $x_2$ will become $b_1+b_2$ ($b_1,b_2$ are coefficients for $x_1$ and $x_2$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
+We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_m$ is replaced by a new formula.
+Only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant coefficient $1$ is introduced, and if we substitute $x_m$ by $x_{m-1}+p$ for some $p$, coefficient of linear occurrence of $x_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $x_{m}$ and $x_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
 Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le c^2$.
-Note that $\delta$ and $B$ are at most $\ell_{10}(ms)$, when we substitute $\ivarx_1$ by $\ivarx_2+\delta-1$ or by $B-1$, the largest constant in the formula is at most $s\cdot  10^{\ell_{10}(ms)}\le ms^2$.
+Note that $\delta$ and $B$ are at most $\ell_{10}(ms)$, when we substitute $\ivarx_m$ by $\ivarx_{m-1}+\delta-1$ or by $B-1$, the largest constant in the formula is at most $s\cdot  10^{\ell_{10}(ms)}\le ms^2$.
 And an inequality is replaced by at most $4\ell_{10}(ms)$ atomic formulas, so $a'_1\le 4\ell_{10}(ms)a$.
@@
 
-Since linear occurrences of $\ivarx_1$ are eliminated using Cooper's algorithm, we combine Oppen's analysis for this sub-procedure:
+Since linear occurrences of $\ivarx_m$ are eliminated using Cooper's algorithm, we combine Oppen's analysis for this sub-procedure:
 
@@
 $${\sf space}\le q \cdot m! \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
-for some large constant $p$. It is also a bound for deterministic time\wuhao{why}.
+for some large constant $p$. \wuhao{time complexity?}.
 

### 2021-09-29 FMCAD2021/1_intro.tex
complexity add a remark

@@
 This can be seen as another contribution of this paper.
-\wuhao{We also analyze the deterministic upper bound of the complexity for the first time, the time complexity of eliminating one quantifier is shown to be triply exponential (Appendix~\ref{app:cpx}).}
+We also analyze the deterministic upper bound of the complexity for the first time, the space complexity of eliminating one quantifier is shown to be 3-EXPSPACE.
 Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

### 2021-09-29 FMCAD2021/3_flatten.tex
complexity add a remark

@@
 
-The function  {\parseInt} takes a decimal string as the input and returns the integer represented by the string\footnote{The {\parseInt} function in scripting languages e.g.
-Javascript is more general in the sense that the base can be a number between 2 and 36.
+The function  {\parseInt} takes a decimal string as the input and returns the integer represented by the string\footnote{The {\parseInt} function in scripting languages, e.g.
+Javascript, is more general in the sense that the base can be a number between 2 and 36.
 Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for readability.}.
@@
 
-{\paexp} extends Presburger arithmetic with the exponential function $10^x$ as well as the (partial) functions $\ell_{10}(\ivarx)$ and $\lambda_{10}(\ivarx)$ (cf.
+{\paexp} extends Presburger arithmetic with the exponential function $10^x$ as well as the (partial) functions $\ell_{10}(\ivarx)$ (cf.
 \cite{Point86}).
-The functions $\ell_{10}(\ivarx)$ and $\lambda_{10}(\ivarx)$ are inductively defined as follows.
-\begin{itemize}
-\item The (partial) function $\ell_{10}(\ivarx)$ is\hide{inductively} defined as follows: For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$. (Note that $\ell_{10}(0)$ is undefined.) 
+The function $\ell_{10}(\ivarx)$ is inductively defined as follows.
+For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$.
+(Note that $\ell_{10}(0)$ is undefined.)
+%and $\lambda_{10}(\ivarx)$ are inductively defined as follows.
+%\begin{itemize}
+%\item The (partial) function $\ell_{10}(\ivarx)$ is\hide{inductively} defined as follows: For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$. (Note that $\ell_{10}(0)$ is undefined.) 
 %
-\item The (partial) function $\lambda_{10}(\ivarx)$ can be defined by $\ell_{10}(\ivarx)$: For $n \ge 1$, $\lambda_{10}(n) = 10^{\ell_{10}(n)}$.
+%\item The (partial) function $\lambda_{10}(\ivarx)$ can be defined by $\ell_{10}(\ivarx)$: For $n \ge 1$, $\lambda_{10}(n) = 10^{\ell_{10}(n)}$.
 (Intuitively, $\lambda_{10}(\ivarx)$ is the maximum power of 10 that is no greater than $\ivarx$.)
-\end{itemize}
+%\end{itemize}
 
@@
  $$
- The semantics of {\paexp} are defined similarly to {\pa} with the difference that variables are interpreted over $\Nat$, that is, an interpretation is a function $I: \free(\phi) \rightarrow \Nat$.
-The quantifier-free as well as existential {\paexp} formulas are defined similarly to {\pa} as well.
+ The semantics of {\paexp} are defined similarly to {\pa} with the difference that variables are interpreted over $\Nat$ (to avoid terms like $10^{-2}$), that is, an interpretation is a function $I: \free(\phi) \rightarrow \Nat$.
+ The quantifier-free and existential {\paexp} formulas are defined similarly to {\pa} as well.
 
@@
 %Moreover, we introduce an abbreviation $\lambda_{10}(\ivarx) \equiv 2^{\ell_{10}(\ivarx)}$.
-We will use the notation $\lambda_{10}(\iterm)$ to denote $10^{\ell_{10}(\iterm)}$.
-From the definition of $\lambda_{10}(\iterm)$, we know that for all $n \ge 1$, $\lambda_{10}(n) \le n < 10\lambda_{10}(n)$ holds.
+We will use the notation $\lambda_{10}(\iterm)$ to denote $10^{\ell_{10}(\iterm)}$. It is easy to show that for all $n \ge 1$, $\lambda_{10}(n) \le n < 10\lambda_{10}(n)$ holds.
 

### 2021-09-29 FMCAD2021/4_dec_proc.tex
complexity add a remark

@@
 Moreover, the procedure presented here slightly improves Point's procedure, in the following two aspects: 1) DNF (disjunctive normal form) was required in Point's procedure, which is not required here, 2) in Point's procedure, the divisibility constraints produced by the elimination of linear occurrences of a variable should be converted to equality constraints (by introducing fresh variables) before the elimination of exponential occurrences of some other variable, which is unnecessary here, since the divisibility constraints are directly dealt with in the elimination of exponential occurrences of variables.
-Furthermore, we give a complexity analysis in Appendix~\ref{app:cpx}, which shows that the elimination of one quantifier has a space complexity of 3-EXPSPACE. \wuhao{relation with PA}
+Furthermore, we give a complexity analysis in Section~\ref{sec-cpx}, which shows that the elimination of one quantifier has a space complexity of 3-EXPSPACE. 
 %

### 2021-09-29 FMCAD2021/5_complexity.tex
complexity add a remark

--- 2021-09-18 FMCAD2021/8_appendix.tex
@@
 %!TEX root = paper.tex
-
-\subsection{Details of the normalization step}\label{app-norm}
-%
-The normalization step comprises the following sub-steps.
-\begin{enumerate}
-\item \textit{NNF transformation} At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form). Moreover, we remove the occurrences of $\neg$ by replacing 
-\hide{(a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, }
-(a) $\neg c | \iterm$ with $ c \nmid \iterm$, 
-(b) $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, 
-(c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, 
-(d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
-%
-\item \textit{replace $\ell_{10}(\iterm)$ terms} Repeat the following procedure, until there are no $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: for each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct.
-Note that if $\iterm$ contains no variables, then $\ell_{10}(\iterm)$ is a constant.
-In this case, we can also assume that $\iterm$ contains $\ivarx$ and perform the same replacements, which helps in the analysis of complexity in \ref{app:cpx}.
-Let the resulting formula be $\varphi''$.
-%
-
-\item \textit{flatten $10^\iterm$ terms} Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $10^{\iterm}$ in $\varphi''$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $10^{\iterm}$ by $10^\ivarz$, moreover, add the constraint $\ivarz = \iterm$ as a conjunct.
-Let $\varphi'''$ denote the resulting formula.
-
-\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c | \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
-Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
-$\iterm_1 > \iterm_2$) with $\iterm_1 \le \iterm_2 - 1$ (resp.
-$\iterm_2 \le \iterm_1-1$).
-Replace ever occurrence of $\iterm_1 = \iterm_2$ wtih $\iterm_2 \le \iterm_1 \wedge \iterm_1 \le \iterm_2$.
-Let $\varphi^\dag$ the resulting formula.
-
-\item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_n$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
-$\exists \vec{\ivarz}\exists \ivarx.\ \varphi^\dag$.
-\end{enumerate}
-
-\subsection{Proof of Lemma~\ref{lem:exp-ineq}}
-
-Note that $\ell_{10}(n)$ is undefined for $n\le 0$ in $\paexp$. For convenience, we define for all $n\le 0$ that $\ell_{10}(n)=\ell_{10}(\max\{1,n\})=0$ in the following proof.
-
-\noindent {\bf Lemma~\ref{lem:exp-ineq}}.
-\emph{Let  
-%
-$$
-\begin{array}{l}
-\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-\hspace{4mm} 
-a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
-\end{array}
-$$
-%
-with $a_i \neq 0$, $A\Def \sum_{j=1}^{i-1}|a_j|$, 
-%$B\Def  \sum_{j=1}^{i}|b_j|$, 
-$B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
-and $\delta\Def  \ell_{10}(A)+3$. 
-\begin{itemize}
-    \item If $a_i > 0$, let $\alpha(\vec{\ivary}) \Def \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$, then 
-    \begin{itemize}
-        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds,
-        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)}  \ge \ivarx_{\sigma(i-1)} +\delta$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold.
-    \end{itemize}
-    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-a_i)$, then 
-    \begin{itemize}
-        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold,
-        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
-    \end{itemize}
-\end{itemize}
-}
-
-We need the following proposition for the proof of Lemma~\ref{lem:exp-ineq}.
-
-\begin{proposition} \label{prop:1}
-If $n\ge m\ge 1$ and $p \ge 2(\ell_{10}(n)- \ell_{10}(m)+1)$, then 
-$n p \le m10^p$ holds.
-\end{proposition}
-
-\begin{proof}
-
-First we show that for any $n' \in \Nat$, if $p \ge 2n'$, then $10^p \ge 10^{n'}10^{(p-n')}\ge 10^{n'}10(p-n')\ge 10^{n'}(5p+5(p-2n'))\ge  10^{n'}p$.
-
-If $ p \ge 2 (\ell_{10}(n) - \ell_{10}(m)+1)$, then  $n p \le 10\lambda_{10}(n) p = 10 * 10^{\ell_{10}(n)} p = 10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p$.
-
-Because $p \ge 2(\ell_{10}(n)-\ell_{10}(m) + 1)$, we deduce that $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  p \le 10^p$.
-Therefore, $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p \le 10^{\ell_{10}(m)} 10^p \le m 10^p$.
-We conclude that $np \le m 10^p$.
-\end{proof}
-
-%Then we give the proof for Theorem~\ref{thm:exp-ineq}.
-\begin{proof}[Proof of Lemma~\ref{lem:exp-ineq}]
-We only prove for the case $a_i > 0$, the other case is symmetric.
-
-Let $A\Def \sum_{j=1}^{i-1}|a_j|$, $B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$, and $\delta\Def  \ell_{10}(A)+3$.
-
-Suppose $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$. Moreover, let $\alpha(\vec{\ivary}) \Def  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$.
-
-Note that
- \begin{equation} 
-   A10^{-\delta} =A 10^{-\ell_{10}(A)-3} = \frac{A}{1000\lambda_{10}(A)} \le \frac{1}{100}.   \label{eq:thm-ineq-1}
- \end{equation}
-
- From $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$ and $\ivarx_{\sigma(i-1)} \ge \ldots \ge \ivarx_{\sigma(1)}$, 
-we know   
-$$-A 10^{\ivarx_{\sigma(i)} - \delta} \le  \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} \le A 10^{\ivarx_{\sigma(i)} - \delta}$$
-and
-$$-(\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)} \le \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} \le (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)}.$$
-
-Moreover, let $n = 100\sum_{j=1}^{i}|b_j|$, $m = 1$, and $p = \ivarx_{\sigma(i)}$, then 
-$n \ge m \ge 1$.
-From $2(\ell_{10}(n)- \ell_{10}(m)+1) = 2 (\ell_{10}(100 \sum_{j=1}^{i}|b_j|) + 1) = 2 (\ell_{10}(\sum_{j=1}^{i}|b_j|) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 2(\ell_{10}(n)- \ell_{10}(m)+1)$.
-Then according to Proposition~\ref{prop:1}, 
-$$100(\sum_{j=1}^{i}|b_j|)\ivarx_{\sigma(i)} = np \le m 10^p = 10^{\ivarx_{\sigma(i)}}.$$
-Thus 
-$(\sum_{j=1}^{i}|b_j|)\ivarx_{\sigma(i)}  \le \frac{1}{100} 10^{\ivarx_{\sigma(i)}}.$
-
-If $\ivarx_{\sigma(i)} \ge   \alpha(\vec{\ivary}) + 2$, then
-
-$
-\begin{array}{l}
-a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \ge \\
-a_i 10^{\ivarx_{\sigma(i)}} - A 10^{\ivarx_{\sigma(i)} - \delta} - (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)}  \ge \\
-a_i 10^{\ivarx_{\sigma(i)}} - \frac{1}{100}10^{\ivarx_{\sigma(i)}} - \frac{1}{100} 10^{\ivarx_{\sigma(i)}} = \\
-(a_i - \frac{1}{50}) 10^{\ivarx_{\sigma(i)}} \ge (a_i - \frac{1}{50}) 10^{ \alpha(\vec{\ivary}) + 2} = \\
-(a_i - \frac{1}{50}) 10^{  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i) + 2} = \\
-\frac{10(a_i - \frac{1}{50})}{10^{\ell_{10}(a_i)} } 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \frac{10(a_i - \frac{1}{50})} {a_i} 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \\
-(10 - \frac{1}{5a_i}) 10^{ \ell_{10}(\iterm(\ivary))+1} > 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \iterm(\ivary).
-\end{array}
-$
-
-Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \emph{does not} hold.
-
-If $\ivarx_{\sigma(i)} \le   \alpha(\vec{\ivary}) - 1$, then
-
-$
-\begin{array}{l}
-a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \\
-a_i 10^{\ivarx_{\sigma(i)}} + A 10^{\ivarx_{\sigma(i)} - \delta} + (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)} \le \\
-a_i 10^{\ivarx_{\sigma(i)}} + \frac{A} {10^\delta}10^{\ivarx_{\sigma(i)}} + \frac{1}{100} 10^{\ivarx_{\sigma(i)}} \le \\
-(a_i + \frac{1}{100} + \frac{1}{100})10^{\ivarx_{\sigma(i)}} = (a_i + \frac{1}{50}) 10^{\ivarx_{\sigma(i)}} \le \\
-(a_i + \frac{1}{50}) 10^{\alpha(\vec{\ivary}) - 1} = (a_i + \frac{1}{50}) 10^{\ell_{10}(\iterm(\ivary))- \ell_{10}(a_i) - 1} = \\
-\frac{a_i + \frac{1}{50}} {10^{\ell_{10}(a_i) + 1}} 10^{\ell_{10}(\iterm(\ivary))} = \frac{a_i + \frac{1}{50}} {10 \lambda_{10}(a_i)} 10^{\ell_{10}(\iterm(\ivary))} \le \\
-\frac{a_i + \frac{1}{50}} {a_i+1} 10^{\ell_{10}(\iterm(\ivary))} \le 10^{\ell_{10}(\iterm(\ivary))} \le \iterm(\ivary).
-\end{array}
-$
-
-Therefore,  in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
-
-\end{proof}
-
-We would like to remark that Lemma~\ref{lem:exp-ineq} still holds when the base of exponential function is changed to any natural number $n\ge 2$.
-
-\subsection{Elimination of exponential occurrences of variables for (in)divisibility atomic formulas}\label{app-div}
-
-Consider a divisiblilty atomic formula 
-%
-$$
-\begin{array}{l}
-\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-d\ \big |\ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
-+ \iterm(\vec{\ivary}) \big)
-\end{array}
-$$
-with $a_i \neq 0$. The indivisibility case can be treated analogously by taking a negation.
-%
-
-Let $d = 2^{r_1} 5^{r_2}  d_0$ such that $d_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $d | (10^rd_0)$.
-
-If $d_0 = 1$, then $10^r$ is divisible by $d = 2^{r_1}5^{r_2}$.
-Thus for every $n \ge r$, $d \ |\ 10^n$.
-Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to
-\[
-\small
-\begin{array}{l}
-\bigvee \limits_{p = 0}^{r - 1} d\ \big | \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
-+ \iterm(\vec{\ivary}) \big)\\
-%
-\vee \big(\ivarx_{\sigma(i)} \ge r \wedge d\ \big | \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
-+ \iterm(\vec{\ivary}) \big)\big),
-\end{array}
-\]
-where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
-
-Next, let us assume $d_0 > 1$.
-Since $10$ and $d_0$ are relatively prime, according to Euler's theorem (cf.
-\cite{HW80}), $10^{\phi(d_0)} \equiv 1 \bmod d_0$, where $\phi$ is the Euler function.
-Suppose $10^{\phi(d_0)} = kd_0 + 1$ for some $k \in \Nat$.
-Then for every $n \in \Nat$ with $n \ge r$, 
-$$
-\begin{array}{l}
-10^{n + \phi(d_0)} \bmod d =10^{n-r} 10^r (kd_0 + 1) \bmod d = \\
-10^{n-r} (k 10^rd_0 + 10^r) \bmod d = \\
-10^{n-r} (0+10^r) \bmod d = 10^n \bmod d.
-\end{array}
-$$
-
-Then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
-\[
-\begin{array}{l}
-\bigvee \limits_{p=0}^{r-1} \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[p/\ivarx_{\sigma(i)}]\ \vee \\
-\left(
-\begin{array}{l}
-\ivarx_{\sigma(i)} \ge r\ \wedge \\
-\bigvee \limits_{q = 0}^{\phi(d_0)-1} 
-\left(
-\begin{array}{l}
-\phi(d_0)\ \big |\ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
-d\ \big | 
-\left(
-\begin{array}{l}
-a_i 10^{r+q} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \\
-\sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})
-\end{array}
-\right) 
-\end{array}
-\right)
-\end{array}
-\right),
-\end{array}
-\]
-where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
 
@@
 We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_m$ is replaced by a new formula.
-Only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant coefficient $1$ is introduced, and if we substitute $x_m$ by $x_{m-1}+p$ for some $p$, coefficient of linear occurrence of $x_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $x_{m}$ and $x_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
+Only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant coefficient $1$ is introduced, and if we substitute $\ivarx_m$ by $\ivarx_{m-1}+p$ for some $p$, coefficient of linear occurrence of $\ivarx_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $\ivarx_{m}$ and $\ivarx_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
 Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le c^2$.
@@
 $${\sf space}\le q \cdot m! \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
-for some large constant $p$. \wuhao{time complexity?}.
+for some large constant $p$.
 
-\subsection{Some additional optimizations of the decision procedure for {\paexp}}\label{app-opt}
-
-\paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}
-
-Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be synchronized.
-That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, B^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$.
-Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.
-
-\paragraph{Avoid the formula-size blow-up by depth-first search}
-
-The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size.
-If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory.
-Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
-If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.
-
-\paragraph{Preprocess with small upper bound}
-
-We believe that if a quantifier-free {\paexp} formula is satisfiable, then most probably it is satisfiable with small values assigned to variables.
-Consequently, as a preprocessing step, we put a small upper bound, e.g.
-the biggest constant occurring in the formula, on the values of variables, and perform a depth-first search, so that a model can be quickly found, if there is any.
-If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.
-
-\subsection{Detailed experiment results on the STRINGHASH benchmark suite}\label{app-exp}
-
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-\begin{table}[ht]
-    \caption{Experimental results on STRINGHASH,  O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
-    $\#$: number of problems, $T$: average time in seconds}
-    \centering
-
-    \renewcommand{\arraystretch}{1.1}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-    \hline
-        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
-        \cline{3-10}
-       &  & $\#$ & $T$ &  $\#$ & $T$ & $\#$ & $T$ &  $\#$ & $T$ \\ 
-       \hline
-       \cline{1-10}
-       \multirow{3}{*}{\scriptsize{$12345(w_1)^+(w_2)^+$}} & S & 5 & 14.0 & 29 & 8.5 & 3 & {\bf  0.1} & {\bf 37} & 9.9 \\
-         \cline{2-10}
-        & U & 0 & - & 0 & - & {\bf 60} & {\bf 1.3} & {\bf 60} & 47.2 \\
-         \cline{2-10}
-        & F & 95 & - & 71 & - & 37 & - & {\bf 3} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{\scriptsize{$\begin{array}{l}12345(w_1)^+ \\
-       \hspace{2mm}(w_2)^+6789\end{array}$}} & S & 11 & 13.0 & 29 & 12.0 & 0 & - & {\bf 37} & {\bf 10.6} \\
-        \cline{2-10}
-        & U & 0 & - & 0 & - & {\bf 63} & {\bf 1.2} & {\bf 63} & 50.0 \\
-         \cline{2-10}
-        & F & 89 & - & 71 & - & 37 & - & {\bf 0} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{\scriptsize{$(w_1)^+(w_2)^+6789$}} & S & 18 & 24.0 & 30 & 9.3 & 2 & {\bf  0.1} & {\bf 41} & 16.1 \\
-      	 \cline{2-10}
-        & U & 0 & - & 1 & 4.0 & {\bf 59} & {\bf 2.5} & {\bf 59} & 45.8 \\
-         \cline{2-10}
-        & F & 82 &  & 69 &  & 39 &  & {\bf 0} &  \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}$}} & S & 82 & 8.7 & {\bf 100} & {\bf 2.2} & 28 & 5.9 & {\bf 100} & 18.5 \\
-        \cline{2-10}
-        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
-         \cline{2-10}
-        & F & 18 & - & {\bf 0} & - & 72 & - & {\bf 0} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}6789$}} & S & 60 & 9.3 & 17 & 7.8 & 3 & {\bf 0.3} & {\bf 100} & 16.0 \\
-        \cline{2-10}
-        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
-         \cline{2-10}
-        & F & 40 & - & 83 & - & 97 & - & {\bf 0} & - \\ \hline
-        \cline{1-10}
-       \multirow{3}{*}{\scriptsize{$\Sigma^*_{\sf num}6789$}} & S & 68 & 5.5 & 27 & 13.0 & 24 & {\bf 9.0} & {\bf 100} & 15.7 \\
-        \cline{2-10}
-        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
-         \cline{2-10}
-        & F & 32 & - & 73 & - & 76 & - & {\bf 0} &- \\
-        \hline
-       \end{tabular}
-           \label{table:string}
-\end{table}
+REMARK: When the original formula has no alternating quantifiers, for example, the formula of the form $\exists \ivarx_m \exists \ivarx_{m-1} \dots\exists \ivarx_1.\varphi(\ivarx_1,\dots,\ivarx_{m-1},\ivarx_m,\vec{\ivary})$.
+Since we can carry out the normalization step for all quantified variable $\ivarx_i(1\le i\le m)$ at the same time and enumerate the linear order for all introduced variables, the complexity of eliminating all quantifers is still 3-EXPSPACE.
+However, if there are  alternating quantifers at the very beginning, say the formula is of the form $\forall \ivarx_2\exists \ivarx_1.\varphi(\ivarx_1,\ivarx_2,\vec{\ivary})$.
+We have to eliminate the occurrences of $\ivarx_1$ first, treating $\ivarx_2$ as a free variable like $\vec{\ivary}$.
+When eliminating exponential occurrences of variables in inequalities (see Lemma~\ref{lem:exp-ineq}), $\ivarx_2$ will be collected in term $\iterm(\ivary)$.
+Since $\iterm(\ivary)$ may change due to the  substitutions of $\ivarx_1$, when we carry out normalization step for $\ivarx_2$, the number of introduced fresh variables is not bounded by the length of original formula.
+Hence, we conjecture that the overall complexity for eliminating all quantifiers in a formula will be non-elementary.

### 2021-09-29 FMCAD2021/7_impl_exp.tex
complexity add a remark

--- 2021-05-22 FMCAD2021/6_impl_exp.tex
@@
 
-\vspace*{-1mm}
+\vspace*{-2mm}
 \subsection{Implementation}
-We implemented the decision procedure in Wolfram Mathematica, and obtain a solver, called the {\paexp}-solver, which is able to solve the satisfiability of quantifier-free {\paexp} formulas.
+We implemented the decision procedure in Wolfram Mathematica, and obtain a solver, called the {\paexp}-solver, which is able to solve the satisfiability of {\paexp} formulas.
 
@@
 
-To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC and STRINGHASH.
+To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC and STRINGHASH\footnote{The benchmarks are available at https://github.com/EcstasyH/PAexp-Solver/tree/main/Benchmark}.
 %\begin{itemize}

### 2021-09-29 FMCAD2021/8_conclusion.tex
complexity add a remark

--- 2021-05-22 FMCAD2021/7_conclusion.tex
@@
 The experiment results show the efficacy of our implementation, compared with the state-of-the-art solvers.
-An interesting open question is a refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.
+\hide{An interesting open question is a refined complexity analysis of Point's procedure, considering the fact that Cooper's quantifier elimination procedure for Presburger arithmetic is triply exponential \cite{Oppen73}.
+}

### 2021-09-29 FMCAD2021/9_appendix.tex
complexity add a remark

--- 2021-09-18 FMCAD2021/8_appendix.tex
@@
 
-\subsection{Complexity analysis}\label{app:cpx}
-
-In this part, we analyse the complexity of the quantifier elimination procedure applied on a formula with one existential quantifer, that is, a formula of the form $\exists \ivarx.
-\varphi(\ivarx, \vec{\ivary})$.
-We will prove that the complexity of eliminating one quantifer has a 3-EXPSPACE upper bound.
-Since the quantifier elimination procedure for $\pa$ works as a subprocedure (in eliminating linear occurrences of variables) and also has a 3-EXPSPACE complexity, the bound may not be easily improved.
-
-We first analyse the changes of formula length in the normalization step.
-For the rest of the decision procedure, we adopt the style in the complexity analysis of \pa \ (cf.
-\cite{Oppen73}).
-The idea is that the upper bound of the formula length can be expressed using the product of the number of atoms, the number of coefficients and the length of the maximum constant.
-So we track the changes of such quantities throughout the procedure.
-
-Given a formula $\exists \ivarx.
-\varphi(\ivarx, \vec{\ivary})$ with length $n$.
-After the normalization step, suppose we obtain a formula with $m$ quantified variables $\exists \ivarx_1 \dots\exists \ivarx_m.
-\varphi'(\ivarx_1,\dots, \ivarx_m, \vec{\ivary})$.
-We show that the length of the new formula is at most $10n$ and the number of quantified variables $m\le n$.
-In each sub-step of normalization, we analyse the worst situation:
-(1) \textit{NNF transformation} suppose all atoms are of the form $\iterm_1=\iterm_2$, then taking negation will double the number of atoms, the length increases to $2n$ at most. 
-(2) \textit{replace $\ell_{10}(\iterm)$ terms} the original formula has at most $n$ terms of the form $\ell_{10}(\iterm)$, for each of them, we introduce a fresh variable and two conjuncts.
-The formula length increases to $4n$.
-(3) \textit{flatten $10^\iterm$ terms} similar to (2), the formula has at most $n$ such terms and for each term, a fresh variable and a conjunct are added. The formula length increases to $5n$. 
-(4) \textit{$\le$ transformation} similar to (1), here we assume all atoms are of the form $\iterm_1  = \iterm_2$, so the length of the formula with increase to $10n$ at most.
-
-The analysis for the normalization step is coarse, for example, the worst case in (1)\&(4) or (2)\&(3) can not happen at the same time.
-However, what we need is that the increased length of the formula is bounded by a constant factor,i.e.
-10.
-In addition, we can also conclude that the number of fresh variables, $m$, is less than $n$ since there are at most $n$ different forms of terms.
-
-After the normalization step, denote the obtained formula by $\exists \vec{\ivarx}.
-\varphi(\vec{\ivarx},\vec{\ivary})$ with quantified variables $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$ and free variables $\vec{\ivary}$.
-$\varphi$ contains no quantifiers.
-Let $n'$ denote the length of the formula.
-
-According to our quantifier elimination procedure,we eliminate the quantified variables one by one: each time select the largest variable among $\ivarx_i,1\le i\le m$, first eliminate the exponential occurrences, then linear occurrences.
-The linear order of quantified variables are given by a permutation.
-For $m$ quantified variables, there are $m!$ possible linear orders.
-
-Suppose the specified linear order is $\ivarx_m\ge \dots \ge \ivarx_1$.
-Let $\exists \ivarx_1\dots \exists\ivarx_{m-k}.\varphi_k$ denote the formula obtained from $\exists \vec{\ivarx}.
-\varphi(\vec{\ivarx},\vec{\ivary})$ by eliminating quantifiers $\exists \ivarx_m,\dots,\exists\ivarx_{m-k+1}$.
-Let $\exists \ivarx_1 \dots \exists \ivarx_{m-k+1}.\varphi'_k$ denote the formula obtained from $\exists \ivarx_1\dots \exists\ivarx_{m-k+1}.\varphi_{k-1}$ by eliminating linear occurrences of $\ivarx_{m-k+1}$.
-Let $\varphi_0$ denote $\varphi$.
-
-Let $c_k$ be the number of distinct $d$ in atoms of the form $d \ \big | \ \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of quantified variables in $\varphi_k$.
-Let $s_k$ be the largest constant (including coefficients) and $a_k$ be the number of atomic formulas in $\varphi_k$.
-Similarly we define $c'_k$, $s'_k$ and $a'_k$ for $\varphi'_k$.
-Let $c,s,a$ be $c_0,s_0,a_0$ respectively.
-
-First we analyse the sub-procedure to eliminate exponential occurrences of $\ivarx_m$. We prove the following lemma.
-
-\begin{lemma}\label{lem:cpx exp}
-    \begin{align}
-        c'_1&\le c^2 \notag \\
-        s'_1&\le ms^2 \notag\\
-        a'_1&\le sa\notag 
-    \end{align} 
-\end{lemma}
-
-\begin{proof}
-
-The analysis is separated into 2 cases where all atoms are of the same form (inequalities atoms or divisiblilty atoms).
-
-If all atoms are inequalities atomic formulas of the form in Lemma~\ref{lem:exp-ineq}.
-We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_m$ is replaced by a new formula.
-Only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant coefficient $1$ is introduced, and if we substitute $x_m$ by $x_{m-1}+p$ for some $p$, coefficient of linear occurrence of $x_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $x_{m}$ and $x_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
-Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le c^2$.
-Note that $\delta$ and $B$ are at most $\ell_{10}(ms)$, when we substitute $\ivarx_m$ by $\ivarx_{m-1}+\delta-1$ or by $B-1$, the largest constant in the formula is at most $s\cdot  10^{\ell_{10}(ms)}\le ms^2$.
-And an inequality is replaced by at most $4\ell_{10}(ms)$ atomic formulas, so $a'_1\le 4\ell_{10}(ms)a$.
-
-If all atoms are divisiblilty atomic formulas of the form $d \ \big | \ \iterm$ or $d \ \nmid \ \iterm$.
-We have $c'_1<2c$ since a divisibility atomic formula  will produce at most two forms of atomic formulas $d \ \big | \ \iterm$ and $\phi(d) \ \big | \ \iterm$.
-Note that any constant in $\iterm$ , say $l$, can be replaced by ($l \bmod d)$, so $s'_1\le s$.
-When $d$ is a large prime number, $\phi(d)=d-1$, a divisibility atomic formula is replaced by roughly $d$ atomic formulas, so $a'_1\le sa$.
-
-Choose larger upper bound for $c'_1$, $s'_1$ and $a'_1$ respectively, then the lemma is proved.
-
-\end{proof}
-
-Since linear occurrences of $\ivarx_m$ are eliminated using Cooper's algorithm, we combine Oppen's analysis for this sub-procedure:
-
-\begin{lemma}\label{lem:cpx oppen}
-    \begin{align}
-        c_1&\le c'^4 \notag \\
-        s_1&\le s'^{4c'} \notag\\
-        a_1&\le {a'}^4 s'^{2c'}\notag 
-    \end{align} 
-\end{lemma}
-
-From Lemma~\ref{lem:cpx exp} and Lemma~\ref{lem:cpx oppen}, we have
-\begin{align}
-    c_1&\le c^8 \notag \\
-    s_1&\le (ms^2)^{4c^2} \notag\\
-    a_1&\le (sa)^4(ms^2)^{2c^2}\notag 
-\end{align}
-
-Assuming $m\le n$, by induction on $k$ we get
-
-\begin{lemma}
-    \begin{align}
-        c_k&\le c^{8^k} \notag \\
-        s_k&\le n^{(4c)^{ 8^k}} s^{(8c)^{ 8^k}} \notag\\
-        a_k&\le a^{4^k}n^{(4c)^{ 8^k}} s^{(8c)^{ 8^k}} \notag
-    \end{align} 
-\end{lemma}
-
-If we assume $c\le n$, $a\le n$ and $s\le n$, the space required to store the quantifier free formula, $\varphi_k$, is bounded by the product of the number of linear orders $m!$, the number of atoms $a_k$, the maximum number of constants $2m+2$ per atom, the maximum amount of space $s_k$ to store each constant and some constant $q$:
-
-$${\sf space}\le q \cdot m! \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
-for some large constant $p$. \wuhao{time complexity?}.
-
 \subsection{Some additional optimizations of the decision procedure for {\paexp}}\label{app-opt}

### 2021-09-29 FMCAD2021/paper.tex
complexity add a remark

--- 2021-05-24 FMCAD2021/paper.tex
@@
 The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-\wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
+The complexity of the quantifier elimination procedure is analyzed for the first time and we show that the elimination of one quantifier is 3-EXPSPACE.
+While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
     We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.
@@
 \vspace*{-4mm}
+
 \section{Decision procedure for {\paexp}}\label{sec-dec}
+\input{4_dec_proc.tex}
 
-\input{4_dec_proc.tex}
+\vspace*{-3mm}
+
+\section{Complexity Analysis}\label{sec-cpx}
+\input{5_complexity.tex}
 
@@
 
-\input{5_optimization.tex}
+\input{6_optimization.tex}
 
@@
 
-\input{6_impl_exp.tex}
+\input{7_impl_exp.tex}
 
@@
 
-\input{7_conclusion.tex}
+\input{8_conclusion.tex}
 
@@
 
-\input{8_appendix.tex}
+\input{9_appendix.tex}
 

### 2021-10-25 FMCAD2021/1_intro.tex
dodify

@@
 The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string-integer conversion constraints to the above class.
-The paper proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
+It proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
 For string-integer conversion, their approach projects the solution to string-integer conversion to
@@
 
-In this paper, we propose a complete solution to this problem.
+In this paper, we give a complete solution to this problem.
 We propose a decision procedure for the class of word equation + flat languages + length constraints + string-integer conversion.
-The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$); then, according to the decidability of $\paexp$, we obtain the decidability of the original satisfiability problem.
+The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$), \wuhao{more precisely, the existential fragment of $\paexp$}; then, according to the decidability of $\paexp$, we obtain the decidability of the original satisfiability problem.
 
@@
 This can be seen as another contribution of this paper.
-We also analyze the deterministic upper bound of the complexity for the first time, the space complexity of eliminating one quantifier is shown to be 3-EXPSPACE.
+\wuhao{We also analyze the deterministic upper bound of complexity for the first time: given a formula with only existential quantifiers, the complexity is 3-EXPSPACE to eliminate all quantifiers}.
 Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).
@@
 \paragraph*{Structure}
-After the preliminaries in Section~\ref{sec:pre}, we describe how to flatten a string constraint with string-integer conversion to an $\paexp$ formula in Section~\ref{sec:string-solving}.
+After the preliminaries in Section~\ref{sec:pre}, we describe how to flatten a string constraint with string-integer conversion to an existential $\paexp$ formula in Section~\ref{sec:string-solving}.
 We present the quantifier elimination procedure for $\paexp$ in Section~\ref{sec-dec} and its optimizations in Section~\ref{sec-opt}.

### 2021-10-25 FMCAD2021/3_flatten.tex
dodify

@@
 \begin{theorem} \label{thm:string-parInt}
-	The satisfiability of  $\strint$ under flat domain restrictions  is decidable. 
+    \wuhao{The satisfiability of  $\strint$ under flat domain restrictions is equivalent to the satisfiability of existential $\paexp$ formulas, and thus is decidable. }
 \end{theorem}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-Now, we introduce a new form of atomic string constraints: 
-a {\parseInt} constraint $\phi$ is of the form 
-$n \sim {\parseInt}(t)$,
-where $n$ is an integer term, $\sim \in \{\le,<,=,>,\ge\}$ and $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$ is a string term.
-$\lVert \phi\rVert \Def  \{I \mid I(n)\sim {\parseInt}(I(t))\}$.
-
-In what follows, we only  consider the problem in the case when $t=x$ and $x$ is restricted to (generic) flat languages. 
-%If $t$ is not a single variable, 
-For the general case $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$, 
-it can be reduced to this special case by induction on the structure of $t$. 
-%we can reduce the problem by separating $t$ (corresponding to the $|w|>2$ case in definition).
-
-Given an $\alpha$-flat language $L$,
-we assume $\alpha=\langle p,q \rangle$ and $L=(w_1)^*...(w_q)^*$,
-where $p,q$ and $w_i(1\le i \le q) $ are known.
-We further assume that $x = (w_1)^{\beta_1} ... (w_q)^{\beta_q}$,
-then we have
-\begin{align}
-    &{\parseInt}(x) \notag \\
-    =&{\parseInt}((w_1)^{\beta_1} ... (w_q)^{\beta_q}) \notag \\
-    =&{\parseInt}((w_1)^{\beta_1} ... (w_{q-1})^{\beta_{q-1}})\cdot 2^{\beta_q |w_q|}
-    + {\parseInt}((w_q)^{\beta_q}) \label{parse}
-\end{align}
-(\ref{parse}) is a recursive expression.
-So we only need to deal with the basic case ${\parseInt}((w_q)^{\beta_q})$, 
-where $w_q \neq \epsilon$
-\begin{align}
-    {\parseInt}((w_q)^{\beta_q}) &= \sum_{i=0}^{\beta_q-1} {\parseInt}(w_q)\cdot 2^{|w_q|\cdot i}  \notag \\
-    &={\parseInt}(w_q)\frac{2^{|w_q|\cdot \beta_q}-1}{2^{|w_q|}-1}
-    \label{parseInt-basic}
-\end{align}
-In (\ref{parseInt-basic}), since $w_q$ and $|w_q|$ are known, 
-they can be regarded as constants.
-The only unknown variable is $\beta_q$.
-
-Combine (\ref{parse}) and (\ref{parseInt-basic}),
-the constraint $n={\parseInt}(x)$ can be expressed by an arithmetic expression with 
-$n$ and $(\beta_1,...,\beta_q)$,
-inevitably with exponential components.
-
-Observe the form of the above equation,
-$a,b,n$ are integer variables and either occur in 
-an exponential term or a linear term.
-This is always the case,
-so the problem can be reduced to 
-the decidability of PA with exponential function.
-
-When $x$ of {\parseInt} constraints is restricted to the
-generic $\alpha$-flat language ($\alpha$ is fixed),
-the (\ref{parse}) and (\ref{parseInt-basic}) still hold 
-but $w_i(1\le i\le q)$ is known.
-However,
-by definition,
-the generic $\alpha$-flat language is the finite union of all $\alpha$-flat languages,
-so we can enumerate all possible values for $w_i$.
-In this way,
-the problem can still be reduced to the decidability of PA with exponential function ($\paexp$), i.e.,
-\begin{theorem} \label{thm:string-parInt}
-If {$\paexp$} is decidable, then the satisfiability (validity) of string constraints with {\parseInt} in which all string variables 
-are ranged over flat strings is decidable. 
-\end{theorem}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%The correctness of the flattening approach is stated below.
-%\begin{proposition}
-%$\llbracket \varphi \rrbracket_{\flatfun_\varphi}$ is nonempty iff $\flatten_{\flatfun_\varphi}(\varphi)$ is satisfiable.
-%\end{proposition}

### 2021-10-25 FMCAD2021/4_dec_proc.tex
dodify

@@
 Moreover, the procedure presented here slightly improves Point's procedure, in the following two aspects: 1) DNF (disjunctive normal form) was required in Point's procedure, which is not required here, 2) in Point's procedure, the divisibility constraints produced by the elimination of linear occurrences of a variable should be converted to equality constraints (by introducing fresh variables) before the elimination of exponential occurrences of some other variable, which is unnecessary here, since the divisibility constraints are directly dealt with in the elimination of exponential occurrences of variables.
-Furthermore, we give a complexity analysis in Section~\ref{sec-cpx}, which shows that the elimination of one quantifier has a space complexity of 3-EXPSPACE. 
 %
 
-However, Point's procedure is quite expensive and a faithful implementation would not scale\footnote{We did implement Point's algorithm and discovered that the implementation could only solve formulas of very small size.}.
-In the next section, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
+However, in the next section we analyse the complexity of Point's procedure to be 3-EXPSPACE, which  is quite expensive and a faithful implementation would not scale\footnote{We did implement Point's algorithm and discovered that the implementation could only solve formulas of very small size.}.
+So in Section~\ref{sec-opt}, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
 
@@
 
+The normalization step comprises the following sub-steps.
+\begin{enumerate}
+\item \textit{NNF transformation} At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form). Moreover, we remove the occurrences of $\neg$ by replacing 
+\hide{(a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, }
+(a) $\neg c | \iterm$ with $ c \nmid \iterm$, 
+(b) $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, 
+(c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, 
+(d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
+%
+\item \textit{replace $\ell_{10}(\iterm)$ terms} Repeat the following procedure, until there are no $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: for each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct.
+Note that if $\iterm$ contains no variables, then $\ell_{10}(\iterm)$ is a constant.
+In this case, we can also assume that $\iterm$ contains $\ivarx$ and perform the same replacements, which helps in the analysis of complexity in \ref{app:cpx}.
+Let the resulting formula be $\varphi''$.
+%
+
+\item \textit{flatten $10^\iterm$ terms} Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $10^{\iterm}$ in $\varphi''$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $10^{\iterm}$ by $10^\ivarz$, moreover, add the constraint $\ivarz = \iterm$ as a conjunct.
+Let $\varphi'''$ denote the resulting formula.
+
+\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c | \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
+$\iterm_1 > \iterm_2$) with $\iterm_1 \le \iterm_2 - 1$ (resp.
+$\iterm_2 \le \iterm_1-1$).
+Replace ever occurrence of $\iterm_1 = \iterm_2$ wtih $\iterm_2 \le \iterm_1 \wedge \iterm_1 \le \iterm_2$.
+Let $\varphi^\dag$ the resulting formula.
+
+\item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_n$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
+$\exists \vec{\ivarz}\exists \ivarx.\ \varphi^\dag$.
+\end{enumerate}
+
+Intuitively, the normalization step first absorbs all negation operators by transforming the formula into NNF.
+Then it removes the occurrences of $\ell_{10}(\iterm)$ where $\ivarx$ occurs in $\iterm$, by encoding them with the exponential function.
+Moreover, for each occurrence of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, it introduces a fresh variable $\ivarz$, replaces $10^\iterm$ with $10^\ivarz$, and adds the equality $\ivarz = \iterm$.
+All equalities and inequalities will be rewritten into the form $\iterm_1\le\iterm_2$.
+Finally, add quantifiers for the introduced fresh variables.
+
+After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
+Denote the negation of $c | \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 The normalization step first removes the occurrences of $\ell_{10}(\iterm)$ where $\ivarx$ occurs in $\iterm$, by encoding them with the exponential function.
@@
 More details can be found in Appendix~\ref{app-norm}.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
     For convenience, let us call these formulas as inequality respectively (in)divisibility atomic formulas.
@@
 The elimination of the exponential occurrences of the (in)divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.
-
-Let us consider 
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Inequality atoms}
+In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these inequality atomic formulas. Let us consider 
 $$
@@
 
-\hide{
-\medskip
-
-\paragraph{Divisibility atomic formulas} Consider
+\subsubsection{Divisibility atoms}
+Consider a divisiblilty atomic formula 
 %
@@
 $$
-with $a_i \neq 0$.
+with $a_i \neq 0$. The indivisibility case can be treated analogously.
 %
 
-Let $c = 2^{r_1} 5^{r_2}  c_0$ such that $c_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $c | (10^rc_0)$.
+Let $d = 2^{r_1} 5^{r_2}  d_0$ such that $d_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $d | (10^rd_0)$.
 
-If $c_0 = 1$, then $10^r$ is divisible by $c = 2^{r_1}5^{r_2}$.
-Thus for every $n \ge r$, $c \ |\ 10^n$.
+If $d_0 = 1$, then $10^r$ is divisible by $d = 2^{r_1}5^{r_2}$.
+Thus for every $n \ge r$, $d \ |\ 10^n$.
 Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to
@@
 where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
-}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-Let $d = 2^rd_0$, $d_0$ is an odd natural number. 
-According to Euler's Theorem, 
-$gcd(2,d_0)=1$ implies
-$2^{\phi(d_0)} \text{mod}\ d_0 = 1$,
-where $\phi$ is the Euler function.
-Consider function $f(n)=(2^n\ \text{mod} \ d)$,
-when $n\ge r$, 
-$f(n)$ becomes a periodic function and its period is a divisor of $\phi(d_0)$ because
-\begin{align}
-    f(n+\phi(d_0)) 
-    &=2^{n+\phi(d_0)}\ \text{mod}\ d &\notag\\
-    &=2^n\cdot 2^{\phi(d_0)}\ \text{mod}\ d &\notag\\
-    &=2^n\cdot (k d_0 + 1) \ \text{mod}\ d &
-    (\mbox{assume } 2^{\phi(d_0)}=k d_0 +1) \notag \\
-    &= 2^n \ \text{mod}\ d &
-    (\mbox{when } n\ge r, 2^n d_0\ \text{mod}\ d = 0 )\notag\\
-    &= f(n) \notag
-\end{align}
-
-When $x_i\le r-1$, we just enumerate all possible value of $x_i$, i.e., 
-$$\rho_5 \Def \bigvee_{p=0}^{r-1} \tau(p,\bar{x},\bar{y})\,.$$
-When $x_i\ge r$, $\tau(x_i,\bar{x},\bar{y})$ is equivalent to
-$$
-\rho_6 \Def \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
-+  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r] \,.
-$$
-            Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.
-
-\textbf{Elim-exp} is the most technical part of our QE algorithm.
-As mentioned, \textbf{Elim-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as inputs, and outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
-The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{Elim-linear}.
-For simplicity, we will abuse notations $x_i$ for $x_{\sigma(i)}$, and $\bar{x}$ for $(x_{\sigma(0)},...,x_{\sigma(i-1)})$ ( remind that $x_{\sigma(i+1)},...,x_n$ have been eliminated already).
-
-After normalization, the formula $\theta(\bar{x},\bar{y})$ contains atoms of three forms corresponding to 
-the predicates $\le$, $|$ and negation of $|$, respectively.
-The problem will be discussed in two cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
-corresponding to \textbf{Elim-exp-ineq} or \textbf{Elim-exp-div}.
-
-%We first give the whole procedure of \textbf{Elim-exp},and then provide the details of the two sub-routines.
-
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{$x_i$ and $\theta$, $x_i$ is larger than $x_0$ to $x_{i-1}$
-    and occurs exponentially in $\theta$} 
-    \KwOut{a formula equivalent to $\theta$ where $x_i$ occurs linearly}
-
-    \While{there is an atom $\tau(x_i,\bar{x},\bar{y})$ contains $2^{x_i}$}
-    {
-        $\Psi:= \textit{False}$\;
-        \eIf{$\tau$ is of the form $a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$}
-        { 
-            \tcp{ \textbf{Elim-exp-ineq} case}
-            $A :=  \sum_{j=0}^{i-1}|a_j|$, $B:= b_i + \sum_{j=0}^{i-1}|b_j|$\;
-            $B':= 2(l_2(B)+3)$, $\delta:=  l_2(A)+3$\;
-            If $a_i>0$ then $\alpha := l_2(t(y))-l_2(a_i)$
-            else $\alpha := l_2(-t(y))-l_2(-a_i)$\;
-
-            $\rho_1 :=  (x_i \le \alpha -1 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$
-
-            $\quad \vee ( x_i \le \alpha -1 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta} (x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))$\;
-
-            $\rho_2 :=  (x_i = \alpha) \wedge \tau[\alpha/x_i]$\;
-            $\rho_3 :=  (x_i = \alpha+1) \wedge \tau[\alpha+1/x_i]$\;
-            $\rho_4 :=  (x_i \ge \alpha+2 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$
-
-            $\quad \vee (x_i \ge \alpha +2 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta}(x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))
-            $\;
-            \eIf{$a_i>0$}
-            {
-                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
-                \vee [x_i \le \alpha -1 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$
-            }
-            {
-                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
-                \vee [x_i \ge \alpha +2 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$    
-            }
-            }
-        {
-            \tcp{the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
-            \tcp{ \textbf{Elim-exp-div} case}
-            \tcp{ let $d=2^r d_0$ where $d_0$ is odd}
-            $\rho_5 := \bigvee_{p=0}^{r-1} [\tau(p,\bar{x},\bar{y})\wedge x=p]$\;
-            $\rho_6 := \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
-            +  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r]
-            $\;
-            $\Psi :=  \rho_5 \vee \rho_6$
-        }
-        replace $\tau$ by $\Psi$ in $\theta$\;
-    }
-    \caption{Elim-exp}
-\end{algorithm}
-
-\subsubsection{Elim-exp-ineq}
-
-This part deals with the case when $\tau$ is an inequality of the form 
-$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
-
-We now try to eliminate $2^{x_i}$ in $\tau$, the idea is to find a bound for $x_i$, 
-either by constants or by other variables.
-We will prove the following theorem.
-
-\begin{theorem} \label{thm:exp-ineq}
-Assume that $x_i$ is the maximal variable in $\{x_0,...,x_{i}\}$ according to the given order, $\bar{x}$ denotes $\{x_0,...,x_{i-1}\}$.
-Given an inequality $\tau(x_i,\bar{x},\bar{y})$ of the form  
-$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
-with $a_i \neq 0$. Define $A\Def \sum_{j=0}^{i-1}|a_j|$, 
-$B\Def |b_i| + \sum_{j=0}^{i-1}|b_j|$, 
-$B'\Def 2(l_2(B)+3)$,
-$\delta\Def  l_2(A)+3$, then 
-\begin{itemize}
-    \item if $a_i > 0$, let $\alpha \Def l_2(t(y))-l_2(a_i)$
-    \begin{itemize}
-        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
-        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
-    \end{itemize}
-    \item if $a_i < 0$, let $\alpha \Def l_2(-t(y))-l_2(-(a_i))$
-    \begin{itemize}
-        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
-        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
-    \end{itemize}
-\end{itemize}
-\end{theorem}
-
-Before proving \textbf{Theorem 1}, we need  the following lemma to estimate linear terms.
-
-\begin{lemma} \label{lem:1}
-For any $n,m\in \mathbb{N}$, if $n\ge m\ge 1$ and $x\ge 2(l_2(n)-l_2(m)+1)$, then 
-$nx\le m2^x$ holds.
-\end{lemma}
-
-\begin{proof}
-First we can prove that for any $N\in \mathbb{N}$, $x\ge 2N \implies 2^N x\le 2^x$.
-Let $N\Def l_2(n)-l_2(m)+1$, then we have $x\ge  2N \implies n x\le 2\lambda_2(n) x\le 2^N \lambda_2(m) x \le m 2^x $.
-\qed
-\end{proof}
-
-Then we give the proof for Theorem~\ref{thm:exp-ineq}.
-\begin{proof}
-We only prove for the $a_i > 0$ case, the other case is analogous. 
-The goal is to find a bound for $x_i$ such that 
-the values of the atoms containing $x_i$ keep constant when $x_i$ is greater than the bound.
-
-Note that $x_i$ is the largest among $\bar{x}$. 
-suppose $x_i > x_{i-1} + \delta$, and let 
- $\delta \Def  l_2(A)+3$, then 
- \begin{equation} 
-   2^{-\delta}A = \frac{A}{8\lambda_2(A)}\le \frac{1}{4}.   \label{eq:thm-ineq-1}
- \end{equation}
- When $x_i \ge B' = 2(l_2(4B)-l_2(1)+1)$, according to Lemma~\ref{lem:1}, 
-\begin{equation} 
-    4 B x_i \le 2^{x_i}. \label{eq:thm-ineq-2}
-\end{equation}
-
-When $x_i \ge \alpha+2$,
-\begin{align}
-    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
-    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
-     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
-    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
-    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
-    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
-      (\mbox{ Def. of } \alpha) & \notag \\
-     & \ge  t(\bar{y}) & \notag 
-\end{align}
-So we conclude that %when $x_i \ge \alpha+2,x_i\ge B',x_i \ge x_{i-1}+\delta$,
-$\tau(x_i,\bar{x},\bar{y})$ keeps \textit{false} in this case.
-
-When $x_i \le \alpha-1$, similarly we have
-\begin{align}
-    \tau(x_i,\bar{x},\bar{y})\notag 
-    & \le  a_i 2^{x_i} + 2^{x_i -\delta}A  + Bx_i &  \notag \\
-    & \le  2^{x_i}(a_i +\frac{1}{4} +\frac{1}{4}) &(\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}}) \notag \\
-   & \le  2\lambda_2(a_i)\frac{\lambda_2(t(\bar{y}))}{2\lambda_2(a_i)} &  (\mbox{ Def. of } \alpha) \notag \\
-   &  \le  t(\bar{y}) & \notag 
-\end{align}
-which indicates that when  $x_i \le \alpha-1$, $\tau(x_i,\bar{x},\bar{y})$ keeps \textit{true}. \qed 
-\end{proof}
-
-\subsubsection{Elim-exp-div}
-
-If $x_i$ occurs exponentially in an atom $\tau(x_i,\bar{x},\bar{y})$ of the form
-$$\tau(x_i,\bar{x},\bar{y}) \Def  d| a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k+t(\bar{y}) \quad (a_i \neq 0)\,,$$
-the algorithm \textbf{Elim-exp-div} outputs an equivalent formula without $2^{x_i}$ terms.
-The idea is that $a_i 2^x$ modulo $d$ is a periodic function when $x$ is large enough and the period can be computed.
-
-Let $d = 2^rd_0$, $d_0$ is an odd natural number. 
-According to Euler's Theorem, 
-$gcd(2,d_0)=1$ implies
-$2^{\phi(d_0)} \text{mod}\ d_0 = 1$,
-where $\phi$ is the Euler function.
-Consider function $f(n)=(2^n\ \text{mod} \ d)$,
-when $n\ge r$, 
-$f(n)$ becomes a periodic function and its period is a divisor of $\phi(d_0)$ because
-\begin{align}
-    f(n+\phi(d_0)) 
-    &=2^{n+\phi(d_0)}\ \text{mod}\ d &\notag\\
-    &=2^n\cdot 2^{\phi(d_0)}\ \text{mod}\ d &\notag\\
-    &=2^n\cdot (k d_0 + 1) \ \text{mod}\ d &
-    (\mbox{assume } 2^{\phi(d_0)}=k d_0 +1) \notag \\
-    &= 2^n \ \text{mod}\ d &
-    (\mbox{when } n\ge r, 2^n d_0\ \text{mod}\ d = 0 )\notag\\
-    &= f(n) \notag
-\end{align}
-
-When $x_i\le r-1$, we just enumerate all possible value of $x_i$, i.e., 
-$$\rho_5 \Def \bigvee_{p=0}^{r-1} \tau(p,\bar{x},\bar{y})\,.$$
-When $x_i\ge r$, $\tau(x_i,\bar{x},\bar{y})$ is equivalent to
-$$
-\rho_6 \Def \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
-+  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r] \,.
-$$
-            Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\znj{\textbf{Solved}.More explanation or a formal proof is needed. }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-\subsection{Elim-linear}
-
-After \textbf{Elim-exp}, 
-we obtain a formula (still denoted by $\theta(x_i,\bar{x},\bar{y})$) without $2^{x_i}$ terms,
-i.e., $x_i$ occurs only linearly.
-Now we wish to construct a formula without $x_i$ equivalent to $\exists x_i.\theta(x_i,\bar{x},\bar{y})$.
-This can be done by 
-following Cooper's QE algorithm for PA,
-treating all $x_j (j<i)$ as free variables (like $\bar{y}$).
-The procedure contains 3 steps.
-
-Step 1: Put $\theta(x_i,\bar{x},\bar{y})$ in NNF form, replace atomic formulas containing symbols other than $\le,|$ by equivalent formulas only with $\le$, that is, $x=a$ by $x\le a \wedge a\le x$, $x< a$ by $x\le a+1$, $x\neq a$ by $x>a \vee x<a$,  $x\geq a$ by $-x \le -a$, and $x>a$ by $-x \le -a-1$.
-In inequality atoms, collect terms of $x_i$
-to one side and guarantee the coefficients of $x_i$ are positive.
-
-Step 2: Let $d$ be the least common multiple of all coefficients of $x_i$. 
-For atoms with $x_i$, multiply all terms by a factor so that the coefficient of $x_i$ is $d$.
-We introduce a fresh variable $x_i'$, replace all occurrence of $dx_i$ by $x_i'$,
-and denote the resulted formula by $\theta'(x_i')$.
-Note that $x_i'$ is a multiple of $d$, 
-so we set $\theta'=\theta'\wedge d|x'$.
-
-We classify all atoms in $\theta$ into the following three sets:  
-$L$ denotes all atoms of the form $t_l(\bar{x},\bar{y})\le x_i'$,
-$U$ denotes all atoms of the form $x_i'\le t_u(\bar{x},\bar{y})$,
-$M$ denotes all atoms of the form $k_m|x_i'+t_m(\bar{x},\bar{y})$ and their negations,
-where $t(\bar{x},\bar{y})$ with a subscript is any term of $\bar{x}$ and $\bar{y}$ and 
-$k_m \in \mathbb{N}$ for $m\in M$. % are integers in $$. 
-%Then all atoms with $x_i'$ are of one of the types $L,U,M$.
-
-Step 3: Let $\delta$ be the least common multiple of $\{k_m \mid m\in M\}$.
-Construct $\theta'_{+\infty}(x_i')$ from $\theta'(x_i')$ by  
- replacing $\textit{true}$ for all atoms in $L$,
-and $\textit{false}$ for all atoms in  $U$.
-%We use $\theta'[t]$ to denote replacing all $x_i'$ by a term $t$. 
-Then $\exists x_i'.\theta'(x_i')$
-is equivalent to
-
-$$\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee 
-\bigvee_{j=0}^{\delta-1} \bigvee_{u\in U} \theta'[t_u(\bar{x},\bar{y})-j/x_i']$$
-
-The first disjunction $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i']$ corresponds to the case where $x_i'$ is large enough 
-so that all atoms in $L$ are \textit{true} and all atoms in  $U$ are \textit{false}. Thus $\theta'_{+\infty}$  contains only divisibility literals.
-If there is a number $n (0\le n \le \delta-1)$ such that $ \theta'_{+\infty}[n/x_i']$ is evaluated to be \textit{true},
-then for every $\lambda\in \mathbb{N}$,
-$\theta'_{+\infty}[n+\lambda\delta/x_i']$ is also evaluated to be \textit{true}.
-Hence, there exists $x_i'$ large enough to satisfy $\theta'$.
-
-The second disjunction corresponds to the case that for 
-some $u\in U, x_i'\le t_u(\bar{x},\bar{y})$ holds.
-In this case, 
-select the minimal $t_u(\bar{x},\bar{y})$
-from atoms in $U$ that holds,
-then there exists a solution for $x_i'$ such that $x_i'$ is in the interval $[t_u(\bar{x},\bar{y})-\delta+1,t_u(\bar{x},\bar{y})]$.
-
-Here we describe Cooper's algorithm in pseudo code.
-We will use $\theta\{\tau'/\tau\}$ to denote substitute all atoms $\tau$ with $\tau'$,
-distinguished from term substitution $\theta[t'/t]$.
-
-\begin{algorithm}[t]
-    \SetAlgoLined
-    \KwIn{$x_i,\theta(x_i,\bar{x},\bar{y})$,
-    where $x_i$ occurs linearly} 
-    \KwOut{A equivalent formula without $x_i$}
-
-    Collect terms of $x_i$ in each atom\;
-    Let $d$ to be the least common multiple of all coefficients of $x_i$\;
-    Multiply each atom by a factor so that coefficient of $x_i $ is $d$\;
-    Let $\theta'(x_i'):= \theta[x_i'/dx_i]$\;
-    $\theta'(x_i') := \theta'(x_i') \wedge d| x_i'$\;
-    \tcp{ Now the atoms have 3 different forms}
-    \tcp{ $t_l(\bar{x},\bar{y})\le x_i'$, $x_i' \le t_u(\bar{x},\bar{y})$, $k_m|x_i'+t_m(\bar{x},\bar{y})$(or its negation)}
-    \tcp{ where $l\in L,u\in U,m\in M$ are the index sets}
-    $\delta :=  lcm\{k_m|m\in M\}$\;
-    \For{all atom $\tau_l \in L$}
-    {
-        $\theta'_{+\infty}:=  \theta'\{\textit{true}/\tau_l\}$
-    }
-    \For{all atom $\tau_u \in U$}
-    {
-        $\theta'_{+\infty}:=  \theta'_{+\infty}\{\textit{false}/\tau_u\}$
-    }
-    output $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee \bigvee_{j=0}^{\delta-1} \bigvee_{u\in U}\theta'[t_u(\bar{x},\bar{y})-j/x_i']$
-    \caption{Elim-linear (Cooper's QE algorithm for PA)}
-\end{algorithm}
-
-\subsection{Compared with Point's Origin Algorithm}
-
-Point's algorithm \cite{Point86} has four steps,
-corresponding to the four steps of our algorithm respectively.
-So we will use the names of our steps to refer to them. 
-The divisible predicate $n|x$ is not included in his theory,
-instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
-The \textbf{Normalization} and \textbf{Enumerate-Orders} steps in two algorithms are similar,
-with some minor differences due to the setting of the theory.
-The main differences between the two algorithms lies in the following aspects.
-
-In \textbf{Elim-exp},
-we adopt the same idea to find sufficient conditions
-and necessary conditions for the formula to holds.
-But our strategy for choosing parameters are easier to understand and use.
-A flaw lies in the $a_i<0$ case in Point's algorithm 
-\footnote{
-In the $a_i<0$ case of \textbf{Elim-exp-ineq},
-Point's algorithm assumes falsely $t(y)\ge 0$.
-The reason for this mistake might be the confusion of 
-two subtraction operators.
-The example mentioned in \ref{sec:3.2} shows this case.
-},
-our \textbf{Elim-exp-ineq} corrects this flaw by 
-treating $a_i<0$ case similarly to the $a_i>0$ case.
-
-\textbf{Elim-linear} in Point's algorithm in some sense includes our
-\textbf{Elim-exp-div}, because divisible predicates is not introduced in his setting.
-In this step,
-Point uses the permutation groups (just like $S_{(n+1)}$ in \textbf{Enumerate-Orders}) and needs recursively renaming of variables $x_i$.
-We instead invoke the QE algorithm of PA without referring to permutation groups, 
-other improved QE algorithm can also be used to simplify this step.
-
-Point's algorithm has more restrictions on the form
-of the formula.
-Before \textbf{Elim-linear} and \textbf{Elim-exp} step,
-it needs to transform the formula into disjunction normal form (DNF)
-and deals with the basic case when input formula $\theta$ is 
-a conjunction of atoms. 
-It is well known that transforming an arbitrary formula into 
-DNF is costly and the length of the formula may increase exponentially.
-So our algorithm remove these limitations and assume no special forms of
-the formula other than NNF.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-
-\subsection{Complexity Analysis}
-\wuhao{To be checked}
-In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
-%\subsection{Complexity Analysis}
-
-In this part, we estimate the complexity of our algorithm.
-It can be divided into two part, 
-the first part convert a string constraint $\phi$ to a PA formula with exponential functions,
-and the second part is the QE procedure on the obtained formula.
-
-Given a finite alphabet $\Sigma$ and $\alpha=\langle p,q \rangle$, the size of $\Sigma$ is denoted by
-$|\Sigma|$. 
-The generic $\alpha$-flat language contains $O(|\Sigma|^{pq})$ $\alpha$-flat languages
-
-\paragraph{Complexity for QE}
-In \textbf{Enumerate-Orders},
-it reduces the origin problem to $(n+1)!$ sub-cases,
-and for each case,
-we recursively invoke \textbf{Elim-exp} and \textbf{Elim-linear}.
-So the key is to analysis the inner \textit{for}
-loop in \textbf{Enumerate-Orders}.
-Note that if we omit the \textbf{Elim-exp} step,
-the sub-cases are equivalent to Cooper's algorithm,
-so we adopt the idea of Oppen's analysis to analyze the complexity
-\cite{Oppen}.
-
-Apply Cooper's algorithm on the quantified PA formula 
-$Q_m x_m Q_{m-1}$ 
-$x_{m-1} ... Q_1 x_1. $
-$F(x_1,...,x_m)$,
-the algorithm repeats $m$ times to eliminate
-$x_1,...,x_m$ one by one.
-Let $F_k \Def Q_m x_m Q_{m-1} 
-x_{m-1} ... Q_{k+1} x_{k+1}. F_k'(x_{k+1},...,x_m)$
-be the formula produced after $k$th iterations of the algorithm. 
-Let $a_k$ denote the number of atoms in $F_k$,
-$c_k$ denote the number of distinct $\delta_i$ in atom $\delta_i | t$ (and its negation) plus the number of distinct coefficients of variables.
-$s_k$ denote the largest value of integer constant.
-Use $a_0,c_0,s_0$ to denote the initial value of $a_k,c_k,s_k$ respectively.
-The following theorem holds.
-
-\begin{theorem}
-for all $k: 1\le k\le m$, the following relations holds
-\begin{itemize}
-    \item $c_k \le {c_{k-1}}^4$
-    \item $s_k \le {s_{k-1}}^{4c_{k-1}}$
-    \item $a_k \le {a_{k-1}}^4 \cdot {s_{k-1}}^{2c_{k-1}}$
-\end{itemize}
-and from the above relations we have 
-\begin{itemize}
-    \item $c_k \le {{c_0}^4}^k$
-    \item $s_k \le {{{s_0}^{4c_0}}^4}^k$
-    \item $a_k \le {{a_0}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
-\end{itemize}
-\end{theorem}
-
-For each $k$,
-the space required to store $F_k$ 
-is $a_k \cdot (m+1) \cdot s_k \cdot q$,
-where $(m+1)$ is the maximum number of constants per atom and $q>1$ is some constant.
-Assume $m\le n, c\le n, a\le n, s\le n$, we obtain the deterministic space complexity is ${{2^2}^2}^{p n log_n}$, which is also the bound for deterministic time.
-
-In our algorithm,
-we use the same denotations and have the following theorem. 
-Note that each iteration corresponds to \textbf{Elim-exp} and \textbf{Elim-linear}.
-\begin{theorem}
-for all $k: 1\le k\le m$, the following relations holds
-\begin{itemize}
-    \item $c_k \le {c_{k-1}}^4$
-    \item $s_k \le {(n s_{k-1})}^{4c_{k-1}}$
-    \item $a_k \le {{(l_2(ns_{k-1})\cdot a_{k-1})}}^4 \cdot {(n s_{k-1})}^{2c_{k-1}}$
-\end{itemize}
-and from the above relations we have 
-\begin{itemize}
-    \item $c_k \le {{c_0}^4}^k$
-    \item $s_k \le {{{n}^{4c_0}}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
-    \item $a_k \le {{a_0}^4}^k \cdot 
-     {{{n}^{4c_0}}^4}^k
-     \cdot {{{s_0}^{4c_0}}^4}^k$
-\end{itemize}
-\end{theorem}
-
-WAIT TO CHECK!
-
-Again assume $m,c,a,s\le n$, which gives 
-$$
-\textit{space}\le q \cdot {n^4}^n \cdot {{n^{(4n)}}^4}^n \cdot
-{{n^{(4n)}}^4}^n \cdot (n+1) \cdot 
-\cdot {{n^{(4n)}}^4}^n \cdot {{n^{(4n)}}^4}^n
-\le 2^{2^{2^{p n \textit{log}n}}}
-$$
-
-Then the upper bound for Cooper's algorithm still holds for the inner \textit{for} loop of \textbf{Enumerate-Orders}.
-When considering \textbf{Normalization} and permutations groups in \textbf{Enumerate-Orders},
-since the super exponential term dominates the number of permutations $n^{O(n)}$,
-this upper bound still holds.
-
-}
+\wuhao{describe the whole procedure using pseudo code}

### 2021-10-25 FMCAD2021/paper.tex
dodify

@@
     In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion.
-The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-The complexity of the quantifier elimination procedure is analyzed for the first time and we show that the elimination of one quantifier is 3-EXPSPACE.
+\wuhao{The approach is built upon a quantifier elimination procedure for the existential fragment of the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
+The complexity of the quantifier elimination procedure is analyzed for the first time and is shown to be 3-EXPSPACE if the formula contains only existential quantifiers}.
 While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.

### 2021-10-26 FMCAD2021/1_intro.tex
structure

@@
 This can be seen as another contribution of this paper.
-\wuhao{We also analyze the deterministic upper bound of complexity for the first time: given a formula with only existential quantifiers, the complexity is 3-EXPSPACE to eliminate all quantifiers}.
+\wuhao{We also analyze the deterministic upper bound of complexity for the first time: given a $\paexp$ formula with only existential quantifiers, eliminating all quantifiers has a 3-EXPSPACE complexity}.
 Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

### 2021-10-26 FMCAD2021/3_flatten.tex
structure

@@
 
-{\paexp} extends Presburger arithmetic with the exponential function $10^x$ as well as the (partial) functions $\ell_{10}(\ivarx)$ (cf.
+{\paexp} extends Presburger arithmetic with the exponential function $10^\ivarx$ as well as the (partial) functions $\ell_{10}(\ivarx)$ (cf.
 \cite{Point86}).
@@
  $$
+ Sometimes we will denote $\neg c | \iterm$ by $c\nmid \iterm$ to avoid the negation operator.
 The semantics of {\paexp} are defined similarly to {\pa} with the difference that variables are interpreted over $\Nat$ (to avoid terms like $10^{-2}$), that is, an interpretation is a function $I: \free(\phi) \rightarrow \Nat$.
@@
 \begin{theorem} \label{thm:string-parInt}
-    \wuhao{The satisfiability of  $\strint$ under flat domain restrictions is equivalent to the satisfiability of existential $\paexp$ formulas, and thus is decidable. }
+    The satisfiability of  $\strint$ under flat domain restrictions is equivalent to the satisfiability of existential $\paexp$ formulas, and thus is decidable.
 \end{theorem}

### 2021-10-26 FMCAD2021/paper.tex
structure

--- 2021-09-29 FMCAD2021/paper.tex
@@
     In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion.
-The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-The complexity of the quantifier elimination procedure is analyzed for the first time and we show that the elimination of one quantifier is 3-EXPSPACE.
+The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
+\wuhao{The complexity of the quantifier elimination procedure is analyzed for the first time and is shown to be 3-EXPSPACE if the formula contains only existential quantifiers}.
 While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.

### 2021-11-08 Acta Informatica2022/1_intro.tex
Acta Informatica format

--- 2021-10-25 FMCAD2021/1_intro.tex
@@
 Many highly efficient solvers for string constraints, to name a few, CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, Z3-Str3~\cite{zheng2013z3}, S3~\cite{trinh2014s3}, Norn~\cite{abdulla2015norn}, Ostrich~\cite{chen2017decidable}, Sloth~\cite{sloth}, ABC~\cite{aydin2018parameterized}, Stranger~\cite{yu2010stranger}, and Trau~\cite{abdulla2018trau}, are developed in the last decade.
-
 The satisfiability of \emph{string-integer conversion constraints}, e.g., $wx=yx \wedge |x| > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
@@
 We propose a decision procedure for the class of word equation + flat languages + length constraints + string-integer conversion.
-The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$), \wuhao{more precisely, the existential fragment of $\paexp$}; then, according to the decidability of $\paexp$, we obtain the decidability of the original satisfiability problem.
+The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$), more precisely, the existential fragment of $\paexp$; then, according to the decidability of $\paexp$, we obtain the decidability of the original satisfiability problem.
 
@@
 This can be seen as another contribution of this paper.
-\wuhao{We also analyze the deterministic upper bound of complexity for the first time: given a formula with only existential quantifiers, the complexity is 3-EXPSPACE to eliminate all quantifiers}.
+We also analyze the deterministic upper bound of complexity for the first time: given a $\paexp$ formula with only existential quantifiers, eliminating all quantifiers has a 3-EXPSPACE complexity.
 Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

### 2021-11-08 Acta Informatica2022/2_preliminary.tex
Acta Informatica format

--- 2021-05-20 FMCAD2021/2_preliminary.tex
@@
 $n$.
-In particular, $|\varepsilon| = 0$.
+In particular, $\vert \varepsilon\vert  = 0$.
 %
@@
  \iterm &\Def & c \mid \ivarx \mid \iterm+\iterm \mid \iterm - \iterm, \\
- \phi & \Def & \iterm \ \op\ \iterm \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,
-
+ \phi & \Def & \iterm \ \op\ \iterm \mid c \vert \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,
  \end{array}
@@
 Given an PA formula $\phi$, and an integer interpretation of $\free(\phi)$, i.e.
-a function $I: \free(\phi) \rightarrow \Int$, we denote by $I \models \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner, with $+$, $-$, and $|$ interpreted as the integer addition, subtraction, and divisibility relation respectively), and call $I$ a \emph{model} of $\phi$.
+a function $I: \free(\phi) \rightarrow \Int$, we denote by $I \models \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner, with $+$, $-$, and $\vert$ interpreted as the integer addition, subtraction, and divisibility relation respectively), and call $I$ a \emph{model} of $\phi$.
 We use $\llbracket \phi \rrbracket$ to denote the set of models of $\phi$.

### 2021-11-08 Acta Informatica2022/3_flatten.tex
Acta Informatica format

--- 2021-10-26 FMCAD2021/3_flatten.tex
@@
 
-\vspace*{-3mm}
 \subsection{String constraints with string-integer conversion ($\strint$)}
@@
     \begin{itemize}
-%        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
+%        \item if $\vertw\vert=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
         \item  if $w=$`$i$' for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}$$($`$i$'$)=i$;
-        \item for $w = w_1$`$i$' for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
+        \item for $w = w_1$`$i$' for $i \in \Sigma_{\textit{num}}$ with $\vert w_1\vert  \ge 1$, 
         ${\parseInt}(w) = 10*{\parseInt}(w_1)+{\parseInt}($`$i$'$)$.
@@
 A $\strint$ formula $\varphi$ is defined by the following rules, where $\len(\sterm)$ denotes the length of a string $\sterm$.
-\vspace{-2mm}
+
 \[
@@
 \iterm & \Def & n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,\\
-\vspace{-2mm}
 \varphi & \Def & \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi,
@@
 \item A length constraint $\phi_l$ is a linear constraint over 
-$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
-where $|\cdot |$ is the length function.
+$Z \cup \{\vert x\vert  \mid x\in X\}$, %the values of $\vert x\vert $ for all $x\in X$,
+where $\vert \cdot \vert $ is the length function.
 We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
@@
 
-\vspace*{-5mm}
 \subsection{An Extension of Presburger Arithmetic with Exponential Functions (\paexp)}
@@
 The syntax of {\paexp} is obtained from that of {\pa} by adding $10^\iterm$ and $\ell_{10}(\iterm)$ to the definition of integer terms. Specifically, {\paexp} formulas are defined by the rules,
-%
-\vspace{-1mm}
+
 $$
@@
 \iterm &\Def& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \\
-\vspace{-1mm}
- \phi &\Def & \iterm \ \op\ \iterm \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi.
+\phi &\Def & \iterm \ \op\ \iterm \mid c \vert \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi.
 \end{array}
 $$
- Sometimes we will denote $\neg c | \iterm$ by $c\nmid \iterm$ to avoid the negation operator.
+
+Sometimes we will denote $\neg c \vert \iterm$ by $c\nmid \iterm$ to avoid the negation operator.
 The semantics of {\paexp} are defined similarly to {\pa} with the difference that variables are interpreted over $\Nat$ (to avoid terms like $10^{-2}$), that is, an interpretation is a function $I: \free(\phi) \rightarrow \Nat$.
@@
 
-\vspace*{-2mm}
 \subsection{Flattening $\strint$ into $\paexp$}
@@
 %
-$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)}).$$
-%
+$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket \vert_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)}).$$
+
 The decoding function above decodes an interpretation of integer and flattening variables $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \Nat$ as a set $\decode_{\flatfun_\varphi}(I_e)$ of interpretations of the $\varphi$'s integer and string variables $(I_s, I_i)$ with $I_s: \svar(\varphi) \rightarrow \Sigma^*$ and $I_i: \ivar(\phi) \rightarrow \Nat$ such that
@@
 \item for $i = k_\svarx$, 
-$$ \hspace{-4mm} \iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{|w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx} } -1 )}{(10^{|w_{\svarx, k_\svarx}|} -1 )}$$ 
-and $\ell_{\svarx, i} = |w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx}$,
+$$\iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{\vert w_{\svarx,k_\svarx}\vert  \#_{\svarx, k_\svarx} } -1 )}{(10^{\vert w_{\svarx, k_\svarx}\vert } -1 )}$$ 
+and $\ell_{\svarx, i} = \vert w_{\svarx,k_\svarx}\vert  \#_{\svarx, k_\svarx}$,
 %
@@
 %
-$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{|w_{\svarx, i} | \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{|w_{\svarx, i}|} -1 )} + \iterm_{\svarx, i+1}$$
+$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{\vert w_{\svarx, i} \vert  \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{\vert w_{\svarx, i}\vert } -1 )} + \iterm_{\svarx, i+1}$$
 %
-and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
+and $\ell_{\svarx, i} = \vert w_{\svarx, i} \vert  \#_{\svarx, i} + \ell_{\svarx, i+1}$.
 %
 \end{itemize}
-Notice that here $|w_{\svarx, \textunderscore } |$ and $\ell_{\svarx, \textunderscore}$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables.
+Notice that here $\vert w_{\svarx, \textunderscore } \vert $ and $\ell_{\svarx, \textunderscore}$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables.
 %
-\item $\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} |w_{\svarx, i} | \#_{\svarx, i}$. 
+\item $\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} \vert w_{\svarx, i} \vert  \#_{\svarx, i}$. 
 \end{itemize}
@@
 \item Replace each integer term $\len(t)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by 
- $\sum \limits_{i \in [m]} \theta_i$, where for every $i \in [m]$, $\theta_i = 1$ if $\alpha_i \in \Sigma_{\textit{num}}$ and $ \theta_i = \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j} | \#_{\alpha_i, j}$ otherwise.
+ $\sum \limits_{i \in [m]} \theta_i$, where for every $i \in [m]$, $\theta_i = 1$ if $\alpha_i \in \Sigma_{\textit{num}}$ and $ \theta_i = \sum \limits_{j \in [k_{\alpha_i}]} \vert w_{\alpha_i, j} \vert  \#_{\alpha_i, j}$ otherwise.
 %
@@
 \begin{itemize}
-\item if $\alpha_m \in \Sigma_{\textit{num}}$, then $\iterm_{t, m} = \alpha_m$ and $\ell_{t, m} = 1$, otherwise, $\iterm_{t, m} = \flatten_{\flatfun_\varphi}(\parseInt(\alpha_m))$ and $\ell_{t, m} = \sum \limits_{j \in [k_{\alpha_m}]} |w_{\alpha_m, j}| \#_{\alpha_m, j}$,
+\item if $\alpha_m \in \Sigma_{\textit{num}}$, then $\iterm_{t, m} = \alpha_m$ and $\ell_{t, m} = 1$, otherwise, $\iterm_{t, m} = \flatten_{\flatfun_\varphi}(\parseInt(\alpha_m))$ and $\ell_{t, m} = \sum \limits_{j \in [k_{\alpha_m}]} \vert w_{\alpha_m, j}\vert  \#_{\alpha_m, j}$,
 %
 \item for $i \in [m-1]$, if $\alpha_i \in \Sigma_{\textit{num}}$, then $\iterm_{t, i} = \alpha_i 10^{\ell_{t, i+1}} + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1}+1$, otherwise, $\iterm_{t, i} = 
-\flatten_{\flatfun_\varphi}(\parseInt(\alpha_i))10^{\ell_{t, i+1}}  + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1} + \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j}| \#_{\alpha_i, j}$.
+\flatten_{\flatfun_\varphi}(\parseInt(\alpha_i))10^{\ell_{t, i+1}}  + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1} + \sum \limits_{j \in [k_{\alpha_i}]} \vert w_{\alpha_i, j}\vert  \#_{\alpha_i, j}$.
 \end{itemize}

### 2021-11-08 Acta Informatica2022/4_dec_proc.tex
Acta Informatica format

--- 2021-10-25 FMCAD2021/4_dec_proc.tex
@@
 \item \textit{NNF transformation} At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form). Moreover, we remove the occurrences of $\neg$ by replacing 
-\hide{(a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, }
-(a) $\neg c | \iterm$ with $ c \nmid \iterm$, 
+\hide{(a) replacing $\neg c \vert  \iterm$ with $\bigvee \limits_{j \in [c-1]} c \vert  (\iterm+j)$, }
+(a) $\neg c \vert  \iterm$ with $ c \nmid \iterm$, 
 (b) $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, 
@@
 
-\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c | \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c \vert  \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
 Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
@@
 
-After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
-Denote the negation of $c | \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
+After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c \vert  \iterm$.
+Denote the negation of $c \vert  \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
 
@@
 It also applies some additional transformations.
-After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c | \iterm$.
-Denote the negation of $c | \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
+After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c \vert  \iterm$.
+Denote the negation of $c \vert  \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
 A formula is \textit{normalized} if it satisfies the above properties.
@@
 
-Recall that after normalization, the atomic formulas are of the form $\iterm_1 \le \iterm_2$, $c | \iterm$ or $c \nmid \iterm$.
+Recall that after normalization, the atomic formulas are of the form $\iterm_1 \le \iterm_2$, $c \vert  \iterm$ or $c \nmid \iterm$.
 Therefore, we can assume that the atomic formulas in $\varphi''_{\sigma,i}$ are  of the form
@@
 or  
-$c \ \big  | \ \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big)$
+$c \ \big  \vert  \ \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big)$
 (or $\nmid$).
@@
 %
-with $a_i \neq 0$, $A\Def \sum_{j=1}^{i-1}|a_j|$, 
-%$B\Def  \sum_{j=1}^{i}|b_j|$, 
-$B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$,
+with $a_i \neq 0$, $A\Def \sum_{j=1}^{i-1}\vert a_j\vert $, 
+%$B\Def  \sum_{j=1}^{i}\vert b_j\vert $, 
+$B \Def 2(\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert )+3)$,
 and $\delta\Def  \ell_{10}(A)+3$. 
@@
 \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-d\ \big |\ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+d\ \big \vert \ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
 + \iterm(\vec{\ivary}) \big)
@@
 
-Let $d = 2^{r_1} 5^{r_2}  d_0$ such that $d_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $d | (10^rd_0)$.
+Let $d = 2^{r_1} 5^{r_2}  d_0$ such that $d_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $d \vert  (10^rd_0)$.
 
 If $d_0 = 1$, then $10^r$ is divisible by $d = 2^{r_1}5^{r_2}$.
-Thus for every $n \ge r$, $d \ |\ 10^n$.
+Thus for every $n \ge r$, $d \ \vert \ 10^n$.
 Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to
@@
 \begin{array}{l}
-\bigvee \limits_{p = 0}^{r - 1} d\ \big | \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
+\bigvee \limits_{p = 0}^{r - 1} d\ \big \vert  \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
 + \iterm(\vec{\ivary}) \big)\\
 %
-\vee \big(\ivarx_{\sigma(i)} \ge r \wedge d\ \big | \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+\vee \big(\ivarx_{\sigma(i)} \ge r \wedge d\ \big \vert  \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
 + \iterm(\vec{\ivary}) \big)\big),
@@
 \begin{array}{l}
-\phi(d_0)\ \big |\ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
-d\ \big | 
+\phi(d_0)\ \big \vert \ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
+d\ \big \vert  
 \left(
@@
 where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
-
-\wuhao{describe the whole procedure using pseudo code}

### 2021-11-08 Acta Informatica2022/5_complexity.tex
Acta Informatica format

--- 2021-09-29 FMCAD2021/5_complexity.tex
@@
 
-Let $c_k$ be the number of distinct $d$ in atoms of the form $d \ \big | \ \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of quantified variables in $\varphi_k$.
+Let $c_k$ be the number of distinct $d$ in atoms of the form $d \ \big \vert  \ \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of quantified variables in $\varphi_k$.
 Let $s_k$ be the largest constant (including coefficients) and $a_k$ be the number of atomic formulas in $\varphi_k$.
@@
 
-If all atoms are divisiblilty atomic formulas of the form $d \ \big | \ \iterm$ or $d \ \nmid \ \iterm$.
-We have $c'_1<2c$ since a divisibility atomic formula  will produce at most two forms of atomic formulas $d \ \big | \ \iterm$ and $\phi(d) \ \big | \ \iterm$.
+If all atoms are divisiblilty atomic formulas of the form $d \ \big \vert  \ \iterm$ or $d \ \nmid \ \iterm$.
+We have $c'_1<2c$ since a divisibility atomic formula  will produce at most two forms of atomic formulas $d \ \big \vert  \ \iterm$ and $\phi(d) \ \big \vert  \ \iterm$.
 Note that any constant in $\iterm$ , say $l$, can be replaced by ($l \bmod d)$, so $s'_1\le s$.

### 2021-11-08 Acta Informatica2022/9_appendix.tex
Acta Informatica format

--- 2021-09-29 FMCAD2021/9_appendix.tex
@@
 %!TEX root = paper.tex
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
     \subsection{Details of the normalization step}\label{app-norm}
@@
     \item \textit{NNF transformation} At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form). Moreover, we remove the occurrences of $\neg$ by replacing 
-\hide{(a) replacing $\neg c | \iterm$ with $\bigvee \limits_{j \in [c-1]} c | (\iterm+j)$, }
-(a) $\neg c | \iterm$ with $ c \nmid \iterm$, 
+    \hide{(a) replacing $\neg c \vert  \iterm$ with $\bigvee \limits_{j \in [c-1]} c \vert  (\iterm+j)$, }
+    (a) $\neg c \vert  \iterm$ with $ c \nmid \iterm$, 
     (b) $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, 
@@
 
-\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c | \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+    \item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c \vert  \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
 Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
@@
     \end{enumerate}
-
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Proof of Lemma~\ref{lem:exp-ineq}}
@@
 
-Let $A\Def \sum_{j=1}^{i-1}|a_j|$, $B \Def 2(\ell_{10}(\sum_{j=1}^{i}|b_j|)+3)$, and $\delta\Def  \ell_{10}(A)+3$.
+Let $A\Def \sum_{j=1}^{i-1}\vert a_j\vert $, $B \Def 2(\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert )+3)$, and $\delta\Def  \ell_{10}(A)+3$.
 
@@
 and
-$$-(\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)} \le \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} \le (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)}.$$
+$$-(\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)} \le \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} \le (\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)}.$$
 
-Moreover, let $n = 100\sum_{j=1}^{i}|b_j|$, $m = 1$, and $p = \ivarx_{\sigma(i)}$, then 
+Moreover, let $n = 100\sum_{j=1}^{i}\vert b_j\vert $, $m = 1$, and $p = \ivarx_{\sigma(i)}$, then 
 $n \ge m \ge 1$.
-From $2(\ell_{10}(n)- \ell_{10}(m)+1) = 2 (\ell_{10}(100 \sum_{j=1}^{i}|b_j|) + 1) = 2 (\ell_{10}(\sum_{j=1}^{i}|b_j|) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 2(\ell_{10}(n)- \ell_{10}(m)+1)$.
+From $2(\ell_{10}(n)- \ell_{10}(m)+1) = 2 (\ell_{10}(100 \sum_{j=1}^{i}\vert b_j\vert ) + 1) = 2 (\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert ) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 2(\ell_{10}(n)- \ell_{10}(m)+1)$.
 Then according to Proposition~\ref{prop:1}, 
-$$100(\sum_{j=1}^{i}|b_j|)\ivarx_{\sigma(i)} = np \le m 10^p = 10^{\ivarx_{\sigma(i)}}.$$
+$$100(\sum_{j=1}^{i}\vert b_j\vert )\ivarx_{\sigma(i)} = np \le m 10^p = 10^{\ivarx_{\sigma(i)}}.$$
 Thus 
-$(\sum_{j=1}^{i}|b_j|)\ivarx_{\sigma(i)}  \le \frac{1}{100} 10^{\ivarx_{\sigma(i)}}.$
+$(\sum_{j=1}^{i}\vert b_j\vert )\ivarx_{\sigma(i)}  \le \frac{1}{100} 10^{\ivarx_{\sigma(i)}}.$
 
@@
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \ge \\
-a_i 10^{\ivarx_{\sigma(i)}} - A 10^{\ivarx_{\sigma(i)} - \delta} - (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)}  \ge \\
+a_i 10^{\ivarx_{\sigma(i)}} - A 10^{\ivarx_{\sigma(i)} - \delta} - (\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)}  \ge \\
 a_i 10^{\ivarx_{\sigma(i)}} - \frac{1}{100}10^{\ivarx_{\sigma(i)}} - \frac{1}{100} 10^{\ivarx_{\sigma(i)}} = \\
@@
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \\
-a_i 10^{\ivarx_{\sigma(i)}} + A 10^{\ivarx_{\sigma(i)} - \delta} + (\sum_{j=1}^{i}|b_j|) \ivarx_{\sigma(i)} \le \\
+a_i 10^{\ivarx_{\sigma(i)}} + A 10^{\ivarx_{\sigma(i)} - \delta} + (\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)} \le \\
 a_i 10^{\ivarx_{\sigma(i)}} + \frac{A} {10^\delta}10^{\ivarx_{\sigma(i)}} + \frac{1}{100} 10^{\ivarx_{\sigma(i)}} \le \\
@@
 
+\hide{
     \subsection{Elimination of exponential occurrences of variables for (in)divisibility atomic formulas}\label{app-div}
@@
     \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-d\ \big |\ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+    d\ \big \vert \ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
     + \iterm(\vec{\ivary}) \big)
@@
 
-Let $d = 2^{r_1} 5^{r_2}  d_0$ such that $d_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $d | (10^rd_0)$.
+    Let $d = 2^{r_1} 5^{r_2}  d_0$ such that $d_0$ is divisible by neither $2$ nor $5$. Moreover, let $r = \max(r_1, r_2)$. Then $d \vert  (10^rd_0)$.
 
     If $d_0 = 1$, then $10^r$ is divisible by $d = 2^{r_1}5^{r_2}$.
-Thus for every $n \ge r$, $d \ |\ 10^n$.
+Thus for every $n \ge r$, $d \ \vert \ 10^n$.
 Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to
@@
     \begin{array}{l}
-\bigvee \limits_{p = 0}^{r - 1} d\ \big | \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
+    \bigvee \limits_{p = 0}^{r - 1} d\ \big \vert  \big(a_i 10^{p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_kp + \sum_{k=1}^{i-1} b_k \ivarx_{\sigma(k)} 
     + \iterm(\vec{\ivary}) \big)\\
     %
-\vee \big(\ivarx_{\sigma(i)} \ge r \wedge d\ \big | \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+    \vee \big(\ivarx_{\sigma(i)} \ge r \wedge d\ \big \vert  \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
     + \iterm(\vec{\ivary}) \big)\big),
@@
     \begin{array}{l}
-\phi(d_0)\ \big |\ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
-d\ \big | 
+    \phi(d_0)\ \big \vert \ (\ivarx_{\sigma(i)} - r - q)\ \wedge \\
+    d\ \big \vert  
     \left(
@@
     where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
-
+}
 \subsection{Some additional optimizations of the decision procedure for {\paexp}}\label{app-opt}

### 2021-11-08 FMCAD2021/paper.tex
Acta Informatica format

@@
 The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986.
-\wuhao{The complexity of the quantifier elimination procedure is analyzed for the first time and is shown to be 3-EXPSPACE if the formula contains only existential quantifiers}.
+The complexity of the quantifier elimination procedure is analyzed for the first time and is shown to be 3-EXPSPACE if the formula contains only existential quantifiers.
 While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.

### 2021-11-08 Acta Informatica2022/1_intro.tex
remove appendix

@@
 This extension is called word equations with regular constraints. 
-However, the satisfiability problem of word equations together with length constraints (e.g., $|x|=|y|+1 \wedge wx=yx$) is still open.
+However, the satisfiability problem of word equations together with length constraints (e.g., $\vert x\vert =\vert y\vert +1 \wedge wx=yx$) is still open.
 The complexity of the satisfiability of word equations with regular constraints was proven to be PSPACE-complete by Plandowski~\cite{plandowski99}, after decades of improvement of the original algorithm by Makanin.
@@
 Many highly efficient solvers for string constraints, to name a few, CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, Z3-Str3~\cite{zheng2013z3}, S3~\cite{trinh2014s3}, Norn~\cite{abdulla2015norn}, Ostrich~\cite{chen2017decidable}, Sloth~\cite{sloth}, ABC~\cite{aydin2018parameterized}, Stranger~\cite{yu2010stranger}, and Trau~\cite{abdulla2018trau}, are developed in the last decade.
-The satisfiability of \emph{string-integer conversion constraints}, e.g., $wx=yx \wedge |x| > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
+The satisfiability of \emph{string-integer conversion constraints}, e.g., $wx=yx \wedge \vert x\vert  > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
 However, this kind of constraints is pervasive in scripting language programs. 
@@
 Although they are useful in handling many practical cases, they fail short when the all string solutions exceed the selected bound.
-For example, a constraint of the form $x.y \neq z  \wedge |x| > 2000$ would be quite challenging to handle using those solvers.
+For example, a constraint of the form $x.y \neq z  \wedge \vert x\vert  > 2000$ would be quite challenging to handle using those solvers.
 
@@
 First, under the restriction, the potential solution space is still infinite, which gives us a higher potential of finding solutions.
-For instance, we can find a solution for $x.y \neq z  \wedge |x| > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is an element of the alphabet.
+For instance, we can find a solution for $x.y \neq z  \wedge \vert x\vert  > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is an element of the alphabet.
 Second, and more importantly, because we can convert the membership problem of a flat language to the satisfiability problem of a Presburger arithmetic formula, the class of word equation + flat languages + length constraints is decidable.

### 2021-11-08 Acta Informatica2022/4_dec_proc.tex
remove appendix

@@
 
-\vspace{-4mm}
+%\vspace{-4mm}
 
@@
 \begin{array}{l}
-\vspace{2mm}
+%\vspace{2mm}
 \left(
@@
 
-\vspace*{-3mm}
+%\vspace*{-3mm}
 \subsection{Normalization}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-The normalization step first removes the occurrences of $\ell_{10}(\iterm)$ where $\ivarx$ occurs in $\iterm$, by encoding them with the exponential function.
-Moreover, for each occurrence of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, it introduces a fresh variable $\ivarz$, replaces $10^\iterm$ with $10^\ivarz$, and adds the equality $\ivarz = \iterm$.
-It also applies some additional transformations.
-After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c \vert  \iterm$.
-Denote the negation of $c \vert  \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
-A formula is \textit{normalized} if it satisfies the above properties.
-More details can be found in Appendix~\ref{app-norm}.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\vspace{-2mm}
+%\vspace{-2mm}
 \subsection{Enumeration of the variable orders}
@@
 
-\vspace*{-3mm}
+%\vspace*{-3mm}
 \subsection{Elimination of  exponential occurrences of variables}\label{sec-elim-exp}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-    For convenience, let us call these formulas as inequality respectively (in)divisibility atomic formulas.
-In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these inequality atomic formulas.
-The elimination of the exponential occurrences of the (in)divisibility formulas are simpler and omitted, which can be found in Appendix~\ref{app-div}.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Inequality atoms}
@@
 \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-\hspace{4mm} 
+%\hspace{4mm} 
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
@@
 
-\vspace{-1mm}
+%\vspace{-1mm}
 \begin{lemma} \label{lem:exp-ineq}
@@
 \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-\hspace{4mm} 
-\vspace{-1mm}
+%\hspace{4mm} 
+%\vspace{-1mm}
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
@@
 
+We need the following proposition for the proof of Lemma~\ref{lem:exp-ineq}.
+
+\begin{proposition} \label{prop:1}
+If $n\ge m\ge 1$ and $p \ge 2(\ell_{10}(n)- \ell_{10}(m)+1)$, then 
+$n p \le m10^p$ holds.
+\end{proposition}
+
+\begin{proof}
+
+First we show that for any $n' \in \Nat$, if $p \ge 2n'$, then $10^p \ge 10^{n'}10^{(p-n')}\ge 10^{n'}10(p-n')\ge 10^{n'}(5p+5(p-2n'))\ge  10^{n'}p$.
+
+If $ p \ge 2 (\ell_{10}(n) - \ell_{10}(m)+1)$, then  $n p \le 10\lambda_{10}(n) p = 10 * 10^{\ell_{10}(n)} p = 10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p$.
+
+Because $p \ge 2(\ell_{10}(n)-\ell_{10}(m) + 1)$, we deduce that $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  p \le 10^p$.
+Therefore, $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p \le 10^{\ell_{10}(m)} 10^p \le m 10^p$.
+We conclude that $np \le m 10^p$.
+\end{proof}
+
+%Then we give the proof for Theorem~\ref{thm:exp-ineq}.
+\begin{proof}[Proof of Lemma~\ref{lem:exp-ineq}]
+We only prove for the case $a_i > 0$, the other case is symmetric.
+
+Let $A\Def \sum_{j=1}^{i-1}\vert a_j\vert $, $B \Def 2(\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert )+3)$, and $\delta\Def  \ell_{10}(A)+3$.
+
+Suppose $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$. Moreover, let $\alpha(\vec{\ivary}) \Def  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$.
+
+Note that
+ \begin{equation} 
+   A10^{-\delta} =A 10^{-\ell_{10}(A)-3} = \frac{A}{1000\lambda_{10}(A)} \le \frac{1}{100}.   \label{eq:thm-ineq-1}
+ \end{equation}
+
+ From $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$ and $\ivarx_{\sigma(i-1)} \ge \ldots \ge \ivarx_{\sigma(1)}$, 
+we know   
+$$-A 10^{\ivarx_{\sigma(i)} - \delta} \le  \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} \le A 10^{\ivarx_{\sigma(i)} - \delta}$$
+and
+$$-(\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)} \le \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} \le (\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)}.$$
+
+Moreover, let $n = 100\sum_{j=1}^{i}\vert b_j\vert $, $m = 1$, and $p = \ivarx_{\sigma(i)}$, then 
+$n \ge m \ge 1$.
+From $2(\ell_{10}(n)- \ell_{10}(m)+1) = 2 (\ell_{10}(100 \sum_{j=1}^{i}\vert b_j\vert ) + 1) = 2 (\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert ) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 2(\ell_{10}(n)- \ell_{10}(m)+1)$.
+Then according to Proposition~\ref{prop:1}, 
+$$100(\sum_{j=1}^{i}\vert b_j\vert )\ivarx_{\sigma(i)} = np \le m 10^p = 10^{\ivarx_{\sigma(i)}}.$$
+Thus 
+$(\sum_{j=1}^{i}\vert b_j\vert )\ivarx_{\sigma(i)}  \le \frac{1}{100} 10^{\ivarx_{\sigma(i)}}.$
+
+If $\ivarx_{\sigma(i)} \ge   \alpha(\vec{\ivary}) + 2$, then
+
+$
+\begin{array}{l}
+a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \ge \\
+a_i 10^{\ivarx_{\sigma(i)}} - A 10^{\ivarx_{\sigma(i)} - \delta} - (\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)}  \ge \\
+a_i 10^{\ivarx_{\sigma(i)}} - \frac{1}{100}10^{\ivarx_{\sigma(i)}} - \frac{1}{100} 10^{\ivarx_{\sigma(i)}} = \\
+(a_i - \frac{1}{50}) 10^{\ivarx_{\sigma(i)}} \ge (a_i - \frac{1}{50}) 10^{ \alpha(\vec{\ivary}) + 2} = \\
+(a_i - \frac{1}{50}) 10^{  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i) + 2} = \\
+\frac{10(a_i - \frac{1}{50})}{10^{\ell_{10}(a_i)} } 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \frac{10(a_i - \frac{1}{50})} {a_i} 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \\
+(10 - \frac{1}{5a_i}) 10^{ \ell_{10}(\iterm(\ivary))+1} > 10^{ \ell_{10}(\iterm(\ivary))+1} \ge \iterm(\ivary).
+\end{array}
+$
+
+Therefore, in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \emph{does not} hold.
+
+If $\ivarx_{\sigma(i)} \le   \alpha(\vec{\ivary}) - 1$, then
+
+$
+\begin{array}{l}
+a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \\
+a_i 10^{\ivarx_{\sigma(i)}} + A 10^{\ivarx_{\sigma(i)} - \delta} + (\sum_{j=1}^{i}\vert b_j\vert ) \ivarx_{\sigma(i)} \le \\
+a_i 10^{\ivarx_{\sigma(i)}} + \frac{A} {10^\delta}10^{\ivarx_{\sigma(i)}} + \frac{1}{100} 10^{\ivarx_{\sigma(i)}} \le \\
+(a_i + \frac{1}{100} + \frac{1}{100})10^{\ivarx_{\sigma(i)}} = (a_i + \frac{1}{50}) 10^{\ivarx_{\sigma(i)}} \le \\
+(a_i + \frac{1}{50}) 10^{\alpha(\vec{\ivary}) - 1} = (a_i + \frac{1}{50}) 10^{\ell_{10}(\iterm(\ivary))- \ell_{10}(a_i) - 1} = \\
+\frac{a_i + \frac{1}{50}} {10^{\ell_{10}(a_i) + 1}} 10^{\ell_{10}(\iterm(\ivary))} = \frac{a_i + \frac{1}{50}} {10 \lambda_{10}(a_i)} 10^{\ell_{10}(\iterm(\ivary))} \le \\
+\frac{a_i + \frac{1}{50}} {a_i+1} 10^{\ell_{10}(\iterm(\ivary))} \le 10^{\ell_{10}(\iterm(\ivary))} \le \iterm(\ivary).
+\end{array}
+$
+
+Therefore,  in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
+
+\end{proof}
+
+We would like to remark that Lemma~\ref{lem:exp-ineq} still holds when the base of exponential function is changed to any natural number $n\ge 2$.
+
 If $a_i > 0$, then the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ can be eliminated by utilizing  Lemma~\ref{lem:exp-ineq} and enumerating the constraints on $\ivarx_{\sigma(i)}$ and $\ivarx_{\sigma(i-1)}$.
@@
 $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
-\vspace{-2mm}
+%\vspace{-2mm}
 \[
@@
  \ivarx_{\sigma(i)} = \ivarx_{\sigma(i-1)} +p\ \wedge \\
- \vspace{-1mm}
+ %\vspace{-1mm}
  \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\ivarx_{\sigma(i-1)}+ p /\ivarx_{\sigma(i)}]

### 2021-11-08 Acta Informatica2022/6_optimization.tex
remove appendix

--- 2021-05-21 FMCAD2021/5_optimization.tex
@@
 
-In the sequel, we present two major optimizations and omit some additional optimizations due to the page limit, which can be found in Appendix~\ref{app-opt}.
+%In the sequel, we present two major optimizations and omit some additional ones, which can be found in Appendix~\ref{app-opt}. 
+In the sequel, we will explain two major optimizations in \ref{para: opt reduce} and \ref{para: opt under appro}. Additional optimizations are listed in \ref{para: opts}.
 
-\vspace*{-3mm}
-\subsection{Reduce the number of enumerated variable orders by over approximation}
+\subsection{Reduce the number of enumerated variable orders by over approximation}\label{para: opt reduce}
 
@@
 
-\vspace*{-3mm}
+%\vspace*{-3mm}
 \subsection{Avoid the elimination of linear occurrences of variables by under approximation}\label{para: opt under appro}
@@
 Recall that Lemma~\ref{lem:exp-ineq} enables us to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ from 
-\vspace{-1mm}
+%\vspace{-1mm}
 $$
@@
 \hspace{4mm} 
-\vspace{-1mm}
+%\vspace{-1mm}
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c.
@@
 %\begin{equation}\label{eqn-org}
-\vspace{-2mm}
+%\vspace{-2mm}
 \[\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le c - b_i  \ivarx_{\sigma(i)}.\]
-\vspace{-1mm}
+%\vspace{-1mm}
 %\end{equation}
@@
 Then we can apply Lemma~\ref{lem:exp-ineq} to the following two inequalities to eliminate $10^{\ivarx_{\sigma(i-1)}}$,
-\vspace{-2mm}
+%\vspace{-2mm}
 \begin{equation}\label{eqn-lb}
@@
 and
-\vspace{-2mm}
+%\vspace{-2mm}
 \begin{equation}\label{eqn-ub}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Additional optimization techniques}\label{para: opts}
+\paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}
+
+Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be synchronized.
+That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, B^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$.
+Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.
+
+\paragraph{Avoid the formula-size blow-up by depth-first search}
+
+The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size.
+If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory.
+Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
+If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.
+
+\paragraph{Preprocess with small upper bound}
+
+We believe that if a quantifier-free {\paexp} formula is satisfiable, then most probably it is satisfiable with small values assigned to variables.
+Consequently, as a preprocessing step, we put a small upper bound, e.g.
+the biggest constant occurring in the formula, on the values of variables, and perform a depth-first search, so that a model can be quickly found, if there is any.
+If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.

### 2021-11-08 Acta Informatica2022/7_impl_exp.tex
remove appendix

--- 2021-09-29 FMCAD2021/7_impl_exp.tex
@@
 
-\vspace*{-2mm}
+%\vspace*{-2mm}
 \subsection{Implementation}
@@
 
-\vspace*{-4mm}
+%\vspace*{-4mm}
 \subsection{Benchmarks}
@@
 
-\vspace*{-4mm}
+%%\vspace*{-4mm}
 \subsection{Experiments}
@@
 \begin{itemize}
-\item over the ARITHMETIC benchmark suite, we compare the {\paexp}-solver against CVC4 and Z3,
-\item over the STRINGHASH benchmark suite, we compare the {\paexp}-solver against CVC4, Z3, and Trau. 
+\item over the ARITHMETIC benchmark suite, we compare the {\paexp}-solver against CVC4 (version 1.8) and Z3 (version 4.8.10),
+\item over the STRINGHASH benchmark suite, we compare the {\paexp}-solver against CVC4, Z3, and Trau\footnote{\url{https://github.com/guluchen/z3/tree/new_trau}}. 
 \end{itemize}
@@
 From the results, we can see that the $\paexp$-solver achieves a good tradeoff between precision and efficiency, although it is slower than the other solvers.
-(More detailed experiment results on STRINGHASH can be found in Appendix~\ref{app-exp}.)
+(More detailed experiment results on STRINGHASH can be found in Table~\ref{table:string}.)
+\wuhao{The two tables may be too big.}
 
@@
 % Merge bound and unbound 
-\vspace*{-3mm}
-\begin{table}[ht]
+%\vspace*{-3mm}
+\setlength\tabcolsep{3pt}
+
+\begin{table*}[ht]
+        \caption{Experimental Results, O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
+    $\#$: number of problems, $T$: average time in seconds}
+    \centering
+    \renewcommand{\arraystretch}{1.1} 
+    \begin{tabular}{ c c c c c c c c c c }
+    \hline
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c}{Z3} & \multicolumn{2}{c}{CVC4} &  \multicolumn{2}{c}{Trau} & \multicolumn{2}{c }{$\paexp$} \\ 
+        &  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$} & \scriptsize{$\#$} & \scriptsize{$T$}  \\ \hline
+        \multirow{4}{*}{
+        \newline \scriptsize{(2,3,3,4)}} & S & 56 & {\bf 0.4} & 42 & 2.3 & - & - &  {\bf 64} & {\bf 0.4} \\
+        \cline{2-10}
+         & U & 69 & {\bf 0.1} & 72 &  {\bf 0.1} & - & - &  {\bf 89} & {\bf 0.1} \\
+         \cline{2-10}
+         & B & - & - & - & - & - & - &  47 & 9.5 \\
+         \cline{2-10}
+         & F & 75 & - & 86 & - & - & - &  {\bf 0} & - \\ 
+         \cline{1-10}
+        \multirow{4}{*}{\scriptsize{(3,4,4,5)}} & S & 33 & {\bf 1.4} & 25 & 2.9 & - & - &  {\bf 52} & 3.3 \\
+        \cline{2-10}
+         & U & 59 & {\bf 0.1} & 60 & {\bf 0.1} & - & - &  {\bf 88} & {\bf 0.1} \\
+         \cline{2-10}
+         & B & -  & - &  - & -  & - & - &  1 & 54.0 \\
+         \cline{2-10}
+         & F & 108 & - & 115 & - & - & - &  {\bf 59} & - \\ 
+         \cline{1-10}
+        \multirow{4}{*}{\scriptsize{(4,5,5,6)}} & S & 35 & {\bf 1.8} & 19 & 6.6 & - & - &  {\bf 47} & 22.4 \\
+        \cline{2-10}
+         & U & 36 & 0.3 & 39 & 0.4 & - & - &  {\bf 72} & {\bf 0.1} \\
+         \cline{2-10}
+         & B & -  & -  & - & - & - & - &  0 & - \\
+         \cline{2-10}
+         & F & 129 & - & 142 & - & - & - &  {\bf 81} & -\\
+         \hline
+         \cline{1-10}
+        \multirow{3}{*}{\scriptsize{Flat}} 
+         & S & 34 & 19.0 & 88 & 12.7 & 5 & {\bf 0.1}& {\bf 115}  & 12.3\\
+        \cline{2-10}
+         & U & 0 & - & 1 & 4.0 & 182 & {\bf 2.5} & {\bf 182} & 47.7 \\
+         \cline{2-10}
+         & F & 266 & - & 211 & - & 113 & - & {\bf 3} & - \\ 
+         \cline{1-10}
+        \multirow{3}{*}{\scriptsize{Non-flat}}
+         & S & 210 & 7.8 & 144 & 4.9 & 55 & 5.9 &{\bf 300} & 16.7 \\
+        \cline{2-10}
+         & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+         & F & 90 & - & 156 & - & 245 & - & {\bf 0} & - \\ 
+         \cline{1-10}
+        \end{tabular}
+            \label{table:arithmetic}
+
+\end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
+    \begin{table*}[ht]
         \caption{Experimental Results, O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
@@
     \hline
-        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c|}{$\paexp$} \\
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c |}{$\paexp$} \\ 
         \cline{3-10}
@@
         \end{tabular}
-            \label{table:arithmetic}
+            \label{table:string}
+\end{table*}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[ht]
+    \caption{Experimental results on STRINGHASH,  O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
+    $\#$: number of problems, $T$: average time in seconds}
+    \centering
+
+    \renewcommand{\arraystretch}{1.1}
+    \begin{tabular}{cccccccccc}
+    \hline
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c}{Z3} & \multicolumn{2}{c}{CVC4} &  \multicolumn{2}{c}{Trau} & \multicolumn{2}{c}{$\paexp$} \\
+       &  & $\#$ & $T$ &  $\#$ & $T$ & $\#$ & $T$ &  $\#$ & $T$ \\ 
+       \hline
+       \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345(w_1)^+(w_2)^+$}} & S & 5 & 14.0 & 29 & 8.5 & 3 & {\bf  0.1} & {\bf 37} & 9.9 \\
+         \cline{2-10}
+        & U & 0 & - & 0 & - & {\bf 60} & {\bf 1.3} & {\bf 60} & 47.2 \\
+         \cline{2-10}
+        & F & 95 & - & 71 & - & 37 & - & {\bf 3} & - \\ 
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$\begin{array}{l}12345(w_1)^+ \\
+       \hspace{2mm}(w_2)^+6789\end{array}$}} & S & 11 & 13.0 & 29 & 12.0 & 0 & - & {\bf 37} & {\bf 10.6} \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & {\bf 63} & {\bf 1.2} & {\bf 63} & 50.0 \\
+         \cline{2-10}
+        & F & 89 & - & 71 & - & 37 & - & {\bf 0} & - \\ 
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$(w_1)^+(w_2)^+6789$}} & S & 18 & 24.0 & 30 & 9.3 & 2 & {\bf  0.1} & {\bf 41} & 16.1 \\
+      	 \cline{2-10}
+        & U & 0 & - & 1 & 4.0 & {\bf 59} & {\bf 2.5} & {\bf 59} & 45.8 \\
+         \cline{2-10}
+        & F & 82 &  & 69 &  & 39 &  & {\bf 0} &  \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}$}} & S & 82 & 8.7 & {\bf 100} & {\bf 2.2} & 28 & 5.9 & {\bf 100} & 18.5 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 18 & - & {\bf 0} & - & 72 & - & {\bf 0} & - \\ 
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}6789$}} & S & 60 & 9.3 & 17 & 7.8 & 3 & {\bf 0.3} & {\bf 100} & 16.0 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 40 & - & 83 & - & 97 & - & {\bf 0} & - \\
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$\Sigma^*_{\sf num}6789$}} & S & 68 & 5.5 & 27 & 13.0 & 24 & {\bf 9.0} & {\bf 100} & 15.7 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 32 & - & 73 & - & 76 & - & {\bf 0} &- \\
+        \hline
+       \end{tabular}
+           \label{table:string}
 \end{table}

### 2022-03-18 Acta Informatica2022/1_intro.tex
.

@@
 
-The emerging of scripting languages boosted the needs of better approaches and tools to ensure program quality.
+The emerging of scripting languages boosted the needs of efficient approaches and tools to ensure program quality.
 Comparing with traditional programming languages, string data type plays a more critical role in its analysis.
@@
 Satisfiability of word equations is a special instance of Hilbert’s 10th problem. 
-In the past, the original motivation of studying word equations was for finding an undecidability proof of Hilbert’s 10th problem. 
+In the past, the original motivation of studying word equations was to find an undecidability proof of Hilbert’s 10th problem. 
 However, the motivation is no longer valid since Makanin finds a decision procedure.
 Recently, driven by the need for program analysis, people started to revisit the problem and its extensions to describe the complete string library APIs in conventional programming languages.
-Many highly efficient solvers for string constraints, to name a few, CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, Z3-Str3~\cite{zheng2013z3}, S3~\cite{trinh2014s3}, Norn~\cite{abdulla2015norn}, Ostrich~\cite{chen2017decidable}, Sloth~\cite{sloth}, ABC~\cite{aydin2018parameterized}, Stranger~\cite{yu2010stranger}, and Trau~\cite{abdulla2018trau}, are developed in the last decade.
+Many highly efficient solvers for string constraints, to name a few, CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, Z3-Str3~\cite{zheng2013z3}, S3~\cite{trinh2014s3}, Norn~\cite{abdulla2015norn}, Ostrich~\cite{chen2017decidable}, Sloth~\cite{sloth}, ABC~\cite{aydin2018parameterized}, Stranger~\cite{yu2010stranger}, Trau~\cite{abdulla2018trau} and so on, are developed in the last decade.
 The satisfiability of \emph{string-integer conversion constraints}, e.g., $wx=yx \wedge \vert x\vert  > \parseInt(y)$, has been proven undecidable in~\cite{DayGHMN18}. 
@@
 This approach is taken in the first-generation solvers such as Hampi~\cite{KiezunGAGHE12} and Kuluza~\cite{SaxenaAHMMS10}.
-Although they are useful in handling many practical cases, they fail short when the all string solutions exceed the selected bound.
+Although they are useful in handling many practical cases, they fail to find an answer when the all string solutions exceed the selected bound.
 For example, a constraint of the form $x.y \neq z  \wedge \vert x\vert  > 2000$ would be quite challenging to handle using those solvers.
@@
 First, under the restriction, the potential solution space is still infinite, which gives us a higher potential of finding solutions.
-For instance, we can find a solution for $x.y \neq z  \wedge \vert x\vert  > 2000$ under a very simple restriction: all variables are in $a^*$, which $a$ is an element of the alphabet.
-Second, and more importantly, because we can convert the membership problem of a flat language to the satisfiability problem of a Presburger arithmetic formula, the class of word equation + flat languages + length constraints is decidable.
+For instance, we can find a solution for $x.y \neq z  \wedge \vert x\vert  > 2000$ under a very simple restriction: all variables are in $a^*$, where $a$ is an element of the alphabet.
+Second, more importantly, because we can convert the membership problem of a flat language to the satisfiability problem of a Presburger arithmetic formula, the class of word equations + flat languages + length constraints is decidable.
 
-The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string-integer conversion constraints to the above class.
-It proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
-For string-integer conversion, their approach projects the solution to string-integer conversion to
-%, in terms of numerical solutions, 
+The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string-integer conversion constraints to the above class,
+which proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
+For string-integer conversion constraints, their approach projects the solution to 
 a finite solution space (in a way similar to the PASS~\cite{goudong2013pass} approach).
@@
 In this paper, we give a complete solution to this problem.
-We propose a decision procedure for the class of word equation + flat languages + length constraints + string-integer conversion.
+We propose a decision procedure for the class of word equations + flat languages + length constraints + string-integer conversion.
 The basic idea of our approach can be sketched as follows: we first reduce the satisfiability problem to the corresponding satisfiability problem of the theory of Presburger Arithmetic with exponential functions (denoted by $\paexp$), more precisely, the existential fragment of $\paexp$; then, according to the decidability of $\paexp$, we obtain the decidability of the original satisfiability problem.
@@
 To remedy this, in \cite{Point86}, Point presented the first quantifier elimination procedure for the satisfiability of $\paexp$.
-Partially attributed to the fact that Point's procedure in \cite{Point86} was presented  in a mathematical and dense way, this quantifier elimination procedure has mostly eluded the attentions of computer science researchers\footnote{There are only two citations in Google Scholar.}.
-To the best of our knowledge, no implementation based on Point's procedure has been available up to now.
+Partially attributed to the fact that Point's procedure in \cite{Point86} was presented  in a mathematical and dense way, this quantifier elimination procedure has mostly eluded the attentions of computer science community\footnote{There are only two citations in Google Scholar.}.
+To the best of our knowledge, no implementation based on Point's procedure was available up to now.
 
-Aiming at introducing Point's procedure to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}).
+Aiming at introducing Point's procedure to computer science community, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science community (Section~\ref{sec-dec}).
 This can be seen as another contribution of this paper.
@@
 Here we illustrate a toy example that mimic the ``mining'' step of block-chain construction. 
-Essentially, given a string hash function ${\sf hash}(w)$, the goal of the mining step is to find a nonce $n$ such that when inserting $n$ to the text to be protected, say $w_1,w_2$, ${\sf hash}(w_1.n.w_2)$ satisfies a certain pattern, e.g., the last $k$ digits are zeros.
+Essentially, given a string hash function ${\sf hash}(w)$, the goal of the mining step is to find a nonce $n$ such that when inserting $n$ to the text to be protected, say $w_1$ and $w_2$, ${\sf hash}(w_1\cdot n\cdot w_2)$ satisfies a certain pattern, e.g., the last $k$ digits are zeros.
 If $w_1$ or $w_2$ are modified, one needs to compute another $n$ satisfies the desired pattern.
@@
 With the optimizations introduced in Section~\ref{sec-opt}, 
-our implementation manages to solve several variants of the string-hash examples as well as some randomly generated problem instances better than the state-of-the-art solvers (Section~\ref{sec-exp}).
+our implementation manages to solve several variants of the string-hash examples as well as some randomly generated arithmetic problem instances better than the state-of-the-art solvers (Section~\ref{sec-exp}).
 
 \paragraph*{Structure}
-After the preliminaries in Section~\ref{sec:pre}, we describe how to flatten a string constraint with string-integer conversion to an existential $\paexp$ formula in Section~\ref{sec:string-solving}.
-We present the quantifier elimination procedure for $\paexp$ in Section~\ref{sec-dec} and its optimizations in Section~\ref{sec-opt}.
+After the preliminaries in Section~\ref{sec:pre}, we show how to flatten a string constraint with string-integer conversion to an existential $\paexp$ formula in Section~\ref{sec:string-solving}.
+We describe the quantifier elimination procedure for $\paexp$ in Section~\ref{sec-dec} and analyse its complexity in Section~\ref{sec-cpx}.
+Several optimization techniques are presented in Section~\ref{sec-opt}.
 Finally, we describe the implementation and experiment results in Section~\ref{sec-exp}.

### 2022-03-18 Acta Informatica2022/2_preliminary.tex
.

@@
 
-An \emph{alphabet} $\Sigma$ is a finite set. Elements of $\Sigma$ are called \emph{letters}.
-%Let $\Sigma$ denote a finite alphabet,
+An \emph{alphabet} $\Sigma$ is a finite set. Each element of $\Sigma$ is called a \emph{letter}.
 A \emph{string} $w$ over $\Sigma$ is a (possibly empty) finite sequence $a_1\ldots a_n$ with $a_i \in \Sigma$ for every $i \in [n]$.
 Let $\varepsilon$ denote the empty string, namely, the empty sequence.
-For a string $w = a_1 \ldots a_n \in \Sigma^*$, let $|w|$ denote the \emph{length} of $w$, i.e.
-$n$.
-In particular, $\vert \varepsilon\vert  = 0$.
+Let $\Sigma^*$ denote the set of all strings over $\Sigma$.
+Let$\Sigma^+$ denote the set of nonempty strings over $\Sigma$. 
+For convenience, we also use $\Sigma_{\epsilon}$ to denote $\Sigma \cup \{\epsilon\}$.
+For a string $w = a_1 \ldots a_n \in \Sigma^*$, let $\len(w)$ denote the \emph{length} of $w$, i.e. $n$. In particular, $\len(\varepsilon)  = 0$.
 %
@@
 %
-Let $\Sigma^*$ denote the set of all strings over $\Sigma$ and $\Sigma^+$ denote the set of nonempty strings over $\Sigma$. 
-For convenience, we also use $\Sigma_{\epsilon}$ to denote $\Sigma \cup \{\epsilon\}$.
-%
 A language $L$ over $\Sigma$ is a subset of $\Sigma^*$.
 
+\hide{
 \paragraph{Presburger Arithmetic} \label{PA}
@@
 We use $\llbracket \phi \rrbracket$ to denote the set of models of $\phi$.
+}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-Presburger Arithmetic (PA) is a first order theory over the signature 
-$\Sigma_\mathbb{N}\Def  \{0,1,+, <, |\}$, where $0,1$ are constants, 
-$+$ is a binary function symbol, $<, |$ are  and $=$ is a binary predicate.
+\paragraph{Presburger Arithmetic} \label{PA}
+Presburger Arithmetic ({\pa}) is the first-order logic of integers with addition. 
+A term of {\pa}, denoted by $\iterm$, is of the form
+$$\iterm \ \Def \ c \mid \ivarx \mid \iterm+\iterm \mid \iterm - \iterm,$$
+where $\ivarx$ and $c$ represent integer variables and integer constants respectively. 
+A formula of {\pa}, denoted by $\phi$, is of the form
+$$\phi \ \Def \ \iterm \ \op\ \iterm \mid c \vert \iterm \mid c \nmid \iterm \mid \neg \phi \mid \phi \wedge \phi \mid \phi \vee \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,$$
+where $\op \in \{=, < , >, \le, \ge\}$.
 
-PA can be axiomatized by the following axioms \cite{PA}
+A \emph{quantifier-free} {\pa} (\qfpa) formula is a {\pa} formula containing no quantifiers. A {\pa} formula is called \emph{existential} if it contains no occurrences of universal quantifiers.
+The set of free variables of $\phi$ is denoted by $\free(\phi)$. 
+We usually write $\phi(\ivarx_1,\cdots, \ivarx_k)$ to denote an PA formula $\phi$ such that $\free(\phi) \subseteq \{\ivarx_1,\cdots, \ivarx_k\}$.
+Given an PA formula $\phi$, and an integer interpretation of $\free(\phi)$, i.e.
+a function $I: \free(\phi) \rightarrow \Int$, we denote by $I \models \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner, with $+$, $-$, $\vert$ and $\nmid$ interpreted as the integer addition, subtraction, divisibility and undivisibility relation respectively), and call $I$ a \emph{model} of $\phi$.
+We use $\llbracket \phi \rrbracket$ to denote the set of models of $\phi$.
 
-\begin{itemize}
-    \item $\forall x, \neg (x+1=0)$
-    \item $\forall x \forall y. x+1=y+1 \to x=y$
-    \item $F(0) \wedge (\forall x. F(x)\to F(x+1)) \to \forall x. F(x)$ 
-    \item $\forall x. x+0=x$
-    \item $\forall x \forall y. x+(y+1)=(x+y)+1$
-\end{itemize}
-Given the domain $\mathbb{N}$,
-the standard interpretation of PA interprets 
-$0,1$ to $0_\mathbb{N},1_\mathbb{N}\in \mathbb{N}$
-and $+,=$ to addition and equality over $\mathbb{N}$.
-We call a PA formula without quantifiers a quantifier-free PA formula.
+We always assume, without loss of expressiveness, that quantifier-free formulas contain no negations, because we can transfrom a formula into negation normal form and negations before atomic formulas can be absorbed by changing the predicates (for example, $\neg (\ivarx \ge \ivary)\equiv \ivarx < \ivary$).
+The only time when we deal with negations is to transform an universal quantifier $\forall$ into an existential one $\neg \exists \neg$.
 
-PA is a decidable theory, 
-and the complexity of decidability is related to 
-the number and locations of quantifiers.
-Generally, 
-the upper bound (on deterministic time and space) 
-for deciding a formula of length $n$ is $2^{2^{2^{p n log(n)}}}$,
-where $p>1$ is a constant\cite{Oppen69}. 
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+It is well-known that {\pa} admits quantifier elimination, for example, Cooper's algorihm\cite{Cooper72}. In Sec~\ref{sec-dec}, we will use this fact and apply Cooper's algorithm as a subprocedure.
 
 \paragraph{Finite state automata}
-A finite state automaton (FA) is a tuple 
+A \emph{finite state automaton} (FA) is a tuple 
 $\Aut=\langle Q,\Sigma,\Delta, q_{\textit{init}}, F\rangle$, 
@@
 A \emph{run} of $\Aut$ on a string $w = a_1 \ldots a_n$ is a sequence $q_0 \xrightarrow{b_1} q_1  \xrightarrow{b_2}  \ldots   \xrightarrow{b_{m-1}}  q_{m-1} \xrightarrow{b_m} q_m$ such that $q_0 = q_{\textit{init}}$, $(q_{i-1}, b_i, q_i) \in \Delta$ for every $i \in [m]$, and $a_1 \ldots a_n = b_1 \ldots b_m$.
-A \emph{run} $q_0 \xrightarrow{b_1} q_1  \xrightarrow{b_2}  \ldots   \xrightarrow{b_{m-1}}  q_{m-1} \xrightarrow{b_m} q_m$ is \emph{accepting} if $q_m  \in F$.
+We say a run $q_0 \xrightarrow{b_1} q_1  \xrightarrow{b_2}  \ldots   \xrightarrow{b_{m-1}}  q_{m-1} \xrightarrow{b_m} q_m$ is \emph{accepting} if $q_m  \in F$.
 A string $w$ is \emph{accepted} by $\Aut$ if there is an accepting run of $\Aut$ on $w$.
@@
 We will present flat languages. This is a high level, language-theoretical view at the flat automata from \cite{Parosh:20:PLDI}. 
-%
-%For integers $k$ and $\ell$ and a string variable $x$, 
-%we define the family of indexed \emph{character variables} $\charvars = \{x_j^i \mid 1 \le i \le k, 1 \le j \le \ell\}$. 
-A \emph{flat language} (FL)  over $\Sigma$
-%with the \emph{period} at most $\ell$ and the \emph{cycle count} $k$ 
-is the set of strings 
+A \emph{flat language} (FL)  over $\Sigma$ is the set of strings 
 that conform to a regular expression of the form

### 2022-03-18 Acta Informatica2022/3_flatten.tex
remove appendix
.

--- 2021-10-25 FMCAD2021/3_flatten.tex
@@
 Then we define the extension of Presburger arithmetic with exponential functions, denoted by $\paexp$.
-Finally, we show how to flatten the string constraints in $\strint$ into the arithmetic constraints in $\paexp$.
+Finally, we show how to flatten the string constraints in $\strint$ into the arithmetic constraints in the existential fragment of $\paexp$.
 
-\vspace*{-3mm}
 \subsection{String constraints with string-integer conversion ($\strint$)}
@@
 
-The function  {\parseInt} takes a decimal string as the input and returns the integer represented by the string\footnote{The {\parseInt} function in scripting languages, e.g.
+The function {\parseInt} takes a decimal string as input and returns the integer represented by the string
+\footnote{The {\parseInt} function in scripting languages, e.g.
 Javascript, is more general in the sense that the base can be a number between 2 and 36.
-Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for readability.}.
-%Since we focus on the decidability,
-%we define a binary version of {\parseInt},
-%which takes a binary string and returns a decimal integer number.
-For example,
-${\parseInt}($`$0123$'$) = {\parseInt}($`$123$'$)=10^2+10*2+3 = 123$. 
-Note here we use the quotation marks to delimit the strings.
-%Clearly,  our decision procedure given in this paper 
-%can be adapted to string constraints with other string to number conversion function without substantial change. 
-%Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
-%from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.
+Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for simplicity.}.
+For example, ${\parseInt}($`$0123$'$) = {\parseInt}($`$123$'$)=1*10^2+2*10+3 = 123$. 
+Note here we use the quotation marks to delimit strings.
 
 Formally, the semantics of the $\parseInt$ function is defined as follows. 
-%\begin{definition} 
 In order to simplify the presentation, we assume all string variables ranging over numerical symbols $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$. 
 Note that one can easily extend our approach to allow arbitrary finite alphabet.
-Then ${\parseInt}: \Sigma_{\textit{num}}^+ \mapsto \Nat$ is recursively defined by
-    for every $w\in \Sigma_{\textit{num}}^+$,
+Then ${\parseInt}: \Sigma_{\textit{num}}^+ \mapsto \Nat$ is recursively defined by, for every $w\in \Sigma_{\textit{num}}^+$,
 \begin{itemize}
-%        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
     \item if $w=$`$i$' for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}$$($`$i$'$)=i$;
-        \item for $w = w_1$`$i$' for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
-        ${\parseInt}(w) = 10*{\parseInt}(w_1)+{\parseInt}($`$i$'$)$.
+    \item for $w = w'\cdot$`$i$' for $i \in \Sigma_{\textit{num}}$ with $\len(w') \ge 1$, 
+        ${\parseInt}(w) = 10*{\parseInt}(w')+{\parseInt}($`$i$'$)$.
 \end{itemize} 
-%\end{definition}
 Note that $\parseInt$ is undefined with $\varepsilon$ as the input.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 In $\strint$, there are two types of variables, i.e. the string variables $\svarx,\svary,\ldots \in \svars$ and the integer variables $\ivarx,\ivary,\ldots \in \ivars$.
-%
 A $\strint$ formula $\varphi$ is defined by the following rules, where $\len(\sterm)$ denotes the length of a string $\sterm$.
-\vspace{-2mm}
 \[
 \begin{array}{l c l}
-\sterm & \Def & a \mid \svarx \mid \sterm \concat \sterm, \\
+\sterm & \Def & a \mid \svarx \mid \sterm \cdot \sterm, \\
 \iterm & \Def & n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,\\
-\vspace{-2mm}
 \varphi & \Def & \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi,
@@
 Let  $\svar(\varphi)$ and $\ivar(\varphi)$ denote the set of string variables and integer variables respectively occurring in $\varphi$.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-\paragraph{String Terms}
-Given a finite alphabet $\Sigma$ and 
-a finite set $X$ of string variables over $\Sigma^*$,
-we define the set of terms $\textit{Terms}(\Sigma,X)$ 
-to be the smallest set satisfying
-\begin{itemize}
-    \item[1] $\Sigma\cup \{\epsilon\} \cup X \subseteq \textit{Terms}(\Sigma,X)$;
-    \item[2] if $t_1,t_2\in \textit{Terms}(\Sigma,X)$, then $t_1 \cdot t_2 \in \textit{Terms}(\Sigma,X)$.
-\end{itemize}
-
-We extend $I_X$ to $\textit{Terms}(\Sigma,X)$ by letting $I_X(\epsilon)=\epsilon$, 
-for $a\in \Sigma, I_X(a)=a$,
-and $I_X(t_1\cdot t_2)= I_X(t_1)\cdot I_X(t_2)$.
-
-\paragraph{String Constraints} \label{par: string constraints}
-Given a constraint $\phi$ and an interpretation $I$,
-$I\models \phi$ denotes that $I$ satisfies $\phi$,
-and $I$ is called a \emph{model} of $\phi$.
-We use $\lVert \phi\rVert$ to denote the set of all models of $\phi$.
-
-We define the following three forms of atomic string constraints:
-\begin{itemize}
-\item An equality constraint $\phi_e$ is of the form 
-$t_1 = t_2$, where $t_1,t_2\in \textit{Terms}(\Sigma,X)$.
-We define $\lVert \phi_e \rVert = \{I\mid I(t_1)=I(t_2)\}$.
-Inequality constraints can be defined analogously.
-
-\item A regular constraint $\phi_r$ is of the form 
-$x\in L(\mathcal{A})$,
-where $x\in \mathbb{X}$ and $\mathcal{A}$ is a finite state automaton.
-We define $\lVert \phi_r \rVert = \{I\mid I(x)\in L(\mathcal{A})\}$.
-
-\item A length constraint $\phi_l$ is a linear constraint over 
-$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
-where $|\cdot |$ is the length function.
-We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
-\end{itemize}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In $\strint$, there are two types of variables, i.e. the string variables $\svarx,\svary,\ldots \in \svars$ and the integer variables $\ivarx,\ivary,\ldots \in \ivars$.
+The syntax of $\strint$ is defined as follows:
+a string term, denoted by $\sterm$, is of the form 
+$$\sterm\  \Def\  a \mid \svarx \mid \sterm \cdot \sterm,$$
+an integer term, denoted by $\iterm$, is of the form
+$$\iterm\  \Def\  n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,$$
+and a $\strint$ formula, denoted by $\varphi$ is of the form 
+$$\varphi\ \Def\ \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi$$
+where $a \in (\Sigma_{\textit{num}})_\varepsilon$, $n \in \Nat$, $\Aut$ is an FA, $\op \in \{=, < , >, \le, \ge\}$ and $\len(\sterm)$ denotes the length of a string $\sterm$. 
+Let us call $\sterm = \sterm$ as string equality constraints, $x \in \Aut$ as regular constraints, $\iterm \op \iterm$ as arithmetic constraints. 
+Let  $\svar(\varphi)$ and $\ivar(\varphi)$ denote the set of string variables and integer variables occurring in $\varphi$ respectively.
 
@@
 
-\begin{example}
-$\svarx \in \Aut \wedge \parseInt(\svarx) = 109\ivarx \wedge \len(\svarx) < 100$, where $\Aut$ is an FA, is an $\strint$ formula.
+\begin{example} Given a FA $\Aut$, the constraint
+$$\svarx \in \Aut \ \wedge\ 
+\parseInt(\svarx) = 109\ivarx\ \wedge\ 
+\len(\svarx) < 100$$
+is a $\strint$ formula.
 \end{example}
-
-%As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
-%essentially a pair of two mappings 
-%$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
-%where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.
 
@@
 whether $\lVert  \varphi \rVert \neq \emptyset$.
-%if not, to compute an interpretation $I$ that satisfies $\Psi$.
 
-\vspace*{-5mm}
 \subsection{An Extension of Presburger Arithmetic with Exponential Functions (\paexp)}
 
-{\paexp} extends Presburger arithmetic with the exponential function $10^x$ as well as the (partial) functions $\ell_{10}(\ivarx)$ (cf.
+{\paexp} extends Presburger arithmetic with two partial functions, the \emph{exponential} function $10^\ivarx$ and the \emph{integer logarithmic} function $\ell_{10}(\ivarx)$ (cf.
 \cite{Point86}).
-The function $\ell_{10}(\ivarx)$ is inductively defined as follows.
-For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$.
-(Note that $\ell_{10}(0)$ is undefined.)
-%and $\lambda_{10}(\ivarx)$ are inductively defined as follows.
-%\begin{itemize}
-%\item The (partial) function $\ell_{10}(\ivarx)$ is\hide{inductively} defined as follows: For $n \ge 1$, $\ell_{10}(n) = m$ iff $10^m \le n < 10^{m+1}$. (Note that $\ell_{10}(0)$ is undefined.) 
-%
-%\item The (partial) function $\lambda_{10}(\ivarx)$ can be defined by $\ell_{10}(\ivarx)$: For $n \ge 1$, $\lambda_{10}(n) = 10^{\ell_{10}(n)}$.
-(Intuitively, $\lambda_{10}(\ivarx)$ is the maximum power of 10 that is no greater than $\ivarx$.)
-%\end{itemize}
+The function $\ell_{10}(\ivarx)$ is defined as follows: for $n \ge 1$, $\ell_{10}(n) = m$ if $10^m \le n < 10^{m+1}$.
+Note that $10^\ivarx$ is undefined for $\ivarx<0$ and $\ell_{10}(\ivarx)$ is undefined for $\ivarx\le 0$.
 
-The syntax of {\paexp} is obtained from that of {\pa} by adding $10^\iterm$ and $\ell_{10}(\iterm)$ to the definition of integer terms. Specifically, {\paexp} formulas are defined by the rules,
-%
-\vspace{-1mm}
- $$
- \begin{array}{ l c l}
- \iterm &\Def& c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) \\
-\vspace{-1mm}
- \phi &\Def & \iterm \ \op\ \iterm \mid c | \iterm \mid \phi \wedge \phi \mid \phi \vee \phi \mid \neg \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi.
- \end{array}
- $$
- The semantics of {\paexp} are defined similarly to {\pa} with the difference that variables are interpreted over $\Nat$ (to avoid terms like $10^{-2}$), that is, an interpretation is a function $I: \free(\phi) \rightarrow \Nat$.
+The syntax of {\paexp} is obtained from that of {\pa} by adding $10^\iterm$ and $\ell_{10}(\iterm)$ to the definition of terms. An {\paexp} term, denoted by $\iterm$, is of the form
+$$\iterm\ \Def\ c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) $$
+where $\ivarx$ is an integer variable and $c\in \mathbb Z$ is an constant integer.
+
+A formula of {\paexp}, denoted by $\phi$, is of the form
+$$\phi \ \Def \ \iterm \ \op\ \iterm \mid c \vert \iterm \mid c \nmid \iterm \mid \neg \phi \mid \phi \wedge \phi \mid \phi \vee \phi \mid \exists \ivarx.\ \phi \mid \forall \ivarx.\ \phi,$$
+where $\op \in \{=, < , >, \le, \ge\}$.
+
+The semantics of {\paexp} are defined similarly to {\pa} with the only difference that variables are interpreted over $\Nat$ rather than $\mathbb Z$, that is, an interpretation is a function $I:\free(\phi)\to \mathbb \Nat$.
+This restriction ensures that terms $10^\iterm$ and $\ell_{10}(\iterm)$ are well-defined: In the \emph{normalization} step of the quantifier elimination procedure, for each $\ell_{10}(\iterm)$ term, we will replace all its occurrences by a fresh variable $\ivarz$, and add a conjunct $10^\ivarz \le \iterm < 10*10^\ivarz$.
+Then $\ivarz\in\Nat$ guarantees that $\iterm \ge 1$ is always satisfied.
+Exponential functions are treated similarly
+(more details can be found in Section~\ref{qe-norm}).
+Note that we can extend the interpretation to $\mathbb Z$ since each integer can be written as the difference of two positive integers.
+
 The quantifier-free and existential {\paexp} formulas are defined similarly to {\pa} as well.
+We also assume that quantifier-free {\paexp} formulas are free of negations, for the same reason we have explained for {\pa}.
+
+For convenience, when working with exponential and logarithmic functions, we use the notation $\lambda_{10}(\iterm)$ to denote $10^{\ell_{10}(\iterm)}$.
+It is easy to show that for all $n \ge 1$, $\lambda_{10}(n) \le n < 10\lambda_{10}(n)$ holds.
 
 \begin{example}
-$10^{\ivarx+ \ivary} + 3 * 10^{\ivary}  = 2 \ivarz +1 $ is an {\paexp} formula.
+$10^{10^\ivarx}+10^{\ivarx+ \ivary} + 3\ivary    = 2 \ivarz +1 $ is an {\paexp} formula.
 \end{example}
 
-%Moreover, we introduce an abbreviation $\lambda_{10}(\ivarx) \equiv 2^{\ell_{10}(\ivarx)}$.
-We will use the notation $\lambda_{10}(\iterm)$ to denote $10^{\ell_{10}(\iterm)}$. It is easy to show that for all $n \ge 1$, $\lambda_{10}(n) \le n < 10\lambda_{10}(n)$ holds.
-
-\vspace*{-2mm}
 \subsection{Flattening $\strint$ into $\paexp$}
@@
 
-A \emph{flat domain restriction} for a string constraint $\varphi$ is a function $\flatfun_\varphi$ that maps each string variable $\svarx \in \svar(\varphi)$ to a flat language $(w_{\svarx,1})^* \cdots (w_{\svarx, k_\svarx})^*$, where $w_{\svarx, i} \in \Sigma_{\textit{num}}^+$ for every $i \in [k_\svarx]$.
-%
+A \emph{flat domain restriction} for a string constraint $\varphi\in \strint$ is a function $\flatfun_\varphi$ that maps each string variable $\svarx \in \svar(\varphi)$ to a flat language $(w_{\svarx,1})^* \cdots (w_{\svarx, k_\svarx})^*$, where $w_{\svarx, i} \in \Sigma_{\textit{num}}^+$ for every $i \in [k_\svarx]$.
 The flattened semantics of $\varphi \in \strint$ is defined as $\llbracket \varphi \rrbracket_{\flatfun_\varphi} = \{(I_s, I_i) \in \llbracket \varphi  \rrbracket \mid \forall \svarx \in \svar(\varphi).\ I_s(x) \in  \flatfun_\varphi(\svarx)\}$.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 The flattening of $\varphi \in \strint$ under a flat domain restriction $\flatfun_\varphi$ is a {\paexp} formula, denoted by $\flatten_{\flatfun_\varphi}(\varphi)$, that encodes its flattened semantics.
-%
 More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and flattening variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$, plus some other auxiliary variables, such that
-%
-$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)}).$$
-%
-The decoding function above decodes an interpretation of integer and flattening variables $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \Nat$ as a set $\decode_{\flatfun_\varphi}(I_e)$ of interpretations of the $\varphi$'s integer and string variables $(I_s, I_i)$ with $I_s: \svar(\varphi) \rightarrow \Sigma^*$ and $I_i: \ivar(\phi) \rightarrow \Nat$ such that
+$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket \vert_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)}).$$
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Under a flat domain restriction $\flatfun_\varphi$, the flattening of $\varphi \in \strint$ is an $\paexp$ formula, denoted by $\flatten_{\flatfun_\varphi}(\varphi)$, that encodes the flattened semantics $\llbracket \varphi \rrbracket_{\flatfun_\varphi}$.
+The idea is that $x$ can be expressed as $(w_{\svarx,1})^{\#_{x,1}} \cdots (w_{\svarx, k_\svarx})^{\#_{x,k_x}}$ with $\#_{x,i}$ denoting the number of repeatitions of $w_{x,i}$,  
+so we can eliminate the occurrences of $x$ in $\varphi$ by introduce a set of integer variables $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$. 
+More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and flattening variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$ plus some other auxiliary variables, such that the interpretations $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \mathbb Z$ of $\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket$and interpretations $(I_s, I_i)$ of $\llbracket \varphi \rrbracket_{\flatfun_\varphi}$ have the following correspondence
+
 \begin{itemize}
@@
 
-The formula $\flatten_{\flatfun_\varphi}(\varphi)$ is constructed inductively by following the structure of $\varphi$: $\flatten_{\flatfun_\varphi}(\varphi_1\ \mathfrak{o}\ \varphi_2) = \flatten_{\flatfun_\varphi}(\varphi_1) \ \mathfrak{o}\  \flatten_{\flatfun_\varphi}(\varphi_2)$, where $\mathfrak{o} \in \{\wedge, \vee\}$, and $\flatten_{\flatfun_\varphi}(\neg \varphi_1) = \neg \flatten_{\flatfun_\varphi}(\varphi_1)$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
+$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket \vert_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)}).$$
+
+The decoding function above decodes an interpretation of integer and flattening variables $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \mathbb Z$ as a set $\decode_{\flatfun_\varphi}(I_e)$ of interpretations of the $\varphi$'s integer and string variables $(I_s, I_i)$ with $I_s: \svar(\varphi) \rightarrow \Sigma^*$ and $I_i: \ivar(\phi) \rightarrow \Nat$ such that
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The formula $\flatten_{\flatfun_\varphi}(\varphi)$ is constructed inductively on the structure of $\varphi$: $\flatten_{\flatfun_\varphi}(\varphi_1\ \mathfrak{o}\ \varphi_2) = \flatten_{\flatfun_\varphi}(\varphi_1) \ \mathfrak{o}\  \flatten_{\flatfun_\varphi}(\varphi_2)$, where $\mathfrak{o} \in \{\wedge, \vee\}$, and $\flatten_{\flatfun_\varphi}(\neg \varphi_1) = \neg \flatten_{\flatfun_\varphi}(\varphi_1)$.
 Therefore, it is sufficient to show how to construct $\flatten_{\flatfun_\varphi}(\varphi)$ for atomic constraints $\varphi$.
 In the sequel, we will show how to construct $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ where $\parseInt(\sterm)$ may occur in $\iterm_1$ or $\iterm_2$.
-The construction of $\flatten_{\flatfun_\varphi}(\varphi)$ for the other atomic constraints is essentially the same as that in \cite{Parosh:20:PLDI} and thus omitted.
+The construction of $\flatten_{\flatfun_\varphi}(\varphi)$ for the other atomic constraints is omitted because it is essentially the same as that in \cite{Parosh:20:PLDI} (we summarize as Theorem~\ref{thm:string}).
+
+\begin{theorem}\label{thm:string}\cite{Parosh:20:PLDI} The satisfiability of Boolean combinations of string equality constraints, regular constraints and arithmetic constraints under flat domain restrictions can be reduced to the satisfiability of existential {\pa}   formulas, thus is decidable.
+\end{theorem}
 
@@
 \item 
-$\flatten_{\flatfun_\varphi}(\parseInt(\svarx)) \Def \iterm_{\svarx,1}$  such that $(\iterm_{\svarx, i})_{i \in [k_\svarx]}$ and $(\ell_{\svarx, i})_{i \in [k_\svarx]}$ are inductively defined as follows:
+$\flatten_{\flatfun_\varphi}(\parseInt(\svarx)) \Def \iterm_{\svarx,1}$  such that $(\iterm_{\svarx, i})_{i \in [k_\svarx]}$ are inductively defined as follows: 
 \begin{itemize}
 \item for $i = k_\svarx$, 
-$$ \hspace{-4mm} \iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{|w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx} } -1 )}{(10^{|w_{\svarx, k_\svarx}|} -1 )}$$ 
-and $\ell_{\svarx, i} = |w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx}$,
+$$\iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{\len(w_{\svarx,k_\svarx})\#_{\svarx, k_\svarx}} -1 )}{(10^{\len(w_{\svarx, k_\svarx}) } -1 )}$$ 
 %
@@
 %
-$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{|w_{\svarx, i} | \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{|w_{\svarx, i}|} -1 )} + \iterm_{\svarx, i+1}$$
+$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{\len(w_{\svarx, i})  \#_{\svarx, i} } -1 ) 10^{\big(\sum_{i+1\le j\le k_x} \len(w_{\svarx, j})  \#_{\svarx, j}\big)}} {(10^{\len(w_{\svarx, i}) } -1 )} + \iterm_{\svarx, i+1}$$
+\end{itemize}
 %
-and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
-%
-\end{itemize}
-Notice that here $|w_{\svarx, \textunderscore } |$ and $\ell_{\svarx, \textunderscore}$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables.
-%
-\item $\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} |w_{\svarx, i} | \#_{\svarx, i}$. 
+\item Notice that here $\len( w_{\svarx, \textunderscore } )$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables. So we have 
+$$\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} \len(w_{\svarx, i})  \#_{\svarx, i}.$$ 
 \end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-Before presenting the construction of $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$, for every $\svarx \in \svar(\varphi)$,
-
-Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by the following procedure.
-\begin{enumerate}
-\item Replace each integer term $\len(t)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by 
- $\sum \limits_{i \in [m]} \theta_i$, where for every $i \in [m]$, $\theta_i = 1$ if $\alpha_i \in \Sigma_{\textit{num}}$ and $ \theta_i = \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j} | \#_{\alpha_i, j}$ otherwise.
-%
-\item  Replace each integer term $\parseInt(\sterm)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by $\iterm_{\parseInt(\sterm)}$, where $\iterm_{\parseInt(\sterm)} = \iterm_{t,1}$ such that $(\iterm_{t,i})_{i \in [m]}$ and $(\ell_{t, i})_{i \in [m]}$ are inductively defined as follows:
-\begin{itemize}
-\item if $\alpha_m \in \Sigma_{\textit{num}}$, then $\iterm_{t, m} = \alpha_m$ and $\ell_{t, m} = 1$, otherwise, $\iterm_{t, m} = \flatten_{\flatfun_\varphi}(\parseInt(\alpha_m))$ and $\ell_{t, m} = \sum \limits_{j \in [k_{\alpha_m}]} |w_{\alpha_m, j}| \#_{\alpha_m, j}$,
-%
-\item for $i \in [m-1]$, if $\alpha_i \in \Sigma_{\textit{num}}$, then $\iterm_{t, i} = \alpha_i 10^{\ell_{t, i+1}} + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1}+1$, otherwise, $\iterm_{t, i} = 
-\flatten_{\flatfun_\varphi}(\parseInt(\alpha_i))10^{\ell_{t, i+1}}  + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1} + \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j}| \#_{\alpha_i, j}$.
-\end{itemize}
-%
-\item Let $\iterm'_1 \op \iterm'_2$ be the formula resulting from the aforementioned replacements.
-(Note that strictly speaking, $\iterm'_1 \op \iterm'_2$ is not a $\paexp$ formula since it contains divisions.) Let $N$ be the least common multiplier of the denominators of $\iterm'_1$ and $\iterm'_2$.
-Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained by multiplying the both sides of $\iterm'_1 \op \iterm'_2$ with $N$,  so that the division operator disappears.
-\end{enumerate}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The following example helps to illustrate the main idea of the flattening technique.
 
@@
 \]
-%Take $n={\parseInt}((11)^a(10)^b)$ for example.
-%\begin{align}
-%    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
-%    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
-%    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
-%\end{align}
 \end{example}
 
-By the above reduction, and the decidability of $\paexp$'s satisfiability (see Theorem~\ref{thm-paexp}), we have 
+By Theorem~\ref{thm:string}, the above reduction, and the decidability of $\paexp$'s satisfiability (see Theorem~\ref{thm-paexp}), we have 
 \begin{theorem} \label{thm:string-parInt}
-    \wuhao{The satisfiability of  $\strint$ under flat domain restrictions is equivalent to the satisfiability of existential $\paexp$ formulas, and thus is decidable. }
+    The satisfiability of  $\strint$ under flat domain restrictions can be reduced to the satisfiability of existential $\paexp$ formulas, and thus is decidable.
 \end{theorem}

### 2022-03-18 Acta Informatica2022/4_dec_proc.tex
.

@@
 Remedying this, Point proposed a quantifier elimination procedure for {\paexp} in \cite{Point86}.
+In this section, we describe how Point's quantifier elimination procedure works.
 
-In this section, we describe how Point's quantifier elimination procedure works. 
 %
@@
 \end{theorem}
-%
+
 Compared to \cite{Point86}, the presentation here is more accessible to computer science researchers.
 Moreover, the procedure presented here slightly improves Point's procedure, in the following two aspects: 1) DNF (disjunctive normal form) was required in Point's procedure, which is not required here, 2) in Point's procedure, the divisibility constraints produced by the elimination of linear occurrences of a variable should be converted to equality constraints (by introducing fresh variables) before the elimination of exponential occurrences of some other variable, which is unnecessary here, since the divisibility constraints are directly dealt with in the elimination of exponential occurrences of variables.
-%
-
-However, in the next section we analyse the complexity of Point's procedure to be 3-EXPSPACE, which  is quite expensive and a faithful implementation would not scale\footnote{We did implement Point's algorithm and discovered that the implementation could only solve formulas of very small size.}.
-So in Section~\ref{sec-opt}, we will propose various optimizations to Point's algorithm, aiming at an efficient implementation.
-
-\smallskip
 
@@
 \ \varphi$ is equivalent to $\neg \exists \ivarx.
-\ \neg \varphi$,  thus in the sequel, we only need to  show that every $\paexp$ formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
+\ \neg \varphi$, to prove Theorem~\ref{thm-paexp}, we only need to show that every existentially quantified $\paexp$ formula $\exists \ivarx.\ \varphi \in \paexp$, where $\varphi$ is quantifier-free, can be transformed into an equivalent quantifier-free formula $\varphi' \in \paexp$.
 
-% when the quantifier elimination problem of $\paexp$  can be easily reduced to the above special case.
+An informal sketch of Point's procedure is presented in Algorithm~\ref{qe-algo}.
 
-% such that $\free(\varphi') = \free(\varphi) \setminus \{\ivarx\}$. 
-%From this, we can easily derive that every {\paexp} formula $\varphi$ can be transformed effectively into an equivalent quantifier-free Presburger arithmetic formula $\varphi'$.
+\hide{Before a formal description of the quantifier elimination procedure, let us use a simple example to illustrate the main idea.}
 
-Before a formal description of the quantifier elimination procedure, let us use a simple example to illustrate the main idea and give an overview of the procedure.
+\begin{algorithm2e}[ht]\label{qe-algo}
+    \SetAlgoLined
+    \KwIn{{\paexp} formula $\exists \ivarx_1.\varphi$, where $\varphi$ is quantifier-free}
+    \KwOut{an equivalent quantifier-free formula}
+    \SetKwFor{ForOrder}{For each}{do}{end}
+    \SetKwRepeat{Do}{do}{while}
 
-%\vspace{-4mm}
+    $\varphi \gets \textrm{normalize } \varphi \textrm{ w.r.t }\ivarx_1 $ (Subsection~\ref{qe-norm});
 
-%\begin{example}
+    \tcp{suppose fresh variables $\ivarx_2,\cdots,\ivarx_n$ are introduced}
+    $\mathcal S_n\gets$ enumerate linear orders among $\ivarx_1,\cdots,\ivarx_n$ (Subsection~\ref{qe-enum})\;
+    \ForOrder{$\sigma\in \mathcal S_n$}
+    {
+        $\varphi'_\sigma  \gets \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$\; 
+        $i\gets n$\;
+        \For{$i> 0$}{
+            eliminate exponential occurrences of $\ivarx_{\sigma(i)}$ in $\varphi'_\sigma$ (Subsection~\ref{qe-elim-exp})\;
+            eliminate linear occurrences of $\ivarx_{\sigma(i)}$ in $\varphi'_\sigma$ (Cooper's QE algorithm)\;
+            \tcp{variable $\ivarx_{\sigma(i)}$ is removed}
+        $i\gets i-1$\;
+        }
+        \tcp{all quantified variables in $\varphi'$ are removed}
+        $\varphi''_\sigma\gets \varphi'_\sigma$\;
+    }
+
+    \Return $\bigvee_{\sigma\in\mathcal S_n}\varphi''_\sigma$\;
+
+    \caption{Quantifer elimination for \paexp}
+\end{algorithm2e}
+
+\hide{
 \subsection{An overview of the quantifier elimination procedure}
-Consider $\varphi \Def \exists \ivarx_2.\ 10^{\ivarx_1 + \ivarx_2} - 10^{\ivarx_2} \le \ivary + 1001$.
 
-At first, we \emph{normalize} $\varphi$ by introducing a fresh variable $\ivarx_3$ for $\ivarx_1 + \ivarx_2$ and get the formula 
+Consider the formula $\varphi \Def \exists \ivarx_2.\ 10^{\ivarx_1 + \ivarx_2} - 10^{\ivarx_2} \le \ivary + 1001$.
+
+At first, we \emph{normalize} $\varphi$ w.r.t. $\ivarx_2$ by introducing a fresh variable $\ivarx_3$ for $\ivarx_1 + \ivarx_2$ and get the formula 
 $$\varphi_1 \Def \exists \ivarx_3 \exists \ivarx_2.\ 10^{\ivarx_3} - 10^{\ivarx_2} \le \ivary + 1001 \wedge \ivarx_3 = \ivarx_1 + \ivarx_2.$$
@@
 %\vspace*{-3mm}
-\subsection{Normalization}
+
+}
+\subsection{Normalization}\label{qe-norm}
 
@@
 \begin{enumerate}
+
+\hide{ % do not allow \neg in PAexp
 \item \textit{NNF transformation} At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form). Moreover, we remove the occurrences of $\neg$ by replacing 
-\hide{(a) replacing $\neg c \vert  \iterm$ with $\bigvee \limits_{j \in [c-1]} c \vert  (\iterm+j)$, }
 (a) $\neg c \vert  \iterm$ with $ c \nmid \iterm$, 
@@
 (d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
-%
+}
+
 \item \textit{replace $\ell_{10}(\iterm)$ terms} Repeat the following procedure, until there are no $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: for each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct.
 Note that if $\iterm$ contains no variables, then $\ell_{10}(\iterm)$ is a constant.
-In this case, we can also assume that $\iterm$ contains $\ivarx$ and perform the same replacements, which helps in the analysis of complexity in \ref{app:cpx}.
+In this case, we can also assume that $\iterm$ contains $\ivarx$ and perform the same replacements, which helps in the analysis of complexity in \ref{sec-cpx}.
 Let the resulting formula be $\varphi''$.
@@
 
-\item \textit{flatten $10^\iterm$ terms} Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $10^{\iterm}$ in $\varphi''$, such that $\iterm$ contains $\ivarx$ but is not $\ivarx$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $10^{\iterm}$ by $10^\ivarz$, moreover, add the constraint $\ivarz = \iterm$ as a conjunct.
+\item \textit{flatten $10^\iterm$ terms} Then repeat the following procedure to $\varphi''$, until for each occurrence of $10^{\iterm}$ with $\ivarx$ occurs in $\iterm$, we have $\iterm = \ivarx$: For each occurrence of the $10^{\iterm}$ in $\varphi''$, such that $\iterm$ contains $\ivarx$ but is not equal to $\ivarx$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $10^{\iterm}$ by $10^\ivarz$, moreover, add the constraint $\ivarz = \iterm$ as a conjunct.
 Let $\varphi'''$ denote the resulting formula.
 
-\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$ ,$c \vert  \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
+\item \textit{$\le$ transformation} Do the following replacements to $\varphi'''$, so that all the atomic formulas in $\varphi^\dag$ are of the form $\iterm_1 \le \iterm_2$, $c \vert  \iterm$ or $c \nmid \iterm$: Replace every occurrence of $\iterm_1 \ge \iterm_2$ with $\iterm_2 \le \iterm_1$.
 Replace every occurrence of $\iterm_1 < \iterm_2$ (resp.
@@
 Replace ever occurrence of $\iterm_1 = \iterm_2$ wtih $\iterm_2 \le \iterm_1 \wedge \iterm_1 \le \iterm_2$.
-Let $\varphi^\dag$ the resulting formula.
+Let $\varphi^\dag$ denote the resulting formula.
 
@@
 
-Intuitively, the normalization step first absorbs all negation operators by transforming the formula into NNF.
-Then it removes the occurrences of $\ell_{10}(\iterm)$ where $\ivarx$ occurs in $\iterm$, by encoding them with the exponential function.
+Intuitively, the normalization step first removes all occurrences of $\ell_{10}(\iterm)$ where $\ivarx$ occurs in $\iterm$, by encoding them with the exponential function.
 Moreover, for each occurrence of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, it introduces a fresh variable $\ivarz$, replaces $10^\iterm$ with $10^\ivarz$, and adds the equality $\ivarz = \iterm$.
@@
 
-After the normalization, the resulting formula is of the following shape: 1) it is in NNF (negation normal form),  2) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 3)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 4) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$ or $c \vert  \iterm$.
-Denote the negation of $c \vert  \iterm$ by $c\nmid t$, then the formula contains no negation symbol.
+After the normalization, the resulting formula is of the following shape: 1) it contains no occurrences of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, 2)  it contains no occurrences of $10^\iterm$ such that $\ivarx$ occurs in $\iterm$, but $\iterm \neq \ivarx$, 3) all the atomic formulas are of the form $\iterm_1 \le \iterm_2$, $c \vert  \iterm$ or $c\nmid \iterm$.
 
 %\vspace{-2mm}
-\subsection{Enumeration of the variable orders}
+\subsection{Enumeration of the variable orders}\label{qe-enum}
 
-Suppose $n-1$ fresh variables are introduced in the normalization procedure, rename the original variable $\ivarx$ and the $n-1$ introduced variables $\ivarx_i,1\le i\le n$.
+Suppose $n-1$ fresh variables are introduced in the normalization procedure, rename the original variable $\ivarx$ and the $n-1$ introduced variables as $\ivarx_i,1\le i\le n$.
 Let the output of the normalization procedure be $\exists \vec{\ivarx}.\ \varphi'$ with $\vec{\ivarx} = (\ivarx_1,\ldots, \ivarx_n)$. 
@@
 
-Assuming a linear order $\sigma \in \mathcal{S}_n$ of $\{\ivarx_1,\ldots, \ivarx_n\}$, we then consider $\varphi'_\sigma  = \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ and eliminate the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(1)}$,  one by one and from $\ivarx_n$ to $\ivarx_1$.
+Assuming a linear order $\sigma \in \mathcal{S}_n$ of $\{\ivarx_1,\ldots, \ivarx_n\}$, we then consider $\varphi'_\sigma  = \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$ and eliminate the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(1)}$,  one by one and from $\ivarx_{\sigma(n)}$ to $\ivarx_{\sigma(1)}$.
+To eliminate the $i$-th quantifer $\exists \ivarx_{\sigma(i)}$, we first eliminate all its exponential occurrences, i.e.
+$10^{\ivarx_{\sigma(i)}}$, using the fact that $\ivarx_{\sigma(i)}$ is the largest among the remaining quantified variables (see Lemma~\ref{lem:exp-ineq}).
+Then linear terms of $\ivarx_{\sigma(i)}$ are eliminated further by applying Cooper's algorithm so that the quantifier $\exists \ivarx_{\sigma(i)}$ can be removed.
 Let $\varphi''_\sigma$ denote the resulting formula.
@@
 
+\hide{
 In the sequel, assuming a linear order $\sigma \in \mathcal{S}_n$, for $i \in [n]$, let $\exists \ivarx_{\sigma(1)} \ldots \exists \ivarx_{\sigma(i)}.\ \varphi''_{\sigma,i}$ be the formula obtained from $\varphi'_\sigma$ by eliminating the quantifiers $\exists \ivarx_{\sigma(n)}$, $\ldots$, $\exists \ivarx_{\sigma(i+1)}$, we show how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\exists \ivarx_{\sigma(1)} \ldots \exists \ivarx_{\sigma(i)}.\ \varphi''_{\sigma,i}$.
@@
 The elimination of linear occurrences of $\ivarx_{\sigma(i)}$ is essentially the quantifier elimination algorithm of {\pa}.
-%$\varphi'_\sigma =  \exists \vec{\ivarx}.\ \varphi' \wedge \bigwedge \limits_{i \in [n-1]} \ivarx_{\sigma(i)} \le \ivarx_{\sigma(i+1)}$.
 
-Note that the order $\ivarx_{\sigma(i)} \ge \ldots \ge \ivarx_{\sigma(1)}$ guarantees the maximality of $\ivarx_{\sigma(i)}$ among $\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}$, which is essential for the elimination of $10^{\ivarx_{\sigma(i)}}$ from $\varphi''_{\sigma,i}$ (see Lemma~\ref{lem:exp-ineq}).
+Note that the order $\ivarx_{\sigma(i)} \ge \ldots \ge \ivarx_{\sigma(1)}$ guarantees the maximality of $\ivarx_{\sigma(i)}$ among $\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}$, which is essential for the elimination of $10^{\ivarx_{\sigma(i)}}$ from $\varphi''_{\sigma,i}$ (see Lemma~\ref{lem:exp-ineq}).}
 
-%\yfc{Explain why only the largest variable can  be eliminated. Which part of the procedure we need this condition? Or if we do not have the order, what can be wrong?}
-
-%\vspace*{-3mm}
-\subsection{Elimination of  exponential occurrences of variables}\label{sec-elim-exp}
+\subsection{Elimination of exponential occurrences of variables}\label{qe-elim-exp}
 
@@
 $c \ \big  \vert  \ \big(a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})\big)$
-(or $\nmid$).
+(or $\nmid$), where $\iterm(\vec{\ivary})$ collects all terms in the formula without $\ivarx_i$.
 
 \subsubsection{Inequality atoms}
-In the sequel, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for these inequality atomic formulas. Let us consider 
-$$
-\begin{array}{l}
-\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-%\hspace{4mm} 
-a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
-\end{array}
-$$
-%
+In the following, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for inequality atomic formulas of the form 
+\begin{equation}\label{tau}
+    \begin{aligned}
+        &\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
+        &a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary})
+    \end{aligned}            
+\end{equation}
+where $a_i\neq 0$, $\iterm(\vec \ivary)$ is the sum of all other terms without $\ivarx_{\sigma(i)}$.
+
 The elimination of the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ relies on the following lemma. 
-Intuitively, the lemma states the fact that if $\ivarx_{\sigma(i)}$ is sufficiently greater than $\ivarx_{\sigma(i-1)}$, then the left-hand-side of $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is \emph{dominated} by $a_i10^{\ivarx_{\sigma(i)}}$, moreover, if $a_i > 0$ and the value of $\ivarx_{\sigma(i)}$ is sufficiently small (resp.
-big), then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds (resp.
-does not hold), similarly for $a_i < 0$.
+Intuitively, 
+The lemma states that when the left-hand-side of the inequality is dominated by the $a_i 10^{\ivarx_{\sigma(i)}}$ term, if $\ivarx_{\sigma(i)}\le \alpha(\vec{\ivary})-1$ or $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2$, it can be determined directly whether the formula is true or false.
 
-%\vspace{-1mm}
 \begin{lemma} \label{lem:exp-ineq}
-Let  
-%
-$$
-\small
-\begin{array}{l}
-\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-%\hspace{4mm} 
-%\vspace{-1mm}
-a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}).
-\end{array}
-$$
-%
-with $a_i \neq 0$, $A\Def \sum_{j=1}^{i-1}\vert a_j\vert $, 
-%$B\Def  \sum_{j=1}^{i}\vert b_j\vert $, 
-$B \Def 2(\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert )+3)$,
-and $\delta\Def  \ell_{10}(A)+3$. 
+Let $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ be of form (\ref{tau}) 
+with $a_i \neq 0$. Let $A\Def 1+\sum_{j=1}^{i-1}\vert a_j\vert$, 
+$B \Def 1+\sum_{j=1}^{i}\vert b_j\vert$, 
+$\delta\Def  \ell_{10}(A)+2$,
+and $\gamma \Def 2\ell_{10}(B)+3$. 
+When $\ivarx_{\sigma(i)} \ge \gamma$ and  $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta$ are all satisfied,
 \begin{itemize}
-    \item If $a_i > 0$, let $\alpha(\vec{\ivary}) \Def \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$, then 
+    \item for $a_i > 0$, let $\alpha(\vec{\ivary}) \Def \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$, then 
     \begin{itemize}
-        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds,
-        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)}  \ge \ivarx_{\sigma(i-1)} +\delta$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold.
+        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds,
+        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold.
     \end{itemize}
-    \item If $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-a_i)$, then 
+    \item for $a_i < 0$, let $\alpha(\vec{\ivary})  \Def \ell_{10}(-\iterm(\ivary))- \ell_{10}(-a_i)$, then 
     \begin{itemize}
-        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold,
-        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta $, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
+        \item if $\ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ \textbf{does not} hold,
+        \item if $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})  +2$, then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
     \end{itemize}
@@
 
-We need the following proposition for the proof of Lemma~\ref{lem:exp-ineq}.
+We need the following two propositions for the proof of Lemma~\ref{lem:exp-ineq}. 
+Proposition~\ref{prop:linear} shows that when $\ivarx$ is large enough, the linear term $n \ivarx$ can always be bound by the exponential term $10^\ivarx$.
+Proposition~\ref{prop:case} can be seen as a special case of Lemma~\ref{lem:exp-ineq} where the left-hand-side of $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ has only one term $a_i 10^{\ivarx_{\sigma(i)}}$.
 
-\begin{proposition} \label{prop:1}
-If $n\ge m\ge 1$ and $p \ge 2(\ell_{10}(n)- \ell_{10}(m)+1)$, then 
-$n p \le m10^p$ holds.
+\begin{proposition} \label{prop:linear}
+Given $n\ge 1$, if $\ivarx \ge 2\ell_{10}(n)+1$, then 
+$n \ivarx  \le 10^\ivarx$ holds.
+\end{proposition}
+
+\begin{proof}
+When $1\le n\le 9$, we have $n\ivarx \le 10\ivarx \le 10^\ivarx$ for all $\ivarx \in\mathbb N$.
+
+For $n\ge 10$, that is, $\ell_{10}(n)\ge 1$, then 
+$$\begin{aligned}
+10^\ivarx&\ge 10^{\ell_{10}(n)+1}10^{\ivarx-\ell_{10}(n)-1}\\
+    &\ge 10^{\ell_{10}(n)+1}10(\ivarx-\ell_{10}(n)-1)\\
+    &\ge 10^{\ell_{10}(n)+1}\ivarx\\
+    &\ge n\ivarx
+\end{aligned}$$
+\end{proof}
+
+\begin{proposition} \label{prop:case}
+    Given $a>0$, 
+    if $\ivarx\le \ell_{10}(\ivary)-\ell_{10}(a)-1$, then $(a+\frac{1}{2})\cdot 10^\ivarx\le\ivary$; 
+    if $\ivarx\ge \ell_{10}(\ivary)-\ell_{10}(a)+2$, then $(a-\frac{1}{2})\cdot 10^\ivarx\ge\ivary$.
 \end{proposition}
@@
 
-First we show that for any $n' \in \Nat$, if $p \ge 2n'$, then $10^p \ge 10^{n'}10^{(p-n')}\ge 10^{n'}10(p-n')\ge 10^{n'}(5p+5(p-2n'))\ge  10^{n'}p$.
+Suppose that $\ivarx\le \ell_{10}(\ivary)-\ell_{10}(a)-1$,
+we have 
+$$
+(a+\frac{1}{2})\cdot 10^\ivarx \le (a+\frac{1}{2})\cdot \frac{\lambda_{10}(\ivary)}{10\lambda_{10}(a)}=\frac{a+\frac{1}{2}}{10\lambda_{10} (a)}\lambda_{10}(\ivary)\le\lambda_{10}(\ivary) \le \ivary. 
+$$
 
-If $ p \ge 2 (\ell_{10}(n) - \ell_{10}(m)+1)$, then  $n p \le 10\lambda_{10}(n) p = 10 * 10^{\ell_{10}(n)} p = 10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p$.
-
-Because $p \ge 2(\ell_{10}(n)-\ell_{10}(m) + 1)$, we deduce that $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  p \le 10^p$.
-Therefore, $10^{\ell_{10}(n)-\ell_{10}(m) + 1}  10^{\ell_{10}(m)} p \le 10^{\ell_{10}(m)} 10^p \le m 10^p$.
-We conclude that $np \le m 10^p$.
+Now, suppose that $\ivarx\ge \ell_{10}(\ivary)-\ell_{10}(a)+2$,
+$$
+(a-\frac{1}{2})\cdot 10^\ivarx \ge (a-\frac{1}{2})\frac{100\lambda_{10}(\ivary)}{\lambda_{10}(a)}=\frac{10(a-\frac{1}{2})}{\lambda_{10}(a)}10\lambda_{10}(\ivary)\ge 10\lambda_{10}(\ivary) \ge \ivary
+$$
 \end{proof}
 
-%Then we give the proof for Theorem~\ref{thm:exp-ineq}.
 \begin{proof}[Proof of Lemma~\ref{lem:exp-ineq}]
-We only prove for the case $a_i > 0$, the other case is symmetric.
+We only prove for the case $a_i > 0$, the $a_i < 0$ case is symmetric. The proof is twofold.
 
-Let $A\Def \sum_{j=1}^{i-1}\vert a_j\vert $, $B \Def 2(\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert )+3)$, and $\delta\Def  \ell_{10}(A)+3$.
+1) First, we prove that if $\ivarx_{\sigma(i)}\ge \gamma$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$, then the left-hand-side of $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is dominated by $a_i 10^{\ivarx_{\sigma(i)}}$, i.e.
+,
+\begin{equation} \label{eq:bound}
+(a_i-\frac{1}{2})10^{\ivarx_{\sigma(i)}}
+< a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)}
+< (a_i+\frac{1}{2})10^{\ivarx_{\sigma(i)}}
+\end{equation}
 
-Suppose $\ivarx_{\sigma(i)} \ge B$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$. Moreover, let $\alpha(\vec{\ivary}) \Def  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$.
+To obtain the formula above, we need to bound the absolute values $\vert\sum_{j=1}^{i-1}a_j10^{\ivarx_{\sigma(j)}}\vert$ and  $\vert\sum_{k=1}^i b_k\ivarx_{\sigma(k)}\vert$ by some fraction of $10^{\ivarx_{\sigma(i)}}$ respectively.
+Here we choose the bound to be $\frac{1}{10}10^{\ivarx_{\sigma(i)}}$.
+
+Using $\ivarx_{\sigma(i)}\ge \ivarx_{\sigma(i-1)}+\delta$ and the linear order $\ivarx_{\sigma(i)}\ge \ivarx_{\sigma(i-1)} \ge \cdots \ge \ivarx_{\sigma(1)}$,
+we can prove that 
+\begin{equation}\label{eq:appr-exp}
+    \begin{aligned}
+        \vert \sum_{j=1}^{i-1}a_j10^{\ivarx_{\sigma(j)}} \vert &\le  \sum_{j=1}^{i-1} \vert a_j\vert 10^{\ivarx_{\sigma(j)}}\\
+        &< A\cdot 10^{\ivarx_{\sigma(i)}-\delta}\\
+        &\le \frac{A}{100\lambda_{10}(A)} 10^{\ivarx_{\sigma(i)}}\\
+        &< \frac{1}{10} 10^{\ivarx_{\sigma(i)}} && \text{(by $\ivarx < 10 \lambda_{10}(\ivarx)$)}.
+    \end{aligned}
+\end{equation}
+
+Similarly, using $\ivarx_{\sigma(i)}\ge \gamma$, we have
+\begin{equation}\label{eq:appr-linear}
+    \begin{aligned}
+        \vert \sum_{k=1}^i b_k\ivarx_{\sigma(k)} \vert &\le \sum_{k=1}^i \vert b_k \vert \ivarx_{\sigma(k)} \\
+        &< B \cdot \ivarx_{\sigma(i)}\\
+        &=\frac{1}{10}\cdot 10B \cdot \ivarx_{\sigma(i)}\\
+        &\le \frac{1}{10} 10^{\ivarx_{\sigma(i)}} &&\text{(by Proposition~\ref{prop:linear}, set $n=10B$)}.     
+    \end{aligned}
+\end{equation}
+
+Combining formula (\ref{eq:appr-exp}) and (\ref{eq:appr-linear}), we have 
+$\vert \sum_{j=1}^{i-1}a_j10^{\ivarx_{\sigma(j)}}
++ \sum_{k=1}^i b_k\ivarx_{\sigma(k)}\vert 
+< \frac{3}{10} 10^{\ivarx_{\sigma(i)}}< \frac{1}{2} 10^{\ivarx_{\sigma(i)}}$, then formula (\ref{eq:bound}) can be easily deduced.
+
+2) Then, we give the sufficient conditions for $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ to hold or not by comparing $\iterm(\vec\ivary)$ with $(a_i \pm\frac{1}{2}) 10^{\ivarx_{\sigma(i)}}$.
+Note that when $\iterm(\vec\ivary)\le 0$, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ does not hold since $0< (a_i-\frac{1}{2})10^{\ivarx_{\sigma(i)}}$, so we only need to consider the $\iterm(\vec\ivary)> 0$ case.
+
+Utilizing Proposition~\ref{prop:case}, we obtain that
+
+\begin{equation}
+    \begin{aligned}
+        \ivarx_{\sigma(i)} \le \alpha(\vec\ivary)-1
+        &\implies (a_i+\frac{1}{2}) \cdot 10^{\ivarx_{\sigma(i)}}\le \iterm(\vec\ivary)\\
+       &\implies \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \textrm{ is satisfied }
+    \end{aligned}
+\end{equation}
+and
+\begin{equation}
+    \begin{aligned}
+        \ivarx_{\sigma(i)} \ge \alpha(\vec\ivary)+2
+        &\implies (a_i-\frac{1}{2})\cdot 10^{\ivarx_{\sigma(i)}}\ge\iterm(\vec\ivary)\\ 
+        &\implies \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \textrm{ is unsatisfied }
+    \end{aligned}
+\end{equation}
+
+Thus the lemma is proved.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
+Let $A\Def \sum_{j=1}^{i-1}\vert a_j\vert $, $\gamma \Def 2(\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert )+3)$, and $\delta\Def  \ell_{10}(A)+3$.
+
+Suppose $\ivarx_{\sigma(i)} \ge \gamma$ and $\ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} + \delta$. Moreover, let $\alpha(\vec{\ivary}) \Def  \ell_{10}(\iterm(\ivary))- \ell_{10}(a_i)$.
 
@@
  \begin{equation} 
-   A10^{-\delta} =A 10^{-\ell_{10}(A)-3} = \frac{A}{1000\lambda_{10}(A)} \le \frac{1}{100}.   \label{eq:thm-ineq-1}
+   A10^{-\delta} =A 10^{-\ell_{10}(A)-3} = \frac{A}{1000\lambda_{10}(A)} \le \frac{1}{100}.   
  \end{equation}
@@
 $n \ge m \ge 1$.
-From $2(\ell_{10}(n)- \ell_{10}(m)+1) = 2 (\ell_{10}(100 \sum_{j=1}^{i}\vert b_j\vert ) + 1) = 2 (\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert ) + 3) = B$, we deduce that $p = \ivarx_{\sigma(i)} \ge B = 2(\ell_{10}(n)- \ell_{10}(m)+1)$.
-Then according to Proposition~\ref{prop:1}, 
+From $2(\ell_{10}(n)- \ell_{10}(m)+1) = 2 (\ell_{10}(100 \sum_{j=1}^{i}\vert b_j\vert ) + 1) = 2 (\ell_{10}(\sum_{j=1}^{i}\vert b_j\vert ) + 3) = \gamma$, we deduce that $p = \ivarx_{\sigma(i)} \ge \gamma = 2(\ell_{10}(n)- \ell_{10}(m)+1)$.
+Then according to Proposition~\ref{prop:linear}, 
 $$100(\sum_{j=1}^{i}\vert b_j\vert )\ivarx_{\sigma(i)} = np \le m 10^p = 10^{\ivarx_{\sigma(i)}}.$$
@@
 $
+Therefore,  in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Therefore,  in this case, $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ holds.
+For $a_i>0$, by lemma~\ref{lem:exp-ineq},
+$\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ (for readability, denoted by $\tau$ below) is equivalent to following formula without exponential occurrences of $\ivarx_{\sigma(i)}$:
+%&&(\ivarx_{\sigma(i)}<\gamma \text{ or } \ivarx_{\sigma(i)}\le \ivarx_{\sigma(i-1)}+\delta)
+\begin{equation}\label{lemma-example}\begin{aligned}
+    &\bigvee_{p=0}^{\gamma-1} (\ivarx_{\sigma(i)}=p \wedge \tau[p/\ivarx_{\sigma(i)}])\\
+    \vee &\bigvee_{p=0}^{\delta-1} (\ivarx_{\sigma(i)}=\ivarx_{\sigma(i-1)}+p\wedge \tau[\ivarx_{\sigma(i-1)}+p/\ivarx_{\sigma(i)}])\\
+    \vee & \bigg( \ivarx_{\sigma(i)}\ge \gamma \wedge \ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)}+\delta \wedge \ivarx_{\sigma(i)}\le \alpha(\vec\ivary)-1\bigg)\\
+    \vee & \bigvee_{p=0,1} \bigg( \ivarx_{\sigma(i)}\ge \gamma \wedge \ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)}+\delta \wedge \ivarx_{\sigma(i)} = \alpha(\vec\ivary)+p \wedge \tau[\alpha(\vec\ivary)+p/\ivarx_{\sigma(i)}]\bigg).
+\end{aligned}\end{equation}
+The elimination of the exponential occurrences of $\ivarx_{\sigma(i)}$ for the case $a_i < 0$ is similar.
 
-\end{proof}
+We would like to make a few remarks about Lemma~\ref{lem:exp-ineq}:
+(1) When $i=1$, we can just ignore the condition $\ivarx_{\sigma(i)}\ge \ivarx_{\sigma(i-1)}+\delta$.
+(2) The constant $1$ in the definition of $A,B$ is to ensure that $\ell_{10}(A),\ell_{10}(B)$ are well-defined. 
+(3) The lemma still holds when the base of exponential function is changed to any natural number $n\ge 2$ (constants in $\delta,\gamma$ should be changed accordingly).
 
-We would like to remark that Lemma~\ref{lem:exp-ineq} still holds when the base of exponential function is changed to any natural number $n\ge 2$.
-
-If $a_i > 0$, then the exponential occurrences of $\ivarx_{\sigma(i)}$ in $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ can be eliminated by utilizing  Lemma~\ref{lem:exp-ineq} and enumerating the constraints on $\ivarx_{\sigma(i)}$ and $\ivarx_{\sigma(i-1)}$.
-Specifically,
-$\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
-%\vspace{-2mm}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hide{
 \[
@@
 \begin{array}{l}
-\bigvee \limits_{p=0}^{B-1} a_i 10^{p}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_i p + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}) \\
-\bigvee \big(\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1  \wedge \ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta \big)\\
+\bigvee \limits_{p=0}^{\gamma-1} a_i 10^{p}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + b_i p + \sum_{k=1}^{i-1}b_k \ivarx_{\sigma(k)} \le \iterm(\vec{\ivary}) \\
+\bigvee \big(\ivarx_{\sigma(i)} \ge \gamma \wedge \ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1  \wedge \ivarx_{\sigma(i)} \ge \ivarx_{\sigma(i-1)} +\delta \big)\\
 %
@@
 \begin{array}{l}
-\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1 \ \wedge \\
+\ivarx_{\sigma(i)} \ge \gamma \wedge \ivarx_{\sigma(i)} \le \alpha(\vec{\ivary})  -1 \ \wedge \\
  \ivarx_{\sigma(i)} = \ivarx_{\sigma(i-1)} +p\ \wedge\\
@@
 \begin{array}{l}
-\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} = \alpha(\vec{\ivary})\ \wedge \\
+\ivarx_{\sigma(i)} \ge \gamma \wedge \ivarx_{\sigma(i)} = \alpha(\vec{\ivary})\ \wedge \\
 \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\alpha(\vec{\ivary})/\ivarx_{\sigma(i)}]
@@
 \begin{array}{l}
-\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} = \alpha(\vec{\ivary})+1\ \wedge \\
+\ivarx_{\sigma(i)} \ge \gamma \wedge \ivarx_{\sigma(i)} = \alpha(\vec{\ivary})+1\ \wedge \\
 \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[\alpha(\vec{\ivary})+1/\ivarx_{\sigma(i)}]
@@
 \begin{array}{l}
-\ivarx_{\sigma(i)} \ge B \wedge \ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2\ \wedge\\
+\ivarx_{\sigma(i)} \ge \gamma \wedge \ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2\ \wedge\\
  \ivarx_{\sigma(i)} = \ivarx_{\sigma(i-1)} +p\ \wedge \\
@@
 \]
-where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.  
-%
-The elimination of the exponential occurrences of $\ivarx_{\sigma(i)}$ for the case $a_i < 0$ is similar.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsubsection{Divisibility atoms}
+
+For divisibility atoms, we utilize its periodic property to enumerate $\ivarx_{\sigma(i)}$ within a finite range.
 Consider a divisiblilty atomic formula 
-%
-$$
-\begin{array}{l}
-\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
-d\ \big \vert \ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+\begin{equation}
+    \begin{aligned}
+        &\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary}) \Def  \\
+        &d\ \big \vert \ \big(a_i 10^{\ivarx_{\sigma(i)}} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
         + \iterm(\vec{\ivary}) \big)               
-\end{array}
-$$
-with $a_i \neq 0$. The indivisibility case can be treated analogously.
-%
+    \end{aligned}
+\end{equation}
+with $a_i \neq 0, d\ge 1$. The indivisibility case can be treated analogously.
 
@@
 
-If $d_0 = 1$, then $10^r$ is divisible by $d = 2^{r_1}5^{r_2}$.
+Since $10$ and $d_0$ are relatively prime, according to Euler's theorem (cf.
+\cite{HW80}), $10^{\phi(d_0)} \equiv 1 \bmod d_0$, where $\phi$ is the Euler function.
+Suppose $10^{\phi(d_0)} = kd_0 + 1$ for some $k \in \Nat$.
+Then for every $n \in \Nat$ with $n \ge r$, 
+$$\begin{aligned}
+    10^{n + \phi(d_0)} 
+    &\equiv 10^{n-r} 10^r (kd_0 + 1) \bmod d\\
+    &\equiv 10^{n-r} (k 10^rd_0 + 10^r) \bmod d\\
+    &\equiv 10^{n-r} (0+10^r) \bmod d\\
+    &\equiv 10^n \bmod d.
+\end{aligned}$$
+
+\hide{
+$$
+\begin{array}{l}
+10^{n + \phi(d_0)} \bmod d =10^{n-r} 10^r (kd_0 + 1) \bmod d = \\
+10^{n-r} (k 10^rd_0 + 10^r) \bmod d = \\
+10^{n-r} (0+10^r) \bmod d = 10^n \bmod d.
+\end{array}
+$$
+}
+
+Then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to the following formula
+
+\begin{equation}\label{lem-div}
+\begin{array}{l}
+\bigvee \limits_{p=0}^{r-1} \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})[p/\ivarx_{\sigma(i)}]\\
+\vee
+\left(
+\begin{array}{l}
+\ivarx_{\sigma(i)} \ge r\ \wedge \\
+\bigvee \limits_{p = 0}^{\phi(d_0)-1} 
+\left(
+\begin{array}{l}
+\phi(d_0)\ \big \vert \ (\ivarx_{\sigma(i)} - r - p)\ \wedge \\
+d\ \big \vert  
+\left(
+\begin{array}{l}
+a_i 10^{r+p} + \sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \\
+\sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} + \iterm(\vec{\ivary})
+\end{array}
+\right) 
+\end{array}
+\right)
+\end{array}
+\right),
+\end{array}
+\end{equation}
+where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear. When $\ivarx_{\sigma(i)}>r$, we only need to enumerate $\ivarx_{\sigma(i)}$ in one period, i.e. from $r$ to $r+\phi(d_0)-1$.
+
+For a special case $d_0 = 1$, since $\phi(1)=1$, (\ref{lem-div}) can be simplified to 
+$$\begin{aligned}
+    \bigvee_{p=0}^{r-1} \tau [p/\ivarx_{\sigma(i)}] \vee  \bigg(\ivarx_{\sigma(i)}\ge r \wedge d\ \big \vert \big(\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i} b_k \ivarx_{\sigma(k)} 
+    + \iterm(\vec{\ivary}) \big)\bigg)
+\end{aligned}.$$
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\hide{
+
+then $10^r$ is divisible by $d = 2^{r_1}5^{r_2}$.
 Thus for every $n \ge r$, $d \ \vert \ 10^n$.
@@
 Next, let us assume $d_0 > 1$. 
-Since $10$ and $d_0$ are relatively prime, according to Euler's theorem (cf.
-\cite{HW80}), $10^{\phi(d_0)} \equiv 1 \bmod d_0$, where $\phi$ is the Euler function.
-Suppose $10^{\phi(d_0)} = kd_0 + 1$ for some $k \in \Nat$.
-Then for every $n \in \Nat$ with $n \ge r$, 
-$$
-\begin{array}{l}
-10^{n + \phi(d_0)} \bmod d =10^{n-r} 10^r (kd_0 + 1) \bmod d = \\
-10^{n-r} (k 10^rd_0 + 10^r) \bmod d = \\
-10^{n-r} (0+10^r) \bmod d = 10^n \bmod d.
-\end{array}
-$$
+}
 
-Then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is equivalent to 
+\hide{
 \[
@@
 where the exponential occurrences of $\ivarx_{\sigma(i)}$ disappear.
+}
+
+\hide{
+\begin{equation}\label{lem-div}
+    \begin{aligned}
+        &\bigvee_{p=0}^{r-1} \tau[p/\ivarx_{\sigma(i)}]\\ 
+        \vee &\left( x_{\sigma(i)}\ge r \wedge 
+            \bigvee_{p=0}^{\phi(d_0)-1} \left(\tau[r+p/\ivarx_{\sigma(i)}]\wedge \phi(d_0)\ \big \vert\ (\ivarx_{\sigma(i)}-r-p) \right)
+            \right).
+    \end{aligned}
+\end{equation}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-03-18 Acta Informatica2022/5_complexity.tex
.

--- 2021-09-29 FMCAD2021/5_complexity.tex
@@
 
-\subsection{Complexity analysis}\label{app:cpx}
+In this section, we analyse the complexity of Point's procedure applied on an existential quantified {\paexp} formula.
 
-In this part, we analyse the complexity of the quantifier elimination procedure applied on a formula with one existential quantifer, that is, a formula of the form $\exists \ivarx.
-\varphi(\ivarx, \vec{\ivary})$.
-We will prove that the complexity of eliminating one quantifer has a 3-EXPSPACE upper bound.
-Since the quantifier elimination procedure for $\pa$ works as a subprocedure (in eliminating linear occurrences of variables) and also has a 3-EXPSPACE complexity, the bound may not be easily improved.
+Consider a formula of the form $$\exists \ivarx^1\cdots\exists \ivarx^m.\varphi(\ivarx^1,\cdots,\ivarx^m,\vec{\ivary})$$ with free variables $\vec{\ivary}$.
+Let $N$ denote the length of the formula.
+We will prove that the space complexity to eliminate all (existential) quantifiers has a 3-EXPSPACE upper bound.
+Since Cooper's quantifier elimination algorithm for $\pa$ works as a subprocedure and also has a 3-EXPSPACE upper bound, this bound may not be easily improved.
 
-We first analyse the changes of formula length in the normalization step.
-For the rest of the decision procedure, we adopt the style in the complexity analysis of \pa \ (cf.
+We first give a brief analysis of the normalization step, in which fresh variables are introduced and the length of the formula grows linearly at most. 
+For the rest of the algorithm, we adopt the strategy of Oppen's analysis of Cooper's algorithm (cf.
 \cite{Oppen73}).
 The idea is that the upper bound of the formula length can be expressed using the product of the number of atoms, the number of coefficients and the length of the maximum constant.
-So we track the changes of such quantities throughout the procedure.
+The critical point of our analysis is that only coefficients of linear occurrences of quantified variables need to be considered.
 
+\paragraph{Normalization} In the normalization step, we normalize the formula w.r.t quantified variables from $\ivarx^m$ to $\ivarx^1$.
+Suppose that fresh variables $\ivarx^i_1,\cdots,\ivarx^i_{n_i}$ are introduced for each $\ivarx^i(1\le i\le m)$, the formula then becomes $$\exists \ivarx^1\exists \ivarx^1_1\cdots\exists\ivarx^1_{n_1}\cdots\exists \ivarx^m\exists \ivarx^m_1\cdots\exists\ivarx^m_{n_m}.\varphi'(\ivarx^1,\cdots,\ivarx^m_{n_m},\vec{\ivary}).$$
+
+Since the number of newly introduced variables is less than the number of occurrences of exponential and logarithmic functions in the original formula, we have at most $m+\sum_{i=1}^m n_i\le N$ quantified variables after normalization.
+
+The increase in the length of the formula during normalization comes from two sources, the conjuncts for each introduced variables and the additional formulas for translating $\neq,=$ relations to $\le$.
+It is not hard to see that both these operations will at most increase the length of the formula by a constant factor.
+
+In the sequel, to avoid redundant symbols, we still use $m$ to denote the number of quantified variables and $N$ to denote the length of the formula after normalization.
+Just keep in mind that $m\le N$.
+
+\hide{
     Given a formula $\exists \ivarx.
@@
 In addition, we can also conclude that the number of fresh variables, $m$, is less than $n$ since there are at most $n$ different forms of terms.
+}
 
-After the normalization step, denote the obtained formula by $\exists \vec{\ivarx}.
-\varphi(\vec{\ivarx},\vec{\ivary})$ with quantified variables $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$ and free variables $\vec{\ivary}$.
-$\varphi$ contains no quantifiers.
-Let $n'$ denote the length of the formula.
+\paragraph{Enumeration of linear orders (the outer for-loop)}
 
-According to our quantifier elimination procedure,we eliminate the quantified variables one by one: each time select the largest variable among $\ivarx_i,1\le i\le m$, first eliminate the exponential occurrences, then linear occurrences.
-The linear order of quantified variables are given by a permutation.
-For $m$ quantified variables, there are $m!$ possible linear orders.
+Denote the normalized formula by $\exists \vec{\ivarx}.
+\varphi(\vec{\ivarx},\vec{\ivary})$ with $\vec{\ivarx} = (\ivarx_1,\dots,\ivarx_m)$.
+According to Point's procedure, we then eliminate the quantified variables one by one: each time select the largest variable among $\ivarx_i,1\le i\le m$, first eliminate the exponential occurrences, then linear occurrences.
+Each linear order of quantified variables is given by a permutation.
+For $m$ quantified variables, there are $m!$ possible linear orders, which means the inner for-loop repeats $m!$ times.
 
+\paragraph{Elimination of quantifiers (the inner for-loop)}
 Suppose the specified linear order is $\ivarx_m\ge \dots \ge \ivarx_1$.
-Let $\exists \ivarx_1\dots \exists\ivarx_{m-k}.\varphi_k$ denote the formula obtained from $\exists \vec{\ivarx}.
-\varphi(\vec{\ivarx},\vec{\ivary})$ by eliminating quantifiers $\exists \ivarx_m,\dots,\exists\ivarx_{m-k+1}$.
-Let $\exists \ivarx_1 \dots \exists \ivarx_{m-k+1}.\varphi'_k$ denote the formula obtained from $\exists \ivarx_1\dots \exists\ivarx_{m-k+1}.\varphi_{k-1}$ by eliminating linear occurrences of $\ivarx_{m-k+1}$.
+Let $\exists \ivarx_1\dots \exists\ivarx_{m-k}.\varphi_k$ denote the formula obtained by eliminating quantifiers $\exists \ivarx_m,\dots,\exists\ivarx_{m-k+1}$ from $\exists \vec{\ivarx}.
+\varphi(\vec{\ivarx},\vec{\ivary})$.
+Let $\varphi'_k$ denote the formula obtained by further eliminating exponential occurrences of $\ivarx_{m-k+1}$ (from $\varphi_k$).
 Let $\varphi_0$ denote $\varphi$.
 
-Let $c_k$ be the number of distinct $d$ in atoms of the form $d \ \big | \ \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of quantified variables in $\varphi_k$.
+Let $c_k$ be the number of distinct divisor $d$ in atoms of the form $d \mid \iterm$ and $d \nmid \iterm$ plus the number of distinct coefficients of \emph{linear occurrences} of quantified variables in $\varphi_k$.
 Let $s_k$ be the largest constant (including coefficients) and $a_k$ be the number of atomic formulas in $\varphi_k$.
 Similarly we define $c'_k$, $s'_k$ and $a'_k$ for $\varphi'_k$.
-Let $c,s,a$ be $c_0,s_0,a_0$ respectively.
 
-First we analyse the sub-procedure to eliminate exponential occurrences of $\ivarx_m$. We prove the following lemma.
+First we analyse the sub-procedure to eliminate exponential occurrences of the inner most quantified variable $\ivarx_m$. We prove the following lemma.
 
@@
     \begin{align}
-        c'_1&\le c^2 \notag \\
-        s'_1&\le ms^2 \notag\\
-        a'_1&\le sa\notag 
+        c'_0&\le c_0^2 \notag \\
+        s'_0&\le ms_0^2 \notag\\
+        a'_0&\le s_0a_0\notag 
     \end{align} 
@@
 \begin{proof}
+The analysis is divided into two cases by assuming all atomic formulas are of the same form (inequalities atoms or divisibility atoms).
 
-The analysis is separated into 2 cases where all atoms are of the same form (inequalities atoms or divisiblilty atoms).
+If all atoms are inequalities of the form in Lemma~\ref{lem:exp-ineq}.
+We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_m$ is replaced by a new formula.
+Note that coefficients of linear occurrences of $\ivarx_i$, for $i\le m-2$, remain changed throughout the substitutions, only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant 1 is introduced as a coefficient for $\ivarx_m$, and if we substitute $\ivarx_m$ by $\ivarx_{m-1}+p$ for some constant $p$, coefficient of linear occurrence of $\ivarx_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $\ivarx_{m}$ and $\ivarx_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
+Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le c_0^2$.
+Note that $\delta$ and $B$ in Lemma~\ref{lem:exp-ineq} are at most $\ell_{10}(ms_0)$.
+When we substitute $\ivarx_m$ by $\ivarx_{m-1}+\delta-1$ or by $B-1$, the largest constant in the formula becomes at most $s_0\cdot  10^{\ell_{10}(ms_0)}\le ms_0^2$.
+And an inequality is replaced by at most $4\ell_{10}(ms_0)$ atomic formulas, so $a'_1\le 4\ell_{10}(ms_0)a_0$.
 
-If all atoms are inequalities atomic formulas of the form in Lemma~\ref{lem:exp-ineq}.
-We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_m$ is replaced by a new formula.
-Only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant coefficient $1$ is introduced, and if we substitute $\ivarx_m$ by $\ivarx_{m-1}+p$ for some $p$, coefficient of linear occurrence of $\ivarx_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $\ivarx_{m}$ and $\ivarx_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
-Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le c^2$.
-Note that $\delta$ and $B$ are at most $\ell_{10}(ms)$, when we substitute $\ivarx_m$ by $\ivarx_{m-1}+\delta-1$ or by $B-1$, the largest constant in the formula is at most $s\cdot  10^{\ell_{10}(ms)}\le ms^2$.
-And an inequality is replaced by at most $4\ell_{10}(ms)$ atomic formulas, so $a'_1\le 4\ell_{10}(ms)a$.
+If all atoms are divisibility atomic formulas of the form $d \ \big \vert  \ \iterm$ or $d \ \nmid \ \iterm$.
+We have $c'_1<2c_0$ because a divisibility atomic formula  will produce at most two forms of atomic formulas $d \ \big \vert  \ \iterm$ and $\phi(d) \ \big \vert  \ \iterm$.
+In a divisibility atom, any constant in the divident $\iterm$ , say $l$, can be replaced by ($l \bmod d)$.
+So we have $s'_1\le s_0$.
+When $d$ is a large prime number, $\phi(d)=d-1$, a divisibility atomic formula is replaced by roughly $d$ atomic formulas, so $a'_1\le s_0a_0$.
 
-If all atoms are divisiblilty atomic formulas of the form $d \ \big | \ \iterm$ or $d \ \nmid \ \iterm$.
-We have $c'_1<2c$ since a divisibility atomic formula  will produce at most two forms of atomic formulas $d \ \big | \ \iterm$ and $\phi(d) \ \big | \ \iterm$.
-Note that any constant in $\iterm$ , say $l$, can be replaced by ($l \bmod d)$, so $s'_1\le s$.
-When $d$ is a large prime number, $\phi(d)=d-1$, a divisibility atomic formula is replaced by roughly $d$ atomic formulas, so $a'_1\le sa$.
-
-Choose larger upper bound for $c'_1$, $s'_1$ and $a'_1$ respectively, then the lemma is proved.
+Choose larger upper bounds for $c'_1$, $s'_1$ and $a'_1$ respectively, then the lemma is proved.
 
@@
 
-Since linear occurrences of $\ivarx_m$ are eliminated using Cooper's algorithm, we combine Oppen's analysis for this sub-procedure:
+When $\varphi_0'$ is transformed into $\varphi_{1}$ by eliminating linear occurrences of $\ivarx_m$, Oppen's analysis gives the following lemma.
 
-\begin{lemma}\label{lem:cpx oppen}
+\begin{lemma}\cite{Oppen73}\label{lem:cpx oppen}
     \begin{align}
-        c_1&\le c'^4 \notag \\
-        s_1&\le s'^{4c'} \notag\\
-        a_1&\le {a'}^4 s'^{2c'}\notag 
+        c_1&\le c_0'^4 \notag \\
+        s_1&\le s_0'^{4c_0'} \notag\\
+        a_1&\le {a_0'}^4 s_0'^{2c_0'}\notag 
     \end{align} 
@@
 
-From Lemma~\ref{lem:cpx exp} and Lemma~\ref{lem:cpx oppen}, we have
+Combining Lemma~\ref{lem:cpx exp} and Lemma~\ref{lem:cpx oppen}, we have
 \begin{align}
-    c_1&\le c^8 \notag \\
-    s_1&\le (ms^2)^{4c^2} \notag\\
-    a_1&\le (sa)^4(ms^2)^{2c^2}\notag 
+    c_1&\le c_0^8 \notag \\
+    s_1&\le (ms_0^2)^{4c_0^2} \notag\\
+    a_1&\le (s_0a_0)^4(m_0s_0^2)^{2c_0^2}\notag 
 \end{align}
 
-Assuming $m\le n$, by induction on $k$ we get
+By induction on $k$ and assuming $m\le N$, we get
 
@@
     \begin{align}
-        c_k&\le c^{8^k} \notag \\
-        s_k&\le n^{(4c)^{ 8^k}} s^{(8c)^{ 8^k}} \notag\\
-        a_k&\le a^{4^k}n^{(4c)^{ 8^k}} s^{(8c)^{ 8^k}} \notag
+        c_k&\le c_0^{8^k} \notag \\
+        s_k&\le N^{(4c_0)^{ 8^k}} s_0^{(8c_0)^{ 8^k}} \notag\\
+        a_k&\le a_0^{4^k}n^{(4c_0)^{ 8^k}} s_0^{(8c_0)^{ 8^k}} \notag
     \end{align} 
@@
 
-If we assume $c\le n$, $a\le n$ and $s\le n$, the space required to store the quantifier free formula, $\varphi_k$, is bounded by the product of the number of linear orders $m!$, the number of atoms $a_k$, the maximum number of constants $2m+2$ per atom, the maximum amount of space $s_k$ to store each constant and some constant $q$:
+If we adopt the assumptions in \cite{Oppen73}, by assuming $c_0\le N$, $a_0\le N$ and $s_0\le N$, the space required to store the quantifier free formula $\varphi_k$, is bounded by the product of the number of linear orders $m!$, the number of atoms $a_k$, the maximum number of constants $2m+2$ per atom, the maximum amount of space $s_k$ to store each constant and some constant $q$.
+So the space complexity is $O(m!
+\cdot a_k \cdot (2m+2) \cdot s_k)=O(2^{2^{2^{pn \log n}}})$, which belongs to 3-EXPSPACE.
 
-$${\sf space}\le q \cdot m! \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
-for some large constant $p$.
+\hide{$${\sf space}\le q \cdot m! \cdot a_k \cdot (2m+2) \cdot s_k \le 2^{2^{2^{pn \log n}}}$$
+for some large constant $p$. }
 
-REMARK: When the original formula has no alternating quantifiers, for example, the formula of the form $\exists \ivarx_m \exists \ivarx_{m-1} \dots\exists \ivarx_1.\varphi(\ivarx_1,\dots,\ivarx_{m-1},\ivarx_m,\vec{\ivary})$.
-Since we can carry out the normalization step for all quantified variable $\ivarx_i(1\le i\le m)$ at the same time and enumerate the linear order for all introduced variables, the complexity of eliminating all quantifers is still 3-EXPSPACE.
-However, if there are  alternating quantifers at the very beginning, say the formula is of the form $\forall \ivarx_2\exists \ivarx_1.\varphi(\ivarx_1,\ivarx_2,\vec{\ivary})$.
-We have to eliminate the occurrences of $\ivarx_1$ first, treating $\ivarx_2$ as a free variable like $\vec{\ivary}$.
-When eliminating exponential occurrences of variables in inequalities (see Lemma~\ref{lem:exp-ineq}), $\ivarx_2$ will be collected in term $\iterm(\ivary)$.
-Since $\iterm(\ivary)$ may change due to the  substitutions of $\ivarx_1$, when we carry out normalization step for $\ivarx_2$, the number of introduced fresh variables is not bounded by the length of original formula.
-Hence, we conjecture that the overall complexity for eliminating all quantifiers in a formula will be non-elementary.
+The above analysis is focused on the existential {\paexp} formulas and is sufficient for our string constraints solving context.
+But what about general {\paexp} formulas that contain alternating quantifiers?
+Unfortunately, directly extending our analysis on alternating quantifers by transforming $\forall \ivarx.\varphi$ into $\neg \exists \ivarx.\neg \varphi$ will not give an elementary upper bound.
+We will explain the main idea without going into details on this issue.
+To eliminate quantifiers in a formula, say $\forall \ivarx_2\exists \ivarx_1.\varphi(\ivarx_1,\ivarx_2,\vec \ivary)$, we first eliminate $\exists \ivarx_1$, treating $\ivarx_2$ and $\vec \ivary$ both as free variables.
+The subprocedure to eliminate exponential occurrences of $\ivarx_1$ is at the cost of introducing logarithmic expressions of $\ivarx_2$ and $\vec \ivary$, which in the worst case increases \emph{exponentially} the number of fresh variables in the normalization step w.r.t $\ivarx_2$.
+We conjecture that the space complexity of quantifier elimination procedure over {\paexp} formulas still has an elementary upper bound, which may require a more detailed analysis or improvements of the algorithm.

### 2022-03-18 Acta Informatica2022/6_optimization.tex
.

@@
 
-In the last section, we illustrated the main idea of the decision procedure for  {\paexp}.
-The decision procedure has a high complexity in the sense that the elimination of the exponential occurrences   of each variable incurs an exponential blow-up, similarly for linear occurrences.
+In the last section we showed the complexity of Point's procedure over existential {\paexp} formulas to be 3-EXPSPACE, which  is quite expensive and a faithful implementation would not scale\footnote{We did implement Point's algorithm and discovered that the implementation could only solve formulas of very small size.}.
+Note that this high complexity holds even for quantifier-free $\paexp$ formulas: For a quantifier-free formula $\varphi$, we solve its satisfiability problem by adding the existential quantifiers for all the variables occurring in $\varphi$, then eliminate the quantifiers one by one, resulting into $\ltrue$ or $\lfalse$ in the end.
+The original formula $\varphi$ is satisfiable if $\ltrue$ is obtained in the end.
+
+\hide{
+In Section~\ref{sec-dec}, we illustrated the main idea of the decision procedure for  {\paexp}.
+In last section, we show that the decision procedure has a high complexity in the sense that the elimination of the exponential occurrences of each variable incurs an exponential blow-up, similarly for linear occurrences.
 Note that this high complexity holds even for quantifier-free formulas containing exponential terms: For a quantifier-free formula $\varphi$ containing exponential terms, we solve its satisfiability problem by adding the existential quantifiers for all the variables occurring in $\varphi$, then eliminate the quantifiers one by one, resulting into $\ltrue$ or $\lfalse$ in the end.
 The original formula $\varphi$ is satisfiable if $\ltrue$ is obtained in the end.
+}
 
@@
 
-%In the sequel, we present two major optimizations and omit some additional ones, which can be found in Appendix~\ref{app-opt}. 
 In the sequel, we will explain two major optimizations in \ref{para: opt reduce} and \ref{para: opt under appro}. Additional optimizations are listed in \ref{para: opts}.
@@
 
-Recall that in the decision procedure for {\paexp}, after the normalization, the variable orders are enumerated and for each order, the exponential and linear occurrences of variables are eliminated.
+Recall that in Point's procedure for {\paexp}, after the normalization, the variable orders are enumerated and for each order, the exponential and linear occurrences of variables are eliminated.
 Since the quantifier elimination is expensive and applied to each possible order of variables, if we could reduce the candidate variable orders in the very beginning, it would facilitate considerable speed-up for the decision procedure.
@@
 
-Recall that Lemma~\ref{lem:exp-ineq} enables us to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ from 
-%\vspace{-1mm}
-$$
+Recall that Lemma~\ref{lem:exp-ineq} enables us to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ from an atomic formula $\tau$ of the form
+\hide{$$
 \begin{array}{l}
 \tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
-\hspace{4mm} 
-%\vspace{-1mm}
 a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c.
 \end{array}
-$$
-Actually, Lemma~\ref{lem:exp-ineq} does more in the sense that all occurrences of $\ivarx_{\sigma(i)}$, including the linear ones, are eliminated from the atomic formulas resulted from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, e.g.
-$\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-1)} + p /\ivarx_{\sigma(i)}]$.
-Then we can continue eliminating the exponential occurrences of $\ivarx_{\sigma(i-1)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-1)} + p /\ivarx_{\sigma(i)}]$, provided that the coefficient of $\ivarx_{\sigma(i-1)}$ therein is nonzero.
+$$}
+\begin{equation}
+    \begin{aligned}
+        &\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
+        &a_i 10^{\ivarx_{\sigma(i)}}+\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c.
+    \end{aligned}            
+\end{equation}
+
+Actually, Lemma~\ref{lem:exp-ineq} does more in the sense that all occurrences of $\ivarx_{\sigma(i)}$, including the linear ones, are eliminated from the atomic formulas resulted from $\tau$, e.g.
+$\tau[\ivarx_{\sigma(i-1)} + p /\ivarx_{\sigma(i)}]$.
+Then we can continue eliminating the exponential occurrences of $\ivarx_{\sigma(i-1)}$ from $\tau[\ivarx_{\sigma(i-1)} + p /\ivarx_{\sigma(i)}]$, provided that the coefficient of $\ivarx_{\sigma(i-1)}$ therein is nonzero.
 Iterating this process would produce a {\pa} formula eventually.
 
-Nevertheless, the side condition of Lemma~\ref{lem:exp-ineq}, namely $a_i \neq 0$, undermines the aforementioned natural idea. If $a_i = 0$, but $b_i \neq 0$, 
+Nevertheless, the condition of Lemma~\ref{lem:exp-ineq}, namely $a_i \neq 0$, undermines the aforementioned natural idea. If $a_i = 0$, but $b_i \neq 0$, 
 %namely, 
@@
 %$$
-then we are unable to utilize Lemma~\ref{lem:exp-ineq} to eliminate the linear occurrences of $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$.
-In this case, the quantifier elimination algorithm of {\pa} has to be applied to eliminate $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, so that later on, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-1)}$, which requires that $\ivarx_{\sigma(i-1)}$ is the maximum variable in the left-hand side of the inequality.
+then we are unable to utilize Lemma~\ref{lem:exp-ineq} to eliminate the linear occurrences of $\ivarx_{\sigma(i)}$ from $\tau$.
+In this case, the quantifier elimination algorithm of {\pa} has to be applied to eliminate $\ivarx_{\sigma(i)}$, so that later on, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-1)}$, which requires that $\ivarx_{\sigma(i-1)}$ is the maximum variable in the left-hand side of the inequality.
 
-To avoid applying the quantifier elimination algorithm of {\pa}, we consider the following under approximation of $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, namely, we additionally assume that $\ivarx_{\sigma(i)} \le 10^{u}$ for some constant bound $u \in \Nat$ with $u \ge 1$.
-Then $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$ can be rewritten as
+To avoid applying the quantifier elimination algorithm of {\pa}, we consider the following under approximation of $\tau$, namely, we additionally assume that $\ivarx_{\sigma(i)} \le 10^{u}$ for some constant bound $u \in \Nat$.
+Then $\tau$ can be rewritten as
 %\begin{equation}\label{eqn-org}
@@
 \end{equation}
-Let $\alpha_1 = \ell_{10}(c_1) -  \ell_{10}(a_{i-1})$ and $\alpha_2 = \ell_{10}(c_2) -  \ell_{10}(a_{i-1})$. Then from Lemma~\ref{lem:exp-ineq}, 
+
+Let $\alpha_1 = \ell_{10}(c_1) -  \ell_{10}(a_{i-1})$ and $\alpha_2 = \ell_{10}(c_2) -  \ell_{10}(a_{i-1})$. Then from Lemma~\ref{lem:exp-ineq}, let $A\Def 1+\sum_{j=1}^{i-2}\vert a_j\vert$, 
+$B \Def 1+\sum_{j=1}^{i-1}\vert b_j\vert$, 
+$\delta\Def  \ell_{10}(A)+2$,
+and $\gamma \Def 2\ell_{10}(B)+3$. 
 \begin{itemize}
-\item if $\ivarx_{\sigma(i-1)} \ge B$, $\ivarx_{\sigma(i-1)} \le \alpha_1 - 1$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-lb}), thus also $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, is evaluated to $\ltrue$,
-\item if $\ivarx_{\sigma(i-1)} \ge B$, $\ivarx_{\sigma(i-1)} \ge \alpha_2 + 2$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-ub}), thus also $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, is evaluated to $\lfalse$.
+\item if $\ivarx_{\sigma(i-1)} \ge \gamma$, $\ivarx_{\sigma(i-1)} \le \alpha_1 - 1$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-lb}), thus also $\tau'$, is evaluated to $\ltrue$,
+\item if $\ivarx_{\sigma(i-1)} \ge \gamma$, $\ivarx_{\sigma(i-1)} \ge \alpha_2 + 2$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-ub}), thus also $\tau'$, is evaluated to $\lfalse$.
 \end{itemize}
-Therefore, $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, thus also $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, is equivalent to 
+
+Therefore, $\tau'$ and $\tau$ are equivalent to 
+\begin{equation}\label{lemma-example}\begin{aligned}
+    &\bigvee_{p=0}^{\gamma-1} (\ivarx_{\sigma(i-1)}=p \wedge \tau[p/\ivarx_{\sigma(i-1)}])\\
+    \vee &\bigvee_{p=0}^{\delta-1} \bigg(\ivarx_{\sigma(i-1)}=\ivarx_{\sigma(i-2)}+p\wedge \tau[\ivarx_{\sigma(i-2)}+p/\ivarx_{\sigma(i-1)}]\wedge \ivarx_{\sigma(i-1)}\le \alpha_1-1\bigg)\\
+    \vee &\bigvee_{p=0}^{\delta-1} \bigg(\ivarx_{\sigma(i-1)}=\ivarx_{\sigma(i-2)}+p\wedge \tau[\ivarx_{\sigma(i-2)}+p/\ivarx_{\sigma(i-1)}]\wedge \ivarx_{\sigma(i-1)}\ge \alpha_2+2\bigg)\\
+    \vee & \bigg( \ivarx_{\sigma(i-1)}\ge \gamma \wedge \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)}+\delta \wedge \ivarx_{\sigma(i-1)}\le \alpha(\vec\ivary)-1\bigg)\\
+    \vee & \bigvee_{p=\alpha_1}^{\alpha_2+1} \bigg( \ivarx_{\sigma(i-1)}\ge \gamma \wedge \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)}+\delta \wedge \ivarx_{\sigma(i-1)} = p \wedge \tau[p/\ivarx_{\sigma(i-1)}]\bigg),
+\end{aligned}\end{equation}
+where all exponential occurrences of $\ivarx_{\sigma(i-1)}$ are eliminated.
+
+\hide{
 \[
@@
 \]
-where all exponential occurrences of $\ivarx_{\sigma(i-1)}$ are eliminated.
+}
 
-Similarly, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-2)}$ from $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}]$ as well as  $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}]$, and so on.
+Similarly, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-2)}$ from $\tau'[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}]$ as well as  $\tau'[p /\ivarx_{\sigma(i-1)}]$, and so on.
 Eventually, we obtain a {\pa} formula.
@@
 
-\hide{
-\paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}
+\subsection{Additional optimization techniques}\label{para: opts}
+\paragraph{Eliminate exponential occurrences in atomic formulas simutaneously}
 
-Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be synchronized.
-That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, B^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$.
-Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-Note that when Lemma~\ref{lem:exp-ineq} is applied to an atomic formula $\tau$ to eliminate exponential occurrences of $\ivarx$, it computes 4 parameters, namely $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$, and obtains an equivalent {\pa} formula according to these parameters.
-Since the parameters are used as boundary conditions in enumeration, we can choose the strictest boundary conditions and apply to all atomic formulas.
-
-More specifically, we first compute $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$ 
-for all atomic formulas that contain $10^\ivarx$.
-Let $\alpha^{\min}_1$ denote the minimal element in $\{\alpha^\tau_1\}_\tau$, let $\alpha^{\max}_2,B^{\max},\delta^{\max}$ denote the maximal element in $\{\alpha^\tau_2\}_\tau$, $\{B^\tau\}_\tau$ and $\{\delta^\tau\}_\tau$ respectively.
-Then, we apply $(\alpha^{\min}_1,\alpha^{\max}_2,B^{\max},\delta^{\max})$ to all atomic formulas $\tau$ in parallel and obtain $\varphi'$ equivalent to $\varphi$.
-
-For instance, assume coefficients of $10^\ivarx$ in atomic formulas of $\varphi$ are all non-negative.
-Let $\varphi^+$ denote the formula obtained by replace atomic formulas where coefficients of $10^\ivarx$ are positive by $\ltrue$.
-The resulted formula $\varphi'$ is
-\[
-\small
-\begin{array}{l}
-\bigvee \limits_{p=0}^{B^{\max}-1} ( \ivarx_{\sigma(i-1)} = p \wedge \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) [p/\ivarx_{\sigma(i-1)}])   \\
-\bigvee \left(
-    \begin{array}{l}
-        \ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  - 1  \wedge \\
-        \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} +\delta^{\max} \wedge \varphi^+  
-    \end{array}
-    \right)\\
-%
-\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
-\left(
-\begin{array}{l}
-\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  -1 \ \wedge \\
- \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
- \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
-\end{array}
-\right)\\
-\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
-\left(
-\begin{array}{l}
-\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \ge \alpha^{\max}_2 + 2 \ \wedge \\
- \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
- \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
-\end{array}
-\right)\\
-\bigvee \bigvee \limits_{p = \alpha^{\min}_1}^{\alpha^{\max}_2 + 1}
-\left(
-\begin{array}{l}
-\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} = p\ \wedge \\
- \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}] 
-\end{array}
-\right),
-\end{array}
-\]
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\hide{
-\paragraph{Avoid the formula-size blow-up by depth-first search}
-
-The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size.
-If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory.
-Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
-If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.
-
-\paragraph{Preprocess with small upper bound}
-
-We believe that if a quantifier-free {\paexp} formula is satisfiable, then most probably it is satisfiable with small values assigned to variables.
-Consequently, as a preprocessing step, we put a small upper bound, e.g.
-the biggest constant occurring in the formula, on the values of variables, and perform a depth-first search, so that a model can be quickly found, if there is any.
-If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.
-}
-
-%\zhilin{stopped here}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-In the last section, 
-we introduced quantifier elimination as a decision procedure for {$\paexp$}. According to theorem \ref{thm:string-parInt}, the satisfiability of {$\strint$} fragment is decidable.
-However, it is not practical to directly apply quantifier elimination when solving constraints over {$\paexp$} due to two main reasons.
-
-First, quantifier elimination does not return a model that satisfies the constraints.
-In constrast, it only leaves a formula with constants and free variables.
-To obtain satisfiability, all free variables need to be treated as existential quantified variables and be eliminated one by one until the formula is equal to \textit{true} or \textit{false}.
-Second, the quantifier elimination procedure has a high complexity, both time and space.
-Note that we alternatively invoke QE-exp and Cooper's algorithm to eliminate exponential terms and linear terms of a quantified variable, the length of the formula grows super-exponentially in the procedure.
-
-In this section, we introduced some optimizations on the QE procedure restricted to quantifier-free fragment in {$\paexp$}.
-Given a quantifier-free formula, we wish to decide its satisfiability and find an assignment for variables in the formula if it exists.
-Since the Cooper's algorithm is the main contribution to the complexity, we try to only eliminate exponential terms using an variation of \textbf{Elim-exp} and send the obtained linear formula to SMT-solver.
-The cost is that the algorithm is no longer complete in some situations, in which case some practical restrictions are needed.
-
-A quantifier-free formula is a boolean combination of constraints.
-Here we allow only equalities and inequalities, divisible predicates are replaced by equalities, for example, $2|x$ is replace to $x=2y$, where $y$ is a fresh variable.
-Since we only eliminate exponential terms, the newly introduced $y$ will not bother.
-For convenience, we call variables that occurs exponentially in the formula \emph{exponential variables}, variables that have only linear occurrences \emph{linear variables}.
-Similarly,
-constraints that have exponential terms are called \textit{exponential constraints}, others are called \textit{linear constraints.}
-
-\paragraph{over-approximation}
-
-Given any formula in $\paexp$, we first invoke \textbf{Normalization} to translate it into a simpler form.
-Then we  try to detect if there are obvious conflicts, for example, if the linear constraints combined are unsatisfiable, the whole formula is of course unsatisfiable.
-
-Suppose the normalized formula $\theta$ contains variables $\{x_i:i\in[n]\}$, so the exponential terms could be $e^{x_i}$.
-For each $e^{x_i}$, we subtitute $e^{x_i}$ by a fresh variable $y_i$ in $\theta$ and add the constraints $(e-1)x_i+1\le y_i$ to get $\theta'$.
-Since $(e-1)x_i+1\le e^{x_i}$ holds for all $e,x\in \Nat$, $\theta'$ is an over-approximation of $\theta$ and if $\theta'$ is unsatisfiable over positive reals, then $\theta$ is unsatisfiable over $\Nat$.
-
-This technique also helps us to rule out impossible orders of exponential variables. In this case, constraints $\bigwedge_{1\le i\le n-1} x_{\sigma(i)}<x_{\sigma(i+1)}$ is added in $\theta'$.
-
-\paragraph{modified \textbf{Elim-exp}}\label{para: modified Elim-exp}
-
-\textbf{Elim-exp} is performed recursively on atoms according to orders of all variables.
-The idea is that if the upper bound of variables is given, it is enough to consider orders of exponential variables and this step can be performed on all atoms in the formula simultaneously.
-
-Assume we have $m$ exponential variables with an order $x_1\le ...\le x_m$ and $n-m$ linear variables $x_{m+1},...,x_n$.
-We further assume that there is an upper bound for all variables, say $x_{m+1},...,x_n< e^\gamma$, the bound may be inferred from constraints or specified manually.
-When we try to eliminate exponential terms of $x_i(1\le i\le m)$, an atom with $e^{x_i}$ is of the form
-
-\begin{equation}\label{eq:modify-Elim-exp}
-    \sum_{j=1}^i a_j e^{x_j}+\sum_{k=1}^n b_j x_j\le t 
-\end{equation}
-where $a_m\neq 0$ and $t$ is a constant.
-Note that the form of (\ref{eq:modify-Elim-exp}) is different from formula in theorem \ref{thm:exp-ineq}, exponential variables $x_{i+1},...,x_{m}$ may appear linearly in formula (\ref{eq:modify-Elim-exp}) because we do not invoke \textbf{Elim-linear} to eliminate linear terms.
-Following the method of theorem \ref{thm:exp-ineq}, we seperate terms in the left hand side into the following form.
-\begin{equation}
-    \sum_{j=1}^i (a_j e^{x_j} + b_j x_j) \le t -\sum_{k=i+1}^n b_k x_k \notag
-\end{equation}
-
-The left hand side $\sum_{j=1}^i (a_j e^{x_j} + b_j x_j)$ can be estimated using the same technique in theorem \ref{thm:exp-ineq}.
-For the right hand side, define  $ t_1 \Def t-\gamma (\sum_{k=i+1}^n b_k)$, $t_2\Def t+\gamma (\sum_{k=i+1}^n b_k)$, we have
-\begin{equation}
-    t_1  \le t - \sum_{k=i+1}^n b_k x_k \le  t_2 \notag
-\end{equation}
-
-Similar to \textbf{Elim-exp}, we will discuss 3 subcases,
-assuming w.l.o.g $a_i>0$, define $\alpha_1 \Def l(t_1)-l(a_i)$, $\alpha_2 \Def l(t_2)-l(a_i)$.
-
-\begin{itemize}
-    \item $x_i\le \alpha_1 -1$, corresponds to $\rho_1$ (in \textbf{Elim-exp}), the atom is evaluated \textit{true}
-    \item $\alpha_1 -1\le x_i \le \alpha_2 +2$, corresponds to $\rho_2$ and $\rho_3$
-    \item $\alpha_2 + 2\le x_i$, corresponds to $\rho_4$, the atom is evaluated \textit{false}
-\end{itemize}
-
-For each atom with $e^{x_i}$, we compute $\alpha_1$ and $\alpha_2$ according to coefficients in the atom.
-Select the minimal $\alpha_1$ and the maximal $\alpha_2$, in this way we can perform \textbf{Elim-exp} simultaneously on all atoms in the formula.
-
-Note that the manually specified bound is for variables that only appear linearly in exponential constraints, if there are no variables of this form, the modified algorithm is still complete.
-
-\paragraph{DFS pruning}
-
-Given an order for exponential variables $x_1\le x_2 \le ...
-\le x_m$, the process of \textbf{Elim-exp} from $x_m$ to $x_1$ can be seen as a search tree with depth $m+1$.
-A depth $i(1\le i\le m)$ non-leaf node is a disjunction, each of its child corresponds to a subcase in \textbf{Elim-exp} of $x_i$, a leaf is labeled by a linear formula in which all exponential terms have been eliminated.
-
-In our implement, we adopt a DFS strategy to search the tree to avoid the explosion of length of the formula.
-When eliminating $x_{i-1}$, the upper bound for $x_i$ in some subcases are passed down for pruning.
-
-\paragraph{pre-search}
-To answer \textit{unsat}, the algorithm must search for all possible orders of exponential variables.
-Intuitively, in most satisfiable cases exponential variables are small, so we first give a small bound (roughly the biggest constants in the formula) for exponential variables to perform a pre-search.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Additional optimization techniques}\label{para: opts}
-\paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}
-
-Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be synchronized.
-That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, B^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$.
+Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be conducted simutaneously.
+That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, \gamma^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$.
 Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.
@@
 
-The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size.
-If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory.
-Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
-If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.
+The {\pa} formula resulting from the elimination of exponential occurrences is essentially a disjunction of large number of disjuncts of small size.
+If we store this large formula naively, then its size quickly blows up and exhausts the memory.
+Alternatively, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts.
+When a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.
 
@@
 
-We believe that if a quantifier-free {\paexp} formula is satisfiable, then most probably it is satisfiable with small values assigned to variables.
+We believe that if a quantifier-free {\paexp} formula is satisfiable, then more likely it may be satisfied with an assignment in which all variables are uniformly bounded.
 Consequently, as a preprocessing step, we put a small upper bound, e.g.
-the biggest constant occurring in the formula, on the values of variables, and perform a depth-first search, so that a model can be quickly found, if there is any.
-If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper $u \ge 1$.
+the biggest constant occurring in the formula, on the values of variables, and perform a depth-first search, so that a model can be quickly found if there is any.
+If this preprocessing is unsuccessful, then we continue the search with the greater upper bound $10^u$ for some proper predefined $u \in \mathbb N$.

### 2022-03-18 Acta Informatica2022/7_impl_exp.tex
.

@@
 
-%\vspace*{-2mm}
 \subsection{Implementation}
-We implemented the decision procedure in Wolfram Mathematica, and obtain a solver, called the {\paexp}-solver, which is able to solve the satisfiability of {\paexp} formulas.
+We implement the decision procedure in Wolfram Mathematica, called the {\paexp}-solver, which is able to solve the satisfiability of {\paexp} formulas.
 
-The {\paexp}-solver takes a quantifier-free {\paexp} formula as the input.
-Moreover, it allows specifying a constant upper bound $10^u$ for the values of variables.
-If a constant upper bound $10^u$ is specified, then the problem is to decide whether there is an assignment of values no more than $10^u$ to variables satisfying the given {\paexp} formula.
+The {\paexp}-solver takes a quantifier-free {\paexp} formula as input.
+Moreover, it allows specifying a upper bound $10^u$ to uniformly bound the values of variables.
+Given a upper bound $10^u$, the problem becomes to decide whether there is an assignment in which all variable values are bounded by $10^u$ satisfying the given formula.
 %
-The outputs of the {\paexp}-solver are either ``SAT'', ``UNSAT'', ``B-UNSAT'', or ``TIMEOUT'', corresponding to the facts that the given formula is satisfiable, unsatisfiable, unsatisfiable up to $10^u$, or the search goes beyond the predetermined time limit.
-If the output is ``SAT'', then a model (namely, an assignment of values to variables) is also returned.
+Outputs of the {\paexp}-solver are either ``SAT'', ``UNSAT'', ``B-UNSAT'', or ``TIMEOUT'', standing for the given formula satisfiable, unsatisfiable, unsatisfiable up to $10^u$, or timeout if the solver does not terminate within the time limit.
+If the output is ``SAT'', then a model (namely, an assignment) is returned.
 
@@
 
-To evaluate the performance of the {\paexp}-solver, we created two benchmark suites, ARITHMETIC and STRINGHASH\footnote{The benchmarks are available at https://github.com/EcstasyH/PAexp-Solver/tree/main/Benchmark}.
+To evaluate the performance of the {\paexp}-solver, we design two benchmark suites, ARITHMETIC and STRINGHASH\footnote{The benchmarks are available at https://github.com/EcstasyH/PAexp-Solver/tree/main/Benchmark}.
 %\begin{itemize}
@@
 Moreover, aiming at a better coverage of the syntactical ingredients of {\paexp}, we randomly choose some problem instances and replace the $\le$ symbol of their first inequalities by $=$.
-The constant upper bound for the values of variables is set to be $10^{20}$, motivated by the fact that the largest 64-bit integer is less than $10^{20}$.
+The constant upper bound for the values of variables is set to be $10^{20}$, as the largest 64-bit integer is less than $10^{20}$.
 We also create an SMTLib2 file for each problem instance, to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
-Because neither CVC4 nor Z3 supports the exponential functions natively, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 1$ and $f(n+1) = 10*f(n)$.
+Because neither CVC4 nor Z3 supports the exponential functions directly, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 1$ and $f(n+1) = 10*f(n)$.
 
@@
 This suite comprises two groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
-For one of them, we restrict the nonce string conforming some flat pattern, while for the other one, we allow any word from $\Sigma^*_{\sf num}$ to be used as the nonce.
+We restrict the nonce strings in one group conforming some flat pattern, while for the other one, we allow any word from $\Sigma^*_{\sf num}$ to be used as the nonce.
 %Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, we assume that $\Sigma =\Sigma_{\sf num}$ and $p = 10$ and the string hash function is encoded by $\parseInt$.
@@
 \begin{itemize}
-\item For flat instances, we generate the {\paexp} formulas corresponding to the string constraints, as the inputs to the {\paexp}-solver.
-%
-\item For non-flat instances, we use flat languages $a^* (b_{1} \ldots b_{k})$ to under approximate 
+\item For flat instances, we generate {\paexp} formulas corresponding to the string constraints, as inputs to the {\paexp}-solver.
+\item For non-flat instances, we use flat languages $a^* (b_{1} \ldots b_{k})$ to under-approximate 
 $\Sigma_{\sf num}^*$,  where $a, b_1,\ldots, b_k \in \Sigma_{\sf num}$. 
@@
 %
-We iterate the following procedure until a model is found or the time limit is reached: Initially, set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a$.
+We iterate the following procedure until a model is found or timeout: Initially, set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a$.
 For each assignment, we encode the resulting string constraint into an {\paexp} formula with only one exponential variable.
@@
 
-We compare the {\paexp}-solver against the state-of-the-art SMT solvers on the generated benchmarks. Specifically, 
+We compare the {\paexp}-solver with the state-of-the-art SMT solvers on the generated benchmarks. Specifically, 
 \begin{itemize}
-\item over the ARITHMETIC benchmark suite, we compare the {\paexp}-solver against CVC4 (version 1.8) and Z3 (version 4.8.10),
-\item over the STRINGHASH benchmark suite, we compare the {\paexp}-solver against CVC4, Z3, and Trau\footnote{\url{https://github.com/guluchen/z3/tree/new_trau}}. 
+\item for the ARITHMETIC benchmark suite, we compare the {\paexp}-solver against CVC4 (version 1.8) and Z3 (version 4.8.10),
+\item for the STRINGHASH benchmark suite, we compare the {\paexp}-solver against CVC4, Z3, and Trau\footnote{\url{https://github.com/guluchen/z3/tree/new_trau}}. 
 \end{itemize}
@@
 
-The experiment results are summarized in Table~\ref{table:arithmetic}, where ``Fail'' means either timeout, unknown, or wrong answers.
+The experiment results are summarized in Table~\ref{table:arithmetic}, 
+%and more detailed results of the STRINGHASH benchmark suite are recorded in Table~\ref{table:string}, 
+where ``Fail'' means either timeout, unknown, or wrong answers.
 
-On the ARITHMETIC benchmark suite, the $\paexp$-solver solves around $20\%$-$60\%$ more instances than Z3, and $30\%$-$100\%$ more instances than CVC4.
-Moreover, the gap becomes bigger as the the sizes of the formulas increase, which demonstrates that the $\paexp$-solver is more scalable in solving formulas of greater sizes.
+For the ARITHMETIC benchmark suite, the $\paexp$-solver solves around $20\%$-$60\%$ more instances than Z3, and $30\%$-$100\%$ more instances than CVC4.
+Moreover, the gap becomes bigger as the the sizes of the formulas increase, which demonstrates that the $\paexp$-solver is more efficient in solving formulas of greater sizes.
 The average time of the $\paexp$-solver is comparable with Z3 and CVC4.
@@
 
-On the STRINGHASH benchmark suite, in overall, the $\paexp$-solver solves significantly more problem instances, especially those satisfiable instances, than Z3, CVC4, and Trau.
+For the STRINGHASH benchmark suite, in overall, the $\paexp$-solver solves significantly more instances, especially those satisfiable instances, than Z3, CVC4, and Trau.
 For instance, for flat regular constraints, the $\paexp$-solver solves almost all 300 problem instances, except 3 of them\footnote{These three instances can actually be solved in 70 seconds.}, while  Z3, CVC4, Trau solve only 34, 89, 187 instances respectively.
@@
 it reports ``UNSAT'' for some satisfiable instances.
-%While the $\paexp$-solver solves more instances than the other solvers,  
 From the results, we can see that the $\paexp$-solver achieves a good tradeoff between precision and efficiency, although it is slower than the other solvers. 
-(More detailed experiment results on STRINGHASH can be found in Table~\ref{table:string}.)
-\wuhao{The two tables may be too big.}
+%(More detailed experiment results of STRINGHASH can be found in Table~\ref{table:string}.)
 
@@
 \begin{table*}[ht]
-        \caption{Experimental Results, O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
+    \caption{Experimental Results.
+The left table summarizes the results for both experiments, and more detailded results of STRINGHASH benchmark suite are presented in the right table.
+O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail,
     $\#$: number of problems, $T$: average time in seconds}
@@
     \renewcommand{\arraystretch}{1.1} 
+    \resizebox{5.5cm}{!}{
         \begin{tabular}{ c c c c c c c c c c }
@@
         \end{tabular}
-            \label{table:arithmetic}
-
-\end{table*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hide{
-    \begin{table*}[ht]
-        \caption{Experimental Results, O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
-    $\#$: number of problems, $T$: average time in seconds}
-    \centering
-    \renewcommand{\arraystretch}{1.1}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-    \hline
-        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c|}{Z3} & \multicolumn{2}{c|}{CVC4} &  \multicolumn{2}{c|}{Trau} & \multicolumn{2}{c |}{$\paexp$} \\ 
-        \cline{3-10}
-        &  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$} & \scriptsize{$\#$} & \scriptsize{$T$}  \\ \hline
-        \multirow{4}{*}{
-        \newline \scriptsize{(2,3,3,4)}} & S & 56 & {\bf 0.4} & 42 & 2.3 & - & - &  {\bf 64} & {\bf 0.4} \\
-        \cline{2-10}
-         & U & 69 & {\bf 0.1} & 72 &  {\bf 0.1} & - & - &  {\bf 89} & {\bf 0.1} \\
-         \cline{2-10}
-         & B & - & - & - & - & - & - &  47 & 9.5 \\
-         \cline{2-10}
-         & F & 75 & - & 86 & - & - & - &  {\bf 0} & - \\ \hline
-         \cline{1-10}
-        \multirow{4}{*}{\scriptsize{(3,4,4,5)}} & S & 33 & {\bf 1.4} & 25 & 2.9 & - & - &  {\bf 52} & 3.3 \\
-        \cline{2-10}
-         & U & 59 & {\bf 0.1} & 60 & {\bf 0.1} & - & - &  {\bf 88} & {\bf 0.1} \\
-         \cline{2-10}
-         & B & -  & - &  - & -  & - & - &  1 & 54.0 \\
-         \cline{2-10}
-         & F & 108 & - & 115 & - & - & - &  {\bf 59} & - \\ \hline
-         \cline{1-10}
-        \multirow{4}{*}{\scriptsize{(4,5,5,6)}} & S & 35 & {\bf 1.8} & 19 & 6.6 & - & - &  {\bf 47} & 22.4 \\
-        \cline{2-10}
-         & U & 36 & 0.3 & 39 & 0.4 & - & - &  {\bf 72} & {\bf 0.1} \\
-         \cline{2-10}
-         & B & -  & -  & - & - & - & - &  0 & - \\
-         \cline{2-10}
-         & F & 129 & - & 142 & - & - & - &  {\bf 81} & -\\
-         \hline
-         \cline{1-10}
-        \multirow{3}{*}{\scriptsize{Flat}} 
-         & S & 34 & 19.0 & 88 & 12.7 & 5 & {\bf 0.1}& {\bf 115}  & 12.3\\
-        \cline{2-10}
-         & U & 0 & - & 1 & 4.0 & 182 & {\bf 2.5} & {\bf 182} & 47.7 \\
-         \cline{2-10}
-         & F & 266 & - & 211 & - & 113 & - & {\bf 3} & - \\ \hline
-         \cline{1-10}
-        \multirow{3}{*}{\scriptsize{Non-flat}}
-         & S & 210 & 7.8 & 144 & 4.9 & 55 & 5.9 &{\bf 300} & 16.7 \\
-        \cline{2-10}
-         & U & 0 & - & 0 & - & 0 & - & 0 & - \\
-         \cline{2-10}
-         & F & 90 & - & 156 & - & 245 & - & {\bf 0} & - \\ \hline
-         \cline{1-10}
-        \end{tabular}
-            \label{table:string}
-\end{table*}
     }
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[ht]
-    \caption{Experimental results on STRINGHASH,  O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
-    $\#$: number of problems, $T$: average time in seconds}
-    \centering
-
-    \renewcommand{\arraystretch}{1.1}
+    \resizebox{5.5cm}{!}
+    {
     \begin{tabular}{cccccccccc}
@@
        \end{tabular}
+    }
+    \label{table:arithmetic}
+\end{table*}
+
+% put two tables together
+\hide{
+\begin{table}[ht]
+    \caption{Experimental results on STRINGHASH,  O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
+    $\#$: number of problems, $T$: average time in seconds}
+    \centering
+    \renewcommand{\arraystretch}{1.1}
+    \resizebox{6cm}{!}
+    {
+    \begin{tabular}{cccccccccc}
+    \hline
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c}{Z3} & \multicolumn{2}{c}{CVC4} &  \multicolumn{2}{c}{Trau} & \multicolumn{2}{c}{$\paexp$} \\
+       &  & $\#$ & $T$ &  $\#$ & $T$ & $\#$ & $T$ &  $\#$ & $T$ \\ 
+       \hline
+       \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345(w_1)^+(w_2)^+$}} & S & 5 & 14.0 & 29 & 8.5 & 3 & {\bf  0.1} & {\bf 37} & 9.9 \\
+         \cline{2-10}
+        & U & 0 & - & 0 & - & {\bf 60} & {\bf 1.3} & {\bf 60} & 47.2 \\
+         \cline{2-10}
+        & F & 95 & - & 71 & - & 37 & - & {\bf 3} & - \\ 
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$\begin{array}{l}12345(w_1)^+ \\
+       \hspace{2mm}(w_2)^+6789\end{array}$}} & S & 11 & 13.0 & 29 & 12.0 & 0 & - & {\bf 37} & {\bf 10.6} \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & {\bf 63} & {\bf 1.2} & {\bf 63} & 50.0 \\
+         \cline{2-10}
+        & F & 89 & - & 71 & - & 37 & - & {\bf 0} & - \\ 
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$(w_1)^+(w_2)^+6789$}} & S & 18 & 24.0 & 30 & 9.3 & 2 & {\bf  0.1} & {\bf 41} & 16.1 \\
+      	 \cline{2-10}
+        & U & 0 & - & 1 & 4.0 & {\bf 59} & {\bf 2.5} & {\bf 59} & 45.8 \\
+         \cline{2-10}
+        & F & 82 &  & 69 &  & 39 &  & {\bf 0} &  \\ \hline
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}$}} & S & 82 & 8.7 & {\bf 100} & {\bf 2.2} & 28 & 5.9 & {\bf 100} & 18.5 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 18 & - & {\bf 0} & - & 72 & - & {\bf 0} & - \\ 
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$12345\Sigma^*_{\sf num}6789$}} & S & 60 & 9.3 & 17 & 7.8 & 3 & {\bf 0.3} & {\bf 100} & 16.0 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 40 & - & 83 & - & 97 & - & {\bf 0} & - \\
+        \cline{1-10}
+       \multirow{3}{*}{\scriptsize{$\Sigma^*_{\sf num}6789$}} & S & 68 & 5.5 & 27 & 13.0 & 24 & {\bf 9.0} & {\bf 100} & 15.7 \\
+        \cline{2-10}
+        & U & 0 & - & 0 & - & 0 & - & 0 & - \\
+         \cline{2-10}
+        & F & 32 & - & 73 & - & 76 & - & {\bf 0} &- \\
+        \hline
+       \end{tabular}
+    }
            \label{table:string}
 \end{table}
+}

### 2022-03-18 Acta Informatica2022/8_conclusion.tex
.

--- 2021-09-29 FMCAD2021/8_conclusion.tex
@@
 In this paper, we proposed a complete flattening approach for string constraints with string-integer conversion and flat regular constraints, based on a quantifier elimination procedure by Point in 1986, for the extension of Presburger arithmetic with exponential functions.
-We gave a more accessible reformulation of Point's procedure and proposed various optimizations.
-Moreover, we achieved the first prototypical implementation of Point's procedure.
+We gave a more accessible reformulation of Point's procedure and for the first time a complexity upper bound.
+Moreover, we proposed various optimizations and achieved the first prototypical implementation of Point's procedure.
 We also did extensive experiments to evaluate the performance of the implementation.

### 2022-04-01 Acta Informatica2022/1_intro.tex
before modify sec 4

@@
 
-The emerging of scripting languages boosted the needs of efficient approaches and tools to ensure program quality.
+The emerging of scripting languages boosts the needs of efficient approaches and tools to ensure program quality.
 Comparing with traditional programming languages, string data type plays a more critical role in its analysis.
@@
 This approach is taken in the first-generation solvers such as Hampi~\cite{KiezunGAGHE12} and Kuluza~\cite{SaxenaAHMMS10}.
-Although they are useful in handling many practical cases, they fail to find an answer when the all string solutions exceed the selected bound.
-For example, a constraint of the form $x.y \neq z  \wedge \vert x\vert  > 2000$ would be quite challenging to handle using those solvers.
+Although they are useful in handling many practical cases, they fail to find an answer when all string solutions exceed the selected bound.
+For example, a constraint of the form $x.y \neq z  \wedge \vert x\vert  > 2000$ would be quite challenging using those solvers.
 
@@
 
-The paper of Abdulla et al.~\cite{Parosh:20:PLDI} has considered adding string-integer conversion constraints to the above class,
-which proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
+The paper of Abdulla et al.~\cite{Parosh:20:PLDI} considered adding string-integer conversion constraints to the above class,
+and proposed an algorithm for a restricted form of flat languages and left the support of general flat languages as an open problem.
 For string-integer conversion constraints, their approach projects the solution to 
-a finite solution space (in a way similar to the PASS~\cite{goudong2013pass} approach).
+a finite solution space (in a way similar to the PASS approach~\cite{goudong2013pass}).
 
@@
 This can be seen as another contribution of this paper.
-We also analyze the deterministic upper bound of complexity for the first time: given a $\paexp$ formula with only existential quantifiers, eliminating all quantifiers has a 3-EXPSPACE complexity.
+We also analyze the deterministic upper bound of complexity for the first time: given an $\paexp$ formula with only existential quantifiers, eliminating all quantifiers has a 3-EXPSPACE complexity.
 Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).
@@
 \paragraph*{Structure}
-After the preliminaries in Section~\ref{sec:pre}, we show how to flatten a string constraint with string-integer conversion to an existential $\paexp$ formula in Section~\ref{sec:string-solving}.
+After introducing preliminaries in Section~\ref{sec:pre}, we present how to flatten a string constraint with string-integer conversion to an existential $\paexp$ formula in Section~\ref{sec:string-solving}.
 We describe the quantifier elimination procedure for $\paexp$ in Section~\ref{sec-dec} and analyse its complexity in Section~\ref{sec-cpx}.
-Several optimization techniques are presented in Section~\ref{sec-opt}.
-Finally, we describe the implementation and experiment results in Section~\ref{sec-exp}.
+Several optimization techniques are provided in Section~\ref{sec-opt}.
+Finally, we report the implementation and experiment results in Section~\ref{sec-exp}.

### 2022-04-01 Acta Informatica2022/3_flatten.tex
before modify sec 4

@@
 an integer term, denoted by $\iterm$, is of the form
-$$\iterm\  \Def\  n \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,$$
+$$\iterm\  \Def\  c \mid \ivarx \mid \len(\sterm) \mid \parseInt(\sterm) \mid \iterm + \iterm \mid \iterm - \iterm,$$
 and a $\strint$ formula, denoted by $\varphi$ is of the form 
-$$\varphi\ \Def\ \sterm = \sterm \mid x \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi$$
-where $a \in (\Sigma_{\textit{num}})_\varepsilon$, $n \in \Nat$, $\Aut$ is an FA, $\op \in \{=, < , >, \le, \ge\}$ and $\len(\sterm)$ denotes the length of a string $\sterm$. 
-Let us call $\sterm = \sterm$ as string equality constraints, $x \in \Aut$ as regular constraints, $\iterm \op \iterm$ as arithmetic constraints. 
+$$\varphi\ \Def\ \sterm = \sterm \mid \sterm \in \Aut \mid \iterm\ \op\ \iterm \mid \varphi \wedge \varphi \mid \varphi \vee \varphi \mid \neg \varphi$$
+where $a \in (\Sigma_{\textit{num}})_\varepsilon$, $c \in \mathbb Z$, $\Aut$ is an FA, $\op \in \{=, < , >, \le, \ge\}$ and $\len(\sterm)$ denotes the length of a string $\sterm$. 
+Let us call $\sterm = \sterm$ as string equality constraints, $\sterm \in \Aut$ as regular constraints, $\iterm \op \iterm$ as arithmetic constraints. 
 Let  $\svar(\varphi)$ and $\ivar(\varphi)$ denote the set of string variables and integer variables occurring in $\varphi$ respectively.
 
-The $\strint$ formulas are interpreted on $I=(I_s, I_i)$ where $I_s$ is a partial function from $\svars$ (the set of string variables) to $\Sigma^*$ and $I_i$ is a partial function from $\ivars$ (the set of integer variables) to $\Nat$.
+The $\strint$ formulas are interpreted on $I=(I_s, I_i)$ where $I_s$ is a partial function from $\svars$ (the set of string variables) to $\Sigma^*$ and $I_i$ is a partial function from $\ivars$ (the set of integer variables) to $\mathbb Z$.
+% nat -> Z
 Moreover, it is required that the domains of $I_s, I_i$ are finite.
@@
 
-\subsection{An Extension of Presburger Arithmetic with Exponential Functions (\paexp)}
+\subsection{Presburger Arithmetic with Exponential Functions (\paexp)}
 
@@
 The syntax of {\paexp} is obtained from that of {\pa} by adding $10^\iterm$ and $\ell_{10}(\iterm)$ to the definition of terms. An {\paexp} term, denoted by $\iterm$, is of the form
-$$\iterm\ \Def\ c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^\iterm \mid \ell_{10}(\iterm) $$
+$$\iterm\ \Def\ c \mid \ivarx \mid \iterm + \iterm \mid \iterm - \iterm \mid 10^{\iterm} \mid \ell_{10}(\iterm) $$
 where $\ivarx$ is an integer variable and $c\in \mathbb Z$ is an constant integer.
+In addition, we require that $10^\iterm$ and $\ell_{10}(\iterm)$ terms are well-defined, which restricts the interpretations of $\iterm$ therein.
 
@@
 
+The semantics of {\paexp} are defined similarly to that of {\pa} with the only difference that partial functions are required to be well-defined.
+The quantifier-free and existential {\paexp} formulas are defined similarly to {\pa} as well.
+We also assume that quantifier-free {\paexp} formulas are free of negations, for the same reason we have explained for {\pa}.
+
+\hide{
 The semantics of {\paexp} are defined similarly to {\pa} with the only difference that variables are interpreted over $\Nat$ rather than $\mathbb Z$, that is, an interpretation is a function $I:\free(\phi)\to \mathbb \Nat$.
-This restriction ensures that terms $10^\iterm$ and $\ell_{10}(\iterm)$ are well-defined: In the \emph{normalization} step of the quantifier elimination procedure, for each $\ell_{10}(\iterm)$ term, we will replace all its occurrences by a fresh variable $\ivarz$, and add a conjunct $10^\ivarz \le \iterm < 10*10^\ivarz$.
+This restriction ensures that terms $10^\iterm$ and $\ell_{10}(\iterm)$ are well-defined: In the \emph{normalization} step of the quantifier elimination procedure, for each $\ell_{10}(\iterm)$ term, we will replace all its occurrences by a fresh variable $\ivarz$, and add a conjunct $10^\ivarz \le \iterm < 10\cdot 10^\ivarz$.
 Then $\ivarz\in\Nat$ guarantees that $\iterm \ge 1$ is always satisfied.
@@
 Note that we can extend the interpretation to $\mathbb Z$ since each integer can be written as the difference of two positive integers.
-
-The quantifier-free and existential {\paexp} formulas are defined similarly to {\pa} as well.
-We also assume that quantifier-free {\paexp} formulas are free of negations, for the same reason we have explained for {\pa}.
+}
 
@@
 The flattening of $\varphi \in \strint$ under a flat domain restriction $\flatfun_\varphi$ is a {\paexp} formula, denoted by $\flatten_{\flatfun_\varphi}(\varphi)$, that encodes its flattened semantics.
-More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and flattening variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$, plus some other auxiliary variables, such that
+More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and flattening variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{l_{\svarx,i} \mid i \in [k_\svarx]\}$, plus some other auxiliary variables, such that
 $$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket \vert_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)}).$$
@@
 Under a flat domain restriction $\flatfun_\varphi$, the flattening of $\varphi \in \strint$ is an $\paexp$ formula, denoted by $\flatten_{\flatfun_\varphi}(\varphi)$, that encodes the flattened semantics $\llbracket \varphi \rrbracket_{\flatfun_\varphi}$.
-The idea is that $x$ can be expressed as $(w_{\svarx,1})^{\#_{x,1}} \cdots (w_{\svarx, k_\svarx})^{\#_{x,k_x}}$ with $\#_{x,i}$ denoting the number of repeatitions of $w_{x,i}$,  
-so we can eliminate the occurrences of $x$ in $\varphi$ by introduce a set of integer variables $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$. 
+The idea is that $x$ can be expressed as $(w_{\svarx,1})^{l_{x,1}} \cdots (w_{\svarx, k_\svarx})^{l_{x,k_x}}$ with $l_{x,i}$ denoting the number of iterations of $w_{x,i}$,  
+so we can eliminate the occurrences of $x$ in $\varphi$ by introduce a set of integer variables $\pvar_{\flatfun_\varphi}(\svarx) = \{l_{\svarx,i} \mid i \in [k_\svarx]\}$. 
 More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and flattening variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$ plus some other auxiliary variables, such that the interpretations $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \mathbb Z$ of $\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket$and interpretations $(I_s, I_i)$ of $\llbracket \varphi \rrbracket_{\flatfun_\varphi}$ have the following correspondence
@@
     \item  
-    for every $ \svarx \in \svar(\varphi)$, $I_s(\svarx) = w_{\svarx,1}^{I_e(\#_{\svarx,1})} \ldots  w_{\svarx,k_\svarx}^{I_e(\#_{\svarx,k_\svarx})}$, 
+    for every $ \svarx \in \svar(\varphi)$, $I_s(\svarx) = w_{\svarx,1}^{I_e(l_{\svarx,1})} \ldots  w_{\svarx,k_\svarx}^{I_e(l_{\svarx,k_\svarx})}$, where $I_e(l_{x,i})\ge 0$ for $1\le i\le k_x$.
     %
@@
 \item for $i = k_\svarx$, 
-$$\iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{\len(w_{\svarx,k_\svarx})\#_{\svarx, k_\svarx}} -1 )}{(10^{\len(w_{\svarx, k_\svarx}) } -1 )}$$ 
+$$\iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{\len(w_{\svarx,k_\svarx})l_{\svarx, k_\svarx}} -1 )}{(10^{\len(w_{\svarx, k_\svarx}) } -1 )}$$ 
 %
@@
 %
-$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{\len(w_{\svarx, i})  \#_{\svarx, i} } -1 ) 10^{\big(\sum_{i+1\le j\le k_x} \len(w_{\svarx, j})  \#_{\svarx, j}\big)}} {(10^{\len(w_{\svarx, i}) } -1 )} + \iterm_{\svarx, i+1}$$
+$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{\len(w_{\svarx, i})  l_{\svarx, i} } -1 ) 10^{\big(\sum_{i+1\le j\le k_x} \len(w_{\svarx, j})  l_{\svarx, j}\big)}} {(10^{\len(w_{\svarx, i}) } -1 )} + \iterm_{\svarx, i+1}$$
 \end{itemize}
 %
-\item Notice that here $\len( w_{\svarx, \textunderscore } )$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables. So we have 
-$$\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} \len(w_{\svarx, i})  \#_{\svarx, i}.$$ 
+\item Notice that here $\len( w_{\svarx, \textunderscore } )$ are constants while $l_{\svarx, \textunderscore}$ are (flattening) variables. So we have 
+$$\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} \len(w_{\svarx, i})  l_{\svarx, i}.$$ 
 \end{itemize}
@@
 & \flatten_{\flatfun_\varphi}(\parseInt(\svarx)  =  2\ivarx)  \\
-\Def & 1 \frac{10^{\#_{\svarx,1}}-1}{10-1}10^{\#_{\svarx,2}}  + 2 \frac{10^{\#_{\svarx,2}}-1}{10-1} = 2\ivarx   \\
-\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} - 10^{\#_{\svarx,2}}  + 2*10^{2\#_{\svarx,2}}- 2 = 18\ivarx \\
-\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} +  10^{\#_{\svarx,2}} = 18\ivarx + 2.
+\Def & 1 \frac{10^{l_{\svarx,1}}-1}{10-1}10^{l_{\svarx,2}}  + 2 \frac{10^{l_{\svarx,2}}-1}{10-1} = 2\ivarx   \\
+\equiv & 10^{l_{\svarx,1}+l_{\svarx,2}} - 10^{l_{\svarx,2}}  + 2*10^{2l_{\svarx,2}}- 2 = 18\ivarx \\
+\equiv & 10^{l_{\svarx,1}+l_{\svarx,2}} +  10^{l_{\svarx,2}} = 18\ivarx + 2.
 \end{array}

### 2022-04-01 Acta Informatica2022/4_dec_proc.tex
.
before modify sec 4

@@
 However, Semënov did not give a concrete quantifier elimination procedure.
-Remedying this, Point proposed a quantifier elimination procedure for {\paexp} in \cite{Point86}.
-In this section, we describe how Point's quantifier elimination procedure works.
+Later, Point proposed a quantifier elimination procedure for {\paexp} in \cite{Point86}.
+In this section, we explain how Point's quantifier elimination procedure works.
 
-%
+\vspace*{-3mm}
+
 \begin{theorem}[\cite{Point86}]
@@
 \end{theorem}
+\vspace*{-3mm}
 
@@
 
-    \caption{Quantifer elimination for \paexp}
+    \caption{Point's quantifer elimination procedure for \paexp}
 \end{algorithm2e}
@@
 
-\hide{ % do not allow \neg in PAexp
-\item \textit{NNF transformation} At first, we transform $\varphi'(\ivarx,\vec{\ivary})$ into the NNF (negation normal form). Moreover, we remove the occurrences of $\neg$ by replacing 
-(a) $\neg c \vert  \iterm$ with $ c \nmid \iterm$, 
-(b) $\neg (\iterm_1 = \iterm_2)$ with $\iterm_1 < \iterm_2 \vee \iterm_2 < \iterm_1$, 
-(c) $\neg (\iterm_1 < \iterm_2)$ with $\iterm_2 \le \iterm_1$, 
-(d) $\neg (\iterm_1 \le \iterm_2)$ with $\iterm_2 < \iterm_1$, and so on.
-}
-
-\item \textit{replace $\ell_{10}(\iterm)$ terms} Repeat the following procedure, until there are no $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: for each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10^{\ivarz + 1}$ as a conjunct.
+\item \textit{replace $\ell_{10}(\iterm)$ terms} Repeat the following procedure, until there are no $\ell_{10}(\iterm)$ with $\ivarx$ occurs in $\iterm$: for each occurrence of $\ell_{10}(\iterm)$ such that $\ivarx$ occurs in $\iterm$, introduce a fresh variable, say $\ivarz$, and replace all occurrences of $\ell_{10}(\iterm)$ by $\ivarz$, moreover, add the constraint $10^\ivarz \le \iterm < 10\cdot 10^\ivarz$ as a conjunct.
 Note that if $\iterm$ contains no variables, then $\ell_{10}(\iterm)$ is a constant.
@@
 \item Let $\vec{\ivarz} = \ivarz_1,\ldots, \ivarz_n$ be an enumeration of the freshly introduced variables. Then the result of the normalization procedure is 
-$\exists \vec{\ivarz}\exists \ivarx.\ \varphi^\dag$.
+$\exists \ivarx\exists \vec{\ivarz}.\ \varphi^\dag$.
 \end{enumerate}
@@
 
-%\vspace{-2mm}
 \subsection{Enumeration of the variable orders}\label{qe-enum}
@@
 In the following, we illustrate how to eliminate the exponential occurrences of $\ivarx_{\sigma(i)}$ for inequality atomic formulas of the form
+
+\vspace*{-2mm}
+
 \begin{equation}\label{tau}
@@
 Intuitively, 
-The lemma states that when the left-hand-side of the inequality is dominated by the $a_i 10^{\ivarx_{\sigma(i)}}$ term, if $\ivarx_{\sigma(i)}\le \alpha(\vec{\ivary})-1$ or $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2$, it can be determined directly whether the formula is true or false.
+the lemma states that when the left-hand-side of $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}, \vec{\ivary})$ is dominated by $a_i 10^{\ivarx_{\sigma(i)}}$ and $\ivarx_{\sigma(i)}\le \alpha(\vec{\ivary})-1$ (or $\ivarx_{\sigma(i)} \ge \alpha(\vec{\ivary})+2$),
+the truth value can be determined directly.
 
@@
 
+\vspace*{-3mm}
+
 \begin{proposition} \label{prop:linear}
@@
 \end{proposition}
+
+\vspace*{-3mm}
 

### 2022-04-01 Acta Informatica2022/5_complexity.tex
before modify sec 4

@@
 \begin{lemma}\label{lem:cpx exp}
-    \begin{align}
+$$c'_0\le {c_0}^2 \qquad s'_0\le m{s_0}^2 \qquad a'_0\le s_0a_0$$ 
+\end{lemma}
+
+\hide{\begin{align}
     c'_0&\le c_0^2 \notag \\
@@
     a'_0&\le s_0a_0\notag 
-    \end{align} 
-\end{lemma}
-
+\end{align}}
 \begin{proof}
@@
 We know that each atomic formula $\tau$ with exponential occurrence of $\ivarx_m$ is replaced by a new formula.
-Note that coefficients of linear occurrences of $\ivarx_i$, for $i\le m-2$, remain changed throughout the substitutions, only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant 1 is introduced as a coefficient for $\ivarx_m$, and if we substitute $\ivarx_m$ by $\ivarx_{m-1}+p$ for some constant $p$, coefficient of linear occurrence of $\ivarx_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $\ivarx_{m}$ and $\ivarx_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
-Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le c_0^2$.
+Note that coefficients of linear occurrences of $\ivarx_i$, for $i\le m-2$, remain unchanged throughout the substitutions, only the coefficients of linear occurrences of $\ivarx_m$ and $\ivarx_{m-1}$ will be changed: constant 1 is introduced as a coefficient for $\ivarx_m$, and if we substitute $\ivarx_m$ by $\ivarx_{m-1}+p$ for some constant $p$, coefficient of linear occurrence of $\ivarx_{m-1}$ will become $b_m+b_{m-1}$ ($b_m,b_{m-1}$ are coefficients for $\ivarx_{m}$ and $\ivarx_{m-1}$ in $\tau$, see Lemma~\ref{lem:exp-ineq} ).
+Since the new coefficient is obtained by adding two linear coefficient together, we have $c'_1\le {c_0}^2$.
 Note that $\delta$ and $B$ in Lemma~\ref{lem:exp-ineq} are at most $\ell_{10}(ms_0)$.
-When we substitute $\ivarx_m$ by $\ivarx_{m-1}+\delta-1$ or by $B-1$, the largest constant in the formula becomes at most $s_0\cdot  10^{\ell_{10}(ms_0)}\le ms_0^2$.
+When we substitute $\ivarx_m$ by $\ivarx_{m-1}+\delta-1$ or by $B-1$, the largest constant in the formula becomes at most $s_0\cdot  10^{\ell_{10}(ms_0)}\le m{s_0}^2$.
 And an inequality is replaced by at most $4\ell_{10}(ms_0)$ atomic formulas, so $a'_1\le 4\ell_{10}(ms_0)a_0$.
@@
 
-Choose larger upper bounds for $c'_1$, $s'_1$ and $a'_1$ respectively, then the lemma is proved.
-
-\end{proof}
+Choose larger upper bounds for $c'_1$, $s'_1$ and $a'_1$ respectively, then the lemma is proved.\end{proof}
 
@@
 \begin{lemma}\cite{Oppen73}\label{lem:cpx oppen}
-    \begin{align}
-        c_1&\le c_0'^4 \notag \\
-        s_1&\le s_0'^{4c_0'} \notag\\
-        a_1&\le {a_0'}^4 s_0'^{2c_0'}\notag 
-    \end{align} 
+$$c_1\le {c_0'}^4\qquad
+s_1\le {s_0'}^{4c_0'}\qquad
+a_1\le {a_0'}^4 {s_0'}^{2c_0'}
+$$
 \end{lemma}
 
+\hide{    \begin{align}
+    c_1&\le {c_0'}^4 \notag \\
+    s_1&\le {s_0'}^{4c_0'} \notag\\
+    a_1&\le {a_0'}^4 {s_0'}^{2c_0'}\notag 
+\end{align} }
+
 Combining Lemma~\ref{lem:cpx exp} and Lemma~\ref{lem:cpx oppen}, we have
-\begin{align}
-    c_1&\le c_0^8 \notag \\
-    s_1&\le (ms_0^2)^{4c_0^2} \notag\\
-    a_1&\le (s_0a_0)^4(m_0s_0^2)^{2c_0^2}\notag 
-\end{align}
+$$c_1\le {c_0}^8\qquad
+s_1\le (m{s_0}^2)^{4{c_0}^2} \qquad
+a_1\le (s_0a_0)^4(m_0{s_0}^2)^{2{c_0}^2} 
+$$
+
+\hide{\begin{align}
+    c_1&\le {c_0}^8 \notag \\
+    s_1&\le (m{s_0}^2)^{4{c_0}^2} \notag\\
+    a_1&\le (s_0a_0)^4(m_0{s_0}^2)^{2{c_0}^2}
+\end{align} }
 
@@
 \begin{lemma}
-    \begin{align}
-        c_k&\le c_0^{8^k} \notag \\
-        s_k&\le N^{(4c_0)^{ 8^k}} s_0^{(8c_0)^{ 8^k}} \notag\\
-        a_k&\le a_0^{4^k}n^{(4c_0)^{ 8^k}} s_0^{(8c_0)^{ 8^k}} \notag
-    \end{align} 
+$$c_k\le {c_0}^{8^k}\qquad
+s_k\le N^{(4c_0)^{ 8^k}} {s_0}^{(8c_0)^{ 8^k}} \qquad
+a_k\le {a_0}^{4^k}n^{(4c_0)^{ 8^k}} s_0^{(8c_0)^{ 8^k}} $$
 \end{lemma}
 
+\hide{    \begin{align}
+    c_k&\le {c_0}^{8^k} \notag \\
+    s_k&\le N^{(4c_0)^{ 8^k}} {s_0}^{(8c_0)^{ 8^k}} \notag\\
+    a_k&\le {a_0}^{4^k}n^{(4c_0)^{ 8^k}} s_0^{(8c_0)^{ 8^k}} \notag
+\end{align} 
+}
 If we adopt the assumptions in \cite{Oppen73}, by assuming $c_0\le N$, $a_0\le N$ and $s_0\le N$, the space required to store the quantifier free formula $\varphi_k$, is bounded by the product of the number of linear orders $m!$, the number of atoms $a_k$, the maximum number of constants $2m+2$ per atom, the maximum amount of space $s_k$ to store each constant and some constant $q$.
@@
 The above analysis is focused on the existential {\paexp} formulas and is sufficient for our string constraints solving context. 
-But what about general {\paexp} formulas that contain alternating quantifiers?
-Unfortunately, directly extending our analysis on alternating quantifers by transforming $\forall \ivarx.\varphi$ into $\neg \exists \ivarx.\neg \varphi$ will not give an elementary upper bound.
-We will explain the main idea without going into details on this issue.
+Unfortunately, directly extending our analysis on general {\paexp} formulas with alternating quantifiers by transforming $\forall \ivarx.\varphi$ into $\neg \exists \ivarx.\neg \varphi$ will not give an elementary upper bound.
+
+\hide{We will explain the main idea without going into details on this issue.
 To eliminate quantifiers in a formula, say $\forall \ivarx_2\exists \ivarx_1.\varphi(\ivarx_1,\ivarx_2,\vec \ivary)$, we first eliminate $\exists \ivarx_1$, treating $\ivarx_2$ and $\vec \ivary$ both as free variables.
 The subprocedure to eliminate exponential occurrences of $\ivarx_1$ is at the cost of introducing logarithmic expressions of $\ivarx_2$ and $\vec \ivary$, which in the worst case increases \emph{exponentially} the number of fresh variables in the normalization step w.r.t $\ivarx_2$.
-We conjecture that the space complexity of quantifier elimination procedure over {\paexp} formulas still has an elementary upper bound, which may require a more detailed analysis or improvements of the algorithm.
+We conjecture that the space complexity of quantifier elimination procedure over {\paexp} formulas still has an elementary upper bound, which may require a more detailed analysis or improvements of the algorithm.}

### 2022-04-01 Acta Informatica2022/6_optimization.tex
before modify sec 4

@@
 
-Let $\alpha_1 = \ell_{10}(c_1) -  \ell_{10}(a_{i-1})$ and $\alpha_2 = \ell_{10}(c_2) -  \ell_{10}(a_{i-1})$. Then from Lemma~\ref{lem:exp-ineq}, let $A\Def 1+\sum_{j=1}^{i-2}\vert a_j\vert$, 
+Let $\alpha_1 = \ell_{10}(c_1) -  \ell_{10}(a_{i-1})$ and $\alpha_2 = \ell_{10}(c_2) -  \ell_{10}(a_{i-1})$. 
+Then from Lemma~\ref{lem:exp-ineq}, let $A\Def 1+\sum_{j=1}^{i-2}\vert a_j\vert$, 
 $B \Def 1+\sum_{j=1}^{i-1}\vert b_j\vert$, 
@@
 and $\gamma \Def 2\ell_{10}(B)+3$. 
+When $\ivarx_{\sigma(i-1)} \ge \gamma$ and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$
+
 \begin{itemize}
-\item if $\ivarx_{\sigma(i-1)} \ge \gamma$, $\ivarx_{\sigma(i-1)} \le \alpha_1 - 1$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-lb}), thus also $\tau'$, is evaluated to $\ltrue$,
-\item if $\ivarx_{\sigma(i-1)} \ge \gamma$, $\ivarx_{\sigma(i-1)} \ge \alpha_2 + 2$, and $\ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} + \delta$, then inequality (\ref{eqn-ub}), thus also $\tau'$, is evaluated to $\lfalse$.
+\item if $\ivarx_{\sigma(i-1)} \le \alpha_1 - 1$, then inequality (\ref{eqn-lb}), thus also $\tau'$, holds.
+\item if $\ivarx_{\sigma(i-1)} \ge \alpha_2 + 2$, then inequality (\ref{eqn-ub}), thus also $\tau'$, \textbf{does not} holds.
 \end{itemize}
@@
 Therefore, $\tau'$ and $\tau$ are equivalent to 
-\begin{equation}\label{lemma-example}\begin{aligned}
+\begin{equation}\label{opt-example}\begin{aligned}
     &\bigvee_{p=0}^{\gamma-1} (\ivarx_{\sigma(i-1)}=p \wedge \tau[p/\ivarx_{\sigma(i-1)}])\\
-    \vee &\bigvee_{p=0}^{\delta-1} \bigg(\ivarx_{\sigma(i-1)}=\ivarx_{\sigma(i-2)}+p\wedge \tau[\ivarx_{\sigma(i-2)}+p/\ivarx_{\sigma(i-1)}]\wedge \ivarx_{\sigma(i-1)}\le \alpha_1-1\bigg)\\
-    \vee &\bigvee_{p=0}^{\delta-1} \bigg(\ivarx_{\sigma(i-1)}=\ivarx_{\sigma(i-2)}+p\wedge \tau[\ivarx_{\sigma(i-2)}+p/\ivarx_{\sigma(i-1)}]\wedge \ivarx_{\sigma(i-1)}\ge \alpha_2+2\bigg)\\
+    \vee &\bigvee_{p=0}^{\delta-1} \bigg(\ivarx_{\sigma(i-1)}=\ivarx_{\sigma(i-2)}+p\wedge \tau[\ivarx_{\sigma(i-2)}+p/\ivarx_{\sigma(i-1)}]\bigg)\\
     \vee & \bigg( \ivarx_{\sigma(i-1)}\ge \gamma \wedge \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)}+\delta \wedge \ivarx_{\sigma(i-1)}\le \alpha(\vec\ivary)-1\bigg)\\
@@
 where all exponential occurrences of $\ivarx_{\sigma(i-1)}$ are eliminated.
+Note that in the last line of (\ref{opt-example}), $\ivarx_{\sigma(i-1)}$ is  enumerated from $\alpha_1$ to $\alpha_2+1$, while in (\ref{lemma-example}) only two term $\alpha(\vec \ivary)$ and $\alpha(\vec \ivary)+1$ are tested for $\ivarx_{\sigma(i)}$.
 

### 2022-04-01 Acta Informatica2022/7_impl_exp.tex
.
before modify sec 4

@@
 \subsection{Implementation}
-We implement the decision procedure in Wolfram Mathematica, called the {\paexp}-solver, which is able to solve the satisfiability of {\paexp} formulas.
+We implement the decision procedure in Wolfram Mathematica, called EXP-solver, which is able to solve the satisfiability of {\paexp} formulas.
 
-The {\paexp}-solver takes a quantifier-free {\paexp} formula as input.
+EXP-solver takes a quantifier-free {\paexp} formula as input.
 Moreover, it allows specifying a upper bound $10^u$ to uniformly bound the values of variables.
@@
 %
-Outputs of the {\paexp}-solver are either ``SAT'', ``UNSAT'', ``B-UNSAT'', or ``TIMEOUT'', standing for the given formula satisfiable, unsatisfiable, unsatisfiable up to $10^u$, or timeout if the solver does not terminate within the time limit.
+Outputs of EXP-solver are either ``SAT'', ``UNSAT'', ``B-UNSAT'', or ``TIMEOUT'', standing for the given formula satisfiable, unsatisfiable, unsatisfiable up to $10^u$, or timeout if the solver does not terminate within the time limit.
 If the output is ``SAT'', then a model (namely, an assignment) is returned.
@@
 
-To evaluate the performance of the {\paexp}-solver, we design two benchmark suites, ARITHMETIC and STRINGHASH\footnote{The benchmarks are available at https://github.com/EcstasyH/PAexp-Solver/tree/main/Benchmark}.
+To evaluate the performance of EXP-solver, we design two benchmark suites, ARITHMETIC and STRINGHASH\footnote{The benchmarks are available at https://github.com/EcstasyH/PAexp-Solver/tree/main/Benchmark}.
 %\begin{itemize}
@@
 We also create an SMTLib2 file for each problem instance, to facilitate the comparison with the state-of-the-art of SMT solvers CVC4 and Z3.
-Because neither CVC4 nor Z3 supports the exponential functions directly, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 1$ and $f(n+1) = 10*f(n)$.
+Because neither CVC4 nor Z3 supports the exponential functions directly, in the SMTLib2 files, we encode $10^x$ as a recursive function $f(x)$ defined by: $f(0) = 1$ and $f(n+1) = 10\cdot f(n)$.
 
@@
 We generate the SMTLib2 files for these string constraints, as inputs to the string constraint solvers.
-On the other hand, for the {\paexp}-solver, we do the following:
+On the other hand, for EXP-solver, we do the following:
 \begin{itemize}
-\item For flat instances, we generate {\paexp} formulas corresponding to the string constraints, as inputs to the {\paexp}-solver.
+\item For flat instances, we generate {\paexp} formulas corresponding to the string constraints, as inputs to EXP-solver.
 \item For non-flat instances, we use flat languages $a^* (b_{1} \ldots b_{k})$ to under-approximate 
@@
 
-We compare the {\paexp}-solver with the state-of-the-art SMT solvers on the generated benchmarks. Specifically, 
+We compare EXP-solver with the state-of-the-art SMT solvers on the generated benchmarks. Specifically, 
 \begin{itemize}
-\item for the ARITHMETIC benchmark suite, we compare the {\paexp}-solver against CVC4 (version 1.8) and Z3 (version 4.8.10),
-\item for the STRINGHASH benchmark suite, we compare the {\paexp}-solver against CVC4, Z3, and Trau\footnote{\url{https://github.com/guluchen/z3/tree/new_trau}}. 
+\item for the ARITHMETIC benchmark suite, we compare EXP-solver against CVC4 (version 1.8) and Z3 (version 4.8.10),
+\item for the STRINGHASH benchmark suite, we compare EXP-solver against CVC4, Z3, and Trau\footnote{\url{https://github.com/guluchen/z3/tree/new_trau}}. 
 \end{itemize}
@@
 
-For the ARITHMETIC benchmark suite, the $\paexp$-solver solves around $20\%$-$60\%$ more instances than Z3, and $30\%$-$100\%$ more instances than CVC4.
-Moreover, the gap becomes bigger as the the sizes of the formulas increase, which demonstrates that the $\paexp$-solver is more efficient in solving formulas of greater sizes.
-The average time of the $\paexp$-solver is comparable with Z3 and CVC4.
-The $\paexp$-solver reports ``B-UNSAT'' for 47 instances of the $(2, 3, 3, 4)$-group, while it does not report ``B-UNSAT'' (except one) for the other two groups.
-If more time is allowed, the $\paexp$-solver is able to report ``B-UNSAT'' for the ``TIMEOUT'' instances.
-
-%which means it can tackle some difficult instances.
-However, there are some instances are answered SAT or UNSAT by other solvers but timeout in $\paexp$-solver, e.g.
-12 instances in 3-4-4-5 group.
-The reason is that $\paexp$-solver must seach for all possible assignments before answering B-UNSAT, and the number of variables with exponential occurrences has a dramatic impact on the time complexity due to the non-elementary complexity upper bound.
-
-For the STRINGHASH benchmark suite, in overall, the $\paexp$-solver solves significantly more instances, especially those satisfiable instances, than Z3, CVC4, and Trau.
-For instance, for flat regular constraints, the $\paexp$-solver solves almost all 300 problem instances, except 3 of them\footnote{These three instances can actually be solved in 70 seconds.}, while  Z3, CVC4, Trau solve only 34, 89, 187 instances respectively.
-Trau gets wrong answers for some problem instances, e.g.
-it reports ``UNSAT'' for some satisfiable instances.
-From the results, we can see that the $\paexp$-solver achieves a good tradeoff between precision and efficiency, although it is slower than the other solvers. 
-%(More detailed experiment results of STRINGHASH can be found in Table~\ref{table:string}.)
-
-% for the pattern $12345\Sigma^*_{\sf num}6789$,  the $\paexp$-solver solves all the problem instances, while Z3, CVC4, and Trau solve only 60, 17, 3 of them.
-%
-% than all other three solvers on both two types of constraints.
-For flat regular constraints, Z3 and CVC fail on most of the instances while Trau may answer UNSAT incorrectly to some satisfiable instances.
-$\paexp$-solver work out all but 3 instances in 60 seconds and an extra experiment shows it will the remaining 3 instances can be solved in 70 seconds.
-For non-flat regular constraints, the results show a different bias for differenct pattern (with or without preffix and suffix) both in Z3 and CVC4, whereas $\paexp$-solver work out all problem instances in similar time.
-
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-% Merge bound and unbound 
-%\vspace*{-3mm}
 \setlength\tabcolsep{3pt}
 
-\begin{table*}[ht]
-    \caption{Experimental Results.
-The left table summarizes the results for both experiments, and more detailded results of STRINGHASH benchmark suite are presented in the right table.
-O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail,
+\begin{table*}[t]
+    \caption{Experimental Results, left: results for both experiments, right: more detailded results of STRINGHASH benchmark suite. O: Output, S:SAT, U: UNSAT, B: Bounded UNSAT, F: Fail, 
     $\#$: number of problems, $T$: average time in seconds}
@@
     \hline
-        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c}{Z3} & \multicolumn{2}{c}{CVC4} &  \multicolumn{2}{c}{Trau} & \multicolumn{2}{c }{$\paexp$} \\ 
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c}{Z3} & \multicolumn{2}{c}{CVC4} &  \multicolumn{2}{c}{Trau} & \multicolumn{2}{c }{EXP-solver} \\ 
         &  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$}  & \scriptsize{$\#$} & \scriptsize{$T$} & \scriptsize{$\#$} & \scriptsize{$T$}  \\ \hline
@@
     \hline
-        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c}{Z3} & \multicolumn{2}{c}{CVC4} &  \multicolumn{2}{c}{Trau} & \multicolumn{2}{c}{$\paexp$} \\
+        \multirow{2}{*}{Group }  & \multirow{2}{*}{O} & \multicolumn{2}{c}{Z3} & \multicolumn{2}{c}{CVC4} &  \multicolumn{2}{c}{Trau} & \multicolumn{2}{c}{EXP-solver} \\
        &  & $\#$ & $T$ &  $\#$ & $T$ & $\#$ & $T$ &  $\#$ & $T$ \\ 
@@
 \end{table*}
+
+For the ARITHMETIC benchmark suite, EXP-solver solves around $20\%$-$60\%$ more instances than Z3, and $30\%$-$100\%$ more instances than CVC4.
+Moreover, the gap becomes bigger as the the sizes of the formulas increase, which demonstrates that EXP-solver is more efficient in solving formulas of greater sizes.
+The average time of EXP-solver is comparable with Z3 and CVC4.
+EXP-solver reports ``B-UNSAT'' for 47 instances of the $(2, 3, 3, 4)$-group, while it does not report ``B-UNSAT'' (except one) for the other two groups.
+If more time is allowed, EXP-solver is able to report ``B-UNSAT'' for the ``TIMEOUT'' instances.
+
+%which means it can tackle some difficult instances.
+However, there are some instances are answered SAT or UNSAT by other solvers but timeout in $\paexp$-solver, e.g.
+12 instances in 3-4-4-5 group.
+The reason is that $\paexp$-solver must seach for all possible assignments before answering B-UNSAT, and the number of variables with exponential occurrences has a dramatic impact on the time complexity due to the non-elementary complexity upper bound.
+
+For the STRINGHASH benchmark suite, in overall, EXP-solver solves significantly more instances, especially those satisfiable instances, than Z3, CVC4, and Trau.
+For instance, for flat regular constraints, EXP-solver solves almost all 300 problem instances, except 3 of them\footnote{These three instances can actually be solved in 70 seconds.}, while  Z3, CVC4, Trau solve only 34, 89, 187 instances respectively.
+Trau gets wrong answers for some problem instances, e.g.
+it reports ``UNSAT'' for some satisfiable instances.
+From the results, we can see that EXP-solver achieves a good tradeoff between precision and efficiency, although it is slower than the other solvers. 
+%(More detailed experiment results of STRINGHASH can be found in Table~\ref{table:string}.)
+
+% for the pattern $12345\Sigma^*_{\sf num}6789$,  the $\paexp$-solver solves all the problem instances, while Z3, CVC4, and Trau solve only 60, 17, 3 of them.
+%
+% than all other three solvers on both two types of constraints.
+For flat regular constraints, Z3 and CVC fail on most of the instances while Trau may answer UNSAT incorrectly to some satisfiable instances.
+$\paexp$-solver work out all but 3 instances in 60 seconds and an extra experiment shows it will the remaining 3 instances can be solved in 70 seconds.
+For non-flat regular constraints, the results show a different bias for differenct pattern (with or without preffix and suffix) both in Z3 and CVC4, whereas $\paexp$-solver work out all problem instances in similar time.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+% Merge bound and unbound 
+%\vspace*{-3mm}
 
